### 2018-09-04 src/biological_data_prep.Rmd
Chapter 1 Biological data preperation. Begin tidying data

---
title: "biological data preperation"
author: "Samantha Andrews"
output: html_notebook
---

#Overview
Preperation of the capelin (*Mallotus villosus*) data obtained from [OBIS](www.iobis.org). This database shows the global occurance of capelin. 
Built with 'r getRversion()'.

#Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("packrat", "rworldmap")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

#Raw Data Clean
Load the raw data, then check use head() to check it looks ok, then names() to see all the column names
```{r load raw data}
rawdata <- read.csv("../data/bio/capelin_obis_raw.csv")
head(rawdata) 
names(rawdata)
```

following [Darwin Core](http://rs.tdwg.org/dwc/terms/) rename the latitude and longitude columns to decimalLatitude and decimalLongitude respectively
```{r dawin core naming convention}
tidydata <- rawdata
names(tidydata)[names(tidydata)=="latitude"] <- "decimalLatitude"
names(tidydata)[names(tidydata)=="longitude"] <- "decimalLongitude"
head(tidydata)
```

The dataset is global and needs to be trimmed to contain just the area of interest. 
Plot all the occurance points
```{r map all occurrence points}
library(rworldmap)
map <- getMap(resolution = "low") #creates an object called map at medium resoultion
plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) #the x and y lim are the long-lat bounds of the map
points(tidydata$decimalLongitude, tidydata$decimalLatitude) #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```
now extract just the Atlantic Canada data.
The lon-lat study area (which captures the whole Atlantic Canadian EEZ up to the north of Resolution Island in the Hudson Strait/ Davis Strait) is -71.0,-43.0,62.0,38.0 
```{r map only atlantic canadian occurrences}
tidydata <- tidydata[tidydata$decimalLongitude > -71, ]
tidydata <- tidydata[tidydata$decimalLongitude < -43, ]
tidydata <- tidydata[tidydata$decimalLatitude > 38, ]
plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) #re-map - this is just atl. canada (after checked subsetting is good)
points(tidydata$decimalLongitude, tidydata$decimalLatitude)
```
ok looks good...save what you have so far to a .csv
```{r}
write.csv(tidydata, file = "../data//bio/capelin_obis_tidy.csv", row.names = FALSE)
```

### 2018-09-04 src/biological_data_prep.Rmd
to fix missing inst_code

@@
 
-#Overview
+# Overview
 Preperation of the capelin (*Mallotus villosus*) data obtained from [OBIS](www.iobis.org). This database shows the global occurance of capelin. 
@@
 
-#Package dependencies
+# Package dependencies
 You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
@@
 
-#Raw Data Clean
+# Raw Data Clean
 Load the raw data, then check use head() to check it looks ok, then names() to see all the column names
@@
 ok looks good...save what you have so far to a .csv
-```{r}
+```{r save tidydata to csv}
 write.csv(tidydata, file = "../data//bio/capelin_obis_tidy.csv", row.names = FALSE)
 ```
+
+# Column removal
+So there are many columns that you don't need.. you want to keep
+id
+decimalLatitude
+decimalLongitude
+datecollected
+institutioncode
+individualcount
+resname
+originalscientificname
+
+```{r select colums needed}
+tidydata <- subset(tidydata, select = c(id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, resname, originalscientificname))
+head(tidydata)
+```
+
+# remove data without month/year info
+
+count NAs in the dataset. Mainly you are interested in the datecollected column (ignore individual count - you know there is a lot missing)
+```{r}
+missingdata <- apply(tidydata, 2, function (x) sum(is.na(x)))
+missingdata
+```
+hmm interesting - you have 771 columns from year month and day with missing data, but datacollected has 0. There may be a formatting issue.
+To take a closer look, subset all rows where year == na
+
+```{r}
+missingdatarows <- tidydata[is.na(tidydata$year), ]
+head(missingdatarows)
+```
+so the NAs in the year column relate to no data (without a na value attached). Remove all rows where year == na
+```{r}
+tidydata <- tidydata[!is.na(tidydata$year), ]
+missingdata2 <- apply(tidydata, 2, function (x) sum(is.na(x)))
+missingdata2
+```
+
+# Month year column creation
+make columns for the years and month based on the datecollected column
+```{r new columns}
+tidydata$year <- format(as.Date(tidydata$datecollected), "%Y") #created a new column called year
+tidydata$month <- format(as.Date(tidydata$datecollected), "%m") #created a new column called month
+tidydata$day <- format(as.Date(tidydata$datecollected), "%d") #created a new column called day 
+head(tidydata)
+```
+
+the year, month, day columns are characters Change to numeric
+```{r new columns 2}
+tidydata$year <- as.numeric(tidydata$year)
+tidydata$month <- as.numeric(tidydata$month)
+tidydata$day <- as.numeric(tidydata$day)
+head(tidydata)
+```
+
+# Institutions
+Lets take a look at the institutioncode to make sure the names make sense...
+```{r}
+tidydata$institutioncode <- as.character(tidydata$institutioncode)
+whosampled <- unique(tidydata$institutioncode) #makes a variable of institutions
+whosampledordered <- sort(whosampled) #just puts the list in 
+whosampledordered
+```
+Ok so a number of these names actually represent the same insitiution (e.g.
+[DFO Regions](http://www.dfo-mpo.gc.ca/regions/index-eng.htm).
+Some details [here](ftp://ftp.aoml.noaa.gov/od/pub/library/methods-branton.pdf) and the first is no institution (you might be able to figure it out though).
+The merges needed are as follows
+"Bedford Institute of Oceanography (BIO)" & "BIO" & "DFO-SF" == DFOMTMS 
+"DFO-IML" & "IML-MLI" & "IML_MLI" & "DFO-NG" == DFOQC 
+"DFO-NFLD" & "North Atlantic Fisheries Centre" == "DFONL" 
+"NMFS-NEFSC" & "NOAA, NMFS, Northeast Fisheries Science Center" == NEFSC
+"The Atlantic reference Centre" & "The Atlantic Reference Centre (ARC)" == ARC 
+"DFO-SG" & "DFO Gulf Fisheries Centre" == DFOGULF
+
+and some you can just make shorter
+"Northwest Atlantic Fisheries Organization (NAFO)" == NAFO
+"DFO-ISDM" == DFOISDM
+"Maine DMR" == MaineDMR
+```{r}
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "Bedford Institute of Oceanography (BIO)", "DFOMTMS")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "BIO", "DFOMTMS")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-SF", "DFOMTMS")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-IML", "DFOQC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "IML-MLI", "DFOQC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-NG", "DFOQC")  
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "IML_MLI", "DFOQC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-NFLD", "DFONL")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "North Atlantic Fisheries Centre", "DFONL")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "NMFS-NEFSC", "NEFSC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "NOAA, NMFS, Northeast Fisheries Science Center", "NEFSC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "The Atlantic Reference Centre (ARC)", "ARC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "The Atlantic reference Centre", "ARC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-SG", "DFOGulf")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO Gulf Fisheries Centre", "DFOGulf")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "Northwest Atlantic Fisheries Organization (NAFO)", "NAFO")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-Central and Arctic, Freshwater Institute (FWI)", "DFOCENARC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-ISDM", "DFOISDM")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "Maine DMR", "MaineDMR")
+
+#recheck
+tidydata$institutioncode <- as.character(tidydata$institutioncode)
+whosampled <- unique(tidydata$institutioncode) #makes a variable of institutions
+whosampledordered <- sort(whosampled) #just puts the list in 
+whosampledordered
+```
+
+Now look at the missing institution. Use resname to see if it helps...
+```{r}
+missinginst <- subset(tidydata, institutioncode == "")
+missinginst
+```
+ok now check for unique resname...
+```{r}
+missinginst$resname <- as.character(missinginst$resname)
+missinginstunique <- unique(tidydata$resname) #makes a variable of institutions
+missinginstuniquedordered <- sort(missinginstunique) #just puts the list in 
+missinginstuniquedordered <- as.character(missinginstuniquedordered)
+missinginstuniquedordered
+```
+Ugh ok there are quite a few here....
+
+Ok soo looking at the OBIS metadata [e.g.](http://www.iobis.org/explore/#/dataset/2327), this is what they should be
+
+Arctic Species Trend Index (ASTI) : Marine == CAFF
+Atlantic Reference Centre Museum of Canadian Atlantic Organisms - Invertebrates and Fishes Data == ARC
+Atlantic Zone Monitoring Program Maritimes Region (AZMP) plankton datasets. In: Fisheries and Oceans Canada - BioChem archive == DFOMTMS
+BioChem zooplankton samples from the Gully, 2006-2007 == DFOMTMS
+BioChem:  Zooplankton of St. Margarets Bay 1968 to 1971. == DFOMTMS
+BioChem: DFO Atlantic Zone Monitoring Program plankton datasets - Newfoundland and Labrador Region (OBIS Canada) ==  DFONL
+BioChem: Fish, Eggs and Larvae in the Gulf of St. Lawrence == DFOQC
+BioChem: Sameoto zooplankton collection == DFOMTMS 
+DFO Central and Arctic Multi-species Stock Assessment Surveys == DFOCENARC
+DFO Gulf Region Groundfish Research Vessel Surveys == DFOGulf
+DFO Maritimes Research Vessel Trawl Surveys Fish Observations == DFOMTMS
+DFO Newfoundland and Labrador Region Ecosystem Trawl Surveys == DFOQC
+DFO Quebec Region MLI museum == DFOQC
+DFO Quebec Region Multispecies bottom trawl surveys == DFOQC
+ECNASAP - East Coast North America Strategic Assessment == DFOMTMS
+Electronic Atlas of Ichthyoplankton on the Scotian Shelf of North America == DFOMTMS
+Fish specimens == ROM
+ICES contaminants and biological effects == ICES
+Ichthyology Collection - Royal Ontario Museum == ROM
+IMR Macroplankton surveys == IMR
+Maine Department of Marine Resources Inshore Trawl Survey 2000–2009 == MaineDMR
+NAFO/ICNAF - Environmental Surveys - NORWESTLANT 1-3, 1963: Fish eggs and larvae == NAFO
+NMNH Vertebrate Zoology Fishes Collections == NMNH
+Northeast Fisheries Science Center Bottom Trawl Survey Data == NEFSC
+Northern Gulf of St. Lawrence Fishes == DFOMTMS
+
+ok so now you want to fill in the missing institutioncode with the correct data...
+
+maybe just populate the missing values with the same text as resname
+...
+if institutioncode == ""
+institutioncode == "resname"
+```{r}
+for i in (1:nrow)
+```

### 2018-09-05 src/biological_data_prep.Rmd
Solve missinginstitution issue

@@
 missinginst$resname <- as.character(missinginst$resname)
-missinginstunique <- unique(tidydata$resname) #makes a variable of institutions
+missinginstunique <- unique(missinginst$resname) #makes a variable of institutions
 missinginstuniquedordered <- sort(missinginstunique) #just puts the list in 
@@
 ```
-Ugh ok there are quite a few here....
 
-Ok soo looking at the OBIS metadata [e.g.](http://www.iobis.org/explore/#/dataset/2327), this is what they should be
-
-Arctic Species Trend Index (ASTI) : Marine == CAFF
-Atlantic Reference Centre Museum of Canadian Atlantic Organisms - Invertebrates and Fishes Data == ARC
-Atlantic Zone Monitoring Program Maritimes Region (AZMP) plankton datasets. In: Fisheries and Oceans Canada - BioChem archive == DFOMTMS
-BioChem zooplankton samples from the Gully, 2006-2007 == DFOMTMS
-BioChem:  Zooplankton of St. Margarets Bay 1968 to 1971. == DFOMTMS
-BioChem: DFO Atlantic Zone Monitoring Program plankton datasets - Newfoundland and Labrador Region (OBIS Canada) ==  DFONL
-BioChem: Fish, Eggs and Larvae in the Gulf of St. Lawrence == DFOQC
-BioChem: Sameoto zooplankton collection == DFOMTMS 
-DFO Central and Arctic Multi-species Stock Assessment Surveys == DFOCENARC
-DFO Gulf Region Groundfish Research Vessel Surveys == DFOGulf
-DFO Maritimes Research Vessel Trawl Surveys Fish Observations == DFOMTMS
-DFO Newfoundland and Labrador Region Ecosystem Trawl Surveys == DFOQC
-DFO Quebec Region MLI museum == DFOQC
-DFO Quebec Region Multispecies bottom trawl surveys == DFOQC
-ECNASAP - East Coast North America Strategic Assessment == DFOMTMS
-Electronic Atlas of Ichthyoplankton on the Scotian Shelf of North America == DFOMTMS
-Fish specimens == ROM
-ICES contaminants and biological effects == ICES
-Ichthyology Collection - Royal Ontario Museum == ROM
-IMR Macroplankton surveys == IMR
-Maine Department of Marine Resources Inshore Trawl Survey 2000–2009 == MaineDMR
-NAFO/ICNAF - Environmental Surveys - NORWESTLANT 1-3, 1963: Fish eggs and larvae == NAFO
-NMNH Vertebrate Zoology Fishes Collections == NMNH
-Northeast Fisheries Science Center Bottom Trawl Survey Data == NEFSC
-Northern Gulf of St. Lawrence Fishes == DFOMTMS
-
-ok so now you want to fill in the missing institutioncode with the correct data...
-
-maybe just populate the missing values with the same text as resname
-...
-if institutioncode == ""
-institutioncode == "resname"
+Ok soo looking at the [OBIS metadata](http://www.iobis.org/explore/#/dataset/4356), the institution is the Conservation of Arctic Flora and Fauna (CAFF)
 ```{r}
-for i in (1:nrow)
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "", "CAFF")
+whosampled <- unique(tidydata$institutioncode) #makes a variable of institutions
+whosampledordered <- sort(whosampled) #just puts the list in 
+whosampledordered
 ```

### 2018-09-05 src/biological_data_prep.Rmd
trimmed by date, finished institutions, removed Icelandic data

--- 2018-09-04 src/biological_data_prep.Rmd
@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("packrat", "rworldmap")
+packages <- c("packrat", "rworldmap", "dplyr" "plyr")
 source("../src/ipak.R")
@@
 library(rworldmap)
-map <- getMap(resolution = "low") #creates an object called map at medium resoultion
+map <- getMap(resolution = "low") #creates an object called map at low resoultion
 plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) #the x and y lim are the long-lat bounds of the map
@@
 tidydata <- tidydata[tidydata$decimalLatitude > 38, ]
-plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) #re-map - this is just atl. canada (after checked subsetting is good)
+plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) 
 points(tidydata$decimalLongitude, tidydata$decimalLatitude)
@@
 ok looks good...save what you have so far to a .csv
+```{r save tidydata to csv}
+write.csv(tidydata, file = "../data/bio/capelin_obis_tidy.csv", row.names = FALSE)
+```
+
+# Column removal
+So there are many columns that you don't need.. you want to keep
+id
+decimalLatitude
+decimalLongitude
+datecollected
+institutioncode
+individualcount
+resname
+originalscientificname
+
+```{r select colums needed}
+tidydata <- subset(tidydata, select = c(id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, resname, originalscientificname))
+head(tidydata)
+```
+
+# remove data without month/year info
+
+count NAs in the dataset. Mainly you are interested in the datecollected column (ignore individual count - you know there is a lot missing)
 ```{r}
-write.csv(tidydata, file = "../data//bio/capelin_obis_tidy.csv", row.names = FALSE)
+missingdata <- apply(tidydata, 2, function (x) sum(is.na(x)))
+missingdata
 ```
+hmm interesting - you have 771 columns from year month and day with missing data, but datacollected has 0. There may be a formatting issue.
+To take a closer look, subset all rows where year == na
+
+```{r}
+missingdatarows <- tidydata[is.na(tidydata$year), ]
+head(missingdatarows)
+```
+so the NAs in the year column relate to no data (without a na value attached). Remove all rows where year == na
+```{r}
+tidydata <- tidydata[!is.na(tidydata$year), ]
+missingdata2 <- apply(tidydata, 2, function (x) sum(is.na(x)))
+missingdata2
+```
+
+# Month year column creation
+make columns for the years and month based on the datecollected column
+```{r new yymmdd columns}
+tidydata$year <- format(as.Date(tidydata$datecollected), "%Y") #created a new column called year
+tidydata$month <- format(as.Date(tidydata$datecollected), "%m") #created a new column called month
+tidydata$day <- format(as.Date(tidydata$datecollected), "%d") #created a new column called day 
+head(tidydata)
+```
+
+the year, month, day columns are characters Change to numeric
+```{r new yymmdd columns 2}
+tidydata$year <- as.numeric(tidydata$year)
+tidydata$month <- as.numeric(tidydata$month)
+tidydata$day <- as.numeric(tidydata$day)
+head(tidydata)
+```
+
+# Remove any duplicate entries
+The dplyr package's distinct function removes any duplicate rows (based on all columns EXCEPT id as this is always going to be unique, and year month day as this is generated from datecollected)
+
+```{r}
+library("dplyr")
+tidydata <- distinct(tidydata, decimalLatitude, decimalLongitude, datecollected, institutioncode, institutioncode, individualcount, resname, originalscientificname)
+```
+
+# Institutions
+Lets take a look at the institutioncode to make sure the names make sense...
+```{r}
+tidydata$institutioncode <- as.character(tidydata$institutioncode)
+whosampled <- unique(tidydata$institutioncode) #makes a variable of institutions
+whosampledordered <- sort(whosampled) #just puts the list in 
+whosampledordered
+```
+Ok so a number of these names actually represent the same insitiution (e.g.
+[DFO Regions](http://www.dfo-mpo.gc.ca/regions/index-eng.htm).
+Some details [here](ftp://ftp.aoml.noaa.gov/od/pub/library/methods-branton.pdf) and the first is no institution (you might be able to figure it out though).
+The merges needed are as follows
+"Bedford Institute of Oceanography (BIO)" & "BIO" & "DFO-SF" == DFOMTMS 
+"DFO-IML" & "IML-MLI" & "IML_MLI" & "DFO-NG" == DFOQC 
+"DFO-NFLD" & "North Atlantic Fisheries Centre" == "DFONL" 
+"NMFS-NEFSC" & "NOAA, NMFS, Northeast Fisheries Science Center" == NEFSC
+"The Atlantic reference Centre" & "The Atlantic Reference Centre (ARC)" == ARC 
+"DFO-SG" & "DFO Gulf Fisheries Centre" == DFOGULF
+
+and some you can just make shorter
+"Northwest Atlantic Fisheries Organization (NAFO)" == NAFO
+"DFO-ISDM" == DFOISDM
+"Maine DMR" == MaineDMR
+```{r consistent institutioncodes}
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "Bedford Institute of Oceanography (BIO)", "DFOMTMS")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "BIO", "DFOMTMS")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-SF", "DFOMTMS")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-IML", "DFOQC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "IML-MLI", "DFOQC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-NG", "DFOQC")  
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "IML_MLI", "DFOQC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-NFLD", "DFONL")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "North Atlantic Fisheries Centre", "DFONL")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "NMFS-NEFSC", "NEFSC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "NOAA, NMFS, Northeast Fisheries Science Center", "NEFSC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "The Atlantic Reference Centre (ARC)", "ARC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "The Atlantic reference Centre", "ARC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-SG", "DFOGulf")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO Gulf Fisheries Centre", "DFOGulf")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "Northwest Atlantic Fisheries Organization (NAFO)", "NAFO")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-Central and Arctic, Freshwater Institute (FWI)", "DFOCENARC")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "DFO-ISDM", "DFOISDM")
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "Maine DMR", "MaineDMR")
+
+#recheck
+tidydata$institutioncode <- as.character(tidydata$institutioncode)
+whosampled <- unique(tidydata$institutioncode) #makes a variable of institutions
+whosampledordered <- sort(whosampled) #just puts the list in 
+whosampledordered
+```
+
+Now look at the missing institution. Use resname to see if it helps...
+```{r check missing institutioncode}
+missinginst <- subset(tidydata, institutioncode == "")
+missinginst
+```
+ok now check for unique resname...
+```{r ID missing institutioncode}
+missinginst$resname <- as.character(missinginst$resname)
+missinginstunique <- unique(missinginst$resname) #makes a variable of institutions
+missinginstuniquedordered <- sort(missinginstunique) #just puts the list in 
+missinginstuniquedordered <- as.character(missinginstuniquedordered)
+missinginstuniquedordered
+```
+
+Ok soo looking at the [OBIS metadata](http://www.iobis.org/explore/#/dataset/4356), the institution is the Conservation of Arctic Flora and Fauna (CAFF)
+```{r fix missing institutioncode}
+tidydata$institutioncode <- replace(tidydata$institutioncode, tidydata$institutioncode == "", "CAFF")
+whosampled <- unique(tidydata$institutioncode) #makes a variable of institutions
+whosampledordered <- sort(whosampled) #just puts the list in 
+whosampledordered
+```
+
+Now lets check the number of records and spatial-temporal distribution of the observations by institution code to make sure none are dodgy
+
+first a table of how many observations each instituioncode has
+```{r institution code analysis - count}
+library(plyr) #to use the count function
+obs_by_ins <- count(tidydata, "institutioncode")
+obs_by_ins
+write.csv(obs_by_ins, file = "../output/bio/no_observations_institutioncode.csv")
+```
+ok so IMR, NAFO, ICES, & USM have very few entries (1, 1, 1, and 3 respectively)
+
+Lets take a look at the spatial breakdown of these institutions.First all points...
+
+```{r}
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "All Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(tidydata$decimalLongitude, tidydata$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/all_occurrences.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+```
+Note there is one point up by iceland that you should get rid of (Icelandic population thought to be seperate from Labrador, but it is unclear if this is true or not).
+
+Map the institutioncode == ARC only data...
+```{r map obs by institutuion}
+ARC_obs <- tidydata[tidydata$institutioncode == "ARC", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Arc Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(ARC_obs$decimalLongitude, ARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/ARC_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+CAFF_obs <- tidydata[tidydata$institutioncode == "CAFF", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "CAFF Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(CAFF_obs$decimalLongitude, CAFF_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/CAFF_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOCENARC_obs <- tidydata[tidydata$institutioncode == "DFOCENARC", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Central & Arctic Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOCENARC_obs$decimalLongitude, DFOCENARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/DFOCENARC_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOGulf_obs <- tidydata[tidydata$institutioncode == "DFOGulf", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Gulf Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOGulf_obs$decimalLongitude, DFOGulf_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/DFOGulf_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOISDM_obs <- tidydata[tidydata$institutioncode == "DFOISDM", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO ISDM Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOISDM_obs$decimalLongitude, DFOISDM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/DFOISDM_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOMTMS_obs <- tidydata[tidydata$institutioncode == "DFOMTMS", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Maritimes Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOMTMS_obs$decimalLongitude, DFOMTMS_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/DFOMTMS_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFONL_obs <- tidydata[tidydata$institutioncode == "DFONL", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Newfouundland & Labrador Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFONL_obs$decimalLongitude, DFONL_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/DFONL_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOQC_obs <- tidydata[tidydata$institutioncode == "DFOQC", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Quebec Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOQC_obs$decimalLongitude, DFOQC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/DFOQC_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+ICES_obs <- tidydata[tidydata$institutioncode == "ICES", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "ICES Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(ICES_obs$decimalLongitude, ICES_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/ICES_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+IMR_obs <- tidydata[tidydata$institutioncode == "IMR", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "IMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(IMR_obs$decimalLongitude, IMR_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/IMR_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+MaineDMR_obs <- tidydata[tidydata$institutioncode == "MaineDMR", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Maine DMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(MaineDMR_obs$decimalLongitude, MaineDMR_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/MaineDMR_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+NAFO_obs <- tidydata[tidydata$institutioncode == "NAFO", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "NAFO Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(NAFO_obs$decimalLongitude, NAFO_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/NAFO_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+NEFSC_obs <- tidydata[tidydata$institutioncode == "NEFSC", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "NEFSC Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(NEFSC_obs$decimalLongitude, NEFSC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/NEFSC_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+ROM_obs <- tidydata[tidydata$institutioncode == "ROM", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/ROM_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+USNM_obs <- tidydata[tidydata$institutioncode == "USNM", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "National Museum of Natural History Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(USNM_obs$decimalLongitude, USNM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/USNM_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+```
+
+Looking at the maps, the ROM, NAFO, IMR, ICES have data next to iceland.
+Since NAFO, IMR, and ICES only have one data point each, you can remove these observations easily from the record...
+```{r}
+tidydata <- tidydata[!tidydata$institutioncode == "NAFO", ]
+tidydata <- tidydata[!tidydata$institutioncode == "IMR", ]
+tidydata <- tidydata[!tidydata$institutioncode == "ICES", ]
+```
+Now for  ROM there are more data points...
+
+write what you have so far to a csv
+```{r}
+write.csv(tidydata, file = "../data/bio/capelin_obis_tidy.csv", row.names = FALSE)
+```

### 2018-09-06 src/biological_data_prep.Rmd
back up before R restart for package installation

@@
 Preperation of the capelin (*Mallotus villosus*) data obtained from [OBIS](www.iobis.org). This database shows the global occurance of capelin.
+
+A note to anyone who might happen to stumble across this...
+I am a beginner in R and have had no exposure to similar languages.
+The code herein is unlikely to be elegant and there is likely more efficient ways of running the code.
+
 Built with 'r getRversion()'.
@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("packrat", "rworldmap", "dplyr" "plyr")
+packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal")
 source("../src/ipak.R")
@@
 
+# Month year column creation
+make columns for the years and month based on the datecollected column
+```{r new yymmdd columns}
+tidydata$year <- format(as.Date(tidydata$datecollected), "%Y") #created a new column called year
+tidydata$month <- format(as.Date(tidydata$datecollected), "%m") #created a new column called month
+tidydata$day <- format(as.Date(tidydata$datecollected), "%d") #created a new column called day 
+head(tidydata)
+```
+
+the year, month, day columns are characters Change to numeric
+```{r new yymmdd columns 2}
+tidydata$year <- as.numeric(tidydata$year)
+tidydata$month <- as.numeric(tidydata$month)
+tidydata$day <- as.numeric(tidydata$day)
+head(tidydata)
+```
+
 # remove data without month/year info
@@
 
-# Month year column creation
-make columns for the years and month based on the datecollected column
-```{r new yymmdd columns}
-tidydata$year <- format(as.Date(tidydata$datecollected), "%Y") #created a new column called year
-tidydata$month <- format(as.Date(tidydata$datecollected), "%m") #created a new column called month
-tidydata$day <- format(as.Date(tidydata$datecollected), "%d") #created a new column called day 
-head(tidydata)
-```
-
-the year, month, day columns are characters Change to numeric
-```{r new yymmdd columns 2}
-tidydata$year <- as.numeric(tidydata$year)
-tidydata$month <- as.numeric(tidydata$month)
-tidydata$day <- as.numeric(tidydata$day)
-head(tidydata)
-```
-
 # Remove any duplicate entries
@@
 library("dplyr")
-tidydata <- distinct(tidydata, decimalLatitude, decimalLongitude, datecollected, institutioncode, institutioncode, individualcount, resname, originalscientificname)
+tidydata <- distinct(tidydata, decimalLatitude, decimalLongitude, datecollected, institutioncode, institutioncode, individualcount, resname, originalscientificname, .keep_all = T) #the .keep_all means that id isn't excluded from the output
+head(tidydata)
 ```
@@
 Since NAFO, IMR, and ICES only have one data point each, you can remove these observations easily from the record...
-```{r}
+```{r remove iceland data}
 tidydata <- tidydata[!tidydata$institutioncode == "NAFO", ]
@@
 ```
+
 Now for  ROM there are more data points...
+Use indentify function which lets you click on a point on the plot
+```{r locate coordinates of iceland point}
+x11() #this opens the plot i a window you can interact with
+plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude"))
+identify(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, labels=ROM_obs$decimalLongitude) #decimalLongitude chosen as plot identifier as the values are unique
+```
+The point that needs to be removed is located at -44.96667.
+Note this is not the full value so use some other locators to remove the row (you could just used id, but pick multiple others just to learn)
 
-write what you have so far to a csv
+```{r remove iceland point}
+tidydata <- tidydata[!(tidydata$resname == "Fish specimens" & tidydata$datecollected == "1965-08-13 12:00:00+01"), ]
+```
+
+now replot all data just to check...
+
+```{r map occurrences less iceland}
+plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "All Occurrences without Iceland", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(tidydata$decimalLongitude, tidydata$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/all_occurrences_less_iceland.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+```
+great! write what you have so far to a csv
 ```{r}
@@
 ```
+
+# Remove dates you don't have environmental data for
+The BIOMER dataset starts in 1998-01 (GLORYS 1993), so cut all occurrences prior to 1998-01
+
+```{r remove occurrences prior to 98/01}
+tidytest <- tidydata
+tidytest <- tidytest[tidytest$year >= 1998, ]
+```
+
+and write to a csv...
+```{r}
+write.csv(tidydata, file = "../data/bio/capelin_obis_tidy.csv", row.names = FALSE)
+```
+
+# Adding new (non-environmental) data to the file
+There are some more columns you want to add...
+- Occurrence - presence/background (1/0)
+- NAFO region
+- Gear type
+- Unique Cell ID (based on the environmental data)
+
+so not to mess with tidydata, create a new variable and write a new .csv
+```{r create new variable/csv for adding data}
+adddata <- tidydata
+write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE)
+```
+
+first create a new column for Occurence, and populate with "1" (because these are all presences)
+```{r}
+adddata$occurrence <- "1"
+head(adddata)
+```
+
+# Add NAFO Region occurrence is from
+NAFO Zones shapefile obtained from [NAFO](https://www.nafo.int/Data/GIS)
+
+```{r}
+library(rgdal)
+addtest <- adddata
+
+```

### 2018-09-06 src/biological_data_prep.Rmd
added NAFO, gear, presence. next onto creating a env. reference grid

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal")
+packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp",)
 source("../src/ipak.R")
@@
 originalscientificname
+collectioncode (this contains some info that might help with finding out gear used in the survey)
 
 ```{r select colums needed}
-tidydata <- subset(tidydata, select = c(id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, resname, originalscientificname))
+tidydata <- subset(tidydata, select = c(id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, resname, originalscientificname, collectioncode))
 head(tidydata)
@@
 ```{r remove occurrences prior to 98/01}
-tidytest <- tidydata
-tidytest <- tidytest[tidytest$year >= 1998, ]
+tidydata <- tidydata[tidydata$year >= 1998, ]
 ```
@@
 
+```{r NAFO shapefile extraction}
+library(rgdal)
+library(sp)
+coordinates(adddata) <- c("decimalLongitude", "decimalLatitude") 
+nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
+proj4string(adddata) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
+adddata$nafo_zone <- over(adddata, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
+head(adddata) #just to check it worked
+```
+
+Great. Lets see if any are missing a NAFO Zone. But first, write to a csv and reload as the data is now some wierd 'spatial points dataframe'...
 ```{r}
-library(rgdal)
-addtest <- adddata
+write.csv(adddata, "../data/bio/capelin_additional_tidy.csv", row.names=FALSE)
+adddata <- read.csv("../data/bio/capelin_additional_tidy.csv")
+nafo_na <- subset(adddata, is.na(adddata$nafo_zone))
+nafo_na
+```
 
+Uh oh there is a lot. A bet a bunch of them are in the Hudson Strait (which is outside the NAFO Zone).Map the nafo_na points...
+```{r}
+plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
 ```
+ok yes - Hudson Strait has many, but there are some near the coast that are probably just missed out due to minor differences in the point location/shapefile boundaries.
+Start with the Hudson Strait ones as this is easy - anything above 56N, nafo_zone == "HudsonStrait"
+
+```{r}
+adddata$nafo_zone <- as.character(adddata$nafo_zone)
+adddata[is.na(adddata)] <- "xx"
+adddata$nafo_zone[adddata$nafo_zone == "xx" & adddata$decimalLatitude > 56] <- "HudsonStrait"
+```
+and map to check
+```{r}
+nafo_na <- subset(adddata, nafo_zone == "xx")
+plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+```
+so now i need to account for the others (either as occuring on land or just in an area that is missing a NAFO Zone tag) which is more tricky... i might just do this bit in ArcGIS quickly.
+```{r}
+write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE)
+```
+So in ArcGIS I saw that these points fall on land. Some are extremely close to the coast, some further inland. There are only 132 of thesepoints, so I will just delete them from the database.
+```{r}
+adddata$nafo_zone[adddata$nafo_zone == "xx"] <- NA #make all xx values NA
+adddata <- adddata[!is.na(adddata$nafo_zone), ]
+write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE)
+```
+
+#Gear type
+To get this information, I need to look at the collectioncode details and the OBIS metadata for each resname and then add.
+```{r}
+adddata$gear <- "x" #create a new column called gear and populate with x
+#now give a list of the resnames...
+whichres <- unique(adddata$resname) #makes a variable of institutions
+whichresordered <- sort(whichres) #just puts the list in 
+whichresordered
+```
+[Arctic Species Trend Index (ASTI) : Marine](http://www.iobis.org/explore/#/dataset/4356). No gear details available
+```{r}
+asti <- subset(adddata, resname == "Arctic Species Trend Index (ASTI) : Marine")
+asti$collectioncode <- as.character(asti$collectioncode)
+asti_col <- unique(asti$collectioncode) #makes a variable of institutions
+asti_colordered <- sort(asti_col) #just puts the list in 
+asti_colordered
+```
+there is no information in the collection code. The Metadata on OBIS also gives no details. gear = NA
+```{r}
+adddata$gear[adddata$resname == "Arctic Species Trend Index (ASTI) : Marine"] <- "NA"
+```
+
+[Atlantic Reference Centre Museum of Canadian Atlantic Organisms - Invertebrates and Fishes Data](http://www.iobis.org/explore/#/dataset/29).
+```{r}
+arcm <- subset(adddata, resname == "Atlantic Reference Centre Museum of Canadian Atlantic Organisms - Invertebrates and Fishes Data")
+arcm$collectioncode <- as.character(arcm$collectioncode)
+arcm_col <- unique(arcm$collectioncode) #makes a variable of institutions
+arcm_colordered <- sort(arcm_col) #just puts the list in 
+arcm_colordered
+```
+No gear details in collectioncode. The Metadata on OBIS also gives no details. gear = NA 
+```{r}
+adddata$gear[adddata$resname == "Atlantic Reference Centre Museum of Canadian Atlantic Organisms - Invertebrates and Fishes Data"] <- "NA"
+```
+
+[Atlantic Zone Monitoring Program Maritimes Region (AZMP) plankton datasets. In: Fisheries and Oceans Canada - BioChem archive](http://iobis.org/explore/#/dataset/2328). 
+```{r}
+azmpmar <- subset(adddata, resname == "Atlantic Zone Monitoring Program Maritimes Region (AZMP) plankton datasets. In: Fisheries and Oceans Canada - BioChem archive")
+azmpmar$collectioncode <- as.character(azmpmar$collectioncode)
+azmpmar_col <- unique(azmpmar$collectioncode) #makes a variable of institutions
+azmpmar_colordered <- sort(azmpmar_col) #just puts the list in 
+azmpmar_colordered
+```
+No gear details in collectioncode. The metadata on OBIS indicates vertical plankton tow 
+```{r}
+adddata$gear[adddata$resname == "Atlantic Zone Monitoring Program Maritimes Region (AZMP) plankton datasets. In: Fisheries and Oceans Canada - BioChem archive"] <- "vertical_plankton_tow"
+```
+
+[BioChem zooplankton samples from the Gully, 2006-2007](http://iobis.org/explore/#/dataset/2327). 
+```{r}
+biogul <- subset(adddata, resname == "BioChem zooplankton samples from the Gully, 2006-2007")
+biogul$collectioncode <- as.character(biogul$collectioncode)
+biogul_col <- unique(biogul$collectioncode) #makes a variable of institutions
+biogul_colordered <- sort(biogul_col) #just puts the list in 
+biogul_colordered
+```
+Along with the OBIS metadata, indicates gear = vertical plankton tow
+```{r}
+adddata$gear[adddata$resname == "BioChem zooplankton samples from the Gully, 2006-2007"] <- "vertical_plankton_tow"
+```
+
+[BioChem: DFO Atlantic Zone Monitoring Program plankton datasets - Newfoundland and Labrador Region (OBIS Canada)](http://iobis.org/explore/#/dataset/2666). 
+```{r}
+azmpnl <- subset(adddata, resname == "BioChem: DFO Atlantic Zone Monitoring Program plankton datasets - Newfoundland and Labrador Region (OBIS Canada")
+azmpnl$collectioncode <- as.character(azmpnl$collectioncode)
+azmpnl_col <- unique(azmpnl$collectioncode) #makes a variable of institutions
+azmpnl_colordered <- sort(azmpnl_col) #just puts the list in 
+azmpnl_colordered
+```
+No details in collectioncode. Obis metadata indicates gear = vertical plankton tow
+```{r}
+adddata$gear[adddata$resname == "BioChem: DFO Atlantic Zone Monitoring Program plankton datasets - Newfoundland and Labrador Region (OBIS Canada)"] <- "vertical_plankton_tow"
+```
+
+[BioChem: Sameoto zooplankton collection](http://iobis.org/explore/#/dataset/2670).Vertical plankton tow
+```{r}
+biosam <- subset(adddata, resname == "BioChem: Sameoto zooplankton collection")
+biosam$collectioncode <- as.character(biosam$collectioncode)
+biosam_col <- unique(biosam$collectioncode) #makes a variable of institutions
+biosam_colordered <- sort(biosam_col) #just puts the list in 
+biosam_colordered
+```
+Insufficient details in collectioncode. Obis metadata indicates gear = vertical plankton tow
+```{r}
+adddata$gear[adddata$resname == "BioChem: Sameoto zooplankton collection"] <- "vertical_plankton_tow"
+```
+
+[DFO Central and Arctic Multi-species Stock Assessment Surveys](http://iobis.org/explore/#/dataset/2293). A
+```{r}
+dfocas <- subset(adddata, resname == "DFO Central and Arctic Multi-species Stock Assessment Surveys")
+dfocas$collectioncode <- as.character(dfocas$collectioncode)
+dfocas_col <- unique(dfocas$collectioncode) #makes a variable of institutions
+dfocas_colordered <- sort(dfocas_col) #just puts the list in 
+dfocas_colordered
+```
+ok these are all types of bottom trawl.
+```{r}
+adddata$gear[adddata$resname == "DFO Central and Arctic Multi-species Stock Assessment Surveys" & adddata$collectioncode == "DFOSurvey_Modified Alfredo-03"] <- "bottom_trawl_alfredo_03"
+adddata$gear[adddata$resname == "DFO Central and Arctic Multi-species Stock Assessment Surveys" & adddata$collectioncode == "DFOSurvey_Modified Cosmos 2600"] <- "bottom_trawl_cosmos_2600"
+adddata$gear[adddata$resname == "DFO Central and Arctic Multi-species Stock Assessment Surveys" & adddata$collectioncode == "DFOSurvey_Modified Modified (21\") Campelen"] <- "bottom_trawl_campelen_21"
+adddata$gear[adddata$resname == "DFO Central and Arctic Multi-species Stock Assessment Surveys" & adddata$collectioncode == "DFOSurvey_Modified Standard (14\") Campelen"] <- "bottom_trawl_campelen_14"
+```
+
+[DFO Gulf Region Groundfish Research Vessel Surveys](bottom trawl).  
+```{r}
+dfogulfas <- subset(adddata, resname == "DFO Gulf Region Groundfish Research Vessel Surveys")
+dfogulfas$collectioncode <- as.character(dfogulfas$collectioncode)
+dfogulfas_col <- unique(dfogulfas$collectioncode) #makes a variable of institutions
+dfogulfas_colordered <- sort(dfogulfas_col) #just puts the list in 
+dfogulfas_colordered
+```
+no gear details in collection code.
+OBIS metadata indicates Three survey vessels and two types of bottom-trawls were used prior to 2003: the E.
+E.
+Prince from 1971-1985 using a Yankee 36 trawl, the Lady Hammond from 1985-1991 and the Alfred Needler from 1992-2002, both using a Western IIA trawl.
+Due to a fire on the Needler, the 2003 survey was conducted by its sister-ship, the Templeman using the Western IIA trawl.
+This survey started late and was incomplete.
+The survey was conducted by both the Needler and the Teleost in 2004 and 2005 and by the Teleost since then, using the Western IIA trawl.
+Note you have not captured ship information above so leave it out here. Your data is 1998 onwards to the gear for all = bottom_trawl_western_IIA
+```{r}
+adddata$gear[adddata$resname == "DFO Gulf Region Groundfish Research Vessel Surveys"] <- "bottom_trawl_western_IIA"
+```
+
+[DFO Maritimes Research Vessel Trawl Surveys Fish Observations](http://iobis.org/explore/#/dataset/237).
+```{r}
+dfomas <- subset(adddata, resname == "DFO Maritimes Research Vessel Trawl Surveys Fish Observations")
+dfomas$collectioncode <- as.character(dfomas$collectioncode)
+dfomas_col <- unique(dfomas$collectioncode) #makes a variable of institutions
+dfomas_colordered <- sort(dfomas_col) #just puts the list in 
+dfomas_colordered
+```
+insuffient information. OBIS metadata just indicates "a variety of trawl gear". gear = "bottom_trawl"
+```{r}
+adddata$gear[adddata$resname == "DFO Maritimes Research Vessel Trawl Surveys Fish Observations"] <- "bottom_trawl"
+```
+
+[DFO Newfoundland and Labrador Region Ecosystem Trawl Surveys](http://iobis.org/explore/#/dataset/2814)
+```{r}
+dfonas <- subset(adddata, resname == "DFO Newfoundland and Labrador Region Ecosystem Trawl Surveys")
+dfonas$collectioncode <- as.character(dfonas$collectioncode)
+dfonas_col <- unique(dfonas$collectioncode) #makes a variable of institutions
+dfonas_colordered <- sort(dfonas_col) #just puts the list in 
+dfonas_colordered
+```
+More useful, OBIS metadata indicates "Starting in the fall of 1995 surveys have been conducted using a Campelen 1800 trawl fitted with rock-hopper foot gear." gear = bottom_trawl_campelen_1800
+```{r}
+adddata$gear[adddata$resname == "DFO Newfoundland and Labrador Region Ecosystem Trawl Surveys"] <- "bottom_trawl_campelen_1800"
+```
+
+[DFO Quebec Region Multispecies bottom trawl surveys](http://iobis.org/explore/#/dataset/2542)
+```{r}
+dfoqas <- subset(adddata, resname == "DFO Quebec Region Multispecies bottom trawl surveys")
+dfoqas$collectioncode <- as.character(dfoqas$collectioncode)
+dfoqas_col <- unique(dfoqas$collectioncode) #makes a variable of institutions
+dfoqas_colordered <- sort(dfoqas_col) #just puts the list in 
+dfoqas_colordered
+```
+no gear details in collectioncode. OBIS metadata indicates Campelen 1800. gear = bottom_trawl_campelen_1800
+```{r}
+adddata$gear[adddata$resname == "DFO Quebec Region Multispecies bottom trawl surveys"] <- "bottom_trawl_campelen_1800"
+```
+
+[Fish specimens](http://iobis.org/explore/#/dataset/2417)
+```{r}
+fishsp <- subset(adddata, resname == "Fish specimens")
+fishsp$collectioncode <- as.character(fishsp$collectioncode)
+fishsp_col <- unique(fishsp$collectioncode) #makes a variable of institutions
+fishsp_colordered <- sort(fishsp_col) #just puts the list in 
+fishsp_colordered
+```
+no gear detail in collectioncode. OBIS metadata indicates these are museum records. gear = NA
+```{r}
+adddata$gear[adddata$resname == "Fish specimens"] <- "NA"
+```
+
+[Ichthyology Collection - Royal Ontario Museum](http://iobis.org/explore/#/dataset/3511)
+```{r}
+iccol <- subset(adddata, resname == "Ichthyology Collection - Royal Ontario Museum")
+iccol$collectioncode <- as.character(iccol$collectioncode)
+iccol_col <- unique(iccol$collectioncode) #makes a variable of institutions
+iccol_colordered <- sort(iccol_col) #just puts the list in 
+iccol_colordered
+```
+no gear details in collectioncode. OBIS metadata indicates museum records. gear = NA
+```{r}
+adddata$gear[adddata$resname == "Ichthyology Collection - Royal Ontario Museum"] <- "NA"
+```
+
+[Maine Department of Marine Resources Inshore Trawl Survey 2000–2009](http://iobis.org/explore/#/dataset/1693)
+```{r}
+mdmr <- subset(adddata, resname == "Maine Department of Marine Resources Inshore Trawl Survey 2000–2009")
+mdmr$collectioncode <- as.character(mdmr$collectioncode)
+mdmr_col <- unique(mdmr$collectioncode) #makes a variable of institutions
+mdmr_colordered <- sort(mdmr_col) #just puts the list in 
+mdmr_colordered
+```
+Insufficient infomation in collectioncode.
+OBIS metadata indicates "a modified version of the shrimp net used in Maine waters.
+It was designed by the vessel owner and net designer Jeff Flagg to fish for a variety of near-bottom dwelling species without targeting any specific component" gear-type = "bottom_trawl"
+```{r}
+adddata$gear[adddata$resname == "Maine Department of Marine Resources Inshore Trawl Survey 2000–2009"] <- "bottom_trawl"
+```
+
+[Northeast Fisheries Science Center Bottom Trawl Survey Data](http://iobis.org/explore/#/dataset/1435)
+```{r}
+nefsu <- subset(adddata, resname == "Northeast Fisheries Science Center Bottom Trawl Survey Data")
+nefsu$collectioncode <- as.character(nefsu$collectioncode)
+nefsu_col <- unique(nefsu$collectioncode) #makes a variable of institutions
+nefsu_colordered <- sort(nefsu_col) #just puts the list in 
+nefsu_colordered
+```
+insufficient detail in collection code. OBIS metadata indicates "bottom trawl gear". gear = "bottom_trawl"
+```{r}
+adddata$gear[adddata$resname == "Northeast Fisheries Science Center Bottom Trawl Survey Data"] <- "bottom_trawl"
+```
+
+[Northern Gulf of St. Lawrence Fishes](http://iobis.org/explore/#/dataset/2671)
+```{r}
+ngsl <- subset(adddata, resname == "Northern Gulf of St. Lawrence Fishes")
+ngsl$collectioncode <- as.character(ngsl$collectioncode)
+ngsl_col <- unique(ngsl$collectioncode) #makes a variable of institutions
+ngsl_colordered <- sort(ngsl_col) #just puts the list in 
+ngsl_colordered
+```
+no gear details in collectioncode. OBIS metadata indicates "Sampling methods used ranged from fish traps and seine nets on the coasts to offshore trawling". gear = NA
+```{r}
+adddata$gear[adddata$resname == "Northern Gulf of St. Lawrence Fishes"] <- "NA"
+```
+
+ok check to see if all of gear is done (ie if there are any x left)
+```{r}
+gearcheck <- subset(adddata, gear == "x")
+gearcheck
+```
+yay - all done.
+
+# createa unique cell ID

### 2018-09-10 src/environmental_data_preperation.Rmd
starting to deal with env. netcdfs

---
title: "environmental_data_preperation"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Preperation of GLORYS and BIOMER netCDFs obtained from [Copernicus](www.marine.copernicus.eu/services-portfolio/access-to-products/). These databases give global oceanographic conditions.
The BIOMER datasets are in netCDF version 3, and the GLORYS in netCDF version 1

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. The code herein is unlikely to be elegant and there is likely more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("RNetCDF", "raster")
source("./src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

# read in a netcdf
```{r}
library("RNetCDF") #this will read the netcdf version 1 formats 
ncpath <- "../data/env/"
ncname <- "CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim"  
ncfname <- paste(ncpath, ncname, ".nc", sep="")
dname <- "salinity"  
ncin <- nc_open(ncfname)
ncin
```
# convert netCDF into a raster brick

```{r}
library(raster)
rp <- projectRaster(ncin,crs="+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m no_defs")
```

```{r}
ncbrick <- brick(ncin)
```

### 2018-09-10 src/environmental_data_preperation.Rmd
wel...not really much progress..

@@
 library("RNetCDF") #this will read the netcdf version 1 formats 
-ncpath <- "../data/env/"
-ncname <- "CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim"  
-ncfname <- paste(ncpath, ncname, ".nc", sep="")
-dname <- "salinity"  
-ncin <- nc_open(ncfname)
-ncin
-```
-# convert netCDF into a raster brick
-
-```{r}
-library(raster)
-rp <- projectRaster(ncin,crs="+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m no_defs")
+library(ncdf4)
+ncbr <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
+ncbr
 ```
 
+reproject...
 ```{r}
-ncbrick <- brick(ncin)
+canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
+library(rgdal)
+rp <- projectRaster(ncbr, crs = canada_eq_ar_proj)
+rp
 ```

### 2018-09-10 src/environmental_data_preperation.Rmd
wel...not really much progress..

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("RNetCDF", "raster")
+packages <- c("RNetCDF", "raster", "rgdal")
 source("./src/ipak.R")
@@
 library("RNetCDF") #this will read the netcdf version 1 formats 
-library(ncdf4)
+library("ncdf4")
 ncbr <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
@@
 canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
-library(rgdal)
+library("rgdal")
 rp <- projectRaster(ncbr, crs = canada_eq_ar_proj)

### 2018-09-11 src/biological_data_prep.Rmd
addition of "see who sampled in which year"
addition of month by instituion

@@
 
+# see who sampled in which year and month
+
+year
+```{r}
+inst_by_yr <- with(adddata, table(year, institutioncode))
+write.csv(inst_by_yr, file = "../output/bio/institutioncode_by_yr.csv")
+```
+
+month
+```{r}
+inst_by_mt <- with(adddata, table(month, institutioncode))
+write.csv(inst_by_mt, file = "../output/bio/institutioncode_by_mth.csv")
+```
+
 # createa unique cell ID

### 2018-09-11 src/environmental_data_preperation.Rmd
Michael Sumner is a legend!

@@
 packages <- c("RNetCDF", "raster", "rgdal")
-source("./src/ipak.R")
+source("../src/ipak.R")
 ipak(packages)
@@
 ```
+
+# michael code
+```{r}
+## use ncdf4 directly
+library(ncdf4)
+## connection to the file
+nc <- nc_open("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
+## desired variable 
+v <- ncvar_get(nc, "salinity")
+## dimension coordinates
+lon <- ncvar_get(nc, "longitude")
+lat <- ncvar_get(nc, "latitude")
+## a template, one map from the lon, lat, and v[,,1]
+template <- raster(list(x = lon, y = lat, z = v[,,1]), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ")
+
+## apply template to brick, and set values to v (but flipped in Y because that's how raster works - lat is increasing in the NetCDF, so we go upside down)
+diy <- setValues(brick(template), v[, ncol(v):1, ])
+
+## target
+prj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
+target <- projectExtent(diy, prj)  ## get a target extent, and we'll set the exact resolution we want
+res(target) <- c(25000, 25000)
+aea <- projectRaster(diy, target)
+writeRaster(aea,'../data/env/aea.tif',options=c('TFW=YES'))
+plot(aea[[36]]) #the number in brackets refers to the depth number
+
+m <- spTransform(rnaturalearth::ne_countries(), prj)
+```

### 2018-09-17 src/environmental_data_preperation.Rmd
?

--- 2018-09-10 src/environmental_data_preperation.Rmd
@@
 ```
+um...

### 2018-09-17 src/environmental_data_preperation.Rmd
create unique cell id layer (based on a GLORYS Layer)

--- 2018-09-10 src/environmental_data_preperation.Rmd
@@
 Note this function checks to see if the packages are installed first.
-Th
 ```{r pre-install packages, message=FALSE}
-packages <- c("RNetCDF", "raster", "rgdal")
-source("./src/ipak.R")
+packages <- c("raster", "rgdal") 
+source("../src/ipak.R")
 ipak(packages)
-suppressMessages(ipak)
 ```
@@
 ```{r}
-library("RNetCDF") #this will read the netcdf version 1 formats 
 library("ncdf4")
@@
 
-reproject...
+# reproject 
+There is currently an issue with the projectRaster command in the raster package. Michael Sumners kindly offered a [temporary solution](https://twitter.com/mdsumner/status/1040500830539935744)
 ```{r}
-canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
+## use ncdf4 directly
+library(ncdf4)
+## connection to the file
+nc <- nc_open("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
+## desired variable 
+v <- ncvar_get(nc, "salinity")
+## dimension coordinates
+lon <- ncvar_get(nc, "longitude")
+lat <- ncvar_get(nc, "latitude")
+## a template, one map from the lon, lat, and v[,,1]
+template <- raster(list(x = lon, y = lat, z = v[,,1]), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ")
+
+## apply template to brick, and set values to v (but flipped in Y because that's how raster works - lat is increasing in the NetCDF, so we go upside down)
+diy <- setValues(brick(template), v[, ncol(v):1, ])
+
+## target
+prj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs" #note -60 used to be -91. Changed to 'straighten up'.
+target <- projectExtent(diy, prj)  ## get a target extent, and we'll set the exact resolution we want
+res(target) <- c(25000, 25000)
+aea <- projectRaster(diy, target)
+plot(aea[[1]]) #the number in brackets refers to the depth number
+aea
+writeRaster(aea[[1]],'../data/env/aea.tif',overwrite=TRUE,options=c('TFW=YES'))
+```
+
+#if projectRaster woudl work properly...
+
+library(raster)
+canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=- +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
 library("rgdal")
@@
 rp
+writeRaster(rp,'../data/env/rp.tif',options=c('TFW=YES')) #this last line jsut saves as a tiff...don't really nead this
+
+----------
+
+#create unique id layer
+ok now you want to create a duplicate of the top layer and then give each cell an unique ID
+
+First 
+```{r}
+aea[[1]]
+unique_cell <- (aea[[1]])
 ```
+
+Because the NA values are currently attributed with some other number and not NA in R (NA is set to -32767 in this particular layer), you need to make sure to exclude the cells which represent NA...
+so anything under 0.
+the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (11,425)
+```{r}
+cell_val_freq <- freq(unique_cell, value=NULL)
+cell_val_freq
+```
+ok so we have 4693 cells that have NA (or in reality a negative number that represents NA). so remove 4693 from ncell. We know that there are 11424 cells, soo... 11424-4693 = 6732 cells have data.
+
+now give a unique ID to each cell
+```{r}
+unique_cell[unique_cell > 0] <- 1:6731
+writeRaster(unique_cell,'../output/env/unique_cell1.tif',overwrite=TRUE, options=c('TFW=YES')) #this is just to test it in ArcGIS
+```

### 2018-09-17 src/biological_data_prep.Rmd
extracting value to point....

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp",)
+packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster")
 source("../src/ipak.R")
@@
 
-# createa unique cell ID
+# create a unique cell ID
+
+A unique cell ID layer has been created in the environmental_data_preperation.Rmd. Now you want to extract the value of that layer to the .csv you are creating...
+
+Load the env. data
+```{r}
+library(raster)
+cell_layer <- raster("../output/env/unique_cell.tif") #this is loading the cell layer that was created and saved as a tif
+plot(cell_layer)
+```
+The occurrence data needs to be converted into s spatialpoints dataframe
+```{r}
+library(sp)
+xy <- adddata[ ,c(2,3)]
+spatial_data <- SpatialPointsDataFrame(coords = xy, data = adddata, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
+spatial_data
+```
+ok now you should probably reproject the data into the same projection as the env. data
+```{r}
+spatial_data_aea <- spTransform(spatial_data, CRS = "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs") #note -60 used to be -91.
+Changed to 'straighten up'
+```
+Ok now plot the env. and point data...
+```{r}
+plot(cell_layer)
+plot(spatial_data_aea$decimalLatitude, spatial_data_aea$decimalLongitude)
+plot(spatial_data$decimalLongitude,spatial_data$decimalLatitude)
+```
+oh it doesn;t work... it should be in meters now?

### 2018-09-17 src/environmental_data_preperation.Rmd
extracting value to point....

@@
 unique_cell[unique_cell > 0] <- 1:6731
-writeRaster(unique_cell,'../output/env/unique_cell1.tif',overwrite=TRUE, options=c('TFW=YES')) #this is just to test it in ArcGIS
+writeRaster(unique_cell,'../output/env/unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
 ```

### 2018-09-18 src/biological_data_prep.Rmd
fixed the issue with projecting the species data to aea

@@
 library(sp)
-xy <- adddata[ ,c(2,3)]
+xy <- adddata[ ,c(3,2)] #3 = decimalLongitude, 2 = decimalLatitude
 spatial_data <- SpatialPointsDataFrame(coords = xy, data = adddata, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
@@
 Changed to 'straighten up'
+spatial_data_aea
 ```
+```{r}
+plot(spatial_data_aea, axes=TRUE, cex.axis=.95)
+write.csv(spatial_data_aea, file = "../output/bio/aea_data.csv", row.names = FALSE)
+
+```
+
 Ok now plot the env. and point data...
@@
 plot(cell_layer)
-plot(spatial_data_aea$decimalLatitude, spatial_data_aea$decimalLongitude)
-plot(spatial_data$decimalLongitude,spatial_data$decimalLatitude)
+points(spatial_data_aea$decimalLongitude, spatial_data_aea$decimalLatitude)
 ```
-oh it doesn;t work... it should be in meters now?
+Write the transformed species data to a new csv (this will also have the coordinates in meters - note meters are under decimalLongitude1 and decimalLatitude1)
+```{r}
+write.csv(spatial_data_aea, file = "../output/bio/adddata_aea.csv", row.names = FALSE)
+```

### 2018-09-18 src/biological_data_prep.Rmd
unique cell id extracted

--- 2018-09-11 src/biological_data_prep.Rmd
@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp",)
+packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble")
 source("../src/ipak.R")
@@
 
-# createa unique cell ID
+# create a unique cell ID
+
+A unique cell ID layer has been created in the environmental_data_preperation.Rmd. Now you want to extract the value of that layer to the .csv you are creating...
+
+Load the env. data
+```{r}
+library(raster)
+cell_layer <- raster("../output/env/unique_cell.tif") #this is loading the cell layer that was created and saved as a tif
+plot(cell_layer)
+```
+and just grab the attributes of the cell_layer
+```{r}
+cell_layer
+```
+
+The occurrence data needs to be converted into spatialpoints dataframe
+```{r}
+library(sp)
+xy <- adddata[ ,c(3,2)] #3 = decimalLongitude, 2 = decimalLatitude
+adddata_aea <- SpatialPointsDataFrame(coords = xy, data = adddata, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
+adddata_aea
+```
+
+```{r}
+adddata_aea <- spTransform(adddata_aea, CRS = "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs") #note -60 used to be -91.
+Changed to 'straighten up'
+adddata_aea
+```
+
+and plot and Write the transformed species data to a new csv (this will also have the coordinates in meters - note meters are under decimalLongitude1 and decimalLatitude1)...
+```{r}
+plot(adddata_aea, axes=TRUE, cex.axis=.95)
+write.csv(adddata_aea, file = "../output/bio/data_aea.csv", row.names = FALSE)
+```
+
+Ok now plot the env. and point data...
+```{r}
+plot(cell_layer)
+points(adddata_aea$decimalLongitude, adddata_aea$decimalLatitude)
+```
+
+ok now the next step is to extract the unique cell value and add to each of the points
+then write a .csv
+then convert back to a normal dataframe
+```{r}
+adddata_aea$cell_id <- extract(x=cell_layer, y = adddata_aea)
+write.csv(adddata_aea, file = "../output/bio/data_aea_cell.csv", row.names = FALSE)
+data_aea <- as.data.frame(adddata_aea)
+head(data_aea)
+```
+looking at the header data, there seems to be some new columns created on the way...
+```{r}
+colnames(data_aea)
+```
+I don't want "optional" - get rid of it!
+```{r}
+data_aea <- subset(data_aea, select = -c(optional))
+colnames(data_aea)
+write.csv(adddata_aea, file = "../output/bio/data_aea_cell.csv", row.names = FALSE)
+```

### 2018-09-18 src/environmental_data_preperation.Rmd
fixed the issue with projecting the species data to aea
unique cell correction - GLORYS and BIOMER matched to create single unique cell

@@
 Preperation of GLORYS and BIOMER netCDFs obtained from [Copernicus](www.marine.copernicus.eu/services-portfolio/access-to-products/). These databases give global oceanographic conditions.
-The BIOMER datasets are in netCDF version 3, and the GLORYS in netCDF version 1
 
@@
 # read in a netcdf
+note this is a GLORYS dataset
 ```{r}
@@
 aea
-writeRaster(aea[[1]],'../data/env/aea.tif',overwrite=TRUE,options=c('TFW=YES'))
+writeRaster(aea[[1]],'../output/env/aea.tif',overwrite=TRUE,options=c('TFW=YES'))
 ```
@@
 rp
-writeRaster(rp,'../data/env/rp.tif',options=c('TFW=YES')) #this last line jsut saves as a tiff...don't really nead this
+writeRaster(rp,'../data/env/rp.tif',options=c('TFW=YES'))
 
@@
 so anything under 0.
-the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (11,425)
+the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (11,424)
 ```{r}
@@
 ```
-ok so we have 4693 cells that have NA (or in reality a negative number that represents NA). so remove 4693 from ncell. We know that there are 11424 cells, soo... 11424-4693 = 6732 cells have data.
+ok so we have 4693 cells that have NA (or in reality a negative number that represents NA). so remove 4693 from ncell. We know that there are 11424 cells, soo... 11424-4693 = 6731 cells have data.
 
@@
 ```
+
+Done!
+
+-----------------------------------------
+#The Biomer Layers...
+
+ok so that unique cell layer was built from a GLORYS layer...
+in the biomer layer needs to match to this layer (and i have reason to think that it does not - the biomer layer was converted by me from an ORCA025 grid, and IIRC when I looked at the layers in ArcGIS there were some differences in the cell boundaries)
+
+#read in a biomer netcdf
+```{r}
+library("ncdf4")
+biobrick <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_BIO_001_018_chl_monthly_mean_1998-01.ncwgs.nc_lonlat.ncdepthtrim.nc")
+biobrick
+```
+
+#Project the biomer file to match the glorys layer
+```{r}
+bio_aea <- projectRaster(biobrick, aea) #note this only projects the top layer
+bio_aea
+writeRaster(bio_aea,'../output/env/bio_aea.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
+```
+
+plot
+```{r}
+plot(bio_aea)
+```
+ok now lets do the same "create unique cell ID" process for the biomer layer as you did for GLORYS. You want to check they match
+
+Because the NA values are currently attributed with some other number and not NA in R (NA is set to -32767 in this particular layer), you need to make sure to exclude the cells which represent NA...
+so anything under 0.
+the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (11,424) 
+```{r}
+bio_unique_cell <- bio_aea
+bio_cell_val_freq <- freq(bio_unique_cell, value=NULL)
+bio_cell_val_freq
+```
+ok so we have 4695 cells that have NA.
+Note that the GLORYS layer has 4693 NA cells....
+a quick visual inspection with ArcGIS reveals one cell likes off the coast of the USA south of Nova Scotia (unique cell ID 4840) and one off the north tip of Nova Scotia (unique cell ID 3438)
+so remove 4840 from ncell. We know that there are 11424 cells, soo... 11424-4695 = 6729 cells have data.
+
+you need to remove cell 3438 and 4695 from the Glorys unique cell layer....
+```{r}
+values(unique_cell)[values(unique_cell) == 3438] = NA
+values(unique_cell)[values(unique_cell) == 4840] = NA
+```
+
+look at the number of NAs again...
+```{r}
+uc_na <- freq(unique_cell,value=NA)
+uc_na
+```
+
+ok great. ok so now give a unique ID to each BIOMER cell
+```{r}
+bio_unique_cell[bio_unique_cell > -1] <- 1:6729 
+writeRaster(bio_unique_cell,'../output/env/bio_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
+bio_unique_cell
+```
+now give a unique ID to the GLORYS layer with the cells removed...
+```{r}
+glo_unique_cell <- unique_cell
+glo_unique_cell[glo_unique_cell > -1] <- 1:6729 
+writeRaster(glo_unique_cell,'../output/env/glo_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
+bio_unique_cell
+```
+
+and done (again)

### 2018-09-18 src/biological_data_prep.Rmd
updated to use new cell id layer

@@
 library(raster)
-cell_layer <- raster("../output/env/unique_cell.tif") #this is loading the cell layer that was created and saved as a tif
+cell_layer <- raster("../output/env/glo_unique_cell.tif") #this is loading the cell layer that was created and saved as a tif
 plot(cell_layer)

### 2018-09-18 src/environmental_data_preperation.Rmd
AMO Data Prep

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("raster", "rgdal") 
+packages <- c("raster", "rgdal", "dplyr") 
 source("../src/ipak.R")
@@
 and done (again)
+
+-------------------------------------------
+
+#AMO data 
+Downloaded from [NOAA](https://www.esrl.noaa.gov/psd/data/timeseries/AMO/) (AMO, unsmoothed, short) as a txt and then converted to a .csv
+
+load raw data
+```{r}
+amoraw <- read.csv("../data/env/amon.us.data.csv", header = FALSE) 
+head(amoraw)
+```
+ok first col = year, the others represent the month. Give the columns some proper names
+```{r}
+colnames(amoraw) <- c("year","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
+head(amoraw)
+```
+
+The data downloaded from the site indidcated that NA values are represented by -99.99. Lets replace those values with NA 
+```{r}
+amoraw <- replace(amoraw,amoraw==-99.99, NA)
+```
+
+Now subset the dataset to start in 1998 (this is when the capelin dataset for the region starts)
+```{r}
+amoraw <- amoraw[amoraw$year >= 1998, ]
+head(amoraw)
+```
+
+And just for fun, plot the AMO... follow the tutorial [here](http://www.stat.pitt.edu/stoffer/tsa4/R_toot.htm)
+```{r}
+amocolumn <- matrix(t(amoraw[ ,2:13]), ncol=1) #month data is in col 2 - 13
+amocolumn1 <- as.vector(amocolumn)
+amotimeseries <- ts(amocolumn1, frequency = 12, start = c(1998,1))
+plot.ts(amotimeseries)
+abline(h = 0)
+dev.copy(png, "../output/env/amo.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+```
+
+ok lets try a different plot, lets try see which years were on average positive (above 0) and which were negative (below 0)
+
+First I will create a matrix that takes all of the raw data and then appends some extra columns - year average value, year phase (warm or cold) and then seasonal values (spring, summer, autum, winter) + seasonal phase (warm cold)
+
+Spring:	1 March - 31 May (col 4:6) Summer: 1 June - 31 August (col 7:9) Autumn: 1 Sept - 31 Nov (col 10:12) Winter: 1 Dec - 28 Feb (col 13, 1:2)
+```{r}
+amophases <- amoraw 
+amophases$YearAvg <- rowMeans(amophases[2:13], na.rm = TRUE) #the month data is in col 2-13
+amophases$SpringAvg <-  rowMeans(amophases[4:6], na.rm = TRUE)
+amophases$SummerAvg <-  rowMeans(amophases[7:9], na.rm = TRUE)
+amophases$AutumnAvg <-  rowMeans(amophases[10:12], na.rm = TRUE)
+# winter has a different format - the columns are in 1, 2 & 3 which means the rowmeans function needs a bit of tweaking
+amophases$WinterAvg <- rowMeans(amophases[, c(13, 2, 3)], na.rm = TRUE) #selects just cols 12, 2, & 3
+```
+
+Now label the phases positive, negative (or avg if any happen to be at 0)
+```{r}
+amophases$YearPhase[amophases$YearAvg ==0] <- "avg"
+amophases$YearPhase[amophases$YearAvg >0] <- "positive"
+amophases$YearPhase[amophases$YearAvg <0] <- "negative"
+amophases$SpringPhase[amophases$SpringAvg ==0] <- "avg"
+amophases$SpringPhase[amophases$SpringAvg >0] <- "positive"
+amophases$SpringPhase[amophases$SpringAvg <0] <- "negative"
+amophases$SummerPhase[amophases$SummerAvg ==0] <- "avg"
+amophases$SummerPhase[amophases$SummerAvg >0] <- "positive"
+amophases$SummerPhase[amophases$SummerAvg <0] <- "negative"
+amophases$AutumnPhase[amophases$AutumnAvg ==0] <- "avg"
+amophases$AutumnPhase[amophases$AutumnAvg >0] <- "positive"
+amophases$AutumnPhase[amophases$AutumnAvg <0] <- "negative"
+amophases$WinterPhase[amophases$WinterAvg ==0] <- "avg"
+amophases$WinterPhase[amophases$WinterAvg >0] <- "positive"
+amophases$WinterPhase[amophases$WinterAvg <0] <- "negative"
+```
+
+now save this as a new csv
+```{r}
+write.csv(amophases, file = "../output/env/amo.csv", row.names = FALSE)
+```

### 2018-09-18 src/biological_data_prep.Rmd
trying to extract amo data to the species dataset...

@@
 ```
+
+# Add AMO Data
+The data has been prepared as per the environmental_data_preperation.Rmd
+
+```{r}
+amo <- read.csv("../output/env/amo.csv", header=TRUE)
+head(amo)
+```
+
+So you want..
+- AMO value at sampling month/year
+- AMO value at previous sampling month/year
+- AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
+
+change month col names to numeric
+
+```{r}
+names(amo)[2:13] = c("1", "2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
+head(amo)
+```
+
+to test this just subset some of your species data...
+```{r}
+test <- data_aea <- data_aea[data_aea$year < 2000,  ]
+```
+
+ok now do something...
+
+subset cols 1:13
+```{r}
+testamo <- amo <- amo[ ,1:13]
+testamo
+```
+
+```{r}
+testamo <- reshape(testamo, 
+  varying = c("1", "2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12"),
+  v.names = "value",
+  direction = "long")
+
+testamo.sort <- testamo[order(testamo$year),]
+testamo
+```
+
+ok... soo now something like
+where test$year = testamo$year AND test$month = testamo$time
+  extract value
+    add to new test$amo_sample

### 2018-09-19 src/biological_data_prep.Rmd
amo_sample added

@@
 
-to test this just subset some of your species data...
+Start with the month data first...subset cols 1:13 (where the year and month data are held)
 ```{r}
-test <- data_aea <- data_aea[data_aea$year < 2000,  ]
-```
-
-ok now do something...
-
-subset cols 1:13
-```{r}
-testamo <- amo <- amo[ ,1:13]
-testamo
+amo_yrmt <- amo[ ,1:13]
+amo_yrmt
 ```
@@
 ```{r}
-testamo <- reshape(testamo, 
+amo_yrmt <- reshape(amo_yrmt, 
   varying = c("1", "2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12"),
@@
 
-testamo.sort <- testamo[order(testamo$year),]
-testamo
+amo_yrmt.sort <- amo_yrmt[order(amo_yrmt$year),]
+amo_yrmt
+```
+rename time to month.. and value to amo_sample
+```{r}
+colnames(amo_yrmt)[2] <- "month"
+colnames(amo_yrmt)[3] <- "amo_sample"
+amo_yrmt
 ```
 
-ok... soo now something like
-where test$year = testamo$year AND test$month = testamo$time
-  extract value
-    add to new test$amo_sample
+now match the year and the month in the two dataframes, and populate the species dataset (data_aea) with amo_sample
+```{r}
+data_aea <- merge(x = data_aea, y = amo_yrmt[ , c("year", "month", "amo_sample")], by = c("year", "month"), all.x = TRUE)
+data_aea
+```
+
+ok

### 2018-09-19 src/environmental_data_preperation.Rmd
trying to extract amo data to the species dataset...
amo_sample added
amo timing correction

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("raster", "rgdal", "dplyr") 
+packages <- c("raster", "rgdal") 
 source("../src/ipak.R")
@@
 
-#if projectRaster woudl work properly...
+#if projectRaster would work properly...
 
@@
 
-Now subset the dataset to start in 1998 (this is when the capelin dataset for the region starts)
+Now subset the dataset to start in 1997 (the capelin data starts in 1998, but you want some previous time period stuff)
 ```{r}
-amoraw <- amoraw[amoraw$year >= 1998, ]
+amoraw <- amoraw[amoraw$year >= 1997, ]
 head(amoraw)
@@
 ```
-
-ok lets try a different plot, lets try see which years were on average positive (above 0) and which were negative (below 0)
 

### 2018-09-19 src/biological_data_prep.Rmd
amo prev

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble")
+packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful")
 source("../src/ipak.R")
@@
 
-amo_yrmt.sort <- amo_yrmt[order(amo_yrmt$year),]
+amo_yrmt <- amo_yrmt[order(amo_yrmt$year),]
 amo_yrmt
@@
 
-ok
+ok next - AMO value at previous sampling month/year
+
+shift the amo_sample column down by 1...
+```{r}
+library(useful)
+amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", len = 1)
+amo_yrmt
+```
+
+Note the shifted values are in a new column called amo_sample.Shifted. rename to amo_prev
+```{r}
+colnames(amo_yrmt)[colnames(amo_yrmt)=="amo_sample.Shifted"] <- "amo_prev"
+amo_yrmt
+```
+
+```{r}
+data_aea <- merge(x = data_aea, y = amoshift[ , c("year", "month", "amo_prev")], by = c("year", "month"), all.x = TRUE)
+data_aea
+```
+
+and now - AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
+
+same process as before - shift row then match
+
+But first!
+
+You need to extract the seasonal values from the amo dataset
+```{r}
+
+```
+
+```{r}
+amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", len = 1)
+amo_yrmt
+```

### 2018-09-19 src/environmental_data_preperation.Rmd
winter calculation correction

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("raster", "rgdal") 
+packages <- c("ncdf4", "raster", "rgdal", "useful") 
 source("../src/ipak.R")
@@
 
-Spring:	1 March - 31 May (col 4:6) Summer: 1 June - 31 August (col 7:9) Autumn: 1 Sept - 31 Nov (col 10:12) Winter: 1 Dec - 28 Feb (col 13, 1:2)
+Spring:	1 March - 31 May (col 4:6) Summer: 1 June - 31 August (col 7:9) Autumn: 1 Sept - 31 Nov (col 10:12) Winter: 1 Dec - 28 Feb (col 13 from prev year, 1:2)
 ```{r}
@@
 amophases$AutumnAvg <-  rowMeans(amophases[10:12], na.rm = TRUE)
-# winter has a different format - the columns are in 1, 2 & 3 which means the rowmeans function needs a bit of tweaking
-amophases$WinterAvg <- rowMeans(amophases[, c(13, 2, 3)], na.rm = TRUE) #selects just cols 12, 2, & 3
+```
+winter has a different format - the columns are in 1 & 2, & 13 from the previous column, which means the rowmeans function needs a bit of tweaking...
+
+1) shift the dec column you created down by 1
+```{r}
+library(useful)
+amophases <- shift.column(data = amophases, columns = "Dec", len = 1, up = FALSE)
+amophases
+write.csv(amophases, file = "../output/env/amo.csv", row.names = FALSE) #for some reason you need to write to a csv and then reload for the next step to work properly...
+```
+
+the shifted data is in Dec.Shifted...
+```{r}
+amophases <- read.csv("../output/env/amo.csv")
+amophases$WinterAvg <- rowMeans(amophases[, c("Dec.Shifted", "Jan", "Feb")], na.rm = TRUE) 
+amophases
 ```

### 2018-09-19 src/biological_data_prep.Rmd
amo data extraction

@@
 library(useful)
-amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", len = 1)
+amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", up = FALSE)
 amo_yrmt
@@
 amo_yrmt
+write.csv(amo_yrmt, "../output/env/amo_prev.csv", row.names = FALSE) #for some reason i need to write to a csv and reload for the next step to work properly
 ```
 
+now grab the data...
 ```{r}
-data_aea <- merge(x = data_aea, y = amoshift[ , c("year", "month", "amo_prev")], by = c("year", "month"), all.x = TRUE)
+amo_prev <- read.csv("../output/env/amo_prev.csv")
+data_aea <- merge(x = data_aea, y = amo_prev[ , c("year", "month", "amo_prev")], by = c("year", "month"), all.x = TRUE)
 data_aea
@@
 ```{r}
-
+amo_winter <- subset(amo, select = c("year", "WinterAvg"))
+amo_winter
 ```
 
+now shift the winter average data 
 ```{r}
-amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", len = 1)
-amo_yrmt
+amo_winter <- shift.column(data = amo_winter, columns = "WinterAvg", up = FALSE)
+amo_winter
+write.csv(amo_winter, "../output/env/amo_winter.csv", row.names = FALSE) #for some reason need to save to csv then reload in the next step...
 ```
+remember your shifted column is called WinterAvg.Shifted
+
+now extract the WinterAvg.Shifted values to the species dataset
+```{r}
+amo_winter <- read.csv("../output/env/amo_winter.csv")
+data_aea <- merge(x = data_aea, y = amo_winter[ , c("year", "WinterAvg.Shifted")], by = c("year"), all.x = TRUE)
+colnames(data_aea)[colnames(data_aea)=="WinterAvg.Shifted"] <- "amo_winter" #change the colname
+data_aea
+```
+ok and write a csv
+```{r}
+write.csv(data_aea,"../output/bio/data_aea_cell_amo.csv", row.names = FALSE)
+```

### 2018-09-19 src/environmental_data_preperation.Rmd
amo data extraction

@@
 
-Now subset the dataset to start in 1997 (the capelin data starts in 1998, but you want some previous time period stuff)
+Now subset the dataset to start in 1996 (the capelin data starts in 1998, but you want some previous time period stuff)
 ```{r}
-amoraw <- amoraw[amoraw$year >= 1997, ]
+amoraw <- amoraw[amoraw$year >= 1996, ]
 head(amoraw)

### 2018-09-19 src/environmental_data_preperation.Rmd
NAO data done

@@
 # Overview
-Preperation of GLORYS and BIOMER netCDFs obtained from [Copernicus](www.marine.copernicus.eu/services-portfolio/access-to-products/). These databases give global oceanographic conditions.
+Preperation of GLORYS and BIOMER netCDFs obtained from [Copernicus](www.marine.copernicus.eu/services-portfolio/access-to-products/). These databases give global oceanographic condition.
+UPDATE: also includes prep of the AMO and NAO data...
 
@@
 
-# read in a netcdf
+# NetCDFs (Biomer and GLORYS)
+read in a netcdf
 note this is a GLORYS dataset
@@
 
-# reproject 
+reproject ...
 There is currently an issue with the projectRaster command in the raster package. Michael Sumners kindly offered a [temporary solution](https://twitter.com/mdsumner/status/1040500830539935744)
@@
 
-#if projectRaster would work properly...
+if projectRaster would work properly...
 
@@
 
-#create unique id layer
+# create unique id layer
 ok now you want to create a duplicate of the top layer and then give each cell an unique ID
@@
 -----------------------------------------
-#The Biomer Layers...
+# The Biomer Layers...
 
@@
 
-#read in a biomer netcdf
+read in a biomer netcdf
 ```{r}
@@
 
-#Project the biomer file to match the glorys layer
+Project the biomer file to match the glorys layer
 ```{r}
@@
 
-#AMO data 
+# AMO data
+
 Downloaded from [NOAA](https://www.esrl.noaa.gov/psd/data/timeseries/AMO/) (AMO, unsmoothed, short) as a txt and then converted to a .csv
@@
 amocolumn1 <- as.vector(amocolumn)
-amotimeseries <- ts(amocolumn1, frequency = 12, start = c(1998,1))
+amotimeseries <- ts(amocolumn1, frequency = 12, start = c(1996,1))
 plot.ts(amotimeseries)
@@
 ```
+
+# NAO
+
+Dataset mes from [NCAR](https://climatedataguide.ucar.edu/climate-data/hurrell-north-atlantic-oscillation-nao-index-pc-based).
+Note I am using the monthly principal component (PC)-based indices of the North Atlantic Oscillation [(NAO)](https://climatedataguide.ucar.edu/sites/default/files/nao_pc_monthly.txt).
+I have already converted this file into a csv and removed unncesessary text.
+Note there are no headers...
+
+```{r}
+naoraw <- read.csv("../data/env/amon.us.data.csv", header = FALSE) 
+head(naoraw)
+```
+
+lets give some column names. col 1  year, 2:13 are the months
+```{r}
+colnames(naoraw) <- c("year","1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
+head(naoraw)
+```
+
+ NA values are represented by -99.99. Change these to NA
+```{r}
+naoraw <- replace(naoraw,naoraw==-99.99, NA)
+```
+
+Now subset the dataset to start in 1996 (the capelin data starts in 1998, but you want some previous time period stuff)
+
+```{r}
+nao <- naoraw[naoraw$year >= 1996, ]
+head(nao) 
+```
+
+And just for fun, plot the NAO follow the tutorial here
+
+```{r}
+naocolumn <- matrix(t(amoraw[ ,2:13]), ncol=1) #month data is in col 2 - 13
+naocolumn1 <- as.vector(naocolumn)
+naotimeseries <- ts(naocolumn1, frequency = 12, start = c(1996,1))
+plot.ts(naotimeseries)
+abline(h = 0)
+dev.copy(png, "../output/env/nao.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+```
+
+First I will create a matrix that takes all of the raw data and then appends some extra columns - year average value, year phase (warm or cold) and then seasonal values (spring, summer, autum, winter) + seasonal phase (warm cold)
+
+Spring:	1 March - 31 May (col 4:6) Summer: 1 June - 31 August (col 7:9) Autumn: 1 Sept - 31 Nov (col 10:12) Winter: 1 Dec - 28 Feb (col 13 from prev year, 1:2)
+
+```{r}
+naophases <- nao 
+naophases$YearAvg <- rowMeans(naophases[2:13], na.rm = TRUE) #the month data is in col 2-13
+naophases$SpringAvg <- rowMeans(naophases[4:6], na.rm = TRUE)
+naophases$SummerAvg <- rowMeans(naophases[7:9], na.rm = TRUE)
+naophases$AutumnAvg <- rowMeans(naophases[10:12], na.rm = TRUE)
+```
+
+winter has a different format - the columns are in 1 & 2, & 13 from the previous column, which means the rowmeans function needs a bit of tweaking...
+
+shift the dec column you created down by 1
+```{r}
+library(useful)
+naophases <- shift.column(data = naophases, columns = "12", len = 1, up = FALSE)
+naophases
+write.csv(naophases, file = "../output/env/nao.csv", row.names = FALSE) #for some reason you need to write to a csv and then reload for the next step to work properly...
+
+```
+
+he shifted data is in 12.Shifted...
+```{r}
+naophases <- read.csv("../output/env/nao.csv")
+naophases$WinterAvg <- rowMeans(naophases[ , c("X12.Shifted", "X1", "X2")], na.rm = TRUE)  #x because apparently all the col names start with x for some reason...
+naophases
+```
+
+Now label the phases positive, negative (or avg if any happen to be at 0)
+
+```{r}
+naophases$YearPhase[naophases$YearAvg ==0] <- "avg"
+naophases$YearPhase[naophases$YearAvg >0] <- "positive"
+naophases$YearPhase[naophases$YearAvg <0] <- "negative"
+naophases$SpringPhase[naophases$SpringAvg ==0] <- "avg"
+naophases$SpringPhase[naophases$SpringAvg >0] <- "positive"
+naophases$SpringPhase[naophases$SpringAvg <0] <- "negative"
+naophases$SummerPhase[naophases$SummerAvg ==0] <- "avg"
+naophases$SummerPhase[naophases$SummerAvg >0] <- "positive"
+naophases$SummerPhase[naophases$SummerAvg <0] <- "negative"
+naophases$AutumnPhase[naophases$AutumnAvg ==0] <- "avg"
+naophases$AutumnPhase[naophases$AutumnAvg >0] <- "positive"
+naophases$AutumnPhase[naophases$AutumnAvg <0] <- "negative"
+naophases$WinterPhase[naophases$WinterAvg ==0] <- "avg"
+naophases$WinterPhase[naophases$WinterAvg >0] <- "positive"
+naophases$WinterPhase[naophases$WinterAvg <0] <- "negative"
+```
+
+now save this as a new csv
+
+```{r}
+write.csv(naophases, file = "../output/env/nao.csv", row.names = FALSE)
+```

### 2018-09-19 src/biological_data_prep.Rmd
NAO data added

@@
 ```
+
+#NAO values
+
+The data has been prepared as per the environmental_data_preperation.Rmd
+
+```{r}
+nao <- read.csv("../output/env/nao.csv", header=TRUE)
+head(nao)
+```
+
+So you want..
+- AMO value at sampling month/year
+- AMO value at previous sampling month/year
+- AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
+
+change month col names to just numeric
+```{r}
+names(nao)[2:13] = c("1", "2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
+head(nao)
+```
+
+Start with the month data first...subset cols 1:13 (where the year and month data are held)
+
+```{r}
+nao_yrmt <- nao[ ,1:13]
+nao_yrmt
+```
+
+```{r}
+nao_yrmt <- reshape(nao_yrmt, 
+  varying = c("1", "2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12"),
+  v.names = "value",
+  direction = "long")
+
+nao_yrmt <- nao_yrmt[order(amo_yrmt$year),]
+nao_yrmt
+```
+
+rename time to month.. and value to nao_sample
+
+```{r}
+colnames(nao_yrmt)[2] <- "month"
+colnames(nao_yrmt)[3] <- "nao_sample"
+nao_yrmt
+```
+
+now match the year and the month in the two dataframes, and populate the species dataset (data_aea) with nao_sample
+
+```{r}
+data_aea <- merge(x = data_aea, y = nao_yrmt[ , c("year", "month", "nao_sample")], by = c("year", "month"), all.x = TRUE)
+data_aea
+```
+
+ok next - NAO value at previous sampling month/year
+
+shift the nao_sample column down by 1...
+
+```{r}
+library(useful)
+nao_yrmt <- shift.column(data = nao_yrmt, columns = "nao_sample", up = FALSE)
+nao_yrmt
+```
+
+Note the shifted values are in a new column called amo_sample.Shifted. rename to amo_prev
+
+```{r}
+colnames(nao_yrmt)[colnames(nao_yrmt)=="nao_sample.Shifted"] <- "nao_prev"
+nao_yrmt
+write.csv(nao_yrmt, "../output/env/nao_prev.csv", row.names = FALSE) #for some reason i need to write to a csv and reload for the next step to work properly
+```
+
+now grab the data...
+
+```{r}
+nao_prev <- read.csv("../output/env/nao_prev.csv")
+data_aea <- merge(x = data_aea, y = nao_prev[ , c("year", "month", "nao_prev")], by = c("year", "month"), all.x = TRUE)
+data_aea
+```
+
+and now - NAO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
+
+same process as before - shift row then match
+
+But first!
+
+You need to extract the seasonal values from the NAO dataset
+
+```{r}
+nao_winter <- subset(nao, select = c("year", "WinterAvg"))
+nao_winter
+```
+
+now shift the winter average data
+
+```{r}
+nao_winter <- shift.column(data = nao_winter, columns = "WinterAvg", up = FALSE)
+nao_winter
+write.csv(nao_winter, "../output/env/nao_winter.csv", row.names = FALSE) #for some reason need to save to csv then reload in the next step...
+
+```
+
+remember your shifted column is called WinterAvg.Shifted
+
+now extract the WinterAvg.Shifted values to the species dataset
+
+```{r}
+nao_winter <- read.csv("../output/env/nao_winter.csv")
+data_aea <- merge(x = data_aea, y = nao_winter[ , c("year", "WinterAvg.Shifted")], by = c("year"), all.x = TRUE)
+colnames(data_aea)[colnames(data_aea)=="WinterAvg.Shifted"] <- "nao_winter" #change the colname
+data_aea
+```
+
+ok and write a csv
+
+```{r}
+write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao.csv", row.names = FALSE)
+```

### 2018-09-20 src/environmental_data_preperation.Rmd
export unique grid as polygon

@@
 writeRaster(unique_cell,'../output/env/unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
+uniquecell_poly <- rasterToPolygons(aea[[1]], fun=NULL, n=4, na.rm=TRUE, digits=12, dissolve=FALSE)
+
+writeOGR(uniquecell_poly, dsn = '.', layer = '../output/env/poly', driver = "ESRI Shapefile")
+
 ```

### 2018-09-28 src/raster_brick_stuff.Rmd
netcdf to rasterbrick - function testing

--- 2018-09-19 src/environmental_data_preperation.Rmd
@@
 ---
-title: "environmental_data_preperation"
+title: "raster_brick_stuff"
 author: "Samantha Andrews"
@@
 # Overview
-Preperation of GLORYS and BIOMER netCDFs obtained from [Copernicus](www.marine.copernicus.eu/services-portfolio/access-to-products/). These databases give global oceanographic conditions.
+Creating a way to automate mass raster_brick creation, with unique names...
 
@@
 Note this function checks to see if the packages are installed first.
-```{r pre-install packages, message=FALSE}
-packages <- c("ncdf4", "raster", "rgdal", "useful") 
+```{r pre-install packages, suppressMessages(expr)}
+packages <- c("rgdal", "raster", "ncdf4", "RNetCDF", "sp", "parallel", "magrittr", "data.table") 
 source("../src/ipak.R")
@@
 
-# read in a netcdf
-note this is a GLORYS dataset
+For the parallel computing, detect cores automatically (leave one free though).
+```{r}
+cores <- detectCores() - 1
+```
+
+Anne is the best - really the best. She is the best that ever was. No Anne bester.
+
 ```{r}
 library("ncdf4")
-ncbr <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
-ncbr
+source("../src/netcdf_to_rasterbrick.R")
+
 ```
-
-# reproject 
-There is currently an issue with the projectRaster command in the raster package. Michael Sumners kindly offered a [temporary solution](https://twitter.com/mdsumner/status/1040500830539935744)
-```{r}
-## use ncdf4 directly
-library(ncdf4)
-## connection to the file
-nc <- nc_open("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
-## desired variable 
-v <- ncvar_get(nc, "salinity")
-## dimension coordinates
-lon <- ncvar_get(nc, "longitude")
-lat <- ncvar_get(nc, "latitude")
-## a template, one map from the lon, lat, and v[,,1]
-template <- raster(list(x = lon, y = lat, z = v[,,1]), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ")
-
-## apply template to brick, and set values to v (but flipped in Y because that's how raster works - lat is increasing in the NetCDF, so we go upside down)
-diy <- setValues(brick(template), v[, ncol(v):1, ])
-
-## target
-prj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs" #note -60 used to be -91. Changed to 'straighten up'.
-target <- projectExtent(diy, prj)  ## get a target extent, and we'll set the exact resolution we want
-res(target) <- c(25000, 25000)
-aea <- projectRaster(diy, target)
-plot(aea[[1]]) #the number in brackets refers to the depth number
-aea
-writeRaster(aea[[1]],'../output/env/aea.tif',overwrite=TRUE,options=c('TFW=YES'))
-```
-
-#if projectRaster would work properly...
-
-library(raster)
-canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=- +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
-library("rgdal")
-rp <- projectRaster(ncbr, crs = canada_eq_ar_proj)
-rp
-writeRaster(rp,'../data/env/rp.tif',options=c('TFW=YES'))
-
-----------
-
-#create unique id layer
-ok now you want to create a duplicate of the top layer and then give each cell an unique ID
-
-First 
-```{r}
-aea[[1]]
-unique_cell <- (aea[[1]])
-```
-
-Because the NA values are currently attributed with some other number and not NA in R (NA is set to -32767 in this particular layer), you need to make sure to exclude the cells which represent NA...
-so anything under 0.
-the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (11,424)
-```{r}
-cell_val_freq <- freq(unique_cell, value=NULL)
-cell_val_freq
-```
-ok so we have 4693 cells that have NA (or in reality a negative number that represents NA). so remove 4693 from ncell. We know that there are 11424 cells, soo... 11424-4693 = 6731 cells have data.
-
-now give a unique ID to each cell
-```{r}
-unique_cell[unique_cell > 0] <- 1:6731 
-writeRaster(unique_cell,'../output/env/unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
-```
-
-Done!
-
------------------------------------------
-#The Biomer Layers...
-
-ok so that unique cell layer was built from a GLORYS layer...
-in the biomer layer needs to match to this layer (and i have reason to think that it does not - the biomer layer was converted by me from an ORCA025 grid, and IIRC when I looked at the layers in ArcGIS there were some differences in the cell boundaries)
-
-#read in a biomer netcdf
-```{r}
-library("ncdf4")
-biobrick <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_BIO_001_018_chl_monthly_mean_1998-01.ncwgs.nc_lonlat.ncdepthtrim.nc")
-biobrick
-```
-
-#Project the biomer file to match the glorys layer
-```{r}
-bio_aea <- projectRaster(biobrick, aea) #note this only projects the top layer
-bio_aea
-writeRaster(bio_aea,'../output/env/bio_aea.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
-```
-
-plot
-```{r}
-plot(bio_aea)
-```
-ok now lets do the same "create unique cell ID" process for the biomer layer as you did for GLORYS. You want to check they match
-
-Because the NA values are currently attributed with some other number and not NA in R (NA is set to -32767 in this particular layer), you need to make sure to exclude the cells which represent NA...
-so anything under 0.
-the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (11,424) 
-```{r}
-bio_unique_cell <- bio_aea
-bio_cell_val_freq <- freq(bio_unique_cell, value=NULL)
-bio_cell_val_freq
-```
-ok so we have 4695 cells that have NA.
-Note that the GLORYS layer has 4693 NA cells....
-a quick visual inspection with ArcGIS reveals one cell likes off the coast of the USA south of Nova Scotia (unique cell ID 4840) and one off the north tip of Nova Scotia (unique cell ID 3438)
-so remove 4840 from ncell. We know that there are 11424 cells, soo... 11424-4695 = 6729 cells have data.
-
-you need to remove cell 3438 and 4695 from the Glorys unique cell layer....
-```{r}
-values(unique_cell)[values(unique_cell) == 3438] = NA
-values(unique_cell)[values(unique_cell) == 4840] = NA
-```
-
-look at the number of NAs again...
-```{r}
-uc_na <- freq(unique_cell,value=NA)
-uc_na
-```
-
-ok great. ok so now give a unique ID to each BIOMER cell
-```{r}
-bio_unique_cell[bio_unique_cell > -1] <- 1:6729 
-writeRaster(bio_unique_cell,'../output/env/bio_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
-bio_unique_cell
-```
-now give a unique ID to the GLORYS layer with the cells removed...
-```{r}
-glo_unique_cell <- unique_cell
-glo_unique_cell[glo_unique_cell > -1] <- 1:6729 
-writeRaster(glo_unique_cell,'../output/env/glo_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
-bio_unique_cell
-```
-
-and done (again)
-
--------------------------------------------
-
-#AMO data 
-Downloaded from [NOAA](https://www.esrl.noaa.gov/psd/data/timeseries/AMO/) (AMO, unsmoothed, short) as a txt and then converted to a .csv
-
-load raw data
-```{r}
-amoraw <- read.csv("../data/env/amon.us.data.csv", header = FALSE) 
-head(amoraw)
-```
-ok first col = year, the others represent the month. Give the columns some proper names
-```{r}
-colnames(amoraw) <- c("year","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
-head(amoraw)
-```
-
-The data downloaded from the site indidcated that NA values are represented by -99.99. Lets replace those values with NA 
-```{r}
-amoraw <- replace(amoraw,amoraw==-99.99, NA)
-```
-
-Now subset the dataset to start in 1996 (the capelin data starts in 1998, but you want some previous time period stuff)
-```{r}
-amoraw <- amoraw[amoraw$year >= 1996, ]
-head(amoraw)
-```
-
-And just for fun, plot the AMO... follow the tutorial [here](http://www.stat.pitt.edu/stoffer/tsa4/R_toot.htm)
-```{r}
-amocolumn <- matrix(t(amoraw[ ,2:13]), ncol=1) #month data is in col 2 - 13
-amocolumn1 <- as.vector(amocolumn)
-amotimeseries <- ts(amocolumn1, frequency = 12, start = c(1998,1))
-plot.ts(amotimeseries)
-abline(h = 0)
-dev.copy(png, "../output/env/amo.png") #this prints a png of the plot
-dev.off() #this turns off the print command
-```
-
-First I will create a matrix that takes all of the raw data and then appends some extra columns - year average value, year phase (warm or cold) and then seasonal values (spring, summer, autum, winter) + seasonal phase (warm cold)
-
-Spring:	1 March - 31 May (col 4:6) Summer: 1 June - 31 August (col 7:9) Autumn: 1 Sept - 31 Nov (col 10:12) Winter: 1 Dec - 28 Feb (col 13 from prev year, 1:2)
-```{r}
-amophases <- amoraw 
-amophases$YearAvg <- rowMeans(amophases[2:13], na.rm = TRUE) #the month data is in col 2-13
-amophases$SpringAvg <-  rowMeans(amophases[4:6], na.rm = TRUE)
-amophases$SummerAvg <-  rowMeans(amophases[7:9], na.rm = TRUE)
-amophases$AutumnAvg <-  rowMeans(amophases[10:12], na.rm = TRUE)
-```
-winter has a different format - the columns are in 1 & 2, & 13 from the previous column, which means the rowmeans function needs a bit of tweaking...
-
-1) shift the dec column you created down by 1
-```{r}
-library(useful)
-amophases <- shift.column(data = amophases, columns = "Dec", len = 1, up = FALSE)
-amophases
-write.csv(amophases, file = "../output/env/amo.csv", row.names = FALSE) #for some reason you need to write to a csv and then reload for the next step to work properly...
-```
-
-the shifted data is in Dec.Shifted...
-```{r}
-amophases <- read.csv("../output/env/amo.csv")
-amophases$WinterAvg <- rowMeans(amophases[, c("Dec.Shifted", "Jan", "Feb")], na.rm = TRUE) 
-amophases
-```
-
-Now label the phases positive, negative (or avg if any happen to be at 0)
-```{r}
-amophases$YearPhase[amophases$YearAvg ==0] <- "avg"
-amophases$YearPhase[amophases$YearAvg >0] <- "positive"
-amophases$YearPhase[amophases$YearAvg <0] <- "negative"
-amophases$SpringPhase[amophases$SpringAvg ==0] <- "avg"
-amophases$SpringPhase[amophases$SpringAvg >0] <- "positive"
-amophases$SpringPhase[amophases$SpringAvg <0] <- "negative"
-amophases$SummerPhase[amophases$SummerAvg ==0] <- "avg"
-amophases$SummerPhase[amophases$SummerAvg >0] <- "positive"
-amophases$SummerPhase[amophases$SummerAvg <0] <- "negative"
-amophases$AutumnPhase[amophases$AutumnAvg ==0] <- "avg"
-amophases$AutumnPhase[amophases$AutumnAvg >0] <- "positive"
-amophases$AutumnPhase[amophases$AutumnAvg <0] <- "negative"
-amophases$WinterPhase[amophases$WinterAvg ==0] <- "avg"
-amophases$WinterPhase[amophases$WinterAvg >0] <- "positive"
-amophases$WinterPhase[amophases$WinterAvg <0] <- "negative"
-```
-
-now save this as a new csv
-```{r}
-write.csv(amophases, file = "../output/env/amo.csv", row.names = FALSE)
-```

### 2018-09-28 src/netcdf_to_rasterbrick.R

@@ -1,0 +1,10 @@
+# Function to take all netCDFs in one folder and transform into raster bricks
+
+netcdf_to_rasterbrick <- function( ncdf_filename_input ) {
+  mcdf_filename_input %>%
+    file.path(../data/env/netcdf, . ) %>%
+    nc_open(.) %>%
+    ncvar_get(.) %>%
+    brick(.) %>%
+    return()
+}

### 2018-09-28 src/raster_brick_stuff.Rmd
probably not useful stuff for my netcdf data extraction...

@@
 # Package dependencies
-You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
-Note this function checks to see if the packages are installed first.
+You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
+Note this function checks to see if the packages are installed first. If they are not installed, it will do so before loading
 ```{r pre-install packages, suppressMessages(expr)}
@@
 
-For the parallel computing, detect cores automatically (leave one free though).
+For the parallel computing (this will make things faster), detect cores automatically (leave one free though).
 ```{r}
@@
 
-```{r}
-library("ncdf4")
-source("../src/netcdf_to_rasterbrick.R")
-
+Tell R where to find your data and where to spit out the output
+```{r data_paths}
+envdata_path <- "../data/env/netcdf"
+projected_path <- "../data/env"
+output_path <- "../output/env"
 ```

### 2018-09-28 src/rasterbrick_to_ae.R

@@ -1,0 +1,7 @@
+# project rasterbricks into albers equal area based on a tiff that has been reprojected already (called aea)
+
+rasterbrick_to_aea <- function( brick_input ) {
+  brick_input %>%
+    projectRaster(. , proj4string(aea)) %>%
+    return(.)
+}

### 2018-10-02 src/raster_brick_stuff.Rmd
more raster loop mess...

@@
 
+Note that the original files have been renamed using Windows Powershell. See OneNote document for coding details. All files are now in the following format:
+
+yyyy_mm_variablename.nc
+
+where variablename is all alpha characters, but varies in length. E.G. 1993_01_mlp.nc
+
 For the parallel computing (this will make things faster), detect cores automatically (leave one free though).
@@
 
-Anne is the best - really the best. She is the best that ever was. No Anne bester.
-
 Tell R where to find your data and where to spit out the output
 ```{r data_paths}
-envdata_path <- "../data/env/netcdf"
-projected_path <- "../data/env"
-output_path <- "../output/env"
+weatherraster_path <- "../data/env/netcdftest"
+countryshape_path <- "./output/env"
+output_path <- "../output/env/nctest"
 ```
+
+# ---- 1. Define Functions.
+
+# -- 1.A. Define Small Subfunctions.
+
+Small function 1) Reads filename & explicitly opens it as a NetCDF file.
+```{r}
+open_netcdf_as_rasterbrick <- function( ncdf_filename_input ) {
+  ncdf_filename_input %>%
+  file.path( weatherraster_path , . ) %>%
+  nc_open( . )  %>%  # Open path as NetCDF file.
+  ncvar_get( . ) %>%  # Get NetCDF file.
+  brick( . ) %>%
+  return( . )
+}
+```
+
+# Small function 2) Transforms the raster brick to our country shapefile.
+```{r}
+match_rainbrick_to_countryshape <- function( brick_input ) {
+  brick_input %>%
+  # Reproject raster brick to the shapefile's coordinate system.
+  projectRaster( . , crs = proj4string( country_shapefile ), method = "ngb" ) %>%
+  return( . )
+}
+```
+
+# Small function 3) Extract data from a raster brick.
+```{r}
+generate_data_from_rasterandshape <- function( brick_input ) {
+	brick_input %>%
+    raster::extract( . , countryshape , TRUE, fun = mean, na.rm = TRUE  )
+    return( . )
+}
+```
+
+# Small function 4) Grab 4-digit year from input filename.
+```{r}
+grab_year_from_inputfile <- function( ncdf_filename_input ) {
+  ncdf_filename_input %>%
+  regexpr("[0-9]+", . ) %>%  # Match 4-digit year.
+  regmatches( ncdf_filename_input , . ) %>%  # Get matched REGEX from input string.
+  return( . )
+}
+```
+
+# --- 1.B. Define "BIG" Function That Extracts Dataset From a NetCDF File.
+```{r}
+generate_datatable_from_rasterbricks <- function( ncdf_filename_input ) {
+	ncdf_filename_input %>%
+	open_netcdf_as_rasterbrick( . ) %>%
+	match_rainbrick_to_countryshape( . ) %>%
+	generate_data_from_rasterandshape( . ) -> country_means_dataframe
+	ncdf_filename_input %>%
+	grab_year_from_inputfile( ) %>%
+	write.csv( country_means_dataframe ,
+				file = file.path( . , output_path ) )
+}
+```
+
+# ---- 2. Main Code: Setup Environment to Run Big Function.
+
+Start with your name of the country shapefile we're referencing.
+```{r}
+"poly.shp" %>%
+file.path(countryshape_path) %>%
+readODG( den = .  , layer = "FID" ) -> countryshape
+raster_file_list <- list.files( path = weatherraster_path ,
+pattern = ".nc" , all.files = FALSE , full.names = FALSE )
+```
+
+Generate list of NetCDF files automatically from our directory.
+Match all files ending in ".nc"
+```{r}
+raster_file_list <- list.files( path = weatherraster_path ,
+pattern = ".nc" , all.files = FALSE , full.names = FALSE )
+```
+
+Run our big function on the list of NetCDF files.
+```{r}
+mclapply( raster_file_list , generate_datatable_from_rasterbricks )
+```
+
+# ---- 3. Assemble .CSV Files using Data.Table and Lapply.
+
+Fetch all files ending in .CSV in out output path.
+```{r}
+csv_file_list <- list.files( path = output_path ,
+                             pattern = ".csv",
+                             all.files = FALSE,
+                             full.names = TRUE,
+                             recursive = FALSE )
+lapply( csv_file_list , fread , sep = "," ) %>%
+rbindlist( . ) -> big_datatable
+write.csv( big_datatable, file = file.path( output_path , "big_file_name.csv") )
+```
+
+# !!!!!
+
+```{r}
+b <- nc_open("../data/env/netcdftest/1993_01_salinity.nc")
+b
+c <- brick(b, lvar = 4)
+c
+pts <- data_aea
+ts <- extract(c, cbind(pts$decimalLongitude, pts$decimalLatitude))
+```

### 2018-10-03 src/help_extract_from_multiple_netCDFs.Rmd
seeking extraction help - tidy code to do this

---
title: "help_extract_from_multiple_netCDFs"
author: "Samantha Andrews"
output: html_notebook
---

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first. If they are not installed, it will do so before loading
```{r pre-install packages, suppressMessages(expr)}
packages <- c("rgdal", "raster", "ncdf4", "RNetCDF", "sp", "parallel", "magrittr", "data.table") 
source("../src/ipak.R")
ipak(packages)
```

# Initial setup

Detect cores automatically, I usually free one up.
```{r}
cores <- detectCores() - 1
```

Define your file paths here.
```{r}
weatherraster_path <- "/path/to/weatherrasterfiles"
countryshape_path <- "/path/to/countryshapefile"
output_path <- "/path/to/outputfiles"
```

# ---- 1. Define Functions.

# -- 1.A. Define Small Subfunctions.

Small function 1) Reads filename & explicitly opens it as a NetCDF file.
```{r}
open_netcdf_as_rasterbrick <- function( ncdf_filename_input ) {
  ncdf_filename_input %>%
  file.path( weatherraster_path , . ) %>%
  nc_open( . )  %>%  
  ncvar_get( . ) %>%   
  brick( . ) %>% # original code: brick( . ,ymn = -0, ymx = 360, xmn = -90, xmx = 90, crs = "the string for your CRS" ) %>%
  return( . )
}
```

Small function 2) Transforms the raster brick to our country shapefile.
```{r}
match_rainbrick_to_countryshape <- function( brick_input ) {
  brick_input %>%
  projectRaster( . , aea) %>% #aea represents an existing rasterbrick that is in the albers equal area   projection
  # Original code included a crop option here - my files are already cropped to the region I need
  return( . )
}
```

# Small function 3) Extract data from a raster brick.
```{r}
generate_data_from_rasterandshape <- function( brick_input ) {
	brick_input %>%
	# Take means according to the countryshape.
	# Make sure df = TRUE , so that output is a dataframe.
    raster::extract( . , countryshape ,
    				f=TRUE, fun = mean, na.rm = TRUE  )
    return( . )
}
```

# Small function 4) Grab 4-digit year from input filename.
grab_year_from_inputfile <- function( ncdf_filename_input ) {

  ncdf_filename_input %>%
  regexpr("[0-9]+", . ) %>%  # Match 4-digit year.
  regmatches( ncdf_filename_input , . ) %>%  # Get matched REGEX from input string.
  return( . )
}

# --- 1.B. Define "BIG" Function That Extracts Dataset From a NetCDF File.

generate_datatable_from_rasterbricks <- function( ncdf_filename_input ) {

	# Note: the only argument is a NetCDF filename.

	# Start with file argument and process with the sub-functions above.
	ncdf_filename_input %>%

	open_netcdf_as_rasterbrick( . ) %>%
	match_rainbrick_to_countryshape( . ) %>%
	generate_data_from_rasterandshape( . ) -> country_means_dataframe

	# Go back to the file input name, create automatic names, and save.
	ncdf_filename_input %>%

	grab_year_from_inputfile( ) %>%
	write.csv( country_means_dataframe ,
				file = file.path( . , output_path ) )
}

# ---- 2. Main Code: Setup Environment to Run Big Function.

# Start with your name of the country shapefile we're referencing.
"country_shapefile_name.shp" %>%
file.path( countryshape_path , ) %>%
readODG( den = .  , layer = "countries" ) -> countryshape

# Generate list of NetCDF files automatically from our directory.
# Match all files ending in ".nc"
raster_file_list <- list.files( path = weatherraster_path ,
pattern = ".nc" , all.files = FALSE , full.names = FALSE )

# Run our big function on the list of NetCDF files.
mclapply( raster_file_list , generate_datatable_from_rasterbricks )

# ---- 3. Assemble .CSV Files using Data.Table and Lapply.

# Fetch all files ending in .CSV in out output path.
csv_file_list <- list.files( path = output_path ,
                             pattern = ".csv",
                             all.files = FALSE,
                             full.names = TRUE,
                             recursive = FALSE )

# Take the list of saved files & "fast read" them into R.
lapply( csv_file_list , fread , sep = "," ) %>%

# Transform the list of read files into a data.table:
rbindlist( . ) -> big_datatable

# Note: Before reassembling the data, or after, you may want
# to manipulate the data so that it is in a more usable format.

# Note: You may want to setkeys() for data.table here.

# Save the big file.
write.csv( big_datatable, file = file.path( output_path , "big_file_name.csv") )

### 2018-10-03 src/biological_data_prep.Rmd
warm/cold year intro
more raster loop mess...
depth layer codes sorted!

@@
 collectioncode (this contains some info that might help with finding out gear used in the survey)
+depth
 
 ```{r select colums needed}
-tidydata <- subset(tidydata, select = c(id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, resname, originalscientificname, collectioncode))
+tidydata <- subset(tidydata, select = c(id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode))
 head(tidydata)
@@
 ```
+
+# depth bins
+the depth layers in the netCDF are literal slices.
+I want to attribute the observations with depth data to the nearest depth layer.
+I have created a .csv which lists all the depth layers, and then created depth bins with min and max values.
+
+load the depth bin csv
+```{r}
+depthbin <- read.csv("../data/env/depth_bin.csv", header = TRUE) 
+head(depthbin)
+
+```
+
+so what you want to do is look at the observation, see if it has a depth value (data_aea$depth) and if so, find the value that lies between depthbin$bin_min and depthbin$bin_max, then extract depthbin$layer_name and add to data_aea$depth_layer, and then extract depthbin$layer_no and add to data_aea$depthlayer_no
+
+first make sure that the data_aea$depth without values == NA (currently they are marked xx)
+```{r}
+data_aea$depth[data_aea$depth == "xx"] <- NA #make all xx values NA
+```
+
+the depth data in data_aea is a factor. This needs to be numeric
+```{r}
+data_aea$depth <- as.numeric(as.character(data_aea$depth))
+str(data_aea)
+```
+
+Now run some fancy if loop (courtesy of [TinglTanglBob] (https://stackoverflow.com/questions/52626995/problem-with-loop-to-add-values-from-one-dataframe-and-to-another-based-on-condi/52627728#52627728))
+```{r}
+check_depth <- function(d_temp)
+{
+  #print(d_temp)
+  if(is.na(d_temp)) return(NA) # if d_temp is na just return NA 
+  layer_name_temp <- depthbin$layer_name[which(depthbin$bin_min <= d_temp & depthbin$bin_max >= d_temp)]
+  if(length(layer_name_temp) > 1) layer_name_temp <- layer_name_temp[1] # in case there are more hits, the first one is taken
+  return(layer_name_temp)
+}
+
+data_aea$depth_layer <- sapply(data_aea$depth, check_depth)
+```
+
+ok now dow the same thing to get your layer_no
+```{r}
+check_depth2 <- function(d_temp)
+{
+  #print(d_temp)
+  if(is.na(d_temp)) return(NA) # if d_temp is na just return NA 
+  layer_no_temp <- depthbin$layer_no[which(depthbin$bin_min <= d_temp & depthbin$bin_max >= d_temp)]
+  if(length(layer_no_temp) > 1) layer_no_temp <- layer_no_temp[1] # in case there are more hits, the first one is taken
+  return(layer_no_temp)
+}
+
+data_aea$depthlayerno <- sapply(data_aea$depth, check_depth2)
+```
+
+# warm/cold year
+
+The warm/cold year index will be taken from a [DFO CSAS report](http://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2018/2018_039-eng.pdf)

### 2018-10-03 src/environmental_data_preperation.Rmd
depth layer codes sorted!

@@
 read in a biomer netcdf
+note this only reads in the top layer which is what i need right now. To get the other layers, add ", lvar = 4" to brick (see code)
 ```{r}
 library("ncdf4")
-biobrick <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_BIO_001_018_chl_monthly_mean_1998-01.ncwgs.nc_lonlat.ncdepthtrim.nc")
+biobrick <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_BIO_001_018_chl_monthly_mean_1998-01.ncwgs.nc_lonlat.ncdepthtrim.nc") # lvar = 4 
 biobrick
@@
 ```{r}
-bio_aea <- projectRaster(biobrick, aea) #note this only projects the top layer
+bio_aea <- projectRaster(biobrick, aea) 
 bio_aea

### 2018-10-03 src/help_extract_from_multiple_netCDFs.Rmd
depth layer codes sorted!

@@
   ncvar_get( . ) %>%   
-  brick( . ) %>% # original code: brick( . ,ymn = -0, ymx = 360, xmn = -90, xmx = 90, crs = "the string for your CRS" ) %>%
+  brick( . , lvar = 4) %>% # lvar = 4 to read in all depth levels (nlayers)
   return( . )
@@
 generate_data_from_rasterandshape <- function( brick_input ) {
+	# this whole section needs to be changed 
   brick_input %>%
@@
 # Small function 4) Grab 4-digit year from input filename.
+```{r}
 grab_year_from_inputfile <- function( ncdf_filename_input ) {
@@
 }
+```
 

### 2018-10-03 src/raster_brick_stuff.Rmd
depth layer codes sorted!

@@
 I am a beginner in R and have had no exposure to similar languages.
-The code herein is unlikely to be elegant and there is likely more efficient ways of running the code.
+I don't know what I'm doing.
+The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.
 
@@
 ```
+
+```{r}
+brickttest <- nc_open("../data/env/netcdftest/1993_01_salinity.nc")
+testbrick <- brick( "../data/env/netcdftest/1993_01_salinity.nc", lvar = 4)
+
+```

### 2018-10-04 src/biological_data_prep.Rmd
short update to coordinates

@@
 now extract just the Atlantic Canada data.
-The lon-lat study area (which captures the whole Atlantic Canadian EEZ up to the north of Resolution Island in the Hudson Strait/ Davis Strait) is -71.0,-43.0,62.0,38.0 
+The lon-lat study area (which captures the whole Atlantic Canadian EEZ up to the north of Resolution Island in the Hudson Strait/ Davis Strait) is -70.0,-43.0,70.0,38.0 
 ```{r map only atlantic canadian occurrences}
-tidydata <- tidydata[tidydata$decimalLongitude > -71, ]
+tidydata <- tidydata[tidydata$decimalLongitude > -70, ]
 tidydata <- tidydata[tidydata$decimalLongitude < -43, ]
 tidydata <- tidydata[tidydata$decimalLatitude > 38, ]
-plot(map, xlim = c(-180, -44), ylim =c(0,80), asp = 1) 
+plot(map, xlim = c(-180, -42), ylim =c(0,80), asp = 1) 
 points(tidydata$decimalLongitude, tidydata$decimalLatitude)
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "All Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "All Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(tidydata$decimalLongitude, tidydata$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Arc Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "Arc Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(ARC_obs$decimalLongitude, ARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "CAFF Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "CAFF Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(CAFF_obs$decimalLongitude, CAFF_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Central & Arctic Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Central & Arctic Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(DFOCENARC_obs$decimalLongitude, DFOCENARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Gulf Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Gulf Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(DFOGulf_obs$decimalLongitude, DFOGulf_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO ISDM Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO ISDM Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(DFOISDM_obs$decimalLongitude, DFOISDM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Maritimes Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Maritimes Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(DFOMTMS_obs$decimalLongitude, DFOMTMS_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Newfouundland & Labrador Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Newfouundland & Labrador Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(DFONL_obs$decimalLongitude, DFONL_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "DFO Quebec Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Quebec Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(DFOQC_obs$decimalLongitude, DFOQC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "ICES Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "ICES Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(ICES_obs$decimalLongitude, ICES_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "IMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "IMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(IMR_obs$decimalLongitude, IMR_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Maine DMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "Maine DMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(MaineDMR_obs$decimalLongitude, MaineDMR_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "NAFO Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "NAFO Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(NAFO_obs$decimalLongitude, NAFO_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "NEFSC Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "NEFSC Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(NEFSC_obs$decimalLongitude, NEFSC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "National Museum of Natural History Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "National Museum of Natural History Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(USNM_obs$decimalLongitude, USNM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 x11() #this opens the plot i a window you can interact with
-plot(map2, xlim = c(-71, -43), ylim =c(39, 62), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude"))
@@
 ```
-The point that needs to be removed is located at -44.96667.
+The point that needs to be removed is located at -44.96667 and -53,483333.
 Note this is not the full value so use some other locators to remove the row (you could just used id, but pick multiple others just to learn)
@@
 tidydata <- tidydata[!(tidydata$resname == "Fish specimens" & tidydata$datecollected == "1965-08-13 12:00:00+01"), ]
+tidydata <- tidydata[!(tidydata$resname == "Fish specimens" & tidydata$datecollected == "1967-10-25 12:00:00+01"), ]
+tidydata <- tidydata[!(tidydata$resname == "Ichthyology Collection - Royal Ontario Museum" & tidydata$datecollected == "1967-10-25 12:00:00+01"), ]
 ```
@@
 ```{r map occurrences less iceland}
-plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "All Occurrences without Iceland", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "All Occurrences without Iceland", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(tidydata$decimalLongitude, tidydata$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 ```{r}
-plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 nafo_na <- subset(adddata, nafo_zone == "xx")
-plot(map2, xlim = c(-71, -43), ylim =c(38, 62), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
@@
 ```{r}
-write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE)
-```
-So in ArcGIS I saw that these points fall on land. Some are extremely close to the coast, some further inland. There are only 132 of thesepoints, so I will just delete them from the database.
-```{r}
-adddata$nafo_zone[adddata$nafo_zone == "xx"] <- NA #make all xx values NA
-adddata <- adddata[!is.na(adddata$nafo_zone), ]
+write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE) #just write all the data you have
+```
+
+So in ArcGIS I saw that some points are extremely close to the coast, but all lie in 4T.
+```{r}
+adddata$nafo_zone[adddata$nafo_zone == "xx"] <- "4T" #make all xx values 4T
+nafo_na <- subset(adddata, nafo_zone == "xx")
+plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(nafo_na$decimalLongitude, nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+```
+all sorted!
+
+```{r}
 write.csv(adddata, file = "../data/bio/capelin_additional_tidy.csv", row.names = FALSE)
@@
 
-The warm/cold year index will be taken from a [DFO CSAS report](http://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2018/2018_039-eng.pdf)
+The warm/cold year index will be taken from a [DFO CSAS report](http://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2018/2018_039-eng.pdf) - Fig 8

### 2018-10-05 src/biological_data_prep.Rmd
env. layer correction and dataset rebuilding complete

@@
 
-Load the env. data
+Load the env. data - you can just use the glorys layer or the biomer layer as they are identical
 ```{r}
@@
 
-first make sure that the data_aea$depth without values == NA (currently they are marked xx)
+the depth data in data_aea is a factor. This needs to be numeric
+```{r}
+data_aea$depth <- as.integer(as.character(data_aea$depth))
+str(data_aea)
+```
+
+make sure that the data_aea$depth without values == NA (currently they are marked xx)
 ```{r}
 data_aea$depth[data_aea$depth == "xx"] <- NA #make all xx values NA
-```
-
-the depth data in data_aea is a factor. This needs to be numeric
-```{r}
-data_aea$depth <- as.numeric(as.character(data_aea$depth))
-str(data_aea)
 ```
@@
 
+and save
+
+```{r}
+write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins.csv", row.names = FALSE)
+```
+
 # warm/cold year

### 2018-10-05 src/environmental_data_preperation.Rmd
env. layer correction and dataset rebuilding complete

@@
 library("ncdf4")
-ncbr <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
+ncbr <- brick("../data/env/1993_01_salinity.nc")
 ncbr
@@
 ## connection to the file
-nc <- nc_open("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
+nc <- nc_open("../data/env/1993_01_salinity.nc")
 ## desired variable 
@@
 so anything under 0.
-the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (11,424)
+the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (15,402)
 ```{r}
@@
 ```
-ok so we have 4693 cells that have NA (or in reality a negative number that represents NA). so remove 4693 from ncell. We know that there are 11424 cells, soo... 11424-4693 = 6731 cells have data.
+ok so we have 7634 cells that have NA (or in reality a negative number that represents NA). so remove 7634 from ncell. We know that there are 15402 cells, soo... 15402-7634 = 7768 cells have data.
 
@@
 ```{r}
-unique_cell[unique_cell > 0] <- 1:6731 
-writeRaster(unique_cell,'../output/env/unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
+unique_cell[unique_cell > 0] <- 1:7768 
+writeRaster(unique_cell,'../output/env/glo_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
 uniquecell_poly <- rasterToPolygons(aea[[1]], fun=NULL, n=4, na.rm=TRUE, digits=12, dissolve=FALSE)
-
-writeOGR(uniquecell_poly, dsn = '.', layer = '../output/env/poly', driver = "ESRI Shapefile")
+writeOGR(uniquecell_poly, dsn = '.', layer = '../output/env/glo_unique_cell_poly', driver = "ESRI Shapefile")
 
@@
 
-ok so that unique cell layer was built from a GLORYS layer...
-in the biomer layer needs to match to this layer (and i have reason to think that it does not - the biomer layer was converted by me from an ORCA025 grid, and IIRC when I looked at the layers in ArcGIS there were some differences in the cell boundaries)
-
-read in a biomer netcdf
+Read in a biomer netcdf
 note this only reads in the top layer which is what i need right now. To get the other layers, add ", lvar = 4" to brick (see code)
@@
 library("ncdf4")
-biobrick <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_BIO_001_018_chl_monthly_mean_1998-01.ncwgs.nc_lonlat.ncdepthtrim.nc") # lvar = 4 
+biobrick <- brick("../data/env/1998_01_chl.nc") # lvar = 4 
 biobrick
@@
 so anything under 0.
-the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (11,424) 
+the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (15402) 
 ```{r}
@@
 ```
-ok so we have 4695 cells that have NA.
-Note that the GLORYS layer has 4693 NA cells....
-a quick visual inspection with ArcGIS reveals one cell likes off the coast of the USA south of Nova Scotia (unique cell ID 4840) and one off the north tip of Nova Scotia (unique cell ID 3438)
-so remove 4840 from ncell. We know that there are 11424 cells, soo... 11424-4695 = 6729 cells have data.
+ok so we have 7652 cells that have NA. Note that the GLORYS layer has 7634 NA cells....  so let's remove the cells from the GLORYS raster that correspond to NA in the BIOMER raster...
 
-you need to remove cell 3438 and 4695 from the Glorys unique cell layer....
 ```{r}
-values(unique_cell)[values(unique_cell) == 3438] = NA
-values(unique_cell)[values(unique_cell) == 4840] = NA
+aea <- overlay(aea, bio_aea, fun = function(x, y){
+  x[is.na(y[])] <- NA
+  return(x)
+})
+plot(aea[[1]])
 ```
 
-look at the number of NAs again...
 ```{r}
-uc_na <- freq(unique_cell,value=NA)
-uc_na
+aea_cut_cell_val_freq <- freq(aea[[1]], value=NULL)
+aea_cut_cell_val_freq
 ```
 
-ok great. ok so now give a unique ID to each BIOMER cell
+great! So let's redo the GLORYS unique cell layer....
+
 ```{r}
-bio_unique_cell[bio_unique_cell > -1] <- 1:6729 
+unique_cell <- aea[[1]]
+unique_cell[unique_cell > 0] <- 1:7750 
+writeRaster(unique_cell,'../output/env/glo_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
+uniquecell_poly <- rasterToPolygons(aea[[1]], fun=NULL, n=4, na.rm=TRUE, digits=12, dissolve=FALSE)
+writeOGR(uniquecell_poly, dsn = '.', layer = '../output/env/glo_unique_cell_poly', driver = "ESRI Shapefile")
+unique_cell
+```
+
+and create a unique cell biomer layer..
+
+```{r}
+bio_unique_cell <- bio_aea
+bio_unique_cell[bio_unique_cell > -1] <- 1:7750 
 writeRaster(bio_unique_cell,'../output/env/bio_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
@@
 ```
-now give a unique ID to the GLORYS layer with the cells removed...
-```{r}
-glo_unique_cell <- unique_cell
-glo_unique_cell[glo_unique_cell > -1] <- 1:6729 
-writeRaster(glo_unique_cell,'../output/env/glo_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
-bio_unique_cell
-```
-
-and done (again)
 

### 2018-10-22 src/environmental_data_preperation.Rmd
mass netcdf to raster bricks

@@
 library("ncdf4")
-ncbr <- brick("../data/env/1993_01_salinity.nc")
+testnc <-  "../data/env/1993_01_salinity.nc"
+ncbr <- brick(testnc,lvar = 4) #lvar = 4 loads the levels
 ncbr 
@@
 
+#```{r}
 library(raster)
-canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=- +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
+canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
 library("rgdal")
-rp <- projectRaster(ncbr, crs = canada_eq_ar_proj)
+rp <- projectRaster (ncbr, crs = canada_eq_ar_proj)
 rp
-writeRaster(rp,'../data/env/rp.tif',options=c('TFW=YES'))
+plot(rp[[1]])
+writeRaster(rp[[1]],'../data/env/rplayer.tif',options=c('TFW=YES'), overwrite=TRUE) 
+#```
 

### 2018-10-22 src/netcdf_to_brick_projected.Rmd
mass netcdf to raster bricks

---
title: "netcdf_to_brick_projected"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Mass netCDF to raster brick, and then reprojection...

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. The code herein is unlikely to be elegant and there is likely more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.

```{r pre-install packages, message=FALSE}
packages <- c("ncdf4")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

Ok first get a list of netcdfs in the folder...
```{r}
netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE)
netcdf_list
```

ok now create a raster brick for each netcdf (with the same file name),

```{r}
no_netcdf <- length(netcdf_list)
netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
library (ncdf4)
for (i in 1:no_netcdf) {
  print(netcdf_name[i]) #this just prings the name of the netCDF R is working on
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  assign(netcdf_name[i], temp_brick) #this asigns the netCDF name to the raster brick
}

```

### 2018-10-22 src/netcdf_to_brick_projected.Rmd
raster brick loop with projection done. working on first stages of extraction

@@
 The code herein is unlikely to be elegant and there is likely more efficient ways of running the code.
+
+SAM - YOU MIGHT WANT TO [PARALLEL PROCESS FOR THE WHOLE DATASET..](https://cran.r-project.org/web/packages/doParallel/vignettes/gettingstartedParallel.pdf) and also [here](https://gis.stackexchange.com/questions/253618/r-multicore-approach-to-extract-raster-values-using-spatial-points)
 
@@
 netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE)
-netcdf_list
 ```
 
-ok now create a raster brick for each netcdf (with the same file name),
+ok now create a raster brick for each netcdf, project, and assing the same file name as the netCDF
 
 ```{r}
-no_netcdf <- length(netcdf_list)
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
 netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
@@
   temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) #aea is an existing raster in the projection I want that I created (environmental_data_preperation)
   assign(netcdf_name[i], temp_brick) #this asigns the netCDF name to the raster brick
@@
 ```
+just a quick test....
+
+```{r}
+`1998_01_salinity.nc`
+plot(`1998_01_salinity.nc` [[8]])
+```
+
+Look up year and month in the occurence data (data_aea)
+year = data_aea$year
+month = data_aea$month
+
+the raster brick file names are 
+character 1:4 = year
+charcater 6:7 = month
+character 9:end = env. variable name
+
+maybe start by splitting the occurence data into year-month? Then you can search for the rasters without going through everything....
+```{r}
+species_obs_date <- split(data_aea,paste(data_aea$year, data_aea$month))
+species_obs_date
+```

### 2018-10-22 src/biological_data_prep.Rmd
feck all progress :(

@@
 
+#add columns for the env. variables
+The ones ypu have to add are...
+
+mlp
+salinty_1
+salinity_depth
+ssh
+temp_1
+temp_depth
+chl_1
+chl_depth
+o2_1
+o2_depth
+
+populate the columns with NA
+```{r}
+data_aea$mlp <- NA
+data_aea$salinty_1 <- NA
+data_aea$salinity_depth <- NA
+data_aea$ssh <- NA
+data_aea$temp_1 <- NA
+data_aea$temp_depth <- NA
+data_aea$chl_1 <- NA
+data_aea$chl_depth <- NA
+data_aea$o2_1 <- NA
+data_aea$o2_depth <- NA
+write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins_env.csv", row.names = FALSE)
+```
+
 # warm/cold year

### 2018-10-22 src/netcdf_to_brick_projected.Rmd
feck all progress :(

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("ncdf4")
+packages <- c("ncdf4", "sp")
 source("../src/ipak.R")
@@
 ```{r}
-species_obs_date <- split(data_aea,paste(data_aea$year, data_aea$month))
-species_obs_date
+library(sp)
+species_obs_date <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")) #convert into a spatialpoints dataframe to do the extraction
+species_obs_date <- spTransform(species_obs_date, CRS = "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs") #need to reproject....
+note -60 used to be -91.
+Changed to 'straighten up'
+split_obs <- split(species_obs_date, paste(species_obs_date$year, species_obs_date$month))
 ```
+
+just check the species_obs_date complete file in a plot
+```{r}
+plot(species_obs_date)
+```
+looks good...
+
+```{r}
+for(i in split_obs){
+  print(i)
+}
+```
+
+hmm what now....
+
+You want to populate the following fields in the species_obs_date spatialdf
+mlp
+salinty_1
+salinity_depth
+ssh
+temp_1
+temp_depth
+chl_1
+chl_depth
+o2_1
+o2_depth
+
+where _1 indicates the surface layer [[1]] and depth is == depthlayerno
+Note mlp and ssh have only one single layer ([[1]])
+
+netcdf_list = list of all netcdfs
+no_netcdf = number of netcdfs created
+netcdf_name = name of netcdfs - will match the name of the raster bricks too.
+
+```{r}
+
+```

### 2018-10-23 src/help_extract_from_multiple_netCDFs.Rmd
mass netcdf to raster bricks
asking for help....

@@
 
-# Initial setup
-
-Detect cores automatically, I usually free one up.
+# use multi-core (if it is an option)
+Detect cores automatically, and free one up.
 ```{r}
@@
 
-Define your file paths here.
+# The short story
+
+I have multiple netCDFs that need to extract values from to points. The problem is, I need R to decide which netCDFs (and which layers) to perform the extraction on.
+
+# The details
+
+I have around 1500 NetCDFs.
+Each netCDF has 4 dimensions (time - T, longitude - X, latitude - Y, depth - Z) and represents a single oceanographic variable for a particular month in a particular year.
+I have given each netCDF the following naming convention:
+
+yyyy_mm_variablename.nc (where variable name can vary in length)
+
+An example of the netCDF is available [here](https://drive.google.com/file/d/1y2XU5RzUBRLgTXkql667a5o_PVk5LMTP/view?usp=sharing) .
+To use the netCDFs in R, I have made a loop to create raster bricks and project them in an equal-area grid from all the netCDFs:
+
 ```{r}
-weatherraster_path <- "/path/to/weatherrasterfiles"
-countryshape_path <- "/path/to/countryshapefile"
-output_path <- "/path/to/outputfiles"
-```
-
-# ---- 1. Define Functions.
-
-# -- 1.A. Define Small Subfunctions.
-
-Small function 1) Reads filename & explicitly opens it as a NetCDF file.
-```{r}
-open_netcdf_as_rasterbrick <- function( ncdf_filename_input ) {
-  ncdf_filename_input %>%
-  file.path( weatherraster_path , . ) %>%
-  nc_open( . )  %>%  
-  ncvar_get( . ) %>%   
-  brick( . , lvar = 4) %>% # lvar = 4 to read in all depth levels (nlayers)
-  return( . )
+netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf) {
+  print(netcdf_name[i]) #this just prings the name of the netCDF R is working on
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) #aea is an existing raster in the projection I want that I created (environmental_data_preperation)
+  assign(netcdf_name[i], temp_brick) #this asigns the netCDF name to the raster brick
 }
@@
 
-Small function 2) Transforms the raster brick to our country shapefile.
+Note that this may not be the most computationally efficient way to do this as I will be creating a lot of raster bricks that will be stored to memory…
+
+I also had a dataset (.csv) containing fish observation (point) data. This dataset (let's call it occurrence) contains the following columns:
+
+occurrence$id (a unique observation ID)
+occurrence$decimalLongitude (Longitude in decimal degree)
+occurrence$decimalLatitude (Latitude in decimal degree)
+occurrence$dLongitudeM (Longitude in meters)
+occurrence$LatitudeM (Latitude in meters)
+occurrence$year (year of observation)
+occurrence$month (month of observation)
+occurrence$depthlayerno (depth of observation - the value corresponds to the depth layers in the raster bricks e.g.
+if occurrence$depthlayerno == 20, then the observation is from the raster brick layer [[20]])
+
+I then have some empty columns that need to be populated with data from the raster bricks, each representing an oceanographic variable and depth layer the data needs to come from. An example is:
+
+occurrence$salinity_1 (extract value to point from the salinity raster brick layer [[1]])
+occurrence$salinity_depth (extract value to point from the salinity raster brick layer [[occurrence$depthlayerno]])
+
 ```{r}
-match_rainbrick_to_countryshape <- function( brick_input ) {
-  brick_input %>%
-  projectRaster( . , aea) %>% #aea represents an existing rasterbrick that is in the albers equal area projection
-  # Original code included a crop option here - my files are already cropped to the region I need
-  return( . )
-}
+occurrence <- read.csv("./data/bio/occurrence_help.csv", header = TRUE)
 ```
 
-# Small function 3) Extract data from a raster brick.
+Most of the raster bricks contain 50 layers - some only have 1 layer. Furthermore not all observations have a depthlayerno value (occurrence$depthlayerno == NA)
+
+A cut-down version of this .csv is available here XXXX.
+
+The dput for occurrence is as follows:
+
 ```{r}
-generate_data_from_rasterandshape <- function( brick_input ) {
-	# this whole section needs to be changed 
-  brick_input %>%
-	# Take means according to the countryshape.
-	# Make sure df = TRUE , so that output is a dataframe.
-    raster::extract( . , countryshape ,
-    				f=TRUE, fun = mean, na.rm = TRUE  )
-    return( . )
-}
+dput(occurrence)
 ```
 
-# Small function 4) Grab 4-digit year from input filename.
+And str(occurrence) is
+
 ```{r}
-grab_year_from_inputfile <- function( ncdf_filename_input ) {
-
-  ncdf_filename_input %>%
-  regexpr("[0-9]+", . ) %>%  # Match 4-digit year.
-  regmatches( ncdf_filename_input , . ) %>%  # Get matched REGEX from input string.
-  return( . )
-}
+str(occurrence)
 ```
 
-# --- 1.B. Define "BIG" Function That Extracts Dataset From a NetCDF File.
+I have converted the occurrence dataset into a SpatialPointsDataFrame, projected it into an equal area grid (to match the raster bricks), and them split the data into year-months (I thought this might be an easier way to tackle the next stage):
 
-generate_datatable_from_rasterbricks <- function( ncdf_filename_input ) {
+```{r}
+library(sp)
+species_obs_date <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")) #convert into a spatialpoints dataframe to do the extraction
+species_obs_date <- spTransform(species_obs_date, CRS = "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs") #need to reproject....
+note -60 used to be -91.
+Changed to 'straighten up'
+split_obs <- split(species_obs_date, paste(species_obs_date$year, species_obs_date$month))
+```
 
-	# Note: the only argument is a NetCDF filename.
+# The problem
 
-	# Start with file argument and process with the sub-functions above.
-	ncdf_filename_input %>%
+What I would like to happen is for each  each observation (row) in occurrences, for each oceanographic variable (e.g.
+using salinity) look up occurrence$year and occurrence$month and locate the appropriate raster brick, and extract values to points from layer [[1]] and add the result to occurrence$salinity_1, then look up occurrence$depthlayerno and extract values to points from layer[[depthlayerno]] and add the result to occurrence$salinity_depth.
 
-	open_netcdf_as_rasterbrick( . ) %>%
-	match_rainbrick_to_countryshape( . ) %>%
-	generate_data_from_rasterandshape( . ) -> country_means_dataframe
+Honestly, I have no idea where to begin.
+I found a rather [neat blog from Nathan Lane](http://nathanlane.info/tutorial/2016/01/02/gisfunctional.html) in which he uses functions to extract values to polygons from multiple raster bricks (but without selecting particular raster bricks to extract the data from).
+Unfortunately adapting it to my needs is way beyond my current skillset (which is a beginner).
 
-	# Go back to the file input name, create automatic names, and save.
-	ncdf_filename_input %>%
-
-	grab_year_from_inputfile( ) %>%
-	write.csv( country_means_dataframe ,
-				file = file.path( . , output_path ) )
-}
-
-# ---- 2. Main Code: Setup Environment to Run Big Function.
-
-# Start with your name of the country shapefile we're referencing.
-"country_shapefile_name.shp" %>%
-file.path( countryshape_path , ) %>%
-readODG( den = .  , layer = "countries" ) -> countryshape
-
-# Generate list of NetCDF files automatically from our directory.
-# Match all files ending in ".nc"
-raster_file_list <- list.files( path = weatherraster_path ,
-pattern = ".nc" , all.files = FALSE , full.names = FALSE )
-
-# Run our big function on the list of NetCDF files.
-mclapply( raster_file_list , generate_datatable_from_rasterbricks )
-
-# ---- 3. Assemble .CSV Files using Data.Table and Lapply.
-
-# Fetch all files ending in .CSV in out output path.
-csv_file_list <- list.files( path = output_path ,
-                             pattern = ".csv",
-                             all.files = FALSE,
-                             full.names = TRUE,
-                             recursive = FALSE )
-
-# Take the list of saved files & "fast read" them into R.
-lapply( csv_file_list , fread , sep = "," ) %>%
-
-# Transform the list of read files into a data.table:
-rbindlist( . ) -> big_datatable
-
-# Note: Before reassembling the data, or after, you may want
-# to manipulate the data so that it is in a more usable format.
-
-# Note: You may want to setkeys() for data.table here.
-
-# Save the big file.
-write.csv( big_datatable, file = file.path( output_path , "big_file_name.csv") )
+Any help you can offer would be gratefully received!

### 2018-10-23 src/netcdf_to_brick_projected.Rmd
asking for help....

@@
 ```{r}
-netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE)
+netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
 ```
@@
 just a quick test....
-
+(also not sure why the names are wrapped in '')
 ```{r}
@@
 
-You want to populate the following fields in the species_obs_date spatialdf
+I want to populate the following fields in the species_obs_date spatialdf
 mlp
@@
 
-where _1 indicates the surface layer [[1]] and depth is == depthlayerno
+where _1 indicates the surface layer [[1]] of each environmental variable, and depth is the == species_obs_date$depthlayerno
 Note mlp and ssh have only one single layer ([[1]])

### 2018-10-23 src/help_extract_from_multiple_netCDFs.Rmd
link to csv
figuring out formatting...

@@
 
-yyyy_mm_variablename.nc (where variable name can vary in length)
+'yyyy_mm_variablename.nc' (where variable name can vary in length)
 
@@
 
-A cut-down version of this .csv is available here XXXX.
+A cut-down version of this .csv is available [here](https://drive.google.com/file/d/13ZTZC48o2i0ZktobfukHjaTMh7Qbrfei/view?usp=sharing).
 

### 2018-10-23 src/help_extract_from_multiple_netCDFs.Rmd
ugh
formatting done

@@
 
-Built with 'r getRversion()'.
+Built with 
+```{r}
+getRversion()
+```
 
@@
 ```{r pre-install packages, suppressMessages(expr)}
-packages <- c("rgdal", "raster", "ncdf4", "RNetCDF", "sp", "parallel", "magrittr", "data.table") 
+packages <- c("ncdf4", "sp") 
 source("../src/ipak.R")
 ipak(packages)
-```
-
-# use multi-core (if it is an option)
-Detect cores automatically, and free one up.
-```{r}
-cores <- detectCores() - 1
 ```
@@
 I have around 1500 NetCDFs.
-Each netCDF has 4 dimensions (time - T, longitude - X, latitude - Y, depth - Z) and represents a single oceanographic variable for a particular month in a particular year.
+Each netCDF has 4 dimensions (`time - T`, `longitude - X`, `latitude - Y`, `depth - Z`) and represents a single oceanographic variable for a particular month in a particular year.
 I have given each netCDF the following naming convention:
 
-'yyyy_mm_variablename.nc' (where variable name can vary in length)
+`yyyy_mm_variablename.nc` (where variable name can vary in length)
 
-An example of the netCDF is available [here](https://drive.google.com/file/d/1y2XU5RzUBRLgTXkql667a5o_PVk5LMTP/view?usp=sharing) .
-To use the netCDFs in R, I have made a loop to create raster bricks and project them in an equal-area grid from all the netCDFs:
+An example of the `netCDF` is available [here](https://drive.google.com/file/d/1y2XU5RzUBRLgTXkql667a5o_PVk5LMTP/view?usp=sharing) .
+To use the netCDFs in R, I have made a loop to create `raster bricks` and project them in an equal-area grid:
 
@@
 
-I also had a dataset (.csv) containing fish observation (point) data. This dataset (let's call it occurrence) contains the following columns:
+I also had a dataset (.csv) containing fish observation (point) data. This dataset (let's call it `occurrence`) contains the following columns:
 
-occurrence$id (a unique observation ID)
-occurrence$decimalLongitude (Longitude in decimal degree)
-occurrence$decimalLatitude (Latitude in decimal degree)
-occurrence$dLongitudeM (Longitude in meters)
-occurrence$LatitudeM (Latitude in meters)
-occurrence$year (year of observation)
-occurrence$month (month of observation)
-occurrence$depthlayerno (depth of observation - the value corresponds to the depth layers in the raster bricks e.g.
-if occurrence$depthlayerno == 20, then the observation is from the raster brick layer [[20]])
+`occurrence$id` (a unique observation ID)
+`occurrence$decimalLongitude` (Longitude in decimal degree)
+`occurrence$decimalLatitude` (Latitude in decimal degree)
+`occurrence$dLongitudeM` (Longitude in meters)
+`occurrence$LatitudeM` (Latitude in meters)
+`occurrence$year` (year of observation)
+`occurrence$month` (month of observation)
+`occurrence$depthlayerno (depth of observation - the value corresponds to the depth layers in the raster bricks e.g.
+if `occurrence$depthlayerno` == `20`, then the observation is from the raster brick layer `[[20]]`)
 
@@
 
-occurrence$salinity_1 (extract value to point from the salinity raster brick layer [[1]])
-occurrence$salinity_depth (extract value to point from the salinity raster brick layer [[occurrence$depthlayerno]])
+`occurrence$salinity_1` (extract value to point from the salinity raster brick layer `[[1]]`)
+`occurrence$salinity_depth` (extract value to point from the salinity raster brick layer `[[occurrence$depthlayerno]]`)
 
@@
 
-Most of the raster bricks contain 50 layers - some only have 1 layer. Furthermore not all observations have a depthlayerno value (occurrence$depthlayerno == NA)
+Most of the raster bricks contain 50 layers - some only have 1 layer. Furthermore not all observations have a depthlayerno value (`occurrence$depthlayerno == NA`)
 
@@
 
-The dput for occurrence is as follows:
+The `dput` for occurrence is as follows:
 
@@
 
-And str(occurrence) is
+And `str(occurrence)` is
 
@@
 
-I have converted the occurrence dataset into a SpatialPointsDataFrame, projected it into an equal area grid (to match the raster bricks), and them split the data into year-months (I thought this might be an easier way to tackle the next stage):
+I have converted the occurrence dataset into a `SpatialPointsDataFrame`, projected it into an equal area grid (to match the raster bricks), and them split the data into year-months (I thought this might be an easier way to tackle the next stage):
 
@@
 
-What I would like to happen is for each  each observation (row) in occurrences, for each oceanographic variable (e.g.
-using salinity) look up occurrence$year and occurrence$month and locate the appropriate raster brick, and extract values to points from layer [[1]] and add the result to occurrence$salinity_1, then look up occurrence$depthlayerno and extract values to points from layer[[depthlayerno]] and add the result to occurrence$salinity_depth.
+What I would like to happen is for each  each observation (row) in `occurrences`, for each oceanographic variable (e.g.
+using salinity) look up `occurrence$year` and `occurrence$month` and locate the appropriate raster brick, and extract values to points from layer `[[1]]` and add the result to `occurrence$salinity_1`, then look up `occurrence$depthlayerno` and extract values to points from layer`[[depthlayerno]]` and add the result to `occurrence$salinity_depth`.
 
@@
 I found a rather [neat blog from Nathan Lane](http://nathanlane.info/tutorial/2016/01/02/gisfunctional.html) in which he uses functions to extract values to polygons from multiple raster bricks (but without selecting particular raster bricks to extract the data from).
-Unfortunately adapting it to my needs is way beyond my current skillset (which is a beginner).
+Unfortunately adapting it to my needs is way beyond my current skillset.
 

### 2018-10-23 src/help_extract_from_multiple_netCDFs.Rmd
html for slr

@@
 author: "Samantha Andrews"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
+  pdf_document: default
 ---
@@
 ```{r pre-install packages, suppressMessages(expr)}
-packages <- c("ncdf4", "sp") 
+packages <- c("ncdf4", "sp", "raster") 
 source("../src/ipak.R")
@@
 ```{r}
+aea <- raster("../output/env/aea.tif")
 netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
@@
 `occurrence$month` (month of observation)
-`occurrence$depthlayerno (depth of observation - the value corresponds to the depth layers in the raster bricks e.g.
-if `occurrence$depthlayerno` == `20`, then the observation is from the raster brick layer `[[20]]`)
+`occurrence$depthlayerno`  (depth of observation - the value corresponds to the depth layers in the raster bricks e.g.
+if `occurrence$depthlayerno`== `20`, then the observation is from the raster brick layer `[[20]]`)
 
@@
 ```{r}
-occurrence <- read.csv("./data/bio/occurrence_help.csv", header = TRUE)
+occurrence <- read.csv("../data/bio/occurrence_help.csv", header = TRUE)
 ```
@@
 library(sp)
-species_obs_date <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")) #convert into a spatialpoints dataframe to do the extraction
+xy <- occurrence[ , c("decimalLongitude", "decimalLatitude")]
+species_obs_date <- SpatialPointsDataFrame(coords = xy, data = occurrence, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")) #convert into a spatialpoints dataframe to do the extraction
 species_obs_date <- spTransform(species_obs_date, CRS = "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs") #need to reproject....

### 2018-10-23 src/netcdf_to_brick_projected.Rmd
html for slr

@@
 library(sp)
+xy <- data_aea[ , c("decimalLongitude", "decimalLatitude")]
 species_obs_date <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")) #convert into a spatialpoints dataframe to do the extraction

### 2018-10-25 src/biological_data_prep.Rmd
bottom depth extraction

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful")
+packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful", "marmap")
 source("../src/ipak.R")

### 2018-10-25 src/environmental_data_preperation.Rmd
feck all progress :(
bottom depth extraction

@@
 ```
+
+# Bathymetric
+
+bottom depth from [GEBCO](www.gebco.net)
+
+```{r}
+library(ncdf4)
+gebco <- nc_open("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
+gebco
+library(marmap)
+bat <-readGEBCO.bathy("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
+summary(bat)
+```
+
+create pretty custom color palette and then plot the bathy laye
+```{r}
+blues <- colorRampPalette(c("lightblue", "cadetblue2", "cadetblue1", "white"))
+plot(bat, n = 1, image = TRUE, bpal = blues(100))
+```
+lovely jubbely
+
+add the species data...
+```{r}
+data_aea <-read.csv( "../output/bio/data_aea_cell_amo_nao_bins_env.csv", header = TRUE) #just for canada pc as not loaded
+plot(bat, n = 1, image = TRUE, bpal = blues(100))
+points(data_aea$decimalLongitude, data_aea$decimalLatitude, pch = 20, col = "yellow") #note these are both the wgs84 images
+```
+and now get depth... note all the points and bathy data is wgs!
+
+```{r}
+data_aea$bottom_depth <- get.depth(bat, data_aea[ ,5:4], locator = FALSE)
+data_aea$bottom_depth <- data_aea$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth.
+This is extracting the depth column and deplacing data_aea$bottom_depth with the correct data.
+head(data_aea)
+```
+
+might be easier just to work form a TIF though...
+
+First project the tiff to equal area
+```{r}
+library(raster)
+geb <- "../data/env/gebco.tif"
+mask1 <- "../data/env/aea.tif" #just for canada pc as file isn't loaded
+aea <- raster(mask1)
+gebr <- brick(geb)
+gebr
+aea
+geb
+geb_aea <- projectRaster(geb, aea) 
+```
+
+```{r GEBCO raster extraction}
+library(rgdal)
+library(sp)
+gebco_coordinates(data_aea) <- c("decimalLongitude", "decimalLatitude") 
+nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
+proj4string(adddata) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
+adddata$nafo_zone <- over(adddata, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
+head(adddata) #just to check it worked
+```

### 2018-10-25 src/environmental_data_preperation.Rmd
GEBCO_AEA

@@
 library(ncdf4)
-gebco <- nc_open("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
-gebco
 library(marmap)
@@
 
-might be easier just to work form a TIF though...
+```{r}
+write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins_btmdepth.csv", row.names = FALSE)
+```
 
-First project the tiff to equal area
+note that you will still need to project the bathymetric layer to aea to do the mapping on...
+
+I THINK THIS WILL BE SOLVED ON MY REAL PC - SAM DEAL WITH THIS!
+
 ```{r}
 library(raster)
-geb <- "../data/env/gebco.tif"
 mask1 <- "../data/env/aea.tif" #just for canada pc as file isn't loaded
 aea <- raster(mask1)
-gebr <- brick(geb)
-gebr
-aea
-geb
-geb_aea <- projectRaster(geb, aea) 
+gebco <- raster("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
+gebco
+prj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
+gebco_aea <- projectRaster(gebco, prj)  
+gebco_aea
 ```
 
-```{r GEBCO raster extraction}
-library(rgdal)
-library(sp)
-gebco_coordinates(data_aea) <- c("decimalLongitude", "decimalLatitude") 
-nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
-proj4string(adddata) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
-adddata$nafo_zone <- over(adddata, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
-head(adddata) #just to check it worked
+```{r}
+plot(gebco_aea)
 ```

### 2018-10-25 src/biological_data_prep.Rmd
obs_by_cell

@@
 
+# observations by cell (in total, and also by time step)
+
+as the title says... how many observations happeend in each cell:
+- in total
+- in each yyyymm
+
+```{r observation by cell - total}
+library(plyr) #to use the count function
+obs_by_cell <- count(data_aea, "cell_id")
+obs_by_cell
+write.csv(obs_by_cell, file = "../output/bio/no_observations_cell.csv")
+```
+
+```{r}
+obs_cell_yymm <- count(data_aea, c("cell_id", "year", "month"))
+obs_cell_yymm
+write.csv(obs_cell_yymm, file = "../output/bio/no_observations_cell_yymm.csv")
+```
+
 # warm/cold year

### 2018-10-26 src/biological_data_prep.Rmd
depth data as per canada, plus time-sliced cell randomisation

@@
 ```
+Now attach these values to data_aea....
+
+total obs for the whole time period
+```{r}
+data_aea <- merge(data_aea, obs_by_cell, by="cell_id")
+head(data_aea)
+```
+rename freq column to something meaningful
+
+```{r}
+names(data_aea)[names(data_aea)=="freq"] <- "total_cell_obs"
+head(data_aea)
+```
+
+time-sliced
+```{r}
+data_aea <- merge(data_aea, obs_cell_yymm, by.x=c("cell_id", "year", "month"), by.y=c("cell_id", "year", "month"))
+names(data_aea)[names(data_aea)=="freq"] <- "yymm_cell_obs"
+head(data_aea)
+```
+
+write to .csv
+```{r}
+write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno.csv", row.names = FALSE)
+```
 

### 2018-10-26 src/environmental_data_preperation.Rmd
depth data as per canada, plus time-sliced cell randomisation

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("ncdf4", "raster", "rgdal", "useful") 
+packages <- c("ncdf4", "raster", "rgdal", "useful", "marmap") 
 source("../src/ipak.R")
@@
 
-note that you will still need to project the bathymetric layer to aea to do the mapping on...
-
-I THINK THIS WILL BE SOLVED ON MY REAL PC - SAM DEAL WITH THIS!
+And now create a raster of Gebco in aea for mapping...
 
@@
 library(raster)
-mask1 <- "../data/env/aea.tif" #just for canada pc as file isn't loaded
-aea <- raster(mask1)
 gebco <- raster("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
 gebco
-prj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
-gebco_aea <- projectRaster(gebco, prj)  
+gebco_aea <- projectRaster(gebco, aea)  
 gebco_aea

### 2018-10-26 src/help_extract_from_multiple_netCDFs.tex
depth data as per canada, plus time-sliced cell randomisation

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={help\_extract\_from\_multiple\_netCDFs},
            pdfauthor={Samantha Andrews},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{help\_extract\_from\_multiple\_netCDFs}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Samantha Andrews}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\begin{document}
\maketitle

A note to anyone who might happen to stumble across this\ldots{} I am a
beginner in R and have had no exposure to similar languages. I don't
know what I'm doing. The code herein is unlikely to be elegant and there
area probably more efficient ways of running the code.

Built with

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{getRversion}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] '3.5.1'
\end{verbatim}

\section{Package dependencies}\label{package-dependencies}

You can install and load them using the following code which uses a
function called
\href{https://gist.github.com/stevenworthington/3178163}{ipak}. Note
this function checks to see if the packages are installed first. If they
are not installed, it will do so before loading

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{packages <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ncdf4"}\NormalTok{, }\StringTok{"sp"}\NormalTok{, }\StringTok{"raster"}\NormalTok{) }
\KeywordTok{source}\NormalTok{(}\StringTok{"../src/ipak.R"}\NormalTok{)}
\KeywordTok{ipak}\NormalTok{(packages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: ncdf4
\end{verbatim}

\begin{verbatim}
## Loading required package: sp
\end{verbatim}

\begin{verbatim}
## Loading required package: raster
\end{verbatim}

\begin{verbatim}
##  ncdf4     sp raster 
##   TRUE   TRUE   TRUE
\end{verbatim}

\section{The short story}\label{the-short-story}

I have multiple netCDFs that need to extract values from to points. The
problem is, I need R to decide which netCDFs (and which layers) to
perform the extraction on.

\section{The details}\label{the-details}

I have around 1500 NetCDFs. Each netCDF has 4 dimensions
(\texttt{time\ -\ T}, \texttt{longitude\ -\ X}, \texttt{latitude\ -\ Y},
\texttt{depth\ -\ Z}) and represents a single oceanographic variable for
a particular month in a particular year. I have given each netCDF the
following naming convention:

\texttt{yyyy\_mm\_variablename.nc} (where variable name can vary in
length)

An example of the \texttt{netCDF} is available
\href{https://drive.google.com/file/d/1y2XU5RzUBRLgTXkql667a5o_PVk5LMTP/view?usp=sharing}{here}
. To use the netCDFs in R, I have made a loop to create
\texttt{raster\ bricks} and project them in an equal-area grid:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{aea <-}\StringTok{ }\KeywordTok{raster}\NormalTok{(}\StringTok{"../output/env/aea.tif"}\NormalTok{)}
\NormalTok{netcdf_list <-}\StringTok{ }\KeywordTok{list.files}\NormalTok{(}\StringTok{"../data/env/netcdf"}\NormalTok{, }\DataTypeTok{pattern =} \StringTok{'*.nc'}\NormalTok{, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{) }\CommentTok{#true means the full path is included}
\NormalTok{no_netcdf <-}\StringTok{ }\KeywordTok{length}\NormalTok{(netcdf_list) }\CommentTok{#for the loop - need to know how many files to cycle through}
\NormalTok{netcdf_name <-}\StringTok{ }\KeywordTok{list.files}\NormalTok{(}\StringTok{"../data/env/netcdf"}\NormalTok{, }\DataTypeTok{pattern =} \StringTok{'*.nc'}\NormalTok{, }\DataTypeTok{full.names =} \OtherTok{FALSE}\NormalTok{) }\CommentTok{#false means the path is not included}
\KeywordTok{library}\NormalTok{ (ncdf4)}
\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{no_netcdf) \{}
  \KeywordTok{print}\NormalTok{(netcdf_name[i]) }\CommentTok{#this just prings the name of the netCDF R is working on}
\NormalTok{  temp_brick <-}\StringTok{ }\KeywordTok{brick}\NormalTok{(netcdf_list[i], }\DataTypeTok{lvar =} \DecValTok{4}\NormalTok{)}
\NormalTok{  temp_brick <-}\StringTok{ }\KeywordTok{projectRaster}\NormalTok{(temp_brick, aea) }\CommentTok{#aea is an existing raster in the projection I want that I created (environmental_data_preperation)}
  \KeywordTok{assign}\NormalTok{(netcdf_name[i], temp_brick) }\CommentTok{#this asigns the netCDF name to the raster brick}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "1998_01_mlp.nc"
## [1] "1998_01_salinity.nc"
## [1] "1998_01_ssh.nc"
## [1] "1998_01_temp.nc"
## [1] "1998_02_mlp.nc"
## [1] "1998_02_salinity.nc"
## [1] "1998_02_ssh.nc"
## [1] "1998_02_temp.nc"
## [1] "1998_03_mlp.nc"
## [1] "1998_03_salinity.nc"
## [1] "1998_03_ssh.nc"
## [1] "1998_03_temp.nc"
## [1] "1998_04_mlp.nc"
## [1] "1998_04_salinity.nc"
## [1] "1998_04_ssh.nc"
## [1] "1998_04_temp.nc"
## [1] "1998_05_mlp.nc"
## [1] "1998_05_salinity.nc"
## [1] "1998_05_ssh.nc"
## [1] "1998_05_temp.nc"
## [1] "1998_06_mlp.nc"
## [1] "1998_06_salinity.nc"
## [1] "1998_06_ssh.nc"
## [1] "1998_06_temp.nc"
## [1] "1998_07_mlp.nc"
## [1] "1998_07_salinity.nc"
## [1] "1998_07_ssh.nc"
## [1] "1998_07_temp.nc"
## [1] "1998_08_mlp.nc"
## [1] "1998_08_salinity.nc"
## [1] "1998_08_ssh.nc"
## [1] "1998_08_temp.nc"
## [1] "1998_09_mlp.nc"
## [1] "1998_09_salinity.nc"
## [1] "1998_09_ssh.nc"
## [1] "1998_09_temp.nc"
## [1] "1998_10_mlp.nc"
## [1] "1998_10_salinity.nc"
## [1] "1998_10_ssh.nc"
## [1] "1998_10_temp.nc"
## [1] "1998_11_mlp.nc"
## [1] "1998_11_salinity.nc"
## [1] "1998_11_ssh.nc"
## [1] "1998_11_temp.nc"
## [1] "1998_12_mlp.nc"
## [1] "1998_12_salinity.nc"
## [1] "1998_12_ssh.nc"
## [1] "1998_12_temp.nc"
\end{verbatim}

Note that this may not be the most computationally efficient way to do
this as I will be creating a lot of raster bricks that will be stored to
memory\ldots{}

I also had a dataset (.csv) containing fish observation (point) data.
This dataset (let's call it \texttt{occurrence}) contains the following
columns:

\texttt{occurrence\$id} (a unique observation ID)
\texttt{occurrence\$decimalLongitude} (Longitude in decimal degree)
\texttt{occurrence\$decimalLatitude} (Latitude in decimal degree)
\texttt{occurrence\$dLongitudeM} (Longitude in meters)
\texttt{occurrence\$LatitudeM} (Latitude in meters)
\texttt{occurrence\$year} (year of observation)
\texttt{occurrence\$month} (month of observation)
\texttt{occurrence\$depthlayerno\ (depth\ of\ observation\ -\ the\ value\ corresponds\ to\ the\ depth\ layers\ in\ the\ raster\ bricks\ e.g.\ if}occurrence\$depthlayerno\texttt{==}20\texttt{,\ then\ the\ observation\ is\ from\ the\ raster\ brick\ layer}{[}{[}20{]}{]}`)

I then have some empty columns that need to be populated with data from
the raster bricks, each representing an oceanographic variable and depth
layer the data needs to come from. An example is:

\texttt{occurrence\$salinity\_1} (extract value to point from the
salinity raster brick layer \texttt{{[}{[}1{]}{]}})
\texttt{occurrence\$salinity\_depth} (extract value to point from the
salinity raster brick layer
\texttt{{[}{[}occurrence\$depthlayerno{]}{]}})

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{occurrence <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"../data/bio/occurrence_help.csv"}\NormalTok{, }\DataTypeTok{header =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Most of the raster bricks contain 50 layers - some only have 1 layer.
Furthermore not all observations have a depthlayerno value
(\texttt{occurrence\$depthlayerno\ ==\ NA})

A cut-down version of this .csv is available
\href{https://drive.google.com/file/d/13ZTZC48o2i0ZktobfukHjaTMh7Qbrfei/view?usp=sharing}{here}.

The \texttt{dput} for occurrence is as follows:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{dput}\NormalTok{(occurrence)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## structure(list(year = c(1998L, 1998L, 1998L, 1998L, 1998L, 1998L, 
## 1998L, 1998L, 1998L, 1998L, 1998L, 1998L, 1998L, 1998L, 1998L
## ), month = c(10L, 10L, 10L, 10L, 11L, 11L, 12L, 12L, 12L, 4L, 
## 4L, 4L, 4L, 4L, 5L), id = c(386669852L, 386669881L, 386669904L, 
## 386411062L, 386674153L, 386415830L, 386417305L, 386417335L, 386417588L, 
## 386403612L, 386402081L, 386402062L, 386402787L, 386401885L, 386405989L
## ), decimalLatitude = c(53.95, 53.63, 53.72, 45.19, 50.26, 48.42, 
## 48.96, 48.95, 47.7, 46.87, 46.38, 46.46, 47.4, 47.12, 45.6), 
##     decimalLongitude = c(-54.91, -55.51, -55.73, -52.85, -55.49, 
##     -49.52, -52.78, -52.81, -52.93, -57.86, -54.94, -54.82, -56.37, 
##     -54.61, -51.68), LongitudeM = c(331170.3733, 294592.1288, 
##     279536.5639, 568120.6623, 321164.228, 775577.4971, 529039.8088, 
##     526968.5028, 532747.635, 164445.6561, 392590.9816, 401204.4947, 
##     275695.0622, 411603.073, 655083.2443), LatitudeM = c(1549820.727, 
##     1511221.289, 1520353.142, 594666.88, 1135604.497, 981130.7983, 
##     1008824.847, 1007486.149, 868898.3383, 751816.2887, 709993.6159, 
##     719487.3038, 815139.5931, 793297.9028, 650032.9389), depthlayerno = c(30L, 
##     28L, 26L, 24L, 35L, 35L, 33L, 32L, 25L, 33L, 31L, 32L, 35L, 
##     31L, 23L), salinty_1 = c(NA, NA, NA, NA, NA, NA, NA, NA, 
##     NA, NA, NA, NA, NA, NA, NA), salinity_depth = c(NA, NA, NA, 
##     NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA), temp_1 = c(NA, 
##     NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA), 
##     temp_depth = c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 
##     NA, NA, NA, NA)), class = "data.frame", row.names = c(NA, 
## -15L))
\end{verbatim}

And \texttt{str(occurrence)} is

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{str}\NormalTok{(occurrence)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 'data.frame':    15 obs. of  12 variables:
##  $ year            : int  1998 1998 1998 1998 1998 1998 1998 1998 1998 1998 ...
##  $ month           : int  10 10 10 10 11 11 12 12 12 4 ...
##  $ id              : int  386669852 386669881 386669904 386411062 386674153 386415830 386417305 386417335 386417588 386403612 ...
##  $ decimalLatitude : num  54 53.6 53.7 45.2 50.3 ...
##  $ decimalLongitude: num  -54.9 -55.5 -55.7 -52.9 -55.5 ...
##  $ LongitudeM      : num  331170 294592 279537 568121 321164 ...
##  $ LatitudeM       : num  1549821 1511221 1520353 594667 1135604 ...
##  $ depthlayerno    : int  30 28 26 24 35 35 33 32 25 33 ...
##  $ salinty_1       : logi  NA NA NA NA NA NA ...
##  $ salinity_depth  : logi  NA NA NA NA NA NA ...
##  $ temp_1          : logi  NA NA NA NA NA NA ...
##  $ temp_depth      : logi  NA NA NA NA NA NA ...
\end{verbatim}

I have converted the occurrence dataset into a
\texttt{SpatialPointsDataFrame}, projected it into an equal area grid
(to match the raster bricks), and them split the data into year-months
(I thought this might be an easier way to tackle the next stage):

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(sp)}
\NormalTok{xy <-}\StringTok{ }\NormalTok{occurrence[ , }\KeywordTok{c}\NormalTok{(}\StringTok{"decimalLongitude"}\NormalTok{, }\StringTok{"decimalLatitude"}\NormalTok{)]}
\NormalTok{species_obs_date <-}\StringTok{ }\KeywordTok{SpatialPointsDataFrame}\NormalTok{(}\DataTypeTok{coords =}\NormalTok{ xy, }\DataTypeTok{data =}\NormalTok{ occurrence, }\DataTypeTok{proj4string =} \KeywordTok{CRS}\NormalTok{(}\StringTok{"+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"}\NormalTok{)) }\CommentTok{#convert into a spatialpoints dataframe to do the extraction}
\NormalTok{species_obs_date <-}\StringTok{ }\KeywordTok{spTransform}\NormalTok{(species_obs_date, }\DataTypeTok{CRS =} \StringTok{"+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"}\NormalTok{) }\CommentTok{#need to reproject.... note -60 used to be -91. Changed to 'straighten up'}
\NormalTok{split_obs <-}\StringTok{ }\KeywordTok{split}\NormalTok{(species_obs_date, }\KeywordTok{paste}\NormalTok{(species_obs_date}\OperatorTok{$}\NormalTok{year, species_obs_date}\OperatorTok{$}\NormalTok{month))}
\end{Highlighting}
\end{Shaded}

\section{The problem}\label{the-problem}

What I would like to happen is for each each observation (row) in
\texttt{occurrences}, for each oceanographic variable (e.g.~using
salinity) look up \texttt{occurrence\$year} and
\texttt{occurrence\$month} and locate the appropriate raster brick, and
extract values to points from layer \texttt{{[}{[}1{]}{]}} and add the
result to \texttt{occurrence\$salinity\_1}, then look up
\texttt{occurrence\$depthlayerno} and extract values to points from
layer\texttt{{[}{[}depthlayerno{]}{]}} and add the result to
\texttt{occurrence\$salinity\_depth}.

Honestly, I have no idea where to begin. I found a rather
\href{http://nathanlane.info/tutorial/2016/01/02/gisfunctional.html}{neat
blog from Nathan Lane} in which he uses functions to extract values to
polygons from multiple raster bricks (but without selecting particular
raster bricks to extract the data from). Unfortunately adapting it to my
needs is way beyond my current skillset.

Any help you can offer would be gratefully received!

\end{document}

### 2018-10-26 src/timesliced_random_cell_selection.Rmd
depth data as per canada, plus time-sliced cell randomisation

---
title: "timesliced_random_cell_selection"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Preperation of the capelin (*Mallotus villosus*) data obtained from [OBIS](www.iobis.org). This database shows the global occurance of capelin.

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("raster")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

you have a unique_cell raster already (glo_unique_cell) (output/env/glo_unique_cell.tif)

```{r}
plot(glo_unique_cell)
```

as a reminder, how many cells are there?
```{r}
glo_unique_cell
```

6729...

ok now you want to generate a centroid for each cell... (a point in the center of each cell)
use the raster layer to 1) transfer the raster into a dataframe, with the values (cell numbers) linked to each centroid with xy coordinates (note as the raster is an aea projection, the xy will be in meters)
```{r}
library(raster)
unique_cell_centroid <- as.data.frame(glo_unique_cell, row.names = NULL, optional = FALSE, xy = TRUE, centroids = TRUE, na.rm = TRUE)
names(unique_cell_centroid)[names(unique_cell_centroid)=="layer.1"] <- "cell_id" #just renames the output
names(unique_cell_centroid)[names(unique_cell_centroid)=="x"] <- "longitude" #just renames the output
names(unique_cell_centroid)[names(unique_cell_centroid)=="y"] <- "latitude" #just renames the output
write.csv(unique_cell_centroid, "../output/env/unique_cell_centroid.csv")
plot(glo_unique_cell)
points(unique_cell_centroid$longitude, unique_cell_centroid$latitude, pch = 10, col = "black")
```
well... the plot doesn't tell me much... but i definately have points!

ok so now I have a base list to work off... I have split the cells with presence data into a list by time-slice (split_obs_cell_yymm) (code in biological_data_prep)

```{r}
test_timeslice <- split_obs_cell_yymm[1]
test_timeslice <- as.data.frame(test_timeslice)
test_time_cell <- unique_cell_centroid
test_time_cell <- test_time_cell[!test_time_cell$cell_id %in% test_timeslice$X1998.10.cell_id, , drop = FALSE]
```
ok that works... but you need to pop into a loop and dynamically name the output (and create a .csv maybe)

```{r}

```

### 2018-10-29 src/timesliced_random_cell_selection.Rmd
progress with loop to find all unoccupied cells in each timeslice (missing correct .csv naming)

@@
 
-ok so now I have a base list to work off... I have split the cells with presence data into a list by time-slice (split_obs_cell_yymm) (code in biological_data_prep)
+ok so now I have a base list to work off...
+
+Split the data_aea into timeslices (year/month).
+All you need is the year, month, cell_id
+```{r}
+cell_yymm <- subset(data_aea, select= c(cell_id, year, month)) #create a new df with just cell_id, yr, month
+split_cell_yymm <- split(cell_yymm, list(cell_yymm$year, cell_yymm$month), drop = TRUE)
+```
+
+What I want to do is for each time slice, create a list of all cells, then remove the ones where observations occured
 
 ```{r}
-test_timeslice <- split_obs_cell_yymm[1]
+test_timeslice <- split_cell_yymm[[1]]
+test_timeslice
 test_timeslice <- as.data.frame(test_timeslice)
+
+timeslice_name <- name(split_cell_yymm(1))
 test_time_cell <- unique_cell_centroid
-test_time_cell <- test_time_cell[!test_time_cell$cell_id %in% test_timeslice$X1998.10.cell_id, , drop = FALSE]
+test_time_cell <- test_time_cell[!test_time_cell$cell_id %in% test_timeslice$cell_id, , drop = FALSE]
+write.csv(test_time_cell, "../output/bio/test_time_cell.csv", row.names = FALSE)
 ```
@@
 
+need a loop to create a copy of the unique_cell_centroid dataframe, and give it the same name as the splitobs_cell_yymm level
+
+this almost works - just need it to assign a name not a list number. note the output goes to /src....
 ```{r}
+no_timeslices <- length(split_cell_yymm) # how many levels (time slices) are in the list
+for (i in 1:no_timeslices){
+  test_timeslice <- split_cell_yymm[[i]]
+  test_timeslice <- as.data.frame(test_timeslice)
+  test_time_cell <- unique_cell_centroid
+  test_time_cell <- test_time_cell[!test_time_cell$cell_id %in% test_timeslice$cell_id, , drop = FALSE]
+  write.csv(test_time_cell, paste0(i, ".csv"), row.names = FALSE)
+}
 

### 2018-10-29 src/timesliced_random_cell_selection.Rmd
loop complete, started cell selection randomisation

--- 2018-10-26 src/timesliced_random_cell_selection.Rmd
@@
 
-ok so now I have a base list to work off... I have split the cells with presence data into a list by time-slice (split_obs_cell_yymm) (code in biological_data_prep)
+ok so now I have a base list to work off...
+
+Split the data_aea into timeslices (year/month).
+All you need is the year, month, cell_id
+```{r}
+cell_yymm <- subset(data_aea, select= c(cell_id, year, month)) #create a new df with just cell_id, yr, month
+split_cell_yymm <- split(cell_yymm, list(cell_yymm$year, cell_yymm$month), drop = TRUE)
+```
+
+What I want to do is for each time slice, create a list of all cells, then remove the ones where observations occured
+ly name the output (and create a .csv with the name of the level being worked on)
+
+Here is a loop to create a copy of the unique_cell_centroid dataframe, and give it the same name as the splitobs_cell_yymm level
+.csv naming code courtesy of [nadizan/stack exchange](https://stackoverflow.com/questions/53045158/r-naming-csv-with-a-list-level-name-in-a-loop?noredirect=1#comment92990285_53045158)
+
+Note that the file are saved in /src. Manually move them to output/bio/background
 
 ```{r}
-test_timeslice <- split_obs_cell_yymm[1]
-test_timeslice <- as.data.frame(test_timeslice)
-test_time_cell <- unique_cell_centroid
-test_time_cell <- test_time_cell[!test_time_cell$cell_id %in% test_timeslice$X1998.10.cell_id, , drop = FALSE]
+no_timeslices <- length(split_cell_yymm) # how many levels (time slices) are in the list
+for (i in 1:no_timeslices){
+  timeslice <- split_cell_yymm[[i]]
+  timeslice <- as.data.frame(test_timeslice)
+  time_cell <- unique_cell_centroid
+  time_cell <- test_time_cell[!test_time_cell$cell_id %in% test_timeslice$cell_id, , drop = FALSE]
+  write.csv(test_time_cell, paste0(names(lapply(split_cell_yymm, names))[i], ".csv"), row.names = FALSE) 
+}
 ```
-ok that works... but you need to pop into a loop and dynamically name the output (and create a .csv maybe)
+
+Ok so now I have a bunch of .csv files containing cells where no observations occurred for a given timeslice.
+Next step is to randomly select cells in each timeslice. Make sure you generate new files for each...
+
+load the .csvs... [code](https://www.reed.edu/data-at-reed/software/R/reading_and_writing.html)
 
 ```{r}
+bckrnd_folder <- "../output/bio/background/"      # path to folder that holds multiple .csv files
+bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
+
+# read in each .csv file in file_list and create a data frame with the same name as the .csv file
+for (i in 1:length(bckrnd_list)){
+  assign(bckrnd_list[i], 
+  read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
+  )
+  #now do something....
+
+  }
 
 ```
+and that something you are going to do is randomly select rows....

### 2018-10-30 src/timesliced_random_cell_selection.Rmd
added depth to the timesliced cell randomisation process

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("raster")
+packages <- c("raster", "dplyr")
 source("../src/ipak.R")
@@
 
-ok so now I have a base list to work off...
+ok so now I have a base list to work off... This gives you a dataset of xy points, but you need xyz (where z is based on the depth layers in the environmental data)
+```{r}
+unique_cell_xyz <- unique_cell_centroid
+unique_cell_xyz$depthlayerno <- NA
+unique_cell_xyz <- unique_cell_xyz[rep(seq_len(nrow(unique_cell_xyz)), each=50),] #this generates 50 repetitions of each row
+unique_cell_xyz$depthlayerno <- rep(1:50) #assign 1:50 on a repeating cycle to the depth column. 50 is chosen because there are 50 depth layers
+unique_cell_xyz$id_depth <- paste(unique_cell_xyz$cell_id, unique_cell_xyz$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID.
+This is to help remove cells later and randomise
+```
 
 Split the data_aea into timeslices (year/month).
-All you need is the year, month, cell_id
+All you need is the year, month, cell_id, and depthlayerno
 ```{r}
-cell_yymm <- subset(data_aea, select= c(cell_id, year, month)) #create a new df with just cell_id, yr, month
+cell_yymm <- subset(data_aea, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month
+cell_yymm$id_depth <- paste(cell_yymm$cell_id, cell_yymm$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID. This is to help remove cells later and randomise
 split_cell_yymm <- split(cell_yymm, list(cell_yymm$year, cell_yymm$month), drop = TRUE)
@@
 
-What I want to do is for each time slice, create a list of all cells, then remove the ones where observations occured
-ly name the output (and create a .csv with the name of the level being worked on)
+What I want to do is for each time slice, create a list of all cells (xyz), then remove the ones where observations occuredand create a .csv with the name of the level being worked on
 
@@
   timeslice <- split_cell_yymm[[i]]
-  timeslice <- as.data.frame(test_timeslice)
-  time_cell <- unique_cell_centroid
-  time_cell <- test_time_cell[!test_time_cell$cell_id %in% test_timeslice$cell_id, , drop = FALSE]
-  write.csv(test_time_cell, paste0(names(lapply(split_cell_yymm, names))[i], ".csv"), row.names = FALSE) 
+  timeslice <- as.data.frame(timeslice)
+  time_cell <- unique_cell_xyz
+  time_cell <- time_cell[!time_cell$id_depth %in% timeslice$id_depth, , drop = FALSE]
+  write.csv(time_cell, paste0(names(lapply(split_cell_yymm, names))[i], ".csv"), row.names = FALSE) 
 }
@@
 
-Ok so now I have a bunch of .csv files containing cells where no observations occurred for a given timeslice.
+Ok so now I have a bunch of .csv files containing cells where no observations occurred for a given timeslice on xyz coordinates.
 Next step is to randomly select cells in each timeslice. Make sure you generate new files for each...
@@
 and that something you are going to do is randomly select rows....
+
+manual attempt:
+```{r}
+library(dplyr)
+test_cell_selection <- read.csv("../output/bio/background/1998.10.csv", header = TRUE) #just load one csv to play with
+test_ran_cell_sel <- sample_n(test_cell_selection, 10) #where 10 = number of rows to sample
+```

### 2018-10-30 src/timesliced_random_cell_selection.Rmd
all of the random! timesliced!

@@
 
-Note that the file are saved in /src. Manually move them to output/bio/background
-
-```{r}
+```{r timesliced unique unsampled}
 no_timeslices <- length(split_cell_yymm) # how many levels (time slices) are in the list
+bckoutput <- "../output/bio/background/raw/" #where the files are to be saved to
 for (i in 1:no_timeslices){
@@
   time_cell <- time_cell[!time_cell$id_depth %in% timeslice$id_depth, , drop = FALSE]
-  write.csv(time_cell, paste0(names(lapply(split_cell_yymm, names))[i], ".csv"), row.names = FALSE) 
+  write.csv(time_cell, paste0(bckoutput, names(lapply(split_cell_yymm, names))[i], ".csv"), row.names = FALSE) 
 }
@@
 ```{r}
-bckrnd_folder <- "../output/bio/background/"      # path to folder that holds multiple .csv files
+library(dplyr)
+bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
 bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
-
+randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
 # read in each .csv file in file_list and create a data frame with the same name as the .csv file
 for (i in 1:length(bckrnd_list)){
-  assign(bckrnd_list[i], 
-  read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
-  )
-  #now do something....
-
+  test_sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
+  test_sel <- sample_n(test_sel, 10000) #where 10000 = number of rows to sample (large sample as per maxent)
+  write.csv(test_sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
 }
@@
 ```
-and that something you are going to do is randomly select rows....
 
-manual attempt:
+ok now i have a ton of files for random background points broken into timeslices.
+I'd like to add a colum for year and one for month to each of these datasets.
+They should be populated based on the name of the file (e.g.
+1998.4.csv would have year = 1998 and month = 4)
+
 ```{r}
-library(dplyr)
-test_cell_selection <- read.csv("../output/bio/background/1998.10.csv", header = TRUE) #just load one csv to play with
-test_ran_cell_sel <- sample_n(test_cell_selection, 10) #where 10 = number of rows to sample
+
 ```

### 2018-10-30 src/timesliced_random_cell_selection.Rmd
random timesliced files given yymm columns

@@
 for (i in 1:length(bckrnd_list)){
-  test_sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
-  test_sel <- sample_n(test_sel, 10000) #where 10000 = number of rows to sample (large sample as per maxent)
-  write.csv(test_sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
+  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
+  sel <- sample_n(test_sel, 10000) #where 10000 = number of rows to sample (large sample as per maxent)
+  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
 }
@@
 ```{r}
+rand_folder <- "../output/bio/background/rand/"      # path to folder that holds multiple .csv files
+rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
+randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
+for(i in 1:length(rand_list)){
+  randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
+  randtemp$year <- substr(rand_list[i], 1, 4)
+  randtemp$month <- substr(rand_list[i], 6, 7)
+  write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
+}
+```
+note that in the output temp file in R shows a . after the month if that month is a single digit. The .csv outputs don't
 
-```
+ok now you can start to populate the rest of the background point information. The background points file should mirror as much as possible the occurence file

### 2018-10-31 src/raster_sjl.Rmd
raster walkthrough for sjl help with code

# load the netcdf and raster packages
```{r}
library(ncdf4)
library(raster)
```

The data are held in netCDF files. We don't work directly with the netCDFs in R. Instad we load a netcdf file and convert it to a raster brick so we can work with it. In this example the netCDF I'll use is called '1993_01_salinity.nc'.I've sent it to you in the email with this code.

```{r}
nc <-  "./data/env/netcdf/1993_01_salinity.nc" # change the file path to where ever your netcdfs are saved
rasbr <- brick(nc,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
rasbr # this will give you some information about the raster brick
```
We no longer care that the data originated from a netCDF file. Everything now is done with the raster brick.

We can plot the different layers inside the raster brick.

Plot the top layer
```{r}
plot(rasbr[[1]]) #the double brackets tell R which layer (which corresponds to my depth) to plot
```

plot layer 20
```{r}
plot(rasbr[[20]]) #the double brackets tell R which layer (which corresponds to my depth) to plot
```

If you look back to the output of 'rasbr' and look at 'coord. ref.', you can see that the CRS (coordinate reference system) is in WGS84 (+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0)

Because I am using these files for Maxent, I need the projection to be an albers equal area. We can project the whole raster brick. There are a couple of ways to do this, but the simplest way is to use an existing file in the projection I want. The file I will use for this is called aea.tif. I've sent it to you in the email with this code as a zip. Note that there are two files in the zip folder - you will need both (but you only call on the .tif in the code) 
```{r}
aea <- raster("../output/env/aea.tif") # change the file path to where ever the tif and .twf files are saved
rasbr_aea <- projectRaster(rasbr, aea) 
plot(rasbr_aea[[1]]) # just to look at the top layer of the raster brick after projection so you can see the difference
```

The next piece of data we need is the observation data itself. I've sent you the observations.csv in the email with this code.
```{r}
obs <- read.csv("../data/bio/observations.csv") #change this path to your .csv location
head(obs)
```

The lon and lat represent meters as oppose to decimal degrees so it is 'albers equal area' ready.
For this exercise, I'm going to ignore year, month, and depth and just do a straightforward extract to points.

First we need to convert 'obs' into a 'spatialpoints dataframe'
```{r}
xy <- obs[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
obssp <- SpatialPointsDataFrame(coords = xy, data = obs, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
obssp
```

We can plot the observations on any layer of the 'rasbr_aea ' brick
```{r}
plot(rasbr_aea[[1]]) #plot layer 1 - the top layer
points(obssp$lon, obssp$lat)
```

The code to extract the valuees from any layer in 'rasbr_aea' is very simple
```{r}
obssp$salinity <- extract(x=rasbr_aea[[1]], y = obssp) #extract from the top layer of the rasbr_aea and add to the obssp dataset in a new column called salinity
head(obssp) # just to take a look at the output
```

I can write this 'spatialpoints dataframe' to a .csv and convert it back to a 'normal dataframe'.
```{r}
write.csv(obssp, file = "../output/bio/observations_sal.csv", row.names = FALSE) # change the path to where ever you would like to save it
obs <- as.data.frame(obssp) # this coverts the 'spatialpoints dataframe' to a 'normal dataframe'
obs # just to take a look at the output
```

for some reason converting the 'spatialpoints dataframe' to a 'normal dataframe' adds a duplicate lon and lat column (the write.csv function does the same). I don't know why, but they are easy to remove so I am not worried about it.

### 2018-10-31 src/raster_sjl.Rmd
finding lon lat wgs84

@@
 
-We can plot the different layers inside the raster brick.
+We can plot the different layers (which represent depth) inside the raster brick.
 
@@
 ```{r}
-plot(rasbr[[1]]) #the double brackets tell R which layer (which corresponds to my depth) to plot
+plot(rasbr[[1]]) #the double brackets tell R which layer to plot
 ```
@@
 ```{r}
-plot(rasbr[[20]]) #the double brackets tell R which layer (which corresponds to my depth) to plot
+plot(rasbr[[20]]) 
 ```
@@
 ```
+A note about obs$depth.
+This is not a measurement, but relates to the layer in the raster bricks.
+E.G.
+the observation in row one was found at rasbr[[15]], the observation in row 3 at rasbr[[17]].
+I created this field myself.
 

### 2018-10-31 src/timesliced_random_cell_selection.Rmd
finding lon lat wgs84

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("raster", "dplyr")
+packages <- c("raster", "dplyr", "data.table", "sp")
 source("../src/ipak.R")
@@
 
+# the show
 you have a unique_cell raster already (glo_unique_cell) (output/env/glo_unique_cell.tif)
@@
   randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
-  randtemp$year <- substr(rand_list[i], 1, 4)
-  randtemp$month <- substr(rand_list[i], 6, 7)
+  randtemp$year <- substr(rand_list[i], 1, 4) #the year is characters 1-4 in the filename
+  randtemp$month <- substr(rand_list[i], 6, 7) #the month is characters 6 & 7 in the filename (except where the month is a single digit)
   write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
@@
 
-ok now you can start to populate the rest of the background point information. The background points file should mirror as much as possible the occurence file
+# populating the background points data with... data
+
+ok now you can start to populate the rest of the background point information.
+The background points file should mirror as much as possible the occurence file.
+Maybe it would be best to merge the files together, and them split them again later?
+would save looping everything to repeat the process through multiple files...
+
+```{r}
+library(data.table)
+filenames  <- list.files(path = rand_folder, pattern="*.csv", full.names = TRUE) 
+background_all <- rbindlist(lapply(filenames,fread))
+write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
+```
+
+get the columns from the occurence data so you know what to add
+```{r}
+occ_column <- colnames(data_aea)
+occ_column
+```
+
+ok so have cell_id, year, month, decimalLongitude.1(meters - in background_all its longitude), decimalLatitude.1 (meters - in background_all its latitude), depth_layer_no.
+
+Easy first one - add an id based on id_depth
+```{r}
+background_all$id <- background_all$id_depth
+write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
+```
+
+ok lets's get the real decimalLatitude and decimalLongitude.
+Will need to plot all point, extract the lonlat coordinates
+
+1) load the file as a spatialdatapointframe. give it the aea coordinate
+
+```{r}
+library(sp)
+xy <- background_all[ ,c(1,2)] #lon lat in col 1 and 2
+sp_background <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
+head(sp_background)
+```
+```{r}
+plot(sp_background)
+```
+
+2) reproject to a wgs
+
+```{r}
+sp_background <- spTransform(sp_background, CRS = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
+plot(sp_background)
+```
+3) extract lonlat, add to dataframe
+
+```{r}
+head(sp_background)
+```
+
+save as csv

### 2018-11-02 src/raster_sjl.Rmd
working towards depth layer selection....

@@
 ```{r}
-nc <-  "./data/env/netcdf/1993_01_salinity.nc" # change the file path to where ever your netcdfs are saved
+nc <-  "../data/env/netcdf/1993_01_salinity.nc" # change the file path to where ever your netcdfs are saved
 rasbr <- brick(nc,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
@@
 ```
+
 We no longer care that the data originated from a netCDF file. Everything now is done with the raster brick.
@@
 ```{r}
-obs <- read.csv("../data/bio/observations.csv") #change this path to your .csv location
+obs <- read.csv("../data/observations3.csv") #change this path to your .csv location
 head(obs)
@@
 I don't know why, but they are easy to remove so I am not worried about it.
+
+#shawn code
+
+SL I AM DOING SOME STUFF HERE :O) 
+THE ABOVE EXTRACTION IS THE HOOK UP. I WANT TO EXTRACT FOR A CERTAIN DEPTH DEFINED BY OBSSP IN ORDER TO AUTOMATE THIS. THIS MEANS YOU NEED
+TO MAKE [[]] LINKED TO DEPTH...LET'S SEE IF WE CAN DO THIS NOW.
+```{r}
+obssp$salinity2 <- extract(x=rasbr_aea[[obssp$depth]], y = obssp)
+obssp$salinity2 # just to take a look at the output
+```
+
+Ok, the above is close but not right. It is printing salinity for depths 15, 22, 17, 18, 15, 16, 16 for each observation record.
+At least this is what I think. Let's check it out.
+```{r}
+obssp$salinity22 <- extract(x=rasbr_aea[[22]], y = obssp) #I expect this to be 33.42. Yep.
+obssp$salinity22
+```
+
+Starting over now I have created a smaller observation file with only 3 records. 
+Gonna create a loop here and iterate over the depths listed in the observation file
+```{r}
+Data_Loop <-NULL #Gonna put my loop data in here
+for (i in obssp$depth) { #cycle through values in depth column (15, 22, 17)
+  #this is the equation you will be calculating for every value of y
+  sal <- extract(x=rasbr_aea[[i]], y = obssp) #create a data list
+
+  #Put your data in a data frame. Don't forget to enclose this within the loop!
+  Data_Loop <- rbind(Data_Loop, data.frame(depth=i, year=obssp$year, month=obssp$month, lon=obssp$lon, lat=obssp$lat, Salinity=sal))
+
+  #print the i value so you can track progress. Good for longer loops
+  print(i)
+  #Close the loop
+  }
+Data_Loop
+```
+
+SL: TRY IT ANOTHER WAY. Problem is that we need to tell x to be at specific depth as specified by obssp$depth and y to be only the lat and lon at that depth.
+The two examples here (one above and one below) are VERY close. Just a matter of tweaking and playing around a bit more. See if you can figure it out.
+```{r}
+Data_Loop <-NULL #Gonna put my loop data in here
+for (i in 1:length(obssp)) { #cycle through values in depth column (15, 22, 17) #NO IT CYCLES THROUGH THE ROWS
+  #this is the equation you will be calculating for every value of y
+  sal <- extract(x=rasbr_aea[[obssp$depth[i]]], y = obssp[i,]) #go with i row in obssp
+
+  #Put your data in a data frame. Don't forget to enclose this within the loop!
+  Data_Loop <- rbind(Data_Loop, data.frame(depth=obssp$depth, year=obssp$year, month=obssp$month, lon=obssp$lon, lat=obssp$lat, Salinity=sal))
+
+  #print the i value so you can track progress. Good for longer loops
+  print(i)
+  #Close the loop
+}
+Data_Loop
+```
+
+# sam tests on shawns loop
+
+let's get the values manually
+```{r}
+test15 <- extract(x=rasbr_aea[[15]], y = obssp)
+test22 <- extract(x=rasbr_aea[[22]], y = obssp)
+test17 <- extract(x=rasbr_aea[[17]], y = obssp)
+test17
+```
+
+this loop replaces the values in obssp$samtest with the last depth layer... ie. the process does not work for each row in obssp in turn
+```{r}
+Data_Loop <-NULL #Gonna put my loop data in here
+for (i in 1:length(obssp)) {  # CYCLES THROUGH THE ROWS
+  de <- obssp$depth[[i]]  #have a variable for the depth layer
+  print(de)
+  obssp$samtest <- extract(x=rasbr_aea[[de]], y = obssp) #this is the equation you will be calculating for every value of y
+
+  #Put your data in a data frame. Don't forget to enclose this within the loop!
+  #Data_Loop <- rbind(Data_Loop, data.frame(depth=obssp$depth, year=obssp$year, month=obssp$month, lon=obssp$lon, lat=obssp$lat, Salinity=sal))
+
+  #print the i value so you can track progress. Good for longer loops
+  #print(i)
+  #Close the loop
+}
+#Data_Loop
+obssp$samtest
+```
+
+take 2
+```{r}
+Data_Loop <-NULL #Gonna put my loop data in here
+for (i in 1:length(obssp)) {  # CYCLES THROUGH THE ROWS
+  de <- obssp$depth[[i]]  #have a variable for the depth layer
+  print(de)
+  obssp$samtest <- extract(x=rasbr_aea[[de]], y = obssp) #this is the equation you will be calculating for every value of y
+
+  #Put your data in a data frame. Don't forget to enclose this within the loop!
+  #Data_Loop <- rbind(Data_Loop, data.frame(depth=obssp$depth, year=obssp$year, month=obssp$month, lon=obssp$lon, lat=obssp$lat, Salinity=sal))
+
+  #print the i value so you can track progress. Good for longer loops
+  #print(i)
+  #Close the loop
+}
+#Data_Loop
+obssp$samtest
+```

### 2018-11-02 src/dataextraction_SL.R

@@ -1,0 +1,107 @@
+# load the netcdf and raster packages
+library(ncdf4)
+library(raster)
+
+
+# The data are held in netCDF files. We don't work directly with the netCDFs in R. Instad we load a netcdf file and convert it to a raster brick so we can work with it. In this example the netCDF I'll # # use is called '1993_01_salinity.nc'.I've sent it to you in the email with this code.
+nc <-  "1993_01_salinity.nc" # change the file path to where ever your netcdfs are saved
+rasbr <- brick(nc,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+rasbr # this will give you some information about the raster brick
+
+# We no longer care that the data originated from a netCDF file. Everything now is done with the raster brick.
+
+# We can plot the different layers (which represent depth) inside the raster brick.
+
+#SL: gonna check some stuff on nc
+print(nc)
+nc2 <- nc_open(nc)
+attributes(nc2$var)
+data <- ncvar_get(nc2, attributes(nc2$var)$names[1]) #get salinity data
+dim(data)
+
+# Plot the top layer
+plot(rasbr[[1]]) #the double brackets tell R which layer to plot
+
+# plot layer 20
+plot(rasbr[[20]])
+
+# If you look back to the output of 'rasbr' and look at 'coord. ref.', you can see that the CRS (coordinate reference system) is in WGS84 (+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0)
+
+
+# Because I am using these files for Maxent, I need the projection to be an albers equal area. We can project the whole raster brick. There are a couple of ways to do this, but the simplest way is to use an existing file in the projection I want. The file I will use for this is called aea.tif. I've sent it to you in the email with this code as a zip. Note that there are two files in the zip folder - # you will need both (but you only call on the .tif in the code)
+aea <- raster("aea.tif") # change the file path to where ever the tif and .twf files are saved
+rasbr_aea <- projectRaster(rasbr, aea)
+plot(rasbr_aea[[1]]) # just to look at the top layer of the raster brick after projection so you can see the difference
+
+
+# The next piece of data we need is the observation data itself. I've sent you the observations.csv in the email with this code.
[…]

### 2018-11-02 src/raster_sjl.Rmd
depth layer selection solved (see take 2 for code)

@@
 ```{r}
-obs <- read.csv("../data/observations3.csv") #change this path to your .csv location
+obs <- read.csv("../data/observations3.csv", header = TRUE) #change this path to your .csv location
 head(obs)
@@
 ```{r}
-test15 <- extract(x=rasbr_aea[[15]], y = obssp)
-test22 <- extract(x=rasbr_aea[[22]], y = obssp)
-test17 <- extract(x=rasbr_aea[[17]], y = obssp)
+testrow1 <- read.csv("../data/row1.csv", header = TRUE)
+testrow1xy <- testrow1[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow1sp <- SpatialPointsDataFrame(coords = testrow1xy, data = testrow1, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow1sp
+test15 <- extract(x=rasbr_aea[[15]], y = testrow1sp)
+test15
+testrow2 <- read.csv("../data/row1.csv", header = TRUE)
+testrow2xy <- testrow2[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow2sp <- SpatialPointsDataFrame(coords = testrow2xy, data = testrow2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow2sp
+test22 <- extract(x=rasbr_aea[[22]], y = testrow2sp)
+test22
+testrow3 <- read.csv("../data/row1.csv", header = TRUE)
+testrow3xy <- testrow3[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow3sp <- SpatialPointsDataFrame(coords = testrow3xy, data = testrow3, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow3sp
+test17 <- extract(x=rasbr_aea[[17]], y = testrow3sp)
 test17
 ```
+
+row 1 (dpeth 15) should be 32.8761...
+row 2 (depth 22) should be 33.10022...
+row 3 (depth 17) should be 32.90343...
 
@@
 take 2
+
 ```{r}
-Data_Loop <-NULL #Gonna put my loop data in here
+#Data_Loop <-NULL #Gonna put my loop data in here
 for (i in 1:length(obssp)) {  # CYCLES THROUGH THE ROWS
   de <- obssp$depth[[i]]  #have a variable for the depth layer
-  print(de)
-  obssp$samtest <- extract(x=rasbr_aea[[de]], y = obssp) #this is the equation you will be calculating for every value of y
+  #print(de)
+  obssp$samtest[i] <- extract(x=rasbr_aea[[de]], y = obssp) #this is the equation you will be calculating for every value of y
 
@@
   #print the i value so you can track progress. Good for longer loops
-  #print(i)
+  print(i)
   #Close the loop
@@
 obssp$samtest
+obssp
 ```
+
+so with this code, the values are correct but I get the following error "number of items to replace is not a multiple of replacement length". I don't know why... but it works...
+
+Ok next step, can you dynamically name the column with the variable in question (or at least the name of the original netcdf)

### 2018-11-02 src/raster_sjl.Rmd
searching through year and month to add correct data...

@@
 obs <- read.csv("../data/observations3.csv", header = TRUE) #change this path to your .csv location
+obs$month <- sprintf("%02d",obs$month) # new addition to add the 0 infront of single digit... test for seeing if you can match the netcdf name for year month selection of bricks
 head(obs)
+
 ```
@@
 
-let's get the values manually
+let's get the values manually at depth
 ```{r}
@@
 
-row 1 (dpeth 15) should be 32.8761...
+row 1 (depth 15) should be 32.8761...
 row 2 (depth 22) should be 33.10022...
 row 3 (depth 17) should be 32.90343...
+
+let's get the values manually at the surface
+```{r}
+testrow1.1 <- extract(x=rasbr_aea[[1]], y = testrow1sp)
+testrow1.1
+testrow2.1 <- extract(x=rasbr_aea[[1]], y = testrow2sp)
+testrow2.1
+testrow3.1 <- extract(x=rasbr_aea[[1]], y = testrow3sp)
+testrow3.1
+```
 
@@
 
-Ok next step, can you dynamically name the column with the variable in question (or at least the name of the original netcdf)
+test with just extracting the surface layer data..
+
+```{r}
+for (i in 1:length(obssp)) {  # CYCLES THROUGH THE ROWS
+  obssp$testsurface[i] <- extract(x=rasbr_aea[[1]], y = obssp) #this is the equation you will be calculating for every value of y
+  print(i)
+  }
+obssp$testsurface
+obssp
+```
+
+scrap the variable name matching idea (instead just make a loop for each variable, and each variable held in a different folder). just try select by year and month.
+note you have changed the dates on the .csv file so reload it
+
+```{r warning=FALSE}
+netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf) {
+  print(netcdf_name[i]) #this just prings the name of the netCDF R is working on
+  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # here I am extracting the first part of the filename deliminated with _ (which is the year)
+  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # here I am extracting the second part of the filename deliminated with _ (which is the month)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) #aea is an existing raster in the projection I want that I created (environmental_data_preperation)
+    for (j in 1:length(obssp)) {  # CYCLES THROUGH THE ROWS
+      yr <- obssp$year[j]
+      mth <- obssp$month[j]
+      if (brkyr == yr &  brkmth == mth){
+        de <- obssp$depth[[j]]  #have a variable for the depth layer
+        obssp$samtest[j] <- extract(x=temp_brick[[de]], y = obssp) #this is the equation you will be calculating for every value of y
+        obssp$samtesttop[j] <- extract(x=temp_brick[[1]], y = obssp) #this is the equation you will be calculating for every value of y
+        print(j)
+      }
+    }
+}
+obssp$samtesttop
+```

### 2018-11-02 src/raster_sjl.Rmd
messing with if statement in loop... not working...

--- 2018-11-02 src/raster_sjl.Rmd
@@
 ```{r}
-#Data_Loop <-NULL #Gonna put my loop data in here
 for (i in 1:length(obssp)) {  # CYCLES THROUGH THE ROWS
   de <- obssp$depth[[i]]  #have a variable for the depth layer
-  #print(de)
-  obssp$samtest[i] <- extract(x=rasbr_aea[[de]], y = obssp) #this is the equation you will be calculating for every value of y
-
-  #Put your data in a data frame. Don't forget to enclose this within the loop!
-  #Data_Loop <- rbind(Data_Loop, data.frame(depth=obssp$depth, year=obssp$year, month=obssp$month, lon=obssp$lon, lat=obssp$lat, Salinity=sal))
-
-  #print the i value so you can track progress. Good for longer loops
+  obssp$salinity_depth[i] <- extract(x=rasbr_aea[[de]], y = obssp) #this is the equation you will be calculating for every value of y
+  obssp$salinity_surface[i] <- extract(x=rasbr_aea[[1]], y = obssp) #this is the equation you will be calculating for every value of y
   print(i)
-  #Close the loop
 }
-#Data_Loop
-obssp$samtest
-obssp
+obssp$salinity_depth
+obssp$salinity_surface
 ```
@@
 
-Ok next step, can you dynamically name the column with the variable in question (or at least the name of the original netcdf)
+test with just extracting the surface layer data..
+
+```{r}
+for (i in 1:length(obssp)) {  # CYCLES THROUGH THE ROWS
+  obssp$testsurface[i] <- extract(x=rasbr_aea[[1]], y = obssp) #this is the equation you will be calculating for every value of y
+  print(i)
+  }
+obssp$salinity_surface
+obssp
+```
+
+scrap the variable name matching idea (instead just make a loop for each variable, and each variable held in a different folder). just try select by year and month.
+note you have changed the dates on the .csv file so reload it
+
+```{r warning=FALSE}
+obs <- read.csv("../data/observations3.csv", header = TRUE) #change this path to your .csv location
+obs$month <- sprintf("%02d",obs$month) # new addition to add the 0 infront of single digit... test for seeing if you can match the netcdf name for year month selection of bricks
+xy <- obs[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+obssp <- SpatialPointsDataFrame(coords = xy, data = obs, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+
+#the loop....
+netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf){  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
+  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # extracting the second part of the netcdf filename (which is the month)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:length(obssp)) {  
+      print(j)
+      de <- obssp$depth[[j]]  # a variable for the observation depth layer
+      yr <- obssp$year[j]  # a variable for the observation year
+      mth <- obssp$month[j]  # a variable for the observation month
+      #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
+        if (brkyr == yr & brkmth == mth){
+          obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
+          obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
+
+        }
+    }
+}
+obssp$salinity_depth
+obssp$salinity_surface
+```
+not extracting the correct values...
+
+check
+
+load in the three netcdfs
+```{r}
+nc1 <-  "../data/env/testnetcdf/salinity/1993_01_salinity.nc" # change the file path to where ever your netcdfs are saved
+row1rast <- brick(nc1,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+nc2 <-  "../data/env/testnetcdf/salinity/1998_01_salinity.nc" # change the file path to where ever your netcdfs are saved
+row2rast <- brick(nc2,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+nc3 <-  "../data/env/testnetcdf/salinity/1998_02_salinity.nc" # change the file path to where ever your netcdfs are saved
+row3rast <- brick(nc3,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+```
+
+let's get the values manually at depth
+```{r}
+testrow1 <- read.csv("../data/row1.csv", header = TRUE)
+testrow1$month <- sprintf("%02d",testrow1$month)
+testrow1xy <- testrow1[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow1sp <- SpatialPointsDataFrame(coords = testrow1xy, data = testrow1, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow1sp
+test15 <- extract(x=row1rast[[15]], y = testrow1sp)
+test15
+testrow2 <- read.csv("../data/row2.csv", header = TRUE)
+testrow2$month <- sprintf("%02d",testrow2$month)
+testrow2xy <- testrow2[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow2sp <- SpatialPointsDataFrame(coords = testrow2xy, data = testrow2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow2sp
+test22 <- extract(x=row2rast[[22]], y = testrow2sp)
+test22
+testrow3 <- read.csv("../data/row3.csv", header = TRUE)
+testrow3$month <- sprintf("%02d",testrow3$month)
+testrow3xy <- testrow3[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow3sp <- SpatialPointsDataFrame(coords = testrow3xy, data = testrow3, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow3sp
+test17 <- extract(x=row3rast[[17]], y = testrow3sp)
+test17
+```
+
+let's get the values manually at the surface
+```{r}
+testrow1.1 <- extract(x=row1rast[[1]], y = testrow1sp)
+testrow1.1
+testrow2.1 <- extract(x=row2rast[[1]], y = testrow2sp)
+testrow2.1
+testrow3.1 <- extract(x=row3rast[[1]], y = testrow3sp)
+testrow3.1
+```

### 2018-11-05 src/timesliced_random_cell_selection.Rmd
can upload

@@
 
-2) reproject to a wgs
+2) reproject to a wgs (use an existing file as a mask)
 
@@
 3) extract lonlat, add to dataframe
+save as csv
+
+use the gebco raster as a mask?
 
 ```{r}
-head(sp_background)
+wgs84_mask <- gebco
+plot(wgs84_mask)
 ```
-
-save as csv

### 2018-11-07 src/raster_sjl.Rmd
lat lon on background points done (+ little progress on the raster brick loop)

@@
 library(raster)
+library(rgdal)
 ```
@@
 plot(rasbr_aea[[1]]) # just to look at the top layer of the raster brick after projection so you can see the difference
+rasbr_aea
 ```
@@
           obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
-
         }
@@
 }
-obssp$salinity_depth
-obssp$salinity_surface
+
+str(t)
 ```
+
 not extracting the correct values...
@@
 ```
+
+shawn suggests creating a table/list of raster bricks with year and month data. Actually need to do this with the netCDFs (as you want to load the netCDF to convert)
+```{r}
+#the loop....
+netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE)
+nctable <-as.data.frame(netcdf_list)
+no_netcdf <- length(netcdf_list) 
+netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) 
+library (ncdf4)
+for (i in 1:no_netcdf){  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
+  nctable$year[i] <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1)
+  nctable$month[i] <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2)
+}
+
+```
+
+a new loop... no this isnt going to work...
+```{r warning=FALSE}
+#the loop....
+library (ncdf4)
+netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+for (j in 1:length(obssp)) {  
+      print(j)
+      de <- obssp$depth[[j]]  # a variable for the observation depth layer
+      yr <- obssp$year[j]  # a variable for the observation year
+      mth <- obssp$month[j]  # a variable for the observation month
+      if (yr == nctable$year & mth == nctable$month) {
+        temp_brick <- brick(netcdf_list[i], lvar = 4)
+        temp_brick <- projectRaster(temp_brick, aea) 
+        for (j in 1:length(obssp)) {  
+        print(j)
+        de <- obssp$depth[[j]]  # a variable for the observation depth layer
+        obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
+        obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
+        }
+      }
+}
+```
+
+hmmm play...
+```{r}
+netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf){  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
+  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # extracting the second part of the netcdf filename (which is the month)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:length(obssp)) {  
+      print(j)
+      de <- obssp$depth[[j]]  # a variable for the observation depth layer
+      yr <- obssp$year[j]  # a variable for the observation year
+      mth <- obssp$month[j]  # a variable for the observation month
+      #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
+        for (k in 1:length(obssp)){
+          if (brkyr == yr & brkmth == mth){
+            obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
+            obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
+          }
+        } 
+      }
+}
+obssp$salinity_depth
+```
+
+netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+for (i in 1:no_netcdf){  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
+  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # extracting the second part of the netcdf filename (which is the month)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:length(obssp)) {  
+      print(j)
+      de <- obssp$depth[[j]]  # a variable for the observation depth layer
+      yr <- obssp$year[j]  # a variable for the observation year
+      mth <- obssp$month[j]  # a variable for the observation month
+      #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
+        if (brkyr == yr & brkmth == mth){
+          obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
+          obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
+        }
+    }
+}
+obssp$salinity_depth
+obssp$salinity_surface

### 2018-11-07 src/timesliced_random_cell_selection.Rmd
lat lon on background points done (+ little progress on the raster brick loop)

--- 2018-10-31 src/timesliced_random_cell_selection.Rmd
@@
 library(sp)
+background_all <- as.data.frame(background_all)
 xy <- background_all[ ,c(1,2)] #lon lat in col 1 and 2
@@
 ```
+
 ```{r}
@@
 ```{r}
+sp_background <- data.frame(test, long=coordinates(sp_background)[,1], lat=coordinates(sp_background)[,2])
 head(sp_background)
@@
 save as csv
+```{r}
+write.csv(adddata_aea, file = "../output/bio/data_aea.csv", row.names = FALSE)
+
+```

### 2018-11-07 src/timesliced_random_cell_selection.Rmd
NAFO region addition to background points

@@
 
+change name of long lat columns to decimalLongitude and decimalLatitude respectively
+```{r}
+sp_background <- as.data.frame(sp_background) #need to go back to a 'normal' dataframe to change col names
+head(sp_background) #note the columns in question have changed name...
+colnames(sp_background)[colnames(sp_background)=="longitude.1"] <- "decimalLongitude"
+colnames(sp_background)[colnames(sp_background)=="latitude.1"] <- "decimalLatitude"
+head(sp_background)
+```
+
 save as csv
 ```{r}
-write.csv(adddata_aea, file = "../output/bio/data_aea.csv", row.names = FALSE)
+write.csv(sp_background, file = "../output/bio/background_all_lonlat.csv", row.names = FALSE)
+```
 
+# Add NAFO Region background point is in
+NAFO Zones shapefile obtained from [NAFO](https://www.nafo.int/Data/GIS)
+
+```{r NAFO shapefile extraction}
+library(rgdal)
+library(sp)
+coordinates(sp_background) <- c("decimalLongitude", "decimalLatitude") 
+nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
+proj4string(sp_background) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
+sp_background$nafo_zone <- over(sp_background, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
+head(sp_background) #just to check it worked
 ```
+
+Great. Lets see if any are missing a NAFO Zone. But first, write to a csv and reload as the data is now some wierd 'spatial points dataframe'...
+```{r}
+write.csv(sp_background, "../output/bio/background_all_lonlat_nafo.csv", row.names=FALSE)
+sp_background <- as.data.frame(sp_background) 
+bc_nafo_na <- subset(sp_background, is.na(sp_background$nafo_zone))
+bc_nafo_na
+```
+ok lots don't have a NAFO region attached. This is probably the Hudson Bay issue again...
+
+Uh oh there is a lot. A bet a bunch of them are in the Hudson Strait (which is outside the NAFO Zone).Map the nafo_na points...
+```{r}
+plot(map2, xlim = c(-65, -40), ylim =c(30, 70), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(bc_nafo_na$decimalLongitude, bc_nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+```
+
+huh that is interesting... I think I want to plot this in Arc and take a closer look...

### 2018-11-08 src/raster_sjl.Rmd
another simple solution done

@@
 
-str(t)
 ```
@@
 row1rast <- brick(nc1,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+row1rast <- projectRaster(row1rast, aea)
+row1rast
 nc2 <-  "../data/env/testnetcdf/salinity/1998_01_salinity.nc" # change the file path to where ever your netcdfs are saved
 row2rast <- brick(nc2,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+row2rast <- projectRaster(row2rast, aea)
+row2rast
 nc3 <-  "../data/env/testnetcdf/salinity/1998_02_salinity.nc" # change the file path to where ever your netcdfs are saved
 row3rast <- brick(nc3,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+row3rast <- projectRaster(row3rast, aea)
+row3rast
 ```
@@
 testrow1 <- read.csv("../data/row1.csv", header = TRUE)
-testrow1$month <- sprintf("%02d",testrow1$month)
+#testrow1$month <- sprintf("%02d",testrow1$month)
 testrow1xy <- testrow1[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
@@
 testrow2 <- read.csv("../data/row2.csv", header = TRUE)
-testrow2$month <- sprintf("%02d",testrow2$month)
+#testrow2$month <- sprintf("%02d",testrow2$month)
 testrow2xy <- testrow2[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
@@
 testrow3 <- read.csv("../data/row3.csv", header = TRUE)
-testrow3$month <- sprintf("%02d",testrow3$month)
+#testrow3$month <- sprintf("%02d",testrow3$month)
 testrow3xy <- testrow3[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
@@
 ```{r}
+obs <- read.csv("../data/observations3.csv", header = TRUE) #change this path to your .csv location
+#obs$month <- sprintf("%02d",obs$month) # new addition to add the 0 infront of single digit... test for seeing if you can match the netcdf name for year month selection of bricks
+xy <- obs[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+obssp <- SpatialPointsDataFrame(coords = xy, data = obs, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+```{r}
 netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
@@
 for (i in 1:no_netcdf){  
-  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
-  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # extracting the first part of the netcdf filename (which is the year)
-  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # extracting the second part of the netcdf filename (which is the month)
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
   temp_brick <- brick(netcdf_list[i], lvar = 4)
   temp_brick <- projectRaster(temp_brick, aea) 
-    for (j in 1:length(obssp)) {  
-      print(j)
+    for (j in 1:nrow(obssp)) {  
+      print (j)
       de <- obssp$depth[[j]]  # a variable for the observation depth layer
-      yr <- obssp$year[j]  # a variable for the observation year
-      mth <- obssp$month[j]  # a variable for the observation month
+      yr <- (obssp$year[j])  # a variable for the observation year
+      mth <- (obssp$month[j])  # a variable for the observation month
       #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
-        for (k in 1:length(obssp)){
+        #for (k in 1:nrow(obssp)){
           if (brkyr == yr & brkmth == mth){
-            obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
-            obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
+            obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp[j, ]) 
+            obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp[j, ]) 
+           }
+
+        #} 
     }
 }
-      }
-}
+```
+
+```{r}
 obssp$salinity_depth
+obssp$salinity_surface
 ```
+done! Next stage... with rows where the year month combo comes up more than once...
 

### 2018-11-08 src/raster_sjl.Rmd
exploration of looping through several years and months that are the same...

@@
 
+```{r}
+obs <- read.csv("../data/observationslong.csv", header = TRUE) #change this path to your .csv location
+#obs$month <- sprintf("%02d",obs$month) # new addition to add the 0 infront of single digit... test for seeing if you can match the netcdf name for year month selection of bricks
+xy <- obs[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+obssp <- SpatialPointsDataFrame(coords = xy, data = obs, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+```{r}
 netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
@@
 netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
 for (i in 1:no_netcdf){  
-  print(netcdf_name[i]) #this just prints the name of the netCDF R is working on
-  brkyr <- sapply(strsplit(netcdf_name[i], "_"), "[[", 1) # extracting the first part of the netcdf filename (which is the year)
-  brkmth <- sapply(strsplit(netcdf_name[i], "_"), "[[", 2) # extracting the second part of the netcdf filename (which is the month)
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
   temp_brick <- brick(netcdf_list[i], lvar = 4)
   temp_brick <- projectRaster(temp_brick, aea) 
-    for (j in 1:length(obssp)) {  
-      print(j)
+    for (j in 1:nrow(obssp)) {  
+      print (j)
       de <- obssp$depth[[j]]  # a variable for the observation depth layer
-      yr <- obssp$year[j]  # a variable for the observation year
-      mth <- obssp$month[j]  # a variable for the observation month
+      yr <- (obssp$year[j])  # a variable for the observation year
+      mth <- (obssp$month[j])  # a variable for the observation month
       #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
+        #for (k in 1:nrow(obssp)){
           if (brkyr == yr & brkmth == mth){
-          obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp) 
-          obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp)
+            obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp[j, ]) 
+            obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp[j, ]) 
+           }
+
+        #} 
     }
 }
-}
+```
+
+```{r}
 obssp$salinity_depth
 obssp$salinity_surface
+```
+output is  
+> obssp$salinity_surface
+[1] 32.82551 33.19321 32.18221 32.23423 32.61295 32.61270 32.39069
+> obssp$salinity_depth
+[1] 32.83107 33.19973       NA 32.51381       NA 32.61270 32.39358
+
+next step - check to see if the values are correct (particularly interested in those NA...)
+
+```{r}
+plot(temp_brick[[40]])
+```
+for the NA - of course - the layer no longer exists at that depth (water shallower than that)

### 2018-11-08 src/raster_sjl.Rmd
working of depth being NA....

@@
 for the NA - of course - the layer no longer exists at that depth (water shallower than that)
+
+check the values for all layers manually
+
+```{r}
+test <- read.csv("../data/observationslong.csv", header = TRUE)
+testrow1 <- test[1, ]
+testrow1
+#testrow1$month <- sprintf("%02d",testrow1$month)
+testrow1xy <- testrow1[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow1sp <- SpatialPointsDataFrame(coords = testrow1xy, data = testrow1, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test10 <- extract(x=row1rast[[10]], y = testrow1sp)
+test10
+testrow2 <- test[2, ]
+#testrow2$month <- sprintf("%02d",testrow2$month)
+testrow2xy <- testrow2[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow2sp <- SpatialPointsDataFrame(coords = testrow2xy, data = testrow2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test12 <- extract(x=row1rast[[12]], y = testrow2sp)
+test12
+testrow3 <- test[3, ]
+testrow3xy <- testrow3[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow3sp <- SpatialPointsDataFrame(coords = testrow3xy, data = testrow3, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow3sp
+test40 <- extract(x=row3rast[[40]], y = testrow3sp)
+test40
+testrow4 <- test[4, ]
+testrow4xy <- testrow4[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow4sp <- SpatialPointsDataFrame(coords = testrow4xy, data = testrow4, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow4sp
+test18 <- extract(x=row3rast[[18]], y = testrow4sp)
+test18
+testrow5 <- test[5, ]
+testrow5xy <- testrow5[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow5sp <- SpatialPointsDataFrame(coords = testrow5xy, data = testrow5, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test27 <- extract(x=row3rast[[27]], y = testrow5sp)
+test27
+testrow6 <- test[6, ]
+testrow6xy <- testrow6[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow6sp <- SpatialPointsDataFrame(coords = testrow6xy, data = testrow6, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test1 <- extract(x=row2rast[[1]], y = testrow6sp)
+test1
+testrow7 <- test[7, ]
+testrow7xy <- testrow7[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow7sp <- SpatialPointsDataFrame(coords = testrow7xy, data = testrow7, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test5 <- extract(x=row2rast[[5]], y = testrow7sp)
+test5
+```
+
+and surface...
+```{r}
+test1.1 <- extract(x=row1rast[[1]], y = testrow1sp)
+test2.1 <- extract(x=row1rast[[1]], y = testrow2sp)
+test3.1 <- extract(x=row3rast[[1]], y = testrow3sp)
+test3.1
+test4.1 <- extract(x=row3rast[[1]], y = testrow4sp)
+test4.1
+test5.1 <- extract(x=row3rast[[1]], y = testrow5sp)
+test5.1
+test6.1 <- extract(x=row2rast[[1]], y = testrow6sp)
+test6.1
+test7.1 <- extract(x=row2rast[[1]], y = testrow7sp)
+test7.1
+```
+pull up values you got from the loop...
+```{r}
+obssp$salinity_depth
+```
+
+```{r}
+obssp$salinity_surface
+```
+ok so the loop works for repetitive years and months
+
+now all the depth values were numbered - in reality I have some NA. How to deal...
+.csv altered to include 1 NA
+
+```{r}
+obs <- read.csv("../data/observationslong.csv", header = TRUE) #change this path to your .csv location
+#obs$month <- sprintf("%02d",obs$month) # new addition to add the 0 infront of single digit... test for seeing if you can match the netcdf name for year month selection of bricks
+xy <- obs[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+obssp <- SpatialPointsDataFrame(coords = xy, data = obs, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+```{r}
+netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf){  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(obssp)) {  
+      print (j)
+      de <- obssp$depth[[j]]  # a variable for the observation depth layer
+      yr <- (obssp$year[j])  # a variable for the observation year
+      mth <- (obssp$month[j])  # a variable for the observation month
+      #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
+        #for (k in 1:nrow(obssp)){
+          if (brkyr == yr & brkmth == mth){
+            obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp[j, ]) 
+            if (is.na(de)){
+              obssp$salinity_depth <- NA
+            } else (de >= 1) {
+              obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp[j, ])
+            }
+           }
+
+        #} 
+    }
+}
+```
+
+```{r}
+obssp$salinity_depth
+```

### 2018-11-09 src/raster_sjl.Rmd
loop works! Just trying something else out

@@
       mth <- (obssp$month[j])  # a variable for the observation month
-      #I think i need an if statement here... if brkyr == yr & brkmth == mth do the data extraction below
-        #for (k in 1:nrow(obssp)){
           if (brkyr == yr & brkmth == mth){
@@
             if (is.na(de)){
-              obssp$salinity_depth <- NA
-            } else (de >= 1) {
+              obssp$salinity_depth[j] <- NA
+            } else  
               obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp[j, ])
           }
-           }
-
-        #} 
     }
@@
 ```{r}
-obssp$salinity_depth
+obssp$salinity_surface
 ```
+above works!
+
+Just play to see if you can get the variable thingy in....
+
+```{r}
+netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf){  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(obssp)) {  
+      print (j)
+      de <- obssp$depth[[j]]  # a variable for the observation depth layer
+      yr <- (obssp$year[j])  # a variable for the observation year
+      mth <- (obssp$month[j])  # a variable for the observation month
+          if (brkyr == yr & brkmth == mth){
+            obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp[j, ]) 
+            if (is.na(de)){
+              obssp$salinity_depth[j] <- NA
+            } else  
+              obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp[j, ])
+          }
+    }
+}
+```

### 2018-11-09 src/raster_sjl.Rmd
variable selection done

@@
   brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
   temp_brick <- brick(netcdf_list[i], lvar = 4)
@@
       mth <- (obssp$month[j])  # a variable for the observation month
-          if (brkyr == yr & brkmth == mth){
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+
+             obssp$temp_surface[j] <- extract(x=temp_brick[[1]], y = obssp[j, ]) 
+              if (is.na(de)){
+                obssp$temp_depth[j] <- NA
+              } else  
+                obssp$temp_depth[j] <- extract(x=temp_brick[[de]], y = obssp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
               obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp[j, ]) 
@@
           }
+
 ```
+
+```{r}
+obssp$temp_depth
+```
+
+check temp
+
+```{r}
+nc1 <-  "../data/env/testnetcdf/salinity/1993_01_temp.nc" # change the file path to where ever your netcdfs are saved
+row1rast <- brick(nc1,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+row1rast <- projectRaster(row1rast, aea)
+row1rast
+nc2 <-  "../data/env/testnetcdf/salinity/1998_01_temp.nc" # change the file path to where ever your netcdfs are saved
+row2rast <- brick(nc2,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+row2rast <- projectRaster(row2rast, aea)
+row2rast
+nc3 <-  "../data/env/testnetcdf/salinity/1998_02_temp.nc" # change the file path to where ever your netcdfs are saved
+row3rast <- brick(nc3,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
+row3rast <- projectRaster(row3rast, aea)
+row3rast
+```
+
+```{r}
+testrow1 <- read.csv("../data/row1.csv", header = TRUE)
+#testrow1$month <- sprintf("%02d",testrow1$month)
+testrow1xy <- testrow1[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow1sp <- SpatialPointsDataFrame(coords = testrow1xy, data = testrow1, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow1sp
+test15 <- extract(x=row1rast[[15]], y = testrow1sp)
+test15
+testrow2 <- read.csv("../data/row2.csv", header = TRUE)
+#testrow2$month <- sprintf("%02d",testrow2$month)
+testrow2xy <- testrow2[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow2sp <- SpatialPointsDataFrame(coords = testrow2xy, data = testrow2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow2sp
+test22 <- extract(x=row2rast[[22]], y = testrow2sp)
+test22
+testrow3 <- read.csv("../data/row3.csv", header = TRUE)
+#testrow3$month <- sprintf("%02d",testrow3$month)
+testrow3xy <- testrow3[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow3sp <- SpatialPointsDataFrame(coords = testrow3xy, data = testrow3, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow3sp
+test17 <- extract(x=row3rast[[17]], y = testrow3sp)
+test17
+```
+
+```{r}
+testrow1.1 <- extract(x=row1rast[[1]], y = testrow1sp)
+testrow1.1
+testrow2.1 <- extract(x=row2rast[[1]], y = testrow2sp)
+testrow2.1
+testrow3.1 <- extract(x=row3rast[[1]], y = testrow3sp)
+testrow3.1
+```
+
+yes! ok now for the bigger test...

### 2018-11-09 src/raster_sjl.Rmd
full loop check - thunderbirds are go!

@@
 yes! ok now for the bigger test...
+
+```{r}
+obs <- read.csv("../data/observationslong.csv", header = TRUE) #change this path to your .csv location
+#obs$month <- sprintf("%02d",obs$month) # new addition to add the 0 infront of single digit... test for seeing if you can match the netcdf name for year month selection of bricks
+xy <- obs[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+obssp <- SpatialPointsDataFrame(coords = xy, data = obs, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+```{r}
+netcdf_list <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/testnetcdf/salinity", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf){  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(obssp)) {  
+      print (j)
+      de <- obssp$depth[[j]]  # a variable for the observation depth layer
+      yr <- (obssp$year[j])  # a variable for the observation year
+      mth <- (obssp$month[j])  # a variable for the observation month
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+
+             obssp$temp_surface[j] <- extract(x=temp_brick[[1]], y = obssp[j, ]) 
+              if (is.na(de)){
+                obssp$temp_depth[j] <- NA
+              } else  
+                obssp$temp_depth[j] <- extract(x=temp_brick[[de]], y = obssp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              obssp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = obssp[j, ]) 
+              if (is.na(de)){
+                obssp$salinity_depth[j] <- NA
+              } else  
+                obssp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = obssp[j, ]) 
+          }
+          } 
+          }
+
+```
+
+```{r}
+obssp$temp_depth
+```
+
+checkt the temp
+
+```{r}
+test <- read.csv("../data/observationslong.csv", header = TRUE)
+testrow1 <- test[1, ]
+testrow1
+#testrow1$month <- sprintf("%02d",testrow1$month)
+testrow1xy <- testrow1[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow1sp <- SpatialPointsDataFrame(coords = testrow1xy, data = testrow1, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test10 <- extract(x=row1rast[[10]], y = testrow1sp)
+test10
+testrow2 <- test[2, ]
+#testrow2$month <- sprintf("%02d",testrow2$month)
+testrow2xy <- testrow2[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow2sp <- SpatialPointsDataFrame(coords = testrow2xy, data = testrow2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test12 <- extract(x=row1rast[[12]], y = testrow2sp)
+test12
+testrow3 <- test[3, ]
+testrow3xy <- testrow3[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow3sp <- SpatialPointsDataFrame(coords = testrow3xy, data = testrow3, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow3sp
+test40 <- extract(x=row3rast[[40]], y = testrow3sp)
+test40
+testrow4 <- test[4, ]
+testrow4xy <- testrow4[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow4sp <- SpatialPointsDataFrame(coords = testrow4xy, data = testrow4, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+testrow4sp
+test18 <- extract(x=row3rast[[18]], y = testrow4sp)
+test18
+testrow5 <- test[5, ]
+testrow5xy <- testrow5[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow5sp <- SpatialPointsDataFrame(coords = testrow5xy, data = testrow5, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test27 <- extract(x=row3rast[[27]], y = testrow5sp)
+test27
+testrow6 <- test[6, ]
+testrow6xy <- testrow6[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow6sp <- SpatialPointsDataFrame(coords = testrow6xy, data = testrow6, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test1 <- extract(x=row2rast[[1]], y = testrow6sp)
+test1
+testrow7 <- test[7, ]
+testrow7xy <- testrow7[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+testrow7sp <- SpatialPointsDataFrame(coords = testrow7xy, data = testrow7, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+test5 <- extract(x=row2rast[[5]], y = testrow7sp)
+test5
+```
+
+```{r}
+test1.1 <- extract(x=row1rast[[1]], y = testrow1sp)
+test1.1
+test2.1 <- extract(x=row1rast[[1]], y = testrow2sp)
+test2.1
+test3.1 <- extract(x=row3rast[[1]], y = testrow3sp)
+test3.1
+test4.1 <- extract(x=row3rast[[1]], y = testrow4sp)
+test4.1
+test5.1 <- extract(x=row3rast[[1]], y = testrow5sp)
+test5.1
+test6.1 <- extract(x=row2rast[[1]], y = testrow6sp)
+test6.1
+test7.1 <- extract(x=row2rast[[1]], y = testrow7sp)
+test7.1
+```
+yes!!

### 2018-11-12 src/biological_data_prep.Rmd
Add loop to extract env. correlatea to real code (prence points only)

@@
 
-#add columns for the env. variables
-The ones ypu have to add are...
-
-mlp
-salinty_1
-salinity_depth
-ssh
-temp_1
-temp_depth
-chl_1
-chl_depth
-o2_1
-o2_depth
-
-populate the columns with NA
-```{r}
-data_aea$mlp <- NA
-data_aea$salinty_1 <- NA
-data_aea$salinity_depth <- NA
-data_aea$ssh <- NA
-data_aea$temp_1 <- NA
-data_aea$temp_depth <- NA
-data_aea$chl_1 <- NA
-data_aea$chl_depth <- NA
-data_aea$o2_1 <- NA
-data_aea$o2_depth <- NA
-write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins_env.csv", row.names = FALSE)
-```
-
 # observations by cell (in total, and also by time step)

### 2018-11-12 src/environmental_data_preperation.Rmd
Add loop to extract env. correlatea to real code (prence points only)

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("ncdf4", "raster", "rgdal", "useful", "marmap") 
+packages <- c("ncdf4", "raster", "rgdal", "useful", "marmap", "sp") 
 source("../src/ipak.R")
@@
 
-And now create a raster of Gebco in aea for mapping...
+And now create a raster of Gebco in aea for mapping once the SDM is done...
 
@@
 ```
+
+# attach environmental correlates
+
+This loop runs through the netcdf files and then looks for which rows in data_aea it should extract the value to point from, and at what depth (netcDF layer)
+
+First you need to put the observations (data_aea) into a spatialpointsdataframe.
+```{r}
+library(sp)
+xy2 <- data_aea[ ,c(18,19)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy2, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+and the loop. Note that MLP and SSH only have a single layer and as such they do not have a statement to extract at depth.
+```{r}
+netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf){  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(data_aea_sp)) {  
+      print (j)
+      de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
+      yr <- (data_aea_sp$year[j])  # a variable for the observation year
+      mth <- (data_aea_sp$month[j])  # a variable for the observation month
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+
+             data_aea_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$temp_depth[j] <- NA
+              } else  
+                data_aea_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              data_aea_sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$salinity_depth[j] <- NA
+              } else  
+                data_aea_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
+              data_aea_sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$chl_depth[j] <- NA
+              } else  
+                data_aea_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
+              data_aea_sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$o2_depth[j] <- NA
+              } else  
+                data_aea_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
+              data_aea_sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
+              data_aea_sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
+
+          }
+    } 
+  }
+}
+
+```

### 2018-11-12 src/env_data_prep.Rmd
i crashed the environmental_data_preperation notebook... recreated in env_data_prep

--- 2018-11-12 src/environmental_data_preperation.Rmd
@@
 template <- raster(list(x = lon, y = lat, z = v[,,1]), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ")
-
 ## apply template to brick, and set values to v (but flipped in Y because that's how raster works - lat is increasing in the NetCDF, so we go upside down)
 diy <- setValues(brick(template), v[, ncol(v):1, ])
-
 ## target
@@
 writeOGR(uniquecell_poly, dsn = '.', layer = '../output/env/glo_unique_cell_poly', driver = "ESRI Shapefile")
-
 ```
@@
 write.csv(naophases, file = "../output/env/nao.csv", row.names = FALSE) #for some reason you need to write to a csv and then reload for the next step to work properly...
-
 ```
@@
 library(sp)
-xy2 <- data_aea[ ,c(18,19)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
-data_aea_sp <- SpatialPointsDataFrame(coords = xy2, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+xy <- data_aea[ ,c(18,19)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
 ```
 
-and the loop. Note that MLP and SSH only have a single layer and as such they do not have a statement to extract at depth.
 ```{r}
@@
 no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
-netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not includedCores
 library (ncdf4)
-for (i in 1:no_netcdf){  
+library(sp)
+library(rgdal)
+library(raster)
+for (i in 1:no_netcdf) {  
   print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
@@
       print (j)
-      de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
+      de <- data_aea_sp$depth[[j]]  # a variable for the observation depth layer
       yr <- (data_aea_sp$year[j])  # a variable for the observation year
@@
           }
+
     }
 }
-}
-
 ```

### 2018-11-12 src/environmental_data_preperation.Rmd
i crashed the environmental_data_preperation notebook... recreated in env_data_prep

@@
     for (j in 1:nrow(data_aea_sp)) {  
-      print (j)
       de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
@@
           }
-    } 
+
     }

### 2018-11-12 src/raster_sjl.Rmd
i crashed the environmental_data_preperation notebook... recreated in env_data_prep

@@
 yes!!
+
+full variables
+
+```{r}
+data_aea <- read.csv("../data/observations3.csv", header = TRUE) #change this path to your .csv location
+xy <- data_aea[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+```{r}
+
+netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not includedCores
+library (ncdf4)
+library(sp)
+library(rgdal)
+#no_cores <- detectCores() -1
+#core_cluster <- makeCluster(no_cores)
+#registerDoParallel(core_cluster)
+for (i in 1:no_netcdf) {  
+  library(raster)
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(data_aea_sp)) {  
+      print (j)
+      de <- data_aea_sp$depth[[j]]  # a variable for the observation depth layer
+      yr <- (data_aea_sp$year[j])  # a variable for the observation year
+      mth <- (data_aea_sp$month[j])  # a variable for the observation month
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+
+             data_aea_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$temp_depth[j] <- NA
+              } else  
+                data_aea_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              data_aea_sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$salinity_depth[j] <- NA
+              } else  
+                data_aea_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
+              data_aea_sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$chl_depth[j] <- NA
+              } else  
+                data_aea_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
+              data_aea_sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$o2_depth[j] <- NA
+              } else  
+                data_aea_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
+              data_aea_sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
+              data_aea_sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
+
+          }
+
+    }
+}
+
+#stopCluster(core_cluster)
+```
+
+```{r}
+data_aea_sp$mlp_surface
+data_aea_sp$salinity_depth
+```

### 2018-11-12 src/env_data_prep.Rmd
env. correlates attached to presence

@@
 library(rgdal)
+for (i in 1:no_netcdf) {  
   library(raster)
-for (i in 1:no_netcdf) {  
   print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
@@
     for (j in 1:nrow(data_aea_sp)) {  
-      print (j)
-      de <- data_aea_sp$depth[[j]]  # a variable for the observation depth layer
+      de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
       yr <- (data_aea_sp$year[j])  # a variable for the observation year
@@
           if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
-
               data_aea_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
@@
 ```
+
+```{r}
+head(data_aea_sp$chl_depth)
+```
+
+change dataset back into a dataframe and then save as a .csv
+
+```{r}
+data_aea <- as.data.frame(data_aea_sp)
+write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env.csv", row.names = FALSE)
+```

### 2018-11-12 src/raster_sjl.Rmd
env. correlates attached to presence

--- 2018-11-09 src/raster_sjl.Rmd
@@
 yes!!
+
+full variables
+
+```{r}
+test <- read.csv("../data/looptest.csv", header = TRUE) #change this path to your .csv location
+xy <- test[ ,c(18,19)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = test, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+```{r}
+
+netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not includedCores
+library (ncdf4)
+library(sp)
+library(rgdal)
+for (i in 1:no_netcdf) {  
+  library(raster)
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(data_aea_sp)) {  
+      de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
+      yr <- (data_aea_sp$year[j])  # a variable for the observation year
+      mth <- (data_aea_sp$month[j])  # a variable for the observation month
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+              data_aea_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$temp_depth[j] <- NA
+              } else  
+                data_aea_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              data_aea_sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$salinity_depth[j] <- NA
+              } else  
+                data_aea_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
+              data_aea_sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$chl_depth[j] <- NA
+              } else  
+                data_aea_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
+              data_aea_sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$o2_depth[j] <- NA
+              } else  
+                data_aea_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
+              data_aea_sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
+              data_aea_sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
+
+          }
+
+    }
+}
+
+```
+
+```{r}
+data_aea_sp$mlp_surface
+data_aea_sp$salinity_depth
+data_aea_sp$o2_surface
+```
+
+```{r}
+str(data_aea)
+```
+
+```{r}
+str(obs)
+```

### 2018-11-13 src/data_exploration.Rmd
removing some test scripts i no longer need, plus data exploration pt 2

---
title: "Data Exploration"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Exploration of all the data collected on the presence points

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there are probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
```{r pre-install packages, message=FALSE}
packages <- c() 
source("../src/ipak.R")
ipak(packages)
```

Read in the presence points

```{r}
presence <- read.csv("../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env.csv", header = TRUE)
head(presence)
```

# Update: Who sampled in which year and which month:

year
```{r}
inst_by_yr <- with(presence, table(year, institutioncode))
write.csv(inst_by_yr, file = "../output/bio/institutioncode_by_yr.csv")
inst_by_yr
```

month
```{r}
inst_by_mt <- with(presence, table(month, institutioncode))
write.csv(inst_by_mt, file = "../output/bio/institutioncode_by_mth.csv")
inst_by_mt
```

### 2018-11-13 src/env_data_prep.Rmd
removing some test scripts i no longer need, plus data exploration pt 2

@@
 I don't know what I'm doing.
-The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.
+The code herein is unlikely to be elegant and there are probably more efficient ways of running the code.
 
@@
 ```
+note that if you wanted to just loop through the netcdfs and create a ton of raster bricks with the same name as the netcdf, the following code you wrote before works
+
+netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf) {
+  print(netcdf_name[i]) #this just prings the name of the netCDF R is working on
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) #aea is an existing raster in the projection I want that I created (environmental_data_preperation)
+  assign(netcdf_name[i], temp_brick) #this asigns the netCDF name to the raster brick
+}
 

### 2018-11-13 src/netcdf_to_rasterbrick.R

@@ -1,10 +1,0 @@
-# Function to take all netCDFs in one folder and transform into raster bricks
-
-netcdf_to_rasterbrick <- function( ncdf_filename_input ) {
-  mcdf_filename_input %>%
-    file.path(../data/env/netcdf, . ) %>%
-    nc_open(.) %>%
-    ncvar_get(.) %>%
-    brick(.) %>%
-    return()
-}

### 2018-11-13 src/rasterbrick_to_ae.R

@@ -1,7 +1,0 @@
-# project rasterbricks into albers equal area based on a tiff that has been reprojected already (called aea)
-
-rasterbrick_to_aea <- function( brick_input ) {
-  brick_input %>%
-    projectRaster(. , proj4string(aea)) %>%
-    return(.)
-}

### 2018-11-13 src/data_exploration.Rmd
packrat update

@@
 Note this function checks to see if the packages are installed first.
-```{r pre-install packages, message=FALSE}
-packages <- c() 
+The "include=FALSE" supresses the package installation text appearing in the document...
+```{r pre-install packages, include=FALSE}
+packages <- c("plyr") 
 source("../src/ipak.R")
@@
 ```
+
+# Update: Observations by Cell (in total and by time-step)
+
+as the title says... how many observations happeend in each cell:
+- in total
+
+```{r observation by cell - total}
+library(plyr) #to use the count function
+obs_by_cell <- count(presence, "cell_id")
+write.csv(obs_by_cell, file = "../output/bio/no_observations_cell.csv")
+head(obs_by_cell)
+```
+
+- in each yyyymm
+```{r}
+obs_cell_yymm <- count(presence, c("cell_id", "year", "month"))
+write.csv(obs_cell_yymm, file = "../output/bio/no_observations_cell_yymm.csv")
+head(obs_cell_yymm)
+```

### 2018-11-14 src/dataextraction_SL.R
removing script no longer needed

@@ -1,107 +1,0 @@
-# load the netcdf and raster packages
-library(ncdf4)
-library(raster)
-
-
-# The data are held in netCDF files. We don't work directly with the netCDFs in R. Instad we load a netcdf file and convert it to a raster brick so we can work with it. In this example the netCDF I'll # # use is called '1993_01_salinity.nc'.I've sent it to you in the email with this code.
-nc <-  "1993_01_salinity.nc" # change the file path to where ever your netcdfs are saved
-rasbr <- brick(nc,lvar = 4) #lvar = 4 is used because the netCDF contains four dimensions. Without this command, the raster brick will only have one level (the top one, which is my surface data)
-rasbr # this will give you some information about the raster brick
-
-# We no longer care that the data originated from a netCDF file. Everything now is done with the raster brick.
-
-# We can plot the different layers (which represent depth) inside the raster brick.
-
-#SL: gonna check some stuff on nc
-print(nc)
-nc2 <- nc_open(nc)
-attributes(nc2$var)
-data <- ncvar_get(nc2, attributes(nc2$var)$names[1]) #get salinity data
-dim(data)
-
-# Plot the top layer
-plot(rasbr[[1]]) #the double brackets tell R which layer to plot
-
-# plot layer 20
-plot(rasbr[[20]])
-
-# If you look back to the output of 'rasbr' and look at 'coord. ref.', you can see that the CRS (coordinate reference system) is in WGS84 (+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0)
-
-
-# Because I am using these files for Maxent, I need the projection to be an albers equal area. We can project the whole raster brick. There are a couple of ways to do this, but the simplest way is to use an existing file in the projection I want. The file I will use for this is called aea.tif. I've sent it to you in the email with this code as a zip. Note that there are two files in the zip folder - # you will need both (but you only call on the .tif in the code)
-aea <- raster("aea.tif") # change the file path to where ever the tif and .twf files are saved
-rasbr_aea <- projectRaster(rasbr, aea)
-plot(rasbr_aea[[1]]) # just to look at the top layer of the raster brick after projection so you can see the difference
-
-
-# The next piece of data we need is the observation data itself. I've sent you the observations.csv in the email with this code.
[…]

### 2018-11-14 src/data_exploration.Rmd
remvoing stuff ive already done (doh)
simple plots of env. correlates

--- 2018-11-13 src/data_exploration.Rmd
@@
 Note this function checks to see if the packages are installed first.
-```{r pre-install packages, message=FALSE}
-packages <- c() 
+The "include=FALSE" supresses the package installation text appearing in the document...
+```{r pre-install packages, include=FALSE}
+packages <- c("plyr", "graphics") 
 source("../src/ipak.R")
@@
 ```
+
+# environmental correlates
+
+1) Get the max, min, an mean values and add into a dataframe
+
+Temperature
+```{r}
+temp_depth_max <- max(presence$temp_depth, na.rm = TRUE)
+temp_depth_min <- min(presence$temp_depth, na.rm = TRUE)
+temp_depth_mean <- mean(presence$temp_depth, na.rm = TRUE)
+temp_surface_max <- max(presence$temp_surface, na.rm = TRUE)
+temp_surface_min <- min(presence$temp_surface, na.rm = TRUE)
+temp_surface_mean <- mean(presence$temp_surface, na.rm = TRUE)
+```
+
+Salinity
+```{r}
+sal_depth_max <- max(presence$salinity_depth, na.rm = TRUE)
+sal_depth_min <- min(presence$salinity_depth, na.rm = TRUE)
+sal_depth_mean <- mean(presence$salinity_depth, na.rm = TRUE)
+sal_surface_max <- max(presence$salinity_surface, na.rm = TRUE)
+sal_surface_min <- min(presence$salinity_surface, na.rm = TRUE)
+sal_surface_mean <- mean(presence$salinity_surface, na.rm = TRUE)
+```
+
+Chl
+```{r}
+chl_depth_max <- max(presence$chl_depth, na.rm = TRUE)
+chl_depth_min <- min(presence$chl_depth, na.rm = TRUE)
+chl_depth_mean <- mean(presence$chl_depth, na.rm = TRUE)
+chl_surface_max <- max(presence$chl_surface, na.rm = TRUE)
+chl_surface_min <- min(presence$chl_surface, na.rm = TRUE)
+chl_surface_mean <- mean(presence$chl_surface, na.rm = TRUE)
+```
+
+O2
+```{r}
+o2_depth_max <- max(presence$o2_depth, na.rm = TRUE)
+o2_depth_min <- min(presence$o2_depth, na.rm = TRUE)
+o2_depth_mean <- mean(presence$o2_depth, na.rm = TRUE)
+o2_surface_max <- max(presence$o2_surface, na.rm = TRUE)
+o2_surface_min <- min(presence$o2_surface, na.rm = TRUE)
+o2_surface_mean <- mean(presence$o2_surface, na.rm = TRUE)
+```
+
+MLP
+```{r}
+mlp_surface_max <- max(presence$mlp_surface, na.rm = TRUE)
+mlp_surface_min <- min(presence$mlp_surface, na.rm = TRUE)
+mlp_surface_mean <- mean(presence$mlp_surface, na.rm = TRUE)
+```
+
+SSH
+```{r}
+ssh_surface_max <- max(presence$ssh_surface, na.rm = TRUE)
+ssh_surface_min <- min(presence$ssh_surface, na.rm = TRUE)
+ssh_surface_mean <- mean(presence$ssh_surface, na.rm = TRUE)
+```
+
+create matrix
+```{r}
+td <- c(temp_depth_max, temp_depth_min, temp_depth_mean)
+ts <- c(temp_surface_max, temp_surface_min, temp_surface_mean)
+sd <- c(sal_depth_max, sal_depth_min, sal_depth_mean)
+ss <- c(sal_surface_max, sal_surface_min, sal_surface_mean)
+cd <- c(sal_depth_max, sal_depth_min, sal_depth_mean)
+cs <- c(sal_surface_max, sal_surface_min, sal_surface_mean)
+od <- c(o2_depth_max, o2_depth_min, o2_depth_mean)
+os <- c(o2_surface_max, o2_surface_min, o2_surface_mean)
+mlp <- c(mlp_surface_max, mlp_surface_min, mlp_surface_mean)
+ssh <- c(ssh_surface_max, ssh_surface_min, ssh_surface_mean)
+env_stats <- rbind(td, ts, sd, ss, cd, cs, od, os, mlp, ssh)
+row.names(env_stats) <- c("Temp Depth", "Temp Surface", "Salinity Depth", "Salinity Surface", "Chl Depth", "Chl Surface", "Oxygen Depth", "Oxygen Surface", "MLP", "SSH") 
+colnames(env_stats) <- c("Max", "Min", "Mean")
+write.csv(env_stats, "../output/env/env_correlates_basic_stats.csv", row.names = TRUE)
+env_stats
+```
+
+Hmm some potentially suspicious values here in
+- Oxygen Depth min
+- mlp  Max
+and maybe also
+- Salinity Depth min
+- Salinity Surface min
+- CHL depth min
+- CHL depth max
+
+lets just see where these values appear and check the netcdf layers.
+Note I will check these values in cygwin using the cdo operator infov on the netcdfs (to ensure there was no issue with processing in R).
+
+get the year and month these values appeared in
+```{r}
+o2d_check_yr <- subset(presence$year, presence$o2_depth == o2_depth_min)
+o2d_check_mth <- subset(presence$month, presence$o2_depth == o2_depth_min)
+```
+O2 depth min is in 2005_08
+
+And value seems correct
+
+```{r}
+mlp_check_yr <- subset(presence$year, presence$mlp_surface == mlp_surface_max)
+mlp_check_mth <- subset(presence$month, presence$mlp_surface == mlp_surface_max)
+```
+mlp max in in 2007_12
+
+And yes the value seems correct
+
+```{r}
+sald_check_yr <- subset(presence$year, presence$salinity_depth == sal_depth_max)
+sald_check_mth <- subset(presence$month, presence$salinity_depth == sal_depth_max)
+```
+year month 2001_05
+
+also seems correct
+
+```{r}
+sals_check_yr <- subset(presence$year, presence$salinity_surface == sal_surface_min)
+sals_check_mth <- subset(presence$month, presence$salinity_surface == sal_surface_min)
+```
+year month 21998_06
+
+ok again...
+
+Leave the rest
+
+#simple plots of env. correlates
+
+temperature
+```{r}
+par(mfrow=c(1,2))
+plot(presence$temp_depth, main = "Temp at Depth (Various)", ylab = "Kelvin")
+plot(presence$temp_surface, main = "Temp at Surface", ylab = "Kelvin")
+dev.copy(png, "../output/env/simple_plots/temperature_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+Hmm not necessarily very helpful. But anyway, lets carry on
+
+salinity
+```{r}
+par(mfrow=c(1,2))
+plot(presence$salinity_depth, main = "Salinity at Depth (Various)", ylab = "Practical Salinity Units")
+plot(presence$salinity_surface, main = "Salinity at Surface", ylab = "Practical Salinity Units")
+dev.copy(png,"../output/env/simple_plots/salinity_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+Chlorophyll
+```{r}
+par(mfrow=c(1,2))
+plot(presence$chl_depth, main = "Chl at Depth (Various)", ylab = "Chl (mmol.m-3)")
+plot(presence$chl_surface, main = "Chl at Surface", ylab = "Chl (mmol.m-3)")
+dev.copy(png,"../output/env/simple_plots/chl_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+O2
+```{r}
+par(mfrow=c(1,2))
+plot(presence$o2_depth, main = "Dissolved Oxygen at Depth (Various)", ylab = "Chl (mmol.m-3)")
+plot(presence$o2_surface, main = "Dissolved Oxygen at Surface", ylab = "O2 (mmol.m-3)")
+dev.copy(png,"../output/env/simple_plots/o2_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+mlp
+```{r}
+plot(presence$mlp_surface, main = "Mixed Layer Thickness", ylab = "Depth (m)")
+dev.copy(png,"../output/env/simple_plots/mlp_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+SSH
+```{r}
+plot(presence$ssh_surface, main = "Sea Surface Height", ylab = "Height (m)")
+dev.copy(png,"../output/env/simple_plots/ssh_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```

### 2018-11-14 src/data_exploration.Rmd
histogram addition + start of boxplots

@@
 ```
+
+XXX SAM - YOU WANT TO THINK ABOUT DOING THIS BY DEPTH RATHER THAN ALL DEPTHS)
+
+```{r}
+unique_depths <- unique(presence$depthlayerno)
+unique_depthdordered <- sort(unique_depths) #just puts the list in oder with no NA
+length(unique_depthdordered)
+```
+ouch - 40 layers... a job for a boxplot probably
+
+```{r}
+
+```
+
+# frequency plots of env. corr
+
+just do surface layers for now
+
+temperature
+```{r}
+hist(presence$temp_surface, main = "Temp at Surface", xlab = "Kelvin")
+dev.copy(png, "../output/env/simple_plots/temperature_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+# simple boxplots of env. corr
+
+Same as above but with boxplots (may provide some more useful info)
+
+XXX SAM YOU WANT TO PUT THE TWO LAYERS ON THE ONE PLOT - NOT TWO SIDE BY SIDE
+
+temperature
+```{r}
+par(mfrow=c(1,2))
+boxplot(presence$temp_depth, main = "Temp at Depth (Various)", ylab = "Kelvin")
+boxplot(presence$temp_surface, main = "Temp at Surface", ylab = "Kelvin")
+dev.copy(png, "../output/env/simple_plots/temperature_boxplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```

### 2018-11-15 src/timesliced_random_cell_selection.Rmd
creation of minimum convex polygon for constraining background points

@@
 # Package dependencies
-You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
-Note this function checks to see if the packages are installed first.
-Th
-```{r pre-install packages, message=FALSE}
-packages <- c("raster", "dplyr", "data.table", "sp")
+You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163).
+Note this function checks to see if the packages are installed first before loading.
+
+```{r pre-install & load packages, include=FALSE}
+packages <- c("raster", "dplyr", "data.table", "sp", "adehabitatHR")
 source("../src/ipak.R")
 ipak(packages)
-suppressMessages(ipak)
 ```
@@
 
-What I want to do is for each time slice, create a list of all cells (xyz), then remove the ones where observations occuredand create a .csv with the name of the level being worked on
+The area covered by these points are huge -much larger than the distribution of my presence points.
+Ideally I need to constrain the area where the background points are selected from to match the entire range of the species.
+The most common method for this is to use a minimum convex hull to estimate the species range, and then use that area to constrain the background point selection.
+Note I am a little worried that I may loose some of the edge data as they are centroids of cells...
+
+```{r}
+mcpbase <- read.csv("../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env.csv", header = TRUE)
+mcpbase_xy <- mcpbase[ , c("decimalLongitude.1", "decimalLatitude.1")] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+mcpbase_sp <- SpatialPointsDataFrame(coords = mcpbase_xy, data = mcpbase, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+mcp_capelin_100 <- mcp(mcpbase_sp[ , 12], percent = 100) #12 is the column with the species name in (the id for the species)
+plot(mcp_capelin_100)
+```
+
+oh tres exciting. Let's overlay the occurrence points on that
+
+```{r}
+plot(mcp_capelin_100)
+points(mcpbase_sp$decimalLongitude.1, mcpbase_sp$decimalLatitude.1)
+```
+
+ok yea, i guess.... still wondering about those edge areas....anyway
+
+What I want to do is for each time slice, create a list of all cells (xyz), then remove the ones where observations occurred and create a .csv with the name of the level being worked on
 

### 2018-11-15 src/timesliced_random_cell_selection.Rmd
package updates

@@
 
-```{r pre-install & load packages, include=FALSE}
-packages <- c("raster", "dplyr", "data.table", "sp", "adehabitatHR")
-source("../src/ipak.R")
-ipak(packages)
+New package attempt to do this.... #include=FALSE
+
+```{r pre-install & load packages, }
+install.packages("pacman")
+library(pacman)
+p_load(raster, dplyr, data.table, sp, adehabitatHR, maptools)
+
+#
+
+#packages <- c("raster", "dplyr", "data.table", "sp", "adehabitatHR", "maptools")
+#source("../src/ipak.R")
+#ipak(packages)
 ```

### 2018-11-15 src/timesliced_random_cell_selection.Rmd
convex hull done - new data being added

--- 2018-11-15 src/timesliced_random_cell_selection.Rmd
@@
 
-```{r pre-install & load packages, include=FALSE}
-packages <- c("raster", "dplyr", "data.table", "sp", "adehabitatHR")
+New package attempt to do this....
+old code 
+packages <- c("raster", "dplyr", "data.table", "sp", "adehabitatHR", "maptools")
 source("../src/ipak.R")
 ipak(packages)
+
+```{r pre-install & load packages, include=FALSE}
+install.packages("pacman")
+library(pacman)
+p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal)
 ```
@@
 ```{r}
+glo_unique_cell <- raster("../output/env/glo_unique_cell.tif") #this is loading the cell layer that was created and saved as a tif
 plot(glo_unique_cell)
@@
 
-6729...
+7750
 
@@
 ```{r}
-library(raster)
 unique_cell_centroid <- as.data.frame(glo_unique_cell, row.names = NULL, optional = FALSE, xy = TRUE, centroids = TRUE, na.rm = TRUE)
@@
 points(unique_cell_centroid$longitude, unique_cell_centroid$latitude, pch = 10, col = "black")
+write.csv(unique_cell_centroid, "../output/env/unique_cell_points.csv", row.names = FALSE)
 ```
 well... the plot doesn't tell me much... but i definately have points!
-
-ok so now I have a base list to work off... This gives you a dataset of xy points, but you need xyz (where z is based on the depth layers in the environmental data)
-```{r}
-unique_cell_xyz <- unique_cell_centroid
-unique_cell_xyz$depthlayerno <- NA
-unique_cell_xyz <- unique_cell_xyz[rep(seq_len(nrow(unique_cell_xyz)), each=50),] #this generates 50 repetitions of each row
-unique_cell_xyz$depthlayerno <- rep(1:50) #assign 1:50 on a repeating cycle to the depth column. 50 is chosen because there are 50 depth layers
-unique_cell_xyz$id_depth <- paste(unique_cell_xyz$cell_id, unique_cell_xyz$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID.
-This is to help remove cells later and randomise
-```
-
-Split the data_aea into timeslices (year/month).
-All you need is the year, month, cell_id, and depthlayerno
-```{r}
-cell_yymm <- subset(data_aea, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month
-cell_yymm$id_depth <- paste(cell_yymm$cell_id, cell_yymm$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID. This is to help remove cells later and randomise
-split_cell_yymm <- split(cell_yymm, list(cell_yymm$year, cell_yymm$month), drop = TRUE)
-```
 
@@
 
-ok yea, i guess.... still wondering about those edge areas....anyway
+ok yea, i guess.... still wondering about those edge areas....anyway save the hull as a polygon
+
+```{r}
+writeOGR(mcp_capelin_100, dsn = '.', layer = '../output/bio/mcp_capelin_100_poly', driver = "ESRI Shapefile")
+```
+
+now need to select points within a polygon....
+```{r}
+unique_cell_centroid <- read.csv("../output/env/unique_cell_centroid.csv",header = TRUE)
+unique_cell_xy <- unique_cell_centroid[ , c("longitude", "latitude")] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+unique_c_sp <- SpatialPointsDataFrame(coords = unique_cell_xy, data = unique_cell_centroid, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+plot(mcp_capelin_100)
+points(unique_c_sp, pch = "red")
+```
+
+```{r}
+mcp_unique_cell <- unique_c_sp[mcp_capelin_100, ] #this is just subsetting 
+plot(mcp_capelin_100)
+points(mcp_unique_cell)
+```
+ok now convert mcp_unique_cell back to a dataframe for further processing, and save the clipped unique cells to a .csv
+```{r}
+mcp_unique_cell <- as.data.frame(mcp_unique_cell)
+write.csv(mcp_unique_cell, "../output/env/unique_cell_centroid_mcp")
+```
+
+ok so now I have a base list to work off... This gives you a dataset of xy points, but you need xyz (where z is based on the depth layers in the environmental data)
+```{r}
+unique_cell_xyz <- mcp_unique_cell
+unique_cell_xyz$depthlayerno <- NA
+unique_cell_xyz <- unique_cell_xyz[rep(seq_len(nrow(unique_cell_xyz)), each=50),] #this generates 50 repetitions of each row
+unique_cell_xyz$depthlayerno <- rep(1:50) #assign 1:50 on a repeating cycle to the depth column. 50 is chosen because there are 50 depth layers
+unique_cell_xyz$id_depth <- paste(unique_cell_xyz$glo_unique_cell, unique_cell_xyz$depthlayerno, sep="") #create a new column that is a unique glo_unique_cell & depth ID.
+This is to help remove cells later and randomise
+```
 
 What I want to do is for each time slice, create a list of all cells (xyz), then remove the ones where observations occurred and create a .csv with the name of the level being worked on
+
+Split the data_aea into timeslices (year/month).
+All you need is the year, month, cell_id, and depthlayerno
+```{r}
+data_aea <- mcpbase
+cell_yymm <- subset(data_aea, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month
+cell_yymm$id_depth <- paste(cell_yymm$cell_id, cell_yymm$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID. This is to help remove cells later and randomise
+split_cell_yymm <- split(cell_yymm, list(cell_yymm$year, cell_yymm$month), drop = TRUE)
+```
 
@@
 ```{r}
-library(dplyr)
 bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
@@
   sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
-  sel <- sample_n(test_sel, 10000) #where 10000 = number of rows to sample (large sample as per maxent)
+  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample (large sample as per maxent)
   write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
 }
-
+head(sel)  
 ```
@@
 }
+head(randtemp)
 ```
@@
 ```{r}
-library(data.table)
 filenames  <- list.files(path = rand_folder, pattern="*.csv", full.names = TRUE) 
-background_all <- rbindlist(lapply(filenames,fread))
+background_all <- rbindlist(lapply(filenames,fread, fill = TRUE))
 write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
+head(background_all)
 ```
@@
 write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
+head(background_all)
 ```
@@
 ```{r}
-library(sp)
 background_all <- as.data.frame(background_all)
-xy <- background_all[ ,c(1,2)] #lon lat in col 1 and 2
+xy <- background_all[ ,c(2,3)] #lon lat in col 1 and 2
 sp_background <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
@@
 ```{r}
-sp_background <- data.frame(test, long=coordinates(sp_background)[,1], lat=coordinates(sp_background)[,2])
+sp_background <- data.frame(sp_background, long=coordinates(sp_background)[,1], lat=coordinates(sp_background)[,2])
 head(sp_background)
@@
 ```{r NAFO shapefile extraction}
-library(rgdal)
-library(sp)
 coordinates(sp_background) <- c("decimalLongitude", "decimalLatitude") 

### 2018-11-15 src/timesliced_random_cell_selection.Rmd
nafo zone correction underway

@@
 library(pacman)
-p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal)
+p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap)
 ```
@@
 
-change name of long lat columns to decimalLongitude and decimalLatitude respectively
+remove existing decimlalLongitude and decimlalLongitude, and long and lat, and optional (duplicate columns) and change name of longitude.2 & latitude.2 columns to decimalLongitude and decimalLatitude respectively
 ```{r}
 sp_background <- as.data.frame(sp_background) #need to go back to a 'normal' dataframe to change col names
-head(sp_background) #note the columns in question have changed name...
-colnames(sp_background)[colnames(sp_background)=="longitude.1"] <- "decimalLongitude"
-colnames(sp_background)[colnames(sp_background)=="latitude.1"] <- "decimalLatitude"
+head(sp_background) 
+sp_background <- subset(sp_background, select = -c(decimalLongitude, decimalLatitude, long, lat, optional))
+colnames(sp_background)[colnames(sp_background)=="longitude.2"] <- "decimalLongitude"
+colnames(sp_background)[colnames(sp_background)=="latitude.2"] <- "decimalLatitude"
 head(sp_background)
@@
 ```{r}
-plot(map2, xlim = c(-65, -40), ylim =c(30, 70), asp = 1, main = "Occurences missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-65, -40), ylim =c(30, 70), asp = 1, main = "Background missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
 points(bc_nafo_na$decimalLongitude, bc_nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/background_missing_nafo.png") #this prints a png of the plot
+dev.off() #this turns off the print comman
 ```
 
-huh that is interesting... I think I want to plot this in Arc and take a closer look...
+Indeed the many of the missing background points are in the hudson strait.
+Others are near the coast and thus probably just call outside the NAFO shapefiles (remember the smapefiles are not all super-perfectly fitting).
+Think I will take this to arcgis for fixing....

### 2018-11-15 src/biological_data_prep.Rmd
removing some test scripts i no longer need, plus data exploration pt 2
moving on to NAO and AMO

@@
 So you want..
-- AMO value at sampling month/year
-- AMO value at previous sampling month/year
-- AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
+- NAO value at sampling month/year
+- NAO value at previous sampling month/year
+- NAO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
 

### 2018-11-15 src/timesliced_random_cell_selection.Rmd
moving on to NAO and AMO

@@
 Think I will take this to arcgis for fixing....
+
+basically what I did was use the selection -> select by attributes tool to find the points where NAFO_region = NA and then manually recode them (using the field calculator), and exported as a .csv.
+
+```{r}
+background_all <- read.csv("../output/bio/background_all_lonlat_nafo2.csv", header = TRUE)
+head(background_all)
+```
+
+#original scientific name
+
+create a column called "originalscientificname" and fill it with "Mallotus villosus ab"
+
+```{r}
+background_all$originalscientificname <- "Mallotus villosus ab"
+head(background_all)
+```
+
+#NAO and AMO values

### 2018-11-16 src/biological_data_prep.Rmd
nao correct
amo done

@@
 amo_yrmt
+write.csv(amo_yrmt, "../output/env/amo_yrmt.csv") # written for reuse with the background points 
 ```
@@
 
-nao_yrmt <- nao_yrmt[order(amo_yrmt$year),]
+nao_yrmt <- nao_yrmt[order(nao_yrmt$year),] #changed amo to nao
 nao_yrmt

### 2018-11-16 src/env_data_prep.Rmd
package updates
NAO correction :(

@@
 ```{r}
-naoraw <- read.csv("../data/env/amon.us.data.csv", header = FALSE) 
+naoraw <- read.csv("../data/env/nao_pc_monthly.csv", header = FALSE) 
 head(naoraw)
@@
 
-And now create a raster of Gebco in aea for mapping once the SDM is done...
+And now create a raster of Gebco in aea for mapping once the SDM is done..
 

### 2018-11-16 src/biological_data_prep.Rmd
NAO correction :(
nao added to background

@@
 
-nao_yrmt <- nao_yrmt[order(nao_yrmt$year),] #changed amo to nao
+nao_yrmt <- nao_yrmt[order(nao_yrmt$year),] 
 nao_yrmt
@@
 data_aea <- merge(x = data_aea, y = nao_yrmt[ , c("year", "month", "nao_sample")], by = c("year", "month"), all.x = TRUE)
-data_aea
+head(data_aea)
 ```
@@
 
-Note the shifted values are in a new column called amo_sample.Shifted. rename to amo_prev
+Note the shifted values are in a new column called amo_sample.Shifted. rename to nao_prev
 
@@
 
-now extract the WinterAvg.Shifted values to the species dataset
+now extract the WinterAvg.Shifted values to the species dataset #####XXXXXXX#####
 

### 2018-11-16 src/timesliced_random_cell_selection.Rmd
nao added to background

@@
 
+check for unique years
+```{r}
+back_yr_uniq <- unique(background_all$year)
+back_yr_uniq
+```
+check for unique months
+```{r}
+back_mt_uniq <- unique(background_all$month)
+back_mt_uniq
+```
+
+check number of observations 
+```{r}
+back_rowcount <- nrow(background_all)
+back_rowcount
+```
+
 get the columns from the occurence data so you know what to add
@@
 
+just double-check they are all gone
 ```{r}
 background_all <- read.csv("../output/bio/background_all_lonlat_nafo2.csv", header = TRUE)
+bc_nafo2_na <- subset(background_all, is.na(background_all$nafo_zone))
+bc_nafo2_na
+```
+great none!
+
+```{r}
+head(background_all)
+```
+ok so some more crap has been added - the year and month both have _ appended.... change this. Also change glo_unique_cell to cell_id
+
+```{r}
+names(background_all)[names(background_all)=="year_"] <- "year"
+names(background_all)[names(background_all)=="month_"] <- "month"
+names(background_all)[names(background_all)=="glo_unique_cell"] <- "cell_id"
 head(background_all)
@@
 background_all$originalscientificname <- "Mallotus villosus ab"
+write.csv(background_all, "../output/bio/background_all_lonlat_nafo2_sciname.csv")
 head(background_all)
@@
 
-#NAO and AMO values
+# Add AMO Data
+The data has been prepared as per the environmental_data_preperation.Rmd, and biological_data_prep.Rmd
+
+So you want..
+- AMO value at sampling month/year
+- AMO value at previous sampling month/year
+- AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
+
+Load the prepared amo_prev.csv file with has the amo at different year/month steps
+```{r}
+amo_prev <- read.csv("../output/env/amo_prev.csv", header = TRUE)
+head(amo_prev)
+```
+
+now match the year and the month in the two dataframes, and populate the background points with amo_sample
+```{r}
+background_all <- merge(x = background_all, y = amo_prev[ , c("year", "month", "amo_sample")], by = c("year", "month"), all.x = TRUE)
+head(background_all)
+```
+
+ok next - AMO value at previous sampling month/year 
+```{r}
+background_all <- merge(x = background_all, y = amo_prev[ , c("year", "month", "amo_prev")], by = c("year", "month"), all.x = TRUE)
+head(background_all)
+```
+
+and now - AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn) (note they are in a different .csv)
+
+```{r}
+amo_winter <- read.csv("../output/env/amo_winter.csv")
+background_all <- merge(x = background_all, y = amo_winter[ , c("year", "WinterAvg.Shifted")], by = c("year"), all.x = TRUE)
+colnames(background_all)[colnames(background_all)=="WinterAvg.Shifted"] <- "amo_winter" #change the colname
+head(background_all)
+```
+
+ok and write a csv
+```{r}
+write.csv(data_aea,"../output/bio/background_all_lonlat_nafo2_sciname_amo.csv", row.names = FALSE)
+```
+
+# NAO values
+The data has been prepared as per the environmental_data_preperation.Rmd and biological_data_prep.Rmd
+
+So you want..
+- NAO value at sampling month/year
+- NAO value at previous sampling month/year
+- NAO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
+
+Load the prepared nao_prev.csv file with has the amo at different year/month steps
+```{r}
+nao_prev <- read.csv("../output/env/nao_prev.csv", header = TRUE)
+head(nao_prev)
+```
+
+now match the year and the month in the two dataframes, and populate the background points with nao_sample
+
+```{r}
+background_all <- merge(x = background_all, y = nao_prev[ , c("year", "month", "nao_sample")], by = c("year", "month"), all.x = TRUE)
+background_all
+```
+
+ok next - NAO value at previous sampling month/year
+```{r}
+background_all <- merge(x = background_all, y = nao_prev[ , c("year", "month", "nao_prev")], by = c("year", "month"), all.x = TRUE)
+background_all
+```
+
+and now - NAO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn). The data is in a different .CSV
+
+```{r}
+nao_winter <- read.csv("../output/env/nao_winter.csv")
+background_all <- merge(x = background_all, y = nao_winter[ , c("year", "WinterAvg.Shifted")], by = c("year"), all.x = TRUE)
+colnames(background_all)[colnames(background_all)=="WinterAvg.Shifted"] <- "nao_winter" #change the colname
+background_all
+```
+
+and write .csv
+```{r}
+write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao.csv", row.names = FALSE)
+```

### 2018-11-19 src/biological_data_prep.Rmd
background depth layer name

@@
 
+rename the lat/long meters columns and remove the duplicate column
+
+```{r}
+names(data_aea)[names(data_aea)=="decimalLongitude.1"] <- "longitude_meters"
+names(data_aea)[names(data_aea)=="decimalLatitude.1"] <- "latitude_meters"
+data_aea <- subset(data_aea, select = -c(decimalLongitude.1.1, decimalLatitude.1.1))
+head(data_aea)
+```
+
 # warm/cold year

### 2018-11-19 src/timesliced_random_cell_selection.Rmd
background depth layer name

@@
 ```
-ok so some more crap has been added - the year and month both have _ appended.... change this. Also change glo_unique_cell to cell_id
+ok so some more crap has been added - the year and month both have _ appended.... change this. Also change 
+- glo_unique_cell to cell_id
+- longitude to longitude_meters
+- latitude to latitude_meters
+
+and remove
+- optional
+- x.1
+- x
+- id_depth
 
@@
 names(background_all)[names(background_all)=="glo_unique_cell"] <- "cell_id"
+names(background_all)[names(background_all)=="longitude"] <- "longitude_meters"
+names(background_all)[names(background_all)=="latitude"] <- "latitude_meters"
+background_all <- subset(background_all, select = -c(optional, X.1, X, id_depth))
 head(background_all)
@@
 ```
+
+still need
+-bottom_depth
+-depth_layer
+-total_cell_obs
+-yymm_cell_obs
+
+#depth layer

### 2018-11-19 src/env_data_prep.Rmd
background bathy data

@@
 ```{r}
-data_aea$bottom_depth <- get.depth(bat, data_aea[ ,5:4], locator = FALSE)
+data_aea$bottom_depth <- get.depth(bat, data_aea[ ,5:4], locator = FALSE) #5 = decimalLongitude, 4 = decimalLatitude
 data_aea$bottom_depth <- data_aea$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth.

### 2018-11-19 src/timesliced_random_cell_selection.Rmd
background bathy data

@@
 library(pacman)
-p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap)
+p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap, ncdf4, marmap)
 ```
@@
 #depth layer
+
+the depth layers in the netCDF are literal slices.
+Currently I have listed the depthlayerno (which is literally which layer in the netcdf).
+I should also add the depth_layer name which gives the depth the slice is supposed to represent.
+Data is stored in depth_bin.csv
+
+load the depth bin csv
+```{r}
+depthbin <- read.csv("../data/env/depth_bin.csv", header = TRUE) 
+head(depthbin)
+```
+
+so basically what i want to do is, where depthbin$layer_no = background_all$depthlayerno, extract depthbin$layer_name and add to background_all$depth_layer. Can use the merge function.
+First need to change depthbin col names to match background_all
+
+```{r}
+names(depthbin)[names(depthbin)=="layer_no"] <- "depthlayerno"
+names(depthbin)[names(depthbin)=="layer_name"] <- "depth_layer" #this one is just so i don't have to change the column name in the background_all file later
+head(depthbin)
+```
+```{r}
+background_all <- merge(x = background_all, y = depthbin[ , c("depthlayerno", "depth_layer")], by.x = ("depthlayerno"), all.x = TRUE)
+head(background_all)
+```
+
+just check some random rows...
+
+```{r}
+randcheck <- background_all[sample(nrow(background_all), 10), ]
+randcheck
+```
+all good - save to .csv
+
+```{r}
+write.csv(background_all, "../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname.csv", row.names = FALSE)
+```
+
+# Bathymetric
+
+bottom depth from [GEBCO](www.gebco.net)
+
+```{r}
+bat <-readGEBCO.bathy("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
+summary(bat)
+```
+
+create pretty custom color palette and then plot the bathy laye
+```{r}
+blues <- colorRampPalette(c("lightblue", "cadetblue2", "cadetblue1", "white"))
+plot(bat, n = 1, image = TRUE, bpal = blues(100))
+```
+lovely jubbely
+
+add the species data...
+```{r}
+plot(bat, n = 1, image = TRUE, bpal = blues(100))
+points(background_all$decimalLongitude, background_all$decimalLatitude, pch = 20, col = "yellow") #note these are both the wgs84 images
+```
+and now get depth... note all the points and bathy data is wgs!
+
+```{r}
+background_all$bottom_depth <- get.depth(bat, background_all[ ,10:11], locator = FALSE)
+background_all$bottom_depth <- background_all$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth.
+This is extracting the depth column and deplacing data_aea$bottom_depth with the correct data.
+write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath.csv", row.names = FALSE)
+head(background_all)
+```
+
+# observations by cell (in total, and also by time step)
+
+as the title says... how many observations happeend in each cell:
+- in total
+- in each yyyymm
+
+```{r observation by cell - total}
+library(plyr) #to use the count function
+obs_by_cell <- count(data_aea, "cell_id")
+obs_by_cell
+write.csv(obs_by_cell, file = "../output/bio/no_observations_cell.csv")
+```
+
+```{r}
+obs_cell_yymm <- count(data_aea, c("cell_id", "year", "month"))
+obs_cell_yymm
+write.csv(obs_cell_yymm, file = "../output/bio/no_observations_cell_yymm.csv")
+```
+Now attach these values to data_aea....
+
+total obs for the whole time period
+```{r}
+data_aea <- merge(data_aea, obs_by_cell, by="cell_id")
+head(data_aea)
+```
+rename freq column to something meaningful
+
+```{r}
+names(data_aea)[names(data_aea)=="freq"] <- "total_cell_obs"
+head(data_aea)
+```
+
+time-sliced
+```{r}
+data_aea <- merge(data_aea, obs_cell_yymm, by.x=c("cell_id", "year", "month"), by.y=c("cell_id", "year", "month"))
+names(data_aea)[names(data_aea)=="freq"] <- "yymm_cell_obs"
+head(data_aea)
+```
+
+write to .csv
+```{r}
+write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno.csv", row.names = FALSE)
+```
+`

### 2018-11-19 src/timesliced_random_cell_selection.Rmd
obs_by_cell inc depth for background

@@
 library(pacman)
-p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap, ncdf4, marmap)
+p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap, ncdf4, marmap, plyr)
 ```
@@
 ```{r observation by cell - total}
-library(plyr) #to use the count function
-obs_by_cell <- count(data_aea, "cell_id")
-obs_by_cell
-write.csv(obs_by_cell, file = "../output/bio/no_observations_cell.csv")
+obs_by_cell_back <- count(background_all, "cell_id")
+write.csv(obs_by_cell_back, file = "../output/bio/back_no_observations_cell.csv")
+head(obs_by_cell_back)
 ```
@@
 ```{r}
-obs_cell_yymm <- count(data_aea, c("cell_id", "year", "month"))
-obs_cell_yymm
-write.csv(obs_cell_yymm, file = "../output/bio/no_observations_cell_yymm.csv")
+obs_cell_yymm_back <- count(background_all, c("cell_id", "year", "month"))
+write.csv(obs_cell_yymm_back, file = "../output/bio/back_no_observations_cell_yymm.csv")
+head(obs_cell_yymm_back)
 ```
+
+```{r}
+obs_cell_yymm_depth_back <- count(background_all, c("cell_id", "year", "month", "depthlayerno"))
+write.csv(obs_cell_yymm_depth_back, file = "../output/bio/back_no_observations_cell_yymm_depth.csv")
+head(obs_cell_yymm_depth_back)
+```
+
 Now attach these values to data_aea....
@@
 ```{r}
-data_aea <- merge(data_aea, obs_by_cell, by="cell_id")
-head(data_aea)
+background_all <- merge(background_all, obs_by_cell_back, by="cell_id")
+head(background_all)
 ```
+
 rename freq column to something meaningful
+```{r}
+names(background_all)[names(background_all)=="freq"] <- "total_cell_obs"
+head(background_all)
+```
+
+just check some random rows...
 
 ```{r}
-names(data_aea)[names(data_aea)=="freq"] <- "total_cell_obs"
-head(data_aea)
+randcheckcell <- background_all[sample(nrow(background_all), 10), ]
+randcheckcell
 ```
+all good
 
@@
 ```{r}
-data_aea <- merge(data_aea, obs_cell_yymm, by.x=c("cell_id", "year", "month"), by.y=c("cell_id", "year", "month"))
-names(data_aea)[names(data_aea)=="freq"] <- "yymm_cell_obs"
-head(data_aea)
+background_all <- merge(background_all, obs_cell_yymm_back, by.x=c("cell_id", "year", "month"), by.y=c("cell_id", "year", "month"))
+background_all <- merge(background_all, obs_cell_yymm_depth_back, by.x=c("cell_id", "year", "month", "depthlayerno"), by.y=c("cell_id", "year", "month", "depthlayerno"))
+names(background_all)[names(background_all)=="freq"] <- "yymm_cell_obs"
+names(background_all)[names(background_all)=="freq"] <- "yymmdepth_cell_obs"
+write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath_cellobs.csv", row.names = FALSE)
+head(background_all)
 ```
 
-write to .csv
+and just check some random rows...
+
 ```{r}
-write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno.csv", row.names = FALSE)
+randcheckyymm <- background_all[sample(nrow(background_all), 10), ]
+randcheckyymm
 ```
-`

### 2018-11-20 src/timesliced_random_cell_selection.Rmd
very painful depth correction...

--- 2018-11-19 src/timesliced_random_cell_selection.Rmd
@@
 library(pacman)
-p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap)
+p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap, ncdf4, marmap, plyr)
 ```
@@
 ```
-ok now convert mcp_unique_cell back to a dataframe for further processing, and save the clipped unique cells to a .csv
+ok now save the clipped unique cells to a .csv
 ```{r}
-mcp_unique_cell <- as.data.frame(mcp_unique_cell)
 write.csv(mcp_unique_cell, "../output/env/unique_cell_centroid_mcp")
@@
 
-ok so now I have a base list to work off... This gives you a dataset of xy points, but you need xyz (where z is based on the depth layers in the environmental data)
+#decimalLat and Lon
+ok lets's get the real decimalLatitude and decimalLongitude.
+Will need to plot all point, extract the lonlat coordinates
+
+reproject to a wgs (use an existing file as a mask)
+
 ```{r}
+sp_mcp_unique_cell <- spTransform(mcp_unique_cell, CRS = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
+plot(sp_mcp_unique_cell)
+```
+extract lonlat, add to dataframe
+
+```{r}
+sp_mcp_unique_cell <- data.frame(sp_mcp_unique_cell, long=coordinates(sp_mcp_unique_cell)[,1], lat=coordinates(sp_mcp_unique_cell)[,2])
+head(sp_mcp_unique_cell)
+```
+
+remove existing long, lat, optional, longitude.1, latitude.1 (duplicate columns) and change name of longitude.2 & latitude.2 columns to decimalLongitude and decimalLatitude respectively, and longitude and latitude to longitude_meters and latitude_meters ande glo_unique_cell to cell_id
+```{r}
+sp_mcp_unique_cell <- as.data.frame(sp_mcp_unique_cell) #need to go back to a 'normal' dataframe to change col names
+head(sp_mcp_unique_cell) 
+sp_mcp_unique_cell <- subset(sp_mcp_unique_cell, select = -c(long, lat, optional, longitude.1, latitude.1))
+colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="longitude.2"] <- "decimalLongitude"
+colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="latitude.2"] <- "decimalLatitude"
+colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="longitude"] <- "longitude_meters"
+colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="latitude"] <- "latitude_meters"
+colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="glo_unique_cell"] <- "cell_id"
+head(sp_mcp_unique_cell)
+```
+
+save as csv
+```{r}
+write.csv(sp_background, file = "../output/bio/unique_cell_centroid_lonlat.csv", row.names = FALSE)
+```
+
+# Add NAFO Region background points in
+NAFO Zones shapefile obtained from [NAFO](https://www.nafo.int/Data/GIS)
+
+```{r NAFO shapefile extraction}
+coordinates(sp_mcp_unique_cell) <- c("decimalLongitude", "decimalLatitude") 
+nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
+proj4string(sp_mcp_unique_cell) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
+sp_mcp_unique_cell$nafo_zone <- over(sp_mcp_unique_cell, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
+head(sp_mcp_unique_cell) #just to check it worked
+```
+
+Great. Lets see if any are missing a NAFO Zone. But first, write to a csv and reload as the data is now some wierd 'spatial points dataframe'...
+```{r}
+write.csv(sp_mcp_unique_cell, "../output/bio/unique_cell_centroid_lonlat_nafo.csv", row.names=FALSE)
+sp_mcp_unique_cell <- as.data.frame(sp_mcp_unique_cell) 
+bc_nafo_na <- subset(sp_mcp_unique_cell, is.na(sp_mcp_unique_cell$nafo_zone))
+bc_nafo_na
+```
+
+ok lots don't have a NAFO region attached. This is probably the Hudson Bay issue again...
+
+Uh oh there is a lot. A bet a bunch of them are in the Hudson Strait (which is outside the NAFO Zone).Map the nafo_na points...
+```{r}
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-65, -40), ylim =c(30, 70), asp = 1, main = "Background missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(bc_nafo_na$decimalLongitude, bc_nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/background_missing_nafo.png") #this prints a png of the plot
+dev.off() #this turns off the print comman
+```
+
+Indeed the many of the missing background points are in the hudson strait.
+Others are near the coast and thus probably just call outside the NAFO shapefiles (remember the smapefiles are not all super-perfectly fitting).
+Think I will take this to arcgis for fixing....
+
+basically what I did was use the selection -> select by attributes tool to find the points where NAFO_region = NA and then manually recode them (using the field calculator), and exported as a .csv.
+
+just double-check they are all gone
+```{r}
+mcp_unique_cell <- read.csv("../output/env/unique_cell_centroid_lonlat_nafo2.csv", header = TRUE)
+bc_nafo2_na <- subset(mcp_unique_cell, is.na(mcp_unique_cell$nafo_zone))
+bc_nafo2_na
+```
+
+# random depth
+
+ok so now I have a base list to work off...
+This gives you a dataset of xy points, but you need xyz (where z is based on the depth layers in the environmental data).
+This should be constrained by the presence depth range
+```{r}
+max_obs_depth <- max(mcpbase$depthlayerno, na.rm = TRUE) #max is 47
+min_obs_depth <- min(mcpbase$depthlayerno, na.rm = TRUE) #min is 1
 unique_cell_xyz <- mcp_unique_cell
 unique_cell_xyz$depthlayerno <- NA
-unique_cell_xyz <- unique_cell_xyz[rep(seq_len(nrow(unique_cell_xyz)), each=50),] #this generates 50 repetitions of each row
-unique_cell_xyz$depthlayerno <- rep(1:50) #assign 1:50 on a repeating cycle to the depth column. 50 is chosen because there are 50 depth layers
-unique_cell_xyz$id_depth <- paste(unique_cell_xyz$glo_unique_cell, unique_cell_xyz$depthlayerno, sep="") #create a new column that is a unique glo_unique_cell & depth ID.
+unique_cell_xyz <- unique_cell_xyz[rep(seq_len(nrow(unique_cell_xyz)), each=47),] #this generates 47  repetitions of each row 47 is chosen because there are 47 sampled depth layers
+unique_cell_xyz$depthlayerno <- rep(1:47) #assign 1:47 on a repeating cycle to the depth column. 47 is chosen because there are 50 depth layers
+unique_cell_xyz$id_depth <- paste(unique_cell_xyz$cell_id, unique_cell_xyz$depthlayerno, sep="") #create a new column that is a unique glo_unique_cell & depth ID.
 This is to help remove cells later and randomise
@@
 data_aea <- mcpbase
-cell_yymm <- subset(data_aea, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month
+cell_yymm <- subset(data_aea, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month, depthlayer no
 cell_yymm$id_depth <- paste(cell_yymm$cell_id, cell_yymm$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID. This is to help remove cells later and randomise
@@
 
+ok lets start by getting rid of some of the duplicated columns right now (put in by arc, various processes..)
+
+```{r}
+head(background_all)
+```
+remove
+FID
+optional
+
+rename
+longitude_ to longitude_meters
+latitude_m to latitude_meters
+decimalLon to decimalLongitude
+decimalLat to decimalLatitude
+
+```{r}
+background_all <- subset(background_all, select = -c(optional, FID))
+names(background_all)[names(background_all)=="longitude_"] <- "longitude_meters"
+names(background_all)[names(background_all)=="latitude_m"] <- "latitude_meters"
+names(background_all)[names(background_all)=="decimalLon"] <- "decimalLongitude"
+names(background_all)[names(background_all)=="decimalLat"] <- "decimalLatitude"
+head(background_all)
+```
+
+# what data to add
+
 get the columns from the occurence data so you know what to add
@@
 
+# create unique observation ID
+
 Easy first one - add an id based on id_depth
 ```{r}
-background_all$id <- background_all$id_depth
+names(background_all)[names(background_all)=="id_depth"] <- "id"
 write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
-head(background_all)
-```
-
-ok lets's get the real decimalLatitude and decimalLongitude.
-Will need to plot all point, extract the lonlat coordinates
-
-1) load the file as a spatialdatapointframe. give it the aea coordinate
-
-```{r}
-background_all <- as.data.frame(background_all)
-xy <- background_all[ ,c(2,3)] #lon lat in col 1 and 2
-sp_background <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
-head(sp_background)
-```
-
-```{r}
-plot(sp_background)
-```
-
-2) reproject to a wgs (use an existing file as a mask)
-
-```{r}
-sp_background <- spTransform(sp_background, CRS = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
-plot(sp_background)
-```
-3) extract lonlat, add to dataframe
-
-```{r}
-sp_background <- data.frame(sp_background, long=coordinates(sp_background)[,1], lat=coordinates(sp_background)[,2])
-head(sp_background)
-```
-
-remove existing decimlalLongitude and decimlalLongitude, and long and lat, and optional (duplicate columns) and change name of longitude.2 & latitude.2 columns to decimalLongitude and decimalLatitude respectively
-```{r}
-sp_background <- as.data.frame(sp_background) #need to go back to a 'normal' dataframe to change col names
-head(sp_background) 
-sp_background <- subset(sp_background, select = -c(decimalLongitude, decimalLatitude, long, lat, optional))
-colnames(sp_background)[colnames(sp_background)=="longitude.2"] <- "decimalLongitude"
-colnames(sp_background)[colnames(sp_background)=="latitude.2"] <- "decimalLatitude"
-head(sp_background)
-```
-
-save as csv
-```{r}
-write.csv(sp_background, file = "../output/bio/background_all_lonlat.csv", row.names = FALSE)
-```
-
-# Add NAFO Region background point is in
-NAFO Zones shapefile obtained from [NAFO](https://www.nafo.int/Data/GIS)
-
-```{r NAFO shapefile extraction}
-coordinates(sp_background) <- c("decimalLongitude", "decimalLatitude") 
-nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
-proj4string(sp_background) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
-sp_background$nafo_zone <- over(sp_background, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
-head(sp_background) #just to check it worked
-```
-
-Great. Lets see if any are missing a NAFO Zone. But first, write to a csv and reload as the data is now some wierd 'spatial points dataframe'...
-```{r}
-write.csv(sp_background, "../output/bio/background_all_lonlat_nafo.csv", row.names=FALSE)
-sp_background <- as.data.frame(sp_background) 
-bc_nafo_na <- subset(sp_background, is.na(sp_background$nafo_zone))
-bc_nafo_na
-```
-ok lots don't have a NAFO region attached. This is probably the Hudson Bay issue again...
-
-Uh oh there is a lot. A bet a bunch of them are in the Hudson Strait (which is outside the NAFO Zone).Map the nafo_na points...
-```{r}
-map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-65, -40), ylim =c(30, 70), asp = 1, main = "Background missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
-points(bc_nafo_na$decimalLongitude, bc_nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
-dev.copy(png, "../output/bio/background_missing_nafo.png") #this prints a png of the plot
-dev.off() #this turns off the print comman
-```
-
-Indeed the many of the missing background points are in the hudson strait.
-Others are near the coast and thus probably just call outside the NAFO shapefiles (remember the smapefiles are not all super-perfectly fitting).
-Think I will take this to arcgis for fixing....
-
-basically what I did was use the selection -> select by attributes tool to find the points where NAFO_region = NA and then manually recode them (using the field calculator), and exported as a .csv.
-
-just double-check they are all gone
-```{r}
-background_all <- read.csv("../output/bio/background_all_lonlat_nafo2.csv", header = TRUE)
-bc_nafo2_na <- subset(background_all, is.na(background_all$nafo_zone))
-bc_nafo2_na
-```
-great none!
-
-```{r}
-head(background_all)
-```
-ok so some more crap has been added - the year and month both have _ appended.... change this. Also change 
-- glo_unique_cell to cell_id
-- longitude to longitude_meters
-- latitude to latitude_meters
-
-and remove
-- optional
-- x.1
-- x
-- id_depth
-
-```{r}
-names(background_all)[names(background_all)=="year_"] <- "year"
-names(background_all)[names(background_all)=="month_"] <- "month"
-names(background_all)[names(background_all)=="glo_unique_cell"] <- "cell_id"
-names(background_all)[names(background_all)=="longitude"] <- "longitude_meters"
-names(background_all)[names(background_all)=="latitude"] <- "latitude_meters"
-background_all <- subset(background_all, select = -c(optional, X.1, X, id_depth))
 head(background_all)
@@
 #depth layer
+
+the depth layers in the netCDF are literal slices.
+Currently I have listed the depthlayerno (which is literally which layer in the netcdf).
+I should also add the depth_layer name which gives the depth the slice is supposed to represent.
+Data is stored in depth_bin.csv
+
+load the depth bin csv
+```{r}
+depthbin <- read.csv("../data/env/depth_bin.csv", header = TRUE) 
+head(depthbin)
+```
+
+so basically what i want to do is, where depthbin$layer_no = background_all$depthlayerno, extract depthbin$layer_name and add to background_all$depth_layer. Can use the merge function.
+First need to change depthbin col names to match background_all
+
+```{r}
+names(depthbin)[names(depthbin)=="layer_no"] <- "depthlayerno"
+names(depthbin)[names(depthbin)=="layer_name"] <- "depth_layer" #this one is just so i don't have to change the column name in the background_all file later
+head(depthbin)
+```
+```{r}
+background_all <- merge(background_all, y = depthbin[ , c("depthlayerno", "depth_layer")], by = "depthlayerno", all.x = TRUE)
+head(background_all)
+```
+
+just check some random rows...
+
+```{r}
+randcheck <- background_all[sample(nrow(background_all), 10), ]
+randcheck
+```
+all good - save to .csv
+
+```{r}
+write.csv(background_all, "../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname.csv", row.names = FALSE)
+```
+
+# Bathymetric
+
+bottom depth from [GEBCO](www.gebco.net)
+
+```{r}
+bat <-readGEBCO.bathy("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
+summary(bat)
+```
+
+create pretty custom color palette and then plot the bathy layer with species data
+```{r}
+blues <- colorRampPalette(c("lightblue", "cadetblue2", "cadetblue1", "white"))
+plot(bat, n = 1, image = TRUE, bpal = blues(100))
+points(background_all$decimalLongitude, background_all$decimalLatitude, pch = 20, col = "yellow") #note these are both the wgs84 images
+```
+lovely jubbely
+
+and now get depth... note all the points and bathy data is wgs!
+
+```{r}
+background_all$bottom_depth <- get.depth(bat, background_all[ ,8:9], locator = FALSE)
+background_all$bottom_depth <- background_all$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth.
+This is extracting the depth column and deplacing data_aea$bottom_depth with the correct data.
+write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath.csv", row.names = FALSE)
+head(background_all)
+```
+
+# observations by cell (in total, and also by time step)
+
+as the title says... how many observations happeend in each cell:
+- in total
+
+```{r observation by cell - total}
+obs_by_cell_back <- count(background_all, "cell_id")
+write.csv(obs_by_cell_back, file = "../output/bio/back_no_observations_cell.csv")
+head(obs_by_cell_back)
+```
+
+- in each yyyymm
+```{r}
+obs_cell_yymm_back <- count(background_all, c("cell_id", "year", "month"))
+write.csv(obs_cell_yymm_back, file = "../output/bio/back_no_observations_cell_yymm.csv")
+head(obs_cell_yymm_back)
+```
+
+- in each yyyymm at depth
+```{r}
+obs_cell_yymm_depth_back <- count(background_all, c("cell_id", "year", "month", "depthlayerno"))
+write.csv(obs_cell_yymm_depth_back, file = "../output/bio/back_no_observations_cell_yymm_depth.csv")
+head(obs_cell_yymm_depth_back)
+```
+
+Now attach these values
+
+total obs for the whole time period
+```{r}
+background_all <- merge(background_all, obs_by_cell_back, by="cell_id")
+head(background_all)
+```
+
+rename freq column to something meaningful
+```{r}
+names(background_all)[names(background_all)=="freq"] <- "total_cell_obs_xy"
+head(background_all)
+```
+
+just check some random rows...
+
+```{r}
+randcheckcell <- background_all[sample(nrow(background_all), 10), ]
+randcheckcell
+```
+all good
+
+time-sliced
+```{r}
+background_all <- merge(background_all, obs_cell_yymm_back, by.x=c("cell_id", "year", "month"), by.y=c("cell_id", "year", "month"))
+names(background_all)[names(background_all)=="freq"] <- "total_cell_obs_xyt"
+background_all <- merge(background_all, obs_cell_yymm_depth_back, by.x=c("cell_id", "year", "month", "depthlayerno"), by.y=c("cell_id", "year", "month", "depthlayerno"))
+names(background_all)[names(background_all)=="freq"] <- "total_cell_obs_xyzt"
+write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath_cellobs.csv", row.names = FALSE)
+head(background_all)
+```
+
+and just check some random rows...
+
+```{r}
+randcheckyymm <- background_all[sample(nrow(background_all), 10), ]
+randcheckyymm
+```
+
+# attach environmental correlates
+
+This loop runs through the netcdf files and then looks for which rows in data_aea it should extract the value to point from, and at what depth (netcDF layer)
+
+First you need to put the background (background_all) into a spatialpointsdataframe.
+```{r}
+background_all2 <- subset(background_all[1:6, ])
+xy <- background_all2[ ,c(7,8)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+head(background_all_sp)
+```
+
+```{r}
+netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+aea <- raster("../output/env/aea.tif") 
+for (i in 1:no_netcdf) {  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(background_all_sp)) {  
+      de <- background_all_sp$depthlayerno[[j]]  # a variable for the observation depth layer
+      yr <- (background_all_sp$year[j])  # a variable for the observation year
+      mth <- (background_all_sp$month[j])  # a variable for the observation month
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+              background_all_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
+              if (is.na(de)){
+                background_all_sp$temp_depth[j] <- NA
+              } else  
+                background_all_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              background_all_sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
+              if (is.na(de)){
+                background_all_sp$salinity_depth[j] <- NA
+              } else  
+                background_all_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
+              background_all_sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
+              if (is.na(de)){
+                background_all_sp$chl_depth[j] <- NA
+              } else  
+                background_all_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
+              background_all_sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
+              if (is.na(de)){
+                background_all_sp$o2_depth[j] <- NA
+              } else  
+                background_all_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
+              background_all_sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
+              background_all_sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ])
+
+          }
+
+    }
+}
+head(background_all_sp)
+```
+
+and just check some random rows...
+
+```{r}
+randcheck <- background_all[sample(nrow(background_all), 5), ]
+randcheck
+```

### 2018-11-23 src/timesliced_random_cell_selection.Rmd
i thin the backgrond env correlates are in.... 4 days to run loop

@@
 ```{r}
-background_all2 <- subset(background_all[1:6, ])
-xy <- background_all2[ ,c(7,8)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
-background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+xy <- background_all[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
 head(background_all_sp)
@@
 }
+write.csv(background_all, "../output/bio/background_complete.csv", row.names = FALSE)
 head(background_all_sp)

### 2018-11-23 src/timesliced_random_cell_selection.Rmd
check on env. correlate extraction (ca marche!)

@@
 }
-write.csv(background_all, "../output/bio/background_complete.csv", row.names = FALSE)
+write.csv(background_all_sp, "../output/bio/background_complete.csv", row.names = FALSE)
+background_all <- as.data.frame(background_all_sp)
 head(background_all_sp)
@@
 ```
+
+```{r}
+randchkxy <- randcheck[ , c("longitude_meters", "latitude_meters")] 
+randchsp <- SpatialPointsDataFrame(coords = randchkxy, data = randcheck, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+
+#row 1
+temp_brick <-  brick("../data/env/netcdf/2002_03_chl.nc", lvar = 4)
+t200203chl <- projectRaster(temp_brick, aea)
+chls <- extract(x=t200203chl[[1]], y = randchsp[1, ]) 
+chls
+chld <- extract(x=t200203chl[[23]], y = randchsp[1, ]) 
+chld 
+temp_brick <-  brick("../data/env/netcdf/2002_03_mlp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+mlps <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
+mlps
+temp_brick <-  brick("../data/env/netcdf/2002_03_o2.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+o2s <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
+o2s
+o2d <- extract(x=tempbrick[[23]], y = randchsp[1, ]) 
+o2d
+temp_brick <-  brick("../data/env/netcdf/2002_03_salinity.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sals <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
+sals
+sald <- extract(x=tempbrick[[23]], y = randchsp[1, ]) 
+sald
+temp_brick <-  brick("../data/env/netcdf/2002_03_temp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+temps <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
+temps
+tempd <- extract(x=tempbrick[[23]], y = randchsp[1, ]) 
+tempd
+temp_brick <-  brick("../data/env/netcdf/2002_03_ssh.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sshs <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
+sshs
+
+#row 2
+temp_brick <-  brick("../data/env/netcdf/2007_07_chl.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+tempbrick
+chls <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
+chls
+chld <- extract(x=tempbrick[[13]], y = randchsp[2, ]) 
+chld 
+temp_brick <-  brick("../data/env/netcdf/2007_07_mlp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+mlps <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
+mlps
+temp_brick <-  brick("../data/env/netcdf/2007_07_o2.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+o2s <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
+o2s
+o2d <- extract(x=tempbrick[[13]], y = randchsp[2, ]) 
+o2d
+temp_brick <-  brick("../data/env/netcdf/2007_07_salinity.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sals <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
+sals
+sald <- extract(x=tempbrick[[13]], y = randchsp[2, ]) 
+sald
+temp_brick <-  brick("../data/env/netcdf/2007_07_temp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+temps <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
+temps
+tempd <- extract(x=tempbrick[[13]], y = randchsp[2, ]) 
+tempd
+temp_brick <-  brick("../data/env/netcdf/2007_07_ssh.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sshs <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
+sshs
+
+#row 3
+temp_brick <-  brick("../data/env/netcdf/2011_08_chl.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+tempbrick
+chls <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
+chls
+chld <- extract(x=tempbrick[[13]], y = randchsp[3, ]) 
+chld 
+temp_brick <-  brick("../data/env/netcdf/2011_08_mlp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+mlps <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
+mlps
+temp_brick <-  brick("../data/env/netcdf/2011_08_o2.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+o2s <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
+o2s
+o2d <- extract(x=tempbrick[[13]], y = randchsp[3, ]) 
+o2d
+temp_brick <-  brick("../data/env/netcdf/2011_08_salinity.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sals <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
+sals
+sald <- extract(x=tempbrick[[13]], y = randchsp[3, ]) 
+sald
+temp_brick <-  brick("../data/env/netcdf/2011_08_temp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+temps <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
+temps
+tempd <- extract(x=tempbrick[[13]], y = randchsp[3, ]) 
+tempd
+temp_brick <-  brick("../data/env/netcdf/2011_08_ssh.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sshs <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
+sshs
+
+#row 4
+temp_brick <-  brick("../data/env/netcdf/2000_05_chl.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+tempbrick
+chls <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
+chls
+chld <- extract(x=tempbrick[[2]], y = randchsp[4, ]) 
+chld 
+temp_brick <-  brick("../data/env/netcdf/2000_05_mlp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+mlps <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
+mlps
+temp_brick <-  brick("../data/env/netcdf/2000_05_o2.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+o2s <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
+o2s
+o2d <- extract(x=tempbrick[[2]], y = randchsp[4, ]) 
+o2d
+temp_brick <-  brick("../data/env/netcdf/2000_05_salinity.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sals <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
+sals
+sald <- extract(x=tempbrick[[2]], y = randchsp[4, ]) 
+sald
+temp_brick <-  brick("../data/env/netcdf/2000_05_temp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+temps <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
+temps
+tempd <- extract(x=tempbrick[[2]], y = randchsp[4, ]) 
+tempd
+temp_brick <-  brick("../data/env/netcdf/2000_05_ssh.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sshs <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
+sshs
+
+#row 5
+temp_brick <-  brick("../data/env/netcdf/2012_06_chl.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+tempbrick
+chls <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
+chls
+chld <- extract(x=tempbrick[[25]], y = randchsp[5, ]) 
+chld 
+temp_brick <-  brick("../data/env/netcdf/2012_06_mlp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+mlps <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
+mlps
+temp_brick <-  brick("../data/env/netcdf/2012_06_o2.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+o2s <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
+o2s
+o2d <- extract(x=tempbrick[[25]], y = randchsp[5, ]) 
+o2d
+temp_brick <-  brick("../data/env/netcdf/2012_06_salinity.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sals <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
+sals
+sald <- extract(x=tempbrick[[25]], y = randchsp[5, ]) 
+sald
+temp_brick <-  brick("../data/env/netcdf/2012_06_temp.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+temps <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
+temps
+tempd <- extract(x=tempbrick[[25]], y = randchsp[5, ]) 
+tempd
+temp_brick <-  brick("../data/env/netcdf/2012_06_ssh.nc", lvar = 4)
+tempbrick <- projectRaster(temp_brick, aea)
+sshs <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
+sshs
+```
+
+it worked!

### 2018-11-26 src/biological_data_prep.Rmd
minor updates to both background and presence datasets (occurrence coded on background, obs by cell xyzt added to presence)

@@
 
+time-sliced by depth
+
+```{r}
+obs_cell_yymm_depth <- count(data_aea, c("cell_id", "year", "month", "depthlayerno"))
+data_aea <- merge(data_aea, obs_cell_yymm_depth, by.x=c("cell_id", "year", "month", "depthlayerno"), by.y=c("cell_id", "year", "month", "depthlayerno"))
+names(data_aea)[names(data_aea)=="freq"] <- "total_cell_obs_xyzt"
+```
+
 write to .csv

### 2018-11-26 src/timesliced_random_cell_selection.Rmd
minor updates to both background and presence datasets (occurrence coded on background, obs by cell xyzt added to presence)

@@
 background_all <- as.data.frame(background_all_sp)
-head(background_all_sp)
+head(background_all)
 ```
@@
 it worked!
+
+# observation occurrence yes or no
+
+create a new column for Occurence, and populate with "0" (because these are all background)
+```{r}
+background_all$occurrence <- "0"
+write.csv(background_all, "../output/bio/background_complete_obs.csv", row.names = FALSE)
+head(background_all)
+```

### 2018-11-26 src/biological_data_prep.Rmd
kelvin to celsius conversions on both datasets

@@
 Note this function checks to see if the packages are installed first.
-Th
+
 ```{r pre-install packages, message=FALSE}
-packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful", "marmap")
+packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful", "marmap", "weathermetrics")
 source("../src/ipak.R")
@@
 
+# add a col for temp variable in celsius (every-day person readable)
+
+```{r}
+data_aea$temp_celsius_depth <- kelvin.to.celsius(data_aea$temp_depth, round = 4) # 4 is the number of decimal places. the values are expressed to 4
+data_aea$temp_celsius_surface <- kelvin.to.celsius(data_aea$temp_surface, round = 4)
+write.csv(data_aea, "../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env_naocor_cels.csv", row.names = FALSE)
+```
+
 # warm/cold year

### 2018-11-26 src/timesliced_random_cell_selection.Rmd
kelvin to celsius conversions on both datasets

@@
 ```{r pre-install & load packages, include=FALSE}
-install.packages("pacman")
-library(pacman)
-p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap, ncdf4, marmap, plyr)
+if (!require("pacman")) install.packages("pacman")
+p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap, ncdf4, marmap, plyr, weathermetrics)
 ```
@@
 ```
+
+# add a col for temp variable in celsius (every-day person readable)
+
+Uses the weathermetrics package
+
+```{r}
+background_all$temp_celsius_depth <- kelvin.to.celsius(background_all$temp_depth, round = 4) # 4 is the number of decimal places. the values are expressed to 4
+background_all$temp_celsius_surface <- kelvin.to.celsius(background_all$temp_surface, round = 4)
+write.csv(background_all, "../output/bio/background_complete_obs_cels.csv", row.names = FALSE)
+```

### 2018-11-26 src/data_exploration.Rmd
env. correlate boxplots per month and per year

@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("plyr", "graphics") 
+packages <- c("plyr", "graphics", "ggplot2") 
 source("../src/ipak.R")
@@
 ```{r}
-presence <- read.csv("../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env.csv", header = TRUE)
+presence <- read.csv("../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env_naocor_cels.csv", header = TRUE)
 head(presence)
@@
 
+```{r}
+hist(presence$temp_dept, main = "Temp at Depth", xlab = "Kelvin")
+dev.copy(png, "../output/env/simple_plots/temperature_depth_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
 # simple boxplots of env. corr
@@
 ```
+
+```{r}
+boxplot(presence$temp_depth ~ presence$month, xlab = "month", ylab = "Temperature (Kelvin)", main = "Temperature at Observation Depth per Month")
+dev.copy(png, "../output/env/simple_plots/temperature_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$chl_depth ~ presence$month, xlab = "month", ylab = "Chlorophyll Concentrations (mmol.m-3)", main = "Chlorophyll at Observation Depth per Month")
+dev.copy(png, "../output/env/simple_plots/chlorophyll_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$salinity_depth ~ presence$month, xlab = "month", ylab = "Salinity Practical Salinity Unit (PSU)", main = "Salinity at Observation Depth per Month")
+dev.copy(png, "../output/env/simple_plots/salinity_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+o2
+
+xxxx
+```{r}
+boxplot(presence$temp_depth ~ presence$year, xlab = "Year", ylab = "Temperature (Kelvin)", main = "Temperature at Observation Depth")
+dev.copy(png, "../output/env/simple_plots/temperature_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```

### 2018-11-26 src/data_exploration.Rmd
finished yr/month bloxplots

@@
 
-o2
+```{r}
+boxplot(presence$o2_depth ~ presence$month, xlab = "month", ylab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)", main = "Dissolved Oxygen at Observation Depth per Month")
+dev.copy(png, "../output/env/simple_plots/o2_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
 
-xxxx
 ```{r}
-boxplot(presence$temp_depth ~ presence$year, xlab = "Year", ylab = "Temperature (Kelvin)", main = "Temperature at Observation Depth")
+boxplot(presence$mlp_surface ~ presence$month, xlab = "month", ylab = "Mixed Layer Thickness (m)", main = "Mixed Layer Thickness at Observation per Month")
+dev.copy(png, "../output/env/simple_plots/mlp_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$ssh_surface ~ presence$month, xlab = "month", ylab = "Sea Surface Height (m)", main = "Sea Surface Height at Observation per Month")
+dev.copy(png, "../output/env/simple_plots/ssh_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$temp_depth ~ presence$year, xlab = "Year", ylab = "Temperature (Kelvin)", main = "Temperature at Observation Depth per Year")
 dev.copy(png, "../output/env/simple_plots/temperature_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
@@
 ```
+
+```{r}
+boxplot(presence$chl_depth ~ presence$year, xlab = "Year", ylab = "Chlorophyll Concentrations (mmol.m-3)", main = "Chlorophyll at Observation Depth per Year")
+dev.copy(png, "../output/env/simple_plots/chl_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$salinity_depth ~ presence$year, xlab = "Year", ylab = "Salinity Practical Salinity Unit (PSU)", main = "Salinity at Observation Depth per Year")
+dev.copy(png, "../output/env/simple_plots/salinity_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$o2_depth ~ presence$year, xlab = "Year", ylab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)", main = "Dissolved Oxygen at Observation Depth per Year")
+dev.copy(png, "../output/env/simple_plots/oxygen_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$mlp ~ presence$year, xlab = "Year", ylab = "Mixed Layer Thickness (m)", main = "Mixed Layer Thickness at Observation per Year")
+dev.copy(png, "../output/env/simple_plots/mlp_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$ssh ~ presence$year, xlab = "Year", ylab = "Sea Surface Height (m)", main = "Sea Surface Height at Observation per Year")
+dev.copy(png, "../output/env/simple_plots/ssh_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$temp_surface ~ presence$month, xlab = "month", ylab = "Temperature (Kelvin)", main = "Temperature at Observation (Surface) per Month")
+dev.copy(png, "../output/env/simple_plots/temperature_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$chl_surface ~ presence$month, xlab = "month", ylab = "Chlorophyll Concentrations (mmol.m-3)", main = "Chlorophyll at Observation (Surface) per Month")
+dev.copy(png, "../output/env/simple_plots/chlorophyll_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$salinity_surface ~ presence$month, xlab = "month", ylab = "Salinity Practical Salinity Unit (PSU)", main = "Salinity at Observation (Surface) per Month")
+dev.copy(png, "../output/env/simple_plots/salinity_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$o2_surface ~ presence$month, xlab = "month", ylab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)", main = "Dissolved Oxygen at Observation (Surface) per Month")
+dev.copy(png, "../output/env/simple_plots/o2_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$mlp_surface ~ presence$month, xlab = "month", ylab = "Mixed Layer Thickness (m)", main = "Mixed Layer Thickness at Observation per Month")
+dev.copy(png, "../output/env/simple_plots/mlp_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$ssh_surface ~ presence$month, xlab = "month", ylab = "Sea Surface Height (m)", main = "Sea Surface Height at Observation per Month")
+dev.copy(png, "../output/env/simple_plots/ssh_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$temp_surface ~ presence$year, xlab = "Year", ylab = "Temperature (Kelvin)", main = "Temperature at Observation (Surface) per Year")
+dev.copy(png, "../output/env/simple_plots/temperature_boxplot_surface_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$chl_surface ~ presence$year, xlab = "Year", ylab = "Chlorophyll Concentrations (mmol.m-3)", main = "Chlorophyll at Observation (Surface) per Year")
+dev.copy(png, "../output/env/simple_plots/chl_boxplot_surface_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$salinity_surface ~ presence$year, xlab = "Year", ylab = "Salinity Practical Salinity Unit (PSU)", main = "Salinity at Observation (Surface) per Year")
+dev.copy(png, "../output/env/simple_plots/salinity_boxplot_surface_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(presence$o2_surface ~ presence$year, xlab = "Year", ylab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)", main = "Dissolved Oxygen at Observation (Surface) per Year")
+dev.copy(png, "../output/env/simple_plots/oxygen_boxplot_surface_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```

### 2018-11-26 src/data_exploration.Rmd
histograms of all env. correlates

@@
 
-just do surface layers for now
-
 temperature
@@
 dev.copy(png, "../output/env/simple_plots/temperature_depth_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+chl
+```{r}
+hist(presence$chl_surface, main = "Chlorophyll at Surface", xlab = "Chlorophyll Concentrations (mmol.m-3)")
+dev.copy(png, "../output/env/simple_plots/chl_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+hist(presence$chl_depth, main = "Chlorophyll at Observation Depth", xlab = "Chlorophyll Concentrations (mmol.m-3)")
+dev.copy(png, "../output/env/simple_plots/chl_depth_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+salinity
+```{r}
+hist(presence$salinity_surface, main = "Salinity at Surface", xlab = "Salinity Practical Salinity Unit (PSU)")
+dev.copy(png, "../output/env/simple_plots/sal_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+hist(presence$salinity_depth, main = "Salinity at Observation Depth", xlab = "Salinity Practical Salinity Unit (PSU)")
+dev.copy(png, "../output/env/simple_plots/sal_depth_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+oxygen
+```{r}
+hist(presence$o2_surface, main = "Dissolved Oxygen at Surface", xlab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)")
+dev.copy(png, "../output/env/simple_plots/o2_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+hist(presence$o2_depth, main = "Dissolved Oxygen at Observation Depth", xlab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)")
+dev.copy(png, "../output/env/simple_plots/o2_depth_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+MLP
+```{r}
+hist(presence$mlp, main = "Mixed Layer Thickness at Observation", xlab = "Mixed Layer Thickness (m)")
+dev.copy(png, "../output/env/simple_plots/mlp_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+ssh
+```{r}
+hist(presence$ssh, main = "Sea Surface Height at Observation", xlab = "Sea Surface Height (m)")
+dev.copy(png, "../output/env/simple_plots/ssh_surface_histogram.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png

### 2018-11-26 src/data_exploration.Rmd
background env correlate histograms

@@
 
+# presence points
 Read in the presence points
@@
 
-# Update: Who sampled in which year and which month:
+## Update: Who sampled in which year and which month:
 
@@
 
-# environmental correlates
+## environmental correlates
 
@@
 
-#simple plots of env. correlates
+## simple plots of env. correlates
 
@@
 
-# simple boxplots of env. corr
+## simple boxplots of env. corr
 
 Same as above but with boxplots (may provide some more useful info)
-
-XXX SAM YOU WANT TO PUT THE TWO LAYERS ON THE ONE PLOT - NOT TWO SIDE BY SIDE
 
@@
 ```
+
+# background points
+Read in the presence points
+
+```{r}
+background <- read.csv("../output/bio/background_complete_obs_cels.csv", header = TRUE)
+head(background)
+```
+
+# environmental correlates
+
+1) Get the max, min, an mean values and add into a dataframe
+
+Temperature
+```{r}
+temp_depth_max_back <- max(background$temp_depth, na.rm = TRUE)
+temp_depth_min_back <- min(background$temp_depth, na.rm = TRUE)
+temp_depth_mean_back <- mean(background$temp_depth, na.rm = TRUE)
+temp_surface_max_back <- max(background$temp_surface, na.rm = TRUE)
+temp_surface_min_back <- min(background$temp_surface, na.rm = TRUE)
+temp_surface_mean_back <- mean(background$temp_surface, na.rm = TRUE)
+```
+
+Salinity
+```{r}
+sal_depth_max_back <- max(background$salinity_depth, na.rm = TRUE)
+sal_depth_min_back <- min(background$salinity_depth, na.rm = TRUE)
+sal_depth_mean_back <- mean(background$salinity_depth, na.rm = TRUE)
+sal_surface_max_back <- max(background$salinity_surface, na.rm = TRUE)
+sal_surface_min_back <- min(background$salinity_surface, na.rm = TRUE)
+sal_surface_mean_back <- mean(background$salinity_surface, na.rm = TRUE)
+```
+
+Chl
+```{r}
+chl_depth_max_back <- max(background$chl_depth, na.rm = TRUE)
+chl_depth_min_back <- min(background$chl_depth, na.rm = TRUE)
+chl_depth_mean_back <- mean(background$chl_depth, na.rm = TRUE)
+chl_surface_max_back <- max(background$chl_surface, na.rm = TRUE)
+chl_surface_min_back <- min(background$chl_surface, na.rm = TRUE)
+chl_surface_mean_back <- mean(background$chl_surface, na.rm = TRUE)
+```
+
+O2
+```{r}
+o2_depth_max_back <- max(background$o2_depth, na.rm = TRUE)
+o2_depth_min_back <- min(background$o2_depth, na.rm = TRUE)
+o2_depth_mean_back <- mean(background$o2_depth, na.rm = TRUE)
+o2_surface_max_back <- max(background$o2_surface, na.rm = TRUE)
+o2_surface_min_back <- min(background$o2_surface, na.rm = TRUE)
+o2_surface_mean_back <- mean(background$o2_surface, na.rm = TRUE)
+```
+
+MLP
+```{r}
+mlp_surface_max_back <- max(background$mlp_surface, na.rm = TRUE)
+mlp_surface_min_back <- min(background$mlp_surface, na.rm = TRUE)
+mlp_surface_mean_back <- mean(background$mlp_surface, na.rm = TRUE)
+```
+
+SSH
+```{r}
+ssh_surface_max_back <- max(background$ssh_surface, na.rm = TRUE)
+ssh_surface_min_back <- min(background$ssh_surface, na.rm = TRUE)
+ssh_surface_mean_back <- mean(background$ssh_surface, na.rm = TRUE)
+```
+
+create matrix
+```{r}
+tdb <- c(temp_depth_max_back, temp_depth_min_back, temp_depth_mean_back)
+tsb <- c(temp_surface_max_back, temp_surface_min_back, temp_surface_mean_back)
+sdb <- c(sal_depth_max_back, sal_depth_min_back, sal_depth_mean_back)
+ssb <- c(sal_surface_max_back, sal_surface_min_back, sal_surface_mean_back)
+cdb <- c(sal_depth_max_back, sal_depth_min_back, sal_depth_mean_back)
+csb <- c(sal_surface_max_back, sal_surface_min_back, sal_surface_mean_back)
+odb <- c(o2_depth_max_back, o2_depth_min_back, o2_depth_mean_back)
+osb <- c(o2_surface_max_back, o2_surface_min_back, o2_surface_mean_back)
+mlpb <- c(mlp_surface_max_back, mlp_surface_min_back, mlp_surface_mean_back)
+sshb <- c(ssh_surface_max_back, ssh_surface_min_back, ssh_surface_mean_back)
+env_stats_back <- rbind(tdb, tsb, sdb, ssb, cdb, csb, odb, osb, mlpb, sshb)
+row.names(env_stats_back) <- c("Temp Depth", "Temp Surface", "Salinity Depth", "Salinity Surface", "Chl Depth", "Chl Surface", "Oxygen Depth", "Oxygen Surface", "MLP", "SSH") 
+colnames(env_stats_back) <- c("Max", "Min", "Mean")
+write.csv(env_stats_back, "../output/env/env_correlates_background_basic_stats.csv", row.names = TRUE)
+env_stats_back
+```

### 2018-11-26 src/timesliced_random_cell_selection.Rmd
background env correlate histograms

@@
 ```{r}
+background_all <- read.csv("../output/bio/background_complete.csv")
 background_all$occurrence <- "0"

### 2018-11-26 src/data_exploration.Rmd
background env correlates boxplots

@@
 ```
+
+temperature
+```{r}
+par(mfrow=c(1,2))
+plot(background$temp_depth, main = "Background Points Temp at Depth (Various)", ylab = "Kelvin")
+plot(background$temp_surface, main = "Background Points Temp at Surface", ylab = "Kelvin")
+dev.copy(png, "../output/env/simple_plots/temperature_background_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+salinity
+```{r}
+par(mfrow=c(1,2))
+plot(background$salinity_depth, main = "Background Points Salinity at Depth (Various)", ylab = "Practical Salinity Units")
+plot(background$salinity_surface, main = "Background Points Salinity at Surface", ylab = "Practical Salinity Units")
+dev.copy(png,"../output/env/simple_plots/salinity_background_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+Chlorophyll
+```{r}
+par(mfrow=c(1,2))
+plot(background$chl_depth, main = "Background Points Chl at Depth (Various)", ylab = "Chl (mmol.m-3)")
+plot(background$chl_surface, main = "Background Points Chl at Surface", ylab = "Chl (mmol.m-3)")
+dev.copy(png,"../output/env/simple_plots/chl_background_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+O2
+```{r}
+par(mfrow=c(1,2))
+plot(background$o2_depth, main = "Background Points Dissolved Oxygen at Depth (Various)", ylab = "Chl (mmol.m-3)")
+plot(background$o2_surface, main = "Background Points Dissolved Oxygen at Surface", ylab = "O2 (mmol.m-3)")
+dev.copy(png,"../output/env/simple_plots/o2_background_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+mlp
+```{r}
+plot(background$mlp_surface, main = "Background Points Mixed Layer Thickness", ylab = "Depth (m)")
+dev.copy(png,"../output/env/simple_plots/mlp_background_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+SSH
+```{r}
+plot(background$ssh_surface, main = "Background Points Sea Surface Height", ylab = "Height (m)")
+dev.copy(png,"../output/env/simple_plots/ssh_background_simpleplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+## frequency plots of env. corr
+
+temperature
+```{r}
+hist(background$temp_surface, main = "Background Temp at Surface", xlab = "Kelvin")
+dev.copy(png, "../output/env/simple_plots/temperature_background_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+hist(background$temp_dept, main = "Background Temp at Depth", xlab = "Kelvin")
+dev.copy(png, "../output/env/simple_plots/temperature_background_depth_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+chl
+```{r}
+hist(background$chl_surface, main = "Background Chlorophyll at Surface", xlab = "Chlorophyll Concentrations (mmol.m-3)")
+dev.copy(png, "../output/env/simple_plots/chl_background_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+hist(background$chl_depth, main = "Background Chlorophyll at Observation Depth", xlab = "Chlorophyll Concentrations (mmol.m-3)")
+dev.copy(png, "../output/env/simple_plots/chl_background_depth_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+salinity
+```{r}
+hist(background$salinity_surface, main = "Background Salinity at Surface", xlab = "Salinity Practical Salinity Unit (PSU)")
+dev.copy(png, "../output/env/simple_plots/sal_background_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+hist(background$salinity_depth, main = "Background Salinity at Observation Depth", xlab = "Salinity Practical Salinity Unit (PSU)")
+dev.copy(png, "../output/env/simple_plots/sal_background_depth_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+oxygen
+```{r}
+hist(background$o2_surface, main = "Background Dissolved Oxygen at Surface", xlab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)")
+dev.copy(png, "../output/env/simple_plots/o2_background_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+hist(background$o2_depth, main = "Background Dissolved Oxygen at Observation Depth", xlab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)")
+dev.copy(png, "../output/env/simple_plots/o2_background_depth_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+MLP
+```{r}
+hist(background$mlp, main = "Background Mixed Layer Thickness at Observation", xlab = "Mixed Layer Thickness (m)")
+dev.copy(png, "../output/env/simple_plots/mlp_background_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+ssh
+```{r}
+hist(background$ssh, main = "Background Sea Surface Height at Observation", xlab = "Sea Surface Height (m)")
+dev.copy(png, "../output/env/simple_plots/ssh_background_surface_histogram.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+## simple boxplots of env. corr
+
+Same as above but with boxplots (may provide some more useful info)
+
+temperature
+
+```{r}
+boxplot(background$temp_depth ~ background$month, xlab = "month", ylab = "Temperature (Kelvin)", main = "Temperature at Background Depth per Month")
+dev.copy(png, "../output/env/simple_plots/temperature_background_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$chl_depth ~ background$month, xlab = "month", ylab = "Chlorophyll Concentrations (mmol.m-3)", main = "Chlorophyll at Background Depth per Month")
+dev.copy(png, "../output/env/simple_plots/chlorophyll_background_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$salinity_depth ~ background$month, xlab = "month", ylab = "Salinity Practical Salinity Unit (PSU)", main = "Salinity at Background Depth per Month")
+dev.copy(png, "../output/env/simple_plots/salinity_background_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$o2_depth ~ background$month, xlab = "month", ylab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)", main = "Dissolved Oxygen at Background Depth per Month")
+dev.copy(png, "../output/env/simple_plots/o2_background_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$mlp_surface ~ background$month, xlab = "month", ylab = "Mixed Layer Thickness (m)", main = "Mixed Layer Thickness at Background per Month")
+dev.copy(png, "../output/env/simple_plots/mlp_background_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$ssh_surface ~ background$month, xlab = "month", ylab = "Sea Surface Height (m)", main = "Sea Surface Height at Background per Month")
+dev.copy(png, "../output/env/simple_plots/ssh_background_boxplot_dept_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$temp_depth ~ background$year, xlab = "Year", ylab = "Temperature (Kelvin)", main = "Temperature at Background Depth per Year")
+dev.copy(png, "../output/env/simple_plots/temperature_background_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$chl_depth ~ background$year, xlab = "Year", ylab = "Chlorophyll Concentrations (mmol.m-3)", main = "Chlorophyll at Background Depth per Year")
+dev.copy(png, "../output/env/simple_plots/chl_background_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$salinity_depth ~ background$year, xlab = "Year", ylab = "Salinity Practical Salinity Unit (PSU)", main = "Salinity at Background Depth per Year")
+dev.copy(png, "../output/env/simple_plots/salinity_background_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$o2_depth ~ background$year, xlab = "Year", ylab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)", main = "Dissolved Oxygen at Background Depth per Year")
+dev.copy(png, "../output/env/simple_plots/oxygen_background_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$mlp ~ background$year, xlab = "Year", ylab = "Mixed Layer Thickness (m)", main = "Mixed Layer Thickness at Background per Year")
+dev.copy(png, "../output/env/simple_plots/mlp_background_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$ssh ~ background$year, xlab = "Year", ylab = "Sea Surface Height (m)", main = "Sea Surface Height at Background per Year")
+dev.copy(png, "../output/env/simple_plots/ssh_background_boxplot_dept_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$temp_surface ~ background$month, xlab = "month", ylab = "Temperature (Kelvin)", main = "Temperature at Background (Surface) per Month")
+dev.copy(png, "../output/env/simple_plots/temperature_background_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$chl_surface ~ background$month, xlab = "month", ylab = "Chlorophyll Concentrations (mmol.m-3)", main = "Chlorophyll at Background (Surface) per Month")
+dev.copy(png, "../output/env/simple_plots/chlorophyll_background_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$salinity_surface ~ background$month, xlab = "month", ylab = "Salinity Practical Salinity Unit (PSU)", main = "Salinity at Background (Surface) per Month")
+dev.copy(png, "../output/env/simple_plots/salinity_background_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$o2_surface ~ background$month, xlab = "month", ylab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)", main = "Dissolved Oxygen at Background (Surface) per Month")
+dev.copy(png, "../output/env/simple_plots/o2_background_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$mlp_surface ~ background$month, xlab = "month", ylab = "Mixed Layer Thickness (m)", main = "Mixed Layer Thickness at Background per Month")
+dev.copy(png, "../output/env/simple_plots/mlp_background_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$ssh_surface ~ background$month, xlab = "month", ylab = "Sea Surface Height (m)", main = "Sea Surface Height at Background per Month")
+dev.copy(png, "../output/env/simple_plots/ssh_background_boxplot_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$temp_surface ~ background$year, xlab = "Year", ylab = "Temperature (Kelvin)", main = "Temperature at Background (Surface) per Year")
+dev.copy(png, "../output/env/simple_plots/temperature_background_boxplot_surface_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$chl_surface ~ background$year, xlab = "Year", ylab = "Chlorophyll Concentrations (mmol.m-3)", main = "Chlorophyll at Background (Surface) per Year")
+dev.copy(png, "../output/env/simple_plots/chl_background_boxplot_surface_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$salinity_surface ~ background$year, xlab = "Year", ylab = "Salinity Practical Salinity Unit (PSU)", main = "Salinity at Background (Surface) per Year")
+dev.copy(png, "../output/env/simple_plots/salinity_background_boxplot_surface_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+boxplot(background$o2_surface ~ background$year, xlab = "Year", ylab = "Mole Concentration of Dissolved Oxygen in Sea Water (mmol.m-3)", main = "Dissolved Oxygen at Background (Surface) per Year")
+dev.copy(png, "../output/env/simple_plots/oxygen_background_boxplot_surface_year.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```

### 2018-11-26 src/data_exploration.Rmd
background_corr plot... not saved as image

@@
 ```
+
+## correlations between background points
+
+check to see if there are any correlations in the env. variables
+
+first subset the env.correlate columns (you don't need everything)
+
+```{r}
+str(background)
+background_env <- subset(background, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+backpairs <- pairs(background_env)
+dev.copy(png, "../output/env/simple_plots/background_corr.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+```

### 2018-11-26 src/data_exploration.Rmd
correlation plots of presence NB output is effectively non-readable

--- 2018-11-26 src/data_exploration.Rmd
@@
 ```
+
+## correlations between background points
+
+check to see if there are any correlations in the env. variables
+
+first subset the env.correlate columns (you don't need everything)
+
+```{r}
+str(background)
+background_env <- subset(background, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+dev.copy(png, "../output/env/simple_plots/background_corr.png") # to automatically save the plot to a png AND show it inline
+backpairs <- pairs(background_env)
+dev.copy(png, "../output/env/simple_plots/background_corr.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+## correlations between presence points
+
+check to see if there are any correlations in the env. variables
+
+first subset the env.correlate columns (you don't need everything)
+
+```{r}
+str(presence)
+pres_env <- subset(presence, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+png("../output/env/simple_plots/presence_corr.png") # to automatically save the plot to a png AND show it inline
+pairs(pres_env)
+dev.off() # stops automatic saving of the plot to a png
+```

### 2018-11-27 src/data_exploration.Rmd
background and presence correlation plots and output

--- 2018-11-26 src/data_exploration.Rmd
@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("plyr", "graphics", "ggplot2") 
+packages <- c("plyr", "graphics", "ggplot2", "corrplot") 
 source("../src/ipak.R")
@@
 ```
+
+## correlations between background points
+
+check to see if there are any correlations in the env. variables
+
+first subset the env.correlate columns (you don't need everything)
+
+```{r}
+background_env <- subset(background, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+background_env_cor <- cor(background_env, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(background_env_cor, "../output/env/background_env_corr.csv", row.names = TRUE)
+background_corrplot <- corrplot(background_env_cor , method = "color", type = "upper", order = "alphabet", tl.cex = 0.8) 
+dev.copy(png,"../output/env/simple_plots/background_envcorr.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+## correlations between presence points
+
+check to see if there are any correlations in the env. variables
+
+first subset the env.correlate columns (you don't need everything) then use cor to get the correlation values, and then corrplot for a visual
+
+```{r}
+pres_env <- subset(presence, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+pres_env_cor <- cor(pres_env, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(pres_env_cor, "../output/env/pres_env_corr.csv", row.names = TRUE)
+pres_corrplot <- corrplot(pres_env_cor , method = "color", type = "upper", order = "alphabet", tl.cex = 0.8) 
+dev.copy(png,"../output/env/simple_plots/pres_envcorr.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+pairs.panels(pres_env_cor)
+```

### 2018-11-27 src/data_exploration.Rmd
density plots + addition of tiff to gitignore

@@
 author: "Samantha Andrews"
-output: html_notebook
+output: 
+  html_notebook: 
+    fig_height: 7
+    fig_width: 10
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 
+to get some density plots all in one graphic, you need to change chunk output to in console, then go to the plot tab, make it fill the screen, then run then make bigger then save :( (probably a better way - this seems to be an issue with RStudio)
+```{r}
+par(mfrow=c(4,4))
+for(i in 1:16){
+  plot(density(background_env[,i], na.rm=T), main = names(background_env)[i])
+}
+```
+PUT THE CHUNK OUTPUT BACK TO INLINE
+
 ## correlations between presence points
@@
 
+to get some density plots all in one graphic, you need to change chunk output to in console, then go to the plot tab, make it fill the screen, then run then make bigger then save :( (probably a better way - this seems to be an issue with RStudio)
 ```{r}
-pairs.panels(pres_env_cor)
+par(mfrow=c(4,4))
+for(i in 1:16){
+  plot(density(pres_env[,i], na.rm=T), main = names(pres_env)[i])
+}
 ```
+PUT THE CHUNK OUTPUT BACK TO INLINE

### 2018-11-27 src/data_exploration.Rmd
density plot with background and presence env. data

@@
 PUT THE CHUNK OUTPUT BACK TO INLINE
+
+## density plot with background and presence env. data
+
+Inspired by Merrow 2013 - top paragraph of page 1063 (are the species observations uniformly distributed over the background, or are they skewed)
+
+```{r}
+ggplot(pres_env, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/temp_depth_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+ggplot(pres_env, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/temp_surface_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+ggplot(pres_env, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/chl_surface_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+ggplot(pres_env, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/chl_depth_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+ggplot(pres_env, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/salinity_surface_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+ggplot(pres_env, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/salinity_depth_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+ggplot(pres_env, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/o2_surface_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+ggplot(pres_env, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/o2_depth_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+ggplot(pres_env, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/mlp_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+ggplot(pres_env, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=background_env, na.rm = TRUE, colour = "blue")
+dev.copy(png,"../output/env/simple_plots/ssh_backvspres.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```

### 2018-11-27 src/env_data_prep.Rmd
simple plots complete

@@
 
- NA values are represented by -99.99. Change these to NA
+ NA values are represented by -999. Change these to NA
 ```{r}
-naoraw <- replace(naoraw,naoraw==-99.99, NA)
+naoraw <- replace(naoraw,naoraw==-999, NA)
 ```
@@
 head(nao) 
+summary(nao)
 ```
@@
 ```{r}
-naocolumn <- matrix(t(amoraw[ ,2:13]), ncol=1) #month data is in col 2 - 13
+naocolumn <- matrix(t(nao[ ,2:13]), ncol=1) #month data is in col 2 - 13 
 naocolumn1 <- as.vector(naocolumn)

### 2018-12-03 src/parallel_env_data.Rmd
trying to get parallel to work...

--- 2018-11-26 src/timesliced_random_cell_selection.Rmd
@@
 ---
-title: "timesliced_random_cell_selection"
+title: "parallel for the env. data grab"
 author: "Samantha Andrews"
-output: html_notebook
+output: 
+  html_notebook: 
+    fig_height: 7
+    fig_width: 10
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 # Overview
-Preperation of the capelin (*Mallotus villosus*) data obtained from [OBIS](www.iobis.org). This database shows the global occurance of capelin.
+Attempt to run my inefficinet giant loop for grabbing env. data to the background points in parallel (need to run again AND maybe need to do more points that will currently take 2 months
 
@@
 # Package dependencies
-You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163).
-Note this function checks to see if the packages are installed first before loading.
-
-New package attempt to do this....
-old code 
-packages <- c("raster", "dplyr", "data.table", "sp", "adehabitatHR", "maptools")
+You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
+Note this function checks to see if the packages are installed first.
+The "include=FALSE" supresses the package installation text appearing in the document...
+```{r pre-install packages, include=FALSE}
+packages <- c("plyr", "graphics", "ggplot2", "corrplot") 
 source("../src/ipak.R")
 ipak(packages)
-
-```{r pre-install & load packages, include=FALSE}
-install.packages("pacman")
-library(pacman)
-p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap, ncdf4, marmap, plyr)
 ```
 
-# the show
-you have a unique_cell raster already (glo_unique_cell) (output/env/glo_unique_cell.tif)
-
-```{r}
-glo_unique_cell <- raster("../output/env/glo_unique_cell.tif") #this is loading the cell layer that was created and saved as a tif
-plot(glo_unique_cell)
-```
-
-as a reminder, how many cells are there?
-```{r}
-glo_unique_cell
-```
-
-7750
-
-ok now you want to generate a centroid for each cell... (a point in the center of each cell)
-use the raster layer to 1) transfer the raster into a dataframe, with the values (cell numbers) linked to each centroid with xy coordinates (note as the raster is an aea projection, the xy will be in meters)
-```{r}
-unique_cell_centroid <- as.data.frame(glo_unique_cell, row.names = NULL, optional = FALSE, xy = TRUE, centroids = TRUE, na.rm = TRUE)
-names(unique_cell_centroid)[names(unique_cell_centroid)=="layer.1"] <- "cell_id" #just renames the output
-names(unique_cell_centroid)[names(unique_cell_centroid)=="x"] <- "longitude" #just renames the output
-names(unique_cell_centroid)[names(unique_cell_centroid)=="y"] <- "latitude" #just renames the output
-write.csv(unique_cell_centroid, "../output/env/unique_cell_centroid.csv")
-plot(glo_unique_cell)
-points(unique_cell_centroid$longitude, unique_cell_centroid$latitude, pch = 10, col = "black")
-write.csv(unique_cell_centroid, "../output/env/unique_cell_points.csv", row.names = FALSE)
-```
-well... the plot doesn't tell me much... but i definately have points!
-
-The area covered by these points are huge -much larger than the distribution of my presence points.
-Ideally I need to constrain the area where the background points are selected from to match the entire range of the species.
-The most common method for this is to use a minimum convex hull to estimate the species range, and then use that area to constrain the background point selection.
-Note I am a little worried that I may loose some of the edge data as they are centroids of cells...
-
-```{r}
-mcpbase <- read.csv("../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env.csv", header = TRUE)
-mcpbase_xy <- mcpbase[ , c("decimalLongitude.1", "decimalLatitude.1")] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
-mcpbase_sp <- SpatialPointsDataFrame(coords = mcpbase_xy, data = mcpbase, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
-mcp_capelin_100 <- mcp(mcpbase_sp[ , 12], percent = 100) #12 is the column with the species name in (the id for the species)
-plot(mcp_capelin_100)
-```
-
-oh tres exciting. Let's overlay the occurrence points on that
-
-```{r}
-plot(mcp_capelin_100)
-points(mcpbase_sp$decimalLongitude.1, mcpbase_sp$decimalLatitude.1)
-```
-
-ok yea, i guess.... still wondering about those edge areas....anyway save the hull as a polygon
-
-```{r}
-writeOGR(mcp_capelin_100, dsn = '.', layer = '../output/bio/mcp_capelin_100_poly', driver = "ESRI Shapefile")
-```
-
-now need to select points within a polygon....
-```{r}
-unique_cell_centroid <- read.csv("../output/env/unique_cell_centroid.csv",header = TRUE)
-unique_cell_xy <- unique_cell_centroid[ , c("longitude", "latitude")] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
-unique_c_sp <- SpatialPointsDataFrame(coords = unique_cell_xy, data = unique_cell_centroid, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
-plot(mcp_capelin_100)
-points(unique_c_sp, pch = "red")
-```
-
-```{r}
-mcp_unique_cell <- unique_c_sp[mcp_capelin_100, ] #this is just subsetting 
-plot(mcp_capelin_100)
-points(mcp_unique_cell)
-```
-ok now save the clipped unique cells to a .csv
-```{r}
-write.csv(mcp_unique_cell, "../output/env/unique_cell_centroid_mcp")
-```
-
-#decimalLat and Lon
-ok lets's get the real decimalLatitude and decimalLongitude.
-Will need to plot all point, extract the lonlat coordinates
-
-reproject to a wgs (use an existing file as a mask)
-
-```{r}
-sp_mcp_unique_cell <- spTransform(mcp_unique_cell, CRS = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
-plot(sp_mcp_unique_cell)
-```
-extract lonlat, add to dataframe
-
-```{r}
-sp_mcp_unique_cell <- data.frame(sp_mcp_unique_cell, long=coordinates(sp_mcp_unique_cell)[,1], lat=coordinates(sp_mcp_unique_cell)[,2])
-head(sp_mcp_unique_cell)
-```
-
-remove existing long, lat, optional, longitude.1, latitude.1 (duplicate columns) and change name of longitude.2 & latitude.2 columns to decimalLongitude and decimalLatitude respectively, and longitude and latitude to longitude_meters and latitude_meters ande glo_unique_cell to cell_id
-```{r}
-sp_mcp_unique_cell <- as.data.frame(sp_mcp_unique_cell) #need to go back to a 'normal' dataframe to change col names
-head(sp_mcp_unique_cell) 
-sp_mcp_unique_cell <- subset(sp_mcp_unique_cell, select = -c(long, lat, optional, longitude.1, latitude.1))
-colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="longitude.2"] <- "decimalLongitude"
-colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="latitude.2"] <- "decimalLatitude"
-colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="longitude"] <- "longitude_meters"
-colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="latitude"] <- "latitude_meters"
-colnames(sp_mcp_unique_cell)[colnames(sp_mcp_unique_cell)=="glo_unique_cell"] <- "cell_id"
-head(sp_mcp_unique_cell)
-```
-
-save as csv
-```{r}
-write.csv(sp_background, file = "../output/bio/unique_cell_centroid_lonlat.csv", row.names = FALSE)
-```
-
-# Add NAFO Region background points in
-NAFO Zones shapefile obtained from [NAFO](https://www.nafo.int/Data/GIS)
-
-```{r NAFO shapefile extraction}
-coordinates(sp_mcp_unique_cell) <- c("decimalLongitude", "decimalLatitude") 
-nafo_zones <- readOGR(dsn=path.expand("../data/bio/nafo_zones"), layer = "nafo_zones_wgs84") #this loads the shapefile
-proj4string(sp_mcp_unique_cell) <- proj4string(nafo_zones) #tells R that the occurrence data is the same projection as the shapefile
-sp_mcp_unique_cell$nafo_zone <- over(sp_mcp_unique_cell, nafo_zones)$ZONE #ZONE is where the zone data is held in teh shapefiles' attributes
-head(sp_mcp_unique_cell) #just to check it worked
-```
-
-Great. Lets see if any are missing a NAFO Zone. But first, write to a csv and reload as the data is now some wierd 'spatial points dataframe'...
-```{r}
-write.csv(sp_mcp_unique_cell, "../output/bio/unique_cell_centroid_lonlat_nafo.csv", row.names=FALSE)
-sp_mcp_unique_cell <- as.data.frame(sp_mcp_unique_cell) 
-bc_nafo_na <- subset(sp_mcp_unique_cell, is.na(sp_mcp_unique_cell$nafo_zone))
-bc_nafo_na
-```
-
-ok lots don't have a NAFO region attached. This is probably the Hudson Bay issue again...
-
-Uh oh there is a lot. A bet a bunch of them are in the Hudson Strait (which is outside the NAFO Zone).Map the nafo_na points...
-```{r}
-map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
-plot(map2, xlim = c(-65, -40), ylim =c(30, 70), asp = 1, main = "Background missing a NAFO Zone", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
-points(bc_nafo_na$decimalLongitude, bc_nafo_na$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
-dev.copy(png, "../output/bio/background_missing_nafo.png") #this prints a png of the plot
-dev.off() #this turns off the print comman
-```
-
-Indeed the many of the missing background points are in the hudson strait.
-Others are near the coast and thus probably just call outside the NAFO shapefiles (remember the smapefiles are not all super-perfectly fitting).
-Think I will take this to arcgis for fixing....
-
-basically what I did was use the selection -> select by attributes tool to find the points where NAFO_region = NA and then manually recode them (using the field calculator), and exported as a .csv.
-
-just double-check they are all gone
-```{r}
-mcp_unique_cell <- read.csv("../output/env/unique_cell_centroid_lonlat_nafo2.csv", header = TRUE)
-bc_nafo2_na <- subset(mcp_unique_cell, is.na(mcp_unique_cell$nafo_zone))
-bc_nafo2_na
-```
-
-# random depth
-
-ok so now I have a base list to work off...
-This gives you a dataset of xy points, but you need xyz (where z is based on the depth layers in the environmental data).
-This should be constrained by the presence depth range
-```{r}
-max_obs_depth <- max(mcpbase$depthlayerno, na.rm = TRUE) #max is 47
-min_obs_depth <- min(mcpbase$depthlayerno, na.rm = TRUE) #min is 1
-unique_cell_xyz <- mcp_unique_cell
-unique_cell_xyz$depthlayerno <- NA
-unique_cell_xyz <- unique_cell_xyz[rep(seq_len(nrow(unique_cell_xyz)), each=47),] #this generates 47  repetitions of each row 47 is chosen because there are 47 sampled depth layers
-unique_cell_xyz$depthlayerno <- rep(1:47) #assign 1:47 on a repeating cycle to the depth column. 47 is chosen because there are 50 depth layers
-unique_cell_xyz$id_depth <- paste(unique_cell_xyz$cell_id, unique_cell_xyz$depthlayerno, sep="") #create a new column that is a unique glo_unique_cell & depth ID.
-This is to help remove cells later and randomise
-```
-
-What I want to do is for each time slice, create a list of all cells (xyz), then remove the ones where observations occurred and create a .csv with the name of the level being worked on
-
-Split the data_aea into timeslices (year/month).
-All you need is the year, month, cell_id, and depthlayerno
-```{r}
-data_aea <- mcpbase
-cell_yymm <- subset(data_aea, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month, depthlayer no
-cell_yymm$id_depth <- paste(cell_yymm$cell_id, cell_yymm$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID. This is to help remove cells later and randomise
-split_cell_yymm <- split(cell_yymm, list(cell_yymm$year, cell_yymm$month), drop = TRUE)
-```
-
-Here is a loop to create a copy of the unique_cell_centroid dataframe, and give it the same name as the splitobs_cell_yymm level
-.csv naming code courtesy of [nadizan/stack exchange](https://stackoverflow.com/questions/53045158/r-naming-csv-with-a-list-level-name-in-a-loop?noredirect=1#comment92990285_53045158)
-
-```{r timesliced unique unsampled}
-no_timeslices <- length(split_cell_yymm) # how many levels (time slices) are in the list
-bckoutput <- "../output/bio/background/raw/" #where the files are to be saved to
-for (i in 1:no_timeslices){
-  timeslice <- split_cell_yymm[[i]]
-  timeslice <- as.data.frame(timeslice)
-  time_cell <- unique_cell_xyz
-  time_cell <- time_cell[!time_cell$id_depth %in% timeslice$id_depth, , drop = FALSE]
-  write.csv(time_cell, paste0(bckoutput, names(lapply(split_cell_yymm, names))[i], ".csv"), row.names = FALSE) 
-}
-```
-
-Ok so now I have a bunch of .csv files containing cells where no observations occurred for a given timeslice on xyz coordinates.
-Next step is to randomly select cells in each timeslice. Make sure you generate new files for each...
-
-load the .csvs... [code](https://www.reed.edu/data-at-reed/software/R/reading_and_writing.html)
-
-```{r}
-bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
-bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
-randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
-# read in each .csv file in file_list and create a data frame with the same name as the .csv file
-for (i in 1:length(bckrnd_list)){
-  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
-  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample (large sample as per maxent)
-  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
-}
-head(sel)  
-```
-
-ok now i have a ton of files for random background points broken into timeslices.
-I'd like to add a colum for year and one for month to each of these datasets.
-They should be populated based on the name of the file (e.g.
-1998.4.csv would have year = 1998 and month = 4)
-
-```{r}
-rand_folder <- "../output/bio/background/rand/"      # path to folder that holds multiple .csv files
-rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
-randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
-for(i in 1:length(rand_list)){
-  randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
-  randtemp$year <- substr(rand_list[i], 1, 4) #the year is characters 1-4 in the filename
-  randtemp$month <- substr(rand_list[i], 6, 7) #the month is characters 6 & 7 in the filename (except where the month is a single digit)
-  write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
-}
-head(randtemp)
-```
-note that in the output temp file in R shows a . after the month if that month is a single digit. The .csv outputs don't
-
-# populating the background points data with... data
-
-ok now you can start to populate the rest of the background point information.
-The background points file should mirror as much as possible the occurence file.
-Maybe it would be best to merge the files together, and them split them again later?
-would save looping everything to repeat the process through multiple files...
-
-```{r}
-filenames  <- list.files(path = rand_folder, pattern="*.csv", full.names = TRUE) 
-background_all <- rbindlist(lapply(filenames,fread, fill = TRUE))
-write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
-head(background_all)
-```
-
-check for unique years
-```{r}
-back_yr_uniq <- unique(background_all$year)
-back_yr_uniq
-```
-check for unique months
-```{r}
-back_mt_uniq <- unique(background_all$month)
-back_mt_uniq
-```
-
-check number of observations 
-```{r}
-back_rowcount <- nrow(background_all)
-back_rowcount
-```
-
-ok lets start by getting rid of some of the duplicated columns right now (put in by arc, various processes..)
-
-```{r}
-head(background_all)
-```
-remove
-FID
-optional
-
-rename
-longitude_ to longitude_meters
-latitude_m to latitude_meters
-decimalLon to decimalLongitude
-decimalLat to decimalLatitude
-
-```{r}
-background_all <- subset(background_all, select = -c(optional, FID))
-names(background_all)[names(background_all)=="longitude_"] <- "longitude_meters"
-names(background_all)[names(background_all)=="latitude_m"] <- "latitude_meters"
-names(background_all)[names(background_all)=="decimalLon"] <- "decimalLongitude"
-names(background_all)[names(background_all)=="decimalLat"] <- "decimalLatitude"
-head(background_all)
-```
-
-# what data to add
-
-get the columns from the occurence data so you know what to add
-```{r}
-occ_column <- colnames(data_aea)
-occ_column
-```
-
-ok so have cell_id, year, month, decimalLongitude.1(meters - in background_all its longitude), decimalLatitude.1 (meters - in background_all its latitude), depth_layer_no.
-
-# create unique observation ID
-
-Easy first one - add an id based on id_depth
-```{r}
-names(background_all)[names(background_all)=="id_depth"] <- "id"
-write.csv(background_all, "../output/bio/background_all.csv", row.names = FALSE)
-head(background_all)
-```
-
-#original scientific name
-
-create a column called "originalscientificname" and fill it with "Mallotus villosus ab"
-
-```{r}
-background_all$originalscientificname <- "Mallotus villosus ab"
-write.csv(background_all, "../output/bio/background_all_lonlat_nafo2_sciname.csv")
-head(background_all)
-```
-
-# Add AMO Data
-The data has been prepared as per the environmental_data_preperation.Rmd, and biological_data_prep.Rmd
-
-So you want..
-- AMO value at sampling month/year
-- AMO value at previous sampling month/year
-- AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
-
-Load the prepared amo_prev.csv file with has the amo at different year/month steps
-```{r}
-amo_prev <- read.csv("../output/env/amo_prev.csv", header = TRUE)
-head(amo_prev)
-```
-
-now match the year and the month in the two dataframes, and populate the background points with amo_sample
-```{r}
-background_all <- merge(x = background_all, y = amo_prev[ , c("year", "month", "amo_sample")], by = c("year", "month"), all.x = TRUE)
-head(background_all)
-```
-
-ok next - AMO value at previous sampling month/year 
-```{r}
-background_all <- merge(x = background_all, y = amo_prev[ , c("year", "month", "amo_prev")], by = c("year", "month"), all.x = TRUE)
-head(background_all)
-```
-
-and now - AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn) (note they are in a different .csv)
-
-```{r}
-amo_winter <- read.csv("../output/env/amo_winter.csv")
-background_all <- merge(x = background_all, y = amo_winter[ , c("year", "WinterAvg.Shifted")], by = c("year"), all.x = TRUE)
-colnames(background_all)[colnames(background_all)=="WinterAvg.Shifted"] <- "amo_winter" #change the colname
-head(background_all)
-```
-
-ok and write a csv
-```{r}
-write.csv(data_aea,"../output/bio/background_all_lonlat_nafo2_sciname_amo.csv", row.names = FALSE)
-```
-
-# NAO values
-The data has been prepared as per the environmental_data_preperation.Rmd and biological_data_prep.Rmd
-
-So you want..
-- NAO value at sampling month/year
-- NAO value at previous sampling month/year
-- NAO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)
-
-Load the prepared nao_prev.csv file with has the amo at different year/month steps
-```{r}
-nao_prev <- read.csv("../output/env/nao_prev.csv", header = TRUE)
-head(nao_prev)
-```
-
-now match the year and the month in the two dataframes, and populate the background points with nao_sample
-
-```{r}
-background_all <- merge(x = background_all, y = nao_prev[ , c("year", "month", "nao_sample")], by = c("year", "month"), all.x = TRUE)
-background_all
-```
-
-ok next - NAO value at previous sampling month/year
-```{r}
-background_all <- merge(x = background_all, y = nao_prev[ , c("year", "month", "nao_prev")], by = c("year", "month"), all.x = TRUE)
-background_all
-```
-
-and now - NAO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn). The data is in a different .CSV
-
-```{r}
-nao_winter <- read.csv("../output/env/nao_winter.csv")
-background_all <- merge(x = background_all, y = nao_winter[ , c("year", "WinterAvg.Shifted")], by = c("year"), all.x = TRUE)
-colnames(background_all)[colnames(background_all)=="WinterAvg.Shifted"] <- "nao_winter" #change the colname
-background_all
-```
-
-and write .csv
-```{r}
-write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao.csv", row.names = FALSE)
-```
-
-still need
--bottom_depth
--depth_layer
--total_cell_obs
--yymm_cell_obs
-
-#depth layer
-
-the depth layers in the netCDF are literal slices.
-Currently I have listed the depthlayerno (which is literally which layer in the netcdf).
-I should also add the depth_layer name which gives the depth the slice is supposed to represent.
-Data is stored in depth_bin.csv
-
-load the depth bin csv
-```{r}
-depthbin <- read.csv("../data/env/depth_bin.csv", header = TRUE) 
-head(depthbin)
-```
-
-so basically what i want to do is, where depthbin$layer_no = background_all$depthlayerno, extract depthbin$layer_name and add to background_all$depth_layer. Can use the merge function.
-First need to change depthbin col names to match background_all
-
-```{r}
-names(depthbin)[names(depthbin)=="layer_no"] <- "depthlayerno"
-names(depthbin)[names(depthbin)=="layer_name"] <- "depth_layer" #this one is just so i don't have to change the column name in the background_all file later
-head(depthbin)
-```
-```{r}
-background_all <- merge(background_all, y = depthbin[ , c("depthlayerno", "depth_layer")], by = "depthlayerno", all.x = TRUE)
-head(background_all)
-```
-
-just check some random rows...
-
-```{r}
-randcheck <- background_all[sample(nrow(background_all), 10), ]
-randcheck
-```
-all good - save to .csv
-
-```{r}
-write.csv(background_all, "../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname.csv", row.names = FALSE)
-```
-
-# Bathymetric
-
-bottom depth from [GEBCO](www.gebco.net)
-
-```{r}
-bat <-readGEBCO.bathy("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
-summary(bat)
-```
-
-create pretty custom color palette and then plot the bathy layer with species data
-```{r}
-blues <- colorRampPalette(c("lightblue", "cadetblue2", "cadetblue1", "white"))
-plot(bat, n = 1, image = TRUE, bpal = blues(100))
-points(background_all$decimalLongitude, background_all$decimalLatitude, pch = 20, col = "yellow") #note these are both the wgs84 images
-```
-lovely jubbely
-
-and now get depth... note all the points and bathy data is wgs!
-
-```{r}
-background_all$bottom_depth <- get.depth(bat, background_all[ ,8:9], locator = FALSE)
-background_all$bottom_depth <- background_all$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth.
-This is extracting the depth column and deplacing data_aea$bottom_depth with the correct data.
-write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath.csv", row.names = FALSE)
-head(background_all)
-```
-
-# observations by cell (in total, and also by time step)
-
-as the title says... how many observations happeend in each cell:
-- in total
-
-```{r observation by cell - total}
-obs_by_cell_back <- count(background_all, "cell_id")
-write.csv(obs_by_cell_back, file = "../output/bio/back_no_observations_cell.csv")
-head(obs_by_cell_back)
-```
-
-- in each yyyymm
-```{r}
-obs_cell_yymm_back <- count(background_all, c("cell_id", "year", "month"))
-write.csv(obs_cell_yymm_back, file = "../output/bio/back_no_observations_cell_yymm.csv")
-head(obs_cell_yymm_back)
-```
-
-- in each yyyymm at depth
-```{r}
-obs_cell_yymm_depth_back <- count(background_all, c("cell_id", "year", "month", "depthlayerno"))
-write.csv(obs_cell_yymm_depth_back, file = "../output/bio/back_no_observations_cell_yymm_depth.csv")
-head(obs_cell_yymm_depth_back)
-```
-
-Now attach these values
-
-total obs for the whole time period
-```{r}
-background_all <- merge(background_all, obs_by_cell_back, by="cell_id")
-head(background_all)
-```
-
-rename freq column to something meaningful
-```{r}
-names(background_all)[names(background_all)=="freq"] <- "total_cell_obs_xy"
-head(background_all)
-```
-
-just check some random rows...
-
-```{r}
-randcheckcell <- background_all[sample(nrow(background_all), 10), ]
-randcheckcell
-```
-all good
-
-time-sliced
-```{r}
-background_all <- merge(background_all, obs_cell_yymm_back, by.x=c("cell_id", "year", "month"), by.y=c("cell_id", "year", "month"))
-names(background_all)[names(background_all)=="freq"] <- "total_cell_obs_xyt"
-background_all <- merge(background_all, obs_cell_yymm_depth_back, by.x=c("cell_id", "year", "month", "depthlayerno"), by.y=c("cell_id", "year", "month", "depthlayerno"))
-names(background_all)[names(background_all)=="freq"] <- "total_cell_obs_xyzt"
-write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath_cellobs.csv", row.names = FALSE)
-head(background_all)
-```
-
-and just check some random rows...
-
-```{r}
-randcheckyymm <- background_all[sample(nrow(background_all), 10), ]
-randcheckyymm
-```
-
-# attach environmental correlates
+# the inefficent loop
 
@@
 ```
-
-and just check some random rows...
-
-```{r}
-randcheck <- background_all[sample(nrow(background_all), 5), ]
-randcheck
-```
-
-```{r}
-randchkxy <- randcheck[ , c("longitude_meters", "latitude_meters")] 
-randchsp <- SpatialPointsDataFrame(coords = randchkxy, data = randcheck, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
-
-#row 1
-temp_brick <-  brick("../data/env/netcdf/2002_03_chl.nc", lvar = 4)
-t200203chl <- projectRaster(temp_brick, aea)
-chls <- extract(x=t200203chl[[1]], y = randchsp[1, ]) 
-chls
-chld <- extract(x=t200203chl[[23]], y = randchsp[1, ]) 
-chld 
-temp_brick <-  brick("../data/env/netcdf/2002_03_mlp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-mlps <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
-mlps
-temp_brick <-  brick("../data/env/netcdf/2002_03_o2.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-o2s <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
-o2s
-o2d <- extract(x=tempbrick[[23]], y = randchsp[1, ]) 
-o2d
-temp_brick <-  brick("../data/env/netcdf/2002_03_salinity.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sals <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
-sals
-sald <- extract(x=tempbrick[[23]], y = randchsp[1, ]) 
-sald
-temp_brick <-  brick("../data/env/netcdf/2002_03_temp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-temps <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
-temps
-tempd <- extract(x=tempbrick[[23]], y = randchsp[1, ]) 
-tempd
-temp_brick <-  brick("../data/env/netcdf/2002_03_ssh.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sshs <- extract(x=tempbrick[[1]], y = randchsp[1, ]) 
-sshs
-
-#row 2
-temp_brick <-  brick("../data/env/netcdf/2007_07_chl.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-tempbrick
-chls <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
-chls
-chld <- extract(x=tempbrick[[13]], y = randchsp[2, ]) 
-chld 
-temp_brick <-  brick("../data/env/netcdf/2007_07_mlp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-mlps <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
-mlps
-temp_brick <-  brick("../data/env/netcdf/2007_07_o2.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-o2s <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
-o2s
-o2d <- extract(x=tempbrick[[13]], y = randchsp[2, ]) 
-o2d
-temp_brick <-  brick("../data/env/netcdf/2007_07_salinity.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sals <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
-sals
-sald <- extract(x=tempbrick[[13]], y = randchsp[2, ]) 
-sald
-temp_brick <-  brick("../data/env/netcdf/2007_07_temp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-temps <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
-temps
-tempd <- extract(x=tempbrick[[13]], y = randchsp[2, ]) 
-tempd
-temp_brick <-  brick("../data/env/netcdf/2007_07_ssh.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sshs <- extract(x=tempbrick[[1]], y = randchsp[2, ]) 
-sshs
-
-#row 3
-temp_brick <-  brick("../data/env/netcdf/2011_08_chl.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-tempbrick
-chls <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
-chls
-chld <- extract(x=tempbrick[[13]], y = randchsp[3, ]) 
-chld 
-temp_brick <-  brick("../data/env/netcdf/2011_08_mlp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-mlps <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
-mlps
-temp_brick <-  brick("../data/env/netcdf/2011_08_o2.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-o2s <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
-o2s
-o2d <- extract(x=tempbrick[[13]], y = randchsp[3, ]) 
-o2d
-temp_brick <-  brick("../data/env/netcdf/2011_08_salinity.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sals <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
-sals
-sald <- extract(x=tempbrick[[13]], y = randchsp[3, ]) 
-sald
-temp_brick <-  brick("../data/env/netcdf/2011_08_temp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-temps <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
-temps
-tempd <- extract(x=tempbrick[[13]], y = randchsp[3, ]) 
-tempd
-temp_brick <-  brick("../data/env/netcdf/2011_08_ssh.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sshs <- extract(x=tempbrick[[1]], y = randchsp[3, ]) 
-sshs
-
-#row 4
-temp_brick <-  brick("../data/env/netcdf/2000_05_chl.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-tempbrick
-chls <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
-chls
-chld <- extract(x=tempbrick[[2]], y = randchsp[4, ]) 
-chld 
-temp_brick <-  brick("../data/env/netcdf/2000_05_mlp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-mlps <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
-mlps
-temp_brick <-  brick("../data/env/netcdf/2000_05_o2.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-o2s <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
-o2s
-o2d <- extract(x=tempbrick[[2]], y = randchsp[4, ]) 
-o2d
-temp_brick <-  brick("../data/env/netcdf/2000_05_salinity.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sals <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
-sals
-sald <- extract(x=tempbrick[[2]], y = randchsp[4, ]) 
-sald
-temp_brick <-  brick("../data/env/netcdf/2000_05_temp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-temps <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
-temps
-tempd <- extract(x=tempbrick[[2]], y = randchsp[4, ]) 
-tempd
-temp_brick <-  brick("../data/env/netcdf/2000_05_ssh.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sshs <- extract(x=tempbrick[[1]], y = randchsp[4, ]) 
-sshs
-
-#row 5
-temp_brick <-  brick("../data/env/netcdf/2012_06_chl.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-tempbrick
-chls <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
-chls
-chld <- extract(x=tempbrick[[25]], y = randchsp[5, ]) 
-chld 
-temp_brick <-  brick("../data/env/netcdf/2012_06_mlp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-mlps <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
-mlps
-temp_brick <-  brick("../data/env/netcdf/2012_06_o2.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-o2s <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
-o2s
-o2d <- extract(x=tempbrick[[25]], y = randchsp[5, ]) 
-o2d
-temp_brick <-  brick("../data/env/netcdf/2012_06_salinity.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sals <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
-sals
-sald <- extract(x=tempbrick[[25]], y = randchsp[5, ]) 
-sald
-temp_brick <-  brick("../data/env/netcdf/2012_06_temp.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-temps <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
-temps
-tempd <- extract(x=tempbrick[[25]], y = randchsp[5, ]) 
-tempd
-temp_brick <-  brick("../data/env/netcdf/2012_06_ssh.nc", lvar = 4)
-tempbrick <- projectRaster(temp_brick, aea)
-sshs <- extract(x=tempbrick[[1]], y = randchsp[5, ]) 
-sshs
-```
-
-it worked!
-
-# observation occurrence yes or no
-
-create a new column for Occurence, and populate with "0" (because these are all background)
-```{r}
-background_all$occurrence <- "0"
-write.csv(background_all, "../output/bio/background_complete_obs.csv", row.names = FALSE)
-head(background_all)
-```

### 2018-12-03 src/parallel_env_data.Rmd
add in test data

@@
   html_notebook: 
-    fig_height: 7
-    fig_width: 10
 editor_options: 
@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("plyr", "graphics", "ggplot2", "corrplot") 
+packages <- c("ncdf4", "raster", "foreach", "doParallel") 
 source("../src/ipak.R")
 ipak(packages)
+```
+
+# this test will use a shorter dataset and a smaller number of netcdfs
+
+```{r}
+test_back <- read.csv("../data/looptest.csv", header = TRUE) #test dataset
+
 ```
@@
 
-First you need to put the background (background_all) into a spatialpointsdataframe.
 ```{r}
-xy <- background_all[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
-background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
-head(background_all_sp)
-```
-
-```{r}
-netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+strt <- Sys.time() #get the start time
+netcdf_list <- list.files("../data/testloopnetcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
 no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
@@
   temp_brick <- projectRaster(temp_brick, aea) 
-    for (j in 1:nrow(background_all_sp)) {  
-      de <- background_all_sp$depthlayerno[[j]]  # a variable for the observation depth layer
-      yr <- (background_all_sp$year[j])  # a variable for the observation year
-      mth <- (background_all_sp$month[j])  # a variable for the observation month
+    for (j in 1:nrow(test_back)) {  
+      de <- test_back$depthlayerno[[j]]  # a variable for the observation depth layer
+      yr <- (test_back$year[j])  # a variable for the observation year
+      mth <- (test_back$month[j])  # a variable for the observation month
           if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
-              background_all_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
+              test_back$temp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
               if (is.na(de)){
-                background_all_sp$temp_depth[j] <- NA
+                test_back$temp_depth[j] <- NA
               } else  
-                background_all_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ])
+                test_back$temp_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ])
           } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
-              background_all_sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
+              test_back$salinity_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
               if (is.na(de)){
-                background_all_sp$salinity_depth[j] <- NA
+                test_back$salinity_depth[j] <- NA
               } else  
-                background_all_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
+                test_back$salinity_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
           } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
-              background_all_sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
+              test_back$chl_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
               if (is.na(de)){
-                background_all_sp$chl_depth[j] <- NA
+                test_back$chl_depth[j] <- NA
               } else  
-                background_all_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
+                test_back$chl_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
           } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
-              background_all_sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
+              test_back$o2_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
               if (is.na(de)){
-                background_all_sp$o2_depth[j] <- NA
+                test_back$o2_depth[j] <- NA
               } else  
-                background_all_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
+                test_back$o2_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
           } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
-              background_all_sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ])
+              test_back$mlp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])
           } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
-              background_all_sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ])
+              test_back$ssh_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])
 
@@
 }
-write.csv(background_all_sp, "../output/bio/background_complete.csv", row.names = FALSE)
-background_all <- as.data.frame(background_all_sp)
-head(background_all)
+write.csv(test_back, "../data/testloopoutput.csv", row.names = FALSE)
+test_back_df <- as.data.frame(test_back)
+head(test_back_df)
+print(Sys.time()-strt) #time it took to run
 ```

### 2018-12-03 src/parallel_env_data.Rmd
not working

@@
 ```{r}
-test_back <- read.csv("../data/looptest.csv", header = TRUE) #test dataset
-
+test_back_or <- read.csv("../data/looptest.csv", header = TRUE) #test dataset
+xy <- test_back_or[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+test_back <- SpatialPointsDataFrame(coords = xy, data = test_back_or, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
 ```
@@
 no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
-netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+netcdf_name <- list.files("../data/testloopnetcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
 aea <- raster("../output/env/aea.tif") 
@@
 ```
+
+# edit test 1
+
+```{r}
+test_back_or <- read.csv("../data/looptest.csv", header = TRUE) #test dataset
+xy <- test_back_or[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+test_back <- SpatialPointsDataFrame(coords = xy, data = test_back_or, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+
+ncores <- detectCores()
+cl <- (ncores)-1
+registerDoParallel(cl)
+
+strt <- Sys.time() #get the start time
+
+netcdf_list <- list.files("../data/testloopnetcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/testloopnetcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+aea <- raster("../output/env/aea.tif") 
+foreach ((i = 1:no_netcdf) , .packages = "raster") %dopar% {  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    foreach (j = 1:nrow(test_back)) %dopar% {  
+      de <- test_back$depthlayerno[[j]]  # a variable for the observation depth layer
+      yr <- (test_back$year[j])  # a variable for the observation year
+      mth <- (test_back$month[j])  # a variable for the observation month
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+              test_back$temp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
+              if (is.na(de)){
+                test_back$temp_depth[j] <- NA
+              } else  
+                test_back$temp_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              test_back$salinity_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
+              if (is.na(de)){
+                test_back$salinity_depth[j] <- NA
+              } else  
+                test_back$salinity_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
+              test_back$chl_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
+              if (is.na(de)){
+                test_back$chl_depth[j] <- NA
+              } else  
+                test_back$chl_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
+              test_back$o2_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
+              if (is.na(de)){
+                test_back$o2_depth[j] <- NA
+              } else  
+                test_back$o2_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
+              test_back$mlp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
+              test_back$ssh_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])
+
+          }
+
+    }
+}
+write.csv(test_back, "../data/testloopoutput.csv", row.names = FALSE)
+test_back_df <- as.data.frame(test_back)
+head(test_back_df)
+stopCluster(cl)
+print(Sys.time()-strt) #time it took to run
+```

### 2018-12-04 outsideproject.Rmd
background point generation with different numbers on same year

---
title: "parallel for the env. data grab"
author: "Samantha Andrews"
output: 
  html_notebook: 
editor_options: 
  chunk_output_type: inline
---

# Overview
Attempt to run my inefficinet giant loop for grabbing env. data to the background points in parallel (need to run again AND maybe need to do more points that will currently take 2 months

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there are probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
The "include=FALSE" supresses the package installation text appearing in the document...
```{r pre-install packages, include=FALSE}
packages <- c("ncdf4", "raster", "foreach", "doParallel", "doMC") 
source("C:/Users/bunny/Documents/GitHub/Chapter_1/src/ipak.R")
ipak(packages)
```

# this test will use a shorter dataset and a smaller number of netcdfs

```{r}
test_back_or <- read.csv("C:/Users/bunny/Documents/GitHub/Chapter_1/data/looptest.csv", header = TRUE) #test dataset
xy <- test_back_or[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
test_back <- SpatialPointsDataFrame(coords = xy, data = test_back_or, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
```

# the inefficent loop

This loop runs through the netcdf files and then looks for which rows in data_aea it should extract the value to point from, and at what depth (netcDF layer)

```{r}

netcdf_list <- list.files("C:/Users/bunny/Documents/GitHub/Chapter_1/data/testloopnetcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("C:/Users/bunny/Documents/GitHub/Chapter_1/data/testloopnetcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("C:/Users/bunny/Documents/GitHub/Chapter_1/output/env/aea.tif") 
for (i in 1:no_netcdf) {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
  for (j in 1:nrow(test_back)) {  
      de <- test_back$depthlayerno[[j]]  # a variable for the observation depth layer
      yr <- (test_back$year[j])  # a variable for the observation year
      mth <- (test_back$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
              test_back$temp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$temp_depth[j] <- NA
              } else  
                test_back$temp_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
              test_back$salinity_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$salinity_depth[j] <- NA
              } else  
                test_back$salinity_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
              test_back$chl_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$chl_depth[j] <- NA
              } else  
                test_back$chl_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
              test_back$o2_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$o2_depth[j] <- NA
              } else  
                test_back$o2_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
              test_back$mlp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
              test_back$ssh_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])

          }

    }
}
write.csv(test_back, "C:/Users/bunny/Documents/GitHub/Chapter_1/data/testloopoutput.csv", row.names = FALSE)
test_back_df <- as.data.frame(test_back)
head(test_back_df)

```

# edit test 1

```{r}
test_back_or <- read.csv("C:/Users/bunny/Documents/GitHub/Chapter_1/data/looptest.csv", header = TRUE) #test dataset
xy <- test_back_or[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
test_back <- SpatialPointsDataFrame(coords = xy, data = test_back_or, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.

ncores <- detectCores()
cl <- (ncores)-1
registerDoParallel(cl)

strt <- Sys.time() #get the start time

netcdf_list <- list.files("C:/Users/bunny/Documents/GitHub/Chapter_1/data/testloopnetcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("C:/Users/bunny/Documents/GitHub/Chapter_1/data/testloopnetcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("C:/Users/bunny/Documents/GitHub/Chapter_1/output/env/aea.tif") 
foreach ((i = 1:no_netcdf) , .packages = c("raster", "ncdf4")) %do% {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(test_back))  {  
      de <- test_back$depthlayerno[[j]]  # a variable for the observation depth layer
      yr <- (test_back$year[j])  # a variable for the observation year
      mth <- (test_back$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
              test_back$temp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$temp_depth[j] <- NA
              } else  
                test_back$temp_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
              test_back$salinity_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$salinity_depth[j] <- NA
              } else  
                test_back$salinity_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
              test_back$chl_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$chl_depth[j] <- NA
              } else  
                test_back$chl_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
              test_back$o2_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$o2_depth[j] <- NA
              } else  
                test_back$o2_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
              test_back$mlp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
              test_back$ssh_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])

          }

    }
}
write.csv(test_back, "C:/Users/bunny/Documents/GitHub/Chapter_1/data/testloopoutput.csv", row.names = FALSE)
test_back_df <- as.data.frame(test_back)
head(test_back_df)
stopCluster(cl)
print(Sys.time()-strt) #time it took to run
```

# the inefficent loop

This loop runs through the netcdf files and then looks for which rows in data_aea it should extract the value to point from, and at what depth (netcDF layer)

```{r}
test_back_or <- read.csv("C:/Users/bunny/Documents/GitHub/Chapter_1/data/looptest.csv", header = TRUE) #test dataset
xy <- test_back_or[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
test_back <- SpatialPointsDataFrame(coords = xy, data = test_back_or, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.

netcdf_list <- list.files("C:/Users/bunny/Documents/GitHub/Chapter_1/data/testloopnetcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("C:/Users/bunny/Documents/GitHub/Chapter_1/data/testloopnetcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("C:/Users/bunny/Documents/GitHub/Chapter_1/output/env/aea.tif") 
for (i in 1:no_netcdf) {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
  for (j in 1:nrow(test_back)) {  
      de <- test_back$depthlayerno[[j]]  # a variable for the observation depth layer
      yr <- (test_back$year[j])  # a variable for the observation year
      mth <- (test_back$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
              test_back$temp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$temp_depth[j] <- NA
              } else  
                test_back$temp_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
              test_back$salinity_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$salinity_depth[j] <- NA
              } else  
                test_back$salinity_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
              test_back$chl_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$chl_depth[j] <- NA
              } else  
                test_back$chl_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
              test_back$o2_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$o2_depth[j] <- NA
              } else  
                test_back$o2_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
              test_back$mlp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
              test_back$ssh_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])

          }

    }
}
write.csv(test_back, "C:/Users/bunny/Documents/GitHub/Chapter_1/data/testloopoutput.csv", row.names = FALSE)
test_back_df <- as.data.frame(test_back)
head(test_back_df)

```

### 2018-12-04 src/background_points_no_check.Rmd
background point generation with different numbers on same year

---
title: "background points - is more better?"
author: "Samantha Andrews"
output: 
  html_notebook: 
editor_options: 
  chunk_output_type: inline
---

# Overview
quick test to see what happens to distribution of env. variables if add more background points (is 10k enough or do I need more)

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there are probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
The "include=FALSE" supresses the package installation text appearing in the document...
```{r pre-install packages, include=FALSE}
packages <- c("ncdf4", "raster", "ggplot2") 
source("../src/ipak.R")
ipak(packages)
```

load one of the raw background.csv files to randomly extract points from (note that these files do not include presence cells, but will do for this test)

```{r}
testbglist <- read.csv("../output/bio/background/raw/2006.2.csv", header = TRUE)
head(testbglist)
```

Now create a bunch of dataframes with different no of random cells selected

```{r}
testbk10000 <- testbglist[sample(nrow(testbglist), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
testbk20000 <- testbglist[sample(nrow(testbglist), 20000), ]  
testbk30000 <- testbglist[sample(nrow(testbglist), 30000), ]  
testbk50000 <- testbglist[sample(nrow(testbglist), 50000), ]  
testbk100000 <- testbglist[sample(nrow(testbglist), 100000), ]  
```

# this test will use a shorter dataset and a smaller number of netcdfs
# the inefficent loop

This loop runs through the netcdf files and then looks for which rows in data_aea it should extract the value to point from, and at what depth (netcDF layer)

```{r}
strt <- Sys.time() #get the start time

xy <- testbk100000[ ,c("longitude_","latitude_m")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
testbk100000sp <- SpatialPointsDataFrame(coords = xy, data = testbk100000, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.

netcdf_list <- list.files("../data/bktstncdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/bktstncdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("../output/env/aea.tif") 
yr <- 2007  # a variable for the observation year
mth <- 10  # a variable for the observation month

for (i in 1:no_netcdf) {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(testbk100000sp)) {  
      de <- testbk100000sp$depthlayerno[[j]]  # a variable for the observation depth layer
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
              testbk100000sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ]) 
              if (is.na(de)){
                testbk100000sp$temp_depth[j] <- NA
              } else  
                testbk100000sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = testbk100000sp[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
              testbk100000sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ]) 
              if (is.na(de)){
                testbk100000sp$salinity_depth[j] <- NA
              } else  
                testbk100000sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = testbk100000sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
              testbk100000sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ]) 
              if (is.na(de)){
                testbk100000sp$chl_depth[j] <- NA
              } else  
                testbk100000sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = testbk100000sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
              testbk100000sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ]) 
              if (is.na(de)){
                testbk100000sp$o2_depth[j] <- NA
              } else  
                testbk100000sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = testbk100000sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
              testbk100000sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
              testbk100000sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ])

          }

    }
}
write.csv(testbk100000sp, "../data/testbk100000output.csv", row.names = FALSE)
test_back_df <- as.data.frame(testbk100000sp)
head(test_back_df)
print(Sys.time()-strt) #time it took to run
```

plot each variable against the different no of backgrounn points
```{r}
test10000 <- as.data.frame(testbk10000sp)
test20000 <- as.data.frame(testbk20000sp)
test50000 <- as.data.frame(testbk50000sp)
#test100000 <- as.data.frame(testbk10000sp)
ggplot(test10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green")
dev.copy(png,"../data/ssh_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(test10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green")
dev.copy(png,"../data/mlp_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

### 2018-12-04 src/background_points_no_check.Rmd
plots of background at diff no

@@
 
+SAM - A BETTER WAY WOULD HAVE BEEN TO RUN THE LOOP ON ALL THE POINTS, THEN RANDOMLY SELECT (QUICKER)
+
 ```{r}
@@
 testbk100000 <- testbglist[sample(nrow(testbglist), 100000), ]  
+testbk190000 <- testbglist[sample(nrow(testbglist), 190000), ]  
 ```
@@
 
-xy <- testbk100000[ ,c("longitude_","latitude_m")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
-testbk100000sp <- SpatialPointsDataFrame(coords = xy, data = testbk100000, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+xy <- testbk190000[ ,c("longitude_","latitude_m")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+testbk190000sp <- SpatialPointsDataFrame(coords = xy, data = testbk190000, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
 
@@
   temp_brick <- projectRaster(temp_brick, aea) 
-    for (j in 1:nrow(testbk100000sp)) {  
-      de <- testbk100000sp$depthlayerno[[j]]  # a variable for the observation depth layer
+    for (j in 1:nrow(testbk190000sp)) {  
+      de <- testbk190000sp$depthlayerno[[j]]  # a variable for the observation depth layer
           if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
-              testbk100000sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ]) 
+              testbk190000sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ]) 
               if (is.na(de)){
-                testbk100000sp$temp_depth[j] <- NA
+                testbk190000sp$temp_depth[j] <- NA
               } else  
-                testbk100000sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = testbk100000sp[j, ])
+                testbk190000sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = testbk190000sp[j, ])
           } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
-              testbk100000sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ]) 
+              testbk190000sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ]) 
               if (is.na(de)){
-                testbk100000sp$salinity_depth[j] <- NA
+                testbk190000sp$salinity_depth[j] <- NA
               } else  
-                testbk100000sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = testbk100000sp[j, ]) 
+                testbk190000sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = testbk190000sp[j, ]) 
           } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
-              testbk100000sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ]) 
+              testbk190000sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ]) 
               if (is.na(de)){
-                testbk100000sp$chl_depth[j] <- NA
+                testbk190000sp$chl_depth[j] <- NA
               } else  
-                testbk100000sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = testbk100000sp[j, ]) 
+                testbk190000sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = testbk190000sp[j, ]) 
           } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
-              testbk100000sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ]) 
+              testbk190000sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ]) 
               if (is.na(de)){
-                testbk100000sp$o2_depth[j] <- NA
+                testbk190000sp$o2_depth[j] <- NA
               } else  
-                testbk100000sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = testbk100000sp[j, ]) 
+                testbk190000sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = testbk190000sp[j, ]) 
           } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
-              testbk100000sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ])
+              testbk190000sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ])
           } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
-              testbk100000sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = testbk100000sp[j, ])
+              testbk190000sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ])
 
@@
 }
-write.csv(testbk100000sp, "../data/testbk100000output.csv", row.names = FALSE)
-test_back_df <- as.data.frame(testbk100000sp)
-head(test_back_df)
+write.csv(testbk190000sp, "../data/testbk190000output.csv", row.names = FALSE)
+#test_back_df <- as.data.frame(testbk100000sp)
+#head(test_back_df)
 print(Sys.time()-strt) #time it took to run
@@
 plot each variable against the different no of backgrounn points
+
 ```{r}
@@
 test50000 <- as.data.frame(testbk50000sp)
-#test100000 <- as.data.frame(testbk10000sp)
-ggplot(test10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green")
+test100000 <- as.data.frame(testbk10000sp)
+
+```
+
+```{r}
+ggplot(test10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
 dev.copy(png,"../data/ssh_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green")
+ggplot(test10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
 dev.copy(png,"../data/mlp_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
+
+ggplot(test10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+dev.copy(png,"../data/temp_surface_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(test10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+dev.copy(png,"../data/temp_depth_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(test10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+dev.copy(png,"../data/salinity_surface_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(test10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+dev.copy(png,"../data/salinity_depth_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(test10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+dev.copy(png,"../data/chl_surface_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(test10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+dev.copy(png,"../data/chl_depth_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(test10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+dev.copy(png,"../data/o2_surface_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(test10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+dev.copy(png,"../data/o2_depth_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
 ```

### 2018-12-05 src/background_points_no_check.Rmd
multiple month background point check

@@
 test100000 <- as.data.frame(testbk10000sp)
-
+test190000 <- as.data.frame(testbk190000sp)
 ```
@@
 ```{r}
-ggplot(test10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/ssh_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/mlp_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/temp_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/temp_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/salinity_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/salinity_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/chl_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/chl_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/o2_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange")
+ggplot(test10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../data/o2_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 ```
+
+Try again for another month - say 1999 02 AND again 2014 06
+
+This time extract the values for all points and then subset
+
+```{r}
+strt <- Sys.time() #get the start time
+
+xy <- testbglist[ ,c("longitude_","latitude_m")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+testbglistsp <- SpatialPointsDataFrame(coords = xy, data = testbglist, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+
+netcdf_list <- list.files("../data/bktstncdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/bktstncdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+aea <- raster("../output/env/aea.tif") 
+yr <- 1999  # a variable for the observation year
+mth <- 02  # a variable for the observation month
+
+for (i in 1:no_netcdf) {  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(testbglistsp)) {  
+      de <- testbglistsp$depthlayerno[[j]]  # a variable for the observation depth layer
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+              testbglistsp$temp_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
+              if (is.na(de)){
+                testbglistsp$temp_depth[j] <- NA
+              } else  
+                testbglistsp$temp_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              testbglistsp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
+              if (is.na(de)){
+                testbglistsp$salinity_depth[j] <- NA
+              } else  
+                testbglistsp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
+              testbglistsp$chl_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
+              if (is.na(de)){
+                testbglistsp$chl_depth[j] <- NA
+              } else  
+                testbglistsp$chl_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
+              testbglistsp$o2_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
+              if (is.na(de)){
+                testbglistsp$o2_depth[j] <- NA
+              } else  
+                testbglistsp$o2_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
+              testbglistsp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
+              testbglistsp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ])
+
+          }
+
+    }
+}
+write.csv(testbglistsp, "../data/env/background_point_check/1999_02/testbglistoutput.csv", row.names = FALSE)
+test_back_df <- as.data.frame(testbglistsp)
+#head(test_back_df)
+print(Sys.time()-strt) #time it took to run
+```
+
+ok now create the first lot of random for 1999_02
+
+```{r}
+testbk100001999 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk200001999 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk300001999 <- test_back_df[sample(nrow(test_back_df), 30000), ]  
+testbk500001999 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk1000001999 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk1900001999 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+```
+
+and plot
+
+```{r}
+ggplot(testbk100001999, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/ssh_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100001999, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/mlp_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100001999, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/temp_surface_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100001999, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/temp_depth_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100001999, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/salinity_surface_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100001999, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/salinity_depth_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100001999, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/chl_surface_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100001999, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/chl_depth_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100001999, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500001999 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/o2_surface_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100001999, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500001999 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/o2_depth_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+And now 2014_06
+
+```{r}
+strt <- Sys.time() #get the start time
+
+xy <- testbglist[ ,c("longitude_","latitude_m")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+testbglistsp <- SpatialPointsDataFrame(coords = xy, data = testbglist, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+
+netcdf_list <- list.files("../data/bktstncdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/bktstncdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+aea <- raster("../output/env/aea.tif") 
+yr <- 2014  # a variable for the observation year
+mth <- 06  # a variable for the observation month
+
+for (i in 1:no_netcdf) {  
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(testbglistsp)) {  
+      de <- testbglistsp$depthlayerno[[j]]  # a variable for the observation depth layer
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+              testbglistsp$temp_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
+              if (is.na(de)){
+                testbglistsp$temp_depth[j] <- NA
+              } else  
+                testbglistsp$temp_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              testbglistsp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
+              if (is.na(de)){
+                testbglistsp$salinity_depth[j] <- NA
+              } else  
+                testbglistsp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
+              testbglistsp$chl_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
+              if (is.na(de)){
+                testbglistsp$chl_depth[j] <- NA
+              } else  
+                testbglistsp$chl_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
+              testbglistsp$o2_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
+              if (is.na(de)){
+                testbglistsp$o2_depth[j] <- NA
+              } else  
+                testbglistsp$o2_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
+              testbglistsp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
+              testbglistsp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ])
+
+          }
+
+    }
+}
+write.csv(testbglistsp, "../data/env/background_point_check/2014_06/testbglistoutput.csv", row.names = FALSE)
+test_back_df <- as.data.frame(testbglistsp)
+#head(test_back_df)
+print(Sys.time()-strt) #time it took to run
+```
+ok now create the first lot of random for 2014_06
+
+```{r}
+testbk100002014 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk200002014 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk300002014 <- test_back_df[sample(nrow(test_back_df), 30000), ]  
+testbk500002014 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk1000002014 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk1900002014 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+```
+
+and plot
+
+```{r}
+ggplot(testbk100002014, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/ssh_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100002014, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/mlp_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100002014, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/temp_surface_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100002014, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/temp_depth_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100002014, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/salinity_surface_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100002014, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/salinity_depth_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100002014, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/chl_surface_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100002014, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/chl_depth_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100002014, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/o2_surface_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(testbk100002014, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../data/o2_depth_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```

### 2018-12-05 src/background_points_no_check.Rmd
3d plot test

@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("ncdf4", "raster", "ggplot2") 
+packages <- c("ncdf4", "raster", "ggplot2", "plotly") 
 source("../src/ipak.R")
@@
 ```
+
+# 3d plot background points
+
+```{r}
+bck2014_06_3d <- plot_ly(x= testbk100002014$longitude_, y = testbk100002014$latitude_m, z = testbk100002014$depthlayerno)
+```

### 2018-12-06 src/timesliced_random_cell_selection.Rmd
line 210 - to redo background

@@
 
+## XXX SAM THIS IS THE CODE THAT NEEDS EDITING - I THINK JUST REMOVE THE LINE #
+
 ```{r timesliced unique unsampled}
@@
   time_cell <- unique_cell_xyz
-  time_cell <- time_cell[!time_cell$id_depth %in% timeslice$id_depth, , drop = FALSE]
+  #time_cell <- time_cell[!time_cell$id_depth %in% timeslice$id_depth, , drop = FALSE]
   write.csv(time_cell, paste0(bckoutput, names(lapply(split_cell_yymm, names))[i], ".csv"), row.names = FALSE) 

### 2018-12-06 src/timesliced_random_cell_selection.Rmd
update to timesclice to remove cell removal process

@@
 This is to help remove cells later and randomise
+write.csv(unique_cell_xyz, "../output/env/unique_cell_centroid_lonlat_nafo2_depth.csv", row.names = FALSE)
 ```
@@
 
-Split the data_aea into timeslices (year/month).
+Split the mcpbase into timeslices (year/month).
 All you need is the year, month, cell_id, and depthlayerno
 ```{r}
-data_aea <- mcpbase
-cell_yymm <- subset(data_aea, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month, depthlayer no
+cell_yymm <- subset(mcpbase, select= c(cell_id, depthlayerno, year, month)) #create a new df with just cell_id, yr, month, depthlayer no
 cell_yymm$id_depth <- paste(cell_yymm$cell_id, cell_yymm$depthlayerno, sep="") #create a new column that is a unique cell_id & depth ID. This is to help remove cells later and randomise

### 2018-12-06 src/timesliced_random_cell_selection.Rmd
changed no of random points to be selected

@@
   sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
-  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample (large sample as per maxent)
+  sel <- sample_n(sel, 20000) #where 20000 = number of rows to sample (more than 10k as some points will have NA values due to depth - I will select 10000 of these without NA values to create a large sample as per maxent)
   write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
@@
 
-check number of observations 
+check number of background points in total 
 ```{r}
@@
 write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath.csv", row.names = FALSE)
-head(background_all)
+head(background_all$bottom_depth)
 ```
@@
 ```
+
+```{r}
+comback <- read.csv("../output/bio/background_complete_obs_cels.csv", header = TRUE)
+backtempna <- subset(comback, is.na(comback$temp_depth))
+backtempna
+backtempnasplit <- split(backtempna, list(backtempna$year, backtempna$month), drop = TRUE)
+```

### 2018-12-13 src/timesliced_random_cell_selection.Rmd
someone turned off pc... blocking code

@@
 
-First you need to put the background (background_all) into a spatialpointsdataframe.
+First you need to create columns for the data to be added to
+
+```{r}
+background_all$temp_depth <- NA
+background_all$temp_surface <- NA
+background_all$salinity_depth <- NA
+background_all$salinity_surface <- NA
+background_all$chl_depth <- NA
+background_all$chl_surface <- NA
+background_all$o2_depth <- NA
+background_all$o2_surface <- NA
+background_all$mlp_surface <- NA
+background_all$ssh_surface <- NA
+```
+
+Second you need to put the background (background_all) into a spatialpointsdataframe.
 ```{r}
@@
 background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+head(background_all_sp)
+```
+
+background_all <- read.csv("./output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath_cellobs.csv", header = TRUE
+
+XXXXXXXXXXX split
+```{r}
+background_all2 <- background_all[background_all$year == 1999, ]
+xy <- background_all2[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
 head(background_all_sp)
@@
 }
-write.csv(background_all_sp, "../output/bio/background_complete.csv", row.names = FALSE)
-background_all <- as.data.frame(background_all_sp)
-head(background_all)
+write.csv(background_all_sp, "../output/bio/background_complete_1999.csv", row.names = FALSE)
+#background_all <- as.data.frame(background_all_sp)
+head(background_all_sp)
 ```

### 2018-12-13 src/background_points_no_check.Rmd

@@
 bck2014_06_3d <- plot_ly(x= testbk100002014$longitude_, y = testbk100002014$latitude_m, z = testbk100002014$depthlayerno)
+bck2014_06_3d
 ```
+
+# 2d plot background points
+
+```{r}
+bck2014_06_2d <- plot(x= testbk100002014$longitude_, y = testbk100002014$latitude_m)
+dev.copy(png,"../data/env/background_point_check/bck10000_2d.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+```

### 2018-12-13 src/data_exploration.Rmd
simple plots complete

@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("plyr", "graphics", "ggplot2", "corrplot") 
+packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA") 
 source("../src/ipak.R")
@@
 head(background)
+test <- background[245:250, ]
+test
 ```
@@
 ```
+
+# NAFO Regions
+
+compare the environmental correlates between different NAFO regions
+
+first see which nafo zones were sampled in each year
+```{r}
+nafo_by_yr <- with(presence, table(year, nafo_zone))
+write.csv(inst_by_yr, file = "../output/bio/nafozone_by_yr.csv")
+nafo_by_yr
+```
+
+and by month
+```{r}
+nafo_by_mth <- with(presence, table(month, nafo_zone))
+write.csv(inst_by_yr, file = "../output/bio/nafozone_by_mth.csv")
+nafo_by_mth
+```
+
+Interesting that there is a point in 1C - this is outside Canadian waters.... anyway
+```{r}
+onec <-  subset(presence, nafo_zone == "1C")
+onec
+```
+
+Let's plot the variables by nafo region/year then by month
+
+```{r}
+pr98 <- subset(presence, year == "1998")
+boxplot(pr98$temp_depth ~ pr98$nafo_zone, xlab = "NAFO Region", ylab = "Temperature (Kelvin)", main = "Temperature at Observation Depth per NAFO Zone")
+```
+
+```{r}
+pr98 <- subset(presence, year == "1998")
+pr11 <- subset(presence, year == "2011")
+par(mfrow=c(2,1))
+boxplot(pr98$temp_depth ~ pr98$nafo_zone, xlab = "NAFO Region", ylab = "Temperature (Kelvin)", main = "Temperature at Observation Depth per NAFO Zone")
+boxplot(pr11$temp_depth ~ pr11$nafo_zone, xlab = "NAFO Region", ylab = "Temperature (Kelvin)", main = "Temperature at Observation Depth per NAFO Zone")
+```
+
+ok scrap the nafo region analysis - it is too mixed up with month (so seeing if region or month is a factor that needs to be factored in to the model is not appropriate)
+
+# check for gear type
+
+what are the unique gear types you have in your presence data, and how many?
+
+```{r observation by cell - total}
+gear_count <- count(presence, "gear")
+write.csv(gear_count, file = "../output/bio/gear_count.csv")
+gear_count
+```
+
+so the vast majority are trawls of some type.
+
+```{r}
+par(mar=c(7,5,1,1))
+boxplot(presence$temp_depth ~ presence$gear, ylab = "Temperature (Kelvin)", main = "Temperature at Observation Depth by gear type", las = 2)
+dev.copy(png, "../output/env/simple_plots/temperature_boxplot_dept_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+looks like not much difference... what about a kruskal wallace test?
+
+```{r}
+kruskal.test(presence$temp_celsius_depth ~ presence$gear)
+```
+Ok so there is a statistically sig difference somewhere in the temp at depth reported by the gear type (Kruskal-Wallis chi-squared = 248.27, df = 7, p-value < 2.2e-16)
+
+To see where the difference(s) are run a Dunn test Zar (2010) states that the Dunn test (in the FSA package) is appropriate for groups with unequal numbers of observations.(Zar, J.H.
+2010.
+Biostatistical Analysis, 5th ed.
+Pearson Prentice Hall: Upper Saddle River, NJ.)
+
+SAM WHEN DUNN IS INSTALLED, GO TO RCOMPANIONS.ORG AND LOOK FOR DUNN - CHECK CHROME HISTORY (PROBABLY UNDER KRUSKAL WALLIS) ALSO MAP THE DISTRIBUTION OF THESE GEAR TYPES
+```{r}
+pairwise.wilcox.test(presence$temp_celsius_depth, presence$gear, p.adj='bonferroni', exact=F)
+```

### 2018-12-19 src/timesliced_random_cell_selection.Rmd
random cell generation (10k)

--- 2018-12-06 src/timesliced_random_cell_selection.Rmd
@@
 bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
-randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
+randbckoutput <- "../output/bio/background/rand20k/" #where the files are to be saved to
 # read in each .csv file in file_list and create a data frame with the same name as the .csv file
@@
 ```{r}
-rand_folder <- "../output/bio/background/rand/"      # path to folder that holds multiple .csv files
+rand_folder <- "../output/bio/background/rand20k/"      # path to folder that holds multiple .csv files
 rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
-randbckoutput <- "../output/bio/background/rand/" #where the files are to be saved to
+randbckoutput <- "../output/bio/background/rand20k/" #where the files are to be saved to
 for(i in 1:length(rand_list)){
@@
 
-First you need to put the background (background_all) into a spatialpointsdataframe.
+First you need to create columns for the data to be added to (otherwise the loop doesn't always work - notably on one PC but not another)
+
+```{r}
+background_all$temp_depth <- NA
+background_all$temp_surface <- NA
+background_all$salinity_depth <- NA
+background_all$salinity_surface <- NA
+background_all$chl_depth <- NA
+background_all$chl_surface <- NA
+background_all$o2_depth <- NA
+background_all$o2_surface <- NA
+background_all$mlp_surface <- NA
+background_all$ssh_surface <- NA
+```
+
+Second you need to put the background (background_all) into a spatialpointsdataframe.
 ```{r}
@@
 background_all <- as.data.frame(background_all_sp)
-head(background_all)
+head(background_all_sp)
 ```
@@
 
+# remove rows with NA values at depth
+
+first rbind... (this you can delete if you had run the code at once)
+```{r}
+back1998 <- read.csv("../output/bio/background_complete_1998.csv", header = TRUE)
+back1999 <- read.csv("../output/bio/background_complete_1999.csv", header = TRUE)
+back2000_2003 <- read.csv("../output/bio/background_complete_2000_2003.csv", header = TRUE)
+back2004_2006 <- read.csv("../output/bio/background_complete_2004_2006.csv", header = TRUE)
+back2007_2011 <- read.csv("../output/bio/background_complete_2007_2011.csv", header = TRUE)
+back2012_2016 <- read.csv("../output/bio/background_complete_2012_2016.csv", header = TRUE)
+
+background_all <- rbind(back1998, back1999, back2000_2003, back2004_2006, back2007_2011, back2012_2016)
+head(background_all)
+```
+
+remove all na values
+
+```{r}
+backna <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
+backna
+background_all[complete.cases(background_all), ] #removes NA values
+write.csv(background_all, "../output/bio/background_complete_na.csv", row.names = FALSE)
+```
+
+randomly select 10000 rows from each timeslice
+```{r}
+#background_all <- read.csv("../output/bio/background_complete_na.csv", header = TRUE)
+#split_cell_back <- split(background_all, list(background_all$year, background_all$month), drop = TRUE)
+
+no_timeslices <- length(split_cell_back) # how many levels (time slices) are in the list
+bckoutput <- "../output/bio/background/rand20k/no_na/" #where the files are to be saved to
+for (i in 1:no_timeslices){
+  timeslice <- split_cell_back[[i]]
+  timeslice <- as.data.frame(timeslice)
+  write.csv(timeslice, paste0(bckoutput, names(lapply(split_cell_back, names))[i], ".csv"), row.names = FALSE) 
+}
+```
+
+```{r}
+bckrnd_folder <- "../output/bio/background/rand20k/no_na/"      # path to folder that holds multiple .csv files
+bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
+randbckoutput <- ". ./output/bio/background/rand20k/no_na_10krand/" #where the files are to be saved to
+# read in each .csv file in file_list and create a data frame with the same name as the .csv file
+for (i in 1:length(bckrnd_list)){
+  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
+  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample 
+  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
+}
+head(sel)
+```
+
 # observation occurrence yes or no

### 2018-12-20 src/testing_point_generation.Rmd
background update

---
title: "pooint extraction test"
output: html_notebook
---

```{r}
testbackground_all <- read.csv("../output/bio/background/raw/2007.3.csv", header = TRUE)
testbackground_all
```

First you need to create columns for the data to be added to (otherwise the loop doesn't always work - notably on one PC but not another)

```{r}
testbackground_all$temp_depth <- NA
testbackground_all$temp_surface <- NA
testbackground_all$salinity_depth <- NA
testbackground_all$salinity_surface <- NA
testbackground_all$chl_depth <- NA
testbackground_all$chl_surface <- NA
testbackground_all$o2_depth <- NA
testbackground_all$o2_surface <- NA
testbackground_all$mlp_surface <- NA
testbackground_all$ssh_surface <- NA
```

Second you need to put the background (testbackground_all) into a spatialpointsdataframe.
```{r}
xy <- testbackground_all[ ,c("longitude_","latitude_m")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude

#xy <- testbackground_all[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
testbackground_all_sp <- SpatialPointsDataFrame(coords = xy, data = testbackground_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
head(testbackground_all_sp)
```

```{r}
netcdf_list <- list.files("../data/env/netcdfbase", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/netcdfbase", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("../output/env/aea.tif") 
for (i in 1:no_netcdf) {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(testbackground_all_sp)) {  
      de <- testbackground_all_sp$depthlayerno[[j]]  # a variable for the observation depth layer
          if (brkvar == "temp.nc"){
              if (is.na(de)){
                testbackground_all_sp$temp_depth[j] <- NA
              } else  
                testbackground_all_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = testbackground_all_sp[j, ])
          } else if (brkvar == "salinity.nc") {
              if (is.na(de)){
                testbackground_all_sp$salinity_depth[j] <- NA
              } else  
                testbackground_all_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = testbackground_all_sp[j, ]) 
          } else if (brkvar == "chl.nc") {
              if (is.na(de)){
                testbackground_all_sp$chl_depth[j] <- NA
              } else  
                testbackground_all_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = testbackground_all_sp[j, ]) 
          } else if (brkvar == "o2.nc") {
              if (is.na(de)){
                testbackground_all_sp$o2_depth[j] <- NA
              } else  
                testbackground_all_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = testbackground_all_sp[j, ]) 
          }

    }
}
write.csv(testbackground_all_sp, "../output/bio/background_point_na_removal.csv", row.names = FALSE)
testbackground_all <- as.data.frame(testbackground_all_sp)
head(testbackground_all_sp)
```

```{r}
backna <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
backna
background_all <- background_all[complete.cases(background_all), ] #removes NA values
backna2 <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
backna2
write.csv(background_all, "../output/bio/background_point_na_removal_nona", row.names = FALSE)
head(testbackground_all)
```

```{r}
testbackground_all <- subset(testbackground_all, select = -c(temp_depth, temp_surface, salinity_depth, salinity_surface, o2_depth, o2_surface, chl_depth, chl_surface, ssh_surface, mlp_surface))
head(testbackground_all)
```
```{r}
testbackground_all <- subset(testbackground_all, select = -c(longitude_.1, latitude_m.1, optional))
head(testbackground_all)
```

```{r}
colnames(testbackground_all)[colnames(testbackground_all)=="decimalLon"] <- "decimalLongitude"
colnames(testbackground_all)[colnames(testbackground_all)=="decimalLat"] <- "decimalLatitude"
colnames(testbackground_all)[colnames(testbackground_all)=="longitude_"] <- "longitude_meters"
colnames(testbackground_all)[colnames(testbackground_all)=="latitude_m"] <- "latitude_meters"
head(testbackground_all)
```

save as csv
```{r}
write.csv(testbackground_all, file = "../output/bio/background_point_na_removal_nona.csv", row.names = FALSE)
```

```{r timesliced unique unsampled}
no_timeslices <- length(split_cell_yymm) # how many levels (time slices) are in the list
bckoutput <- "../output/bio/background/raw/" #where the files are to be saved to
for (i in 1:no_timeslices){
  timeslice <- split_cell_yymm[[i]]
  timeslice <- as.data.frame(timeslice)
  time_cell <- unique_cell_xyz
  #time_cell <- time_cell[!time_cell$id_depth %in% timeslice$id_depth, , drop = FALSE]
  write.csv(time_cell, paste0(bckoutput, names(lapply(split_cell_yymm, names))[i], ".csv"), row.names = FALSE) 
}
```

```{r}
bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand10kreboot/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample 
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(sel)
```

```{r}
rand_folder <- "../output/bio/background/rand10kreboot/"      # path to folder that holds multiple .csv files
rand_list <- list.files(path = rand_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand10kreboot/" #where the files are to be saved to
for(i in 1:length(rand_list)){
  randtemp <- read.csv(paste(rand_folder, rand_list[i], sep=''))
  randtemp$year <- substr(rand_list[i], 1, 4) #the year is characters 1-4 in the filename
  randtemp$month <- substr(rand_list[i], 6, 7) #the month is characters 6 & 7 in the filename (except where the month is a single digit)
  write.csv(randtemp, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(randtemp)
```

### 2018-12-20 src/timesliced_random_cell_selection.Rmd
background update

@@
   sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
-  sel <- sample_n(sel, 20000) #where 20000 = number of rows to sample (more than 10k as some points will have NA values due to depth - I will select 10000 of these without NA values to create a large sample as per maxent)
+  sel <- sample_n(sel, 10000) #where 20000 = number of rows to sample (more than 10k as some points will have NA values due to depth - I will select 10000 of these without NA values to create a large sample as per maxent)
   write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
@@
 ```{r}
-xy <- background_all[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
-background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+background_all2 <- background_all[background_all$year == 2007, ]
+xy <- background_all2[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+
+#xy <- background_all[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
 head(background_all_sp)
@@
 }
-write.csv(background_all_sp, "../output/bio/background_complete.csv", row.names = FALSE)
+write.csv(background_all_sp, "../output/bio/background_complete_2007.csv", row.names = FALSE)
 background_all <- as.data.frame(background_all_sp)
@@
 backna
-background_all[complete.cases(background_all), ] #removes NA values
+background_all <- background_all[complete.cases(background_all), ] #removes NA values
+backna2 <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
+backna2
 write.csv(background_all, "../output/bio/background_complete_na.csv", row.names = FALSE)
@@
 #background_all <- read.csv("../output/bio/background_complete_na.csv", header = TRUE)
-#split_cell_back <- split(background_all, list(background_all$year, background_all$month), drop = TRUE)
+split_cell_back <- split(background_all, list(background_all$year, background_all$month), drop = TRUE)
 
@@
 bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
-randbckoutput <- ". ./output/bio/background/rand20k/no_na_10krand/" #where the files are to be saved to
+randbckoutput <- "../output/bio/background/rand20k/no_na_10krand/" #where the files are to be saved to
 # read in each .csv file in file_list and create a data frame with the same name as the .csv file
@@
 
+crap - the years from 2007 onwards have less than 10k points to choose from...
+create a random 30k for each of these month/year periods
+
+```{r}
+filenames  <- list.files(path = randbckoutput, pattern="*.csv", full.names = TRUE) 
+background_all <- rbindlist(lapply(filenames,fread, fill = TRUE))
+write.csv(background_all, "../output/bio/background_all_complete_na2.csv", row.names = FALSE)
+head(background_all)
+```
 # observation occurrence yes or no
@@
 ```{r}
-background_all <- read.csv("../output/bio/background_complete.csv")
 background_all$occurrence <- "0"

### 2018-12-20 src/testing_point_generation.Rmd
new background point generation  - text needs to be merged with actual code

@@
 ```{r timesliced unique unsampled}
+testbackground_all <- read.csv("../output/bio/background_point_na_removal_nona.csv", header = TRUE)
 no_timeslices <- length(split_cell_yymm) # how many levels (time slices) are in the list
@@
   timeslice <- as.data.frame(timeslice)
-  time_cell <- unique_cell_xyz
+  time_cell <- testbackground_all
   #time_cell <- time_cell[!time_cell$id_depth %in% timeslice$id_depth, , drop = FALSE]

### 2018-12-21 src/testing_point_generation.Rmd
repair 2...

@@
 ```{r}
-backna <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
+testbackground_all <- read.csv("../output/bio/background_point_na_removal.csv", header = TRUE)
+backna <- count(rowSums(!is.na(testbackground_all)))
+backna <- sum(!complete.cases(testbackground_all)) #this just counts the number of rows that contain NA values
 backna
-background_all <- background_all[complete.cases(background_all), ] #removes NA values
-backna2 <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
+testbackground_all <- testbackground_all[complete.cases(testbackground_all), ] #removes NA values
+backna2 <- sum(complete.cases(testbackground_all)) #this just counts the number of rows that contain NA values
 backna2
-write.csv(background_all, "../output/bio/background_point_na_removal_nona", row.names = FALSE)
+write.csv(testbackground_all, "../output/bio/background_point_na_removal_nona.csv", row.names = FALSE)
 head(testbackground_all)
@@
 ```{r}
-testbackground_all <- subset(testbackground_all, select = -c(temp_depth, temp_surface, salinity_depth, salinity_surface, o2_depth, o2_surface, chl_depth, chl_surface, ssh_surface, mlp_surface))
+testbackground_all <- subset(testbackground_all, select = -c(temp_depth, salinity_depth, o2_depth, chl_depth))
 head(testbackground_all)
@@
 ```{r}
-testbackground_all <- subset(testbackground_all, select = -c(longitude_.1, latitude_m.1, optional))
+testbackground_all <- subset(testbackground_all, select = -c(longitude_.1, latitude_m.1, optional, optional.1))
 head(testbackground_all)

### 2018-12-21 src/timesliced_random_cell_selection.Rmd
repair 2...

@@
 ```{r}
+rand_folder <- "../output/bio/background/rand10kreboot/"      # path to folder that holds multiple .csv files
 filenames  <- list.files(path = rand_folder, pattern="*.csv", full.names = TRUE) 
@@
 ```{r}
+background_all <- read.csv("../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname.csv", header = TRUE)
 background_all$bottom_depth <- get.depth(bat, background_all[ ,8:9], locator = FALSE)
@@
 write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath.csv", row.names = FALSE)
-head(background_all$bottom_depth)
+head(background_all)
 ```
@@
 write.csv(background_all_sp, "../output/bio/background_complete_2007.csv", row.names = FALSE)
-background_all <- as.data.frame(background_all_sp)
+background_all2 <- as.data.frame(background_all_sp)
 head(background_all_sp)
@@
 
-remove all na values
+remove all na values THIS CAN BE EDITED TO DOUBLE CHECK FOR NA VALUES
 
 ```{r}
-backna <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
+backna <- sum(!complete.cases(background_all2)) #this just counts the number of rows that contain NA values
 backna
-background_all <- background_all[complete.cases(background_all), ] #removes NA values
+background_all2 <- background_all2[complete.cases(background_all2), ] #removes NA values
 backna2 <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values

### 2019-01-08 src/background_points_no_check.Rmd
updated to use new background point list

@@
 
-load one of the raw background.csv files to randomly extract points from (note that these files do not include presence cells, but will do for this test)
+load the raw background.csv file with all points to randomly extract points from (../output/env/unique_cell_centroid_lonlat_nafo2_depth.csv)
 
 ```{r}
-testbglist <- read.csv("../output/bio/background/raw/2006.2.csv", header = TRUE)
+testbglist <- read.csv("../output/env/unique_cell_centroid_lonlat_nafo2_depth.csv", header = TRUE)
 head(testbglist)
@@
 
-Now create a bunch of dataframes with different no of random cells selected
-
-SAM - A BETTER WAY WOULD HAVE BEEN TO RUN THE LOOP ON ALL THE POINTS, THEN RANDOMLY SELECT (QUICKER)
-
 ```{r}
-testbk10000 <- testbglist[sample(nrow(testbglist), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
-testbk20000 <- testbglist[sample(nrow(testbglist), 20000), ]  
-testbk30000 <- testbglist[sample(nrow(testbglist), 30000), ]  
-testbk50000 <- testbglist[sample(nrow(testbglist), 50000), ]  
-testbk100000 <- testbglist[sample(nrow(testbglist), 100000), ]  
-testbk190000 <- testbglist[sample(nrow(testbglist), 190000), ]  
+backobsno <- nrow(testbglist)
+backobsno
 ```
 
-# this test will use a shorter dataset and a smaller number of netcdfs
 # the inefficent loop
@@
 
+Start with 2007_10 data
+
 ```{r}
@@
 
-xy <- testbk190000[ ,c("longitude_","latitude_m")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
-testbk190000sp <- SpatialPointsDataFrame(coords = xy, data = testbk190000, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+xy <- testbglist[ ,c("longitude_","latitude_m")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
+testbglistsp <- SpatialPointsDataFrame(coords = xy, data = testbglist, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
 
@@
 }
-write.csv(testbk190000sp, "../data/testbk190000output.csv", row.names = FALSE)
-#test_back_df <- as.data.frame(testbk100000sp)
-#head(test_back_df)
+write.csv(testbglistsp, "../data/env/background_point_check/200710_allbackgroundpoints.csv", row.names = FALSE)
+test_back_df <- as.data.frame(testbglistsp)
 print(Sys.time()-strt) #time it took to run
@@
 
-plot each variable against the different no of backgrounn points
+ok now create the first lot of random for 2007_10
 
 ```{r}
-test10000 <- as.data.frame(testbk10000sp)
-test20000 <- as.data.frame(testbk20000sp)
-test50000 <- as.data.frame(testbk50000sp)
-test100000 <- as.data.frame(testbk10000sp)
-test190000 <- as.data.frame(testbk190000sp)
+testbk100001999 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk200001999 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk300001999 <- test_back_df[sample(nrow(test_back_df), 30000), ]  
+testbk500001999 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk1000001999 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk1900001999 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
 ```
+
+plot each variable against the different no of background points
 
@@
 ggplot(test10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/ssh_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_ssh_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(test10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/mlp_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_mlp_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(test10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/temp_surface_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_temp_surface_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(test10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/temp_depth_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_temp_depth_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(test10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/salinity_surface_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_salinity_surface_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(test10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/salinity_depth_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_salinity_depth_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(test10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/chl_surface_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_chl_surface_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(test10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/chl_depth_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_chl_depth_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(test10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/o2_surface_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_o2_surface_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(test10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/o2_depth_back_no.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/200710_o2_depth_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 }
-write.csv(testbglistsp, "../data/env/background_point_check/1999_02/testbglistoutput.csv", row.names = FALSE)
+write.csv(testbglistsp, "../data/env/background_point_check/199902_allbackgroundpoints.csv", row.names = FALSE)
 test_back_df <- as.data.frame(testbglistsp)
-#head(test_back_df)
 print(Sys.time()-strt) #time it took to run
@@
 ggplot(testbk100001999, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/ssh_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_ssh_back.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100001999, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/mlp_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_mlp.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100001999, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/temp_surface_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_temp_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100001999, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/temp_depth_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_temp_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100001999, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/salinity_surface_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_salinity_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100001999, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/salinity_depth_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_salinity_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100001999, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/chl_surface_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_chl_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100001999, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/chl_depth_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_chl_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100001999, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500001999 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/o2_surface_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_o2_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100001999, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500001999 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/o2_depth_back_no_1999.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,".../output/env/background_point_check/199902_o2_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 }
-write.csv(testbglistsp, "../data/env/background_point_check/2014_06/testbglistoutput.csv", row.names = FALSE)
+write.csv(testbglistsp, "../data/env/background_point_check/201406_allbackgroundpoints.csv", row.names = FALSE)
 test_back_df <- as.data.frame(testbglistsp)
-#head(test_back_df)
 print(Sys.time()-strt) #time it took to run
@@
 ggplot(testbk100002014, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/ssh_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_ssh.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100002014, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/mlp_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_mlp.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100002014, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/temp_surface_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_temp_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100002014, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/temp_depth_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_temp_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100002014, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/salinity_surface_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_salinity_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100002014, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/salinity_depth_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_salinity_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100002014, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/chl_surface_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_chl_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100002014, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/chl_depth_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_chl_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100002014, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/o2_surface_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_o2_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ggplot(testbk100002014, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../data/o2_depth_back_no_2014.png") # to automatically save the plot to a png AND show it inline
+dev.copy(png,"../output/env/background_point_check/201406_o2_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png

### 2019-01-08 src/background_points_no_check.Rmd
updated for new background dataset (inc new output folder which is more organised)

@@
 
-netcdf_list <- list.files("../data/bktstncdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+netcdf_list <- list.files("../data/env/bktstncdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
 no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
-netcdf_name <- list.files("../data/bktstncdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+netcdf_name <- list.files("../data/env/bktstncdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
 aea <- raster("../output/env/aea.tif") 
@@
   temp_brick <- projectRaster(temp_brick, aea) 
-    for (j in 1:nrow(testbk190000sp)) {  
-      de <- testbk190000sp$depthlayerno[[j]]  # a variable for the observation depth layer
+    for (j in 1:nrow(testbglistsp)) {  
+      de <- testbglistsp$depthlayerno[[j]]  # a variable for the observation depth layer
           if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
-              testbk190000sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ]) 
+              testbglistsp$temp_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
               if (is.na(de)){
-                testbk190000sp$temp_depth[j] <- NA
+                testbglistsp$temp_depth[j] <- NA
               } else  
-                testbk190000sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = testbk190000sp[j, ])
+                testbglistsp$temp_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ])
           } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
-              testbk190000sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ]) 
+              testbglistsp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
               if (is.na(de)){
-                testbk190000sp$salinity_depth[j] <- NA
+                testbglistsp$salinity_depth[j] <- NA
               } else  
-                testbk190000sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = testbk190000sp[j, ]) 
+                testbglistsp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ]) 
           } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
-              testbk190000sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ]) 
+              testbglistsp$chl_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
               if (is.na(de)){
-                testbk190000sp$chl_depth[j] <- NA
+                testbglistsp$chl_depth[j] <- NA
               } else  
-                testbk190000sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = testbk190000sp[j, ]) 
+                testbglistsp$chl_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ]) 
           } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
-              testbk190000sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ]) 
+              testbglistsp$o2_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ]) 
               if (is.na(de)){
-                testbk190000sp$o2_depth[j] <- NA
+                testbglistsp$o2_depth[j] <- NA
               } else  
-                testbk190000sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = testbk190000sp[j, ]) 
+                testbglistsp$o2_depth[j] <- extract(x=temp_brick[[de]], y = testbglistsp[j, ]) 
           } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
-              testbk190000sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ])
+              testbglistsp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ])
           } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
-              testbk190000sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = testbk190000sp[j, ])
+              testbglistsp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = testbglistsp[j, ])
 

### 2019-01-08 src/timesliced_random_cell_selection.Rmd
??

@@
 ```{r}
-background_all2 <- background_all[background_all$year == 2007, ]
+background_all2 <- background_all[background_all$year == 2009, ]
 xy <- background_all2[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
@@
 }
-write.csv(background_all_sp, "../output/bio/background_complete_2007.csv", row.names = FALSE)
+write.csv(background_all_sp, "../output/bio/background_complete_2009.csv", row.names = FALSE)
 background_all2 <- as.data.frame(background_all_sp)

### 2019-01-08 src/BACKUPtimesliced_random_cell_selection.Rmd

--- 2019-01-08 src/timesliced_random_cell_selection.Rmd
@@
 ```{r}
+background_all <- read.csv("../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath_cellobs.csv", header = TRUE)
 background_all$temp_depth <- NA
@@
 ```{r}
-background_all2 <- background_all[background_all$year == 2009, ]
+background_all2 <- background_all[background_all$year == 1999 & background_all$month == 7, ]
 xy <- background_all2[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
@@
 }
-write.csv(background_all_sp, "../output/bio/background_complete_2009.csv", row.names = FALSE)
+write.csv(background_all_sp, "../output/bio/background_complete_1999.7.csv", row.names = FALSE)
 background_all2 <- as.data.frame(background_all_sp)
@@
 ```{r}
-back1998 <- read.csv("../output/bio/background_complete_1998.csv", header = TRUE)
-back1999 <- read.csv("../output/bio/background_complete_1999.csv", header = TRUE)
-back2000_2003 <- read.csv("../output/bio/background_complete_2000_2003.csv", header = TRUE)
-back2004_2006 <- read.csv("../output/bio/background_complete_2004_2006.csv", header = TRUE)
-back2007_2011 <- read.csv("../output/bio/background_complete_2007_2011.csv", header = TRUE)
-back2012_2016 <- read.csv("../output/bio/background_complete_2012_2016.csv", header = TRUE)
+back1998 <- read.csv("../output/bio/background/correct_random_files/background_complete_1998.csv", header = TRUE)
+back1999 <- read.csv("../output/bio/background/correct_random_files/background_complete_1999.csv", header = TRUE)
+back2000 <- read.csv("../output/bio/background/correct_random_files/background_complete_2000.csv", header = TRUE)
+back2001 <- read.csv("../output/bio/background/correct_random_files/background_complete_2001.csv", header = TRUE)
+back2002 <- read.csv("../output/bio/background/correct_random_files/background_complete_2002.csv", header = TRUE)
+back2003 <- read.csv("../output/bio/background/correct_random_files/background_complete_2003.csv", header = TRUE)
+back2004 <- read.csv("../output/bio/background/correct_random_files/background_complete_2004.csv", header = TRUE)
+back2005 <- read.csv("../output/bio/background/correct_random_files/background_complete_2005.csv", header = TRUE)
+back2006 <- read.csv("../output/bio/background/correct_random_files/background_complete_2006.csv", header = TRUE)
+back2007 <- read.csv("../output/bio/background_complete_2007.csv", header = TRUE)
+back2008 <- read.csv("../output/bio/background_complete_2008.csv", header = TRUE)
+back2009 <- read.csv("../output/bio/background_complete_2009.csv", header = TRUE)
+back2010 <- read.csv("../output/bio/background_complete_2010.csv", header = TRUE)
+back2011 <- read.csv("../output/bio/background_complete_2011.csv", header = TRUE)
+back2012 <- read.csv("../output/bio/background_complete_2012.csv", header = TRUE)
+back2013 <- read.csv("../output/bio/background_complete_2013v2.csv", header = TRUE)
+back2014 <- read.csv("../output/bio/background_complete_2014.csv", header = TRUE)
+back2015 <- read.csv("../output/bio/background_complete_2015.csv", header = TRUE)
 
-background_all <- rbind(back1998, back1999, back2000_2003, back2004_2006, back2007_2011, back2012_2016)
+background_all <- rbind(back1998, back1999, back2000, back2001, back2002, back2003, back2004, back2005, back2006, back2007, back2008, back2009, back2010, back2011, back2012, back2013, back2014, back2015)
 head(background_all)
+write.csv(background_all, "../output/bio/background_all_complete.csv", row.names = FALSE)
+```
+
+the prior files...
+
+```{r}
+#back1 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.1.csv", header = TRUE)
+#back2 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.2.csv", header = TRUE)
+back3 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.3.csv", header = TRUE)
+back4 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.4.csv", header = TRUE)
+back5 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.5.csv", header = TRUE)
+back6 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.6.csv", header = TRUE)
+back7 <- read.csv("../output/bio/background_complete_1999.7.csv", header = TRUE)
+#back8 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.8.csv", header = TRUE)
+back9 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.9.csv", header = TRUE)
+back10 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.10.csv", header = TRUE)
+back11 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.11.csv", header = TRUE)
+back12 <- read.csv("../output/bio/background/rand20k/no_na_10krand/1999.12.csv", header = TRUE)
+
+back1998 <- rbind(back3, back4, back5, back6, back7, back9, back10, back11, back12)
+#back1998 <- rbind(back1, back2, back3, back4, back5, back6, back7, back8, back9, back10, back11, back12)
+write.csv(back1998, "../output/bio/background/correct_random_files/background_complete_1999.csv", row.names = FALSE)
 ```
@@
 ```{r}
-backna <- sum(!complete.cases(background_all2)) #this just counts the number of rows that contain NA values
+backna <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
 backna
-background_all2 <- background_all2[complete.cases(background_all2), ] #removes NA values
-backna2 <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
-backna2
-write.csv(background_all, "../output/bio/background_complete_na.csv", row.names = FALSE)
 ```
 
-randomly select 10000 rows from each timeslice
-```{r}
-#background_all <- read.csv("../output/bio/background_complete_na.csv", header = TRUE)
-split_cell_back <- split(background_all, list(background_all$year, background_all$month), drop = TRUE)
-
-no_timeslices <- length(split_cell_back) # how many levels (time slices) are in the list
-bckoutput <- "../output/bio/background/rand20k/no_na/" #where the files are to be saved to
-for (i in 1:no_timeslices){
-  timeslice <- split_cell_back[[i]]
-  timeslice <- as.data.frame(timeslice)
-  write.csv(timeslice, paste0(bckoutput, names(lapply(split_cell_back, names))[i], ".csv"), row.names = FALSE) 
-}
-```
-
-```{r}
-bckrnd_folder <- "../output/bio/background/rand20k/no_na/"      # path to folder that holds multiple .csv files
-bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
-randbckoutput <- "../output/bio/background/rand20k/no_na_10krand/" #where the files are to be saved to
-# read in each .csv file in file_list and create a data frame with the same name as the .csv file
-for (i in 1:length(bckrnd_list)){
-  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
-  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample 
-  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
-}
-head(sel)
-```
-
-crap - the years from 2007 onwards have less than 10k points to choose from...
-create a random 30k for each of these month/year periods
+XXX THIS NEEDS EDITING AND MOVING TOO XXXX
 

### 2019-01-09 src/background_points_no_check.Rmd
Updated for new background data. All plots complete.

@@
 ```{r}
-testbk100001999 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
-testbk200001999 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
-testbk300001999 <- test_back_df[sample(nrow(test_back_df), 30000), ]  
-testbk500001999 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
-testbk1000001999 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
-testbk1900001999 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
 ```
@@
 ```{r}
-ggplot(test10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
-dev.copy(png,"../output/env/background_point_check/200710_ssh_back_no.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png, "../output/env/background_point_check/200710_ssh_back_no.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(test10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/200710_mlp_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/200710_temp_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/200710_temp_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/200710_salinity_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/200710_salinity_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/200710_chl_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/200710_chl_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/200710_o2_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(test10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=test20000 , na.rm = TRUE, colour = "blue") + geom_density(data=test50000 , na.rm = TRUE, colour = "green") + geom_density(data=test100000 , na.rm = TRUE, colour = "orange") + geom_density(data=test190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/200710_o2_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 ```{r}
-ggplot(testbk100001999, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_ssh_back.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_ssh_back.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(testbk100001999, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_mlp.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_mlp.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(testbk100001999, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_temp_surface.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_temp_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(testbk100001999, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_temp_depth.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_temp_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(testbk100001999, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_salinity_surface.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_salinity_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(testbk100001999, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_salinity_depth.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_salinity_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(testbk100001999, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_chl_surface.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_chl_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(testbk100001999, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_chl_depth.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_chl_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(testbk100001999, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500001999 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_o2_surface.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500001999 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_o2_surface.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
 
-ggplot(testbk100001999, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200001999 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500001999 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000001999 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900001999 , na.rm = TRUE, colour = "pink")
-dev.copy(png,".../output/env/background_point_check/199902_o2_depth.png") # to automatically save the plot to a png AND show it inline
+ggplot(testbk10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500001999 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+dev.copy(png,"../output/env/background_point_check/199902_o2_depth.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
@@
 ```
+
 ok now create the first lot of random for 2014_06
@@
 ```{r}
-testbk100002014 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
-testbk200002014 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
-testbk300002014 <- test_back_df[sample(nrow(test_back_df), 30000), ]  
-testbk500002014 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
-testbk1000002014 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
-testbk1900002014 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
 ```
@@
 ```{r}
-ggplot(testbk100002014, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_ssh.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk100002014, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_mlp.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk100002014, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_temp_surface.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk100002014, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_temp_depth.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk100002014, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_salinity_surface.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk100002014, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_salinity_depth.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk100002014, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_chl_surface.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk100002014, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_chl_depth.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk100002014, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_o2_surface.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk100002014, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk200002014 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk500002014 , na.rm = TRUE, colour = "green") + geom_density(data=testbk1000002014 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk1900002014 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
 dev.copy(png,"../output/env/background_point_check/201406_o2_depth.png") # to automatically save the plot to a png AND show it inline
@@
 ```{r}
-bck2014_06_3d <- plot_ly(x= testbk100002014$longitude_, y = testbk100002014$latitude_m, z = testbk100002014$depthlayerno)
+bck2014_06_3d <- plot_ly(x= testbk10000$longitude_, y = testbk10000$latitude_m, z = testbk10000$depthlayerno)
 bck2014_06_3d
@@
 ```{r}
-bck2014_06_2d <- plot(x= testbk100002014$longitude_, y = testbk100002014$latitude_m)
-dev.copy(png,"../data/env/background_point_check/bck10000_2d.png") # to automatically save the plot to a png AND show it inline
+bck10000points_2d <- plot(x= testbk10000$longitude_, y = testbk10000$latitude_m)
+dev.copy(png,"../data/env/background_point_check/bck10000points_2d.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png
-
 ```

### 2019-01-09 src/data_exploration.Rmd
background correlations rerun with corrected data

@@
 head(background)
-test <- background[245:250, ]
-test
+#test <- background[245:250, ]
+#test
 ```
@@
 
-check to see if there are any correlations in the env. variables
+check to see if there are any correlations in the env. variables for the background points
 
@@
 ```{r}
-kruskal.test(presence$temp_celsius_depth ~ presence$gear)
+kruskal.test(presence$temp_depth ~ presence$gear)
 ```
@@
 Biostatistical Analysis, 5th ed.
-Pearson Prentice Hall: Upper Saddle River, NJ.)
+Pearson Prentice Hall: Upper Saddle River, NJ.) http://rcompanion.org/rcompanion/d_06.html
 
@@
 ```{r}
-pairwise.wilcox.test(presence$temp_celsius_depth, presence$gear, p.adj='bonferroni', exact=F)
+pairwise.wilcox.test(presence$temp_depth, presence$gear, p.adj='bonferroni', exact=F)
 ```
+
+dunn test (not working)
+```{r}
+gear_temp_depthdunn <- dunnTest(presence$temp_depth, presence$gear)
+
+write.csv(gear_temp_depthdunn, "../output/env/gear_temp_depthdunn.csv", row.names = TRUE)
+```

### 2019-01-10 src/data_exploration.Rmd
figuring out if gear is a factor...

@@
 
+add nafo region and gear type into the mix
+
+first subset the env.correlate columns + bottom_depth (you don't need everything)
+
+```{r}
+background_envbotdepth <- subset(background, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
+background_envbotdepth_cor <- cor(background_envbotdepth, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(background_envbotdepth_cor, "../output/env/background_envbotdepth_cor.csv", row.names = TRUE)
+background_corrplot <- corrplot(background_envbotdepth_cor , method = "color", type = "upper", order = "alphabet", tl.cex = 0.8) 
+dev.copy(png,"../output/env/simple_plots/background_envbotdepth_cor.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
 ## correlations between presence points
@@
 
+map the gear usage in Arcgis (gear_type_map)
+
+create a table of gear use by month
+
+```{r}
+gearby_mth <- with(presence, table(month, gear))
+write.csv(gearby_mth, file = "../output/bio/gear_by_mth.csv")
+gearby_mth
+```
+
 ```{r}
@@
 
-dunn test (not working)
-```{r}
-gear_temp_depthdunn <- dunnTest(presence$temp_depth, presence$gear)
-
-write.csv(gear_temp_depthdunn, "../output/env/gear_temp_depthdunn.csv", row.names = TRUE)
-```
+dunn test 
+```{r}
+gear_temp_depthdunn <- dunnTest(presence$temp_depth, presence$gear, list = TRUE)
+gear_temp_depthdunn
+
+write.csv(table, "../output/env/gear_temp_depthdunn.csv", row.names = TRUE)
+```

### 2019-01-14 src/biological_data_prep.Rmd
minor edit to how gear == NA reported (changed to unknown)
investingating presence data for missing values

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful", "marmap", "weathermetrics")
+packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful", "marmap", "weathermetrics", "fastmatch")
 source("../src/ipak.R")
@@
 
+#gear type adjustment
+
+should not have gear not reported listed as NA (for some analysis). Change gear == NA to unknown
+
+```{r}
+data_aea$gear <- as.character(data_aea$gear)
+data_aea$gear[is.na(data_aea$gea)] <- "unknown"
+write.csv(data_aea, "../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env_naocor_cels.csv", row.names = FALSE)
+```
+
 # warm/cold year
@@
 The warm/cold year index will be taken from a [DFO CSAS report](http://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2018/2018_039-eng.pdf) - Fig 8
+
+# checking data for NA's
+
+The presence points have a lot of NAs in the env. correlates. I want to:
+
+-ID and remove rows where all values == NA
+-ID observations with no depth data that are missing surface values (will need to be removed)
+-ID observations with depth data but are missing some surface and/or depth values
+
+how many rows(observations do you have)
+```{r}
+nrow(data_aea) # there are 11516 rows
+```
+
+check for no rows (observations) with NAs
+
+```{r}
+presence_noNAcount <- sum(!complete.cases(data_aea)) #this just counts the number of rows that contain NA values
+presence_noNAcount
+```
+I have 4687 observations that contain NAs
+
+1) find ones that have no env. data at all. these will need to be removed straight off.
+
+There are 11516 rows in the original presence-point dataset
+
+find the column numbers that correspond to the env. data (use fastmatch package as is quicker)
+
+```{r}
+fmatch("chl_surface", names(data_aea)) #28
+fmatch("chl_depth", names(data_aea)) #29
+fmatch("mlp_surface", names(data_aea)) #30
+fmatch("o2_surface", names(data_aea)) #31
+fmatch("o2_depth", names(data_aea)) #32
+fmatch("salinity_surface", names(data_aea)) #33
+fmatch("salinity_depth", names(data_aea)) #34
+fmatch("ssh_surface", names(data_aea)) #35
+fmatch("temp_surface", names(data_aea)) #36
+fmatch("temp_depth", names(data_aea)) #37
+```
+ok so the env. data i am intersted in is in cols 28:37
+
+Out of curiosoity, map the location of these points. Note the difference between this and the code below is the addition of ! on rowSums
+
+```{r}
+presence_novals <- data_aea[!rowSums(is.na(data_aea[ , 28:37])) !=10, ] #4:13 represent the columns w/ env. data, 10 is the number of columns to check (4 to 13 inclusive)
+write.csv(presence_novals, "../output/bio/presence_points_without_envdata.csv", row.names = FALSE)
+nrow(presence_novals) #just to check it has mapped
+```
+
+So I had 123 rows without any env. data (11516 - 11393).
+
+plot
+```{r map presences without any env. data}
+map <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map, xlim = c(-70, -43), ylim =c(38,70), asp = 1) #the x and y lim are the long-lat bounds of the map
+points(presence_novals$decimalLongitude, presence_novals$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+```
+
+Ok so they are along the coast, so the points probably fell outside the boxes.
+
+Now remove the values
+```{r}
+data_aea <- data_aea[rowSums(is.na(data_aea[ , 28:37])) !=10, ] #4:13 represent the columns w/ env. data, 10 is the number of columns to check (4 to 13 inclusive)
+write.csv(data_aea, "../output/bio/presence_nodata_rm")
+nrow(data_aea)
+```
+
+ok so the next step is to make sure that any of the data you have without a depth value does have all relevant surface values
+
+maybe subset
+
+```{r}
+presence_nodepth = data_aea[is.na(data_aea$depth), ]
+nrow(presence_nodepth) #1215 rows
+head(presence_nodepth)
+```
+
+go through each value and see which are missing data
+```{r}
+presence_nodepth_nochl <- presence_nodepth[is.na(presence_nodepth$chl_surface), ]
+nrow(presence_nodepth_nochl) #just to check it has mapped
+presence_nodepth_noo2 <- presence_nodepth[is.na(presence_nodepth$o2_surface), ]
+nrow(presence_nodepth_noo2) #just to check it has mapped
+presence_nodepth_nosalinity <- presence_nodepth[is.na(presence_nodepth$salinity_surface), ]
+nrow(presence_nodepth_nosalinity) #just to check it has mapped
+presence_nodepth_notemp <- presence_nodepth[is.na(presence_nodepth$temp_surface), ]
+nrow(presence_nodepth_notemp) #just to check it has mapped
+presence_nodepth_nossh <- presence_nodepth[is.na(presence_nodepth$ssh_surface), ]
+nrow(presence_nodepth_nossh) #just to check it has mapped
+presence_nodepth_nomlp <- presence_nodepth[is.na(presence_nodepth$mlp_surface), ]
+nrow(presence_nodepth_nomlp) #just to check it has mapped
+```
+
+great! all of the values that have no depth data have surface data.
+
+now on to the data which has depth data
+
+```{r}
+presence_wdepth = data_aea[!is.na(data_aea$depth), ]
+nrow(presence_wdepth) #10178 rows
+head(presence_wdepth)
+```
+
+go through each value and see which are missing data. First the surface values
+```{r}
+presence_wdepth_nochl <- presence_wdepth[is.na(presence_wdepth$chl_surface), ]
+nrow(presence_wdepth_nochl) #just to check it has mapped
+presence_wdepth_noo2 <- presence_wdepth[is.na(presence_wdepth$o2_surface), ]
+nrow(presence_wdepth_noo2) #just to check it has mapped
+presence_wdepth_nosalinity <- presence_wdepth[is.na(presence_wdepth$salinity_surface), ]
+nrow(presence_wdepth_nosalinity) #just to check it has mapped
+presence_wdepth_notemp <- presence_wdepth[is.na(presence_wdepth$temp_surface), ]
+nrow(presence_wdepth_notemp) #just to check it has mapped
+presence_wdepth_nossh <- presence_wdepth[is.na(presence_wdepth$ssh_surface), ]
+nrow(presence_wdepth_nossh) #just to check it has mapped
+presence_wdepth_nomlp <- presence_wdepth[is.na(presence_wdepth$mlp_surface), ]
+nrow(presence_wdepth_nomlp) #just to check it has mapped
+```
+
+Great! All have surface values. now to the depth values
+
+```{r}
+presence_wdepth_nochl <- presence_wdepth[is.na(presence_wdepth$chl_depth), ]
+nrow(presence_wdepth_nochl) #just to check it has mapped
+presence_wdepth_noo2 <- presence_wdepth[is.na(presence_wdepth$o2_depth), ]
+nrow(presence_wdepth_noo2) #just to check it has mapped
+presence_wdepth_nosalinity <- presence_wdepth[is.na(presence_wdepth$salinity_depth), ]
+nrow(presence_wdepth_nosalinity) #just to check it has mapped
+presence_wdepth_notemp <- presence_wdepth[is.na(presence_wdepth$temp_depth), ]
+nrow(presence_wdepth_notemp) #just to check it has mapped
+```
+
+damn... so
+
+chl_depth = 2394 missing values
+o2_depth = 2394 missing values
+salinity_depth = 3133 missing values
+temp_depth = 3109 missing values

### 2019-01-14 src/biological_data_prep.Rmd
depth missing values

@@
 temp_depth = 3109 missing values
+
+This is interesting...
+chl and o2 are missing the same number of points - they come from the biomer dataset; salinity and temp come from glorys and have different values.
+
+I wonder how many of these points represent the same lat-long-depth... need to subset where chl_depth OR o2_depth OR salinity_depth OR temp_depth == OR These columns are not together
+
+```{r}
+presence_wdepth2 <-  presence_wdepth[!(is.na(presence_wdepth$chl_depth) & is.na(presence_wdepth$o2_depth) & is.na(presence_wdepth$salinity_depth) & is.na(presence_wdepth$temp_depth)) , ]
+nrow(presence_wdepth2)
+```

### 2019-01-14 src/biological_data_prep.Rmd
moving through presence point check - rerun loop just in case

--- 2019-01-14 src/biological_data_prep.Rmd
@@
 temp_depth = 3109 missing values
+
+This is interesting...
+chl and o2 are missing the same number of points - they come from the biomer dataset; salinity and temp come from glorys and have different values.
+
+I wonder how many of these points represent the same lat-long-depth... need to subset where chl_depth OR o2_depth OR salinity_depth OR temp_depth == NA These columns are not together
+
+```{r}
+presence_wdepth2 <-  presence_wdepth[!complete.cases(presence_wdepth[28:37]), ]
+nrow(presence_wdepth2) #3349
+```
+ok so there are 3349 points without any data.let's map these
+
+```{r map presences with depth without any env. data at depth}
+map <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map, xlim = c(-70, -43), ylim =c(38,70), asp = 1) #the x and y lim are the long-lat bounds of the map
+points(presence_wdepth2$decimalLongitude, presence_wdepth2$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+```
+hmm everywhere...
+
+Is there a particular time-slice that the values appear in?
+
+```{r}
+npresence_wdepth2mmyy <- with(presence_wdepth2, table(year, month))
+npresence_wdepth2mmyy
+```
+
+hmm there is a lot... how does this compare to the original dataset
+
+```{r}
+data_aeammyy <- with(data_aea, table(year, month))
+data_aeammyy
+```
+
+ok so 2014_07 and 1999_07 are both missing all data. Re-run the loop for these two time-periods (just add the nectcdfs for these files in renetcdf folder).
+
+First you need to put the observations (data_aea) into a spatialpointsdataframe.
+
+```{r}
+fmatch("longitude_meters", names(data_aea)) #19
+fmatch("latitude_meters", names(data_aea)) #20
+```
+
+```{r}
+library(sp)
+xy <- data_aea[ ,c(19,20)] # This is to tell R where the coordinates are (in column 19 and 20). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+```{r}
+netcdf_list <- list.files("../data/env/renetcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/renetcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not includedCores
+library (ncdf4)
+library(sp)
+library(rgdal)
+for (i in 1:no_netcdf) {  
+  library(raster)
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(data_aea_sp)) {  
+      de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
+      yr <- (data_aea_sp$year[j])  # a variable for the observation year
+      mth <- (data_aea_sp$month[j])  # a variable for the observation month
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+              if (is.na(de)){
+                data_aea_sp$temp_depth[j] <- NA
+              } else  
+                data_aea_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              if (is.na(de)){
+                data_aea_sp$salinity_depth[j] <- NA
+              } else  
+                data_aea_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
+              if (is.na(de)){
+                data_aea_sp$chl_depth[j] <- NA
+              } else  
+                data_aea_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
+              if (is.na(de)){
+                data_aea_sp$o2_depth[j] <- NA
+              }
+          }
+
+    }
+}
+testdata_aea <- as.data.frame(data_aea_sp)
+
+```
+
+lets see if that made any difference...
+
+now on to the data which has depth data
+
+```{r}
+presence_wdepthv2 = data_aea[!is.na(data_aea$depth), ]
+nrow(presence_wdepthv2) #10178 rows - same as before
+head(presence_wdepthv2)
+```
+
+```{r}
+npresence_wdepth2mmyy2 <- with(presence_wdepthv2, table(year, month))
+npresence_wdepth2mmyy2
+```
+
+no change...
+
+ok is there a depth layer in particular that is an issue?
+maybe some are missing a depthlayerno?
+
+```{r}
+presence_wdepth_nolayerno <- presence_wdepth2[is.na(presence_wdepth2$depthlayerno), ]
+nrow(presence_wdepth_nolayerno) #0 
+head(presence_wdepth_nolayerno)
+```
+ok so all have a depthlayerno attached
+
+check for count of each depth layer
+
+```{r}
+whichdepths <- unique(presence_wdepth2$depthlayerno) #makes a variable of institutions
+whichdepthsordered <- sort(whichdepths) #just puts the list in 
+whichdepthsordered
+```
+quite a wide group... look at frequency
+
+```{r count depthlayerno for data with depth but missing values}
+whichdepthscount <- count(presence_wdepth2, "depthlayerno")
+whichdepthscount
+```
+very variable...
+
+How many rows are missing all depth data?
+
+```{r}
+presence_wdepthonly <- subset(presence_wdepth2, select = -c(mlp_surface, ssh_surface, temp_surface, salinity_surface, o2_surface, chl_surface))
+head(presence_wdepthonly)
+```
+
+get the col numbers for the depth data
+```{r}
+fmatch("chl_depth", names(presence_wdepthonly)) #28
+fmatch("o2_depth", names(presence_wdepthonly)) #29
+fmatch("salinity_depth", names(presence_wdepthonly)) #30
+fmatch("temp_depth", names(presence_wdepthonly)) #31
+```
+
+now subset rows where all depth data is missing
+```{r}
+presence_wdepthonly_allmissing <-  presence_wdepthonly[(is.na(presence_wdepthonly$temp_depth)) & (is.na(presence_wdepthonly$o2_depth)) & (is.na(presence_wdepthonly$salinity_depth)) & (is.na(presence_wdepthonly$chl_depth)), ]
+nrow(presence_wdepthonly_allmissing) #2178
+head(presence_wdepthonly_allmissing)
+```
+
+so there are 2178 out of 3349 observations that are missing all data for depth. so what about the remaining 1171 points?
+
+want to remove from the observations in presence_wdepthonly (all 3349 observations) the ones that appear in presence_wdepthonly_allmissing (2178 observations)
+
+some code from [stack exchange](https://stackoverflow.com/questions/10907359/r-remove-rows-from-one-data-frame-that-are-in-another)
+
+```{r}
+presence_wdepthonly_somemissing <-  rbind(presence_wdepthonly_allmissing, presence_wdepthonly)
+presence_wdepthonly_somemissing <- presence_wdepthonly_somemissing[!duplicated(presence_wdepthonly_somemissing, fromLast = FALSE) & !duplicated(presence_wdepthonly_somemissing, fromLast = TRUE), ]
+nrow(presence_wdepthonly_somemissing)
+head(presence_wdepthonly_somemissing)
+```
+
+ok before going further, just rerun the loop again on all the data...
+
+First you need to put the observations (data_aea) into a spatialpointsdataframe.
+```{r}
+library(sp)
+xy <- data_aea[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+```{r}
+netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not includedCores
+library (ncdf4)
+library(sp)
+library(rgdal)
+for (i in 1:no_netcdf) {  
+  library(raster)
+  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
+  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
+  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) 
+    for (j in 1:nrow(data_aea_sp)) {  
+      de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
+      yr <- (data_aea_sp$year[j])  # a variable for the observation year
+      mth <- (data_aea_sp$month[j])  # a variable for the observation month
+          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
+              data_aea_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$temp_depth[j] <- NA
+              } else  
+                data_aea_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
+              data_aea_sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$salinity_depth[j] <- NA
+              } else  
+                data_aea_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
+              data_aea_sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$chl_depth[j] <- NA
+              } else  
+                data_aea_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
+              data_aea_sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
+              if (is.na(de)){
+                data_aea_sp$o2_depth[j] <- NA
+              } else  
+                data_aea_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
+          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
+              data_aea_sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
+          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
+              data_aea_sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
+
+          }
+
+    }
+}
+data_aea <- as.data.frame(data_aea_sp)
+write.csv(data_aea,"../output/bio/presence_points_without_envdata_relooped.csv", row.names = FALSE)
+head(data_aea_sp$chl_depth)
+```
+note that if you wanted to just loop through the netcdfs and create a ton of raster bricks with the same name as the netcdf, the following code you wrote before works
+
+netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
+no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
+netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
+library (ncdf4)
+for (i in 1:no_netcdf) {
+  print(netcdf_name[i]) #this just prings the name of the netCDF R is working on
+  temp_brick <- brick(netcdf_list[i], lvar = 4)
+  temp_brick <- projectRaster(temp_brick, aea) #aea is an existing raster in the projection I want that I created (environmental_data_preperation)
+  assign(netcdf_name[i], temp_brick) #this asigns the netCDF name to the raster brick
+}
+
+```{r}
+
+```
+
+change dataset back into a dataframe and then save as a .csv
+
+```{r}
+
+```

### 2019-01-14 src/biological_data_prep.Rmd
re-ran loop. confirmed some dont have data

@@
 ```
-note that if you wanted to just loop through the netcdfs and create a ton of raster bricks with the same name as the netcdf, the following code you wrote before works
-
-netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
-no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
-netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
-library (ncdf4)
-for (i in 1:no_netcdf) {
-  print(netcdf_name[i]) #this just prings the name of the netCDF R is working on
-  temp_brick <- brick(netcdf_list[i], lvar = 4)
-  temp_brick <- projectRaster(temp_brick, aea) #aea is an existing raster in the projection I want that I created (environmental_data_preperation)
-  assign(netcdf_name[i], temp_brick) #this asigns the netCDF name to the raster brick
-}
-
-```{r}
-
-```
-
-change dataset back into a dataframe and then save as a .csv
-
-```{r}
-
-```
+
+see if it has made a difference
+
+```{r}
+presence_wdepthv3 = data_aea[!is.na(data_aea$depth), ]
+nrow(presence_wdepthv3) #10178 rows - same as before whcih is good
+head(presence_wdepthv3)
+```
+
+```{r}
+presence_wdepth_nochl3 <- presence_wdepthv3[is.na(presence_wdepthv3$chl_depth), ]
+nrow(presence_wdepth_nochl3) #just to check it has mapped
+presence_wdepth_noo23 <- presence_wdepthv3[is.na(presence_wdepthv3$o2_depth), ]
+nrow(presence_wdepth_noo23) #just to check it has mapped
+presence_wdepth_nosalinity3 <- presence_wdepthv3[is.na(presence_wdepthv3$salinity_depth), ]
+nrow(presence_wdepth_nosalinity3) #just to check it has mapped
+presence_wdepth_notemp3 <- presence_wdepthv3[is.na(presence_wdepthv3$temp_depth), ]
+nrow(presence_wdepth_notemp3) #just to check it has mapped
+```
+
+It used to be
+[1] 2394
+[1] 2394
+[1] 3133
+[1] 3109
+so actually a minor change (maybe this is because you added the dec 2015 data in - it wasn't available before). ok need to redo checks...

### 2019-01-15 src/background_points_no_check.Rmd
minor updates - just reran some plots with newer data

@@
 ```{r}
-ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Sea Surface Height Above Geoid (meters)")
 dev.copy(png, "../output/env/background_point_check/200710_ssh_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Mixed layer thickness (MLP) (meters)")
 dev.copy(png,"../output/env/background_point_check/200710_mlp_back_no.png") # to automatically save the plot to a png AND show it inline 
@@
 
-ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")  + labs(x = "Temperature at surface (kelvin)")
 dev.copy(png,"../output/env/background_point_check/200710_temp_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Temperature at sampling depth (kelvin)")
 dev.copy(png,"../output/env/background_point_check/200710_temp_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Salinity at surface (kelvin)")
 dev.copy(png,"../output/env/background_point_check/200710_salinity_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Salinity at sampling depth (kelvin)")
 dev.copy(png,"../output/env/background_point_check/200710_salinity_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Chlorophyll concentration at surface (mmol.m-3)")
 dev.copy(png,"../output/env/background_point_check/200710_chl_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Chlorophyll concentration at sampling depth (mmol.m-3)")
 dev.copy(png,"../output/env/background_point_check/200710_chl_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Dissolved oxygen concentration at surface (mmol.m-3)")
 dev.copy(png,"../output/env/background_point_check/200710_o2_surface_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Dissolved oxygen concentration at sampling depth (mmol.m-3)")
 dev.copy(png,"../output/env/background_point_check/200710_o2_depth_back_no.png") # to automatically save the plot to a png AND show it inline
@@
 ```{r}
-ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Sea Surface Height Above Geoid (meters)")
 dev.copy(png,"../output/env/background_point_check/199902_ssh_back.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")  + labs(x = "Mixed layer thickness (MLP) (meters)")
 dev.copy(png,"../output/env/background_point_check/199902_mlp.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Temperature at surface (kelvin)")
 dev.copy(png,"../output/env/background_point_check/199902_temp_surface.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Temperature at sampling depth (kelvin)")
 dev.copy(png,"../output/env/background_point_check/199902_temp_depth.png") # to automatically save the plot to a png AND show it inline
@@
 ```{r}
-ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Sea Surface Height Above Geoid (meters)")
 dev.copy(png,"../output/env/background_point_check/201406_ssh.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")  + labs(x = "Mixed layer thickness (MLP) (meters)")
 dev.copy(png,"../output/env/background_point_check/201406_mlp.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")+ labs(x = "Temperature at surface (kelvin)")
 dev.copy(png,"../output/env/background_point_check/201406_temp_surface.png") # to automatically save the plot to a png AND show it inline
@@
 
-ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
+ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")  + labs(x = "Temperature at sampling depth (kelvin)")
 dev.copy(png,"../output/env/background_point_check/201406_temp_depth.png") # to automatically save the plot to a png AND show it inline
@@
 ```{r}
-bck10000points_2d <- plot(x= testbk10000$longitude_, y = testbk10000$latitude_m)
-dev.copy(png,"../data/env/background_point_check/bck10000points_2d.png") # to automatically save the plot to a png AND show it inline
+bck10000points_2d <- plot(x= testbk10000$longitude_, y = testbk10000$latitude_m, xlab = "Longitude (meters)", ylab = "Latitude (meters")
+dev.copy(png,"../output/env/background_point_check/bck10000points_2d.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png

### 2019-01-15 src/data_exploration.Rmd
minor updates - just reran some plots with newer data

@@
 # Overview
-Exploration of all the data collected on the presence points
+Exploration of all the data collected on the presence points + randomly generated background points
 
@@
 ```{r}
-presence <- read.csv("../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env_naocor_cels.csv", header = TRUE)
+presence <- read.csv("../output/bio/presence_points_without_envdata_relooped.csv", header = TRUE)
 head(presence)
@@
 ```
-ouch - 40 layers... a job for a boxplot probably
+ouch - 39 layers... a job for a boxplot probably
 
@@
 head(background)
-#test <- background[245:250, ]
-#test
 ```
@@
 ```{r}
-pres_env <- subset(presence, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+pres_env <- subset(presence, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 pres_env_cor <- cor(pres_env, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 nafo_by_yr <- with(presence, table(year, nafo_zone))
-write.csv(inst_by_yr, file = "../output/bio/nafozone_by_yr.csv")
+write.csv(nafo_by_yr, file = "../output/bio/nafozone_by_yr.csv")
 nafo_by_yr
@@
 ```{r}
-nafo_by_mth <- with(presence, table(month, nafo_zone))
-write.csv(inst_by_yr, file = "../output/bio/nafozone_by_mth.csv")
+nafo_by_mth <- with(presence, table(nafo_zone, month))
+write.csv(nafo_by_mth, file = "../output/bio/nafozone_by_mth.csv")
 nafo_by_mth

### 2019-01-16 src/data_exploration.Rmd
addition of some individual box plots

@@
 
-temperature
-```{r}
-par(mfrow=c(1,2))
-boxplot(presence$temp_depth, main = "Temp at Depth (Various)", ylab = "Kelvin")
-boxplot(presence$temp_surface, main = "Temp at Surface", ylab = "Kelvin")
-dev.copy(png, "../output/env/simple_plots/temperature_boxplot.png") # to automatically save the plot to a png AND show it inline
+individual plots of each variable 
+```{r}
+par(mfrow=c(1,2))
+boxplot(presence$temp_depth, main = "Temperature at Depth (Various)", ylab = "Kelvin")
+boxplot(presence$temp_surface, main = "Temperature at Surface", ylab = "Kelvin")
+dev.copy(png, "../output/env/simple_plots/temp_boxplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+par(mfrow=c(1,2))
+boxplot(presence$salinity_depth, main = "Salinity at Depth (Various)", ylab = "PSU")
+boxplot(presence$salinity_surface, main = "Salinity at Surface", ylab = "PSU")
+dev.copy(png, "../output/env/simple_plots/salinity_boxplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+par(mfrow=c(1,2))
+boxplot(presence$o2_depth, main = "O2 at Depth (Various)", ylab = "mmol.m-3")
+boxplot(presence$o2_surface, main = "O2 at Surface", ylab = "mmol.m-3")
+dev.copy(png, "../output/env/simple_plots/o2_boxplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+par(mfrow=c(1,2))
+boxplot(presence$chl_depth, main = "Chlorphyll at Depth (Various)", ylab = "mmol.m-3")
+boxplot(presence$chl_surface, main = "Chlorphyll at Surface", ylab = "mmol.m-3")
+dev.copy(png, "../output/env/simple_plots/chl_boxplot.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+par(mfrow=c(1,2))
+boxplot(presence$mlp_surface, main = "Density Mixed Layer Thickness", ylab = "meters")
+boxplot(presence$ssh_surface, main = "Sea Surface Height", ylab = "meters")
+dev.copy(png, "../output/env/simple_plots/mlp_ssh_boxplot.png") # to automatically save the plot to a png AND show it inline
 dev.off() # stops automatic saving of the plot to a png

### 2019-01-16 src/biological_data_prep.Rmd
glorys bottom info added to presence data (but bottom layer no is not good)

@@
 so actually a minor change (maybe this is because you added the dec 2015 data in - it wasn't available before). ok need to redo checks...
+
+# glorys bathymetric data
+
+so i just found that glorys also has a bathymetric data layer...
+
+First you need to put the observations (data_aea) into a spatialpointsdataframe.
+```{r}
+xy <- data_aea[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+then covert the netcdf with the bathy layer (GLO-MFC_001_025_bathy.nclonlat.nc) to a brick and project to the albers equal area
+```{r}
+bathyrast <- raster("../data/env/GLO-MFC_001_025_bathy.nclonlat.nc") #because it is a single level file
+aea <- raster("../output/env/aea.tif") 
+bathyrast <- projectRaster(bathyrast, aea)
+plot(bathyrast)
+```
+
+now extract and add to data_aea
+```{r}
+data_aea_sp$bottom_depth_glorys <- extract(x = bathyrast, y = data_aea_sp)
+head(data_aea_sp)
+```
+
+ok great - now let's do the same thing for the bottom_layer_no (yes Glorys has a layer for this as well!)
+
+ covert the netcdf with the bottom layer nor (GLO-GLO-MFC_001_025_bottomlayerno.nclonlat.nc) to a raster and project to the albers equal area
+```{r}
+botlayernorast <- raster("../data/env/GLO-MFC_001_025_bottomlayerno.nclonlat.nc") #because it is a single level file
+botlayernorast <- projectRaster(botlayernorast, aea)
+plot(botlayernorast)
+```
+
+now extract and add to data_aea_sp, but this time also change the sp to a normal dataframe and save as a .csv
+```{r}
+data_aea_sp$bottom_layerno_glorys <- extract(x = botlayernorast, y = data_aea_sp)
+write.csv(data_aea_sp, file = "presence_points_without_envdata_relooped_glbathy.csv", row.names = FALSE) # change the path to where ever you would like to save it
+data_aea <- as.data.frame(data_aea_sp) # this coverts the 'spatialpoints dataframe' to a 'normal dataframe'
+head(data_aea)
+```
+
+ok bottom layer no is not right

### 2019-01-17 src/biological_data_prep.Rmd
datasets merged, vif run, finished with missing data stuff

--- 2019-01-14 src/biological_data_prep.Rmd
@@
 ```{r}
-names(data_aea)[names(data_aea)=="freq"] <- "total_cell_obs"
+names(data_aea)[names(data_aea)=="freq"] <- "total_cell_obs_xy"
 head(data_aea)
@@
 data_aea <- merge(data_aea, obs_cell_yymm, by.x=c("cell_id", "year", "month"), by.y=c("cell_id", "year", "month"))
-names(data_aea)[names(data_aea)=="freq"] <- "yymm_cell_obs"
+names(data_aea)[names(data_aea)=="freq"] <- "total_cell_obs_xyt"
 head(data_aea)
@@
 o2_depth = 2394 missing values
-salinity_depth = 3133 missing values
+salinity_depth = 3109 missing values
 temp_depth = 3109 missing values
@@
 This is interesting...
-chl and o2 are missing the same number of points - they come from the biomer dataset; salinity and temp come from glorys and have different values.
+chl and o2 are missing the same number of points - they come from the biomer dataset; salinity and temp come from glorys and have the same number of points.
 
@@
 presence_wdepth2 <-  presence_wdepth[!complete.cases(presence_wdepth[28:37]), ]
-nrow(presence_wdepth2) #3349
-```
-ok so there are 3349 points without any data.let's map these
+nrow(presence_wdepth2) #3325
+```
+ok so there are 3325 points without any data.let's map these
 
@@
 
-ok so 2014_07 and 1999_07 are both missing all data. Re-run the loop for these two time-periods (just add the nectcdfs for these files in renetcdf folder).
+ok so 2014_07 and 1999_07 are both missing all data. Re-run the loop for this time-periods(just add the nectcdfs for these files in renetcdf folder).
 
@@
 
-lets see if that made any difference...
+lets see if that made any difference... nope ok so it wasnt that i was missing the files
 
@@
 
-so there are 2178 out of 3349 observations that are missing all data for depth. so what about the remaining 1171 points?
-
-want to remove from the observations in presence_wdepthonly (all 3349 observations) the ones that appear in presence_wdepthonly_allmissing (2178 observations)
-
-some code from [stack exchange](https://stackoverflow.com/questions/10907359/r-remove-rows-from-one-data-frame-that-are-in-another)
-
-```{r}
-presence_wdepthonly_somemissing <-  rbind(presence_wdepthonly_allmissing, presence_wdepthonly)
-presence_wdepthonly_somemissing <- presence_wdepthonly_somemissing[!duplicated(presence_wdepthonly_somemissing, fromLast = FALSE) & !duplicated(presence_wdepthonly_somemissing, fromLast = TRUE), ]
-nrow(presence_wdepthonly_somemissing)
-head(presence_wdepthonly_somemissing)
-```
-
-ok before going further, just rerun the loop again on all the data...
+ok before going further, just rerun the loop again on all the data just to make sure there wasn't some silly fail...
 
@@
 [1] 2394
-[1] 3133
 [1] 3109
-so actually a minor change (maybe this is because you added the dec 2015 data in - it wasn't available before). ok need to redo checks...
+[1] 3109
+so no change.
+
+ok well i think the answer here is that if i am going to use depth data, i should get rid of those with missing data. if i am going to use surface only data, then i can keep all points.
+
+# glorys bathymetric data
+
+so i just found that glorys also has a bathymetric data layer...
+
+First you need to put the observations (data_aea) into a spatialpointsdataframe.
+```{r}
+xy <- data_aea[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+```
+
+then covert the netcdf with the bathy layer (GLO-MFC_001_025_bathy.nclonlat.nc) to a brick and project to the albers equal area
+```{r}
+bathyrast <- raster("../data/env/GLO-MFC_001_025_bathy.nclonlat.nc") #because it is a single level file
+aea <- raster("../output/env/aea.tif") 
+bathyrast <- projectRaster(bathyrast, aea)
+plot(bathyrast)
+```
+
+now extract and add to data_aea
+```{r}
+data_aea_sp$bottom_depth_glorys <- extract(x = bathyrast, y = data_aea_sp)
+data_aea <- as.data.frame(data_aea_sp) # this coverts the 'spatialpoints dataframe' to a 'normal dataframe'
+write.csv(data_aea, file = "../output/bio/presence_points_without_envdata_relooped_glbathy.csv", row.names = FALSE) # change the path to where ever you would like to save it
+head(data_aea)
+```

### 2019-01-17 src/data_exploration.Rmd
datasets merged, vif run, finished with missing data stuff

@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA") 
+packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car") 
 source("../src/ipak.R")
@@
 ```{r}
-presence <- read.csv("../output/bio/presence_points_without_envdata_relooped.csv", header = TRUE)
+presence <- read.csv("../output/bio/presence_points_without_envdata_relooped_glbathy.csv", header = TRUE)
 head(presence)
@@
 ```{r}
-background <- read.csv("../output/bio/background_complete_obs_cels.csv", header = TRUE)
+background <- read.csv("../output/bio/background_complete_obs_cels_globot.csv", header = TRUE)
 head(background)
@@
 ```
+
+#vif
+
+for this you need the joined dataset. this was created under mergig_datasets.src and the file is ../output/bio/presab.csv
+
+```{r}
+presab <- read.csv("../output/bio/presab.csv", header = TRUE)
+```
+
+```{r}
+vif_allpresab <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allpresab, "../output/bio/vif_allpresab.csv", row.names = TRUE)
+vif_allpresab
+```
+interpret - see https://stats.stackexchange.com/questions/70679/which-variance-inflation-factor-should-i-be-using-textgvif-or-textgvif/96584#96584
+To make GVIFs comparable across dimensions, we suggested using GVIF^(1/(2*Df)), where Df is the number of coefficients in the subset (ref fox and motette 1992 in zotero)
+or the 2 continuous variables, GVIFˆ(1/(2*Df)) (which is basically the square root of the VIF/GVIF value as DF=1) is the proportional change of the standard error and confidence interval of their coefficients due to the level of collinearity.
+The GVIFˆ(1/(2*Df)) value of the categorical variable is a similar measure for the reduction in precision of the coefficients' estimation due to collinearity.
+apparently i just need to square GVIF^(1/(2*Df)) and then use the normal VIF "rule of thumb"...
+
+```{r}
+vif_allpresab_sq <- read.csv("../output/bio/vif_allpresab.csv", header = TRUE)
+vif_allpresab_sq$GVIF2Dfsq <- vif_allpresab_sq$GVIF..1..2.Df..^2
+write.csv(vif_allpresab_sq, "../output/bio/vif_allpresab_sq.csv", row.names = TRUE)
+vif_allpresab_sq
+```

### 2019-01-21 src/data_exploration.Rmd
plots of who sampled on final dataset

@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car") 
+packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car", "rworldmap") 
 source("../src/ipak.R")
@@
 
+# remap who sampled
+
+Now lets check the number of records and spatial-temporal distribution of the observations by institution code to make sure none are dodgy
+
+first a table of how many observations each instituioncode has
+```{r institution code analysis - count}
+obs_by_ins <- count(presence, "institutioncode")
+obs_by_ins
+write.csv(obs_by_ins, file = "../output/bio/samplinginstitutions/no_observations_institutioncode.csv")
+```
+ok so IMR, NAFO, ICES, & USM have very few entries (1, 1, 1, and 3 respectively)
+
+Lets take a look at the spatial breakdown of these institutions.First all points...
+
+```{r}
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "All Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(presence$decimalLongitude, presence$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/all_occurrences.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+```
+Note there is one point up by iceland that you should get rid of (Icelandic population thought to be seperate from Labrador, but it is unclear if this is true or not).
+
+Map the institutioncode == ARC only data...
+```{r map obs by institutuion}
+ARC_obs <- presence[presence$institutioncode == "ARC", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "Arc Occurrences", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(ARC_obs$decimalLongitude, ARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/ARC_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOCENARC_obs <- presence[presence$institutioncode == "DFOCENARC", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Central & Arctic Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOCENARC_obs$decimalLongitude, DFOCENARC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/DFOCENARC_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOGulf_obs <- presence[presence$institutioncode == "DFOGulf", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Gulf Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOGulf_obs$decimalLongitude, DFOGulf_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/DFOGulf_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOISDM_obs <- presence[presence$institutioncode == "DFOISDM", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO ISDM Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOISDM_obs$decimalLongitude, DFOISDM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/DFOISDM_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOMTMS_obs <- presence[presence$institutioncode == "DFOMTMS", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Maritimes Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOMTMS_obs$decimalLongitude, DFOMTMS_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/DFOMTMS_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFONL_obs <- presence[presence$institutioncode == "DFONL", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Newfouundland & Labrador Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFONL_obs$decimalLongitude, DFONL_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/DFONL_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+DFOQC_obs <- presence[presence$institutioncode == "DFOQC", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "DFO Quebec Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(DFOQC_obs$decimalLongitude, DFOQC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/DFOQC_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+MaineDMR_obs <- presence[presence$institutioncode == "MaineDMR", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "Maine DMR Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(MaineDMR_obs$decimalLongitude, MaineDMR_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/MaineDMR_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+NEFSC_obs <- presence[presence$institutioncode == "NEFSC", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "NEFSC Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(NEFSC_obs$decimalLongitude, NEFSC_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/NEFSC_occurrences_all.png") #this prints a png of the plot
+dev.off() #this turns off the print command
+
+ROM_obs <- presence[presence$institutioncode == "ROM", ]
+map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
+plot(map2, xlim = c(-70, -43), ylim =c(39, 70), asp = 1, main = "Royal Ontario Museum Occurrences",  col = "cornsilk") #the x and y lim are the long-lat bounds of the map
+points(ROM_obs$decimalLongitude, ROM_obs$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
+dev.copy(png, "../output/bio/samplinginstitutions/ROM_occurrences_all.png") #this prints a png of the plot
+```
+
 # check for gear type

### 2019-01-23 src/data_exploration.Rmd
additional vif

@@
 ```
+
+As per SLR suggestion, rerun without gear
+
+```{r}
+vif_allbutgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutgear, "../output/bio/vif_allbutgear.csv", row.names = TRUE)
+vif_allbutgear
+```
+
+As per SLR suggestion, rerun without gear and most highly correlated variables
+
+```{r}
+vif_allbutgearhighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbutgearhighlycorr, "../output/bio/vif_allbutgearhighlycorr.csv", row.names = TRUE)
+vif_allbutgearhighlycorr
+```
+
+As per SLR suggestion, rerun with gear but without most highly correlated variables
+
+```{r}
+vif_allbuthighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter + gear, data = presab))
+write.csv(vif_allbuthighlycorr, "../output/bio/vif_allbuthighlycorr.csv", row.names = TRUE)
+vif_allbuthighlycorr
+
+vif_allbuthighlycorr_sq <- read.csv("../output/bio/vif_allbuthighlycorr.csv", header = TRUE)
+vif_allbuthighlycorr_sq$GVIF2Dfsq <- vif_allbuthighlycorr_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbuthighlycorr_sq, "../output/bio/vif_allbuthighlycorr_sq.csv", row.names = TRUE)
+vif_allbuthighlycorr_sq
+```
+
+ok remove one variable at a time - leave gear in
+
+remove amo_prev
+```{r}
+vif_allbutamo_prev <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbutamo_prev, "../output/bio/vif_allbutamo_prev.csv", row.names = TRUE)
+vif_allbutamo_prev
+
+vif_allbutamo_prev_sq <- read.csv("../output/bio/vif_allbutamo_prev.csv", header = TRUE)
+vif_allbutamo_prev_sq$GVIF2Dfsq <- vif_allbutamo_prev_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbutamo_prev_sq, "../output/bio/vif_allbutamo_prev_sq.csv", row.names = TRUE)
+vif_allbutamo_prev_sq
+```
+and leave gear out
+
+remove amo_prev + gear
+```{r}
+vif_allbutamo_prevgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbutamo_prevgear, "../output/bio/vif_allbutamo_prevgear.csv", row.names = TRUE)
+vif_allbutamo_prevgear
+```
+
+remove chl_depth
+```{r}
+vif_allbutchl_depth <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutchl_depth, "../output/bio/vif_allbutchl_depth.csv", row.names = TRUE)
+vif_allbutchl_depth
+
+vif_allbutchl_depth_sq <- read.csv("../output/bio/vif_allbutchl_depth.csv", header = TRUE)
+vif_allbutchl_depth_sq$GVIF2Dfsq <- vif_allbutchl_depth_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbutchl_depth_sq, "../output/bio/vif_allbutchl_depth_sq.csv", row.names = TRUE)
+vif_allbutchl_depth_sq
+```
+
+remove chl_depth and gear
+```{r}
+vif_allbutchl_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutchl_depthgear, "../output/bio/vif_allbutchl_depthgear.csv", row.names = TRUE)
+vif_allbutchl_depthgear
+```
+
+remove ssh_surface
+```{r}
+vif_allbutssh_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutssh_surface, "../output/bio/vif_allbutssh_surface.csv", row.names = TRUE)
+vif_allbutssh_surface
+
+vif_allbutssh_surface_sq <- read.csv("../output/bio/vif_allbutssh_surface.csv", header = TRUE)
+vif_allbutssh_surface_sq$GVIF2Dfsq <- vif_allbutssh_surface_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbutssh_surface_sq, "../output/bio/vif_allbutssh_surface_sq.csv", row.names = TRUE)
+vif_allbutssh_surface_sq
+```
+
+remove ssh_surface & gear
+```{r}
+vif_allbutssh_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutssh_surfacegear, "../output/bio/vif_allbutssh_surfacegear.csv", row.names = TRUE)
+vif_allbutssh_surfacegear
+```
+
+remove o2_surface
+```{r}
+vif_allbuto2_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuto2_surface, "../output/bio/vif_allbuto2_surface.csv", row.names = TRUE)
+vif_allbuto2_surface
+
+vif_allbuto2_surface_sq <- read.csv("../output/bio/vif_allbuto2_surface.csv", header = TRUE)
+vif_allbuto2_surface_sq$GVIF2Dfsq <- vif_allbuto2_surface_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbuto2_surface_sq, "../output/bio/vif_allbuto2_surface_sq.csv", row.names = TRUE)
+vif_allbuto2_surface_sq
+```
+
+remove o2_surface & gear
+```{r}
+vif_allbuto2_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuto2_surfacegear, "../output/bio/vif_allbuto2_surfacegear.csv", row.names = TRUE)
+vif_allbuto2_surfacegear
+```
+
+so ther is a question between temp-surface and o2_depth (-0.51923)... let's look at this
+
+with gear
+```{r}
+vif_allbuttemp_surface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuttemp_surface, "../output/bio/vif_allbuttemp_surface.csv", row.names = TRUE)
+vif_allbuttemp_surface
+
+vif_allbuttemp_surface_sq <- read.csv("../output/bio/vif_allbuttemp_surface.csv", header = TRUE)
+vif_allbuttemp_surface_sq$GVIF2Dfsq <- vif_allbuttemp_surface_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbuttemp_surface_sq, "../output/bio/vif_allbuttemp_surface_sq.csv", row.names = TRUE)
+vif_allbuttemp_surface_sq
+
+vif_allo2_depth <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allo2_depth, "../output/bio/vif_allo2_depth.csv", row.names = TRUE)
+vif_allo2_depth
+
+vif_allo2_depth_sq <- read.csv("../output/bio/vif_allo2_depth.csv", header = TRUE)
+vif_allo2_depth_sq$GVIF2Dfsq <- vif_allo2_depth_sq$GVIF..1..2.Df..^2
+write.csv(vif_allo2_depth_sq, "../output/bio/vif_allo2_depth_sq", row.names = TRUE)
+vif_allo2_depth_sq
+```
+
+without gear
+```{r}
+vif_allbuttemp_surfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuttemp_surfacegear, "../output/bio/vif_allbuttemp_surfacegear.csv", row.names = TRUE)
+vif_allbuttemp_surfacegear
+
+vif_allo2_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allo2_depthgear, "../output/bio/vif_allo2_depthgear.csv", row.names = TRUE)
+vif_allo2_depthgear
+```
+
+#maxent?
+
+```{r}
+library("raster")
+library("dismo")
+library("rgeos")
+```

### 2019-01-23 src/data_exploration.Rmd
ID cells (xyzt) with >1 observation

@@
 vif_allo2_depth_sq$GVIF2Dfsq <- vif_allo2_depth_sq$GVIF..1..2.Df..^2
-write.csv(vif_allo2_depth_sq, "../output/bio/vif_allo2_depth_sq", row.names = TRUE)
+write.csv(vif_allo2_depth_sq, "../output/bio/vif_allo2_depth_sq.csv", row.names = TRUE)
 vif_allo2_depth_sq
@@
 ```{r}
-vif_allbuttemp_surfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+vif_allbuttemp_surfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
 write.csv(vif_allbuttemp_surfacegear, "../output/bio/vif_allbuttemp_surfacegear.csv", row.names = TRUE)
@@
 
-vif_allo2_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+vif_allo2_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
 write.csv(vif_allo2_depthgear, "../output/bio/vif_allo2_depthgear.csv", row.names = TRUE)
 vif_allo2_depthgear
+```
+
+```{r}
+vif_test <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+vif_test
+```
+
+# check to see how many cells have >1 point per timeslice
+
+this needs to be 3D so you need to create a cell_id_3d
+```{r}
+presab$cell_id_3d <- paste(presab$cell_id, presab$depthlayerno, sep="_") #create a new column that is a unique cell_id & depth ID. 
+head(presab)
+```
+and now do one with a timeslice label
+```{r}
+presab$cell_id_xyzt <- paste(presab$cell_id_3d, presab$year, presab$month, sep="_") #create a new column that is a unique cell_id & depth ID. 
+head(presab)
+```
+
+the data is in presab$total_cell_obs_xyt 
+```{r}
+unique_pointpercell <- unique(presab$total_cell_obs_xyzt)
+unique_pointpercell
+```
+bugger... ok so lets do a frequency table
+```{r}
+celobsmore1 <- subset(presab, total_cell_obs_xyzt > 1)
+pointpercell <- with(celobsmore1, table(cell_id_xyzt, total_cell_obs_xyzt))
+write.csv(pointpercell, file = "../output/bio/pointpercell.csv")
+nrow(pointpercell)
+pointpercell
+```
+
+so there are 949 3d timesliced cells that contain more than one occurrence. 
+- different sampling days
+- different coordinates
+- risk of duplicate entries (but only if two instituions inputted the same data with slightly different coordinates and different reserch program names - see cell3985_18_2000_7 below as example)
+
+```{r}
+cell3985_18_2000_7  <- subset(celobsmore1, cell_id_xyzt == "3985_18_2000_7")
+cell3985_18_2000_7 
 ```

### 2019-01-24 src/data_exploration.Rmd
countpercell
more and more vif monthly

--- 2019-01-21 src/data_exploration.Rmd
@@
 ```
+
+As per SLR suggestion, rerun without gear
+
+```{r vif all but gear}
+vif_allbutgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutgear, "../output/bio/vif_allbutgear.csv", row.names = TRUE)
+vif_allbutgear
+```
+
+As per SLR suggestion, rerun without gear and most highly correlated variables
+
+```{r vif without gear + high corr}
+vif_allbutgearhighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbutgearhighlycorr, "../output/bio/vif_allbutgearhighlycorr.csv", row.names = TRUE)
+vif_allbutgearhighlycorr
+```
+
+As per SLR suggestion, rerun with gear but without most highly correlated variables
+
+```{r vif all but high corr}
+vif_allbuthighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter + gear, data = presab))
+write.csv(vif_allbuthighlycorr, "../output/bio/vif_allbuthighlycorr.csv", row.names = TRUE)
+vif_allbuthighlycorr
+
+vif_allbuthighlycorr_sq <- read.csv("../output/bio/vif_allbuthighlycorr.csv", header = TRUE)
+vif_allbuthighlycorr_sq$GVIF2Dfsq <- vif_allbuthighlycorr_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbuthighlycorr_sq, "../output/bio/vif_allbuthighlycorr_sq.csv", row.names = TRUE)
+vif_allbuthighlycorr_sq
+```
+
+ok remove one variable at a time - leave gear in
+
+remove amo_prev
+```{r vif all but amoprev}
+vif_allbutamo_prev <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbutamo_prev, "../output/bio/vif_allbutamo_prev.csv", row.names = TRUE)
+vif_allbutamo_prev
+
+vif_allbutamo_prev_sq <- read.csv("../output/bio/vif_allbutamo_prev.csv", header = TRUE)
+vif_allbutamo_prev_sq$GVIF2Dfsq <- vif_allbutamo_prev_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbutamo_prev_sq, "../output/bio/vif_allbutamo_prev_sq.csv", row.names = TRUE)
+vif_allbutamo_prev_sq
+```
+and leave gear out
+
+remove amo_prev + gear
+```{r vif all but amoprev + gear}
+vif_allbutamo_prevgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbutamo_prevgear, "../output/bio/vif_allbutamo_prevgear.csv", row.names = TRUE)
+vif_allbutamo_prevgear
+```
+
+remove chl_depth
+```{r vif all but chldepth}
+vif_allbutchl_depth <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutchl_depth, "../output/bio/vif_allbutchl_depth.csv", row.names = TRUE)
+vif_allbutchl_depth
+
+vif_allbutchl_depth_sq <- read.csv("../output/bio/vif_allbutchl_depth.csv", header = TRUE)
+vif_allbutchl_depth_sq$GVIF2Dfsq <- vif_allbutchl_depth_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbutchl_depth_sq, "../output/bio/vif_allbutchl_depth_sq.csv", row.names = TRUE)
+vif_allbutchl_depth_sq
+```
+
+remove chl_depth and gear
+```{r vif all but chldepth+gear}
+vif_allbutchl_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutchl_depthgear, "../output/bio/vif_allbutchl_depthgear.csv", row.names = TRUE)
+vif_allbutchl_depthgear
+```
+
+remove ssh_surface
+```{r vif all but ssh}
+vif_allbutssh_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutssh_surface, "../output/bio/vif_allbutssh_surface.csv", row.names = TRUE)
+vif_allbutssh_surface
+
+vif_allbutssh_surface_sq <- read.csv("../output/bio/vif_allbutssh_surface.csv", header = TRUE)
+vif_allbutssh_surface_sq$GVIF2Dfsq <- vif_allbutssh_surface_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbutssh_surface_sq, "../output/bio/vif_allbutssh_surface_sq.csv", row.names = TRUE)
+vif_allbutssh_surface_sq
+```
+
+remove ssh_surface & gear
+```{r vif all but ssh+gear}
+vif_allbutssh_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutssh_surfacegear, "../output/bio/vif_allbutssh_surfacegear.csv", row.names = TRUE)
+vif_allbutssh_surfacegear
+```
+
+remove o2_surface
+```{r vif all but o2surface}
+vif_allbuto2_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuto2_surface, "../output/bio/vif_allbuto2_surface.csv", row.names = TRUE)
+vif_allbuto2_surface
+
+vif_allbuto2_surface_sq <- read.csv("../output/bio/vif_allbuto2_surface.csv", header = TRUE)
+vif_allbuto2_surface_sq$GVIF2Dfsq <- vif_allbuto2_surface_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbuto2_surface_sq, "../output/bio/vif_allbuto2_surface_sq.csv", row.names = TRUE)
+vif_allbuto2_surface_sq
+```
+
+remove o2_surface & gear
+```{r vif all but o2surface+gear}
+vif_allbuto2_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuto2_surfacegear, "../output/bio/vif_allbuto2_surfacegear.csv", row.names = TRUE)
+vif_allbuto2_surfacegear
+```
+
+as per SLR chat jan 23:
+
+remove salinity_surface only
+```{r vif all but salsurface}
+vif_allbutsalinitysurface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutsalinitysurface, "../output/bio/vif_allbutsalinitysurface.csv", row.names = TRUE)
+vif_allbutsalinitysurface
+
+vif_allbutsalinitysurface_sq <- read.csv("../output/bio/vif_allbutsalinitysurface.csv", header = TRUE)
+vif_allbutsalinitysurface_sq$GVIF2Dfsq <- vif_allbutsalinitysurface_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbutsalinitysurface_sq, "../output/bio/vif_allbutsalinitysurface_sq.csv", row.names = TRUE)
+vif_allbutsalinitysurface_sq
+```
+and without gear
+
+```{r vif all but salsurface+gear}
+vif_allbutsalinitysurfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutsalinitysurfacegear, "../output/bio/vif_allbutsalinitysurfacegear.csv", row.names = TRUE)
+vif_allbutsalinitysurfacegear
+```
+
+now remove temp_surface plus highly correlated
+```{r vif allbut high corr + temp surface/gear/nogear}
+#with gear
+vif_allbuthighcorrtempsurface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbuthighcorrtempsurface, "../output/bio/vif_allbuthighcorrtempsurface.csv", row.names = TRUE)
+vif_allbuthighcorrtempsurface
+
+vif_allbuthighcorrtempsurface_sq <- read.csv("../output/bio/vif_allbuthighcorrtempsurface.csv", header = TRUE)
+vif_allbuthighcorrtempsurface_sq$GVIF2Dfsq <- vif_allbuthighcorrtempsurface_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbuthighcorrtempsurface_sq, "../output/bio/vif_allbuthighcorrtempsurface_sq.csv", row.names = TRUE)
+vif_allbuthighcorrtempsurface_sq
+
+#without gear
+vif_allbuthighcorrtempsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbuthighcorrtempsurfacegear, "../output/bio/vif_allbuthighcorrtempsurfacegear.csv", row.names = TRUE)
+vif_allbuthighcorrtempsurfacegear
+```
+
+now remove temp_surface plus salinity_surface
+```{r vif allbut temp + salinity surface/gear/nogear}
+#with gear
+vif_allbuttempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuttempsalsurface, "../output/bio/vif_allbuttempsalsurface.csv", row.names = TRUE)
+vif_allbuttempsalsurface
+
+vif_allbuttempsalsurface_sq <- read.csv("../output/bio/vif_allbuttempsalsurface.csv", header = TRUE)
+vif_allbuttempsalsurface_sq$GVIF2Dfsq <- vif_allbuttempsalsurface_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbuttempsalsurface_sq, "../output/bio/vif_allbuttempsalsurface_sq.csv", row.names = TRUE)
+vif_allbuttempsalsurface_sq
+
+#without gear
+vif_allbuttempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuttempsalsurfacegear, "../output/bio/vif_allbuttempsalsurfacegear.csv", row.names = TRUE)
+vif_allbuttempsalsurfacegear
+```
+
+now remove temp_surface plus salinity_surface + highly correlated
+```{r vif allbut high corr +  temp + salinity surface/gear/nogear}
+#with gear
+vif_allbuthighcorrtempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbuthighcorrtempsalsurface, "../output/bio/vif_allbuthighcorrtempsalsurface.csv", row.names = TRUE)
+vif_allbuthighcorrtempsalsurface
+
+vif_allbuthighcorrtempsalsurface_sq <- read.csv("../output/bio/vif_allbuthighcorrtempsalsurface.csv", header = TRUE)
+vif_allbuthighcorrtempsalsurface_sq$GVIF2Dfsq <- vif_allbuthighcorrtempsalsurface_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbuthighcorrtempsalsurface_sq, "../output/bio/vif_allbuthighcorrtempsalsurface_sq.csv", row.names = TRUE)
+vif_allbuthighcorrtempsalsurface_sq
+
+#without gear
+vif_allbuthighcorrtempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbuthighcorrtempsalsurfacegear, "../output/bio/vif_allbuthighcorrtempsalsurfacegear.csv", row.names = TRUE)
+vif_allbuthighcorrtempsalsurfacegear
+```
+
+Just out of curiosity, a vif will all but but atmos drivers
+
+```{r}
+vif_allbutnaoamo <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear, data = presab))
+write.csv(vif_allbutnaoamo, "../output/bio/vif_allbutnaoamo.csv", row.names = TRUE)
+vif_allbutnaoamo
+
+vif_allbutnaoamo_sq <- read.csv("../output/bio/vif_allbutnaoamo.csv", header = TRUE)
+vif_allbutnaoamo_sq$GVIF2Dfsq <- vif_allbutnaoamo_sq$GVIF..1..2.Df..^2
+write.csv(vif_allbutnaoamo_sq, "../output/bio/vif_allbutnaoamo_sq.csv", row.names = TRUE)
+vif_allbutnaoamo_sq
+```
+
+# check to see how many cells have >1 point per timeslice
+
+this needs to be 3D so you need to create a cell_id_3d
+```{r}
+presab$cell_id_3d <- paste(presab$cell_id, presab$depthlayerno, sep="_") #create a new column that is a unique cell_id & depth ID. 
+head(presab)
+```
+and now do one with a timeslice label
+```{r}
+presab$cell_id_xyzt <- paste(presab$cell_id_3d, presab$year, presab$month, sep="_") #create a new column that is a unique cell_id & depth ID. 
+head(presab)
+```
+
+the data is in presab$total_cell_obs_xyt 
+```{r}
+unique_pointpercell <- unique(presab$total_cell_obs_xyzt)
+unique_pointpercell
+```
+bugger... ok so lets do a frequency table
+```{r}
+celobsmore1 <- subset(presab, total_cell_obs_xyzt > 1)
+pointpercell <- with(celobsmore1, table(cell_id_xyzt, total_cell_obs_xyzt))
+write.csv(pointpercell, file = "../output/bio/pointpercell.csv")
+nrow(pointpercell)
+pointpercell
+```
+
+so there are 949 3d timesliced cells that contain more than one occurrence. 
+- different sampling days
+- different coordinates
+- risk of duplicate entries (but only if two instituions inputted the same data with slightly different coordinates and different reserch program names - see cell3985_18_2000_7 below as example)
+
+```{r}
+cell3985_18_2000_7  <- subset(celobsmore1, cell_id_xyzt == "3985_18_2000_7")
+cell3985_18_2000_7 
+```
+
+# monthly spearmans correlations and VIF
+
+curious - does correlations/vif alter between months?
+
+First split the dataset into monthly
+
+```{r subset monthly presab}
+janpresab <- subset(presab, month == "1")
+febpresab <- subset(presab, month == "2")
+marpresab <- subset(presab, month == "3")
+aprpresab <- subset(presab, month == "4")
+maypresab <- subset(presab, month == "5")
+junpresab <- subset(presab, month == "6")
+julpresab <- subset(presab, month == "7")
+augpresab <- subset(presab, month == "8")
+seppresab <- subset(presab, month == "9")
+octpresab <- subset(presab, month == "10")
+novpresab <- subset(presab, month == "11")
+decpresab <- subset(presab, month == "12")
+```
+
+now get the background points for each month (for spearmans)
+```{r monthly background points}
+janback <- subset(janpresab, occurrence == "0")
+febback <- subset(febpresab, occurrence == "0")
+marback <- subset(marpresab, occurrence == "0")
+aprback <- subset(aprpresab, occurrence == "0")
+mayback <- subset(maypresab, occurrence == "0")
+junback <- subset(junpresab, occurrence == "0")
+julback <- subset(julpresab, occurrence == "0")
+augback <- subset(augpresab, occurrence == "0")
+sepback <- subset(seppresab, occurrence == "0")
+octback <- subset(octpresab, occurrence == "0")
+novback <- subset(novpresab, occurrence == "0")
+decback <- subset(decpresab, occurrence == "0")
+```
+
+run the correlations on each month's background points
+
+```{r monthly spearman correlations}
+janback <- subset(janback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+janback_cor <- cor(janback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(janback_cor, "../output/env/janback_cor.csv", row.names = TRUE)
+
+febback <- subset(febback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+febback_cor <- cor(febback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(febback_cor, "../output/env/febback_cor.csv", row.names = TRUE)
+
+marback <- subset(marback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+marback_cor <- cor(marback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(marback_cor, "../output/env/marback_cor.csv", row.names = TRUE)
+
+aprback <- subset(aprback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+aprback_cor <- cor(aprback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(aprback_cor, "../output/env/aprback_cor.csv", row.names = TRUE)
+
+mayback <- subset(mayback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+mayback_cor <- cor(mayback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(mayback_cor, "../output/env/mayback_cor.csv", row.names = TRUE)
+
+junback <- subset(junback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+junback_cor <- cor(junback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(junback_cor, "../output/env/junback_cor.csv", row.names = TRUE)
+
+julback <- subset(julback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+julback_cor <- cor(julback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(julback_cor, "../output/env/julback_cor.csv", row.names = TRUE)
+
+augback <- subset(augback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+augback_cor <- cor(augback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(augback_cor, "../output/env/augback_cor.csv", row.names = TRUE)
+
+sepback <- subset(sepback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+sepback_cor <- cor(sepback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(sepback_cor, "../output/env/sepback_cor.csv", row.names = TRUE)
+
+octback <- subset(octback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+octback_cor <- cor(octback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(octback_cor, "../output/env/octback_cor.csv", row.names = TRUE)
+
+novback <- subset(novback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+novback_cor <- cor(novback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(novback_cor, "../output/env/novback_cor.csv", row.names = TRUE)
+
+decback <- subset(decback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+decback_cor <- cor(decback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
+complete.obs means na.rm in this package.
+#using spearmans as suitable for non-linear relationships
+write.csv(decback_cor, "../output/env/decback_cor.csv", row.names = TRUE)
+```
+
+#maxent?
+
+```{r}
+library("raster")
+library("dismo")
+library("rgeos")
+```

### 2019-01-28 src/data_exploration.Rmd
oceanographic variables by nafo region (presence points)

@@
 first see which nafo zones were sampled in each year
-```{r}
+```{r presence nafo by year}
 nafo_by_yr <- with(presence, table(year, nafo_zone))
@@
 and by month
-```{r}
+```{r prsence nafo by month}
 nafo_by_mth <- with(presence, table(nafo_zone, month))
@@
 onec
+```
+
+## density plot with background and presence env. data by NAFO region
+
+What I want to see if if there is a marked difference between the env.
+correlates of the presence points between NAFO regions.
+This is also to try deal with sampling bias (b/c the whole region is not uniformly sampled in each month, but rather nafo regions have a strong month bias)
+
+first create NAFO-region datasets
+
+```{r presence by nafo}
+nafo0a <- subset(presence, nafo_zone == "0A")
+nafo0b <- subset(presence, nafo_zone == "0B")
+nafo1c <- subset(presence, nafo_zone == "1C")
+nafo2g <- subset(presence, nafo_zone == "2G")
+nafo2h <- subset(presence, nafo_zone == "2H")
+nafo2j <- subset(presence, nafo_zone == "2J")
+nafo3k <- subset(presence, nafo_zone == "3K")
+nafo3l <- subset(presence, nafo_zone == "3L")
+nafo3m <- subset(presence, nafo_zone == "3M")
+nafo3n <- subset(presence, nafo_zone == "3N")
+nafo3o <- subset(presence, nafo_zone == "3O")
+nafo3ps <- subset(presence, nafo_zone == "3Ps")
+nafo4r <- subset(presence, nafo_zone == "4R")
+nafo4s <- subset(presence, nafo_zone == "4S")
+nafo4t <- subset(presence, nafo_zone == "4T")
+nafo4vn <- subset(presence, nafo_zone == "4Vn")
+nafo4vs <- subset(presence, nafo_zone == "4Vs")
+nafo4w <- subset(presence, nafo_zone == "4W")
+nafo4x <- subset(presence, nafo_zone == "4X")
+nafo5y <- subset(presence, nafo_zone == "5Y")
+nafo5ze <- subset(presence, nafo_zone == "5ZE")
+nafohudson <- subset(presence, nafo_zone == "HudsonStrait")
+```
+
+plot by each variable, less 3m (2 samples) 1c (one sample) and 5ze (zero samples?!)
+```{r presence by nafo by variable}
+ggplot(nafo0a, aes(x = temp_surface, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/temp_surface_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(nafo0a, aes(x = temp_depth, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/temp_depth_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(nafo0a, aes(x = chl_surface, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/chl_surface_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(nafo0a, aes(x = chl_depth, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/chl_depth_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(nafo0a, aes(x = salinity_surface, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/salinity_surface_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(nafo0a, aes(x = salinity_depth, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/salinity_depth_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(nafo0a, aes(x = o2_surface, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/o2_surface_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(nafo0a, aes(x = salinity_depth, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/o2_depth_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(nafo0a, aes(x = mlp_surface, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/mlp_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(nafo0a, aes(x = ssh_surface, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/ssh_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
 ```

### 2019-01-29 src/data_exploration.Rmd
fuckit

@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car", "rworldmap") 
+packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car", "rworldmap", "fastmatch") 
 source("../src/ipak.R")
@@
 
-ggplot(nafo0a, aes(x = salinity_depth, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+ggplot(nafo0a, aes(x = o2_depth, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
 dev.copy(png,"../output/env/env_by_nafo_plots/o2_depth_nafo.png") # to automatically save the plot to a png AND show it inline
@@
 
+What I want to see if if there is a marked difference between the env.
+correlates of the presence points between gear type used.
+This is also to try deal with detection bias between gear type (b/c the whole region is not uniformly sampled by the same gear type)
+
+first create gear datasets
+
+```{r presence by gear}
+presence$gear <- as.character(presence$gear)
+bottom_trawl <- subset(presence, gear == "bottom_trawl")
+bottom_trawl_alfredo_03 <- subset(presence, gear == "bottom_trawl_alfredo_03")
+bottom_trawl_campelen_14 <- subset(presence, gear == "bottom_trawl_campelen_14")
+bottom_trawl_campelen_1800 <- subset(presence, gear == "bottom_trawl_campelen_1800")
+bottom_trawl_campelen_21 <- subset(presence, gear == "bottom_trawl_campelen_21")
+bottom_trawl_cosmos_2600 <- subset(presence, gear == "bottom_trawl_cosmos_2600")
+bottom_trawl_western_IIA <- subset(presence, gear == "bottom_trawl_western_IIA")
+unknown <- subset(presence, gear == "unknown")
+vertical_plankton_tow <- subset(presence, gear == "vertical_plankton_tow")
+```
+
+plot by each variable, less 3m (2 samples) 1c (one sample) and 5ze (zero samples?!)
+```{r presence by gear by variable}
+ggplot(bottom_trawl, aes(x = temp_surface, colour = gear)) + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/temp_surface_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(bottom_trawl, aes(x = temp_depth, colour = gear)) + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/temp_depth_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(bottom_trawl, aes(x = chl_surface, colour = gear)) + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/chl_surface_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(bottom_trawl, aes(x = chl_depth, colour = gear))  + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/chl_depth_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(bottom_trawl, aes(x = salinity_surface, colour = gear)) + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/salinity_surface_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(bottom_trawl, aes(x = salinity_depth, colour = gear)) + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/salinity_depth_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(bottom_trawl, aes(x = o2_surface, colour = gear)) + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/o2_surface_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(bottom_trawl, aes(x = o2_depth, colour = gear)) + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/o2_depth_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(bottom_trawl, aes(x = mlp_surface, colour = gear)) + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/mlp_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(bottom_trawl, aes(x = ssh_surface, colour = gear)) + geom_density(na.rm = TRUE) + geom_density(data=bottom_trawl_alfredo_03, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_14, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_1800, na.rm = TRUE) + geom_density(data=bottom_trawl_campelen_21, na.rm = TRUE) + geom_density(data=bottom_trawl_cosmos_2600, na.rm = TRUE) + geom_density(data=bottom_trawl_western_IIA, na.rm = TRUE) + geom_density(data=unknown, na.rm = TRUE) + geom_density(data=vertical_plankton_tow, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_gear_plots/ssh_gear.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
 ```{r}
@@
 this needs to be 3D so you need to create a cell_id_3d
-```{r}
+```{r cell_id_3d}
 presab$cell_id_3d <- paste(presab$cell_id, presab$depthlayerno, sep="_") #create a new column that is a unique cell_id & depth ID. 
@@
 and now do one with a timeslice label
-```{r}
+```{r cell_id_xyzt}
 presab$cell_id_xyzt <- paste(presab$cell_id_3d, presab$year, presab$month, sep="_") #create a new column that is a unique cell_id & depth ID. 
 head(presab)
-```
-
-the data is in presab$total_cell_obs_xyt 
-```{r}
-unique_pointpercell <- unique(presab$total_cell_obs_xyzt)
+write.csv(presab, "../output/bio/presab_cellid_xyzt.csv", row.names = FALSE)
+```
+
+because background points can land in the same xyzt, subset the dataset to be only presences
+
+```{r}
+presencedup <- subset(presab, occurrence == "1")
+nrow(presencedup)
+```
+
+the data is in presencedup$total_cell_obs_xyzt
+```{r unique_pointpercell}
+unique_pointpercell <- unique(presencedup$total_cell_obs_xyzt)
 unique_pointpercell
@@
 bugger... ok so lets do a frequency table
-```{r}
-celobsmore1 <- subset(presab, total_cell_obs_xyzt > 1)
-pointpercell <- with(celobsmore1, table(cell_id_xyzt, total_cell_obs_xyzt))
-write.csv(pointpercell, file = "../output/bio/pointpercell.csv")
-nrow(pointpercell)
-pointpercell
-```
-
-so there are 949 3d timesliced cells that contain more than one occurrence. 
+```{r unique_pointpercell frequencies}
+pointpercellcount <- count(celobsmore1, "total_cell_obs_xyzt")
+pointpercellcount$no_cells <- pointpercellcount$freq / pointpercellcount$total_cell_obs_xyzt
+write.csv(pointpercellcount, file = "../output/bio/pointpercellcount.csv")
+pointpercellcount
+```
+
+so there are 949 3d timesliced cells that contain more than one occurrence, and 2042 rows. 
 - different sampling days
@@
 cell3985_18_2000_7 
+```
+
+Ideally you only want one observation cell cell_id_xyzt. First see if any are missing depth data
+
+find the column numbers that correspond to the env. data (use fastmatch package as is quicker)
+
+```{r}
+fmatch("chl_surface", names(celobsmore1)) #28
+fmatch("chl_depth", names(celobsmore1)) #29
+fmatch("mlp_surface", names(celobsmore1)) #30
+fmatch("o2_surface", names(celobsmore1)) #31
+fmatch("o2_depth", names(celobsmore1)) #32
+fmatch("salinity_surface", names(celobsmore1)) #33
+fmatch("salinity_depth", names(celobsmore1)) #34
+fmatch("ssh_surface", names(celobsmore1)) #35
+fmatch("temp_surface", names(celobsmore1)) #36
+fmatch("temp_depth", names(celobsmore1)) #37
+```
+
+then subset any with NA values
+
+```{r celobsmore1_novals}
+celobsmore1_novals <-  celobsmore1[!complete.cases(celobsmore1[28:37]), ] #28:37 represent the columns w/ env. data
+nrow(celobsmore1_novals) #just to check how many there are - 1095 out of 2042
+```
+
+ok and subset with all values
+
+```{r celobsmore1_allval}
+celobsmore1_allvals <- celobsmore1[complete.cases(celobsmore1[28:37]), ] #28:37 represent the columns w/ env. data
+nrow(celobsmore1_allvals) #947 out of 2042
+```
+
+ok so how many unique cells do we have with complete data...
+
+```{r}
+pointpercellcount_alldata <- count(celobsmore1_allvals, "total_cell_obs_xyzt")
+pointpercellcount_alldata$no_cells <- pointpercellcount_alldata$freq / pointpercellcount_alldata$total_cell_obs_xyzt
+write.csv(pointpercellcount_alldata, file = "../output/bio/pointpercellcount_alldata.csv")
+pointpercellcount_alldata
+```
+
+and see which cell_id_xyzt appear in both celobsmore1_allvals and celobsmore1_novals
+```{r}
+celobsmore1_allnovals <- inner_join(celobsmore1_allvals, celobsmore1_novals)
+celobsmore1_allnovals 
+```
+None.... which is good. I'd expect observations in the same xyzt to have the same values
+
+so, I can successfully reduce the observations down to one per xyzt
+
+```{r}
+fmatch("cell_id_xyzt", names(celobsmore1)) #46
+duplicatesbegone <- celobsmore1[!duplicated(celobsmore1[ ,46]), ] #creates a df with all duplicates gone (now 1 obs per cell)
+nrow(duplicatesbegone) #949 - good
+presabnodup <- subset(presab, total_cell_obs_xyzt == "1") #remove all rows that have more than one obs
+nrow(presabnodup) #1549351 - this adds up 
+presab <- rbind(presabnodup, duplicatesbegone)
+nrow(presab) #1550300 - great!
+
+#recalculate the total_cell_obs_xyzt column
+
+names(presab)[names(presab)=="total_cell_obs_xyzt"] <- "XXtotal_cell_obs_xyzt" #rename col
+obs_cell_yymm_depth_dup <- count(presab, cell_id, year, month, depthlayerno)
+test <- merge(presab, obs_cell_yymm_depth_dup, by.x=c("cell_id", "year", "month", "depthlayerno"), by.y=c("cell_id", "year", "month", "depthlayerno"))
+names(test)[names(test)=="n"] <- "total_cell_obs_xyzt" #rename col
+
+write.csv(presab, "../output/bio/presab_cellid_xyzt_nodup.csv")
+rm(presabnodup, duplicatesbegone, celobsmore1, celobsmore1_novals, celobsmore1_allnovals, celobsmore1_allvals) #just because they are bigish variables
 ```

### 2019-01-30 src/biological_data_prep.Rmd
moved where to deal with duplicates.

@@
 ```
+
+# remove any duplicate entries
+
+Check to see how many cells have >1 point per timeslice. This needs to be timesliced and 3D
+
+```{r cell_id_3d}
+data_aea$cell_id_3d <- paste(data_aea$cell_id, data_aea$depthlayerno, sep="_") #create a new column that is a unique cell_id & depth ID. 
+head(data_aea)
+```
+
+and now do one with a timeslice label
+```{r cell_id_xyzt}
+data_aea$cell_id_xyzt <- paste(data_aea$cell_id_3d, data_aea$year, data_aea$month, sep="_") #create a new column that is a unique cell_id & depth ID. 
+head(data_aea)
+write.csv(data_aea, "../output/bio/presence_points_without_envdata_relooped_glbathy_cellid_xyzt.csv", row.names = FALSE)
+```
+
+ok lets see how many have more than one obs per cell
+```{r unique_pointpercell}
+unique_pointpercell <- unique(data_aea$total_cell_obs_xyzt)
+unique_pointpercell
+```
+
+bugger... ok so lets do a frequency table
+```{r unique_pointpercell frequencies}
+celobsmore1 <- subset(data_aea, total_cell_obs_xyzt >1)
+celobsmore1 #2042 rows
+pointpercellcount <- count(celobsmore1, "total_cell_obs_xyzt")
+pointpercellcount$no_cells <- pointpercellcount$freq / pointpercellcount$total_cell_obs_xyzt
+write.csv(pointpercellcount, file = "../output/bio/pointpercellcount_pres.csv")
+pointpercellcount
+```
+
+so there are 949 3d timesliced cells that contain more than one occurrence, and 2042 rows. 
+- different sampling days
+- different coordinates
+- risk of duplicate entries (but only if two instituions inputted the same data with slightly different coordinates and different reserch program names - see cell3985_18_2000_7 below as example)
+
+Ideally you only want one observation cell cell_id_xyzt. First see if any are missing depth data
+
+find the column numbers that correspond to the env. data (use fastmatch package as is quicker)
+
+```{r}
+fmatch("chl_surface", names(celobsmore1)) #28
+fmatch("chl_depth", names(celobsmore1)) #29
+fmatch("mlp_surface", names(celobsmore1)) #30
+fmatch("o2_surface", names(celobsmore1)) #31
+fmatch("o2_depth", names(celobsmore1)) #32
+fmatch("salinity_surface", names(celobsmore1)) #33
+fmatch("salinity_depth", names(celobsmore1)) #34
+fmatch("ssh_surface", names(celobsmore1)) #35
+fmatch("temp_surface", names(celobsmore1)) #36
+fmatch("temp_depth", names(celobsmore1)) #37
+```
+
+then subset any with NA values
+
+```{r celobsmore1_novals}
+celobsmore1_novals <-  celobsmore1[!complete.cases(celobsmore1[28:37]), ] #28:37 represent the columns w/ env. data
+nrow(celobsmore1_novals) #just to check how many there are - 1095 out of 2042
+```
+
+ok and subset with all values
+
+```{r celobsmore1_allval}
+celobsmore1_allvals <- celobsmore1[complete.cases(celobsmore1[28:37]), ] #28:37 represent the columns w/ env. data
+nrow(celobsmore1_allvals) #947 out of 2042
+```
+
+ok so how many unique cells do we have with complete data...
+
+```{r}
+pointpercellcount_alldata <- count(celobsmore1_allvals, "total_cell_obs_xyzt")
+pointpercellcount_alldata$no_cells <- pointpercellcount_alldata$freq / pointpercellcount_alldata$total_cell_obs_xyzt
+write.csv(pointpercellcount_alldata, file = "../output/bio/pointpercellcount_alldata.csv")
+pointpercellcount_alldata
+```
+
+and see which cell_id_xyzt appear in both celobsmore1_allvals and celobsmore1_novals
+```{r}
+celobsmore1_allnovals <- inner_join(celobsmore1_allvals, celobsmore1_novals)
+celobsmore1_allnovals 
+```
+None.... which is good. I'd expect observations in the same xyzt to have the same values
+
+so, I can successfully reduce the observations down to one per xyzt
+
+```{r}
+fmatch("cell_id_xyzt", names(celobsmore1)) #50
+duplicatesbegone <- celobsmore1[!duplicated(celobsmore1[ ,50]), ] #creates a df with all duplicates gone (now 1 obs per cell)
+nrow(duplicatesbegone) #949 - good
+presabnodup <- subset(data_aea, total_cell_obs_xyzt == "1") #remove all rows that have more than one obs
+nrow(presabnodup) #9351 - this adds up 
+data_aea <- rbind(presabnodup, duplicatesbegone)
+nrow(data_aea) #10300 - great!
+```
+
+#recalculate the total_cell_obs_xyzt column
+```{r}
+obs_cell_yymm_depth <- count(data_aea, c("cell_id", "year", "month", "depthlayerno"))
+data_aea <- merge(data_aea, obs_cell_yymm_depth, by.x=c("cell_id", "year", "month", "depthlayerno"), by.y=c("cell_id", "year", "month", "depthlayerno"))
+names(data_aea)[names(data_aea)=="freq"] <- "total_cell_obs_xyzt"
+head(data_aea)
+```
+
+check for unique values(should be 1 only)
+```{r}
+unique_pointpercell <- unique(data_aea$total_cell_obs_xyzt)
+unique_pointpercell
+```
+Great! Save
+
+```{r}
+write.csv(data_aea, file = "../output/bio/presence_points_without_envdata_relooped_glbathy_nodup.csv", row.names = FALSE) # change the path to where ever you would like to save it
+```

### 2019-01-30 src/data_exploration.Rmd
moved where to deal with duplicates.

@@
 
-# presence points
-Read in the presence points
-
-```{r}
-presence <- read.csv("../output/bio/presence_points_without_envdata_relooped_glbathy.csv", header = TRUE)
-head(presence)
+# read in all data
+
+and subset into background and presences
+```{r}
+presab <- read.csv("../output/bio/presab.csv", header = TRUE)
+presence <- subset(presab, occurrence == "1")
+background <- subset(presab, occurrence == "0")
 ```
@@
 # background points
-Read in the presence points
-
-```{r}
-background <- read.csv("../output/bio/background_complete_obs_cels_globot.csv", header = TRUE)
-head(background)
-```
 
@@
 ```{r presence nafo by year}
+presence$nafo_zone <- as.character(presence$nafo_zone)
 nafo_by_yr <- with(presence, table(year, nafo_zone))
@@
 
-Interesting that there is a point in 1C - this is outside Canadian waters.... anyway
-```{r}
-onec <-  subset(presence, nafo_zone == "1C")
-onec
+Interesting that there is a point in 1C - this is outside Canadian waters remove from the dataset
+```{r}
+presab <- presab[!(presab$nafo_zone == "1C" & presab$occurrence == "1"), ]
+write.csv(presab, "../output/bio/presab.csv", row.names = FALSE)
+presence <- presence[!(presence$nafo_zone == "1C"), ]
+```
+
+and by month again
+```{r prsence nafo by month2}
+nafo_by_mth <- with(presence, table(nafo_zone, month))
+write.csv(nafo_by_mth, file = "../output/bio/nafozone_by_mth.csv")
+nafo_by_mth
 ```
@@
 nafo0b <- subset(presence, nafo_zone == "0B")
-nafo1c <- subset(presence, nafo_zone == "1C")
 nafo2g <- subset(presence, nafo_zone == "2G")
@@
 
-plot by each variable, less 3m (2 samples) 1c (one sample) and 5ze (zero samples?!)
+plot by each variable, less 3m (2 samples) and 5ze (2 samples)
 ```{r presence by nafo by variable}
@@
 ```
-ok so IMR, NAFO, ICES, & USM have very few entries (1, 1, 1, and 3 respectively)
+ok so NEFSC and ROM only have one point each
 
@@
 
-for this you need the joined dataset. this was created under mergig_datasets.src and the file is ../output/bio/presab.csv
-
-```{r}
-presab <- read.csv("../output/bio/presab.csv", header = TRUE)
-```
+for this you need the joined dataset.
 
@@
 ```
-
-# check to see how many cells have >1 point per timeslice
-
-this needs to be 3D so you need to create a cell_id_3d
-```{r cell_id_3d}
-presab$cell_id_3d <- paste(presab$cell_id, presab$depthlayerno, sep="_") #create a new column that is a unique cell_id & depth ID. 
-head(presab)
-```
-and now do one with a timeslice label
-```{r cell_id_xyzt}
-presab$cell_id_xyzt <- paste(presab$cell_id_3d, presab$year, presab$month, sep="_") #create a new column that is a unique cell_id & depth ID. 
-head(presab)
-write.csv(presab, "../output/bio/presab_cellid_xyzt.csv", row.names = FALSE)
-```
-
-because background points can land in the same xyzt, subset the dataset to be only presences
-
-```{r}
-presencedup <- subset(presab, occurrence == "1")
-nrow(presencedup)
-```
-
-the data is in presencedup$total_cell_obs_xyzt
-```{r unique_pointpercell}
-unique_pointpercell <- unique(presencedup$total_cell_obs_xyzt)
-unique_pointpercell
-```
-bugger... ok so lets do a frequency table
-```{r unique_pointpercell frequencies}
-pointpercellcount <- count(celobsmore1, "total_cell_obs_xyzt")
-pointpercellcount$no_cells <- pointpercellcount$freq / pointpercellcount$total_cell_obs_xyzt
-write.csv(pointpercellcount, file = "../output/bio/pointpercellcount.csv")
-pointpercellcount
-```
-
-so there are 949 3d timesliced cells that contain more than one occurrence, and 2042 rows. 
-- different sampling days
-- different coordinates
-- risk of duplicate entries (but only if two instituions inputted the same data with slightly different coordinates and different reserch program names - see cell3985_18_2000_7 below as example)
-
-```{r}
-cell3985_18_2000_7  <- subset(celobsmore1, cell_id_xyzt == "3985_18_2000_7")
-cell3985_18_2000_7 
-```
-
-Ideally you only want one observation cell cell_id_xyzt. First see if any are missing depth data
-
-find the column numbers that correspond to the env. data (use fastmatch package as is quicker)
-
-```{r}
-fmatch("chl_surface", names(celobsmore1)) #28
-fmatch("chl_depth", names(celobsmore1)) #29
-fmatch("mlp_surface", names(celobsmore1)) #30
-fmatch("o2_surface", names(celobsmore1)) #31
-fmatch("o2_depth", names(celobsmore1)) #32
-fmatch("salinity_surface", names(celobsmore1)) #33
-fmatch("salinity_depth", names(celobsmore1)) #34
-fmatch("ssh_surface", names(celobsmore1)) #35
-fmatch("temp_surface", names(celobsmore1)) #36
-fmatch("temp_depth", names(celobsmore1)) #37
-```
-
-then subset any with NA values
-
-```{r celobsmore1_novals}
-celobsmore1_novals <-  celobsmore1[!complete.cases(celobsmore1[28:37]), ] #28:37 represent the columns w/ env. data
-nrow(celobsmore1_novals) #just to check how many there are - 1095 out of 2042
-```
-
-ok and subset with all values
-
-```{r celobsmore1_allval}
-celobsmore1_allvals <- celobsmore1[complete.cases(celobsmore1[28:37]), ] #28:37 represent the columns w/ env. data
-nrow(celobsmore1_allvals) #947 out of 2042
-```
-
-ok so how many unique cells do we have with complete data...
-
-```{r}
-pointpercellcount_alldata <- count(celobsmore1_allvals, "total_cell_obs_xyzt")
-pointpercellcount_alldata$no_cells <- pointpercellcount_alldata$freq / pointpercellcount_alldata$total_cell_obs_xyzt
-write.csv(pointpercellcount_alldata, file = "../output/bio/pointpercellcount_alldata.csv")
-pointpercellcount_alldata
-```
-
-and see which cell_id_xyzt appear in both celobsmore1_allvals and celobsmore1_novals
-```{r}
-celobsmore1_allnovals <- inner_join(celobsmore1_allvals, celobsmore1_novals)
-celobsmore1_allnovals 
-```
-None.... which is good. I'd expect observations in the same xyzt to have the same values
-
-so, I can successfully reduce the observations down to one per xyzt
-
-```{r}
-fmatch("cell_id_xyzt", names(celobsmore1)) #46
-duplicatesbegone <- celobsmore1[!duplicated(celobsmore1[ ,46]), ] #creates a df with all duplicates gone (now 1 obs per cell)
-nrow(duplicatesbegone) #949 - good
-presabnodup <- subset(presab, total_cell_obs_xyzt == "1") #remove all rows that have more than one obs
-nrow(presabnodup) #1549351 - this adds up 
-presab <- rbind(presabnodup, duplicatesbegone)
-nrow(presab) #1550300 - great!
-
-#recalculate the total_cell_obs_xyzt column
-
-names(presab)[names(presab)=="total_cell_obs_xyzt"] <- "XXtotal_cell_obs_xyzt" #rename col
-obs_cell_yymm_depth_dup <- count(presab, cell_id, year, month, depthlayerno)
-test <- merge(presab, obs_cell_yymm_depth_dup, by.x=c("cell_id", "year", "month", "depthlayerno"), by.y=c("cell_id", "year", "month", "depthlayerno"))
-names(test)[names(test)=="n"] <- "total_cell_obs_xyzt" #rename col
-
-write.csv(presab, "../output/bio/presab_cellid_xyzt_nodup.csv")
-rm(presabnodup, duplicatesbegone, celobsmore1, celobsmore1_novals, celobsmore1_allnovals, celobsmore1_allvals) #just because they are bigish variables
-```
+not much diff..
 
@@
 
+Run a VIF for each month (with and without gear)
+
+```{r}
+
+```
+
 #maxent?

### 2019-01-30 src/data_exploration.Rmd
monthly vifs

@@
 ```{r}
-gearby_mth <- with(presence, table(month, gear))
+gearby_mth <- with(presence, table(gear, month))
 write.csv(gearby_mth, file = "../output/bio/gear_by_mth.csv")
@@
 ```{r subset monthly presab}
+presab$gear <- as.character(presab$gear)
 janpresab <- subset(presab, month == "1")
@@
 
+#with gear - for jan there is none because there is only one category of gear
+
+#without gear
+vif_nogear_jan <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface, data = janpresab))
+write.csv(vif_nogear_jan, "../output/bio/monthly_vifs/vif_nogear_jan.csv", row.names = TRUE)
+vif_nogear_jan
+
+#with gear
+vif_gear_feb <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface, data = febpresab))
+write.csv(vif_gear_feb, "../output/bio/monthly_vifs/vif_gear_feb.csv", row.names = TRUE)
+vif_gear_feb
+
+vif_gear_feb <- read.csv("../output/bio/monthly_vifs/vif_gear_feb.csv", header = TRUE)
+vif_gear_feb$GVIF2Dfsq <- vif_gear_feb$GVIF..1..2.Df..^2
+write.csv(vif_gear_feb, "../output/bio/monthly_vifs/vif_gear_feb.csv", row.names = TRUE)
+vif_gear_feb
+
+#without gear
+vif_nogear_feb <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface, data = febpresab))
+write.csv(vif_nogear_feb, "../output/bio/monthly_vifs/vif_nogear_feb.csv", row.names = TRUE)
+vif_nogear_feb
+
+#with gear
+vif_gear_mar <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = marpresab))
+write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
+vif_gear_mar
+
+vif_gear_mar <- read.csv("../output/bio/monthly_vifs/vif_gear_mar.csv", header = TRUE)
+vif_gear_mar$GVIF2Dfsq <- vif_gear_mar$GVIF..1..2.Df..^2
+write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
+vif_gear_mar
+
+#without gear
+vif_nogear_mar <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = marpresab))
+write.csv(vif_nogear_mar, "../output/bio/monthly_vifs/vif_nogear_mar.csv", row.names = TRUE)
+vif_nogear_mar
+
+#with gear
+vif_gear_apr <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
+write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
+vif_gear_apr
+
+vif_gear_apr <- read.csv("../output/bio/monthly_vifs/vif_gear_apr.csv", header = TRUE)
+vif_gear_apr$GVIF2Dfsq <- vif_gear_apr$GVIF..1..2.Df..^2
+write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
+vif_gear_apr
+
+#without gear
+vif_nogear_apr <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
+write.csv(vif_nogear_apr, "../output/bio/monthly_vifs/vif_nogear_apr.csv", row.names = TRUE)
+vif_nogear_apr
+
+#with gear
+vif_gear_may <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = maypresab))
+write.csv(vif_gear_may, "../output/bio/monthly_vifs/vif_gear_may.csv", row.names = TRUE)
+vif_gear_may
+
+#without gear
+vif_nogear_may <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = maypresab))
+write.csv(vif_nogear_may, "../output/bio/monthly_vifs/vif_nogear_may.csv", row.names = TRUE)
+vif_nogear_may
+
+#with gear
+vif_gear_jun <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+write.csv(vif_gear_jun, "../output/bio/monthly_vifs/vif_gear_jun.csv", row.names = TRUE)
+vif_gear_jun
+
+#without gear
+vif_nogear_jun <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+write.csv(vif_nogear_jun, "../output/bio/monthly_vifs/vif_nogear_jun.csv", row.names = TRUE)
+vif_nogear_jun
+
+#with gear
+vif_gear_jul <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = julpresab))
+write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
+vif_gear_jul
+
+vif_gear_jul <- read.csv("../output/bio/monthly_vifs/vif_gear_jul.csv", header = TRUE)
+vif_gear_jul$GVIF2Dfsq <- vif_gear_jul$GVIF..1..2.Df..^2
+write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
+vif_gear_jul
+
+#without gear
+vif_nogear_jul <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = julpresab))
+write.csv(vif_nogear_jul, "../output/bio/monthly_vifs/vif_nogear_jul.csv", row.names = TRUE)
+vif_nogear_jul
+
+#with gear
+vif_gear_aug <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = augpresab))
+write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
+vif_gear_aug
+
+vif_gear_aug <- read.csv("../output/bio/monthly_vifs/vif_gear_aug.csv", header = TRUE)
+vif_gear_aug$GVIF2Dfsq <- vif_gear_aug$GVIF..1..2.Df..^2
+write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
+vif_gear_aug
+
+#without gear
+vif_nogear_aug <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = augpresab))
+write.csv(vif_nogear_aug, "../output/bio/monthly_vifs/vif_nogear_aug.csv", row.names = TRUE)
+vif_nogear_aug
+
+#with gear
+vif_gear_sep <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = seppresab))
+write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
+vif_gear_sep
+
+vif_gear_sep <- read.csv("../output/bio/monthly_vifs/vif_gear_sep.csv", header = TRUE)
+vif_gear_sep$GVIF2Dfsq <- vif_gear_sep$GVIF..1..2.Df..^2
+write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
+vif_gear_sep
+
+#without gear
+vif_nogear_sep <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = seppresab))
+write.csv(vif_nogear_sep, "../output/bio/monthly_vifs/vif_nogear_sep.csv", row.names = TRUE)
+vif_nogear_sep
+
+#with gear
+vif_gear_oct <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
+write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
+vif_gear_oct
+
+vif_gear_oct <- read.csv("../output/bio/monthly_vifs/vif_gear_oct.csv", header = TRUE)
+vif_gear_oct$GVIF2Dfsq <- vif_gear_oct$GVIF..1..2.Df..^2
+write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
+vif_gear_oct
+
+#without gear
+vif_nogear_oct <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
+write.csv(vif_nogear_oct, "../output/bio/monthly_vifs/vif_nogear_oct.csv", row.names = TRUE)
+vif_nogear_oct
+
+#with gear
+vif_gear_nov <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = novpresab))
+write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
+vif_gear_nov
+
+vif_gear_nov <- read.csv("../output/bio/monthly_vifs/vif_gear_nov.csv", header = TRUE)
+vif_gear_nov$GVIF2Dfsq <- vif_gear_nov$GVIF..1..2.Df..^2
+write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
+vif_gear_nov
+
+#without gear
+vif_nogear_nov <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = novpresab))
+write.csv(vif_nogear_nov, "../output/bio/monthly_vifs/vif_nogear_nov.csv", row.names = TRUE)
+vif_nogear_nov
+
+#with gear
+vif_gear_dec <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = decpresab))
+write.csv(vif_gear_dec, "../output/bio/monthly_vifs/vif_gear_dec.csv", row.names = TRUE)
+vif_gear_dec
+
+#without gear
+vif_nogear_dec <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = decpresab))
+write.csv(vif_nogear_dec, "../output/bio/monthly_vifs/vif_nogear_dec.csv", row.names = TRUE)
+vif_nogear_dec
 ```

### 2019-01-30 src/data_exploration.Rmd
arse

@@
 
+spearmans indicates mlp_surface, ssh_surface, salinity_surface, temp_surface, 02_surface all highly correlated every month. remove from each model and rerun
+Run a VIF for each month (with and without gear)
+
+```{r}
+
+#with gear - for jan there is none because there is only one category of gear
+
+#without gear
+vif_nogear_jan <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth, data = janpresab))
+write.csv(vif_nogear_jan, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_jan.csv", row.names = TRUE)
+vif_nogear_jan
+
+#with gear
+vif_gear_feb <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth, data = febpresab))
+write.csv(vif_gear_feb, "../output/bio/monthly_vifs/nohighcorr_vif_gear_feb.csv", row.names = TRUE)
+vif_gear_feb
+
+vif_gear_feb <- read.csv("../output/bio/monthly_vifs/highcorrvif_gear_feb.csv", header = TRUE)
+vif_gear_feb$GVIF2Dfsq <- vif_gear_feb$GVIF..1..2.Df..^2
+write.csv(vif_gear_feb, "../output/bio/monthly_vifs/nohighcorr_vif_gear_feb.csv", row.names = TRUE)
+vif_gear_feb
+
+#without gear
+vif_nogear_feb <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth, data = febpresab))
+write.csv(vif_nogear_feb, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_feb.csv", row.names = TRUE)
+vif_nogear_feb
+
+#with gear
+vif_gear_mar <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = marpresab))
+write.csv(vif_gear_mar, "../output/bio/monthly_vifs/nohighcorr_vif_gear_mar.csv", row.names = TRUE)
+vif_gear_mar
+
+vif_gear_mar <- read.csv("../output/bio/monthly_vifs/nohighcorr_vif_gear_mar.csv", header = TRUE)
+vif_gear_mar$GVIF2Dfsq <- vif_gear_mar$GVIF..1..2.Df..^2
+write.csv(vif_gear_mar, "../output/bio/monthly_vifs/nohighcorr_vif_gear_mar.csv", row.names = TRUE)
+vif_gear_mar
+
+#without gear
+vif_nogear_mar <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = marpresab))
+write.csv(vif_nogear_mar, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_mar.csv", row.names = TRUE)
+vif_nogear_mar
+
+#with gear
+vif_gear_apr <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
+write.csv(vif_gear_apr, "../output/bio/monthly_vifs/nohighcorr_vif_gear_apr.csv", row.names = TRUE)
+vif_gear_apr
+
+vif_gear_apr <- read.csv("../output/bio/monthly_vifs/vif_gear_apr.csv", header = TRUE)
+vif_gear_apr$GVIF2Dfsq <- vif_gear_apr$GVIF..1..2.Df..^2
+write.csv(vif_gear_apr, "../output/bio/monthly_vifs/nohighcorr_vif_gear_apr.csv", row.names = TRUE)
+vif_gear_apr
+
+#without gear
+vif_nogear_apr <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
+write.csv(vif_nogear_apr, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_apr.csv", row.names = TRUE)
+vif_nogear_apr
+
+#with gear
+vif_gear_may <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = maypresab))
+write.csv(vif_gear_may, "../output/bio/monthly_vifs/nohighcorr_vif_gear_may.csv", row.names = TRUE)
+vif_gear_may
+
+#without gear
+vif_nogear_may <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = maypresab))
+write.csv(vif_nogear_may, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_may.csv", row.names = TRUE)
+vif_nogear_may
+
+#with gear
+vif_gear_jun <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+write.csv(vif_gear_jun, "../output/bio/monthly_vifs/nohighcorr_vif_gear_jun.csv", row.names = TRUE)
+vif_gear_jun
+
+#without gear
+vif_nogear_jun <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+write.csv(vif_nogear_jun, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_jun.csv", row.names = TRUE)
+vif_nogear_jun
+
+#with gear
+vif_gear_jul <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = julpresab))
+write.csv(vif_gear_jul, "../output/bio/monthly_vifs/nohighcorr_vif_gear_jul.csv", row.names = TRUE)
+vif_gear_jul
+
+vif_gear_jul <- read.csv("../output/bio/monthly_vifs/nohighcorr_vif_gear_jul.csv", header = TRUE)
+vif_gear_jul$GVIF2Dfsq <- vif_gear_jul$GVIF..1..2.Df..^2
+write.csv(vif_gear_jul, "../output/bio/monthly_vifs/nohighcorr_vif_gear_jul.csv", row.names = TRUE)
+vif_gear_jul
+
+#without gear
+vif_nogear_jul <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = julpresab))
+write.csv(vif_nogear_jul, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_jul.csv", row.names = TRUE)
+vif_nogear_jul
+
+#with gear
+vif_gear_aug <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = augpresab))
+write.csv(vif_gear_aug, "../output/bio/monthly_vifs/nohighcorr_vif_gear_aug.csv", row.names = TRUE)
+vif_gear_aug
+
+vif_gear_aug <- read.csv("../output/bio/monthly_vifs/nohighcorr_vif_gear_aug.csv", header = TRUE)
+vif_gear_aug$GVIF2Dfsq <- vif_gear_aug$GVIF..1..2.Df..^2
+write.csv(vif_gear_aug, "../output/bio/monthly_vifs/nohighcorr_vif_gear_aug.csv", row.names = TRUE)
+vif_gear_aug
+
+#without gear
+vif_nogear_aug <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = augpresab))
+write.csv(vif_nogear_aug, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_aug.csv", row.names = TRUE)
+vif_nogear_aug
+
+#with gear
+vif_gear_sep <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = seppresab))
+write.csv(vif_gear_sep, "../output/bio/monthly_vifs/nohighcorr_vif_gear_sep.csv", row.names = TRUE)
+vif_gear_sep
+
+vif_gear_sep <- read.csv("../output/bio/monthly_vifs/nohighcorr_vif_gear_sep.csv", header = TRUE)
+vif_gear_sep$GVIF2Dfsq <- vif_gear_sep$GVIF..1..2.Df..^2
+write.csv(vif_gear_sep, "../output/bio/monthly_vifs/nohighcorr_vif_gear_sep.csv", row.names = TRUE)
+vif_gear_sep
+
+#without gear
+vif_nogear_sep <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = seppresab))
+write.csv(vif_nogear_sep, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_sep.csv", row.names = TRUE)
+vif_nogear_sep
+
+#with gear
+vif_gear_oct <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
+write.csv(vif_gear_oct, "../output/bio/monthly_vifs/nohighcorr_vif_gear_oct.csv", row.names = TRUE)
+vif_gear_oct
+
+vif_gear_oct <- read.csv("../output/bio/monthly_vifs/nohighcorr_vif_gear_oct.csv", header = TRUE)
+vif_gear_oct$GVIF2Dfsq <- vif_gear_oct$GVIF..1..2.Df..^2
+write.csv(vif_gear_oct, "../output/bio/monthly_vifs/nohighcorr_vif_gear_oct.csv", row.names = TRUE)
+vif_gear_oct
+
+#without gear
+vif_nogear_oct <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
+write.csv(vif_nogear_oct, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_oct.csv", row.names = TRUE)
+vif_nogear_oct
+
+#with gear
+vif_gear_nov <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = novpresab))
+write.csv(vif_gear_nov, "../output/bio/monthly_vifs/nohighcorr_vif_gear_nov.csv", row.names = TRUE)
+vif_gear_nov
+
+vif_gear_nov <- read.csv("../output/bio/monthly_vifs/nohighcorr_vif_gear_nov.csv", header = TRUE)
+vif_gear_nov$GVIF2Dfsq <- vif_gear_nov$GVIF..1..2.Df..^2
+write.csv(vif_gear_nov, "../output/bio/monthly_vifs/nohighcorr_vif_gear_nov.csv", row.names = TRUE)
+vif_gear_nov
+
+#without gear
+vif_nogear_nov <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = novpresab))
+write.csv(vif_nogear_nov, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_nov.csv", row.names = TRUE)
+vif_nogear_nov
+
+#with gear
+vif_gear_dec <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = decpresab))
+write.csv(vif_gear_dec, "../output/bio/monthly_vifs/nohighcorr_vif_gear_dec.csv", row.names = TRUE)
+vif_gear_dec
+
+#without gear
+vif_nogear_dec <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = decpresab))
+write.csv(vif_nogear_dec, "../output/bio/monthly_vifs/nohighcorr_vif_nogear_dec.csv", row.names = TRUE)
+vif_nogear_dec
+```
+
 #maxent?

### 2019-01-31 src/data_exploration.Rmd
subset backgrond data to get 10k in total per month

--- 2019-01-30 src/data_exploration.Rmd
@@
 year
-```{r}
+```{r observations by institute-year}
 inst_by_yr <- with(presence, table(year, institutioncode))
@@
 month
-```{r}
+```{r observations by institute-month}
 inst_by_mt <- with(presence, table(month, institutioncode))
@@
 inst_by_mt
+```
+
+and just a year-month table
+```{r observations by year-month}
+year_by_mt <- with(presence, table(year, month))
+write.csv(year_by_mt, file = "../output/bio/year_by_mth.csv")
+year_by_mt
 ```
@@
 
-```{r}
-
+```{r monthly vif with and without gear}
+
+#with gear - for jan there is none because there is only one category of gear
+
+#without gear
+vif_nogear_jan <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = janpresab))
+write.csv(vif_nogear_jan, "../output/bio/monthly_vifs/vif_nogear_jan.csv", row.names = TRUE)
+vif_nogear_jan
+
+#without gear
+vif_nogear_feb <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = febpresab))
+write.csv(vif_nogear_feb, "../output/bio/monthly_vifs/vif_nogear_feb.csv", row.names = TRUE)
+vif_nogear_feb
+
+#with gear
+vif_gear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
+write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
+vif_gear_mar
+
+vif_gear_mar <- read.csv("../output/bio/monthly_vifs/vif_gear_mar.csv", header = TRUE)
+vif_gear_mar$GVIF2Dfsq <- vif_gear_mar$GVIF..1..2.Df..^2
+write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
+vif_gear_mar
+
+#without gear
+vif_nogear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
+write.csv(vif_nogear_mar, "../output/bio/monthly_vifs/vif_nogear_mar.csv", row.names = TRUE)
+vif_nogear_mar
+
+#with gear
+vif_gear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
+write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
+vif_gear_apr
+
+vif_gear_apr <- read.csv("../output/bio/monthly_vifs/vif_gear_apr.csv", header = TRUE)
+vif_gear_apr$GVIF2Dfsq <- vif_gear_apr$GVIF..1..2.Df..^2
+write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
+vif_gear_apr
+
+#without gear
+vif_nogear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
+write.csv(vif_nogear_apr, "../output/bio/monthly_vifs/vif_nogear_apr.csv", row.names = TRUE)
+vif_nogear_apr
+
+#with gear
+vif_gear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
+write.csv(vif_gear_may, "../output/bio/monthly_vifs/vif_gear_may.csv", row.names = TRUE)
+vif_gear_may
+
+#without gear
+vif_nogear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
+write.csv(vif_nogear_may, "../output/bio/monthly_vifs/vif_nogear_may.csv", row.names = TRUE)
+vif_nogear_may
+
+#with gear
+vif_gear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
+write.csv(vif_gear_jun, "../output/bio/monthly_vifs/vif_gear_jun.csv", row.names = TRUE)
+vif_gear_jun
+
+#without gear
+vif_nogear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
+write.csv(vif_nogear_jun, "../output/bio/monthly_vifs/vif_nogear_jun.csv", row.names = TRUE)
+vif_nogear_jun
+
+#with gear
+vif_gear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
+write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
+vif_gear_jul
+
+vif_gear_jul <- read.csv("../output/bio/monthly_vifs/vif_gear_jul.csv", header = TRUE)
+vif_gear_jul$GVIF2Dfsq <- vif_gear_jul$GVIF..1..2.Df..^2
+write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
+vif_gear_jul
+
+#without gear
+vif_nogear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
+write.csv(vif_nogear_jul, "../output/bio/monthly_vifs/vif_nogear_jul.csv", row.names = TRUE)
+vif_nogear_jul
+
+#with gear
+vif_gear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
+write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
+vif_gear_aug
+
+vif_gear_aug <- read.csv("../output/bio/monthly_vifs/vif_gear_aug.csv", header = TRUE)
+vif_gear_aug$GVIF2Dfsq <- vif_gear_aug$GVIF..1..2.Df..^2
+write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
+vif_gear_aug
+
+#without gear
+vif_nogear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
+write.csv(vif_nogear_aug, "../output/bio/monthly_vifs/vif_nogear_aug.csv", row.names = TRUE)
+vif_nogear_aug
+
+#with gear
+vif_gear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
+write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
+vif_gear_sep
+
+vif_gear_sep <- read.csv("../output/bio/monthly_vifs/vif_gear_sep.csv", header = TRUE)
+vif_gear_sep$GVIF2Dfsq <- vif_gear_sep$GVIF..1..2.Df..^2
+write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
+vif_gear_sep
+
+#without gear
+vif_nogear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
+write.csv(vif_nogear_sep, "../output/bio/monthly_vifs/vif_nogear_sep.csv", row.names = TRUE)
+vif_nogear_sep
+
+#with gear
+vif_gear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
+write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
+vif_gear_oct
+
+vif_gear_oct <- read.csv("../output/bio/monthly_vifs/vif_gear_oct.csv", header = TRUE)
+vif_gear_oct$GVIF2Dfsq <- vif_gear_oct$GVIF..1..2.Df..^2
+write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
+vif_gear_oct
+
+#without gear
+vif_nogear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
+write.csv(vif_nogear_oct, "../output/bio/monthly_vifs/vif_nogear_oct.csv", row.names = TRUE)
+vif_nogear_oct
+
+#with gear
+vif_gear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
+write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
+vif_gear_nov
+
+vif_gear_nov <- read.csv("../output/bio/monthly_vifs/vif_gear_nov.csv", header = TRUE)
+vif_gear_nov$GVIF2Dfsq <- vif_gear_nov$GVIF..1..2.Df..^2
+write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
+vif_gear_nov
+
+#without gear
+vif_nogear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
+write.csv(vif_nogear_nov, "../output/bio/monthly_vifs/vif_nogear_nov.csv", row.names = TRUE)
+vif_nogear_nov
+
+#with gear
+vif_gear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
+write.csv(vif_gear_dec, "../output/bio/monthly_vifs/vif_gear_dec.csv", row.names = TRUE)
+vif_gear_dec
+
+#without gear
+vif_nogear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
+write.csv(vif_nogear_dec, "../output/bio/monthly_vifs/vif_nogear_dec.csv", row.names = TRUE)
+vif_nogear_dec
+```
+
+spearmans indicates mlp_surface, ssh_surface, salinity_surface, temp_surface, 02_surface all highly correlated every month. remove from each model and rerun
+Run a VIF for each month (with and without gear)
+
+```{r monthly vif with and without gear and highly correlated}
+
+#with gear - for jan there is none because there is only one category of gear
+
+#without gear
+vif_nogear_jan <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = janpresab))
+write.csv(vif_nogear_jan, "../output/bio/monthly_vifs/vif_nogear_jan.csv", row.names = TRUE)
+vif_nogear_jan
+
+#without gear
+vif_nogear_feb <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = febpresab))
+write.csv(vif_nogear_feb, "../output/bio/monthly_vifs/vif_nogear_feb.csv", row.names = TRUE)
+vif_nogear_feb
+
+#with gear
+vif_gear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
+write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
+vif_gear_mar
+
+vif_gear_mar <- read.csv("../output/bio/monthly_vifs/vif_gear_mar.csv", header = TRUE)
+vif_gear_mar$GVIF2Dfsq <- vif_gear_mar$GVIF..1..2.Df..^2
+write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
+vif_gear_mar
+
+#without gear
+vif_nogear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
+write.csv(vif_nogear_mar, "../output/bio/monthly_vifs/vif_nogear_mar.csv", row.names = TRUE)
+vif_nogear_mar
+
+#with gear
+vif_gear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
+write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
+vif_gear_apr
+
+vif_gear_apr <- read.csv("../output/bio/monthly_vifs/vif_gear_apr.csv", header = TRUE)
+vif_gear_apr$GVIF2Dfsq <- vif_gear_apr$GVIF..1..2.Df..^2
+write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
+vif_gear_apr
+
+#without gear
+vif_nogear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
+write.csv(vif_nogear_apr, "../output/bio/monthly_vifs/vif_nogear_apr.csv", row.names = TRUE)
+vif_nogear_apr
+
+#with gear
+vif_gear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
+write.csv(vif_gear_may, "../output/bio/monthly_vifs/vif_gear_may.csv", row.names = TRUE)
+vif_gear_may
+
+#without gear
+vif_nogear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
+write.csv(vif_nogear_may, "../output/bio/monthly_vifs/vif_nogear_may.csv", row.names = TRUE)
+vif_nogear_may
+
+#with gear
+vif_gear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
+write.csv(vif_gear_jun, "../output/bio/monthly_vifs/vif_gear_jun.csv", row.names = TRUE)
+vif_gear_jun
+
+#without gear
+vif_nogear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
+write.csv(vif_nogear_jun, "../output/bio/monthly_vifs/vif_nogear_jun.csv", row.names = TRUE)
+vif_nogear_jun
+
+#with gear
+vif_gear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
+write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
+vif_gear_jul
+
+vif_gear_jul <- read.csv("../output/bio/monthly_vifs/vif_gear_jul.csv", header = TRUE)
+vif_gear_jul$GVIF2Dfsq <- vif_gear_jul$GVIF..1..2.Df..^2
+write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
+vif_gear_jul
+
+#without gear
+vif_nogear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
+write.csv(vif_nogear_jul, "../output/bio/monthly_vifs/vif_nogear_jul.csv", row.names = TRUE)
+vif_nogear_jul
+
+#with gear
+vif_gear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
+write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
+vif_gear_aug
+
+vif_gear_aug <- read.csv("../output/bio/monthly_vifs/vif_gear_aug.csv", header = TRUE)
+vif_gear_aug$GVIF2Dfsq <- vif_gear_aug$GVIF..1..2.Df..^2
+write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
+vif_gear_aug
+
+#without gear
+vif_nogear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
+write.csv(vif_nogear_aug, "../output/bio/monthly_vifs/vif_nogear_aug.csv", row.names = TRUE)
+vif_nogear_aug
+
+#with gear
+vif_gear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
+write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
+vif_gear_sep
+
+vif_gear_sep <- read.csv("../output/bio/monthly_vifs/vif_gear_sep.csv", header = TRUE)
+vif_gear_sep$GVIF2Dfsq <- vif_gear_sep$GVIF..1..2.Df..^2
+write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
+vif_gear_sep
+
+#without gear
+vif_nogear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
+write.csv(vif_nogear_sep, "../output/bio/monthly_vifs/vif_nogear_sep.csv", row.names = TRUE)
+vif_nogear_sep
+
+#with gear
+vif_gear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
+write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
+vif_gear_oct
+
+vif_gear_oct <- read.csv("../output/bio/monthly_vifs/vif_gear_oct.csv", header = TRUE)
+vif_gear_oct$GVIF2Dfsq <- vif_gear_oct$GVIF..1..2.Df..^2
+write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
+vif_gear_oct
+
+#without gear
+vif_nogear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
+write.csv(vif_nogear_oct, "../output/bio/monthly_vifs/vif_nogear_oct.csv", row.names = TRUE)
+vif_nogear_oct
+
+#with gear
+vif_gear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
+write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
+vif_gear_nov
+
+vif_gear_nov <- read.csv("../output/bio/monthly_vifs/vif_gear_nov.csv", header = TRUE)
+vif_gear_nov$GVIF2Dfsq <- vif_gear_nov$GVIF..1..2.Df..^2
+write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
+vif_gear_nov
+
+#without gear
+vif_nogear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
+write.csv(vif_nogear_nov, "../output/bio/monthly_vifs/vif_nogear_nov.csv", row.names = TRUE)
+vif_nogear_nov
+
+#with gear
+vif_gear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
+write.csv(vif_gear_dec, "../output/bio/monthly_vifs/vif_gear_dec.csv", row.names = TRUE)
+vif_gear_dec
+
+#without gear
+vif_nogear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
+write.csv(vif_nogear_dec, "../output/bio/monthly_vifs/vif_nogear_dec.csv", row.names = TRUE)
+vif_nogear_dec
 ```

### 2019-01-31 src/data_exploration.Rmd
rerun correlations etc on background10k pm

@@
 ```{r}
-presab <- read.csv("../output/bio/presab.csv", header = TRUE)
+presab <- read.csv("../output/bio/presab_10k.csv", header = TRUE)
 presence <- subset(presab, occurrence == "1")
@@
 
-add nafo region and gear type into the mix
+add nafo region and gear type into the mix CANT!!
 
@@
 presab <- presab[!(presab$nafo_zone == "1C" & presab$occurrence == "1"), ]
-write.csv(presab, "../output/bio/presab.csv", row.names = FALSE)
+write.csv(presab, "../output/bio/presab_10k.csv", row.names = FALSE)
 presence <- presence[!(presence$nafo_zone == "1C"), ]
@@
 ```
+XXXSAM FROM HERE - NEED TO ADD PLOTS TO RESULTS DOC THEN CONTINUE
 

### 2019-02-04 src/data_exploration.Rmd
monthly spearmans rerun with selected data + minor alteration to the folders output sent to

--- 2019-01-31 src/data_exploration.Rmd
@@
 
-looks like not much difference... what about a kruskal wallace test?
+what about a kruskal wallace test?
 
@@
 vif_allpresab <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allpresab, "../output/bio/vif_allpresab.csv", row.names = TRUE)
+write.csv(vif_allpresab, "../output/bio/vif/vif_allpresab.csv", row.names = TRUE)
 vif_allpresab
@@
 ```{r}
-vif_allpresab_sq <- read.csv("../output/bio/vif_allpresab.csv", header = TRUE)
+vif_allpresab_sq <- read.csv("../output/bio/vif/vif_allpresab.csv", header = TRUE)
 vif_allpresab_sq$GVIF2Dfsq <- vif_allpresab_sq$GVIF..1..2.Df..^2
-write.csv(vif_allpresab_sq, "../output/bio/vif_allpresab_sq.csv", row.names = TRUE)
+write.csv(vif_allpresab_sq, "../output/bio/vif/vif_allpresab.csv", row.names = TRUE)
 vif_allpresab_sq
@@
 vif_allbutgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbutgear, "../output/bio/vif_allbutgear.csv", row.names = TRUE)
+write.csv(vif_allbutgear, "../output/bio/vif/vif_allbutgear.csv", row.names = TRUE)
 vif_allbutgear
 ```
-
+xx
 As per SLR suggestion, rerun without gear and most highly correlated variables
@@
 ```{r vif without gear + high corr}
-vif_allbutgearhighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
-write.csv(vif_allbutgearhighlycorr, "../output/bio/vif_allbutgearhighlycorr.csv", row.names = TRUE)
+vif_allbutgearhighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbutgearhighlycorr, "../output/bio/vif/vif_allbutgearhighlycorr.csv", row.names = TRUE)
 vif_allbutgearhighlycorr
@@
 ```{r vif all but high corr}
-vif_allbuthighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter + gear, data = presab))
-write.csv(vif_allbuthighlycorr, "../output/bio/vif_allbuthighlycorr.csv", row.names = TRUE)
+vif_allbuthighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter + gear, data = presab))
+write.csv(vif_allbuthighlycorr, "../output/bio/vif/vif_allbuthighlycorr.csv", row.names = TRUE)
 vif_allbuthighlycorr
 
-vif_allbuthighlycorr_sq <- read.csv("../output/bio/vif_allbuthighlycorr.csv", header = TRUE)
+vif_allbuthighlycorr_sq <- read.csv("../output/bio/vif/vif_allbuthighlycorr.csv", header = TRUE)
 vif_allbuthighlycorr_sq$GVIF2Dfsq <- vif_allbuthighlycorr_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbuthighlycorr_sq, "../output/bio/vif_allbuthighlycorr_sq.csv", row.names = TRUE)
+write.csv(vif_allbuthighlycorr_sq, "../output/bio/vif/vif_allbuthighlycorr.csv", row.names = TRUE)
 vif_allbuthighlycorr_sq
@@
 ok remove one variable at a time - leave gear in
+
+remove bottom_depth
+```{r vif all but botdepth}
+vif_allbutbot_prev <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutbot_prev, "../output/bio/vif/vif_allbutbot_prev.csv", row.names = TRUE)
+
+vif_allbutbot_prev <- read.csv("../output/bio/vif/vif_allbutbot_prev.csv", header = TRUE)
+vif_allbutbot_prev$GVIF2Dfsq <- vif_allbutbot_prev$GVIF..1..2.Df..^2
+write.csv(vif_allbutbot_prev, "../output/bio/vif/vif_allbutbot_prev.csv", row.names = TRUE)
+vif_allbutbot_prev
+```
+
+and leave gear out
+
+remove bottom_depth + gear
+```{r vif all but botddepth + gear}
+vif_allbutbot_prevgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbutbot_prevgear, "../output/bio/vif/vif_allbutbot_prevgear.csv", row.names = TRUE)
+vif_allbutbot_prevgear
+```
 
@@
 vif_allbutamo_prev <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
-write.csv(vif_allbutamo_prev, "../output/bio/vif_allbutamo_prev.csv", row.names = TRUE)
+write.csv(vif_allbutamo_prev, "../output/bio/vif/vif_allbutamo_prev.csv", row.names = TRUE)
+
+vif_allbutamo_prev <- read.csv("../output/bio/vif/vif_allbutamo_prev.csv", header = TRUE)
+vif_allbutamo_prev$GVIF2Dfsq <- vif_allbutamo_prev$GVIF..1..2.Df..^2
+write.csv(vif_allbutamo_prev, "../output/bio/vif/vif_allbutamo_prev.csv", row.names = TRUE)
 vif_allbutamo_prev
-
-vif_allbutamo_prev_sq <- read.csv("../output/bio/vif_allbutamo_prev.csv", header = TRUE)
-vif_allbutamo_prev_sq$GVIF2Dfsq <- vif_allbutamo_prev_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbutamo_prev_sq, "../output/bio/vif_allbutamo_prev_sq.csv", row.names = TRUE)
-vif_allbutamo_prev_sq
 ```
@@
 vif_allbutamo_prevgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
-write.csv(vif_allbutamo_prevgear, "../output/bio/vif_allbutamo_prevgear.csv", row.names = TRUE)
+write.csv(vif_allbutamo_prevgear, "../output/bio/vif/vif_allbutamo_prevgear.csv", row.names = TRUE)
 vif_allbutamo_prevgear
@@
 vif_allbutchl_depth <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbutchl_depth, "../output/bio/vif_allbutchl_depth.csv", row.names = TRUE)
+write.csv(vif_allbutchl_depth, "../output/bio/vif/vif_allbutchl_depth.csv", row.names = TRUE)
+
+vif_allbutchl_depth <- read.csv("../output/bio/vif/vif_allbutchl_depth.csv", header = TRUE)
+vif_allbutchl_depth$GVIF2Dfsq <- vif_allbutchl_depth$GVIF..1..2.Df..^2
+write.csv(vif_allbutchl_depth, "../output/bio/vif/vif_allbutchl_depth.csv", row.names = TRUE)
 vif_allbutchl_depth
-
-vif_allbutchl_depth_sq <- read.csv("../output/bio/vif_allbutchl_depth.csv", header = TRUE)
-vif_allbutchl_depth_sq$GVIF2Dfsq <- vif_allbutchl_depth_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbutchl_depth_sq, "../output/bio/vif_allbutchl_depth_sq.csv", row.names = TRUE)
-vif_allbutchl_depth_sq
 ```
@@
 vif_allbutchl_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbutchl_depthgear, "../output/bio/vif_allbutchl_depthgear.csv", row.names = TRUE)
+write.csv(vif_allbutchl_depthgear, "../output/bio/vif/vif_allbutchl_depthgear.csv", row.names = TRUE)
 vif_allbutchl_depthgear
@@
 vif_allbutssh_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbutssh_surface, "../output/bio/vif_allbutssh_surface.csv", row.names = TRUE)
+write.csv(vif_allbutssh_surface, "../output/bio/vif/vif_allbutssh_surface.csv", row.names = TRUE)
+
+vif_allbutssh_surface <- read.csv("../output/bio/vif/vif_allbutssh_surface.csv", header = TRUE)
+vif_allbutssh_surface$GVIF2Dfsq <- vif_allbutssh_surface$GVIF..1..2.Df..^2
+write.csv(vif_allbutssh_surface, "../output/bio/vif/vif_allbutssh_surface.csv", row.names = TRUE)
 vif_allbutssh_surface
-
-vif_allbutssh_surface_sq <- read.csv("../output/bio/vif_allbutssh_surface.csv", header = TRUE)
-vif_allbutssh_surface_sq$GVIF2Dfsq <- vif_allbutssh_surface_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbutssh_surface_sq, "../output/bio/vif_allbutssh_surface_sq.csv", row.names = TRUE)
-vif_allbutssh_surface_sq
 ```
@@
 vif_allbutssh_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbutssh_surfacegear, "../output/bio/vif_allbutssh_surfacegear.csv", row.names = TRUE)
+write.csv(vif_allbutssh_surfacegear, "../output/bio/vif/vif_allbutssh_surfacegear.csv", row.names = TRUE)
 vif_allbutssh_surfacegear
@@
 vif_allbuto2_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbuto2_surface, "../output/bio/vif_allbuto2_surface.csv", row.names = TRUE)
+write.csv(vif_allbuto2_surface, "../output/bio/vif/vif_allbuto2_surface.csv", row.names = TRUE)
+
+vif_allbuto2_surface <- read.csv("../output/bio/vif/vif_allbuto2_surface.csv", header = TRUE)
+vif_allbuto2_surface$GVIF2Dfsq <- vif_allbuto2_surface$GVIF..1..2.Df..^2
+write.csv(vif_allbuto2_surface, "../output/bio/vif/vif_allbuto2_surface.csv", row.names = TRUE)
 vif_allbuto2_surface
-
-vif_allbuto2_surface_sq <- read.csv("../output/bio/vif_allbuto2_surface.csv", header = TRUE)
-vif_allbuto2_surface_sq$GVIF2Dfsq <- vif_allbuto2_surface_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbuto2_surface_sq, "../output/bio/vif_allbuto2_surface_sq.csv", row.names = TRUE)
-vif_allbuto2_surface_sq
 ```
@@
 vif_allbuto2_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbuto2_surfacegear, "../output/bio/vif_allbuto2_surfacegear.csv", row.names = TRUE)
+write.csv(vif_allbuto2_surfacegear, "../output/bio/vif/vif_allbuto2_surfacegear.csv", row.names = TRUE)
 vif_allbuto2_surfacegear
@@
 vif_allbutsalinitysurface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbutsalinitysurface, "../output/bio/vif_allbutsalinitysurface.csv", row.names = TRUE)
+write.csv(vif_allbutsalinitysurface, "../output/bio/vif/vif_allbutsalinitysurface.csv", row.names = TRUE)
+
+vif_allbutsalinitysurface <- read.csv("../output/bio/vif/vif_allbutsalinitysurface.csv", header = TRUE)
+vif_allbutsalinitysurface$GVIF2Dfsq <- vif_allbutsalinitysurface$GVIF..1..2.Df..^2
+write.csv(vif_allbutsalinitysurface, "../output/bio/vif/vif_allbutsalinitysurface.csv", row.names = TRUE)
 vif_allbutsalinitysurface
-
-vif_allbutsalinitysurface_sq <- read.csv("../output/bio/vif_allbutsalinitysurface.csv", header = TRUE)
-vif_allbutsalinitysurface_sq$GVIF2Dfsq <- vif_allbutsalinitysurface_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbutsalinitysurface_sq, "../output/bio/vif_allbutsalinitysurface_sq.csv", row.names = TRUE)
-vif_allbutsalinitysurface_sq
 ```
@@
 vif_allbutsalinitysurfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbutsalinitysurfacegear, "../output/bio/vif_allbutsalinitysurfacegear.csv", row.names = TRUE)
+write.csv(vif_allbutsalinitysurfacegear, "../output/bio/vif/vif_allbutsalinitysurfacegear.csv", row.names = TRUE)
 vif_allbutsalinitysurfacegear
+```
+
+remove temp_surface only
+```{r vif all but tempsurface}
+vif_allbuttempsurface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuttempsurface, "../output/bio/vif/vif_allbuttempsurface.csv", row.names = TRUE)
+
+vif_allbuttempsurface <- read.csv("../output/bio/vif/vif_allbuttempsurface.csv", header = TRUE)
+vif_allbuttempsurface$GVIF2Dfsq <- vif_allbuttempsurface$GVIF..1..2.Df..^2
+write.csv(vif_allbuttempsurface, "../output/bio/vif/vif_allbuttempsurface.csv", row.names = TRUE)
+vif_allbuttempsurface
+
+vif_allbuttempsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
+write.csv(vif_allbuttempsurfacegear, "../output/bio/vif/vif_allbuttempsurfacegear.csv", row.names = TRUE)
+vif_allbuttempsurfacegear
+
 ```
@@
 #with gear
-vif_allbuthighcorrtempsurface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
-write.csv(vif_allbuthighcorrtempsurface, "../output/bio/vif_allbuthighcorrtempsurface.csv", row.names = TRUE)
+vif_allbuthighcorrtempsurface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbuthighcorrtempsurface, "../output/bio/vif/vif_allbuthighcorrtempsurface.csv", row.names = TRUE)
+
+vif_allbuthighcorrtempsurface <- read.csv("../output/bio/vif/vif_allbuthighcorrtempsurface.csv", header = TRUE)
+vif_allbuthighcorrtempsurface$GVIF2Dfsq <- vif_allbuthighcorrtempsurface$GVIF..1..2.Df..^2
+write.csv(vif_allbuthighcorrtempsurface, "../output/bio/vif/vif_allbuthighcorrtempsurface.csv", row.names = TRUE)
 vif_allbuthighcorrtempsurface
 
-vif_allbuthighcorrtempsurface_sq <- read.csv("../output/bio/vif_allbuthighcorrtempsurface.csv", header = TRUE)
-vif_allbuthighcorrtempsurface_sq$GVIF2Dfsq <- vif_allbuthighcorrtempsurface_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbuthighcorrtempsurface_sq, "../output/bio/vif_allbuthighcorrtempsurface_sq.csv", row.names = TRUE)
-vif_allbuthighcorrtempsurface_sq
-
-#without gear
-vif_allbuthighcorrtempsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
-write.csv(vif_allbuthighcorrtempsurfacegear, "../output/bio/vif_allbuthighcorrtempsurfacegear.csv", row.names = TRUE)
+#without gear
+vif_allbuthighcorrtempsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbuthighcorrtempsurfacegear, "../output/bio/vif/vif_allbuthighcorrtempsurfacegear.csv", row.names = TRUE)
 vif_allbuthighcorrtempsurfacegear
@@
 vif_allbuttempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbuttempsalsurface, "../output/bio/vif_allbuttempsalsurface.csv", row.names = TRUE)
+write.csv(vif_allbuttempsalsurface, "../output/bio/vif/vif_allbuttempsalsurface.csv", row.names = TRUE)
+
+vif_allbuttempsalsurface <- read.csv("../output/bio/vif/vif_allbuttempsalsurface.csv", header = TRUE)
+vif_allbuttempsalsurface$GVIF2Dfsq <- vif_allbuttempsalsurface$GVIF..1..2.Df..^2
+write.csv(vif_allbuttempsalsurface, "../output/bio/vif/vif_allbuttempsalsurface.csv", row.names = TRUE)
 vif_allbuttempsalsurface
 
-vif_allbuttempsalsurface_sq <- read.csv("../output/bio/vif_allbuttempsalsurface.csv", header = TRUE)
-vif_allbuttempsalsurface_sq$GVIF2Dfsq <- vif_allbuttempsalsurface_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbuttempsalsurface_sq, "../output/bio/vif_allbuttempsalsurface_sq.csv", row.names = TRUE)
-vif_allbuttempsalsurface_sq
-
 #without gear
 vif_allbuttempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
-write.csv(vif_allbuttempsalsurfacegear, "../output/bio/vif_allbuttempsalsurfacegear.csv", row.names = TRUE)
+write.csv(vif_allbuttempsalsurfacegear, "../output/bio/vif/vif_allbuttempsalsurfacegear.csv", row.names = TRUE)
 vif_allbuttempsalsurfacegear
@@
 #with gear
-vif_allbuthighcorrtempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
-write.csv(vif_allbuthighcorrtempsalsurface, "../output/bio/vif_allbuthighcorrtempsalsurface.csv", row.names = TRUE)
+vif_allbuthighcorrtempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbuthighcorrtempsalsurface, "../output/bio/vif/vif_allbuthighcorrtempsalsurface.csv", row.names = TRUE)
+
+vif_allbuthighcorrtempsalsurface <- read.csv("../output/bio/vif/vif_allbuthighcorrtempsalsurface.csv", header = TRUE)
+vif_allbuthighcorrtempsalsurface$GVIF2Dfsq <- vif_allbuthighcorrtempsalsurface$GVIF..1..2.Df..^2
+write.csv(vif_allbuthighcorrtempsalsurface, "../output/bio/vif/vif_allbuthighcorrtempsalsurface.csv", row.names = TRUE)
 vif_allbuthighcorrtempsalsurface
 
-vif_allbuthighcorrtempsalsurface_sq <- read.csv("../output/bio/vif_allbuthighcorrtempsalsurface.csv", header = TRUE)
-vif_allbuthighcorrtempsalsurface_sq$GVIF2Dfsq <- vif_allbuthighcorrtempsalsurface_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbuthighcorrtempsalsurface_sq, "../output/bio/vif_allbuthighcorrtempsalsurface_sq.csv", row.names = TRUE)
-vif_allbuthighcorrtempsalsurface_sq
-
-#without gear
-vif_allbuthighcorrtempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
-write.csv(vif_allbuthighcorrtempsalsurfacegear, "../output/bio/vif_allbuthighcorrtempsalsurfacegear.csv", row.names = TRUE)
+#without gear
+vif_allbuthighcorrtempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
+write.csv(vif_allbuthighcorrtempsalsurfacegear, "../output/bio/vif/vif_allbuthighcorrtempsalsurfacegear.csv", row.names = TRUE)
 vif_allbuthighcorrtempsalsurfacegear
@@
 vif_allbutnaoamo <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear, data = presab))
-write.csv(vif_allbutnaoamo, "../output/bio/vif_allbutnaoamo.csv", row.names = TRUE)
+write.csv(vif_allbutnaoamo, "../output/bio/vif/vif_allbutnaoamo.csv", row.names = TRUE)
+
+vif_allbutnaoamo <- read.csv("../output/bio/vif/vif_allbutnaoamo.csv", header = TRUE)
+vif_allbutnaoamo$GVIF2Dfsq <- vif_allbutnaoamo$GVIF..1..2.Df..^2
+write.csv(vif_allbutnaoamo, "../output/bio/vif/vif_allbutnaoamo.csv", row.names = TRUE)
 vif_allbutnaoamo
-
-vif_allbutnaoamo_sq <- read.csv("../output/bio/vif_allbutnaoamo.csv", header = TRUE)
-vif_allbutnaoamo_sq$GVIF2Dfsq <- vif_allbutnaoamo_sq$GVIF..1..2.Df..^2
-write.csv(vif_allbutnaoamo_sq, "../output/bio/vif_allbutnaoamo_sq.csv", row.names = TRUE)
-vif_allbutnaoamo_sq
 ```
@@
 ```{r monthly spearman correlations}
-janback <- subset(janback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+janback <- subset(janback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 janback_cor <- cor(janback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-febback <- subset(febback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+febback <- subset(febback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 febback_cor <- cor(febback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-marback <- subset(marback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+marback <- subset(marback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 marback_cor <- cor(marback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-aprback <- subset(aprback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+aprback <- subset(aprback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 aprback_cor <- cor(aprback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-mayback <- subset(mayback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+mayback <- subset(mayback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 mayback_cor <- cor(mayback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-junback <- subset(junback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+junback <- subset(junback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 junback_cor <- cor(junback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-julback <- subset(julback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+julback <- subset(julback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 julback_cor <- cor(julback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-augback <- subset(augback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+augback <- subset(augback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 augback_cor <- cor(augback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-sepback <- subset(sepback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+sepback <- subset(sepback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 sepback_cor <- cor(sepback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-octback <- subset(octback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+octback <- subset(octback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 octback_cor <- cor(octback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-novback <- subset(novback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+novback <- subset(novback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 novback_cor <- cor(novback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.
@@
 
-decback <- subset(decback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
+decback <- subset(decback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
 decback_cor <- cor(decback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot.

### 2019-02-05 src/data_exploration.Rmd
monthly vif finished. selection of variables done. cleared out observations with missing needed variables

@@
 vif_nogear_jan <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = janpresab))
-write.csv(vif_nogear_jan, "../output/bio/monthly_vifs/vif_nogear_jan.csv", row.names = TRUE)
+write.csv(vif_nogear_jan, "../output/bio/vif/monthly_vifs/vif_nogear_jan.csv", row.names = TRUE)
 vif_nogear_jan
@@
 vif_nogear_feb <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = febpresab))
-write.csv(vif_nogear_feb, "../output/bio/monthly_vifs/vif_nogear_feb.csv", row.names = TRUE)
+write.csv(vif_nogear_feb, "../output/bio/vif/monthly_vifs/vif_nogear_feb.csv", row.names = TRUE)
 vif_nogear_feb
@@
 vif_gear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
-write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
+write.csv(vif_gear_mar, "../output/bio/vif/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
 vif_gear_mar
 
-vif_gear_mar <- read.csv("../output/bio/monthly_vifs/vif_gear_mar.csv", header = TRUE)
+vif_gear_mar <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_mar.csv", header = TRUE)
 vif_gear_mar$GVIF2Dfsq <- vif_gear_mar$GVIF..1..2.Df..^2
-write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
+write.csv(vif_gear_mar, "../output/bio/vif/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
 vif_gear_mar
@@
 vif_nogear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
-write.csv(vif_nogear_mar, "../output/bio/monthly_vifs/vif_nogear_mar.csv", row.names = TRUE)
+write.csv(vif_nogear_mar, "../output/bio/vif/monthly_vifs/vif_nogear_mar.csv", row.names = TRUE)
 vif_nogear_mar
@@
 vif_gear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
-write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
+write.csv(vif_gear_apr, "../output/bio/vif/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
 vif_gear_apr
 
-vif_gear_apr <- read.csv("../output/bio/monthly_vifs/vif_gear_apr.csv", header = TRUE)
+vif_gear_apr <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_apr.csv", header = TRUE)
 vif_gear_apr$GVIF2Dfsq <- vif_gear_apr$GVIF..1..2.Df..^2
-write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
+write.csv(vif_gear_apr, "../output/bio/vif/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
 vif_gear_apr
@@
 vif_nogear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
-write.csv(vif_nogear_apr, "../output/bio/monthly_vifs/vif_nogear_apr.csv", row.names = TRUE)
+write.csv(vif_nogear_apr, "../output/bio/vif/monthly_vifs/vif_nogear_apr.csv", row.names = TRUE)
 vif_nogear_apr
@@
 vif_gear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
-write.csv(vif_gear_may, "../output/bio/monthly_vifs/vif_gear_may.csv", row.names = TRUE)
+write.csv(vif_gear_may, "../output/bio/vif/monthly_vifs/vif_gear_may.csv", row.names = TRUE)
 vif_gear_may
@@
 vif_nogear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
-write.csv(vif_nogear_may, "../output/bio/monthly_vifs/vif_nogear_may.csv", row.names = TRUE)
+write.csv(vif_nogear_may, "../output/bio/vif/monthly_vifs/vif_nogear_may.csv", row.names = TRUE)
 vif_nogear_may
@@
 vif_gear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
-write.csv(vif_gear_jun, "../output/bio/monthly_vifs/vif_gear_jun.csv", row.names = TRUE)
+write.csv(vif_gear_jun, "../output/bio/vif/monthly_vifs/vif_gear_jun.csv", row.names = TRUE)
 vif_gear_jun
@@
 vif_nogear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
-write.csv(vif_nogear_jun, "../output/bio/monthly_vifs/vif_nogear_jun.csv", row.names = TRUE)
+write.csv(vif_nogear_jun, "../output/bio/vif/monthly_vifs/vif_nogear_jun.csv", row.names = TRUE)
 vif_nogear_jun
@@
 vif_gear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
-write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
+write.csv(vif_gear_jul, "../output/bio/vif/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
 vif_gear_jul
 
-vif_gear_jul <- read.csv("../output/bio/monthly_vifs/vif_gear_jul.csv", header = TRUE)
+vif_gear_jul <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_jul.csv", header = TRUE)
 vif_gear_jul$GVIF2Dfsq <- vif_gear_jul$GVIF..1..2.Df..^2
-write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
+write.csv(vif_gear_jul, "../output/bio/vif/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
 vif_gear_jul
@@
 vif_nogear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
-write.csv(vif_nogear_jul, "../output/bio/monthly_vifs/vif_nogear_jul.csv", row.names = TRUE)
+write.csv(vif_nogear_jul, "../output/bio/vif/monthly_vifs/vif_nogear_jul.csv", row.names = TRUE)
 vif_nogear_jul
@@
 vif_gear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
-write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
+write.csv(vif_gear_aug, "../output/bio/vif/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
 vif_gear_aug
 
-vif_gear_aug <- read.csv("../output/bio/monthly_vifs/vif_gear_aug.csv", header = TRUE)
+vif_gear_aug <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_aug.csv", header = TRUE)
 vif_gear_aug$GVIF2Dfsq <- vif_gear_aug$GVIF..1..2.Df..^2
-write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
+write.csv(vif_gear_aug, "../output/bio/vif/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
 vif_gear_aug
@@
 vif_nogear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
-write.csv(vif_nogear_aug, "../output/bio/monthly_vifs/vif_nogear_aug.csv", row.names = TRUE)
+write.csv(vif_nogear_aug, "../output/bio/vif/monthly_vifs/vif_nogear_aug.csv", row.names = TRUE)
 vif_nogear_aug
@@
 vif_gear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
-write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
+write.csv(vif_gear_sep, "../output/bio/vif/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
 vif_gear_sep
 
-vif_gear_sep <- read.csv("../output/bio/monthly_vifs/vif_gear_sep.csv", header = TRUE)
+vif_gear_sep <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_sep.csv", header = TRUE)
 vif_gear_sep$GVIF2Dfsq <- vif_gear_sep$GVIF..1..2.Df..^2
-write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
+write.csv(vif_gear_sep, "../output/bio/vif/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
 vif_gear_sep
@@
 vif_nogear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
-write.csv(vif_nogear_sep, "../output/bio/monthly_vifs/vif_nogear_sep.csv", row.names = TRUE)
+write.csv(vif_nogear_sep, "../output/bio/vif/monthly_vifs/vif_nogear_sep.csv", row.names = TRUE)
 vif_nogear_sep
@@
 vif_gear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
-write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
+write.csv(vif_gear_oct, "../output/bio/vif/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
 vif_gear_oct
 
-vif_gear_oct <- read.csv("../output/bio/monthly_vifs/vif_gear_oct.csv", header = TRUE)
+vif_gear_oct <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_oct.csv", header = TRUE)
 vif_gear_oct$GVIF2Dfsq <- vif_gear_oct$GVIF..1..2.Df..^2
-write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
+write.csv(vif_gear_oct, "../output/bio/vif/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
 vif_gear_oct
@@
 vif_nogear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
-write.csv(vif_nogear_oct, "../output/bio/monthly_vifs/vif_nogear_oct.csv", row.names = TRUE)
+write.csv(vif_nogear_oct, "../output/bio/vif/monthly_vifs/vif_nogear_oct.csv", row.names = TRUE)
 vif_nogear_oct
@@
 vif_gear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
-write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
+write.csv(vif_gear_nov, "../output/bio/vif/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
 vif_gear_nov
 
-vif_gear_nov <- read.csv("../output/bio/monthly_vifs/vif_gear_nov.csv", header = TRUE)
+vif_gear_nov <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_nov.csv", header = TRUE)
 vif_gear_nov$GVIF2Dfsq <- vif_gear_nov$GVIF..1..2.Df..^2
-write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
+write.csv(vif_gear_nov, "../output/bio/vif/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
 vif_gear_nov
@@
 vif_nogear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
-write.csv(vif_nogear_nov, "../output/bio/monthly_vifs/vif_nogear_nov.csv", row.names = TRUE)
+write.csv(vif_nogear_nov, "../output/bio/vif/monthly_vifs/vif_nogear_nov.csv", row.names = TRUE)
 vif_nogear_nov
@@
 vif_gear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
-write.csv(vif_gear_dec, "../output/bio/monthly_vifs/vif_gear_dec.csv", row.names = TRUE)
+write.csv(vif_gear_dec, "../output/bio/vif/monthly_vifs/vif_gear_dec.csv", row.names = TRUE)
 vif_gear_dec
@@
 vif_nogear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
-write.csv(vif_nogear_dec, "../output/bio/monthly_vifs/vif_nogear_dec.csv", row.names = TRUE)
+write.csv(vif_nogear_dec, "../output/bio/vif/monthly_vifs/vif_nogear_dec.csv", row.names = TRUE)
 vif_nogear_dec
@@
 
-spearmans indicates mlp_surface, ssh_surface, salinity_surface, temp_surface, 02_surface all highly correlated every month. remove from each model and rerun
+spearmans indicates chl_depth, mlp_surface, ssh_surface, temp_surface, o2_surface, salinity_surface, and bottom_depth, amo_prev (and for Jun and Oct chl_surface; and for jan amo_winter, ampo_prev, amo_sample, and NAO_prev, and feb amo_winter, ampo_prev, amo_sample, nao winter, and NAO_prev) all  correlated every month.
+remove from each model and rerun
 Run a VIF for each month (with and without gear)
@@
 #without gear
-vif_nogear_jan <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = janpresab))
-write.csv(vif_nogear_jan, "../output/bio/monthly_vifs/vif_nogear_jan.csv", row.names = TRUE)
-vif_nogear_jan
-
-#without gear
-vif_nogear_feb <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = febpresab))
-write.csv(vif_nogear_feb, "../output/bio/monthly_vifs/vif_nogear_feb.csv", row.names = TRUE)
-vif_nogear_feb
-
-#with gear
-vif_gear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
-write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
-vif_gear_mar
-
-vif_gear_mar <- read.csv("../output/bio/monthly_vifs/vif_gear_mar.csv", header = TRUE)
-vif_gear_mar$GVIF2Dfsq <- vif_gear_mar$GVIF..1..2.Df..^2
-write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
-vif_gear_mar
-
-#without gear
-vif_nogear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
-write.csv(vif_nogear_mar, "../output/bio/monthly_vifs/vif_nogear_mar.csv", row.names = TRUE)
-vif_nogear_mar
-
-#with gear
-vif_gear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
-write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
-vif_gear_apr
-
-vif_gear_apr <- read.csv("../output/bio/monthly_vifs/vif_gear_apr.csv", header = TRUE)
-vif_gear_apr$GVIF2Dfsq <- vif_gear_apr$GVIF..1..2.Df..^2
-write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
-vif_gear_apr
-
-#without gear
-vif_nogear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
-write.csv(vif_nogear_apr, "../output/bio/monthly_vifs/vif_nogear_apr.csv", row.names = TRUE)
-vif_nogear_apr
-
-#with gear
-vif_gear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
-write.csv(vif_gear_may, "../output/bio/monthly_vifs/vif_gear_may.csv", row.names = TRUE)
-vif_gear_may
-
-#without gear
-vif_nogear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
-write.csv(vif_nogear_may, "../output/bio/monthly_vifs/vif_nogear_may.csv", row.names = TRUE)
-vif_nogear_may
-
-#with gear
-vif_gear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
-write.csv(vif_gear_jun, "../output/bio/monthly_vifs/vif_gear_jun.csv", row.names = TRUE)
-vif_gear_jun
-
-#without gear
-vif_nogear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
-write.csv(vif_nogear_jun, "../output/bio/monthly_vifs/vif_nogear_jun.csv", row.names = TRUE)
-vif_nogear_jun
-
-#with gear
-vif_gear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
-write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
-vif_gear_jul
-
-vif_gear_jul <- read.csv("../output/bio/monthly_vifs/vif_gear_jul.csv", header = TRUE)
-vif_gear_jul$GVIF2Dfsq <- vif_gear_jul$GVIF..1..2.Df..^2
-write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
-vif_gear_jul
-
-#without gear
-vif_nogear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
-write.csv(vif_nogear_jul, "../output/bio/monthly_vifs/vif_nogear_jul.csv", row.names = TRUE)
-vif_nogear_jul
-
-#with gear
-vif_gear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
-write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
-vif_gear_aug
-
-vif_gear_aug <- read.csv("../output/bio/monthly_vifs/vif_gear_aug.csv", header = TRUE)
-vif_gear_aug$GVIF2Dfsq <- vif_gear_aug$GVIF..1..2.Df..^2
-write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
-vif_gear_aug
-
-#without gear
-vif_nogear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
-write.csv(vif_nogear_aug, "../output/bio/monthly_vifs/vif_nogear_aug.csv", row.names = TRUE)
-vif_nogear_aug
-
-#with gear
-vif_gear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
-write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
-vif_gear_sep
-
-vif_gear_sep <- read.csv("../output/bio/monthly_vifs/vif_gear_sep.csv", header = TRUE)
-vif_gear_sep$GVIF2Dfsq <- vif_gear_sep$GVIF..1..2.Df..^2
-write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
-vif_gear_sep
-
-#without gear
-vif_nogear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
-write.csv(vif_nogear_sep, "../output/bio/monthly_vifs/vif_nogear_sep.csv", row.names = TRUE)
-vif_nogear_sep
-
-#with gear
-vif_gear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
-write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
-vif_gear_oct
-
-vif_gear_oct <- read.csv("../output/bio/monthly_vifs/vif_gear_oct.csv", header = TRUE)
-vif_gear_oct$GVIF2Dfsq <- vif_gear_oct$GVIF..1..2.Df..^2
-write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
-vif_gear_oct
-
-#without gear
-vif_nogear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
-write.csv(vif_nogear_oct, "../output/bio/monthly_vifs/vif_nogear_oct.csv", row.names = TRUE)
-vif_nogear_oct
-
-#with gear
-vif_gear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
-write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
-vif_gear_nov
-
-vif_gear_nov <- read.csv("../output/bio/monthly_vifs/vif_gear_nov.csv", header = TRUE)
-vif_gear_nov$GVIF2Dfsq <- vif_gear_nov$GVIF..1..2.Df..^2
-write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
-vif_gear_nov
-
-#without gear
-vif_nogear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
-write.csv(vif_nogear_nov, "../output/bio/monthly_vifs/vif_nogear_nov.csv", row.names = TRUE)
-vif_nogear_nov
-
-#with gear
-vif_gear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
-write.csv(vif_gear_dec, "../output/bio/monthly_vifs/vif_gear_dec.csv", row.names = TRUE)
-vif_gear_dec
-
-#without gear
-vif_nogear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
-write.csv(vif_nogear_dec, "../output/bio/monthly_vifs/vif_nogear_dec.csv", row.names = TRUE)
-vif_nogear_dec
-```
+vif_nogear_cor_jan <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface+ nao_sample + nao_winter, data = janpresab))
+write.csv(vif_nogear_cor_jan, "../output/bio/vif/monthly_vifs/vif_nogear_cor_jan.csv", row.names = TRUE)
+vif_nogear_cor_jan
+
+#without gear
+vif_nogear_cor_feb <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface  + nao_sample, data = febpresab))
+write.csv(vif_nogear_cor_feb, "../output/bio/vif/monthly_vifs/vif_nogear_cor_feb.csv", row.names = TRUE)
+vif_nogear_cor_feb
+
+#with gear
+vif_gear_cor_mar <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = marpresab))
+write.csv(vif_gear_cor_mar, "../output/bio/vif/monthly_vifs/vif_gear_cor_mar.csv", row.names = TRUE)
+
+vif_gear_cor_mar <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_cor_mar.csv", header = TRUE)
+vif_gear_cor_mar$GVIF2Dfsq <- vif_gear_cor_mar$GVIF..1..2.Df..^2
+write.csv(vif_gear_cor_mar, "../output/bio/vif/monthly_vifs/vif_gear_cor_mar.csv", row.names = TRUE)
+vif_gear_cor_mar
+
+#without gear
+vif_nogear_cor_mar <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = marpresab))
+write.csv(vif_nogear_cor_mar, "../output/bio/vif/monthly_vifs/vif_nogear_cor_mar.csv", row.names = TRUE)
+vif_nogear_cor_mar
+
+#with gear
+vif_gear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
+write.csv(vif_gear_cor_apr, "../output/bio/vif/monthly_vifs/vif_gear_cor_apr.csv", row.names = TRUE)
+
+vif_gear_cor_apr <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_cor_apr.csv", header = TRUE)
+vif_gear_cor_apr$GVIF2Dfsq <- vif_gear_cor_apr$GVIF..1..2.Df..^2
+write.csv(vif_gear_cor_apr, "../output/bio/vif/monthly_vifs/vif_gear_cor_apr.csv", row.names = TRUE)
+vif_gear_cor_apr
+
+#without gear
+vif_nogear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
+write.csv(vif_nogear_cor_apr, "../output/bio/vif/monthly_vifs/vif_nogear_cor_apr.csv", row.names = TRUE)
+vif_nogear_cor_apr
+
+#with gear
+vif_gear_cor_may <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = maypresab))
+write.csv(vif_gear_cor_may, "../output/bio/vif/monthly_vifs/vif_gear_cor_may.csv", row.names = TRUE)
+
+#without gear
+vif_nogear_cor_may <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = maypresab))
+write.csv(vif_nogear_cor_may, "../output/bio/vif/monthly_vifs/vif_nogear_cor_may.csv", row.names = TRUE)
+vif_nogear_cor_may
+
+#with gear
+vif_gear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+write.csv(vif_gear_cor_jun, "../output/bio/vif/monthly_vifs/vif_gear_cor_jun.csv", row.names = TRUE)
+vif_gear_cor_jun
+
+#without gear
+vif_nogear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+write.csv(vif_nogear_cor_jun, "../output/bio/vif/monthly_vifs/vif_nogear_cor_jun.csv", row.names = TRUE)
+vif_nogear_cor_jun
+
+#with gear
+vif_gear_cor_jul <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = julpresab))
+write.csv(vif_gear_cor_jul, "../output/bio/vif/monthly_vifs/vif_gear_cor_jul.csv", row.names = TRUE)
+
+vif_gear_cor_jul <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_cor_jul.csv", header = TRUE)
+vif_gear_cor_jul$GVIF2Dfsq <- vif_gear_cor_jul$GVIF..1..2.Df..^2
+write.csv(vif_gear_cor_jul, "../output/bio/vif/monthly_vifs/vif_gear_cor_jul.csv", row.names = TRUE)
+vif_gear_cor_jul
+
+#without gear
+vif_nogear_cor_jul <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = julpresab))
+write.csv(vif_nogear_cor_jul, "../output/bio/vif/monthly_vifs/vif_nogear_cor_jul.csv", row.names = TRUE)
+vif_nogear_cor_jul
+
+#with gear
+vif_gear_cor_aug <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = augpresab))
+write.csv(vif_gear_cor_aug, "../output/bio/vif/monthly_vifs/vif_gear_cor_aug.csv", row.names = TRUE)
+
+vif_gear_cor_aug <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_cor_aug.csv", header = TRUE)
+vif_gear_cor_aug$GVIF2Dfsq <- vif_gear_cor_aug$GVIF..1..2.Df..^2
+write.csv(vif_gear_cor_aug, "../output/bio/vif/monthly_vifs/vif_gear_cor_aug.csv", row.names = TRUE)
+vif_gear_cor_aug
+
+#without gear
+vif_nogear_cor_aug <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = augpresab))
+write.csv(vif_nogear_cor_aug, "../output/bio/vif/monthly_vifs/vif_nogear_cor_aug.csv", row.names = TRUE)
+vif_nogear_cor_aug
+
+#with gear
+vif_gear_cor_sep <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = seppresab))
+write.csv(vif_gear_cor_sep, "../output/bio/vif/monthly_vifs/vif_gear_cor_sep.csv", row.names = TRUE)
+
+vif_gear_cor_sep <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_cor_sep.csv", header = TRUE)
+vif_gear_cor_sep$GVIF2Dfsq <- vif_gear_cor_sep$GVIF..1..2.Df..^2
+write.csv(vif_gear_cor_sep, "../output/bio/vif/monthly_vifs/vif_gear_cor_sep.csv", row.names = TRUE)
+vif_gear_cor_sep
+
+#without gear
+vif_nogear_cor_sep <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = seppresab))
+write.csv(vif_nogear_cor_sep, "../output/bio/vif/monthly_vifs/vif_nogear_cor_sep.csv", row.names = TRUE)
+vif_nogear_cor_sep
+
+#with gear
+vif_gear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
+write.csv(vif_gear_cor_oct, "../output/bio/vif/monthly_vifs/vif_gear_cor_oct.csv", row.names = TRUE)
+vif_gear_cor_oct
+
+vif_gear_cor_oct <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_cor_oct.csv", header = TRUE)
+vif_gear_cor_oct$GVIF2Dfsq <- vif_gear_cor_oct$GVIF..1..2.Df..^2
+write.csv(vif_gear_cor_oct, "../output/bio/vif/monthly_vifs/vif_gear_cor_oct.csv", row.names = TRUE)
+vif_gear_cor_oct
+
+#without gear
+vif_nogear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
+write.csv(vif_nogear_cor_oct, "../output/bio/vif/monthly_vifs/vif_nogear_cor_oct.csv", row.names = TRUE)
+vif_nogear_cor_oct
+
+#with gear
+vif_gear_cor_nov <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = novpresab))
+write.csv(vif_gear_cor_nov, "../output/bio/vif/monthly_vifs/vif_gear_cor_nov.csv", row.names = TRUE)
+
+vif_gear_cor_nov <- read.csv("../output/bio/vif/monthly_vifs/vif_gear_cor_nov.csv", header = TRUE)
+vif_gear_cor_nov$GVIF2Dfsq <- vif_gear_cor_nov$GVIF..1..2.Df..^2
+write.csv(vif_gear_cor_nov, "../output/bio/vif/monthly_vifs/vif_gear_cor_nov.csv", row.names = TRUE)
+vif_gear_cor_nov
+
+#without gear
+vif_nogear_cor_nov <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = novpresab))
+write.csv(vif_nogear_cor_nov, "../output/bio/vif/monthly_vifs/vif_nogear_cor_nov.csv", row.names = TRUE)
+vif_nogear_cor_nov
+
+#with gear
+vif_gear_cor_dec <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = decpresab))
+write.csv(vif_gear_cor_dec, "../output/bio/vif/monthly_vifs/vif_gear_cor_dec.csv", row.names = TRUE)
+vif_gear_cor_dec
+
+#without gear
+vif_nogear_cor_dec <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = decpresab))
+write.csv(vif_nogear_cor_dec, "../output/bio/vif/monthly_vifs/vif_nogear_cor_dec.csv", row.names = TRUE)
+vif_nogear_cor_dec
+```
+
+in the presences, which rows are missing either temp_depth, salinity_depth, or o2_depth?
+```{r}
+fmatch("o2_depth", names(presence)) #32
+fmatch("salinity_depth", names(presence)) #34
+fmatch("temp_depth", names(presence)) #37
+```
+```{r}
+presence_wdepth_noo2 <- presence[is.na(presence$o2_depth), ]
+nrow(presence_wdepth_noo2) #just to check it has mapped
+presence_wdepth_nosalinity <- presence[is.na(presence$salinity_depth), ]
+nrow(presence_wdepth_nosalinity) #just to check it has mapped
+presence_wdepth_notemp <- presence[is.na(presence$temp_depth), ]
+nrow(presence_wdepth_notemp) #just to check it has mapped
+```
+
+ok remove all rows where temp_depth is missing (3729 rows)
+```{r}
+presab_missingvals <- presab[!is.na(presab$temp_depth), ]
+nrow(presab_missingvals) #126570 - correct
+```
+check for missing vals
+```{r}
+presence_wdepth_noo2 <- presab_missingvals[is.na(presab_missingvals$o2_depth), ]
+nrow(presence_wdepth_noo2) #just to check it has mapped
+presence_wdepth_nosalinity <- presab_missingvals[is.na(presab_missingvals$salinity_depth), ]
+nrow(presence_wdepth_nosalinity) #just to check it has mapped
+presence_wdepth_notemp <- presab_missingvals[is.na(presab_missingvals$temp_depth), ]
+nrow(presence_wdepth_notemp) #just to check it has mapped
+```
+ok now remove row where o2_depth is missing
+```{r}
+presab_missingvals <- presab_missingvals[!is.na(presab_missingvals$o2_depth), ]
+nrow(presab_missingvals) #126360 - correct
+```
+ok as summaries observations per month
+
+```{r}
+presab_missingvals_obs <- subset(presab_missingvals, occurrence == "1")
+obs_by_month <- count(presab_missingvals_obs, "month")
+obs_by_month
+```
+
+and before it was...
+
+```{r}
+obs_by_month_pres <- count(presence, "month")
+obs_by_month_pres
+```
+
+just check the obs that you removed (saved as presence_na.csv) to see if the reported depth is deeper than the gebco derrived bottom depth
+
+```{r}
+presence_na_greater <- subset(presence_NA, depth_layer > bottom_depth_glorys)
+presence_na_greater 
+```
+
+ok potentially i might be able to claw back 150 points...im not sure its worth it
 

### 2019-02-06 src/data_exploration.Rmd
i hate summary stats

@@
 
-plot by each variable, less 3m (2 samples) and 5ze (2 samples)
+plot by each variable, less 3m (2 Sample) and 5ze (2 samples)
 ```{r presence by nafo by variable}
@@
 nrow(presab_missingvals) #126360 - correct
+write.csv(presab_missingvals, "../output/bio/presab_alldata.csv", row.names = FALSE)
+prese_nomissing <- subset(presab_missingvals, occurrence == "1")
+write.csv(prese_nomissing, "../output/bio/pres_nomissing.csv", row.names = FALSE) # this is just to do some mapping in arcgis
 ```
@@
 
-ok potentially i might be able to claw back 150 points...im not sure its worth it
+ok potentially i might be able to claw back 150 points...im not sure its worth it.
+
+ok so redo the occurrence by month/nafo region
+
+```{r prsence nafo by month with complete data}
+nafo_by_mth <- with(presab_missingvals_obs, table(nafo_zone, month))
+write.csv(nafo_by_mth, file = "../output/bio/nafozone_by_mth_missingvalsout.csv")
+nafo_by_mth
+```
+and year-month table
+
+and just a year-month table
+```{r observations by year-month with complete data}
+year_by_mt <- with(presab_missingvals_obs, table(year, month))
+write.csv(year_by_mt, file = "../output/bio/year_by_mth_missingvalsout.csv")
+year_by_mt
+```
+
+ok now as per MJ/Chan, count of cell sampled by year, and then cell sampled by month (do this on the data with obs with missing env. data out)
+
+```{r}
+timescellsampledtotal <- count(presab_missingvals_obs, "cell_id_3d")
+write.csv(timescellsampledtotal, "../output/bio/chan/timescellsampledtotal.csv", row.names = FALSE)
+names(timescellsampledtotal)[names(timescellsampledtotal)=="freq"] <- "samples"
+summarytimescellsampledtotal <- count(timescellsampledtotal, "samples")
+write.csv(timescellsampledtotal, "../output/bio/chan/summarytimescellsampledtotal.csv", row.names = FALSE)
+```
+
+ok now by year
+
+```{r}
+year <- subset(presab_missingvals_obs, year == "1998")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled1998.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled1998.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "1999")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled1999.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled1999.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2000")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2000.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2000.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2001")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2001.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2001.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2002")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2002.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2002.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2003")
+timescellsampledyr <- count(year, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2003.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2003.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2004")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2004.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2004.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2005")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2005.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2005.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2006")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2006.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2006.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2007")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2007.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2007.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2008")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2008.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2008.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2009")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2009.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2009.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2010")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2010.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2010.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2011")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2011.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2011.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2012")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2012.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2012.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2013")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2013.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2013.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2014")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2014.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2014.csv", row.names = FALSE)
+
+year <- subset(presab_missingvals_obs, year == "2015")
+timescellsampledyr <- count(year, "cell_id_3d")
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2015.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2015.csv", row.names = FALSE)
+```
+
+by month
+```{r}
+month <- subset(presab_missingvals_obs, month == "1")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled1.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled1.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "2")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled2.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled2.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "3")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled3.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled3.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "4")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled4.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled4.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "5")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled5.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled5.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "6")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled6.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled6.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "7")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled7.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled7.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "8")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled8.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled8.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "9")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled9.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled9.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "10")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled10.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled10.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "11")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled11.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled11.csv", row.names = FALSE)
+
+month <- subset(presab_missingvals_obs, month == "12")
+timescellsampledyr <- count(month, "cell_id_3d")
+write.csv(timescellsampledyr, file = "../output/bio/chan/timescellsampled12.csv")
+names(timescellsampledyr)[names(timescellsampledyr)=="freq"] <- "samples"
+summaryyear <- count(timescellsampledyr, "samples")
+write.csv(summaryyear, "../output/bio/chan/summarytimescellsampled12.csv", row.names = FALSE)
+
+timescellsampledyr$cell_id_3d <- as.character(timescellsampledyr$cell_id_3d)
+timescellsampledyr$cell_id <- as.numeric(sapply(strsplit(timescellsampledyr$cell_id_3d, "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
+head(timescellsampledyr)
+```
+
+```{r summary stats for env variable by year and month for presence points}
+temp_depth <- describeBy(presence$temp_depth, presence$year, mat = TRUE, digits = 2)
+temp_depth <- subset(temp_depth, select = c(group1, min, max, mean, range))
+names(temp_depth)[names(temp_depth)=="group1"] <- "year"
+write.csv(temp_depth, "../output/env/env_sumstats/temp_depth_presence_summarystats_year.csv", row.names = FALSE)
+
+temp_depth <- describeBy(presence$temp_depth, presence$month, mat = TRUE, digits = 2)
+temp_depth <- subset(temp_depth, select = c(group1, min, max, mean, range))
+names(temp_depth)[names(temp_depth)=="group1"] <- "month"
+write.csv(temp_depth, "../output/env/env_sumstats/temp_depth_presence_summarystats_month.csv", row.names = FALSE)
+
+temp_surface <- describeBy(presence$temp_surface, presence$year, mat = TRUE, digits = 2)
+temp_surface <- subset(temp_surface, select = c(group1, min, max, mean, range))
+names(temp_surface)[names(temp_surface)=="group1"] <- "year"
+write.csv(temp_surface, "../output/env/env_sumstats/temp_dsurface_presence_summarystats_year.csv", row.names = FALSE)
+
+temp_surface <- describeBy(presence$temp_depth, presence$month, mat = TRUE, digits = 2)
+temp_surface <- subset(temp_surface, select = c(group1, min, max, mean, range))
+names(temp_surface)[names(temp_surface)=="group1"] <- "month"
+write.csv(temp_surface, "../output/env/env_sumstats/temp_surface_presence_summarystats_month.csv", row.names = FALSE)
+
+salinity_depth <- describeBy(presence$salinity_depth, presence$year, mat = TRUE, digits = 2)
+salinity_depth <- subset(salinity_depth, select = c(group1, min, max, mean, range))
+names(salinity_depth)[names(salinity_depth)=="group1"] <- "year"
+write.csv(salinity_depth, "../output/env/env_sumstats/salinity_depth_presence_summarystats_year.csv", row.names = FALSE)
+
+salinity_depth <- describeBy(presence$salinity_depth, presence$month, mat = TRUE, digits = 2)
+salinity_depth <- subset(salinity_depth, select = c(group1, min, max, mean, range))
+names(salinity_depth)[names(salinity_depth)=="group1"] <- "month"
+write.csv(salinity_depth, "../output/env/env_sumstats/salinity_depth_presence_summarystats_month.csv", row.names = FALSE)
+
+salinity_surface <- describeBy(presence$salinity_surface, presence$year, mat = TRUE, digits = 2)
+salinity_surface <- subset(salinity_surface, select = c(group1, min, max, mean, range))
+names(salinity_surface)[names(salinity_surface)=="group1"] <- "year"
+write.csv(salinity_surface, "../output/env/env_sumstats/salinity_dsurface_presence_summarystats_year.csv", row.names = FALSE)
+
+salinity_surface <- describeBy(presence$salinity_surface, presence$month, mat = TRUE, digits = 2)
+salinity_surface <- subset(salinity_surface, select = c(group1, min, max, mean, range))
+names(salinity_surface)[names(salinity_surface)=="group1"] <- "month"
+write.csv(salinity_surface, "../output/env/env_sumstats/salinity_surface_presence_summarystats_month.csv", row.names = FALSE)
+
+o2_depth <- describeBy(presence$o2_depth, presence$year, mat = TRUE, digits = 2)
+o2_depth <- subset(o2_depth, select = c(group1, min, max, mean, range))
+names(o2_depth)[names(o2_depth)=="group1"] <- "year"
+write.csv(o2_depth, "../output/env/env_sumstats/o2_depth_presence_summarystats_year.csv", row.names = FALSE)
+
+o2_depth <- describeBy(presence$o2_depth, presence$month, mat = TRUE, digits = 2)
+o2_depth <- subset(o2_depth, select = c(group1, min, max, mean, range))
+names(o2_depth)[names(o2_depth)=="group1"] <- "month"
+write.csv(o2_depth, "../output/env/env_sumstats/o2_depth_presence_summarystats_month.csv", row.names = FALSE)
+
+o2_surface <- describeBy(presence$o2_surface, presence$year, mat = TRUE, digits = 2)
+o2_surface <- subset(o2_surface, select = c(group1, min, max, mean, range))
+names(o2_surface)[names(o2_surface)=="group1"] <- "year"
+write.csv(o2_surface, "../output/env/env_sumstats/o2_dsurface_presence_summarystats_year.csv", row.names = FALSE)
+
+o2_surface <- describeBy(presence$o2_surface, presence$month, mat = TRUE, digits = 2)
+o2_surface <- subset(o2_surface, select = c(group1, min, max, mean, range))
+names(o2_surface)[names(o2_surface)=="group1"] <- "month"
+write.csv(o2_surface, "../output/env/env_sumstats/o2_surface_presence_summarystats_month.csv", row.names = FALSE)
+
+chl_depth <- describeBy(presence$chl_depth, presence$year, mat = TRUE, digits = 2)
+chl_depth <- subset(chl_depth, select = c(group1, min, max, mean, range))
+names(chl_depth)[names(chl_depth)=="group1"] <- "year"
+write.csv(chl_depth, "../output/env/env_sumstats/chl_depth_presence_summarystats_year.csv", row.names = FALSE)
+
+chl_depth <- describeBy(presence$chl_depth, presence$month, mat = TRUE, digits = 2)
+chl_depth <- subset(chl_depth, select = c(group1, min, max, mean, range))
+names(chl_depth)[names(chl_depth)=="group1"] <- "month"
+write.csv(chl_depth, "../output/env/env_sumstats/chl_depth_presence_summarystats_month.csv", row.names = FALSE)
+
+chl_surface <- describeBy(presence$chl_surface, presence$year, mat = TRUE, digits = 2)
+chl_surface <- subset(chl_surface, select = c(group1, min, max, mean, range))
+names(chl_surface)[names(chl_surface)=="group1"] <- "year"
+write.csv(chl_surface, "../output/env/env_sumstats/chl_surface_presence_summarystats_year.csv", row.names = FALSE)
+
+chl_surface <- describeBy(presence$chl_surface, presence$month, mat = TRUE, digits = 2)
+chl_surface <- subset(chl_surface, select = c(group1, min, max, mean, range))
+names(chl_surface)[names(chl_surface)=="group1"] <- "month"
+write.csv(chl_surface, "../output/env/env_sumstats/chl_surface_presence_summarystats_month.csv", row.names = FALSE)
+
+mlp <- describeBy(presence$mlp_surface, presence$year, mat = TRUE, digits = 2)
+mlp <- subset(mlp, select = c(group1, min, max, mean, range))
+names(mlp)[names(mlp)=="group1"] <- "year"
+write.csv(mlp, "../output/env/env_sumstats/mlp_presence_summarystats_year.csv", row.names = FALSE)
+
+mlp <- describeBy(presence$mlp_surface, presence$month, mat = TRUE, digits = 2)
+mlp <- subset(mlp, select = c(group1, min, max, mean, range))
+names(mlp)[names(mlp)=="group1"] <- "month"
+write.csv(mlp, "../output/env/env_sumstats/mlp_presence_summarystats_month.csv", row.names = FALSE)
+
+ssh <- describeBy(presence$ssh_surface, presence$year, mat = TRUE, digits = 2)
+ssh <- subset(ssh, select = c(group1, min, max, mean, range))
+names(ssh)[names(chl_depth)=="group1"] <- "year"
+write.csv(ssh, "../output/env/env_sumstats/ssh_presence_summarystats_year.csv", row.names = FALSE)
+
+ssh <- describeBy(presence$ssh_surface, presence$month, mat = TRUE, digits = 2)
+ssh <- subset(ssh, select = c(group1, min, max, mean, range))
+names(ssh)[names(chl_depth)=="group1"] <- "month"
+write.csv(ssh, "../output/env/env_sumstats/ssh_presence_summarystats_month.csv", row.names = FALSE)
+
+amo_sample <- describeBy(presence$amo_sample, presence$year, mat = TRUE, digits = 2)
+amo_sample <- subset(amo_sample, select = c(group1, min, max, mean, range))
+names(amo_sample)[names(amo_sample)=="group1"] <- "year"
+write.csv(amo_sample, "../output/env/env_sumstats/amo_sample_presence_summarystats_year.csv", row.names = FALSE)
+
+amo_sample <- describeBy(presence$amo_sample, presence$month, mat = TRUE, digits = 2)
+amo_sample <- subset(amo_sample, select = c(group1, min, max, mean, range))
+names(amo_sample)[names(chl_depth)=="group1"] <- "month"
+write.csv(amo_sample, "../output/env/env_sumstats/amo_sample_presence_summarystats_month.csv", row.names = FALSE)
+
+amo_prev <- describeBy(presence$amo_prev, presence$year, mat = TRUE, digits = 2)
+amo_prev <- subset(amo_prev, select = c(group1, min, max, mean, range))
+names(amo_prev)[names(amo_prev)=="group1"] <- "year"
+write.csv(amo_prev, "../output/env/env_sumstats/amo_prev_presence_summarystats_year.csv", row.names = FALSE)
+
+amo_prev <- describeBy(presence$amo_prev, presence$month, mat = TRUE, digits = 2)
+amo_prev <- subset(amo_prev, select = c(group1, min, max, mean, range))
+names(amo_prev)[names(chl_depth)=="group1"] <- "month"
+write.csv(amo_prev, "../output/env/env_sumstats/amo_prev_presence_summarystats_month.csv", row.names = FALSE)
+
+amo_winter <- describeBy(presence$amo_winter, presence$year, mat = TRUE, digits = 2)
+amo_winter <- subset(amo_winter, select = c(group1, min, max, mean, range))
+names(amo_winter)[names(amo_winter)=="group1"] <- "year"
+write.csv(amo_winter, "../output/env/env_sumstats/amo_winter_presence_summarystats_year.csv", row.names = FALSE)
+
+amo_winter <- describeBy(presence$amo_winter, presence$month, mat = TRUE, digits = 2)
+amo_winter <- subset(amo_winter, select = c(group1, min, max, mean, range))
+names(amo_winter)[names(chl_depth)=="group1"] <- "month"
+write.csv(amo_winter, "../output/env/env_sumstats/amo_winter_presence_summarystats_month.csv", row.names = FALSE)
+
+nao_sample <- describeBy(presence$nao_sample, presence$year, mat = TRUE, digits = 2)
+nao_sample <- subset(nao_sample, select = c(group1, min, max, mean, range))
+names(nao_sample)[names(nao_sample)=="group1"] <- "year"
+write.csv(nao_sample, "../output/env/env_sumstats/nao_sample_presence_summarystats_year.csv", row.names = FALSE)
+
+nao_sample <- describeBy(presence$nao_sample, presence$month, mat = TRUE, digits = 2)
+nao_sample <- subset(nao_sample, select = c(group1, min, max, mean, range))
+names(nao_sample)[names(chl_depth)=="group1"] <- "month"
+write.csv(nao_sample, "../output/env/env_sumstats/nao_sample_presence_summarystats_month.csv", row.names = FALSE)
+
+nao_prev <- describeBy(presence$nao_prev, presence$year, mat = TRUE, digits = 2)
+nao_prev <- subset(nao_prev, select = c(group1, min, max, mean, range))
+names(nao_prev)[names(nao_prev)=="group1"] <- "year"
+write.csv(nao_prev, "../output/env/env_sumstats/nao_prev_presence_summarystats_year.csv", row.names = FALSE)
+
+nao_prev <- describeBy(presence$nao_prev, presence$month, mat = TRUE, digits = 2)
+nao_prev <- subset(nao_prev, select = c(group1, min, max, mean, range))
+names(nao_prev)[names(chl_depth)=="group1"] <- "month"
+write.csv(nao_prev, "../output/env/env_sumstats/nao_prev_presence_summarystats_month.csv", row.names = FALSE)
+
+nao_winter <- describeBy(presence$nao_winter, presence$year, mat = TRUE, digits = 2)
+nao_winter <- subset(nao_winter, select = c(group1, min, max, mean, range))
+names(nao_winter)[names(nao_winter)=="group1"] <- "year"
+write.csv(nao_winter, "../output/env/env_sumstats/nao_winter_presence_summarystats_year.csv", row.names = FALSE)
+
+nao_winter <- describeBy(presence$nao_winter, presence$month, mat = TRUE, digits = 2)
+nao_winter <- subset(nao_winter, select = c(group1, min, max, mean, range))
+names(nao_winter)[names(chl_depth)=="group1"] <- "month"
+write.csv(nao_winter, "../output/env/env_sumstats/nao_winter_presence_summarystats_month.csv", row.names = FALSE)
+```
+
+and on background points
+
+```{r summary stats for env variable by year and month for background points}
+temp_depth <- describeBy(background$temp_depth, background$year, mat = TRUE, digits = 2)
+temp_depth <- subset(temp_depth, select = c(group1, min, max, mean, range))
+names(temp_depth)[names(temp_depth)=="group1"] <- "year"
+write.csv(temp_depth, "../output/env/env_sumstats/temp_depth_background_summarystats_year.csv", row.names = FALSE)
+
+temp_depth <- describeBy(background$temp_depth, background$month, mat = TRUE, digits = 2)
+temp_depth <- subset(temp_depth, select = c(group1, min, max, mean, range))
+names(temp_depth)[names(temp_depth)=="group1"] <- "month"
+write.csv(temp_depth, "../output/env/env_sumstats/temp_depth_background_summarystats_month.csv", row.names = FALSE)
+
+temp_surface <- describeBy(background$temp_surface, background$year, mat = TRUE, digits = 2)
+temp_surface <- subset(temp_surface, select = c(group1, min, max, mean, range))
+names(temp_surface)[names(temp_surface)=="group1"] <- "year"
+write.csv(temp_surface, "../output/env/env_sumstats/temp_dsurface_background_summarystats_year.csv", row.names = FALSE)
+
+temp_surface <- describeBy(background$temp_depth, background$month, mat = TRUE, digits = 2)
+temp_surface <- subset(temp_surface, select = c(group1, min, max, mean, range))
+names(temp_surface)[names(temp_surface)=="group1"] <- "month"
+write.csv(temp_surface, "../output/env/env_sumstats/temp_surface_background_summarystats_month.csv", row.names = FALSE)
+
+salinity_depth <- describeBy(background$salinity_depth, background$year, mat = TRUE, digits = 2)
+salinity_depth <- subset(salinity_depth, select = c(group1, min, max, mean, range))
+names(salinity_depth)[names(salinity_depth)=="group1"] <- "year"
+write.csv(salinity_depth, "../output/env/env_sumstats/salinity_depth_background_summarystats_year.csv", row.names = FALSE)
+
+salinity_depth <- describeBy(background$salinity_depth, background$month, mat = TRUE, digits = 2)
+salinity_depth <- subset(salinity_depth, select = c(group1, min, max, mean, range))
+names(salinity_depth)[names(salinity_depth)=="group1"] <- "month"
+write.csv(salinity_depth, "../output/env/env_sumstats/salinity_depth_background_summarystats_month.csv", row.names = FALSE)
+
+salinity_surface <- describeBy(background$salinity_surface, background$year, mat = TRUE, digits = 2)
+salinity_surface <- subset(salinity_surface, select = c(group1, min, max, mean, range))
+names(salinity_surface)[names(salinity_surface)=="group1"] <- "year"
+write.csv(salinity_surface, "../output/env/env_sumstats/salinity_dsurface_background_summarystats_year.csv", row.names = FALSE)
+
+salinity_surface <- describeBy(background$salinity_surface, background$month, mat = TRUE, digits = 2)
+salinity_surface <- subset(salinity_surface, select = c(group1, min, max, mean, range))
+names(salinity_surface)[names(salinity_surface)=="group1"] <- "month"
+write.csv(salinity_surface, "../output/env/env_sumstats/salinity_surface_background_summarystats_month.csv", row.names = FALSE)
+
+o2_depth <- describeBy(background$o2_depth, background$year, mat = TRUE, digits = 2)
+o2_depth <- subset(o2_depth, select = c(group1, min, max, mean, range))
+names(o2_depth)[names(o2_depth)=="group1"] <- "year"
+write.csv(o2_depth, "../output/env/env_sumstats/o2_depth_background_summarystats_year.csv", row.names = FALSE)
+
+o2_depth <- describeBy(background$o2_depth, background$month, mat = TRUE, digits = 2)
+o2_depth <- subset(o2_depth, select = c(group1, min, max, mean, range))
+names(o2_depth)[names(o2_depth)=="group1"] <- "month"
+write.csv(o2_depth, "../output/env/env_sumstats/o2_depth_background_summarystats_month.csv", row.names = FALSE)
+
+o2_surface <- describeBy(background$o2_surface, background$year, mat = TRUE, digits = 2)
+o2_surface <- subset(o2_surface, select = c(group1, min, max, mean, range))
+names(o2_surface)[names(o2_surface)=="group1"] <- "year"
+write.csv(o2_surface, "../output/env/env_sumstats/o2_dsurface_background_summarystats_year.csv", row.names = FALSE)
+
+o2_surface <- describeBy(background$o2_surface, background$month, mat = TRUE, digits = 2)
+o2_surface <- subset(o2_surface, select = c(group1, min, max, mean, range))
+names(o2_surface)[names(o2_surface)=="group1"] <- "month"
+write.csv(o2_surface, "../output/env/env_sumstats/o2_surface_background_summarystats_month.csv", row.names = FALSE)
+
+chl_depth <- describeBy(background$chl_depth, background$year, mat = TRUE, digits = 2)
+chl_depth <- subset(chl_depth, select = c(group1, min, max, mean, range))
+names(chl_depth)[names(chl_depth)=="group1"] <- "year"
+write.csv(chl_depth, "../output/env/env_sumstats/chl_depth_background_summarystats_year.csv", row.names = FALSE)
+
+chl_depth <- describeBy(background$chl_depth, background$month, mat = TRUE, digits = 2)
+chl_depth <- subset(chl_depth, select = c(group1, min, max, mean, range))
+names(chl_depth)[names(chl_depth)=="group1"] <- "month"
+write.csv(chl_depth, "../output/env/env_sumstats/chl_depth_background_summarystats_month.csv", row.names = FALSE)
+
+chl_surface <- describeBy(background$chl_surface, background$year, mat = TRUE, digits = 2)
+chl_surface <- subset(chl_surface, select = c(group1, min, max, mean, range))
+names(chl_surface)[names(chl_surface)=="group1"] <- "year"
+write.csv(chl_surface, "../output/env/env_sumstats/chl_surface_background_summarystats_year.csv", row.names = FALSE)
+
+chl_surface <- describeBy(background$chl_surface, background$month, mat = TRUE, digits = 2)
+chl_surface <- subset(chl_surface, select = c(group1, min, max, mean, range))
+names(chl_surface)[names(chl_surface)=="group1"] <- "month"
+write.csv(chl_surface, "../output/env/env_sumstats/chl_surface_background_summarystats_month.csv", row.names = FALSE)
+
+mlp <- describeBy(background$mlp_surface, background$year, mat = TRUE, digits = 2)
+mlp <- subset(mlp, select = c(group1, min, max, mean, range))
+names(mlp)[names(mlp)=="group1"] <- "year"
+write.csv(mlp, "../output/env/env_sumstats/mlp_background_summarystats_year.csv", row.names = FALSE)
+
+mlp <- describeBy(background$mlp_surface, background$month, mat = TRUE, digits = 2)
+mlp <- subset(mlp, select = c(group1, min, max, mean, range))
+names(mlp)[names(mlp)=="group1"] <- "month"
+write.csv(mlp, "../output/env/env_sumstats/mlp_background_summarystats_month.csv", row.names = FALSE)
+
+ssh <- describeBy(background$ssh_surface, background$year, mat = TRUE, digits = 2)
+ssh <- subset(ssh, select = c(group1, min, max, mean, range))
+names(ssh)[names(chl_depth)=="group1"] <- "year"
+write.csv(ssh, "../output/env/env_sumstats/ssh_background_summarystats_year.csv", row.names = FALSE)
+
+ssh <- describeBy(background$ssh_surface, background$month, mat = TRUE, digits = 2)
+ssh <- subset(ssh, select = c(group1, min, max, mean, range))
+names(ssh)[names(chl_depth)=="group1"] <- "month"
+write.csv(ssh, "../output/env/env_sumstats/ssh_background_summarystats_month.csv", row.names = FALSE)
+
+amo_sample <- describeBy(background$amo_sample, background$year, mat = TRUE, digits = 2)
+amo_sample <- subset(amo_sample, select = c(group1, min, max, mean, range))
+names(amo_sample)[names(amo_sample)=="group1"] <- "year"
+write.csv(amo_sample, "../output/env/env_sumstats/amo_sample_background_summarystats_year.csv", row.names = FALSE)
+
+amo_sample <- describeBy(background$amo_sample, background$month, mat = TRUE, digits = 2)
+amo_sample <- subset(amo_sample, select = c(group1, min, max, mean, range))
+names(amo_sample)[names(chl_depth)=="group1"] <- "month"
+write.csv(amo_sample, "../output/env/env_sumstats/amo_sample_background_summarystats_month.csv", row.names = FALSE)
+
+amo_prev <- describeBy(background$amo_prev, background$year, mat = TRUE, digits = 2)
+amo_prev <- subset(amo_prev, select = c(group1, min, max, mean, range))
+names(amo_prev)[names(amo_prev)=="group1"] <- "year"
+write.csv(amo_prev, "../output/env/env_sumstats/amo_prev_background_summarystats_year.csv", row.names = FALSE)
+
+amo_prev <- describeBy(background$amo_prev, background$month, mat = TRUE, digits = 2)
+amo_prev <- subset(amo_prev, select = c(group1, min, max, mean, range))
+names(amo_prev)[names(chl_depth)=="group1"] <- "month"
+write.csv(amo_prev, "../output/env/env_sumstats/amo_prev_background_summarystats_month.csv", row.names = FALSE)
+
+amo_winter <- describeBy(background$amo_winter, background$year, mat = TRUE, digits = 2)
+amo_winter <- subset(amo_winter, select = c(group1, min, max, mean, range))
+names(amo_winter)[names(amo_winter)=="group1"] <- "year"
+write.csv(amo_winter, "../output/env/env_sumstats/amo_winter_background_summarystats_year.csv", row.names = FALSE)
+
+amo_winter <- describeBy(background$amo_winter, background$month, mat = TRUE, digits = 2)
+amo_winter <- subset(amo_winter, select = c(group1, min, max, mean, range))
+names(amo_winter)[names(chl_depth)=="group1"] <- "month"
+write.csv(amo_winter, "../output/env/env_sumstats/amo_winter_background_summarystats_month.csv", row.names = FALSE)
+
+nao_sample <- describeBy(background$nao_sample, background$year, mat = TRUE, digits = 2)
+nao_sample <- subset(nao_sample, select = c(group1, min, max, mean, range))
+names(nao_sample)[names(nao_sample)=="group1"] <- "year"
+write.csv(nao_sample, "../output/env/env_sumstats/nao_sample_background_summarystats_year.csv", row.names = FALSE)
+
+nao_sample <- describeBy(background$nao_sample, background$month, mat = TRUE, digits = 2)
+nao_sample <- subset(nao_sample, select = c(group1, min, max, mean, range))
+names(nao_sample)[names(chl_depth)=="group1"] <- "month"
+write.csv(nao_sample, "../output/env/env_sumstats/nao_sample_background_summarystats_month.csv", row.names = FALSE)
+
+nao_prev <- describeBy(background$nao_prev, background$year, mat = TRUE, digits = 2)
+nao_prev <- subset(nao_prev, select = c(group1, min, max, mean, range))
+names(nao_prev)[names(nao_prev)=="group1"] <- "year"
+write.csv(nao_prev, "../output/env/env_sumstats/nao_prev_background_summarystats_year.csv", row.names = FALSE)
+
+nao_prev <- describeBy(background$nao_prev, background$month, mat = TRUE, digits = 2)
+nao_prev <- subset(nao_prev, select = c(group1, min, max, mean, range))
+names(nao_prev)[names(chl_depth)=="group1"] <- "month"
+write.csv(nao_prev, "../output/env/env_sumstats/nao_prev_background_summarystats_month.csv", row.names = FALSE)
+
+nao_winter <- describeBy(background$nao_winter, background$year, mat = TRUE, digits = 2)
+nao_winter <- subset(nao_winter, select = c(group1, min, max, mean, range))
+names(nao_winter)[names(nao_winter)=="group1"] <- "year"
+write.csv(nao_winter, "../output/env/env_sumstats/nao_winter_background_summarystats_year.csv", row.names = FALSE)
+
+nao_winter <- describeBy(background$nao_winter, background$month, mat = TRUE, digits = 2)
+nao_winter <- subset(nao_winter, select = c(group1, min, max, mean, range))
+names(nao_winter)[names(chl_depth)=="group1"] <- "month"
+write.csv(nao_winter, "../output/env/env_sumstats/nao_winter_background_summarystats_month.csv", row.names = FALSE)
+```
 

### 2019-02-11 src/data_exploration.Rmd
update to one vif + some early maxent setup with dismo (just to test - maxent will be run in a seperate script)

@@
 #with gear
-vif_gear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
+vif_gear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
 write.csv(vif_gear_cor_apr, "../output/bio/vif/monthly_vifs/vif_gear_cor_apr.csv", row.names = TRUE)
@@
 #without gear
-vif_nogear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
+vif_nogear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
 write.csv(vif_nogear_cor_apr, "../output/bio/vif/monthly_vifs/vif_nogear_cor_apr.csv", row.names = TRUE)
@@
 #with gear
-vif_gear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+vif_gear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
 write.csv(vif_gear_cor_jun, "../output/bio/vif/monthly_vifs/vif_gear_cor_jun.csv", row.names = TRUE)
@@
 #without gear
-vif_nogear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+vif_nogear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
 write.csv(vif_nogear_cor_jun, "../output/bio/vif/monthly_vifs/vif_nogear_cor_jun.csv", row.names = TRUE)
@@
 #with gear
-vif_gear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
+vif_gear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
 write.csv(vif_gear_cor_oct, "../output/bio/vif/monthly_vifs/vif_gear_cor_oct.csv", row.names = TRUE)
@@
 #without gear
-vif_nogear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
+vif_nogear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
 write.csv(vif_nogear_cor_oct, "../output/bio/vif/monthly_vifs/vif_nogear_cor_oct.csv", row.names = TRUE)
@@
 
+# chan/mj - samples per cell by year and month. 
 ok now as per MJ/Chan, count of cell sampled by year, and then cell sampled by month (do this on the data with obs with missing env. data out)
+note this is a summary of the month/year - cells can be sampled more than one, but never more than once at the same month/year
 
@@
 
-```{r}
-library("raster")
-library("dismo")
-library("rgeos")
-```
+This test will be run on january data
+
+subset the presab dataset to january only
+
+```{r}
+janpresabmax <- subset(presab_missingvals, month == "1")
+head(janpresabmax)
+```
+
+ok so maxent files need to be in a particular format: species name, x (longitude), y (latitude), env variable, env variable,....
+
+Maxent doesn't have a way to deal with 3D locations (it only takes xy coordinates), however it is possible to use dummy xy coordinates (see appendicies of Elith 2011 A statistical explanation of MaxEnt for ecologists).
+In my case, I will use the 'id' (the unique id assigned to the sample) for x and "cell_id_xyzt" (the unique cell id on xyzt) for y.
+I will also want to split the dataset by presence and background, so include here
+
+```{r}
+janpresabmax <- subset (janpresabmax, select = c(occurrence, originalscientificname, id, cell_id_xyzt, temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_winter))
+head(janpresabmax)
+```
+
+now split into presence and background
+
+```{r}
+janpresmax <- subset(janpresabmax, occurrence == "1")
+janpresmax <- subset (janpresmax, select = -c(occurrence))
+janabmax <- subset(janpresabmax, occurrence == "0")
+janabmax <- subset (janabmax, select = -c(occurrence))
+```
+
+ok so for maxent you need to install the jar file.
+see https://cran.r-project.org/web/packages/dismo/vignettes/sdm.pdf but note that the location of the jar file has changed http://biodiversityinformatics.amnh.org/open_source/maxent/
+
+citation: Steven J.
+Phillips, Miroslav Dudík, Robert E.
+Schapire.
+[Internet] Maxent software for modeling species niches and distributions (Version 3.4.1).
+Available from url: http://biodiversityinformatics.amnh.org/open_source/maxent/.
+Accessed on 2019-2-11.
+
+```{r}
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+```
+
+```{r}
+library(dismo)
+library(rJava)
+jar <- paste(system.file(package="dismo"), "C:/Users/bunny/Documents/GitHub/Chapter_1/packrat/lib/x86_64-w64-mingw32/3.5.1/dismo/java/maxent.jar", sep='')
+testmax <- maxent(janpresmax, janabmax)
+```

### 2019-02-12 src/data_exploration.Rmd
very minor changes... currently working in BASH

@@
 janpresabmax <- subset (janpresabmax, select = c(occurrence, originalscientificname, id, cell_id_xyzt, temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_winter))
+colnames(janpresabmax)[colnames(janpresabmax)=="cell_id_xyzt"] <- "x"
+colnames(janpresabmax)[colnames(janpresabmax)=="id"] <- "y"
+janpresabmax$x <- gsub("_", "", janpresabmax$x) #this is needed as maxent thinks these are coordinates
 head(janpresabmax)
@@
 janpresmax <- subset (janpresmax, select = -c(occurrence))
+write.csv(janpresmax, "../output/bio/maxenttestjanpr.csv", row.names = FALSE)
 janabmax <- subset(janpresabmax, occurrence == "0")
 janabmax <- subset (janabmax, select = -c(occurrence))
+write.csv(janabmax, "../output/bio/maxenttestjanab.csv", row.names = FALSE)
 ```
@@
 jar <- paste(system.file(package="dismo"), "C:/Users/bunny/Documents/GitHub/Chapter_1/packrat/lib/x86_64-w64-mingw32/3.5.1/dismo/java/maxent.jar", sep='')
+p
 testmax <- maxent(janpresmax, janabmax)

### 2019-02-20 src/data_exploration.Rmd
playing with patterns and stack creation for prediction layers

@@
 
+loop creating layer stacks
+```{r}
+jan98_86.9 <- stack(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*_86.9294*.asc"), full.names = TRUE))    # this folder currently contains other dummy files.
+its just a test to automate stacks
+names(jan98_86.9) # now need to rename to match maxent expectations
+names(jan98_86.9) <- (pattern = "*chl_surface*" paste0("chl_surface"))
+```
+
 ```{r}
@@
 jar <- paste(system.file(package="dismo"), "C:/Users/bunny/Documents/GitHub/Chapter_1/packrat/lib/x86_64-w64-mingw32/3.5.1/dismo/java/maxent.jar", sep='')
-p
-testmax <- maxent(janpresmax, janabmax)
-```
+
+testmax <- maxent(janpresabmax, janpresabmax$occurrence)
+
+```

### 2019-02-21 src/data_exploration.Rmd
stack building - could do with loop

--- 2019-02-12 src/data_exploration.Rmd
@@
 
+zt stratified stack with layer names that match swd
+```{r}
+chl_surface <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*chl_surface_86.9294*.asc"), full.names = TRUE))
+names(chl_surface) <- "chl_surface"
+nao_sample <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*nao_sample_86.9294*.asc"), full.names = TRUE))
+names(nao_sample) <- "nao_sample"
+nao_winter <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*nao_winter_86.9294*.asc"), full.names = TRUE))
+names(nao_winter) <- "nao_winter"
+o2_depth <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*o2_depth_86.9294*.asc"), full.names = TRUE)) 
+names(o2_depth) <- "o2_depth"
+salinity_depth <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*salinity_depth_86.9294*.asc"), full.names = TRUE))
+names(salinity_depth) <- "salinity_depth"
+temp_depth <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*temp_depth_86.9294*.asc"), full.names = TRUE))
+names(temp_depth) <- "temp_depth"
+
+stack1998_01_86.9294 <- stack(chl_surface, nao_sample, nao_winter, o2_depth, salinity_depth, temp_depth)
+
+```
+
+asclist <- stack(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*_86.9294*.asc"), full.names = TRUE))    # this folder currently contains other dummy files.
+its just a test to automate stacks
+ascfile <- names(asclist[[1]])
+ascyear <- as.character(sapply(strsplit(ascfile[[1]], "_"), "[[", 1))
+ascmonth <- as.character(sapply(strsplit(ascfile[[1]], "_"), "[[", 2))
+ascdepth <- as.character(sapply(strsplit(ascfile[[1]], "_"), "[[", 5))
+ascname <- paste(ascyear, ascmonth, ascdepth, sep = "_") #once you figure out how to rename the layers in the stack, you will rename the stack this () or not - maybe just manual
+asclist # now need to rename to match maxent expectations
+
 ```{r}
@@
 jar <- paste(system.file(package="dismo"), "C:/Users/bunny/Documents/GitHub/Chapter_1/packrat/lib/x86_64-w64-mingw32/3.5.1/dismo/java/maxent.jar", sep='')
-p
-testmax <- maxent(janpresmax, janabmax)
-```
+file.exists(jar)
+testmax <- maxent(janpresabmax, janpresabmax$occurrence)
+
+```

### 2019-02-21 src/data_exploration.Rmd
loop creation - could do with unique year and unique month something... but stack name..

--- 2019-02-12 src/data_exploration.Rmd
@@
 
+zt stratified stack with layer names that match swd
+```{r}
+ascname <- list.files("../data/jan_test_86.9", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+noasc <- length(ascname)
+  year <- "1998"
+  month <- "01"
+  depth <- "86.9294.asc"
+  tempvar <- "temp"
+  chlvar <- "chl"
+  o2var <- "o2"
+  salvar <- "salinity"
+  depthvar <- "depth"
+  surfacevar <- "surface"
+  naovar <- "nao"
+  amovar <- "amo"
+  samplevar <- "sample"
+  wintervar <- "winter"
+
+for (i in 1:length(asclist)){
+  ascyear <- (sapply(strsplit(ascname[i], "_"), "[[", 1))
+  ascmonth <- (sapply(strsplit(ascname[i], "_"), "[[", 2))
+  ascdepth <- (sapply(strsplit(ascname[i], "_"), "[[", 5))
+  ascvar1 <- (sapply(strsplit(ascname[i], "_"), "[[", 3))
+  ascvar2 <- (sapply(strsplit(ascname[i], "_"), "[[", 4))
+  ascfile <- ascname[i]
+
+  if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
+    #print(ascfile)
+    temp_depth <- raster(ascfile)
+    names(temp_depth) <- "temp_depth"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
+    chl_surface <- raster(ascfile)
+    names(chl_surface) <- "chl_surface"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
+    o2_depth <- raster(ascfile)
+    names(o2_depth) <- "o2_depth"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
+    salinity_depth <- raster(ascfile)
+    names(salinity_depth) <- "salinity_depth"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == samplevar){
+    nao_sample <- raster(ascfile)
+    names(nao_sample <- "nao_sample")
+  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == wintervar){
+    nao_winter <- raster(ascfile)
+    names(nao_winter) <- "nao_winter"
+  }
+  st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+
+}
+
+chl_surface <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*chl_surface_86.9294*.asc"), full.names = TRUE))
+names(chl_surface) <- "chl_surface"
+nao_sample <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*nao_sample_86.9294*.asc"), full.names = TRUE))
+names(nao_sample) <- "nao_sample"
+nao_winter <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*nao_winter_86.9294*.asc"), full.names = TRUE))
+names(nao_winter) <- "nao_winter"
+o2_depth <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*o2_depth_86.9294*.asc"), full.names = TRUE)) 
+names(o2_depth) <- "o2_depth"
+salinity_depth <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*salinity_depth_86.9294*.asc"), full.names = TRUE))
+names(salinity_depth) <- "salinity_depth"
+temp_depth <- raster(list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*temp_depth_86.9294*.asc"), full.names = TRUE))
+names(temp_depth) <- "temp_depth"
+
+stack1998_01_86.9294 <- stack(chl_surface, nao_sample, nao_winter, o2_depth, salinity_depth, temp_depth)
+
+```
+
+asclist <- (list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*_86.9294*.asc"), full.names = TRUE))    # this folder currently contains other dummy files.
+its just a test to automate stacks
+ascfile <- names(asclist[[1]])
+ascyear <- as.character(sapply(strsplit(ascfile[[1]], "_"), "[[", 1))
+ascmonth <- as.character(sapply(strsplit(ascfile[[1]], "_"), "[[", 2))
+ascdepth <- as.character(sapply(strsplit(ascfile[[1]], "_"), "[[", 5))
+ascname <- paste(ascyear, ascmonth, ascdepth, sep = "_") #once you figure out how to rename the layers in the stack, you will rename the stack this () or not - maybe just manual
+asclist # now need to rename to match maxent expectations
+
 ```{r}

### 2019-02-21 src/data_exploration.Rmd
make loop clever

@@
 
-asclist <- (list.files("../data/jan_test_86.9", pattern = glob2rx("*1998_01*_86.9294*.asc"), full.names = TRUE))    # this folder currently contains other dummy files.
-its just a test to automate stacks
-ascfile <- names(asclist[[1]])
-ascyear <- as.character(sapply(strsplit(ascfile[[1]], "_"), "[[", 1))
-ascmonth <- as.character(sapply(strsplit(ascfile[[1]], "_"), "[[", 2))
-ascdepth <- as.character(sapply(strsplit(ascfile[[1]], "_"), "[[", 5))
-ascname <- paste(ascyear, ascmonth, ascdepth, sep = "_") #once you figure out how to rename the layers in the stack, you will rename the stack this () or not - maybe just manual
-asclist # now need to rename to match maxent expectations
+```{r}
+```{r}
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+```
+
+messing with code to see if can do for unique year month depth
+```{r}
+ascname <- list.files("../data/jan_test_86.9", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+noasc <- length(ascname)
+  year <- "1998"
+  month <- "01"
+  depth <- "86.9294.asc"
+  tempvar <- "temp"
+  chlvar <- "chl"
+  o2var <- "o2"
+  salvar <- "salinity"
+  depthvar <- "depth"
+  surfacevar <- "surface"
+  naovar <- "nao"
+  amovar <- "amo"
+  samplevar <- "sample"
+  wintervar <- "winter"
+
+for (i in 1:length(asclist)){
+  ascyear <- (sapply(strsplit(ascname[i], "_"), "[[", 1))
+  ascmonth <- (sapply(strsplit(ascname[i], "_"), "[[", 2))
+  ascdepth <- (sapply(strsplit(ascname[i], "_"), "[[", 5))
+  ascvar1 <- (sapply(strsplit(ascname[i], "_"), "[[", 3))
+  ascvar2 <- (sapply(strsplit(ascname[i], "_"), "[[", 4))
+  ascfile <- ascname[i]
+
+  if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
+    #print(ascfile)
+    temp_depth <- raster(ascfile)
+    names(temp_depth) <- "temp_depth"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
+    chl_surface <- raster(ascfile)
+    names(chl_surface) <- "chl_surface"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
+    o2_depth <- raster(ascfile)
+    names(o2_depth) <- "o2_depth"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
+    salinity_depth <- raster(ascfile)
+    names(salinity_depth) <- "salinity_depth"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == samplevar){
+    nao_sample <- raster(ascfile)
+    names(nao_sample <- "nao_sample")
+  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == wintervar){
+    nao_winter <- raster(ascfile)
+    names(nao_winter) <- "nao_winter"
+  }
+  st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+
+}
+```
 

### 2019-02-24 src/stack_test.Rmd
playing with stacking yyyy_mm_ddx

messing with code to see if can do for unique year month depth
```{r}
ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
rasters <- lapply(paste0(ascname), raster)

year <- "X1998"
month <- "01"
depth <- "86.9294"
tempvar <- "temp"
chlvar <- "chl"
o2var <- "o2"
salvar <- "salinity"
depthvar <- "depth"
surfacevar <- "surface"
naovar <- "nao"
amovar <- "amo"
samplevar <- "sample"
wintervar <- "winter"

for (i in 1:length(rasters)){
  ascyear <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 1))
  ascmonth <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 2))
  ascdepth <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 5))
  ascvar1 <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 3))
  ascvar2 <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 4))
  ascfile <- rasters[i]

    if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
    #print(ascfile)
    temp_depth <- ascfile[[1]]
    names(temp_depth) <- "temp_depth"
  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
    chl_surface <- ascfile[[1]]
    names(chl_surface) <- "chl_surface"
  } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
    o2_depth <- ascfile[[1]]
    names(o2_depth) <- "o2_depth"
  } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
    salinity_depth <- ascfile[[1]]
    names(salinity_depth) <- "salinity_depth"
  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == wintervar){
    nao_winter <- ascfile[[1]]
    names(nao_winter) <- "nao_winter"
  } else if (ascyear == year & ascmonth == month  & ascvar1 == naovar & ascvar2 == samplevar){
    nao_sample <- ascfile[[1]]
    names(nao_sample) <- "nao_sample"

  }
   # st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
  }
#st19980186.9
```

### 2019-02-24 src/stack_test.Rmd
...

@@
+This works
+
+```{r}
+ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+rasters <- lapply(paste0(ascname), raster)
+
+year <- "X1998"
+month <- "01"
+depth <- "86.9294"
+tempvar <- "temp"
+chlvar <- "chl"
+o2var <- "o2"
+salvar <- "salinity"
+depthvar <- "depth"
+surfacevar <- "surface"
+naovar <- "nao"
+amovar <- "amo"
+samplevar <- "sample"
+wintervar <- "winter"
+
+for (i in 1:length(rasters)){
+  ascyear <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 1))
+  ascmonth <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 2))
+  ascdepth <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 5))
+  ascvar1 <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 3))
+  ascvar2 <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 4))
+  ascfile <- rasters[i]
+
+    if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
+    #print(ascfile)
+    temp_depth <- ascfile[[1]]
+    names(temp_depth) <- "temp_depth"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
+    chl_surface <- ascfile[[1]]
+    names(chl_surface) <- "chl_surface"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
+    o2_depth <- ascfile[[1]]
+    names(o2_depth) <- "o2_depth"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
+    salinity_depth <- ascfile[[1]]
+    names(salinity_depth) <- "salinity_depth"
+  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == wintervar){
+    nao_winter <- ascfile[[1]]
+    names(nao_winter) <- "nao_winter"
+  } else if (ascyear == year & ascmonth == month  & ascvar1 == naovar & ascvar2 == samplevar){
+    nao_sample <- ascfile[[1]]
+    names(nao_sample) <- "nao_sample"
+
+  }
+   st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+  }
+st19980186.9
+```
+
 messing with code to see if can do for unique year month depth
+
+need tidyr package
 ```{r}
+unyears <- as.character(c(1992, 1998)) 
+x <- "X"
+unyears <- paste0(x, unyears, collapse = NULL) #this adds an X to each year to match with raster naming
+unmonths <- c(01, 03)
+undepths <- c(10, 86.9294)
+
+unlayer <- crossing(unyears, unmonths, undepths)
+
+for (uny in unyears){
+  for (unm in unmonths){
+    for (und in undepths){
+      year <- unyears[uny]
+      month <- unmonths[unm]
+      depth <- undepths[und]
+    }
+  }
+}
+
 ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
@@
 #st19980186.9
-```

### 2019-02-25 src/stack_test.Rmd
maxent test run

@@
 ```{r}
-unyears <- as.character(c(1992, 1998)) 
-x <- "X"
-unyears <- paste0(x, unyears, collapse = NULL) #this adds an X to each year to match with raster naming
-unmonths <- c(01, 03)
-undepths <- c(10, 86.9294)
+unyears <- as.character(c("1992", "1998")) 
+unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
+unmonths <- c("01", "03")
+undepths <- c("10", "86.9294")
 
 unlayer <- crossing(unyears, unmonths, undepths)
-
-for (uny in unyears){
-  for (unm in unmonths){
-    for (und in undepths){
-      year <- unyears[uny]
-      month <- unmonths[unm]
-      depth <- undepths[und]
-    }
-  }
-}
+unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used
 
@@
 
-year <- "X1998"
-month <- "01"
-depth <- "86.9294"
+#year <- "X1998"
+#month <- "01"
+#depth <- "86.9294"
 tempvar <- "temp"
@@
 
+for (k in 1:nrow(unlayer)){
+    print(unlayer$stackname[k])
+    year <- unlayer$unyears[k]
+    month <- unlayer$unmonths[k]
+    depth <- unlayer$undepths[k]
     for (i in 1:length(rasters)){
@@
     ascvar2 <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 4))
-  ascfile <- rasters[i]
+    ascfile <- rasters[[i]]
 
@@
     names(nao_sample) <- "nao_sample"
+  }
+}
+  }
 
-  }
-   # st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
-  }
+test <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+
+#stacklist <- as.list(unlayer$stackname)
+#stacklist[2] <- temp_depth
+
 #st19980186.9
+
+#stacktest <- read.csv("../output/bio/stackttest.csv", header = TRUE)
+#stacktest$chl <- as.character(stacktest$chl)
+#stacktest
+
+```
+
+```{r}
+var <- raster(stacktest$chl[1])
+var
+a <- raster(var)
+a
+```
+
+#maxent
+maxnet throwing errors...
+```{r}
+janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
+janpres$occ <- 1
+janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
+janab$occ <- 0
+prab <- rbind(janpres, janab)
+prab$originalscientificname <- "cap" #because species name needs to be the same
+pa <- prab$occ
+prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
+prab
+
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
+pre1 <- predict(mod1, test,  args='outputformat=cloglog')
+plot(pre1)
+plot(mod1)
+
+```

### 2019-02-26 src/stack_test.Rmd
stack predict loop done

--- 2019-02-24 src/stack_test.Rmd
@@
-messing with code to see if can do for unique year month depth
+This works
+
 ```{r}
@@
   }
-   # st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+   st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
   }
-#st19980186.9
+st19980186.9
 ```
+
+messing with code to see if can do for unique year month depth
+
+need tidyr package
+```{r}
+unyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
+#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
+unmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
+undepths <- c("0.505760014", "1.55585527", "2.66768169", "3.85627985" , "5.14036131", "6.5430336", "8.09251881", "9.82275009", "11.7736797" , "13.9910383", "16.525322" , "19.4298019", "22.757616", "26.5583", "30.8745613", "35.7402039", "41.1800232", "47.211895", "53.8506355", "61.1128387" , "69.0216827" , "77.6111603", "86.9294281", "97.0413132", "108.030281", "120", "133.075821", "147.40625", "163.164459" , "180.549927", "199.789963", "221.141174", "244.890625" , "271.356384", "300.887512", "333.862823" , "370.688477", "411.793854", "457.62561", "508.639893", "565.292297", "628.026001", "697.258667", "773.368286", "856.678955", "947.447876", "1045.85425", "1151.99121", "1265.86145", "1387.37695", "1045.854", "1151.99097", "1265.86096") #this is biomer + the 3 glorys layers that are different...
+but that's not the best way because you are creating extra depth rows here...
+actually recommend you rename the asc files in windows powershell to match
+
+unlayer <- crossing(unyears, unmonths, undepths)
+unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used
+unlayer$temp_depth <- NA
+unlayer$temp_surface <- NA
+unlayer$chl_depth <- NA
+unlayer$chl_surface <- NA
+unlayer$o2_depth <- NA
+unlayer$o2_surface <- NA
+unlayer$salinity_depth <- NA
+unlayer$salinity_surface <- NA
+unlayer$nao_sample <- NA
+unlayer$nao_prev <- NA
+unlayer$nao_winter <- NA
+unlayer$amo_sample <- NA
+unlayer$amo_prev <- NA
+unlayer$amo_winter <- NA
+
+ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+year <- "X1998"
+month <- "01"
+depth <- "86.9294"
+tempvar <- "temp"
+chlvar <- "chl"
+o2var <- "o2"
+salvar <- "salinity"
+depthvar <- "depth"
+surfacevar <- "surface"
+naovar <- "nao"
+amovar <- "amo"
+samplevar <- "sample"
+wintervar <- "winter"
+
+for (k in 1:nrow(unlayer)){
+    print(unlayer$stackname[k])
+    year <- unlayer$unyears[k]
+    month <- unlayer$unmonths[k]
+    depth <- unlayer$undepths[k]
+    for (i in 1:length(ascname)){
+
+    ascyear <- (sapply(strsplit(ascname[[i]], "_"), "[[", 1))
+    ascmonth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 2))
+    ascdepth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 5))
+    ascdepth <- gsub(".asc", "", ascdepth)
+    ascvar1 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 3))
+    ascvar2 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 4))
+    ascfile <- ascpath[[i]]
+    print(ascfile)
+
+    if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
+    unlayer$temp_depth[k] <- ascfile
+  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
+   unlayer$chl_surface[k] <- ascfile
+ } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
+  unlayer$o2_depth[k] <- ascfile
+ } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
+  unlayer$salinity_depth[k] <- ascfile
+ } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == wintervar){
+  unlayer$nao_winter[k] <- ascfile
+ } else if (ascyear == year & ascmonth == month  & ascvar1 == naovar & ascvar2 == samplevar){
+  unlayer$nao_sample[k] <- ascfile
+ }
+
+    }
+    write.csv(unlayer, "../data/env/asc_layerlist.csv", row.names = TRUE)
+}
+
+# now loop to create stacks
+unlayer <- read.csv("../data/env/asc_layerlist.csv", header = TRUE)
+unlayer[] <- lapply(unlayer, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+
+for (td in 1:nrow(unlayer)){
+  temp_depth <- raster(unlayer$temp_depth[td])
+  names(temp_depth) <- "temp_depth"
+  chl_surface <- raster(unlayer$chl_surface[td])
+  names(chl_surface) <- "chl_surface"
+  o2_depth <- raster(unlayer$o2_depth[td])
+  names(o2_depth) <- "o2_depth"
+  salinity_depth <- raster(unlayer$salinity_depth[td])
+  names(salinity_depth) <- "salinity_depth"
+  nao_sample <- raster(unlayer$nao_sample[td])
+  names(nao_sample) <- "nao_sample"
+  nao_winter <- raster(unlayer$nao_winter[td])
+  names(nao_winter) <- "nao_winter"
+  stackn <- unlayer$stackname[td]
+  test <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+  #stackSave(test,unlayer$stackname[td])
+  pred <- predict(mod1, test, file = paste0("../output/maxentpredict/", stackn), overwrite=TRUE) 
+  preplot <- plot(pred, png(paste0("../output/maxentpredict/", stackn, ".png")))
+  paste("maxplot_", stackn, ".png", sep = "") # to automatically save the plot to a png AND show it inline
+
+}
+dev.off() # stops automatic saving of the plot to a png
+# now call the stack...
+{
+stack <- stackOpen("st19920310.stk")
+}
+```
+
+  preplot <- plot(pred, png(paste0(stackn, ".png")))
+
+```{r}
+var <- raster(stacktest$chl[1])
+var
+a <- raster(var)
+a
+```
+
+#maxent
+maxnet throwing errors...
+```{r}
+janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
+janpres$occ <- 1
+janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
+janab$occ <- 0
+prab <- rbind(janpres, janab)
+prab$originalscientificname <- "cap" #because species name needs to be the same
+pa <- prab$occ
+prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
+prab
+
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
+pre1 <- predict(mod1, test) #,  args='outputformat=cloglog'
+plot(pre1)
+plot(mod1)
+
+plot(pre1)
+```

### 2019-02-27 src/arcpy_amowinter_asciiloop.py
batch ascii creation

@@ -1,0 +1,50 @@
+# -*- coding: utf-8 -*-
+# ---------------------------------------------------------------------------
+# arcpy_amowinter_asciiloop.py
+# Created on: 2019-02-27 12:15:51.00000
+#   (generated by ArcGIS/ModelBuilder)
+# Description: 
+# ---------------------------------------------------------------------------
+
+# Import arcpy module
+import arcpy
+
+# Load required toolboxes
+arcpy.ImportToolbox("Model Functions")
+
+
+# Local variables:
+ssh_Layer1_tif = "C:\\Users\\bunny\\Documents\\gdrive\\ArcGIS\\ssh_Layer1.tif"
+amo_winter_csv = "C:\\Users\\bunny\\Documents\\GitHub\\Chapter_1\\output\\env\\amo_winter.csv"
+nao_winter_dbf = amo_winter_csv
+ArcGIS = "C:\\Users\\bunny\\Documents\\gdrive\\ArcGIS"
+I_nao_winter_year = nao_winter_dbf
+year = I_nao_winter_year
+month = I_nao_winter_year
+winter_val__2_ = I_nao_winter_year
+conout = "C:\\Users\\bunny\\Documents\\ArcGIS\\Default.gdb\\conout"
+v_year___month__amo_winter_depth__asc = "C:\\Users\\bunny\\Documents\\gdrive\\ArcGIS\\climate_ascii\\%year%_%month%_amo_winter%depth%.asc"
+Value = "2018"
+depth = "_01"
+
+# Process: Table to Table
+arcpy.TableToTable_conversion(amo_winter_csv, ArcGIS, "nao_winter.dbf", "", "year \"year\" true true false 4 Long 0 0 ,First,#,C:\\Users\\bunny\\Documents\\GitHub\\Chapter_1\\output\\env\\amo_winter.csv,year,-1,-1;WinterAvg \"WinterAvg\" true true false 8 Double 0 0 ,First,#,C:\\Users\\bunny\\Documents\\GitHub\\Chapter_1\\output\\env\\amo_winter.csv,WinterAvg,-1,-1;WinterAvgS \"WinterAvgShifted\" true true false 8 Double 0 0 ,First,#,C:\\Users\\bunny\\Documents\\GitHub\\Chapter_1\\output\\env\\amo_winter.csv,WinterAvgShifted,-1,-1", "")
+
+# Process: Iterate Row Selection
[…]

### 2019-03-05 src/stack_test.Rmd
ascii-file database

--- 2019-02-24 src/stack_test.Rmd
@@
 This works
+
+need raster rgdal netcdf4 tidyr
 
@@
 ```{r}
-unyears <- as.character(c(1992, 1998)) 
-x <- "X"
-unyears <- paste0(x, unyears, collapse = NULL) #this adds an X to each year to match with raster naming
-unmonths <- c(01, 03)
-undepths <- c(10, 86.9294)
+unyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
+#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
+unmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
+undepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38") #this list is based on how the files are named when split into depth layers.
+note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers
 
 unlayer <- crossing(unyears, unmonths, undepths)
+unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used
+unlayer$temp_depth <- NA
+unlayer$temp_surface <- NA
+unlayer$chl_depth <- NA
+unlayer$chl_surface <- NA
+unlayer$o2_depth <- NA
+unlayer$o2_surface <- NA
+unlayer$salinity_depth <- NA
+unlayer$salinity_surface <- NA
+unlayer$nao_sample <- NA
+unlayer$nao_prev <- NA
+unlayer$nao_winter <- NA
+unlayer$amo_sample <- NA
+unlayer$amo_prev <- NA
+unlayer$amo_winter <- NA
 
-for (uny in unyears){
-  for (unm in unmonths){
-    for (und in undepths){
-      year <- unyears[uny]
-      month <- unmonths[unm]
-      depth <- undepths[und]
-    }
-  }
-}
+ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
 
-ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
-rasters <- lapply(paste0(ascname), raster)
-
-year <- "X1998"
-month <- "01"
-depth <- "86.9294"
+#year <- "X1998"
+#month <- "01"
+#depth <- "86.9294"
 tempvar <- "temp"
@@
 
-for (i in 1:length(rasters)){
-  ascyear <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 1))
-  ascmonth <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 2))
-  ascdepth <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 5))
-  ascvar1 <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 3))
-  ascvar2 <- (sapply(strsplit(names(rasters[[i]]), "_"), "[[", 4))
-  ascfile <- rasters[i]
+for (k in 1:nrow(unlayer)){
+    print(unlayer$stackname[k])
+    year <- unlayer$unyears[k]
+    month <- unlayer$unmonths[k]
+    depth <- unlayer$undepths[k]
+    for (i in 1:length(ascname)){
+
+    ascyear <- (sapply(strsplit(ascname[[i]], "_"), "[[", 1))
+    ascmonth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 2))
+    ascdepth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 5))
+    ascdepth <- gsub(".asc", "", ascdepth)
+    ascvar1 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 3))
+    ascvar2 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 4))
+    ascfile <- ascpath[[i]]
+    print(ascfile)
 
     if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
-    #print(ascfile)
-    temp_depth <- ascfile[[1]]
-    names(temp_depth) <- "temp_depth"
+    unlayer$temp_depth[k] <- ascfile
   } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
-    chl_surface <- ascfile[[1]]
-    names(chl_surface) <- "chl_surface"
+   unlayer$chl_surface[k] <- ascfile
  } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
-    o2_depth <- ascfile[[1]]
-    names(o2_depth) <- "o2_depth"
+  unlayer$o2_depth[k] <- ascfile
  } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
-    salinity_depth <- ascfile[[1]]
-    names(salinity_depth) <- "salinity_depth"
-  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == wintervar){
-    nao_winter <- ascfile[[1]]
-    names(nao_winter) <- "nao_winter"
+  unlayer$salinity_depth[k] <- ascfile
+ } else if (ascyear == year & ascvar1 == naovar & ascvar2 == wintervar){
+  unlayer$nao_winter[k] <- ascfile
  } else if (ascyear == year & ascmonth == month  & ascvar1 == naovar & ascvar2 == samplevar){
-    nao_sample <- ascfile[[1]]
-    names(nao_sample) <- "nao_sample"
+  unlayer$nao_sample[k] <- ascfile
+ }
 
     }
-   # st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+    write.csv(unlayer, "../data/env/asc_layerlist.csv", row.names = TRUE)
 }
-#st19980186.9
+
+# now loop to create stacks
+unlayer <- read.csv("../data/env/asc_layerlist.csv", header = TRUE)
+unlayer[] <- lapply(unlayer, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+aea <- raster("../output/env/aea.tif")
+
+for (td in 1:nrow(unlayer)){
+  temp_depth <- raster(unlayer$temp_depth[td])
+  #crs(temp_depth) <- +proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 
+  names(temp_depth) <- "temp_depth"
+  #temp_depth <- projectRaster(temp_depth, aea)
+  chl_surface <- raster(unlayer$chl_surface[td])
+  names(chl_surface) <- "chl_surface"
+  o2_depth <- raster(unlayer$o2_depth[td])
+  names(o2_depth) <- "o2_depth"
+  salinity_depth <- raster(unlayer$salinity_depth[td])
+  names(salinity_depth) <- "salinity_depth"
+  nao_sample <- raster(unlayer$nao_sample[td])
+  names(nao_sample) <- "nao_sample"
+  nao_winter <- raster(unlayer$nao_winter[td])
+  names(nao_winter) <- "nao_winter"
+  stackn <- unlayer$stackname[td]
+  test <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+  #stackSave(test,unlayer$stackname[td])
+  pred <- predict(mod1, test, file = paste0("../output/maxentpredict/", stackn), overwrite=TRUE) 
+  preplot <- plot(pred, png(paste0("../output/maxentpredict/", stackn, ".png")))
+  paste("maxplot_", stackn, ".png", sep = "") # to automatically save the plot to a png AND show it inline
+
+}
+dev.off() # stops automatic saving of the plot to a png
+# now call the stack...
+{
+stack <- stackOpen("st19920310.stk")
+}
+```
+
+  preplot <- plot(pred, png(paste0(stackn, ".png")))
+
+```{r}
+var <- raster(stacktest$chl[1])
+var
+a <- raster(var)
+a
+```
+
+#maxent
+maxnet throwing errors...
+```{r}
+janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
+janpres$occ <- 1
+janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
+janab$occ <- 0
+prab <- rbind(janpres, janab)
+prab$originalscientificname <- "cap" #because species name needs to be the same
+pa <- prab$occ
+prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
+prab
+
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
+pre1 <- predict(mod1, test) #,  args='outputformat=cloglog'
+plot(pre1)
+plot(mod1)
+
+o2_depth <- raster("../data/jan_test_86.9/2005_01_o2.nc_0.50576.asc")
+names(o2_depth) <- "o2_depth"
+
+test2 <- stack(temp_depth, salinity_depth, o2_depth, chl_surface)
+prab2 <- subset(prab, select = -c(nao_sample, nao_winter)) #ok so for maxent in r just need the variables
+mod2 <- maxent(x = prab2, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
+pre2 <- predict(mod2, test2)
+plot(pre2)
+plot(pre1)
+```
+
+```{r}
+unyears <- as.character(c("1992", "1998")) 
+unmonths <- c("01", "03")
+undepths <- c("10", "86.9294")
+unlayer <- crossing(unyears, unmonths, undepths)
+unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used
+
+unlayer$temp_depth <- NA
+unlayer$temp_surface <- NA
+unlayer$chl_depth <- NA
+unlayer$chl_surface <- NA
+unlayer$o2_depth <- NA
+unlayer$o2_surface <- NA
+unlayer$salinity_depth <- NA
+unlayer$salinity_surface <- NA
+unlayer$nao_sample <- NA
+unlayer$nao_prev <- NA
+unlayer$nao_winter <- NA
+unlayer$amo_sample <- NA
+unlayer$amo_prev <- NA
+unlayer$amo_winter <- NA
+
+tempvar <- "temp"
+chlvar <- "chl"
+o2var <- "o2"
+salvar <- "salinity"
+depthvar <- "depth"
+surfacevar <- "0.50576"
+naovar <- "nao"
+amovar <- "amo"
+samplevar <- "sample"
+wintervar <- "winter"
+
+ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+for (k in 1:nrow(unlayer)){
+    year <- unlayer$unyears[k]
+    month <- unlayer$unmonths[k]
+    depth <- unlayer$undepths[k]
+
+    for (i in 1:length(ascname)){
+    ascyear <- (sapply(strsplit(ascname[[i]], "_"), "[[", 1))
+    ascmonth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 2))
+    ascvar <- (sapply(strsplit(ascname[[i]], "_"), "[[", 3))
+    ascdepth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 4))
+    ascdepth <- gsub(".asc", "", ascdepth)
+    ascfile <- ascpath[[i]]
+    print(ascfile)
+    if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+      unlayer$temp_depth[k] <- ascfile
+    } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+        unlayer$temp_surface[i] <- ascfile
+    }
+    }
+}
+
+```

### 2019-03-05 src/asc_file_database.Rmd
ascii file update

---
title: "asc_file_database"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Preperation database to extract and stack asc files

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there are probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.

```{r pre-install packages, message=FALSE}
packages <- c("tidyr")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

```{r}
ascyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
ascmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
ascdepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38") 
#this depth list is based on how the files are named when split into depth layers. note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers

asclayer <- crossing(ascyears, ascmonths, ascdepths)
asclayer$stackname <- paste0("st", asclayer$ascyears, asclayer$ascmonths, asclayer$ascdepths) #this creates the name of the stack that will be used
asclayer$temp_depth <- NA
asclayer$temp_surface <- NA
asclayer$chl_depth <- NA
asclayer$chl_surface <- NA
asclayer$o2_depth <- NA
asclayer$o2_surface <- NA
asclayer$salinity_depth <- NA
asclayer$salinity_surface <- NA
asclayer$ssh_surface <- NA
asclayer$mlp_surface <- NA
asclayer$nao_sample <- NA
asclayer$nao_prev <- NA
asclayer$nao_winter <- NA
asclayer$amo_sample <- NA
asclayer$amo_prev <- NA
asclayer$amo_winter <- NA

ascname <- list.files("../data/env/ascii/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/env/ascii/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp"
chlvar <- "chl"
o2var <- "o2"
salvar <- "salinity"
mlpvar <- "mlp"
sshvar <- "ssh"
naosamplevar <- "naosample"
amosamplevar <- "amosample"
naoprev <- "naoprev"
amoprev <- "amoprev"
naowinter <- "naowinter"
amowinter <- "amowinter"
climatedepth <- "1"
surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]
    depth <- asclayer$ascdepths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
        ascdepth <- gsub(".asc", "", ascdepth)
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascfile <- ascpath[k]
        print(ascfile)

        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
          asclayer$temp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
          asclayer$salinity_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
          asclayer$chl_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2lvar & ascdepth == depth){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
          asclayer$o2_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
          asclayer$mlp_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
          asclayer$ssh_surface[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naosample){
          asclayer$nao_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amosample){
          asclayer$amo_sample[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naoprev){
          asclayer$nao_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amoprev){
          asclayer$amo_prev[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == naowinter){
          asclayer$nao_winter[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == amowinter){
          asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, "../data/env/asc_layerlist.csv", row.names = TRUE)
}
```

### 2019-03-07 src/asc_file_database.Rmd
arghhhhh!!!

@@
 
-asclayer <- crossing(ascyears, ascmonths, ascdepths)
-asclayer$stackname <- paste0("st", asclayer$ascyears, asclayer$ascmonths, asclayer$ascdepths) #this creates the name of the stack that will be used
-asclayer$temp_depth <- NA
-asclayer$temp_surface <- NA
-asclayer$chl_depth <- NA
-asclayer$chl_surface <- NA
-asclayer$o2_depth <- NA
-asclayer$o2_surface <- NA
-asclayer$salinity_depth <- NA
-asclayer$salinity_surface <- NA
-asclayer$ssh_surface <- NA
-asclayer$mlp_surface <- NA
-asclayer$nao_sample <- NA
-asclayer$nao_prev <- NA
-asclayer$nao_winter <- NA
-asclayer$amo_sample <- NA
-asclayer$amo_prev <- NA
-asclayer$amo_winter <- NA
+asclayerall <- crossing(ascyears, ascmonths, ascdepths)
+asclayerall$stackname <- paste0("st", asclayerall$ascyears, asclayerall$ascmonths, asclayerall$ascdepths) #this creates the name of the stack that will be used
+asclayerall$temp_depth <- NA
+asclayerall$temp_surface <- NA
+asclayerall$chl_depth <- NA
+asclayerall$chl_surface <- NA
+asclayerall$o2_depth <- NA
+asclayerall$o2_surface <- NA
+asclayerall$salinity_depth <- NA
+asclayerall$salinity_surface <- NA
+asclayerall$ssh_surface <- NA
+asclayerall$mlp_surface <- NA
+asclayerall$nao_sample <- NA
+asclayerall$nao_prev <- NA
+asclayerall$nao_winter <- NA
+asclayerall$amo_sample <- NA
+asclayerall$amo_prev <- NA
+asclayerall$amo_winter <- NA
 
-ascname <- list.files("../data/env/ascii/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
-ascpath <- list.files("../data/env/ascii/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+```
 
-tempvar <- "temp"
-chlvar <- "chl"
-o2var <- "o2"
-salvar <- "salinity"
-mlpvar <- "mlp"
-sshvar <- "ssh"
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 1998)
+
+ascname <- list.files("../data/env/ascii/ascii1998/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii1998/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
 naosamplevar <- "naosample"
 amosamplevar <- "amosample"
-naoprev <- "naoprev"
-amoprev <- "amoprev"
-naowinter <- "naowinter"
-amowinter <- "amowinter"
-climatedepth <- "1"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
 surfacevar <- "0.50576"
@@
         ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
         ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
         ascdepth <- gsub(".asc", "", ascdepth)
-        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
         ascfile <- ascpath[k]
-        print(ascfile)
 
@@
         } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
-          asclayer$temp_depth[i] <- ascfile
+          asclayer$salinity_depth[i] <- ascfile
         } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
@@
           asclayer$chl_surface[i] <- ascfile
-        } else if(ascyear == year & ascmonth == month & ascvar == o2lvar & ascdepth == depth){
-          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
         } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
@@
           asclayer$ssh_surface[i] <- ascfile
-        } else if(ascyear == year & ascmonth == month & ascvar == naosample){
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
           asclayer$nao_sample[i] <- ascfile
-        } else if(ascyear == year & ascmonth == month & ascvar == amosample){
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
           asclayer$amo_sample[i] <- ascfile
-        } else if(ascyear == year & ascmonth == month & ascvar == naoprev){
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
           asclayer$nao_prev[i] <- ascfile
-        } else if(ascyear == year & ascmonth == month & ascvar == amoprev){
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
           asclayer$amo_prev[i] <- ascfile
-        } else if(ascyear == year & ascmonth == month & ascvar == naowinter){
+        } else if(ascyear == year & ascvar == naowintervar){
           asclayer$nao_winter[i] <- ascfile
-        } else if(ascyear == year & ascmonth == month & ascvar == amowinter){
+        } else if(ascyear == year & ascvar == amowintervar){
           asclayer$amo_winter[i] <- ascfile
@@
     }
-  write.csv(asclayer, "../data/env/asc_layerlist.csv", row.names = TRUE)
+  write.csv(asclayer, "../data/env/asc_layerlist1998.csv", row.names = TRUE)
 }
 ```
+
+```{r}
+#!!!!!!  CHANGE ASCLAYER ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 1999)
+
+ascname <- list.files("../data/env/ascii/ascii1999/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii1999/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist1999.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2000)
+
+ascname <- list.files("../data/env/ascii/ascii2000/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2000/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2000.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2001)
+
+ascname <- list.files("../data/env/ascii/ascii2001/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2001/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2001.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2002)
+
+ascname <- list.files("../data/env/ascii/ascii2002/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2002/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2002.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2003)
+
+ascname <- list.files("../data/env/ascii/ascii2003/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2003/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2003.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2004)
+
+ascname <- list.files("../data/env/ascii/ascii2004/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2004/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2004.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2005)
+
+ascname <- list.files("../data/env/ascii/ascii2005/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2005/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2005.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2006)
+
+ascname <- list.files("../data/env/ascii/ascii2006/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2006/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2006.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2007)
+
+ascname <- list.files("../data/env/ascii/ascii2007/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2007/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2007.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2008)
+
+ascname <- list.files("../data/env/ascii/ascii2008/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2008/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2008.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2009)
+
+ascname <- list.files("../data/env/ascii/ascii2009/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2009/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2009.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2010)
+
+ascname <- list.files("../data/env/ascii/ascii2010/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2010/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2010.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2011)
+
+ascname <- list.files("../data/env/ascii/ascii2011/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2011/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2011.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2012)
+
+ascname <- list.files("../data/env/ascii/ascii2012/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2012/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2012.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2103)
+
+ascname <- list.files("../data/env/ascii/ascii2013/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2013/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2013.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2014)
+
+ascname <- list.files("../data/env/ascii/ascii2014/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2014/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2014.csv", row.names = TRUE)
+}
+```
+
+```{r}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2015)
+
+ascname <- list.files("../data/env/ascii/ascii2015/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
+ascpath <- list.files("../data/env/ascii/ascii2015/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
+
+tempvar <- "temp.nc"
+chlvar <- "chl.nc"
+o2var <- "o2.nc"
+salvar <- "salinity.nc"
+mlpvar <- "mlp.nc"
+sshvar <- "ssh.nc"
+naosamplevar <- "naosample"
+amosamplevar <- "amosample"
+naoprevvar <- "naoprev"
+amoprevvar <- "amoprev"
+naowintervar <- "naowinter"
+amowintervar <- "amowinter"
+surfacevar <- "0.50576"
+
+for (i in 1:nrow(asclayer)){
+    print(asclayer$stackname[i])
+    year <- asclayer$ascyears[i]
+    month <- asclayer$ascmonths[i]
+    depth <- asclayer$ascdepths[i]
+
+    for (k in 1:length(ascname)){
+        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
+        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
+        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
+        ascdepth <- (sapply(strsplit(ascname[k], "_"), "[[", 4))
+        ascdepth <- gsub(".asc", "", ascdepth)
+        ascfile <- ascpath[k]
+
+        if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
+          asclayer$temp_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
+          asclayer$temp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == depth){
+          asclayer$salinity_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == salvar & ascdepth == surfacevar){
+          asclayer$salinity_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == depth){
+          asclayer$chl_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == chlvar & ascdepth == surfacevar){
+          asclayer$chl_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == depth){
+          asclayer$o2_depth[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == o2var & ascdepth == surfacevar){
+          asclayer$o2_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == mlpvar){
+          asclayer$mlp_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == sshvar){
+          asclayer$ssh_surface[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+          asclayer$nao_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+          asclayer$amo_sample[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+          asclayer$nao_prev[i] <- ascfile
+        } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
+          asclayer$amo_prev[i] <- ascfile
+        } else if(ascyear == year & ascvar == naowintervar){
+          asclayer$nao_winter[i] <- ascfile
+        } else if(ascyear == year & ascvar == amowintervar){
+          asclayer$amo_winter[i] <- ascfile
+        }
+    }
+  write.csv(asclayer, "../data/env/asc_layerlist2015.csv", row.names = TRUE)
+}
+```

### 2019-03-07 src/data_exploration.Rmd
stack predict loop done
arghhhhh!!!

@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car", "rworldmap", "fastmatch") 
+packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car", "rworldmap", "fastmatch", "tables") 
 source("../src/ipak.R")
@@
 
+and by year
+
+```{r}
+gearby_yr <- with(presence, table(gear, year))
+write.csv(gearby_yr, file = "../output/bio/gear_by_yr.csv")
+gearby_yr
+```
+
 What I want to see if if there is a marked difference between the env.
@@
 ```{r}
-```{r}
 system.file("java", package="dismo") #finding location to put maxent jar file

### 2019-03-11 src/stack_test.Rmd
looping model and plot predictions (note some files coming as empty .png)

@@
 # now loop to create stacks
-unlayer <- read.csv("../data/env/asc_layerlist.csv", header = TRUE)
+janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
+janpres$occ <- 1
+janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
+janab$occ <- 0
+prab <- rbind(janpres, janab)
+prab$originalscientificname <- "cap" #because species name needs to be the same
+pa <- prab$occ
+prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
+prab
+
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
+pre1 <- predict(mod1, test) #,  args='outputformat=cloglog'
+plot(pre1)
+plot(mod1)
+
+unlayer <- read.csv("../data/env/asc_layerlist1998.csv", header = TRUE)
+unlayer <- subset(unlayer, ascmonths == 1)
+unlayer <- unlayer[-c(1), ]
 unlayer[] <- lapply(unlayer, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
@@
 # now call the stack...
-{
-stack <- stackOpen("st19920310.stk")
-}
+
 ```

### 2019-03-11 src/stack_test.Rmd
stack loop prediction complete. model fiddling now

@@
 ```
+
+```{r}
+# now loop to create stacks
+janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
+janpres$occ <- 1
+janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
+janab$occ <- 0
+prab <- rbind(janpres, janab)
+prab$originalscientificname <- "cap" #because species name needs to be the same
+pa <- prab$occ
+prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
+prab
+
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
+
+unlayer <- read.csv("../data/env/asc_layerlist1998.csv", header = TRUE)
+unlayer <- subset(unlayer, ascmonths == 1)
+unlayer <- unlayer[-c(1), ]
+unlayer[] <- lapply(unlayer, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+aea <- raster("../output/env/aea.tif")
+
+for (td in 1:nrow(unlayer)){
+  temp_depth <- raster(unlayer$temp_depth[td])
+  crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(temp_depth) <- "temp_depth"
+  temp_depth <- projectRaster(temp_depth, aea)
+  chl_surface <- raster(unlayer$chl_surface[td])
+  crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(chl_surface) <- "chl_surface"
+  chl_surface <- projectRaster(chl_surface, aea)
+  o2_depth <- raster(unlayer$o2_depth[td])
+  crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(o2_depth) <- "o2_depth"
+  o2_depth <- projectRaster(o2_depth, aea)
+  salinity_depth <- raster(unlayer$salinity_depth[td])
+  crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(salinity_depth) <- "salinity_depth"
+  salinity_depth <- projectRaster(salinity_depth, aea)
+  nao_sample <- raster(unlayer$nao_sample[td])
+  crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(nao_sample) <- "nao_sample"
+  nao_sample <- projectRaster(nao_sample, aea)
+  nao_winter <- raster(unlayer$nao_winter[td])
+  crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+  names(nao_winter) <- "nao_winter"
+  nao_winter <- projectRaster(nao_winter, aea)
+  stackn <- unlayer$stackname[td]
+  test <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+  #stackSave(test,unlayer$stackname[td])
+  pred <- predict(mod1, test, file = paste0("../output/maxentpredict/", stackn), overwrite=TRUE) 
+  paste("maxplot_", stackn, ".png", sep = "") # to automatically save the plot to a png AND show it inline
+  preplot <- plot(pred, png(paste0("../output/maxentpredict/", stackn, ".png")))
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+```

### 2019-03-12 src/stack_test.Rmd
minor change

@@
 ```
+
+testing model fiddling
+
+```{r}
+janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
+janpres$occ <- 1
+janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
+janab$occ <- 0
+prab <- rbind(janpres, janab)
+prab$originalscientificname <- "cap" #because species name needs to be the same
+pa <- prab$occ
+prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
+prab
+
+test <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+mod1 <- maxent(x = prab, p = pa)
+
+pre1 <- predict(mod1, test, args='outputformat=raw' ) #, args = c("outputformat=cloglog")
+preplot <- plot(pre1)
+preplot
+
+```

### 2019-03-13 src/stack_test.Rmd
model tweaking

--- 2019-03-11 src/stack_test.Rmd
@@
 ```
+
+testing model fiddling
+
+```{r}
+janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
+janpres$occ <- 1
+janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
+janab$occ <- 0
+prab <- rbind(janpres, janab)
+prab$originalscientificname <- "cap" #because species name needs to be the same
+pa <- prab$occ
+prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
+prab
+#write.csv(prab, "../output/bio/maxjan98test.csv", row.names = FALSE)
+
+stacklist <- read.csv("../data/env/asc_layerlist1998.csv", header = TRUE)
+stacklist <- subset(stacklist, stackname == "st199801000120")
+stacklist[] <- lapply(stacklist, as.character)
+stacklist
+aea <- raster("../output/env/aea.tif")
+
+temp_depth <- raster(stacklist$temp_depth)
+crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(temp_depth) <- "temp_depth"
+temp_depth <- projectRaster(temp_depth, aea)
+chl_surface <- raster(stacklist$chl_surface)
+crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(chl_surface) <- "chl_surface"
+chl_surface <- projectRaster(chl_surface, aea)
+o2_depth <- raster(stacklist$o2_depth)
+crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(o2_depth) <- "o2_depth"
+o2_depth <- projectRaster(o2_depth, aea)
+salinity_depth <- raster(stacklist$salinity_depth)
+crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(salinity_depth) <- "salinity_depth"
+salinity_depth <- projectRaster(salinity_depth, aea)
+nao_sample <- raster(stacklist$nao_sample)
+crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(nao_sample) <- "nao_sample"
+nao_sample <- projectRaster(nao_sample, aea)
+nao_winter <- raster(stacklist$nao_winter)
+crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+names(nao_winter) <- "nao_winter"
+nao_winter <- projectRaster(nao_winter, aea)
+
+test <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+
+mod1 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000")) #maxiterations increased to ensure model convergence
+mod1
+
+pre1 <- predict(mod1, test, args='outputformat=raw' ) 
+preplot1 <- plot(pre1)
+preplot1
+
+pre2 <- predict(mod1, test, args='outputformat=logistic' ) 
+preplot2 <- plot(pre2)
+preplot2
+
+pre3 <- predict(mod1, test, args='outputformat=cloglog' ) 
+preplot3 <- plot(pre3)
+preplot3
+
+#------------------------
+
+mod2 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "defaultprevalence=0.8")) # changing the prev seems to do nothing
+mod2
+
+mod3 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "defaultprevalence=0.1")) # changing the prev seems to do nothing
+mod3
+
+pre2.2 <- predict(mod2, test, args='outputformat=logistic' ) 
+preplot2.2 <- plot(pre2.2)
+preplot2.2
+
+pre2.3 <- predict(mod3, test, args='outputformat=logistic' ) 
+preplot2.3 <- plot(pre2.3)
+preplot2.3
+
+#-----------------------------
+
+mod4 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE")) 
+mod4
+
+mod5 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "maximumbackground=100")) #because i am using swd, this command is overwridden.  
+mod5
+
+pre5 <- predict(mod5, test, args='outputformat=logistic' ) 
+preplot5 <- plot(pre5)
+preplot5
+```

### 2019-03-14 src/data_exploration.Rmd
playing with maxent variable selection

@@
 
-```{r}
-library(dismo)
-library(rJava)
-jar <- paste(system.file(package="dismo"), "C:/Users/bunny/Documents/GitHub/Chapter_1/packrat/lib/x86_64-w64-mingw32/3.5.1/dismo/java/maxent.jar", sep='')
-file.exists(jar)
-testmax <- maxent(janpresabmax, janpresabmax$occurrence)
-
-```
+# maxentvariableselection package
+
+https://cran.r-project.org/web/packages/MaxentVariableSelection/MaxentVariableSelection.pdf

### 2019-03-14 src/stack_test.Rmd
playing with maxent variable selection

@@
 preplot5
+
+#-----------------------
+
+mod6 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "replicates=5", "replicatetype=crossvalidate")) #replicates with crossvalidate result in multiple map outputs (one for each replicate).
+It is not possible to 'merge' these maps https://stackoverflow.com/questions/28885987/r-how-to-get-one-summary-prediction-map-instead-for-5-when-using-5-fold-cros
+mod6
+
+pre6 <- predict(mod6, test, args='outputformat=logistic' ) 
+preplot6 <- plot(pre6)
+preplot6
+
+#-----------------------
+strt <- Sys.time() #get the start time
+mod7 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4")) #threads should speed things up...
+mod7
+
+pre7 <- predict(mod7, test, args='outputformat=logistic' ) 
+preplot7 <- plot(pre7)
+preplot7
+print(Sys.time()-strt) #time it took to run
+
+#43 secs
+
+strt <- Sys.time() #get the start time
+mod7.2 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE"))
+mod7.2
+
+pre7.2 <- predict(mod7.2, test, args='outputformat=logistic' ) 
+preplot7.2 <- plot(pre7)
+preplot7.2
+print(Sys.time()-strt) #time it took to run
+
+#1.75 mins
+
+#conf that threads does indeed speed up process
+
+#---------------------
+#follwoing radosavljevic 2014
+
+mod8.1 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=0.25")) #playing with multiplier
+mod8.1
+
+pre8.1 <- predict(mod8.1, test, args='outputformat=logistic' ) 
+preplot8.1 <- plot(pre8.1)
+preplot8.1
+
+mod8 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4")) #default setting for betam is 1
+mod8
+
+pre8 <- predict(mod8, test, args='outputformat=logistic' ) 
+preplot8 <- plot(pre8)
+preplot8
+
+mod8.4 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=1.25")) #playing with multiplier
+mod8.4
+
+pre8.4 <- predict(mod8.4, test, args='outputformat=logistic' ) 
+preplot8.4 <- plot(pre8.4)
+preplot8.4
+
+mod8.2 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=2")) 
+mod8.2
+
+pre8.2 <- predict(mod8.2, test, args='outputformat=logistic' ) 
+preplot8.2 <- plot(pre8.2)
+preplot8.2
+
+mod8.3 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=6")) 
+mod8.3
+
+pre8.3 <- predict(mod8.3, test, args='outputformat=logistic' ) 
+preplot8.3 <- plot(pre8.3)
+preplot8.3
+
 ```
+
+one maxent run with all variables
+
+```{r}
+presaball <- read.csv("../output/bio/presab_10k.csv", header = TRUE)
+janpresaball <- subset(presaball, month == 1)
+janpresaball <- subset(janpresaball, select = c(occurrence, temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter))
+paall <- janpresaball$occurrence
+janpresaball <- subset(janpresaball, select = -c(occurrence))
+
+stacklist <- read.csv("../data/env/asc_layerlist1998.csv", header = TRUE)
+stacklist <- subset(stacklist, stackname == "st199801000120")
+stacklist[] <- lapply(stacklist, as.character)
+stacklist
+aea <- raster("../output/env/aea.tif")
+
+temp_depth <- raster(stacklist$temp_depth)
+crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(temp_depth) <- "temp_depth"
+temp_depth <- projectRaster(temp_depth, aea)
+temp_surface <- raster(stacklist$temp_surface)
+crs(temp_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(temp_surface) <- "temp_surface"
+temp_surface <- projectRaster(temp_surface, aea)
+chl_depth <- raster(stacklist$chl_depth)
+crs(chl_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(chl_depth) <- "chl_depth"
+chl_depth <- projectRaster(chl_depth, aea)
+chl_surface <- raster(stacklist$chl_surface)
+crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(chl_surface) <- "chl_surface"
+chl_surface <- projectRaster(chl_surface, aea)
+o2_depth <- raster(stacklist$o2_depth)
+crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(o2_depth) <- "o2_depth"
+o2_depth <- projectRaster(o2_depth, aea)
+o2_surface <- raster(stacklist$o2_surface)
+crs(o2_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(o2_surface) <- "o2_surface"
+o2_surface <- projectRaster(o2_surface, aea)
+salinity_depth <- raster(stacklist$salinity_depth)
+crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(salinity_depth) <- "salinity_depth"
+salinity_depth <- projectRaster(salinity_depth, aea)
+salinity_surface <- raster(stacklist$salinity_surface)
+crs(salinity_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(salinity_surface) <- "salinity_surface"
+salinity_surface <- projectRaster(salinity_surface, aea)
+ssh_surface <- raster(stacklist$ssh_surface)
+crs(ssh_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(ssh_surface) <- "ssh_surface"
+ssh_surface <- projectRaster(ssh_surface, aea)
+mlp_surface <- raster(stacklist$mlp_surface)
+crs(mlp_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(mlp_surface) <- "mlp_surface"
+mlp_surface <- projectRaster(mlp_surface, aea)
+nao_sample <- raster(stacklist$nao_sample)
+crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(nao_sample) <- "nao_sample"
+nao_sample <- projectRaster(nao_sample, aea)
+nao_prev <- raster(stacklist$nao_prev)
+crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(nao_prev) <- "nao_prev"
+nao_prev <- projectRaster(nao_prev, aea)
+nao_winter <- raster(stacklist$nao_winter)
+crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+names(nao_winter) <- "nao_winter"
+nao_winter <- projectRaster(nao_winter, aea)
+amo_sample <- raster(stacklist$amo_sample)
+crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(amo_sample) <- "amo_sample"
+amo_sample <- projectRaster(amo_sample, aea)
+amo_prev <- raster(stacklist$amo_prev)
+crs(amo_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(amo_prev) <- "amo_prev"
+amo_prev <- projectRaster(amo_prev, aea)
+amo_winter <- raster(stacklist$amo_winter)
+crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+names(amo_winter) <- "amo_winter"
+amo_winter <- projectRaster(amo_winter, aea)
+
+testall <- stack(temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter)
+
+system.file("java", package="dismo") #finding location to put maxent jar file
+system("java -version")
+
+mod1all <- maxent(x = janpresaball, p = paall, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4")) 
+mod1all
+
+pre1all <- predict(mod1all, testall, args='outputformat=logistic' ) 
+preplot1all <- plot(pre1all)
+preplot1all
+
+```
+
+# maxentvariableselection package
+
+https://cran.r-project.org/web/packages/MaxentVariableSelection/MaxentVariableSelection.pdf + https://cran.r-project.org/web/packages/MaxentVariableSelection/vignettes/MaxentVariableSelection.pdf
+
+```{r}
+janaball <- read.csv("../output/bio/background_10k.csv", header = TRUE)
+janpresall <- read.csv("../output/bio/pres_nomissing.csv", header = TRUE)
+janpresaball <- rbind.fill(janaball, janpresall)
+janpresaball <- subset(janpresaball, month == 1)
+janpresaball <- subset(janpresaball, select = c(occurrence, longitude_meters, latitude_meters, temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter))
+names(janpresaball)[names(janpresaball)=="occurrence"] <- "species"
+names(janpresaball)[names(janpresaball)=="longitude_meters"] <- "longitude"
+names(janpresaball)[names(janpresaball)=="latitude_meters"] <- "latitude"
+janpresall <- subset(janpresaball, species == 1)
+janpresall$species <- replace(janpresall$species, janpresall$species == "1", "fish")
+write.csv(janpresall, "../output/maxentvariableselection/presence.csv", row.names = FALSE)
+janaball <- subset(janpresaball, species == 0)
+janpresall$species <- replace(janpresall$species, janpresall$species == "0", "bg")
+write.csv(janaball, "../output/maxentvariableselection/background.csv", row.names = FALSE)
+
+VariableSelection(maxent="C:/Users/bunny/Downloads/maxent/maxent.jar", outdir=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/output"), gridfolder=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/testgrid"), occurrencelocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/presence.csv"), backgroundlocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/background.csv"), additionalargs="nolinear noquadratic noproduct nothreshold noautofeature", contributionthreshold=5, correlationthreshold=0.9, betamultiplier=seq(0.5, 4, 0.25))
+```

### 2019-03-20 src/SGB_code.Rmd
some maxentvariableselection playing, and update to potential change to background point extraction

---
title: "simon_test_loop"
output: html_notebook
---
-----
This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}
plot(cars) # if you put a # in the code block, it makes the text after a comment!
```

Add a new chunk by clicking the *Insert Chunk* button on the tool-bar or by pressing *Ctrl+Alt+I*. SIMON - THIS DOESN'T WORK SO WELL ON WINDOWS.ON THE TOOLBAR GO TO TOOL - MODIFY KEYBOARD SHORTCUTS - SEARCH FOR INSERT CHUNK AND CHANGE TO WHATEVER YOU WANT (I USE CTRL + ALT + K)

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed

Package dependencies

You can load them using the following code which uses a function called ipak. I've popped the code for 
Note this function checks to see if the packages are installed first.

```{r pre-install packages, message=FALSE}
packages <- c("netcdf", "rgdal", "sp", "raster")
source("./ipak.R")
ipak(packages)
suppressMessages(ipak)
```
---
OK this is fun ...
Hay Samsam I likes this coding format, being able to fully describe your code separately but still linked to the code is great.

I have put away my crampons and rope and I think I survived the learning curve, wow your a good teacher.
I also did some you-tubing and even looked at the first level of an R course but quickly realised the best way to learn for me is to dive in and learn as I go :-

OK I don't know how well this will work but I am going to try and keep this document going as I work out, what it is I am going to be able to do here. 
Haven't decided yet if your going to get to read this or not so the narrative is probably going to jump from first person reminders to comments I think might help you, also the tense is going to jump as I go back and update stuff ...

For me this is a new final target, none business/database, a new coding language (R) and a new data structure (raster files, netcdf) and I'm sure many others I have yet to find out about.

All that said I do have experience in data manipulation and other coding languages, so here goes ...

Look at Sam's original code and especially her loop which she is concerned about.

Basic requirements:
Take a specific netcdf file, make it into a brick, extract values by :-
the variable of the netcdf,
xy location,
layer,
time period

(Maybe some indexing could be done ...)
Then add that data to specific columns on a data-frame

All based on data provided in another file :-
xyz location of the point
layer
(More indexing ...)

Final goals 
Deal with worse-case scenario,
~200,000 points at every month at every year for every variable

Need to look at options for :-
Splitting the csv file and the netcdf files into smaller time periods
Whole process needs to be as fire and forget as possible, if it can't be reduced hugely in time.

Enter the, in Sam's words 'horribly inefficient loop (but works)'

Focus on, it works, it's a great starting point, I could not have got started without something like this!
---
```{r}
start_time <- Sys.time()
netcdf_list <- list.files("./data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("./data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("./data/env/aea.tif") 
for (i in 1:no_netcdf) {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(background_all_sp)) {  
      de <- background_all_sp$depthlayerno[[j]]  # a variable for the observation depth layer
      yr <- (background_all_sp$year[j])  # a variable for the observation year
      mth <- (background_all_sp$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
              background_all_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
              if (is.na(de)){
                background_all_sp$temp_depth[j] <- NA
              } else  
                background_all_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
              background_all_sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
              if (is.na(de)){
                background_all_sp$salinity_depth[j] <- NA
              } else  
                background_all_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
              background_all_sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
              if (is.na(de)){
                background_all_sp$chl_depth[j] <- NA
              } else  
                background_all_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
              background_all_sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ]) 
              if (is.na(de)){
                background_all_sp$o2_depth[j] <- NA
              } else  
                background_all_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
              background_all_sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
              background_all_sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = background_all_sp[j, ])

          }

    }
}
write.csv(background_all_sp, "./output/points_with_data.csv", row.names = FALSE) #this ... writes a csv file so i have a permanent record that I can work with
end_time <- Sys.time()
print(end_time - start_time)
```
---
First thing I think I will do is try to understand what this is doing and why, best way is to do that is try and rewrite it using coding practices I like to use :-
KISS - Keep it simple, stupid
Write as little code as possible, 
try not to repeat code, if you find similar code think about a function that deals with the differences.
Only evaluate stuff once if possible, each evaluation takes processing time where as referring to a memory location or variable takes very little.
---
```{r}
#Basic rework of the existing logic
start_time <- Sys.time()
background_all2 <- read.csv("./data/simon_test_points.csv", header = TRUE)
xy <- background_all2[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. 
background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
head(background_all_sp)

netcdf_list <- list.files("./data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("./data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("./data/env/aea.tif")
print(paste("start time",start_time,sep=" "))
for (i in 1:no_netcdf) {
  netcdf_active <- netcdf_name[i]
  print(netcdf_active) #this just prints the name of the netCDF R is working one
  netcdf_split <- strsplit(netcdf_name[i], "_")
  brkyr <- unlist(netcdf_split)[1] # extracting the first part of the netcdf filename (which is the year)
  brkmth <-unlist(netcdf_split)[2] # extracting the second part of the netcdf filename (which is the month)
  brkvar <- unlist(netcdf_split)[3] # extracting the third part of the netcdf (inc.nc)
  fileIndex = paste(brkyr,brkmth,sep="_")
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea)
  #j <- nrow(background_all_sp)
  #while(j > 0){
  for (j in 1:nrow(background_all_sp)) {
    de <- background_all_sp$depthlayerno[[j]]  # a variable for the observation depth layer
    yr <- (background_all_sp$year[j])  # a variable for the observation year
    mth <- (background_all_sp$month[j])  # a variable for the observation month
    objIndex <- paste(yr,mth,sep="_")
    if(fileIndex == objIndex){
      temp_brick_1 <- extract(x=temp_brick[[1]], y = background_all_sp[j, ])
      if (!(brkvar == "mlp.nc") & !(brkvar == "ssh.nc")){
        if (is.na(de)){
          temp_brick_de <- NA
        } else {
          temp_brick_de <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
        }
      }
      if(brkvar == "temp.nc"){
        background_all_sp$temp_surface[j] <- temp_brick_1
        background_all_sp$temp_depth[j] <-  temp_brick_de
      } else if (brkvar == "salinity.nc") {
        background_all_sp$salinity_surface[j] <- temp_brick_1 
        background_all_sp$salinity_depth[j] <- temp_brick_de
      } else if (brkvar == "chl.nc") {
        background_all_sp$chl_surface[j] <- temp_brick_1   
        background_all_sp$chl_depth[j] <- temp_brick_de 
      } else if (brkvar == "o2.nc") {
        background_all_sp$o2_surface[j] <- temp_brick_1
        background_all_sp$o2_depth[j] <- temp_brick_de
      } else if (brkvar == "mlp.nc") {
        background_all_sp$mlp_surface[j] <- temp_brick_1
      } else if (brkvar == "ssh.nc") {
        background_all_sp$ssh_surface[j] <- temp_brick_1
      }
    }
  }
}
write.csv(background_all_sp, "./output/points_with_data_1.csv", row.names = FALSE) #this ... writes a csv file so i have a permanent record that I can work with
end_time <- Sys.time()
print(end_time - start_time)
```
---
OK hopefully this is still recognisable, basically I have simplified the logic a little and tried to cut down on evaluations by setting data to variables, I have also added some simple time metrics so I can track my progress.
Also added some to Sam's code as there is no point me making it, supposedly better code if it is then slower than what she had.
Also renamed my output files so that I can compare each against the original to check that I am not introducing data errors without realising it.

First run was almost identical to the original but at I have to get to grips with the studio environment.
In most environments I have worked with when you finish running the code the variables and data disappear but here they  persist, will need to work out how to clear unwanted stuff out and to make sure I don't trip up by having a value in a variable that I am not updating any more but relying on in my code!

First thing I noticed is the netcdf files contain allot of data but we only want a small amount so I wonder if I can find a way of only pulling the files that we need and then mapping the data onto the data brick ...
---
```{r}
#Attempt to reverse loops.
start_time <- Sys.time()

background_all2 <- read.csv("./data/simon_test_points.csv", header = TRUE)
xy <- background_all2[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. 
background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
head(background_all_sp)

netcdf_list <- list.files("./data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("./data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("./data/env/aea.tif")

# Loop through required data list rows
for (j in 1:nrow(background_all_sp)) {
    de <- background_all_sp$depthlayerno[[j]]  # a variable for the observation depth layer
    yr <- (background_all_sp$year[j])  # a variable for the observation year
    mth <- (background_all_sp$month[j])  # a variable for the observation month
    objIndex <- paste(yr,mth,sep="_")
	objType <- c("chl.nc","mlp.nc","o2.nc","salinity.nc","ssh.nc","temp.nc")
	intTypeMax <- length(objType)
	# Loop through required data columns
	#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
	# tried while loops but makes little difference
	#intType <- intTypeMax
	#while(intType > 1){
	#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
	for (intType in 1:intTypeMax){
		brkvar <- unlist(objType)[intType]
		netcdf_name_req <- paste(objIndex,brkvar,sep="_")
		print(paste("search for",netcdf_name_req,sep=" "))
		# Check if you have a file for the relevant row and relevant column data
		indexLoc <- which(sapply(netcdf_name, function(e) is.element(netcdf_name_req, e)))
		if (indexLoc > 0){
			# load data
			print(netcdf_name_req)
			temp_brick <- brick(netcdf_list[indexLoc], lvar = 4)
			temp_brick <- projectRaster(temp_brick, aea) 
			temp_brick_1 =  extract(x=temp_brick[[1]], y = background_all_sp[j, ])
			if (!(brkvar == "mlp.nc") & !(brkvar == "ssh.nc")){
				if (is.na(de)){
					temp_brick_de <- NA
				} else {
					temp_brick_de <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
				}
			}
			 if(brkvar == "temp.nc"){
				background_all_sp$temp_surface[j] <- temp_brick_1
				background_all_sp$temp_depth[j] <-  temp_brick_de
			  } else if (brkvar == "salinity.nc") {
				background_all_sp$salinity_surface[j] <- temp_brick_1 
				background_all_sp$salinity_depth[j] <- temp_brick_de
			  } else if (brkvar == "chl.nc") {
				background_all_sp$chl_surface[j] <- temp_brick_1   
				background_all_sp$chl_depth[j] <- temp_brick_de 
			  } else if (brkvar == "o2.nc") {
				background_all_sp$o2_surface[j] <- temp_brick_1
				background_all_sp$o2_depth[j] <- temp_brick_de
			  } else if (brkvar == "mlp.nc") {
				background_all_sp$mlp_surface[j] <- temp_brick_1
			  } else if (brkvar == "ssh.nc") {
				background_all_sp$ssh_surface[j] <- temp_brick_1
			  }
		}
		#intType <- intType - 1 #required for While loop
	}
}
write.csv(background_all_sp, "./output/points_with_data_2.csv", row.names = FALSE) #this ... writes a csv file so i have a permanent record that I can work with
end_time <- Sys.time()
print(end_time - start_time)
```
---
Ouch !!
That took nearly three times as long as the original but at least I now understand why and it has given me a better understanding of the data and the underlying problem.
Very strange to not have a clear idea of the data, I'm used to looking at tables of data on databases this all seems like black box data, need to look more into how to read these netcdf files.

What it showed me is how out of step the two sets of files are ...

Note:
---
I tried converting the for loops to while loops as I found in my Sudoku coding that made a huge difference with large loops, it does very little evaluation when starting a new loop where as a for loop can do loads if your not careful.
But as the loops are so small it made very little difference just left one commented in to show the difference.

---
Went back to previous code and added some more metrics to work out how efficient it is.
---
```{r}
#Basic rework of the existing logic
no_all_netcdf_rows <- 0
no_all_netcdf_used_rows <- 0
start_time <- Sys.time()
background_all2 <- read.csv("./data/simon_test_points.csv", header = TRUE)
xy <- background_all2[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. 
background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
head(background_all_sp)

netcdf_list <- list.files("./data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("./data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("./data/env/aea.tif")
print(paste("start time",start_time,sep=" "))
for (i in 1:no_netcdf) {
  netcdf_active <- netcdf_name[i]
  print(netcdf_active) #this just prints the name of the netCDF R is working one
  netcdf_split <- strsplit(netcdf_name[i], "_")
  brkyr <- unlist(netcdf_split)[1] # extracting the first part of the netcdf filename (which is the year)
  brkmth <-unlist(netcdf_split)[2] # extracting the second part of the netcdf filename (which is the month)
  brkvar <- unlist(netcdf_split)[3] # extracting the third part of the netcdf (inc.nc)
  fileIndex = paste(brkyr,brkmth,sep="_")
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea)
  no_netcdf_rows <- nrow(background_all_sp)
  no_all_netcdf_rows <- no_all_netcdf_rows + no_netcdf_rows
  for (j in 1:no_nedcdf_rows) {
    de <- background_all_sp$depthlayerno[[j]]  # a variable for the observation depth layer
    yr <- (background_all_sp$year[j])  # a variable for the observation year
    mth <- (background_all_sp$month[j])  # a variable for the observation month
    objIndex <- paste(yr,mth,sep="_")
    if(fileIndex == objIndex){
      no_all_netcdf_used_rows <- no_all_netcdf_used_rows + 1
      temp_brick_1 <- extract(x=temp_brick[[1]], y = background_all_sp[j, ])
      if (!(brkvar == "mlp.nc") & !(brkvar == "ssh.nc")){
        if (is.na(de)){
          temp_brick_de <- NA
        } else {
          temp_brick_de <- extract(x=temp_brick[[de]], y = background_all_sp[j, ]) 
        }
      }
      if(brkvar == "temp.nc"){
        background_all_sp$temp_surface[j] <- temp_brick_1
        background_all_sp$temp_depth[j] <-  temp_brick_de
      } else if (brkvar == "salinity.nc") {
        background_all_sp$salinity_surface[j] <- temp_brick_1 
        background_all_sp$salinity_depth[j] <- temp_brick_de
      } else if (brkvar == "chl.nc") {
        background_all_sp$chl_surface[j] <- temp_brick_1   
        background_all_sp$chl_depth[j] <- temp_brick_de 
      } else if (brkvar == "o2.nc") {
        background_all_sp$o2_surface[j] <- temp_brick_1
        background_all_sp$o2_depth[j] <- temp_brick_de
      } else if (brkvar == "mlp.nc") {
        background_all_sp$mlp_surface[j] <- temp_brick_1
      } else if (brkvar == "ssh.nc") {
        background_all_sp$ssh_surface[j] <- temp_brick_1
      }
    }
  }
}
write.csv(background_all_sp, "./output/points_with_data_3.csv", row.names = FALSE) #this ... writes a csv file so i have a permanent record that I can work with
end_time <- Sys.time()
print(end_time - start_time)
print(paste("Out of",no_netcdf,"netcdf files that contain all together",no_all_netcdf_rows,"rows"))
print(paste("Only",no_all_netcdf_used_rows,"are used"))
```
---
OK this gave me the result that I thought it might.
Out of 18 netcdf files that contain all together 198 rows
Only 66 rows are used

If this was a database project all I would do is create a query to reduce my data set to only matching data before I started to work with it ...
Also I would have the subset of data indexed :-
type, year, month, depth ext.

I wonder what people use to do this, looking at the link Sam gave me it looks like he uses a shaping file, but I worked out that is only to change his data into the county he is working on.

looked at :-
[link](https://cran.r-project.org/web/packages/futureheatwaves/vignettes/starting_from_netcdf.html)
[link](http://geog.uoregon.edu/bartlein/courses/geog490/week04-netCDF.html)
[link](https://www.youtube.com/watch?v=DvfTZSJLRfw)
Trying their examples using our data
---
Worried about package conflicts so found this if I need it :-
It is possible to have multiple versions of a package loaded at once (for example, if you have a development version and a stable version in different libraries). To detach and guarantee that all copies are detached, use this function.

```{r}
detach_package <- function(pkg, character.only = FALSE)
{
  if(!character.only)
  {
    pkg <- deparse(substitute(pkg))
  }
  search_item <- paste("package", pkg, sep = ":")
  while(search_item %in% search())
  {
    detach(search_item, unload = TRUE, character.only = TRUE)
  }
}
```

Usage is, for example

detach_package(vegan)
or

detach_package("vegan", TRUE)

to simply remove an object use
rm(object_name)

Back to the examples

---

Loading required packages 
```{r}
packages <- c("ncdf4.helpers","readr","tidyr","ggplot2")
source("./ipak.R")
ipak(packages)
suppressMessages(ipak)

library(ncdf4)
library(ncdf4.helpers)
library(PCICt)

library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

trying out with one file
Got bored repeating the same code for each variable I wanted to check so wrote some small functions
---
```{r}
netcdf_filepath <- paste0("./data/env/netcdf/",
                            "1998_1_salinity.nc")
get_variable_from_Netcdflink <- function(netcdflink,strVariable,PrintType = 0)
{
  # This function takes three variables
  #netcdflink, this an object that has been opened useing nc_open
  #strVariable, the required variable
  #PrintType, defaults to no printing
  # 1 a summary
  # 2 full print

  object_return = ncvar_get(netcdflink, varid = strVariable)
  if (PrintType == 1)
  {
    print(summary(object_return))
  }else if (PrintType == 1)
  {
    print(object_return)
  }
  return(object_return)
}

print_from_netcdflink <- function(netcdflink,PrintType = 0)
{
  # This function takes two variables
  #netcdflink, this an object that has been opened useing nc_open
  #PrintType, defaults to no printing 
  # 1 Filename,title, Origin date and calander type
  # 2 full print

  if (PrintType == 1)
  {
    print(paste("File name,",netcdflink$filename))
    print(paste("title",ncatt_get(netcdflink,0,"title")))
    print(paste("date type of file",netcdflink$dim$time$units))
    print(paste("calander type of file,",netcdflink$dim$time$calendar))
  }else if (PrintType == 2)
  {
    print(netcdflink)
  }
}

netcdf_output <- nc_open(netcdf_filepath)
print_from_netcdflink(netcdf_output,1)
netcdf_data <- ncvar_get(netcdf_output, attributes(netcdf_output$var)$names[1])

salinity <- get_variable_from_Netcdflink(netcdf_output,"salinity",0)
# salinity <- ncvar_get(netcdf_output,varid = "salinity")
# print(salinity)

time <- get_variable_from_Netcdflink(netcdf_output,"time",0)
longitude <- get_variable_from_Netcdflink(netcdf_output,"longitude",2)
latitude <- get_variable_from_Netcdflink(netcdf_output,"latitude",2)
depth <- get_variable_from_Netcdflink(netcdf_output,"depth",0)

time_index <- which.min(abs(time - 421140))
longitude_index <- which.min(abs(longitude - -60.48507517))
latitude_index <- which.min(abs(latitude - 62.37859726))
depth_index <- which.min(abs(depth - 14))
print(longitude_index)
print(latitude_index)
print(depth_index)

netcdf_req_data <- netcdf_data[longitude_index,latitude_index,depth_index]
print(netcdf_req_data)
nc_close(netcdf_output)

```

Hmm some interesting stuff but still struggling to get to grips with this data.
I had hoped I might have been able to get the same value that is on line one of the test point data under the column salinity_depth

I think we want to be able get to the stage where we can read only the points that are required directly into the data used to create the csv data so a bigger csv would read more data and smaller less. That way I am hoping Sam can get an idea of a performance to data ratio.

First things first, maybe I should try and wrap what we have now in function to see if I can get some performance benefits.
---
```{r}
#Basic rework of code now uses functions and mclapply for parallel running (supposedly)

#Detect cores automatically, I usually free one up.
cores <- detectCores() - 1

no_all_netcdf_rows <- 0
no_all_netcdf_used_rows <- 0
no_netcdf_rows <- 0
intRow <- 0

start_time <- Sys.time()
print(paste("start time",start_time,sep=" "))
background_all2 <- read.csv("./data/simon_test_points.csv", header = TRUE)
xy <- background_all2[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. 
background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
netcdf_Path <- "./data/env/netcdf"
netcdf_list <- list.files("./data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
aea <- raster("./data/env/aea.tif")

deal_With_netncf_file <- function(strFilePath)
{
  strFileName <- basename(strFilePath)
  netcdf_active <- strFileName
  print(netcdf_active) #this just prints the name of the netCDF R is working one
  netcdf_split <- strsplit(strFileName, "_")
  brkyr <- unlist(netcdf_split)[1] # extracting the first part of the netcdf filename (which is the year)
  brkmth <-unlist(netcdf_split)[2] # extracting the second part of the netcdf filename (which is the month)
  brkvar <<- unlist(netcdf_split)[3] # extracting the third part of the netcdf (inc.nc)
  fileIndex <<- paste(brkyr,brkmth,sep="_")
  FileTimeStart <- Sys.time()
  temp_brick <<- brick(strFilePath, lvar = 4)
  FileTimeEnd <- Sys.time()
  #print(paste("File read took",FileTimeEnd - FileTimeStart))
  RasterTimeStart <- Sys.time()
  temp_brick <<- projectRaster(temp_brick, aea)
  RasterTimeEnd <- Sys.time()
  #print(paste("Raster took",RasterTimeEnd - RasterTimeStart))
  no_req_rows <- nrow(background_all_sp)
  RowsTimeStart <- Sys.time()
  #for (j in 1:no_req_rows) {
  #  deal_With_netncf_row(j)
    req_row_list <- 1:no_req_rows
    mclapply(req_row_list,deal_With_netncf_row)
  #}
  RowsTimeEnd <- Sys.time()
  #print(paste("Rows took",RowsTimeEnd - RowsTimeStart))
}
deal_With_netncf_row <- function(intRow)
{
  de <- background_all_sp$depthlayerno[[intRow]]  # a variable for the observation depth layer
  yr <- (background_all_sp$year[intRow])  # a variable for the observation year
  mth <- (background_all_sp$month[intRow])  # a variable for the observation month
  objIndex <- paste(yr,mth,sep="_")
  if(fileIndex == objIndex){
    brkvar_split <- strsplit(brkvar, "\\.")
    strColumn <- unlist(brkvar_split)[1]
    temp_brick_1 <- raster::extract(x=temp_brick[[1]], y = background_all_sp[intRow, ])
    if (!is.null(background_all_sp[[paste0(strColumn,"_surface")]])){
      background_all_sp[[paste0(strColumn,"_surface")]][intRow] <<- temp_brick_1
    } else {
      #print(paste("column",paste0(strColumn,"_surface"),"not found on datasheet"))
    }
    #print(paste("temp_brick_1",temp_brick_1,"row",intRow))
    if (!is.null(background_all_sp[[paste0(strColumn,"_depth")]])){
      if (is.na(de)){
        temp_brick_de <- NA
      } else {
        temp_brick_de <- raster::extract(x=temp_brick[[de]], y = background_all_sp[intRow, ]) 
      }
      background_all_sp[[paste0(strColumn,"_depth")]][intRow] <<- temp_brick_de
      #print(paste("temp_brick_de",temp_brick_de,"row",intRow))
    } else {
      #print(paste("column",paste0(strColumn,"_depth"),"not found on datasheet"))
    }

  } else {
    #print(paste("ObjectIndex",objIndex,"Unread row",intRow))
  }
}

c.deal_With_netncf_file <- compiler::cmpfun(deal_With_netncf_file)
c.deal_With_netncf_row <- compiler::cmpfun(deal_With_netncf_row)

mclapply(netcdf_list,deal_With_netncf_file) # Applies the list of files to deal with function'

write.csv(background_all_sp, "./output/points_with_data_4.csv", row.names = FALSE) #this ... writes a csv file so i have a permanent record that I can work with
end_time <- Sys.time()
print(end_time - start_time)
```
---
I have removed the if check on the files because now what it does is takes the first part of the file name and uses that with "_surface" or "_depth" to set the background_all_sp object.
That way as long as the files you are loading have matching columns you can expand your table without the need to change the code.
Put in mclapply to run lists in parallel for the file and the row
Compiled functions

But it does not not seem to improve the speed that much.
The two parts that seem to take the most time are the file read and projectRaster, but I don't know enough about the data to speed that up.
I even tried using ncdf4 which allows you to link to the file rather than loading it but again I don't know enough about your data to get it working.

There seems to be a magical filtering process with the aea object but I dont understand it enough to get that working with the nc_open options.

----

### 2019-03-20 src/stack_test.Rmd
some maxentvariableselection playing, and update to potential change to background point extraction

@@
 
+with all variables
 ```{r}
@@
 
-VariableSelection(maxent="C:/Users/bunny/Downloads/maxent/maxent.jar", outdir=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/output"), gridfolder=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/testgrid"), occurrencelocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/presence.csv"), backgroundlocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/background.csv"), additionalargs="nolinear noquadratic noproduct nothreshold noautofeature", contributionthreshold=5, correlationthreshold=0.9, betamultiplier=seq(0.5, 4, 0.25))
+VariableSelection(maxent="C:/Users/bunny/Downloads/maxent/maxent.jar", outdir=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/output"), gridfolder=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/testgrid"), occurrencelocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/presence.csv"), backgroundlocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/background.csv"), additionalargs="nolinear noquadratic noproduct nothreshold noautofeature", contributionthreshold=5, correlationthreshold=0.9, betamultiplier=seq(0.5, 3, 0.25))
 ```
+
+with my manually selected variables for jan (temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+```{r}
+janaball <- read.csv("../output/bio/background_10k.csv", header = TRUE)
+janpresall <- read.csv("../output/bio/pres_nomissing.csv", header = TRUE)
+janpresaball_red <- rbind.fill(janaball, janpresall)
+janpresaball_red <- subset(janpresaball_red, month == 1)
+janpresaball_red <- subset(janpresaball_red, select = c(occurrence, longitude_meters, latitude_meters, temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter))
+names(janpresaball_red)[names(janpresaball_red)=="occurrence"] <- "species"
+names(janpresaball_red)[names(janpresaball_red)=="longitude_meters"] <- "longitude"
+names(janpresaball_red)[names(janpresaball_red)=="latitude_meters"] <- "latitude"
+janpresall_red <- subset(janpresaball_red, species == 1)
+janpresall_red$species <- replace(janpresall_red$species, janpresall_red$species == "1", "fish")
+write.csv(janpresall_red, "../output/maxentvariableselection/presence_red.csv", row.names = FALSE)
+janaball_red <- subset(janpresaball_red, species == 0)
+janpresall_red$species <- replace(janpresall_red$species, janpresall_red$species == "0", "bg")
+write.csv(janaball_red, "../output/maxentvariableselection/background_red.csv", row.names = FALSE)
+
+VariableSelection(maxent="C:/Users/bunny/Downloads/maxent/maxent.jar", outdir=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/output"), gridfolder=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/testgrid"), occurrencelocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/presence_red.csv"), backgroundlocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/background_red.csv"), additionalargs="nolinear noquadratic noproduct nothreshold noautofeature", contributionthreshold=5, correlationthreshold=0.9, betamultiplier=seq(0.25, 3, 0.25))
+```
+
+so this is all lovely...
+the two runs suggest b-multipliers of <1 (0.75 and 0.25 respectively).
+In my test runs above, 0.25 gives a 'tighter fit' thatn 1 (the default).
+Apparently changin the b-multiplier helps reduce over and underfitting...
+"the default setting in Maxent is 1.
+Smaller values will constrain the model more, whereas larger numbers will relax the model, making it “smoother”......
+
+maybe try enmval?

### 2019-03-20 src/stack_test.Rmd
minor mod
playing with random test data %

@@
 
-pre8.1 <- predict(mod8.1, test, args='outputformat=logistic' ) 
+pre8.1 <- predict(mod8.1, test, args=c('outputformat=logistic'), file="../output/maxtest.asc") # here i added an asc output so i could pop into arc for a test
 preplot8.1 <- plot(pre8.1)
@@
 
-maybe try enmval?
+maybe try enmeval? NO CAN'T USE WITH SWD https://groups.google.com/forum/#!topic/maxent/lIT4tY_6toY

### 2019-03-21 src/stack_test.Rmd
creating aic + summary table of aic and auc at different regularization values

@@
   #stackSave(test,unlayer$stackname[td])
-  pred <- predict(mod1, test, file = paste0("../output/maxentpredict/", stackn), overwrite=TRUE) 
+  pred <- predict(mod1, test, file = paste0("../output/maxentpredict/", stackn, ".asc"), overwrite=TRUE) #sam you just added.asc but have not tested
   paste("maxplot_", stackn, ".png", sep = "") # to automatically save the plot to a png AND show it inline
@@
 
+#--------------------------
+
+#play with test and training data - note will want to give your own file in the 'real model'
+mod9 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30")) 
+mod9
+
+pre9 <- predict(mod9, test, args='outputformat=logistic' ) 
+preplot9 <- plot(pre9)
+preplot9
+
+#ok this works but doesnt show much difference in the map
+
+#------------------------
+
+#run a model with all years data and project as you have been 
+presaball <- read.csv("../output/bio/presab_10k.csv", header = TRUE)
+subpresaball <- subset(presaball, select = c(occurrence, temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter))
+subpaall <- subpresaball$occurrence
+subpresaball <- subset(subpresaball, select = -c(occurrence))
+
+mod10 <- maxent(x = subpresaball, p = subpaall, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30")) 
+mod10
+
+pre10 <- predict(mod10, test, args='outputformat=logistic' ) 
+preplot10 <- plot(pre10)
+preplot10
+
+# well this makes a terrible mess... the habitat is predicted to be everywhere and the response curves a mess! Was just curious.
+
+#--------------------
+
+#back to fiddling
+
+mod11 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "replicatetype=boo", "replicates=10")) 
+mod11
+
+pre11 <- predict(mod11, test, args='outputformat=logistic' ) 
+preplot11 <- plot(pre11)
+preplot11
+
+#-------------------
+
+#aic fiddling
+
+betamval <- 0.25
+mod9 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.25"), path = "../output/maxent/output/mod9foraic")
+mod9
+
+pre9 <- predict(mod9, test, args='outputformat=logistic' ) 
+preplot9 <- plot(pre9)
+preplot9
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE)
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
+AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))
+
 ```
@@
 
+try changing some of the settings here - paerticularly you want aic
+```{r}
+VariableSelection(maxent="C:/Users/bunny/Downloads/maxent/maxent.jar", outdir=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/output"), gridfolder=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/testgrid"), occurrencelocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/presence_red.csv"), backgroundlocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/background_red.csv"), additionalargs="nohinge", contributionthreshold=5, correlationthreshold=0.9, betamultiplier=seq(0.25, 2, 0.25))
+```
+no you cant - you need to use the grids...
+
 maybe try enmeval? NO CAN'T USE WITH SWD https://groups.google.com/forum/#!topic/maxent/lIT4tY_6toY
+
+# for mj thing - no points per nafo zone
+
+I previously worked out the area for each nafo zone in arcgis. This caclulates the  number of points per nafo zone
+
+```{r}
+allpoints <- read.csv("../output/bio/background/correct_random_files/background_point_na_removal_nona.csv", header = TRUE) #you may want to move the files once you figure out what mess you created..
+nafo_points <- count(allpoints, "nafo_zone")
+total_points <- sum(nafo_points$freq) #191055 -correct
+nafo_points$points_percentage <- (nafo_points$freq/total_points)*100
+
+#combine with the arcgis output...
+nafovolume <- read.csv("../data/env/volume_all_nafo.csv", header = TRUE) #note the hudson strait points will be invalid as there is no polygon to calculate volume on. change to NA
+nafo_points$volume <- nafovolume$Volume
+nafo_points$volume_percentage <- nafovolume$volume_perc
+
+write.csv(nafo_points, file = "../output/env/no_points_per_nafo.csv")
+```

### 2019-03-21 src/stack_test.Rmd
aic refix

@@
 
+AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE)
+
 betamval <- 0.25
-mod9 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.25"), path = "../output/maxent/output/mod9foraic")
-mod9
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.25"), path = "../output/maxent/output/mod9foraic")
 
-pre9 <- predict(mod9, test, args='outputformat=logistic' ) 
-preplot9 <- plot(pre9)
-preplot9
+#now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 0.50
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.50"), path = "../output/maxent/output/mod9foraic")
 
@@
 
-AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE)
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 0.75
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.75"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1.25
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.25"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1.50
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.50"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1.75
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.75"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2.25
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.25"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2.50
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.50"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2.75
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.75"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3.25
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.25"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3.50
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.50"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3.75
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.75"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 4
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=4"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
 for (i in 1:nrow(AIClist)){

### 2019-03-26 src/stack_test.Rmd
happy with maxent model fiddling.

@@
 
-mod11 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "replicatetype=boo", "replicates=10")) 
+mod11 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "replicates=10")) #replicates added
 mod11
@@
 
+#probably dont want replicates - no real way to merge the outputs. saw somehting online that said ideally you want to run just one model
+
 #-------------------
 
-#aic fiddling
-
-AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE)
-
-betamval <- 0.25
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.25"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 0.50
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.50"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 0.75
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.75"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 1
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 1.25
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.25"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 1.50
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.50"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 1.75
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.75"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 2
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 2.25
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.25"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 2.50
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.50"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 2.75
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.75"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 3
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 3.25
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.25"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 3.50
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.50"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 3.75
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.75"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-betamval <- 4
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=4"), path = "../output/maxent/output/mod9foraic")
-
-#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
-
-#count number of non-zero coefficinets, excluding the final four rows
-coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
-coefile <- head(coefile, -4)
-coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-nocoeff <- nrow(coefile2)
-
-#no presences used in model (less those used for testing)
-maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
-notrainingpoints <- maxresults$X.Training.samples
-testauc <- maxresults$Test.AUC
-trainauc <- maxresults$Training.AUC
-
-#natural log
-samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
-samplepred$nat_log <- log(samplepred$Raw.prediction)
-sumsamplepred <- sum(samplepred$nat_log)
-
-for (i in 1:nrow(AIClist)){
-  if(AIClist$regularization[i] == betamval){
-    AIClist$Test.AUC[i] <- testauc
-    AIClist$Train.AUC[i] <- trainauc
-    AIClist$n[i] <- notrainingpoints
-    AIClist$coefficients[i] <- nocoeff
-    AIClist$sum.of.logs[i] <- sumsamplepred
-
-  }
-}
-
-AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
-AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))
+mod12 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "addsamplestobackground=TRUE")) #add to samples because field study with biased data - see https://esajournals.onlinelibrary.wiley.com/doi/10.1890/10-1171.1.
+note this should be the default but keep here as a note
+mod12
+
+pre12 <- predict(mod12, test, args='outputformat=logistic' ) 
+preplot12 <- plot(pre12)
+preplot12
+
+#ok think i am good. the think you will want to change is the beta value - depending on the results of the AIC and when you figure out the test/training data partion, will need to add testsamplesfile
 
@@
 ```
+
+#AIC for regularization
+
+```{r}
+#aic fiddling
+
+AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE)
+
+betamval <- 0.25
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.25"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 0.50
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.50"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 0.75
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.75"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1.25
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.25"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1.50
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.50"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1.75
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.75"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2.25
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.25"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2.50
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.50"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2.75
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.75"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3.25
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.25"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3.50
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.50"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3.75
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.75"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 4
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=4"), path = "../output/maxent/output/mod9foraic")
+
+#now based on glover-kapfer pdf with excel manual.... lets try automate some of this..
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv("../output/maxent/output/mod9foraic/maxentResults.csv", header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv("../output/maxent/output/mod9foraic/species_samplePredictions.csv", header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
+AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))
+
+write.csv(AIClist, "../output/bio/AICforMaxentReg.csv", row.names = FALSE)
+```

### 2019-03-26 src/asc_file_database.Rmd
all raster stack links done
asc outputting

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("tidyr")
+packages <- c("tidyr", "rJava", "dismo")
 source("../src/ipak.R")
@@
 
-```{r}
+```{r base list or ascii layers}
 ascyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
@@
 
-```
-
-```{r}
+write.csv(asclayerall, "../data/env/asc_stacklist.csv", row.names = FALSE)
+```
+
+```{r asclist for 1998}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 1999}
 #!!!!!!  CHANGE ASCLAYER ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2000}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2001}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2002}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2003}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2004}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2005}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2006}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2007}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2008}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2009}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2010}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2011}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2012}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
-#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
-
-asclayer <- subset(asclayerall, ascyears == 2103)
+```{r asclist for 2013}
+#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+
+asclayer <- subset(asclayerall, ascyears == 2013)
 
@@
 
-```{r}
+```{r asclist for 2014}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
@@
 
-```{r}
+```{r asclist for 2015}
 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

### 2019-03-26 src/stack_test.Rmd
asc outputting

@@
 
-need tidyr package
-```{r}
-unyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
-#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
-unmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
-undepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38") #this list is based on how the files are named when split into depth layers.
-note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers
-
-unlayer <- crossing(unyears, unmonths, undepths)
-unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used
-unlayer$temp_depth <- NA
-unlayer$temp_surface <- NA
-unlayer$chl_depth <- NA
-unlayer$chl_surface <- NA
-unlayer$o2_depth <- NA
-unlayer$o2_surface <- NA
-unlayer$salinity_depth <- NA
-unlayer$salinity_surface <- NA
-unlayer$nao_sample <- NA
-unlayer$nao_prev <- NA
-unlayer$nao_winter <- NA
-unlayer$amo_sample <- NA
-unlayer$amo_prev <- NA
-unlayer$amo_winter <- NA
-
-ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
-ascpath <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
-
-#year <- "X1998"
-#month <- "01"
-#depth <- "86.9294"
-tempvar <- "temp"
-chlvar <- "chl"
-o2var <- "o2"
-salvar <- "salinity"
-depthvar <- "depth"
-surfacevar <- "surface"
-naovar <- "nao"
-amovar <- "amo"
-samplevar <- "sample"
-wintervar <- "winter"
-
-for (k in 1:nrow(unlayer)){
-    print(unlayer$stackname[k])
-    year <- unlayer$unyears[k]
-    month <- unlayer$unmonths[k]
-    depth <- unlayer$undepths[k]
-    for (i in 1:length(ascname)){
-
-    ascyear <- (sapply(strsplit(ascname[[i]], "_"), "[[", 1))
-    ascmonth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 2))
-    ascdepth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 5))
-    ascdepth <- gsub(".asc", "", ascdepth)
-    ascvar1 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 3))
-    ascvar2 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 4))
-    ascfile <- ascpath[[i]]
-    print(ascfile)
-
-    if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
-    unlayer$temp_depth[k] <- ascfile
-  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
-   unlayer$chl_surface[k] <- ascfile
- } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
-  unlayer$o2_depth[k] <- ascfile
- } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
-  unlayer$salinity_depth[k] <- ascfile
- } else if (ascyear == year & ascvar1 == naovar & ascvar2 == wintervar){
-  unlayer$nao_winter[k] <- ascfile
- } else if (ascyear == year & ascmonth == month  & ascvar1 == naovar & ascvar2 == samplevar){
-  unlayer$nao_sample[k] <- ascfile
- }
-
-    }
-    write.csv(unlayer, "../data/env/asc_layerlist.csv", row.names = TRUE)
-}
-
 # now loop to create stacks
@@
 plot(mod1)
+
+need tidyr package
+```{r}
 
@@
 
-  preplot <- plot(pred, png(paste0(stackn, ".png")))
-
-```{r}
-var <- raster(stacktest$chl[1])
-var
-a <- raster(var)
-a
-```
-
 #maxent
@@
 ```{r}
-unyears <- as.character(c("1992", "1998")) 
-unmonths <- c("01", "03")
-undepths <- c("10", "86.9294")
-unlayer <- crossing(unyears, unmonths, undepths)
-unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used
-
-unlayer$temp_depth <- NA
-unlayer$temp_surface <- NA
-unlayer$chl_depth <- NA
-unlayer$chl_surface <- NA
-unlayer$o2_depth <- NA
-unlayer$o2_surface <- NA
-unlayer$salinity_depth <- NA
-unlayer$salinity_surface <- NA
-unlayer$nao_sample <- NA
-unlayer$nao_prev <- NA
-unlayer$nao_winter <- NA
-unlayer$amo_sample <- NA
-unlayer$amo_prev <- NA
-unlayer$amo_winter <- NA
-
-tempvar <- "temp"
-chlvar <- "chl"
-o2var <- "o2"
-salvar <- "salinity"
-depthvar <- "depth"
-surfacevar <- "0.50576"
-naovar <- "nao"
-amovar <- "amo"
-samplevar <- "sample"
-wintervar <- "winter"
-
-ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
-ascpath <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
-
-for (k in 1:nrow(unlayer)){
-    year <- unlayer$unyears[k]
-    month <- unlayer$unmonths[k]
-    depth <- unlayer$undepths[k]
-
-    for (i in 1:length(ascname)){
-    ascyear <- (sapply(strsplit(ascname[[i]], "_"), "[[", 1))
-    ascmonth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 2))
-    ascvar <- (sapply(strsplit(ascname[[i]], "_"), "[[", 3))
-    ascdepth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 4))
-    ascdepth <- gsub(".asc", "", ascdepth)
-    ascfile <- ascpath[[i]]
-    print(ascfile)
-    if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
-      unlayer$temp_depth[k] <- ascfile
-    } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
-        unlayer$temp_surface[i] <- ascfile
-    }
-    }
-}
-
-```
-
-```{r}
 # now loop to create stacks
@@
 
-system.file("java", package="dismo") #finding location to put maxent jar file
-system("java -version")
 mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
 
+#THIS IS THE GOOD CODE!
 unlayer <- read.csv("../data/env/asc_layerlist1998.csv", header = TRUE)
@@
 
-```{r}
+```{r fiddling with maxent args}
 janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
@@
 
-```{r}
+```{r maxent with all variables}
 presaball <- read.csv("../output/bio/presab_10k.csv", header = TRUE)
@@
 with all variables
-```{r}
+```{r maxent variable selection package - not used}
 janaball <- read.csv("../output/bio/background_10k.csv", header = TRUE)
@@
 
-```{r}
+```{r AIC for regularization selection}
 #aic fiddling

### 2019-03-27 src/stack_test.Rmd
stuff

@@
 
-need tidyr package
-```{r}
-
-unlayer <- read.csv("../data/env/asc_layerlist1998.csv", header = TRUE)
-unlayer <- subset(unlayer, ascmonths == 1)
-unlayer <- unlayer[-c(1), ]
-unlayer[] <- lapply(unlayer, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
-aea <- raster("../output/env/aea.tif")
-
-for (td in 1:nrow(unlayer)){
-  temp_depth <- raster(unlayer$temp_depth[td])
-  #crs(temp_depth) <- +proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 
-  names(temp_depth) <- "temp_depth"
-  #temp_depth <- projectRaster(temp_depth, aea)
-  chl_surface <- raster(unlayer$chl_surface[td])
-  names(chl_surface) <- "chl_surface"
-  o2_depth <- raster(unlayer$o2_depth[td])
-  names(o2_depth) <- "o2_depth"
-  salinity_depth <- raster(unlayer$salinity_depth[td])
-  names(salinity_depth) <- "salinity_depth"
-  nao_sample <- raster(unlayer$nao_sample[td])
-  names(nao_sample) <- "nao_sample"
-  nao_winter <- raster(unlayer$nao_winter[td])
-  names(nao_winter) <- "nao_winter"
-  stackn <- unlayer$stackname[td]
-  test <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
-  #stackSave(test,unlayer$stackname[td])
-  pred <- predict(mod1, test, file = paste0("../output/maxentpredict/", stackn), overwrite=TRUE) 
-  preplot <- plot(pred, png(paste0("../output/maxentpredict/", stackn, ".png")))
-  paste("maxplot_", stackn, ".png", sep = "") # to automatically save the plot to a png AND show it inline
-
-}
-dev.off() # stops automatic saving of the plot to a png
-# now call the stack...
-
-```
-
 #maxent
@@
 prab
-
-mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))
 
@@
   #stackSave(test,unlayer$stackname[td])
-  pred <- predict(mod1, test, file = paste0("../output/maxentpredict/", stackn, ".asc"), overwrite=TRUE) #sam you just added.asc but have not tested
+  pred <- predict(mod1, test, file = paste0("../output/maxentpredict/", stackn, ".asc"), overwrite=FALSE) 
   paste("maxplot_", stackn, ".png", sep = "") # to automatically save the plot to a png AND show it inline
-  preplot <- plot(pred, png(paste0("../output/maxentpredict/", stackn, ".png")))
-  dev.off() # stops automatic saving of the plot to a png
+  #preplot <- plot(pred, png(paste0("../output/maxentpredict/", stackn, ".png"))) #keep these out if using xvalidation
+  #dev.off() # stops automatic saving of the plot to a png #keep these out if using xvalidation
 }
@@
 strt <- Sys.time() #get the start time
-mod7 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4")) #threads should speed things up...
+mod7 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8")) #threads should speed things up...
 mod7
@@
 
-mod8.1 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=0.25")) #playing with multiplier
+mod8.1 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "betamultiplier=0.25")) #playing with multiplier
 mod8.1
@@
 
-mod8 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4")) #default setting for betam is 1
+mod8 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8")) #default setting for betam is 1
 mod8
@@
 
-mod8.4 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=1.25")) #playing with multiplier
+mod8.4 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "betamultiplier=1.25")) #playing with multiplier
 mod8.4
@@
 
-mod8.2 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=2")) 
+mod8.2 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "betamultiplier=2")) 
 mod8.2
@@
 
-mod8.3 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=6")) 
+mod8.3 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "betamultiplier=6")) 
 mod8.3
@@
 #play with test and training data - note will want to give your own file in the 'real model'
-mod9 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30")) 
+mod9 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30")) 
 mod9
@@
 
-mod10 <- maxent(x = subpresaball, p = subpaall, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30")) 
+mod10 <- maxent(x = subpresaball, p = subpaall, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30")) 
 mod10
@@
 
-mod11 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "replicates=10")) #replicates added
+mod11 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "replicates=10")) #replicates added
 mod11
@@
 
-mod12 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "addsamplestobackground=TRUE")) #add to samples because field study with biased data - see https://esajournals.onlinelibrary.wiley.com/doi/10.1890/10-1171.1.
+mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "replicatetype=crossvalidate", "replicates=5")) #add to samples because field study with biased data - see https://esajournals.onlinelibrary.wiley.com/doi/10.1890/10-1171.1.
 note this should be the default but keep here as a note
 mod12
+
+mod12[2]
 
@@
 
-mod1all <- maxent(x = janpresaball, p = paall, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4")) 
+mod1all <- maxent(x = janpresaball, p = paall, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8")) 
 mod1all
@@
 betamval <- 0.25
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.25"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=0.25"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 0.50
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.50"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=0.50"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 0.75
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=0.75"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=0.75"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 1
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=1"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 1.25
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.25"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=1.25"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 1.50
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.50"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=1.50"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 1.75
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=1.75"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=1.75"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 2
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=2"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 2.25
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.25"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=2.25"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 2.50
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.50"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=2.50"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 2.75
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=2.75"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=2.75"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 3
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=3"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 3.25
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.25"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=3.25"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 3.50
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.50"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=3.50"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 3.75
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=3.75"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=3.75"), path = "../output/maxent/output/mod9foraic")
 
@@
 betamval <- 4
-modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "randomtestpoints=30", "betamultiplier=4"), path = "../output/maxent/output/mod9foraic")
+modaic <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "randomtestpoints=30", "betamultiplier=4"), path = "../output/maxent/output/mod9foraic")
 
@@
 ```
+
+#subset background points
+
+```{r}
+# now loop to create stacks
+janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
+janpres$occ <- 1
+janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
+janab$occ <- 0
+
+#subset2500 randomly
+janab25 <- sample_n(janab, 2500)
+
+prab <- rbind(janpres, janab25)
+prab$originalscientificname <- "cap" #because species name needs to be the same
+pa <- prab$occ
+prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
+
+mod13 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE"))
+
+#THIS IS THE GOOD CODE!
+unlayer <- read.csv("../data/env/asc_layerlist1998.csv", header = TRUE)
+unlayer <- subset(unlayer, ascmonths == 1)
+unlayer <- unlayer[-c(1), ]
+unlayer[] <- lapply(unlayer, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+aea <- raster("../output/env/aea.tif")
+
+for (td in 1:nrow(unlayer)){
+  temp_depth <- raster(unlayer$temp_depth[td])
+  crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(temp_depth) <- "temp_depth"
+  temp_depth <- projectRaster(temp_depth, aea)
+  chl_surface <- raster(unlayer$chl_surface[td])
+  crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(chl_surface) <- "chl_surface"
+  chl_surface <- projectRaster(chl_surface, aea)
+  o2_depth <- raster(unlayer$o2_depth[td])
+  crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(o2_depth) <- "o2_depth"
+  o2_depth <- projectRaster(o2_depth, aea)
+  salinity_depth <- raster(unlayer$salinity_depth[td])
+  crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(salinity_depth) <- "salinity_depth"
+  salinity_depth <- projectRaster(salinity_depth, aea)
+  nao_sample <- raster(unlayer$nao_sample[td])
+  crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+  names(nao_sample) <- "nao_sample"
+  nao_sample <- projectRaster(nao_sample, aea)
+  nao_winter <- raster(unlayer$nao_winter[td])
+  crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+  names(nao_winter) <- "nao_winter"
+  nao_winter <- projectRaster(nao_winter, aea)
+  stackn <- unlayer$stackname[td]
+  test <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+  #stackSave(test,unlayer$stackname[td])
+  pred <- predict(mod13, test, file = paste0("../output/maxentpredict/", stackn, ".asc"), overwrite=FALSE) 
+  paste("maxplot_", stackn, ".png", sep = "") # to automatically save the plot to a png AND show it inline
+  #preplot <- plot(pred, png(paste0("../output/maxentpredict/", stackn, ".png"))) #keep these out if using xvalidation
+  #dev.off() # stops automatic saving of the plot to a png #keep these out if using xvalidation
+}
+```

### 2019-04-03 src/maxent_avec_xvalidation.Rmd
please please please

---
title: "maxent model avec xvalidation"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

Built with 'r getRversion()'..

# Package dependencies
You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). Note this function checks to see if the packages are installed first before loading.

```{r pre-install & load packages, include=FALSE}
packages <- c("dismo")
source("../src/ipak.R")
ipak(packages)
```

#load the presence-background dataset 
("../output/bio/presab_nafo_yr.csv")

```{r load pres-background dataset}
presback <- read.csv("./output/bio/presab_nafo_yr.csv", header = TRUE)
colnames(presback)
```

this dataset contains a lot of columns you don't need for the maxent models. Remove them here

```{r remove unnecessary columns}
presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt))
colnames(presback)
```

#monthly model data selection and prep
For each monthly model, subset the data to the month you want, and the variables you want (Based on spearmans and VIF)

jan variables to keep: temp_depth, salinty_depth, o2_depth, chl_surface, nao_sample, nao_winter 
```{r january data}
janprback <- subset(presback, month == "1")
janprback <- subset(janprback, select = c(occurrence, year, temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_winter))
```

now each month contains NAs - lets remove these (maxent principle doesnt allow for missing data - http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.15.5607&rep=rep1&type=pdf)

```{r}
janprback <- janprback[!is.na(janprback$temp_depth), ]
janprback <- janprback[!is.na(janprback$salinity_depth), ]
janprback <- janprback[!is.na(janprback$o2_depth), ]
janprback <- janprback[!is.na(janprback$chl_surface), ]
janprback <- janprback[!is.na(janprback$nao_sample), ]
janprback <- janprback[!is.na(janprback$nao_winter), ]
```

#create temporally differentiated testing/training data
as per elith and rad.....

```{r} 
#actually you will not use this
noyears <- unique(janprback$year)
train2003 <- subset(janprback, year = "2003")
test2003 <-  subset(janprback, year == "2003")
train2004 <- subset(janprback, year == "2004")
train2005 <- subset(janprback, year == "2005")
train2006 <- subset(janprback, year == "2006")
```

normal model..
mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE")) 
mod12

```{r train and test on 2003 inc AIC for reg val}
train2003 <- subset(janprback, year != "2003")
test2003 <-  subset(janprback, year == "2003")
train2003pa <- train2003$occurrence
test2003p <- subset(test2003, occurrence == "1")
test2003b <- subset(test2003, occurrence == "0")
train2003  <- subset(train2003, select = -c(year, occurrence))

mod <- maxent(x = train2003, p = train2003pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE"))
mod
ev <- evaluate(p = test2003p, a= test2003b, model = mod)
ev
threshold(ev)
```

# beta value (regularization) selection

mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "replicatetype=crossvalidate", "replicates=5")) #add to samples because field study with biased data - see https://esajournals.onlinelibrary.wiley.com/doi/10.1890/10-1171.1. note this should be the default but keep here as a note

### 2019-04-03 src/stack_test.Rmd
new background dataset
please please please

@@
 
-#count number of non-zero coefficinets, excluding the final four rows
+#count number of non-zero coefficients, excluding the final four rows
 coefile <- read.csv("../output/maxent/output/mod9foraic/species.lambdas", header = FALSE)
@@
 ```
+
+from dismo thing... but with my bad code... seems to need to have 
+```{r}
+# file with presence points
+fnames <- list.files(path=paste(system.file(package="dismo"), '/ex', sep=''),
+pattern='grd', full.names=TRUE )
+predictors <- stack(fnames)
+occurence <- paste(system.file(package="dismo"), '/ex/bradypus.csv', sep='')
+occ <- read.table(occurence, header=TRUE, sep=',')[,-1]
+
+fold <- kfold(occ, k=5)
+occtest <- occ[fold == 1, ]
+occtrain <- occ[fold != 1, ]
+
+occtrain <- train2003
+occtrainpa <- occtrain$occurrence
+occtrain <- subset(occtrain, select = -c(year, occurrence))
+
+me2 <- maxent(x = occtrain, p = occtrainpa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE"))
+me2
+
+occtest <- test2003
+occtrain <- train2003
+e1 <- evaluate(p= occtrain, a= occtest, me2) #@models[[2]]
+e1
+
+pvtest <- data.frame(extract(predictors, occtest))
+avtest <- data.frame(extract(predictors, bg))
+e2 <- evaluate(me2, p=pvtest, a=avtest)
+e2
+
+testp <- predict(me2, pvtest)
+testa <- predict(me2, avtest)
+e3 <- evaluate(p=testp, a=testa)
+e3
+threshold(e3)
+plot(e3, 'ROC')
+
+```

### 2019-04-03 src/maxent_avec_xvalidation.Rmd
aic injection

@@
 ```{r pre-install & load packages, include=FALSE}
-packages <- c("dismo")
+packages <- c("dismo", "stringr")
 source("../src/ipak.R")
@@
 ```{r load pres-background dataset}
-presback <- read.csv("./output/bio/presab_nafo_yr.csv", header = TRUE)
+presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
 colnames(presback)
@@
 ```{r remove unnecessary columns}
-presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt))
+presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt))
 colnames(presback)
@@
 
-```{r}
+```{r jan remove na vals}
 janprback <- janprback[!is.na(janprback$temp_depth), ]
@@
 
-#create temporally differentiated testing/training data
-as per elith and rad.....
-
-```{r} 
-#actually you will not use this
-noyears <- unique(janprback$year)
-train2003 <- subset(janprback, year = "2003")
-test2003 <-  subset(janprback, year == "2003")
-train2004 <- subset(janprback, year == "2004")
-train2005 <- subset(janprback, year == "2005")
-train2006 <- subset(janprback, year == "2006")
-```
-
 normal model..
@@
 ```{r train and test on 2003 inc AIC for reg val}
-train2003 <- subset(janprback, year != "2003")
-test2003 <-  subset(janprback, year == "2003")
-train2003pa <- train2003$occurrence
-test2003p <- subset(test2003, occurrence == "1")
-test2003b <- subset(test2003, occurrence == "0")
-train2003  <- subset(train2003, select = -c(year, occurrence))
+timeslice <- "jan2003"
+yr <- str_sub(timeslice, -4)
+train <- subset(janprback, year != yr)
+trainpa <- train$occurrence
+test <-  subset(janprback, year == yr)
+testp <- subset(test, occurrence == "1")
+testp <- subset(testp, select = -c(year, occurrence))
+testb <- subset(test, occurrence == "0")
+testb <- subset(testb, select = -c(year, occurrence))
+train  <- subset(train, select = -c(year, occurrence))
 
-mod <- maxent(x = train2003, p = train2003pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE"))
-mod
-ev <- evaluate(p = test2003p, a= test2003b, model = mod)
-ev
-threshold(ev)
+AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE)
+
+#netam = 0.25
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.25"))
+
+#now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1.... lets try automate some of this..
+
+#count number of non-zero coefficients, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+#betamval <- 0.50
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.50"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 0.75
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.75"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=1"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1.25
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=1.25"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1.50
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=1.5"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 1.75
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=1.75"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=2"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2.25
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=2.25"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2.50
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=2.5"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 2.75
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=2.75"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv")
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=3"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv")
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3.25
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=3.25"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3.50
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=3.5"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 3.75
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=3.75"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+betamval <- 4
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=4"))
+
+#count number of non-zero coefficinets, excluding the final four rows
+coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
+coefile <- head(coefile, -4)
+coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+nocoeff <- nrow(coefile2)
+
+#no presences used in model (less those used for testing)
+maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
+notrainingpoints <- maxresults$X.Training.samples
+testauc <- maxresults$Test.AUC
+trainauc <- maxresults$Training.AUC
+
+#natural log
+samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+samplepred$nat_log <- log(samplepred$Raw.prediction)
+sumsamplepred <- sum(samplepred$nat_log)
+
+for (i in 1:nrow(AIClist)){
+  if(AIClist$regularization[i] == betamval){
+    AIClist$Test.AUC[i] <- testauc
+    AIClist$Train.AUC[i] <- trainauc
+    AIClist$n[i] <- notrainingpoints
+    AIClist$coefficients[i] <- nocoeff
+    AIClist$sum.of.logs[i] <- sumsamplepred
+
+  }
+}
+
+AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
+AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))
+
+write.csv(AIClist, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+
 ```
 
-# beta value (regularization) selection
+``
 
-mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "replicatetype=crossvalidate", "replicates=5")) #add to samples because field study with biased data - see https://esajournals.onlinelibrary.wiley.com/doi/10.1890/10-1171.1.
-note this should be the default but keep here as a note
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE"))
+mod
+
+# this is after beta has been selected for
+ev <- evaluate(p = testp, a= testb, model = mod)
+str(ev)
+auc <- ev@auc
+cor <- ev@cor
+evaluation <- cbind(auc, cor)
+write.csv(evaluation, paste0("../output/maxent/models/", timeslice, "/eval.csv"), row.names = FALSE) 
+thres <- threshold(ev)
+write.csv(thres, paste0("../output/maxent/models/", timeslice, "/threshold.csv"), row.names = FALSE)
+```

### 2019-04-04 src/maxent_avec_xvalidation.Rmd
removal of auc testing as there is no testing at this stage
looped model selection

--- 2019-04-03 src/maxent_avec_xvalidation.Rmd
@@
 
-```{r train and test on 2003 inc AIC for reg val}
-train2003 <- subset(janprback, year != "2003")
-test2003 <-  subset(janprback, year == "2003")
-train2003pa <- train2003$occurrence
-test2003p <- subset(test2003, occurrence == "1")
-test2003b <- subset(test2003, occurrence == "0")
-train2003  <- subset(train2003, select = -c(year, occurrence))
+```{r train and test on jan inc AIC for reg val}
+modelmonth <- cbind(paste0("jan",unique(janprback$year)))
+AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
+betavallist <- cbind(AIClist$regularization)
 
-mod <- maxent(x = train2003, p = train2003pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE"))
-mod
-ev <- evaluate(p = test2003p, a= test2003b, model = mod)
-ev
-threshold(ev)
+for (year in 1:nrow(modelmonth)){
+  timeslice <- modelmonth[year] #testing timeslice
+  yr <- str_sub(timeslice, -4) #the year being used for testing
+  train <- subset(janprback, year != yr) #training data is all years but testing
+  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
+  test <-  subset(janprback, year == yr) #creates the testing dataset
+  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
+  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
+  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+
+  for (h in 1:nrow(betavallist)){
+    betamval <-  betavallist[h]
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
+
+    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
+
+    #count number of non-zero coefficients, excluding the final four rows
+    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    coefile <- head(coefile, -4)
+    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+    nocoeff <- nrow(coefile2)
+    bick <- sum(as.numeric(coefile2[, 2]))
+    bic <- (bick*log(sum(trainpa == 1)) - 2*11)
+
+    #no presences used in model (less those used for testing)
+    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    notrainingpoints <- maxresults$X.Training.samples
+    testauc <- maxresults$Test.AUC
+    trainauc <- maxresults$Training.AUC
+
+    #natural log
+    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    samplepred$nat_log <- log(samplepred$Raw.prediction)
+    sumsamplepred <- sum(samplepred$nat_log)
+
+    #test auc scores and other lovely things you might want to include
+    ev <- evaluate(p = testp, a= testb, model = mod)
+    evauc <- ev@auc
+    evcor <- ev@cor
+    evpcor <- ev@pcor
+    evnp <- ev@np
+    evna <- ev@na
+    evmaxtprtnr <- threshold(ev, 'spec_sens')
+    evmaxkap <- threshold(ev, 'kappa')
+    evno_omission <- threshold(ev, 'no_omission')
+    evprevalence <- threshold(ev, 'prevalence')
+    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+
+    for (i in 1:nrow(AIClist)){
+      if(AIClist$regularization[i] == betamval){
+      AIClist$Train.AUC[i] <- trainauc
+      AIClist$n[i] <- notrainingpoints
+      AIClist$coefficients[i] <- nocoeff
+      AIClist$sum.of.logs[i] <- sumsamplepred
+      AIClist$Test.AUC[i] <- evauc
+      AIClist$evcor[i] <- evcor
+      AIClist$evpcor[i] <- evpcor
+      AIClist$nopresence[i] <- evnp
+      AIClist$nobackground[i] <- evna
+      AIClist$maxsumsenspec[i] <- evmaxtprtnr
+      AIClist$maxsumsenspec[i] <- evmaxkap
+      AIClist$no_omission[i] <- evno_omission
+      AIClist$equal_sens_spec[i] <- evequal_sens_spec
+      AIClist$bic[i] <- bic #not sure this is correct...
+      }
+    }
+  }
+
+AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
+AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))
+
+write.csv(AIClist, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+
+}
+
 ```
 
-# beta value (regularization) selection
+next step... select the models you want for each year and run those models again (primarily based on lowest AICc but check test AUC and other things too)
 
-mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "replicatetype=crossvalidate", "replicates=5")) #add to samples because field study with biased data - see https://esajournals.onlinelibrary.wiley.com/doi/10.1890/10-1171.1.
-note this should be the default but keep here as a note
+jan2003: b = 1
+jan2004: b = 1
+jan2005: b = 1.5
+jan2006: b = 2.25
+
+for (year in 1:nrow(modelmonth)){
+  timeslice <- modelmonth[year] #testing timeslice
+  yr <- str_sub(timeslice, -4) #the year being used for testing
+  train <- subset(janprback, year != yr) #training data is all years but testing
+  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
+  test <-  subset(janprback, year == yr) #creates the testing dataset
+  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
+  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
+  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+
+  for (h in 1:nrow(betavallist)){
+    betamval <-  betavallist[h]
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
+
+    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
+
+    #count number of non-zero coefficients, excluding the final four rows
+    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    coefile <- head(coefile, -4)
+    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+    nocoeff <- nrow(coefile2)
+    bick <- sum(as.numeric(coefile2[, 2]))
+    bic <- (bick*log(sum(trainpa == 1)) - 2*11)
+
+    #no presences used in model (less those used for testing)
+    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    notrainingpoints <- maxresults$X.Training.samples
+    testauc <- maxresults$Test.AUC
+    trainauc <- maxresults$Training.AUC
+
+    #natural log
+    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    samplepred$nat_log <- log(samplepred$Raw.prediction)
+    sumsamplepred <- sum(samplepred$nat_log)
+
+    #test auc scores and other lovely things you might want to include
+    ev <- evaluate(p = testp, a= testb, model = mod)
+    evauc <- ev@auc
+    evcor <- ev@cor
+    evpcor <- ev@pcor
+    evnp <- ev@np
+    evna <- ev@na
+    evmaxtprtnr <- threshold(ev, 'spec_sens')
+    evmaxkap <- threshold(ev, 'kappa')
+    evno_omission <- threshold(ev, 'no_omission')
+    evprevalence <- threshold(ev, 'prevalence')
+    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+
+    for (i in 1:nrow(AIClist)){
+      if(AIClist$regularization[i] == betamval){
+      AIClist$Train.AUC[i] <- trainauc
+      AIClist$n[i] <- notrainingpoints
+      AIClist$coefficients[i] <- nocoeff
+      AIClist$sum.of.logs[i] <- sumsamplepred
+      AIClist$Test.AUC[i] <- evauc
+      AIClist$evcor[i] <- evcor
+      AIClist$evpcor[i] <- evpcor
+      AIClist$nopresence[i] <- evnp
+      AIClist$nobackground[i] <- evna
+      AIClist$maxsumsenspec[i] <- evmaxtprtnr
+      AIClist$maxsumsenspec[i] <- evmaxkap
+      AIClist$no_omission[i] <- evno_omission
+      AIClist$equal_sens_spec[i] <- evequal_sens_spec
+      AIClist$bic[i] <- bic #not sure this is correct...
+      }
+    }
+  }
+
+AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
+AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))
+
+write.csv(AIClist, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+
+}

### 2019-04-04 src/stack_test.Rmd
looped model selection

@@
 
-try changing some of the settings here - paerticularly you want aic
+try changing some of the settings here - particularly you want aic
 ```{r}

### 2019-04-04 src/maxent_avec_xvalidation.Rmd
selected model run

@@
 modelmonth <- cbind(paste0("jan",unique(janprback$year)))
-AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
-betavallist <- cbind(AIClist$regularization)
+aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
+betavallist <- cbind(aicetc$regularization)
 
@@
 
-    for (i in 1:nrow(AIClist)){
-      if(AIClist$regularization[i] == betamval){
-      AIClist$Train.AUC[i] <- trainauc
-      AIClist$n[i] <- notrainingpoints
-      AIClist$coefficients[i] <- nocoeff
-      AIClist$sum.of.logs[i] <- sumsamplepred
-      AIClist$Test.AUC[i] <- evauc
-      AIClist$evcor[i] <- evcor
-      AIClist$evpcor[i] <- evpcor
-      AIClist$nopresence[i] <- evnp
-      AIClist$nobackground[i] <- evna
-      AIClist$maxsumsenspec[i] <- evmaxtprtnr
-      AIClist$maxsumsenspec[i] <- evmaxkap
-      AIClist$no_omission[i] <- evno_omission
-      AIClist$equal_sens_spec[i] <- evequal_sens_spec
-      AIClist$bic[i] <- bic #not sure this is correct...
+    for (i in 1:nrow(aicetc)){
+      if(aicetc$regularization[i] == betamval){
+      aicetc$Train.AUC[i] <- trainauc
+      aicetc$n[i] <- notrainingpoints
+      aicetc$coefficients[i] <- nocoeff
+      aicetc$sum.of.logs[i] <- sumsamplepred
+      aicetc$Test.AUC[i] <- evauc
+      aicetc$evcor[i] <- evcor
+      aicetc$evpcor[i] <- evpcor
+      aicetc$nopresence[i] <- evnp
+      aicetc$nobackground[i] <- evna
+      aicetc$maxsumsenspec[i] <- evmaxtprtnr
+      aicetc$maxsumsenspec[i] <- evmaxkap
+      aicetc$no_omission[i] <- evno_omission
+      aicetc$equal_sens_spec[i] <- evequal_sens_spec
+      aicetc$bic[i] <- bic #not sure this is correct...
       }
@@
 
-AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
-AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))
+aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
+aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
 
-write.csv(AIClist, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
 
@@
 
+```{r}
+modelmonth <- cbind(paste0("jan",unique(janprback$year)))
+aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
 for (year in 1:nrow(modelmonth)){
@@
 
-  for (h in 1:nrow(betavallist)){
-    betamval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
+  if (timeslice == "jan2003"){
+    betaval <- 1
+  } else if (timeslice == "jan2004"){
+    betaval <- 1
+  } else if (timeslice == "jan2005"){
+    betaval <- 1.5
+  } else if (timeslice == "jan2006"){
+    betaval <- 2.25
+  }
+
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
 
@@
 
-    for (i in 1:nrow(AIClist)){
-      if(AIClist$regularization[i] == betamval){
-      AIClist$Train.AUC[i] <- trainauc
-      AIClist$n[i] <- notrainingpoints
-      AIClist$coefficients[i] <- nocoeff
-      AIClist$sum.of.logs[i] <- sumsamplepred
-      AIClist$Test.AUC[i] <- evauc
-      AIClist$evcor[i] <- evcor
-      AIClist$evpcor[i] <- evpcor
-      AIClist$nopresence[i] <- evnp
-      AIClist$nobackground[i] <- evna
-      AIClist$maxsumsenspec[i] <- evmaxtprtnr
-      AIClist$maxsumsenspec[i] <- evmaxkap
-      AIClist$no_omission[i] <- evno_omission
-      AIClist$equal_sens_spec[i] <- evequal_sens_spec
-      AIClist$bic[i] <- bic #not sure this is correct...
-      }
-    }
+      aicetc$Train.AUC <- trainauc
+      aicetc$n[i] <- notrainingpoints
+      aicetc$coefficients <- nocoeff
+      aicetc$sum.of.logs <- sumsamplepred
+      aicetc$Test.AUC <- evauc
+      aicetc$evcor <- evcor
+      aicetc$evpcor <- evpcor
+      aicetc$nopresence <- evnp
+      aicetc$nobackground <- evna
+      aicetc$maxsumsenspec <- evmaxtprtnr
+      aicetc$maxsumsenspec <- evmaxkap
+      aicetc$no_omission <- evno_omission
+      aicetc$equal_sens_spec <- evequal_sens_spec
+      aicetc$bic <- bic #not sure this is correct...
+      aicetc$beta <- betaval
+
+aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
+aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+
+write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+
 }
 
-AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
-AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))
-
-write.csv(AIClist, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
-
-}
+```

### 2019-04-04 src/asc_file_database.Rmd
asc list by month

@@
 ```
+
+it may be better to have ascii lists per month (since models are per month)
+
+```{r}
+a1998 <- read.csv("../data/env/asc_layerlist1998.csv", header = TRUE)
+a1999 <- read.csv("../data/env/asc_layerlist1999.csv", header = TRUE)
+a2000 <- read.csv("../data/env/asc_layerlist2000.csv", header = TRUE)
+a2001 <- read.csv("../data/env/asc_layerlist2001.csv", header = TRUE)
+a2002 <- read.csv("../data/env/asc_layerlist2002.csv", header = TRUE)
+a2003 <- read.csv("../data/env/asc_layerlist2003.csv", header = TRUE)
+a2004 <- read.csv("../data/env/asc_layerlist2004.csv", header = TRUE)
+a2005 <- read.csv("../data/env/asc_layerlist2005.csv", header = TRUE)
+a2006 <- read.csv("../data/env/asc_layerlist2006.csv", header = TRUE)
+a2007 <- read.csv("../data/env/asc_layerlist2007.csv", header = TRUE)
+a2008 <- read.csv("../data/env/asc_layerlist2008.csv", header = TRUE)
+a2009 <- read.csv("../data/env/asc_layerlist2009.csv", header = TRUE)
+a2010 <- read.csv("../data/env/asc_layerlist2010.csv", header = TRUE)
+a2011 <- read.csv("../data/env/asc_layerlist2011.csv", header = TRUE)
+a2012 <- read.csv("../data/env/asc_layerlist2012.csv", header = TRUE)
+a2013 <- read.csv("../data/env/asc_layerlist2013.csv", header = TRUE)
+a2014 <- read.csv("../data/env/asc_layerlist2014.csv", header = TRUE)
+a2015 <- read.csv("../data/env/asc_layerlist2015.csv", header = TRUE)
+
+as <- rbind(a1998, a1999, a2000, a2001, a2002, a2003, a2004, a2005, a2006, a2007, a2008, a2009, a2010, a2011, a2012, a2013, a2014, a2015)
+
+asm <- split(as, as$ascmonths)
+write.csv(asm$`1`, "../data/env/asc_layerlistjan.csv", row.names = TRUE)
+write.csv(asm$`2`, "../data/env/asc_layerlistfeb.csv", row.names = TRUE)
+write.csv(asm$`3`, "../data/env/asc_layerlistmar.csv", row.names = TRUE)
+write.csv(asm$`4`, "../data/env/asc_layerlistapr.csv", row.names = TRUE)
+write.csv(asm$`5`, "../data/env/asc_layerlistmay.csv", row.names = TRUE)
+write.csv(asm$`6`, "../data/env/asc_layerlistjun.csv", row.names = TRUE)
+write.csv(asm$`7`, "../data/env/asc_layerlistjul.csv", row.names = TRUE)
+write.csv(asm$`8`, "../data/env/asc_layerlistaug.csv", row.names = TRUE)
+write.csv(asm$`9`, "../data/env/asc_layerlistsep.csv", row.names = TRUE)
+write.csv(asm$`10`, "../data/env/asc_layerlistoct.csv", row.names = TRUE)
+write.csv(asm$`11`, "../data/env/asc_layerlistnov.csv", row.names = TRUE)
+write.csv(asm$`12`, "../data/env/asc_layerlistdec.csv", row.names = TRUE)
+```

### 2019-04-04 src/maxent_avec_xvalidation.Rmd
model creation and prediction with one variable change

@@
 
-```{r}
-modelmonth <- cbind(paste0("jan",unique(janprback$year)))
+```{r final models plus predict}
+mth <- "jan"
+modelmonth <- cbind(paste0(mth, unique(janprback$year)))
 aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
+unlayer_yr <- read.csv("../data/env/asc_layerlistjan.csv", header = TRUE)
+
 for (year in 1:nrow(modelmonth)){
@@
 
+    unlayer_yr <- subset(unlayer, ascyears == yr)
+    unlayer_yr <- unlayer_yr[-c(1), ]
+    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+    aea <- raster("../output/env/aea.tif")
+
+    for (td in 1:nrow(unlayer_yr)){
+      temp_depth <- raster(unlayer_yr$temp_depth[td])
+      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(temp_depth) <- "temp_depth"
+      temp_depth <- projectRaster(temp_depth, aea)
+      chl_surface <- raster(unlayer_yr$chl_surface[td])
+      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(chl_surface) <- "chl_surface"
+      chl_surface <- projectRaster(chl_surface, aea)
+      o2_depth <- raster(unlayer_yr$o2_depth[td])
+      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(o2_depth) <- "o2_depth"
+      o2_depth <- projectRaster(o2_depth, aea)
+      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
+      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(salinity_depth) <- "salinity_depth"
+      salinity_depth <- projectRaster(salinity_depth, aea)
+      nao_sample <- raster(unlayer_yr$nao_sample[td])
+      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(nao_sample) <- "nao_sample"
+      nao_sample <- projectRaster(nao_sample, aea)
+      nao_winter <- raster(unlayer_yr$nao_winter[td])
+      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      names(nao_winter) <- "nao_winter"
+      nao_winter <- projectRaster(nao_winter, aea)
+      stackn <- unlayer_yr$stackname[td]
+      stack <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+      #stackSave(test,unlayer_yr$stackname[td])
+      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mth, "/", stackn, ".asc"), overwrite=TRUE) 
+      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mth, "/", stackn, ".png"))) 
+      dev.off() # stops automatic saving of the plot to a png
+    }
 }
@@
 ```
+
+ok! now to average results

### 2019-04-04 src/maxent_avec_xvalidation.Rmd
next steps

--- 2019-04-04 src/maxent_avec_xvalidation.Rmd
@@
 ```{r train and test on jan inc AIC for reg val}
-modelmonth <- cbind(paste0("jan",unique(janprback$year)))
+mth <- "jan"
+modelmonth <- cbind(paste0(mth ,unique(janprback$year)))
 aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
@@
 
-```{r}
-modelmonth <- cbind(paste0("jan",unique(janprback$year)))
+```{r final models plus predict}
+mth <- "feb"
+modelmonth <- cbind(paste0(mth, unique(janprback$year)))
 aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
+unlayer <- read.csv(paste0("../data/env/asc_layerlist", mth, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
+
 for (year in 1:nrow(modelmonth)){
@@
 
+    unlayer_yr <- subset(unlayer, ascyears == yr)
+    unlayer_yr <- unlayer_yr[-c(1), ]
+    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+    aea <- raster("../output/env/aea.tif")
+
+    for (td in 1:nrow(unlayer_yr)){
+      temp_depth <- raster(unlayer_yr$temp_depth[td])
+      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(temp_depth) <- "temp_depth"
+      temp_depth <- projectRaster(temp_depth, aea)
+      chl_surface <- raster(unlayer_yr$chl_surface[td])
+      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(chl_surface) <- "chl_surface"
+      chl_surface <- projectRaster(chl_surface, aea)
+      o2_depth <- raster(unlayer_yr$o2_depth[td])
+      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(o2_depth) <- "o2_depth"
+      o2_depth <- projectRaster(o2_depth, aea)
+      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
+      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(salinity_depth) <- "salinity_depth"
+      salinity_depth <- projectRaster(salinity_depth, aea)
+      nao_sample <- raster(unlayer_yr$nao_sample[td])
+      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(nao_sample) <- "nao_sample"
+      nao_sample <- projectRaster(nao_sample, aea)
+      nao_winter <- raster(unlayer_yr$nao_winter[td])
+      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      names(nao_winter) <- "nao_winter"
+      nao_winter <- projectRaster(nao_winter, aea)
+      stackn <- unlayer_yr$stackname[td]
+      stack <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+      #stackSave(test,unlayer_yr$stackname[td])
+      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mth, "/", stackn, ".asc"), overwrite=TRUE) 
+      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mth, "/", stackn, ".png"))) 
+      dev.off() # stops automatic saving of the plot to a png
     }
+}
+```
 
-```
+ok! now to average results:
+- the asc files per depth layer
+- the AICforMaxentReg.csv 
+- the maxentResults.csv

### 2019-04-04 src/maxent_avec_xvalidation.Rmd
avg model results done

@@
 
+#model selection
 next step... select the models you want for each year and run those models again (primarily based on lowest AICc but check test AUC and other things too)
@@
 
+#model averageing
+
 ok! now to average results:
-- the asc files per depth layer
+
 - the AICforMaxentReg.csv 
 - the maxentResults.csv
+- the asc files per depth layer
+
+open all maxentReg files
+
+```{r avg maxentReg}
+#mth <- "jan"
+#modelmonth <- cbind(paste0(mth, unique(janprback$year)))
+for (i in 1:length(modelmonth)){
+  timeslice <- modelmonth[i]
+  yr <- str_sub(timeslice, -4) #the year being used for testing
+  if(yr == 2003){
+    aic2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    } else if (yr == 2004){
+    aic2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    } else if (yr == 2005){
+    aic2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    } else if (yr == 2006){
+    aic2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+  }
+}
+
+aic <- rbind(aic2003, aic2004, aic2005, aic2006)
+aicm <- colMeans(aic)
+aicm$mod <- "avg"
+avgAIC <- rbind(aic, aicm)
+
+write.csv(avgAIC, paste0("../output/maxent/models/", mth, "avgAIC.csv"), row.names = FALSE)
+```
+
+```{r avg maxres}
+#mth <- "jan"
+#modelmonth <- cbind(paste0(mth, unique(janprback$year)))
+for (i in 1:length(modelmonth)){
+  timeslice <- modelmonth[i]
+  yr <- str_sub(timeslice, -4) #the year being used for testing
+  if(yr == 2003){
+    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2004){
+    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2005){
+    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2006){
+    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+  }
+}
+
+maxr <- rbind(maxr2003, maxr2004, maxr2005, maxr2006)
+maxr$Species <- 1
+maxrm <- colMeans(maxr)
+maxrm$Species <- "avg"
+maxrm <- rbind(maxr, maxrm)
+
+write.csv(maxrm, paste0("../output/maxent/models/", mth, "avgmaxentresults.csv"), row.names = FALSE)
+```

### 2019-04-04 src/maxent_avec_xvalidation.Rmd
raster avging started

@@
 ```
+
+for this bit may need to stack rasters by depth... this will need some working with the name
+get unique filenames into a list
+ignore the first X characters in the filename (yr_mth_) and stack based on the last bit of the filename
+then do something probably with raster package
+
+```{r avg asc layers}
+asc_list <- list.files("../output/maxent/predictions/jan", pattern = '*.asc', full.names = TRUE) #true means the full path is included
+
+```

### 2019-04-05 src/maxent_avec_xvalidation.Rmd
asc_sum raster looping + terrible loop to make database

@@
     nocoeff <- nrow(coefile2)
-    bick <- sum(as.numeric(coefile2[, 2]))
-    bic <- (bick*log(sum(trainpa == 1)) - 2*11)
+    bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
+    bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
 
@@
 ```{r final models plus predict}
-mth <- "feb"
+mth <- "jan"
 modelmonth <- cbind(paste0(mth, unique(janprback$year)))
@@
       aicetc$Train.AUC <- trainauc
-      aicetc$n[i] <- notrainingpoints
+      aicetc$n <- notrainingpoints
       aicetc$coefficients <- nocoeff
@@
     unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
+    stackdeplst[[year]] <- stackdep
+    #write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
     aea <- raster("../output/env/aea.tif")
@@
 }
+
+stdep <- do.call(rbind, stackdeplst)
+write.csv(stdep, paste0("../output/maxent/predictions/", mth, "/stack_depth.csv"), row.names = FALSE)
 ```
@@
 ```{r avg asc layers}
-asc_list <- list.files("../output/maxent/predictions/jan", pattern = '*.asc', full.names = TRUE) #true means the full path is included
+asc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "*.asc", full.names = TRUE)
+depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
+names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
+depthlayers$depth <- as.character(depthlayers$depth)
+
+colno <- 1
+
+for (d in 1:nrow(depthlayers)){
+  dep <- depthlayers$depth[d]
+  for (a in 1:length(asc_list)){
+    asc <- asc_list[a]
+    if (grepl(dep, asc)){
+      depthlayers[d , paste0("a", colno)] <- asc
+        colno <- colno+1
+    }
+  }
+}
+write.csv(depthlayers, (paste0("../output/maxent/predictions/", mth, "/asclayerbydepth.csv")), row.names = TRUE)
+
+for (x in 1:nrow(depthlayers)){
+  asclay <- depthlayers[1, ]
+  asclay <- asclay[!is.na(asclay)]
+  asclay <- as.data.frame(rbind(asclay))
+  asclay$V1 <- as.character(asclay$V1)
+  asclay$V2 <- as.character(asclay$V2)
+  asclay$V3 <- as.character(asclay$V3)
+  asclay$V4 <- as.character(asclay$V4)
+  asclay$V5 <- as.character(asclay$V5)
+  asclaydep <- asclay$V1
+
+  asc2003 <- raster(asclay$V2)
+  asc2004 <- raster(asclay$V3)
+  asc2005 <- raster(asclay$V4)
+  asc2006 <- raster(asclay$V5)
+
+  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
+  avg <- calc(stkasc, fun = mean)
+
+  writeRaster(avg, (paste0("../output/maxent/predictions/", mth, "/avg_", asclaydep, ".asc")), overwrite=TRUE)
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mth, "/avg_", asclaydep, ".png"))) 
+  dev.off()
+}
 

### 2019-04-05 src/maxent_avec_xvalidation.Rmd
horrible raster stack code!

@@
 for (x in 1:nrow(depthlayers)){
-  asclay <- depthlayers[1, ]
+  asclay <- depthlayers[x, ]
   asclay <- asclay[!is.na(asclay)]
@@
 
+ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mth), glob2rx(pattern = "avg_*.asc"), full.names = TRUE)
+
+asc120  <- raster(paste0("../output/maxent/predictions/", mth, "/avg_120.asc"))
+asc41.18 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_41.18.asc"))
+asc1.55586 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_1.55586.asc"))
+asc1045.85 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_1045.85.asc"))
+asc108.03 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_108.03.asc"))
+asc11.7737 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_11.7737.asc"))
+asc1151.99 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_1151.99.asc"))
+asc1265.86 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_1265.86.asc"))
+asc13.991 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_13.991.asc"))
+asc133.076 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_133.076.asc"))
+asc1387.38 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_1387.38.asc"))
+asc147.406 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_147.406.asc"))
+asc16.5253 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_16.5253.asc"))
+asc163.165 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_163.165.asc"))
+asc180.55 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_180.55.asc"))
+asc19.4298 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_19.4298.asc"))
+asc199.79 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_199.79.asc"))
+asc2.66768 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_2.66768.asc"))
+asc22.7576 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_22.7576.asc"))
+asc221.141 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_221.141.asc"))
+asc244.891 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_244.891.asc"))
+asc26.5583 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_26.5583.asc"))
+asc271.356 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_271.356.asc"))
+asc3.85628 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_3.85628.asc"))
+asc30.8746 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_30.8746.asc"))
+asc300.888 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_300.888.asc"))
+asc333.863 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_333.863.asc"))
+asc35.7402 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_35.7402.asc"))
+asc370.689 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_370.689.asc"))
+asc411.794 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_411.794.asc"))
+asc457.626 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_457.626.asc"))
+asc47.2119 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_47.2119.asc"))
+asc5.14036 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_5.14036.asc"))
+asc508.64 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_508.64.asc"))
+asc53.8506 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_53.8506.asc"))
+asc565.292 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_565.292.asc"))
+asc6.54303 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_6.54303.asc"))
+asc61.1128 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_61.1128.asc"))
+asc628.026 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_628.026.asc"))
+asc69.0217 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_69.0217.asc"))
+asc697.259 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_697.259.asc"))
+asc77.6112 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_77.6112.asc"))
+asc773.368 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_773.368.asc"))
+asc8.09252 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_8.09252.asc"))
+asc856.679 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_856.679.asc"))
+asc86.9294 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_86.9294.asc"))
+asc9.82275 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_9.82275.asc"))
+asc947.448 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_947.448.asc"))
+asc97.0413 <- raster(paste0("../output/maxent/predictions/", mth, "/avg_97.0413.asc"))
+
+stackavg <- stack(asc120, asc41.18, asc1.55586, asc1045.85, asc108.03, asc11.7737, asc1151.99, asc1265.86, asc13.991, asc133.076, asc1387.38, asc147.406, asc16.5253, asc163.165, asc180.55, asc19.4298, asc199.79, asc2.66768, asc22.7576, asc221.141, asc244.891, asc26.5583, asc271.356, asc3.85628, asc30.8746, asc300.888, asc333.863, asc35.7402, asc370.689, asc411.794, asc457.626, asc47.2119, asc5.14036, asc508.64, asc53.8506, asc565.292, asc6.54303, asc61.1128, asc628.026, asc69.0217, asc697.259, asc77.6112, asc773.368, asc8.09252, asc856.679, asc86.9294, asc9.82275, asc947.448, asc97.0413)
+
+sum <- calc(stkasc, fun = sum)
+avg <- calc(stkasc, fun = mean)
+  writeRaster(avg, (paste0("../output/maxent/predictions/", mth, "/avg_", mth, ".asc")), overwrite=TRUE)
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mth, "/avg_", mth, ".png"))) 
+  dev.off()
+
 ```
+
+asc120
+asc41.18
+asc1.55586
+asc1045.85
+asc108.03
+asc11.7737
+asc1151.99
+asc1265.86
+asc13.991
+asc133.076
+asc1387.38
+asc147.406
+asc16.5253
+asc163.165
+asc180.55
+asc19.4298
+asc199.79
+asc2.66768
+asc22.7576
+asc221.141
+asc244.891
+asc26.5583
+asc271.356
+asc3.85628
+asc30.8746
+asc300.888
+asc333.863
+asc35.7402
+asc370.689
+asc411.794
+asc457.626
+asc47.2119
+asc5.14036
+asc508.64
+asc53.8506
+asc565.292
+asc6.54303
+asc61.1128
+asc628.026
+asc69.0217
+asc697.259
+asc77.6112
+asc773.368
+asc8.09252
+asc856.679
+asc86.9294
+asc9.82275
+asc947.448
+asc97.0413

### 2019-04-08 src/maxent_avec_xvalidation.Rmd
minor change to get cloglog output

@@
       #stackSave(test,unlayer_yr$stackname[td])
-      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mth, "/", stackn, ".asc"), overwrite=TRUE) 
+      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mth, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
       prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mth, "/", stackn, ".png"))) 
@@
 
-  writeRaster(avg, (paste0("../output/maxent/predictions/", mth, "/avg_", asclaydep, ".asc")), overwrite=TRUE)
+  writeRaster(avg, (paste0("../output/maxent/predictions/", mth, "/avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cloglog')
   avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mth, "/avg_", asclaydep, ".png"))) 

### 2019-04-11 src/data_exploration.Rmd
minor alteration to make one-stop date change for running models

@@
 ```{r pre-install packages, include=FALSE}
-packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car", "rworldmap", "fastmatch", "tables") 
+packages <- c("plyr", "graphics", "ggplot2", "corrplot", "FSA", "car", "rworldmap", "fastmatch", "tables", "dplyr") 
 source("../src/ipak.R")
@@
 ```{r}
-presab <- read.csv("../output/bio/presab_10k.csv", header = TRUE)
+presab <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
 presence <- subset(presab, occurrence == "1")
@@
 Temperature
-```{r}
+```{r presence temp stats}
 temp_depth_max <- max(presence$temp_depth, na.rm = TRUE)
@@
 Salinity
-```{r}
+```{r presence salinity stats}
 sal_depth_max <- max(presence$salinity_depth, na.rm = TRUE)
@@
 Chl
-```{r}
+```{r presence chl stats}
 chl_depth_max <- max(presence$chl_depth, na.rm = TRUE)
@@
 O2
-```{r}
+```{r presence o2 stats}
 o2_depth_max <- max(presence$o2_depth, na.rm = TRUE)
@@
 
+## plots of env. correlates by depth layer
+the idea here is to see if there are any stark differences between the niche occupied by capelin at different depths
+
+plot presences by unique depth & env. corr
+```{r plot presences by unique depth & env. corr}
+unique_depths <- unique(presence$depthlayer)
+unique_depths
+
+d1 <- subset(presence, depthlayerno == "1")
+d10 <- subset(presence, depthlayerno == "10")
+d11 <- subset(presence, depthlayerno == "11")
+d12 <- subset(presence, depthlayerno == "12")
+d13 <- subset(presence, depthlayerno == "13")
+d14 <- subset(presence, depthlayerno == "14")
+d15 <- subset(presence, depthlayerno == "15")
+d16 <- subset(presence, depthlayerno == "16")
+d17 <- subset(presence, depthlayerno == "17")
+d18 <- subset(presence, depthlayerno == "18")
+d19 <- subset(presence, depthlayerno == "19")
+d20 <- subset(presence, depthlayerno == "20")
+d21 <- subset(presence, depthlayerno == "21")
+d22 <- subset(presence, depthlayerno == "22")
+d23 <- subset(presence, depthlayerno == "23")
+d24 <- subset(presence, depthlayerno == "24")
+d25 <- subset(presence, depthlayerno == "25")
+d26 <- subset(presence, depthlayerno == "26")
+d27 <- subset(presence, depthlayerno == "27")
+d28 <- subset(presence, depthlayerno == "28")
+d29 <- subset(presence, depthlayerno == "29")
+d30 <- subset(presence, depthlayerno == "30")
+d31 <- subset(presence, depthlayerno == "31")
+d32 <- subset(presence, depthlayerno == "32")
+d33 <- subset(presence, depthlayerno == "33")
+d34 <- subset(presence, depthlayerno == "34")
+d35 <- subset(presence, depthlayerno == "35")
+d36 <- subset(presence, depthlayerno == "36")
+d37 <- subset(presence, depthlayerno == "37")
+d38 <- subset(presence, depthlayerno == "38")
+d39 <- subset(presence, depthlayerno == "39")
+d40 <- subset(presence, depthlayerno == "40")
+d41 <- subset(presence, depthlayerno == "41")
+d42 <- subset(presence, depthlayerno == "42")
+d43 <- subset(presence, depthlayerno == "43")
+d44 <- subset(presence, depthlayerno == "44")
+d45 <- subset(presence, depthlayerno == "45")
+d46 <- subset(presence, depthlayerno == "46")
+d47 <- subset(presence, depthlayerno == "47")
+
+gtemp <- ggplot(d1, aes(x = temp_depth, colour = depth_layer)) + geom_density(na.rm = TRUE) + geom_density(data=d13, na.rm = TRUE) + geom_density(data=d14, na.rm = TRUE) + geom_density(data=d15, na.rm = TRUE) + geom_density(data=d16, na.rm = TRUE) + geom_density(data=d17, na.rm = TRUE) + geom_density(data=d18, na.rm = TRUE) + geom_density(data=d19, na.rm = TRUE) + geom_density(data=d20, na.rm = TRUE) + geom_density(data=d21, na.rm = TRUE) + geom_density(data=d22, na.rm = TRUE) + geom_density(data=d23, na.rm = TRUE) + geom_density(data=d24, na.rm = TRUE) + geom_density(data=d25, na.rm = TRUE) + geom_density(data=d26, na.rm = TRUE) + geom_density(data=d27, na.rm = TRUE) + geom_density(data=d28, na.rm = TRUE) + geom_density(data=d29, na.rm = TRUE) + geom_density(data=d30, na.rm = TRUE) + geom_density(data=d31, na.rm = TRUE) + geom_density(data=d32, na.rm = TRUE) + geom_density(data=d33, na.rm = TRUE) + geom_density(data=d34, na.rm = TRUE) + geom_density(data=d35, na.rm = TRUE) + geom_density(data=d36, na.rm = TRUE) + geom_density(data=d37, na.rm = TRUE) + geom_density(data=d38, na.rm = TRUE) + geom_density(data=d39, na.rm = TRUE) + geom_density(data=d40, na.rm = TRUE) + geom_density(data=d41, na.rm = TRUE) + geom_density(data=d42, na.rm = TRUE)
+dev.copy(png,"../output/env/simple_plots/by_depth/temp_depth.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+#subtemp_depth <- subset(presence, select = c(depth_layer, temp_depth))
+
+tdep <- plot(presence$temp_depth, presence$depth_layer)
+
+```
+
+ggplot(nafo0a, aes(x = temp_surface, colour = nafo_zone)) + geom_density(na.rm = TRUE) + geom_density(data=nafo0b, na.rm = TRUE) + geom_density(data=nafo2g, na.rm = TRUE) + geom_density(data=nafo2h, na.rm = TRUE) + geom_density(data=nafo2j, na.rm = TRUE) + geom_density(data=nafo3k, na.rm = TRUE) + geom_density(data=nafo3l, na.rm = TRUE) + geom_density(data=nafo3n, na.rm = TRUE) + geom_density(data=nafo3o, na.rm = TRUE) + geom_density(data=nafo3ps, na.rm = TRUE) + geom_density(data=nafo4r, na.rm = TRUE) + geom_density(data=nafo4s, na.rm = TRUE) + geom_density(data=nafo4t, na.rm = TRUE) + geom_density(data=nafo4vn, na.rm = TRUE) + geom_density(data=nafo4vs, na.rm = TRUE) + geom_density(data=nafo4w, na.rm = TRUE) + geom_density(data=nafo4x, na.rm = TRUE) + geom_density(data=nafo5y, na.rm = TRUE) + geom_density(data=nafohudson, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_nafo_plots/temp_surface_nafo.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
 # background points
 
-# environmental correlates
+## environmental correlates
 
@@
 ```
-
-```{r}
-system.file("java", package="dismo") #finding location to put maxent jar file
-system("java -version")
-```
-
-messing with code to see if can do for unique year month depth
-```{r}
-ascname <- list.files("../data/jan_test_86.9", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks
-noasc <- length(ascname)
-  year <- "1998"
-  month <- "01"
-  depth <- "86.9294.asc"
-  tempvar <- "temp"
-  chlvar <- "chl"
-  o2var <- "o2"
-  salvar <- "salinity"
-  depthvar <- "depth"
-  surfacevar <- "surface"
-  naovar <- "nao"
-  amovar <- "amo"
-  samplevar <- "sample"
-  wintervar <- "winter"
-
-for (i in 1:length(asclist)){
-  ascyear <- (sapply(strsplit(ascname[i], "_"), "[[", 1))
-  ascmonth <- (sapply(strsplit(ascname[i], "_"), "[[", 2))
-  ascdepth <- (sapply(strsplit(ascname[i], "_"), "[[", 5))
-  ascvar1 <- (sapply(strsplit(ascname[i], "_"), "[[", 3))
-  ascvar2 <- (sapply(strsplit(ascname[i], "_"), "[[", 4))
-  ascfile <- ascname[i]
-
-  if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
-    #print(ascfile)
-    temp_depth <- raster(ascfile)
-    names(temp_depth) <- "temp_depth"
-  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
-    chl_surface <- raster(ascfile)
-    names(chl_surface) <- "chl_surface"
-  } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
-    o2_depth <- raster(ascfile)
-    names(o2_depth) <- "o2_depth"
-  } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
-    salinity_depth <- raster(ascfile)
-    names(salinity_depth) <- "salinity_depth"
-  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == samplevar){
-    nao_sample <- raster(ascfile)
-    names(nao_sample <- "nao_sample")
-  } else if (ascyear == year & ascmonth == month & ascvar1 == naovar & ascvar2 == wintervar){
-    nao_winter <- raster(ascfile)
-    names(nao_winter) <- "nao_winter"
-  }
-  st19980186.9 <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
-
-}
-```
-
-# maxentvariableselection package
-
-https://cran.r-project.org/web/packages/MaxentVariableSelection/MaxentVariableSelection.pdf

### 2019-04-11 src/maxent_avec_xvalidation.Rmd
minor alteration to make one-stop date change for running models

--- 2019-04-05 src/maxent_avec_xvalidation.Rmd
@@
 ---
-title: "maxent model avec xvalidation"
+title: "maxent model avec xvalidation (year-based test training)"
 author: "Samantha Andrews"
@@
 ```{r pre-install & load packages, include=FALSE}
-packages <- c("dismo", "stringr")
+packages <- c("dismo", "stringr", "rJava")
 source("../src/ipak.R")
@@
 
-jan variables to keep: temp_depth, salinty_depth, o2_depth, chl_surface, nao_sample, nao_winter 
-```{r january data}
-janprback <- subset(presback, month == "1")
-janprback <- subset(janprback, select = c(occurrence, year, temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_winter))
+```{r}
+janvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_winter")
+marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+
+mthnavar_lst <- list(jan = janvar, mar = marvar, apr = aprvar, may = mayvar, jun = junvar, jul = julvar, aug = augvar, sep = sepvar, oct = octvar, nov = novvar, dec = decvar)
+```
+
+select mont
+```{r month of interest -change}
+mthna <- "mar"
+mthno <- "3"
+```
+
+```{r subsest data by month}
+prback <- subset(presback, month == mthno)
+prback <- subset(prback, select = mthnavar_lst[[mthna]])
 ```
@@
 
-```{r jan remove na vals}
-janprback <- janprback[!is.na(janprback$temp_depth), ]
-janprback <- janprback[!is.na(janprback$salinity_depth), ]
-janprback <- janprback[!is.na(janprback$o2_depth), ]
-janprback <- janprback[!is.na(janprback$chl_surface), ]
-janprback <- janprback[!is.na(janprback$nao_sample), ]
-janprback <- janprback[!is.na(janprback$nao_winter), ]
+```{r remove na vals}
+prback <- na.omit(prback)
 ```
@@
 ```{r train and test on jan inc AIC for reg val}
-mth <- "jan"
-modelmonth <- cbind(paste0(mth ,unique(janprback$year)))
+modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
 aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
@@
   yr <- str_sub(timeslice, -4) #the year being used for testing
-  train <- subset(janprback, year != yr) #training data is all years but testing
+  train <- subset(prback, year != yr) #training data is all years but testing
   trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-  test <-  subset(janprback, year == yr) #creates the testing dataset
+  test <-  subset(prback, year == yr) #creates the testing dataset
   testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
@@
 ```{r final models plus predict}
-mth <- "jan"
-modelmonth <- cbind(paste0(mth, unique(janprback$year)))
+modelmonth <- cbind(paste0(mthna, unique(prback$year)))
 aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
-unlayer <- read.csv(paste0("../data/env/asc_layerlist", mth, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
+unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
 
@@
       #stackSave(test,unlayer_yr$stackname[td])
-      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mth, "/", stackn, ".asc"), overwrite=TRUE) 
-      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mth, "/", stackn, ".png"))) 
+      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
+      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", stackn, ".png"))) 
       dev.off() # stops automatic saving of the plot to a png
@@
 stdep <- do.call(rbind, stackdeplst)
-write.csv(stdep, paste0("../output/maxent/predictions/", mth, "/stack_depth.csv"), row.names = FALSE)
+write.csv(stdep, paste0("../output/maxent/predictions/", mthna, "/stack_depth.csv"), row.names = FALSE)
 ```
@@
 ```{r avg maxentReg}
-#mth <- "jan"
-#modelmonth <- cbind(paste0(mth, unique(janprback$year)))
+
 for (i in 1:length(modelmonth)){
@@
 
-write.csv(avgAIC, paste0("../output/maxent/models/", mth, "avgAIC.csv"), row.names = FALSE)
+write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
 ```
@@
 ```{r avg maxres}
-#mth <- "jan"
-#modelmonth <- cbind(paste0(mth, unique(janprback$year)))
+
 for (i in 1:length(modelmonth)){
@@
 
-write.csv(maxrm, paste0("../output/maxent/models/", mth, "avgmaxentresults.csv"), row.names = FALSE)
+write.csv(maxrm, paste0("../output/maxent/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
 ```
@@
 get unique filenames into a list
-ignore the first X characters in the filename (yr_mth_) and stack based on the last bit of the filename
+ignore the first X characters in the filename (yr_mthna_) and stack based on the last bit of the filename
 then do something probably with raster package
@@
 ```{r avg asc layers}
-asc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "*.asc", full.names = TRUE)
+asc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "*.asc", full.names = TRUE)
 depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
@@
 }
-write.csv(depthlayers, (paste0("../output/maxent/predictions/", mth, "/asclayerbydepth.csv")), row.names = TRUE)
+write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/asclayerbydepth.csv")), row.names = TRUE)
 
 for (x in 1:nrow(depthlayers)){
-  asclay <- depthlayers[1, ]
+  asclay <- depthlayers[x, ]
   asclay <- asclay[!is.na(asclay)]
@@
 
-  writeRaster(avg, (paste0("../output/maxent/predictions/", mth, "/avg_", asclaydep, ".asc")), overwrite=TRUE)
-  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mth, "/avg_", asclaydep, ".png"))) 
+  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cloglog')
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".png"))) 
   dev.off()
@@
 
+ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), glob2rx(pattern = "avg_*.asc"), full.names = TRUE)
+
+asc120  <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_120.asc"))
+asc41.18 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_41.18.asc"))
+asc1.55586 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1.55586.asc"))
+asc1045.85 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1045.85.asc"))
+asc108.03 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_108.03.asc"))
+asc11.7737 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_11.7737.asc"))
+asc1151.99 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1151.99.asc"))
+asc1265.86 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1265.86.asc"))
+asc13.991 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_13.991.asc"))
+asc133.076 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_133.076.asc"))
+asc1387.38 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1387.38.asc"))
+asc147.406 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_147.406.asc"))
+asc16.5253 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_16.5253.asc"))
+asc163.165 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_163.165.asc"))
+asc180.55 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_180.55.asc"))
+asc19.4298 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_19.4298.asc"))
+asc199.79 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_199.79.asc"))
+asc2.66768 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_2.66768.asc"))
+asc22.7576 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_22.7576.asc"))
+asc221.141 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_221.141.asc"))
+asc244.891 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_244.891.asc"))
+asc26.5583 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_26.5583.asc"))
+asc271.356 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_271.356.asc"))
+asc3.85628 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_3.85628.asc"))
+asc30.8746 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_30.8746.asc"))
+asc300.888 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_300.888.asc"))
+asc333.863 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_333.863.asc"))
+asc35.7402 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_35.7402.asc"))
+asc370.689 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_370.689.asc"))
+asc411.794 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_411.794.asc"))
+asc457.626 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_457.626.asc"))
+asc47.2119 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_47.2119.asc"))
+asc5.14036 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_5.14036.asc"))
+asc508.64 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_508.64.asc"))
+asc53.8506 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_53.8506.asc"))
+asc565.292 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_565.292.asc"))
+asc6.54303 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_6.54303.asc"))
+asc61.1128 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_61.1128.asc"))
+asc628.026 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_628.026.asc"))
+asc69.0217 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_69.0217.asc"))
+asc697.259 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_697.259.asc"))
+asc77.6112 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_77.6112.asc"))
+asc773.368 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_773.368.asc"))
+asc8.09252 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_8.09252.asc"))
+asc856.679 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_856.679.asc"))
+asc86.9294 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_86.9294.asc"))
+asc9.82275 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_9.82275.asc"))
+asc947.448 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_947.448.asc"))
+asc97.0413 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_97.0413.asc"))
+
+stackavg <- stack(asc120, asc41.18, asc1.55586, asc1045.85, asc108.03, asc11.7737, asc1151.99, asc1265.86, asc13.991, asc133.076, asc1387.38, asc147.406, asc16.5253, asc163.165, asc180.55, asc19.4298, asc199.79, asc2.66768, asc22.7576, asc221.141, asc244.891, asc26.5583, asc271.356, asc3.85628, asc30.8746, asc300.888, asc333.863, asc35.7402, asc370.689, asc411.794, asc457.626, asc47.2119, asc5.14036, asc508.64, asc53.8506, asc565.292, asc6.54303, asc61.1128, asc628.026, asc69.0217, asc697.259, asc77.6112, asc773.368, asc8.09252, asc856.679, asc86.9294, asc9.82275, asc947.448, asc97.0413)
+
+sum <- calc(stkasc, fun = sum)
+avg <- calc(stkasc, fun = mean)
+  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".asc")), overwrite=TRUE)
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".png"))) 
+  dev.off()
+
 ```
+
+asc120
+asc41.18
+asc1.55586
+asc1045.85
+asc108.03
+asc11.7737
+asc1151.99
+asc1265.86
+asc13.991
+asc133.076
+asc1387.38
+asc147.406
+asc16.5253
+asc163.165
+asc180.55
+asc19.4298
+asc199.79
+asc2.66768
+asc22.7576
+asc221.141
+asc244.891
+asc26.5583
+asc271.356
+asc3.85628
+asc30.8746
+asc300.888
+asc333.863
+asc35.7402
+asc370.689
+asc411.794
+asc457.626
+asc47.2119
+asc5.14036
+asc508.64
+asc53.8506
+asc565.292
+asc6.54303
+asc61.1128
+asc628.026
+asc69.0217
+asc697.259
+asc77.6112
+asc773.368
+asc8.09252
+asc856.679
+asc86.9294
+asc9.82275
+asc947.448
+asc97.0413

### 2019-04-12 src/maxent_avec_xvalidation.Rmd
start listing reg values needed
minor updates to code for march

@@
 ```{r pre-install & load packages, include=FALSE}
+options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
 packages <- c("dismo", "stringr", "rJava")
@@
 
-jan2003: b = 1
-jan2004: b = 1
-jan2005: b = 1.5
-jan2006: b = 2.25
+KEEP A NOTE OF THESE IN ONENOTE
+
+mar2003 b = 1
+
+mar1999 b = 1.25
+mar2002 b = 1.25
+mar2008 b = 1.25
+mar2009 b = 1.25
+
+mar2005 b = 1.5
+mar2006 b = 1.5
+
+mar2000 b = 1.75
+
+mar2001 b= 2.25
+mar2007 b = 2.25
+
+mar2010 b = 3.5
 
@@
 
-  if (timeslice == "jan2003"){
+  if (timeslice == "mar2003"){
     betaval <- 1
-  } else if (timeslice == "jan2004"){
-    betaval <- 1
-  } else if (timeslice == "jan2005"){
+  } else if (timeslice == "mar1999"| timeslice == "mar2002" | timeslice == "mar2008" | timeslice == "mar2009"){
+    betaval <- 1.25
+  } else if (timeslice == "mar2005" | timeslice == "mar2006"){
     betaval <- 1.5
-  } else if (timeslice == "jan2006"){
+  } else if (timeslice == "mar2000"){
+    betaval <- 1.75
+  } else if (timeslice == "mar2001" | timeslice == "mar2007"){
     betaval <- 2.25
+  } else if (timeslice == "mar2010"){
+    betaval <- 3.5
   }
@@
     unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+    stackdeplst <- list()
     stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
@@
       nao_sample <- projectRaster(nao_sample, aea)
+      nao_prev <- raster(unlayer_yr$nao_prev[td])
+      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(nao_prev) <- "nao_prev"
+      nao_prev <- projectRaster(nao_prev, aea)
       nao_winter <- raster(unlayer_yr$nao_winter[td])
@@
       nao_winter <- projectRaster(nao_winter, aea)
+      amo_sample <- raster(unlayer_yr$amo_sample[td])
+      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(amo_sample) <- "amo_sample"
+      amo_sample <- projectRaster(amo_sample, aea)
+      amo_winter <- raster(unlayer_yr$amo_winter[td])
+      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      names(amo_winter) <- "amo_winter"
+      amo_winter <- projectRaster(amo_winter, aea)
       stackn <- unlayer_yr$stackname[td]
-      stack <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
+      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
       #stackSave(test,unlayer_yr$stackname[td])

### 2019-04-12 src/maxent_avec_xvalidation.Rmd
april

@@
 options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava")
+packages <- c("dismo", "stringr", "rJava", "raster")
 source("../src/ipak.R")
@@
 ```{r month of interest -change}
-mthna <- "mar"
-mthno <- "3"
+mthna <- "apr"
+mthno <- "4"
 ```
@@
 
-mar2003 b = 1
-
-mar1999 b = 1.25
-mar2002 b = 1.25
-mar2008 b = 1.25
-mar2009 b = 1.25
-
-mar2005 b = 1.5
-mar2006 b = 1.5
-
-mar2000 b = 1.75
-
-mar2001 b= 2.25
-mar2007 b = 2.25
-
-mar2010 b = 3.5
+y1998 = 2.75
+y1999 = 2.25
+y2000 = 2
+y2001 = 1.75
+y2002 = 2
+y2003 = 2
+y2004 = 2
+y2005 = 2.25
+y2006 = 1.75
+y2007 = 1.75
+y2008 = 2.25
+y2009 = 2.25
+y2010 = 2.5
+y2011 =  2.5
 
@@
 
-  if (timeslice == "mar2003"){
-    betaval <- 1
-  } else if (timeslice == "mar1999"| timeslice == "mar2002" | timeslice == "mar2008" | timeslice == "mar2009"){
-    betaval <- 1.25
-  } else if (timeslice == "mar2005" | timeslice == "mar2006"){
-    betaval <- 1.5
-  } else if (timeslice == "mar2000"){
+  if (timeslice == "y2001" | timeslice == "y2006" | timeslice == "y2007"){
     betaval <- 1.75
-  } else if (timeslice == "mar2001" | timeslice == "mar2007"){
+  } else if (timeslice == "y2000"| timeslice == "y2002" | timeslice == "y2003" | timeslice == "y2004"){
+    betaval <- 2
+  } else if (timeslice == "y1999" | timeslice == "y2005" | timeslice == "y2008" | timeslice == "y2009"){
     betaval <- 2.25
-  } else if (timeslice == "mar2010"){
-    betaval <- 3.5
+  } else if (timeslice == "y2010" | timeslice == "y2011"){
+    betaval <- 2.5
+  } else if (timeslice == "y1998"){
+    betaval <- 2.75
   }
@@
   yr <- str_sub(timeslice, -4) #the year being used for testing
-  if(yr == 2003){
+  if(yr == 1998){
+    aic1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 1999){
+    aic1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2000){
+    aic2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2001){
+    aic2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2002){
+    aic2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2003){
     aic2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2004){
     aic2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2005){
     aic2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2006){
     aic2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2007){
+    aic2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2008){
+    aic2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2009){
+    aic2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2010){
+    aic2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2011){
+    aic2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2012){
+    aic2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2013){
+    aic2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2014){
+    aic2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    } else if (yr == 2015){
+    aic2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
   }
@@
 
-aic <- rbind(aic2003, aic2004, aic2005, aic2006)
+aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
+maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)
+
 aicm <- colMeans(aic)
@@
 
-write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
-```
-
-```{r avg maxres}
-
-for (i in 1:length(modelmonth)){
-  timeslice <- modelmonth[i]
-  yr <- str_sub(timeslice, -4) #the year being used for testing
-  if(yr == 2003){
-    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2004){
-    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2005){
-    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2006){
-    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-  }
-}
-
-maxr <- rbind(maxr2003, maxr2004, maxr2005, maxr2006)
 maxr$Species <- 1
@@
 
+write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
 write.csv(maxrm, paste0("../output/maxent/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
@@
   asclay <- as.data.frame(rbind(asclay))
-  asclay$V1 <- as.character(asclay$V1)
-  asclay$V2 <- as.character(asclay$V2)
+  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
+  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
   asclay$V3 <- as.character(asclay$V3)
@@
   asclay$V5 <- as.character(asclay$V5)
+  asclay$V6 <- as.character(asclay$V6)
+  asclay$V7 <- as.character(asclay$V7)
+  asclay$V8 <- as.character(asclay$V8)
+  asclay$V9 <- as.character(asclay$V9)
+  asclay$V10 <- as.character(asclay$V10)
+  asclay$V11 <- as.character(asclay$V11)
+  asclay$V12 <- as.character(asclay$V12)
+  asclay$V13 <- as.character(asclay$V13)
+  asclay$V14 <- as.character(asclay$V14)
+  asclay$V15 <- as.character(asclay$V15)
+
   asclaydep <- asclay$V1
 
-  asc2003 <- raster(asclay$V2)
-  asc2004 <- raster(asclay$V3)
-  asc2005 <- raster(asclay$V4)
-  asc2006 <- raster(asclay$V5)
+  asc1998 <- raster(asclay$V2)
+  asc1999 <- raster(asclay$V3)
+  asc2000 <- raster(asclay$V4)
+  asc2001 <- raster(asclay$V5)
+  asc2002 <- raster(asclay$V6)
+  asc2003 <- raster(asclay$V7)
+  asc2004 <- raster(asclay$V8)
+  asc2005 <- raster(asclay$V9)
+  asc2006 <- raster(asclay$V10)
+  asc2007 <- raster(asclay$V11)
+  asc2008 <- raster(asclay$V12)
+  asc2009 <- raster(asclay$V13)
+  asc2010 <- raster(asclay$V14)
+  asc2011 <- raster(asclay$V15)
 
-  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
+  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)
   avg <- calc(stkasc, fun = mean)
@@
 
-sum <- calc(stkasc, fun = sum)
-avg <- calc(stkasc, fun = mean)
-  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".asc")), overwrite=TRUE)
-  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".png"))) 
+#stsum <- calc(stkasc, fun=sum)
+stavg <- calc(stkasc, fun = mean)
+  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".asc")), overwrite=TRUE)
+  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".png"))) 
   dev.off()

### 2019-04-12 src/maxent_avec_xvalidation.Rmd
minor change to avg

@@
 #stsum <- calc(stkasc, fun=sum)
-stavg <- calc(stkasc, fun = mean)
+stavg <- calc(stackavg, fun = mean, na.rm = TRUE)
   writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".asc")), overwrite=TRUE)

### 2019-04-12 src/maxent_avec_xvalidation.Rmd
minor update to sum

--- 2019-04-12 src/maxent_avec_xvalidation.Rmd
@@
 
-#stsum <- calc(stkasc, fun=sum)
-stavg <- calc(stkasc, fun = mean)
+stsum <- sum(stackavg, na.rm = TRUE)
+  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/sum_", mthna, ".asc")), overwrite=TRUE)
+  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/sum_", mthna, ".png"))) 
+  dev.off()
+stavg <- calc(stackavg, fun = mean, na.rm = TRUE)
   writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".asc")), overwrite=TRUE)

### 2019-04-15 src/maxent_avec_xvalidation.Rmd
minor change to get min pred area

@@
 options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster")
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat")
 source("../src/ipak.R")
@@
 
-```{r train and test on jan inc AIC for reg val}
+```{r train and test inc AIC for reg val}
 modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
@@
     evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+
+    #calculate minimal predicted area using ecospat
+    aicect$mpa <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
 
@@
       aicetc$maxsumsenspec <- evmaxtprtnr
-      aicetc$maxsumsenspec <- evmaxkap
+      aicetc$maxkap <- evmaxkap
       aicetc$no_omission <- evno_omission
+      aicetc$prevalence <- evprevalence
       aicetc$equal_sens_spec <- evequal_sens_spec
@@
     aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+
+    #calculate minimal predicted area using ecospat
+    aicetc$mpa <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
 
@@
       #stackSave(test,unlayer_yr$stackname[td])
-      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
-      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", stackn, ".png"))) 
+     # pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
+      predcum <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cumulative') 
+     # prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", stackn, ".png"))) 
+      prplotcum <- plot(predcum, png(paste0("../output/maxent/predictions/", mthna, "/", stackn, ".png"))) 
       dev.off() # stops automatic saving of the plot to a png

### 2019-04-16 src/maxent_avec_xvalidation.Rmd
playing with aic for paper 2... doesnt work (cant loop through the folds)

@@
     evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+    evsensitivity <- threshold(ev, 'sensitivity')
+
+      aicetc$train_AUC <- trainauc
+      aicetc$no_trainng_points <- notrainingpoints
+      aicetc$test_coefficients <- nocoeff
+      aicetc$test_sum.of.logs <- sumsamplepred
+      aicetc$test_AUC <- evauc
+      aicetc$test_evcor <- evcor
+      aicetc$test_evpcor <- evpcor
+      aicetc$test_nopresence <- evnp
+      aicetc$test_nobackground <- evna
+      aicetc$test_maxsumsenspec <- evmaxtprtnr
+      aicetc$test_maxkap <- evmaxkap
+      aicetc$test_no_omission <- evno_omission
+      aicetc$test_prevalence <- evprevalence
+      aicetc$test_equal_sens_spec <- evequal_sens_spec
+      aicetc$test_sensitivity <- evsensitivity
+      aicetc$test_bic <- bic #not sure this is correct...
+      aicetc$beta <- betaval
+
+    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
+    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
 
     #calculate minimal predicted area using ecospat
-    aicect$mpa <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
-
-    for (i in 1:nrow(aicetc)){
-      if(aicetc$regularization[i] == betamval){
-      aicetc$Train.AUC[i] <- trainauc
-      aicetc$n[i] <- notrainingpoints
-      aicetc$coefficients[i] <- nocoeff
-      aicetc$sum.of.logs[i] <- sumsamplepred
-      aicetc$Test.AUC[i] <- evauc
-      aicetc$evcor[i] <- evcor
-      aicetc$evpcor[i] <- evpcor
-      aicetc$nopresence[i] <- evnp
-      aicetc$nobackground[i] <- evna
-      aicetc$maxsumsenspec[i] <- evmaxtprtnr
-      aicetc$maxsumsenspec[i] <- evmaxkap
-      aicetc$no_omission[i] <- evno_omission
-      aicetc$equal_sens_spec[i] <- evequal_sens_spec
-      aicetc$bic[i] <- bic #not sure this is correct...
-      }
+    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
       }
@@
     evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+    evsensitivity <- threshold(ev, 'sensitivity')
 
-      aicetc$Train.AUC <- trainauc
-      aicetc$n <- notrainingpoints
-      aicetc$coefficients <- nocoeff
-      aicetc$sum.of.logs <- sumsamplepred
-      aicetc$Test.AUC <- evauc
-      aicetc$evcor <- evcor
-      aicetc$evpcor <- evpcor
-      aicetc$nopresence <- evnp
-      aicetc$nobackground <- evna
-      aicetc$maxsumsenspec <- evmaxtprtnr
-      aicetc$maxkap <- evmaxkap
-      aicetc$no_omission <- evno_omission
-      aicetc$prevalence <- evprevalence
-      aicetc$equal_sens_spec <- evequal_sens_spec
-      aicetc$bic <- bic #not sure this is correct...
+      aicetc$train_AUC <- trainauc
+      aicetc$no_trainng_points <- notrainingpoints
+      aicetc$test_coefficients <- nocoeff
+      aicetc$test_sum.of.logs <- sumsamplepred
+      aicetc$test_AUC <- evauc
+      aicetc$test_evcor <- evcor
+      aicetc$test_evpcor <- evpcor
+      aicetc$test_nopresence <- evnp
+      aicetc$test_nobackground <- evna
+      aicetc$test_maxsumsenspec <- evmaxtprtnr
+      aicetc$test_maxkap <- evmaxkap
+      aicetc$test_no_omission <- evno_omission
+      aicetc$test_prevalence <- evprevalence
+      aicetc$test_equal_sens_spec <- evequal_sens_spec
+      aicetc$test_sensitivity <- evsensitivity
+      aicetc$test_bic <- bic #not sure this is correct...
       aicetc$beta <- betaval
@@
     #calculate minimal predicted area using ecospat
-    aicetc$mpa <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
+    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
 
@@
       #stackSave(test,unlayer_yr$stackname[td])
-     # pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
-      predcum <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cumulative') 
-     # prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", stackn, ".png"))) 
-      prplotcum <- plot(predcum, png(paste0("../output/maxent/predictions/", mthna, "/", stackn, ".png"))) 
+      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
+      predcum <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
+      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
+      dev.off() # stops automatic saving of the plot to a png
+      prplotcum <- plot(predcum, png(paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".png"))) 
       dev.off() # stops automatic saving of the plot to a png
@@
 
-```{r avg asc layers}
-asc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "*.asc", full.names = TRUE)
+```{r avg asc layers cloglog}
+asc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "cloglog*.asc", full.names = TRUE)
 depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
@@
 }
-write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/asclayerbydepth.csv")), row.names = TRUE)
+write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)
 
@@
 
-asc120  <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_120.asc"))
-asc41.18 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_41.18.asc"))
-asc1.55586 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1.55586.asc"))
-asc1045.85 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1045.85.asc"))
-asc108.03 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_108.03.asc"))
-asc11.7737 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_11.7737.asc"))
-asc1151.99 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1151.99.asc"))
-asc1265.86 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1265.86.asc"))
-asc13.991 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_13.991.asc"))
-asc133.076 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_133.076.asc"))
-asc1387.38 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1387.38.asc"))
-asc147.406 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_147.406.asc"))
-asc16.5253 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_16.5253.asc"))
-asc163.165 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_163.165.asc"))
-asc180.55 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_180.55.asc"))
-asc19.4298 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_19.4298.asc"))
-asc199.79 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_199.79.asc"))
-asc2.66768 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_2.66768.asc"))
-asc22.7576 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_22.7576.asc"))
-asc221.141 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_221.141.asc"))
-asc244.891 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_244.891.asc"))
-asc26.5583 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_26.5583.asc"))
-asc271.356 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_271.356.asc"))
-asc3.85628 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_3.85628.asc"))
-asc30.8746 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_30.8746.asc"))
-asc300.888 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_300.888.asc"))
-asc333.863 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_333.863.asc"))
-asc35.7402 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_35.7402.asc"))
-asc370.689 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_370.689.asc"))
-asc411.794 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_411.794.asc"))
-asc457.626 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_457.626.asc"))
-asc47.2119 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_47.2119.asc"))
-asc5.14036 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_5.14036.asc"))
-asc508.64 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_508.64.asc"))
-asc53.8506 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_53.8506.asc"))
-asc565.292 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_565.292.asc"))
-asc6.54303 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_6.54303.asc"))
-asc61.1128 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_61.1128.asc"))
-asc628.026 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_628.026.asc"))
-asc69.0217 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_69.0217.asc"))
-asc697.259 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_697.259.asc"))
-asc77.6112 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_77.6112.asc"))
-asc773.368 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_773.368.asc"))
-asc8.09252 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_8.09252.asc"))
-asc856.679 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_856.679.asc"))
-asc86.9294 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_86.9294.asc"))
-asc9.82275 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_9.82275.asc"))
-asc947.448 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_947.448.asc"))
-asc97.0413 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_97.0413.asc"))
+clog_asc120  <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_120.asc"))
+clog_asc41.18 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_41.18.asc"))
+clog_asc1.55586 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1.55586.asc"))
+clog_asc1045.85 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1045.85.asc"))
+clog_asc108.03 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_108.03.asc"))
+clog_asc11.7737 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_11.7737.asc"))
+clog_asc1151.99 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1151.99.asc"))
+clog_asc1265.86 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1265.86.asc"))
+clog_asc13.991 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_13.991.asc"))
+clog_asc133.076 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_133.076.asc"))
+clog_asc1387.38 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1387.38.asc"))
+clog_asc147.406 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_147.406.asc"))
+clog_asc16.5253 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_16.5253.asc"))
+clog_asc163.165 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_163.165.asc"))
+clog_asc180.55 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_180.55.asc"))
+clog_asc19.4298 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_19.4298.asc"))
+clog_asc199.79 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_199.79.asc"))
+clog_asc2.66768 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_2.66768.asc"))
+clog_asc22.7576 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_22.7576.asc"))
+clog_asc221.141 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_221.141.asc"))
+clog_asc244.891 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_244.891.asc"))
+clog_asc26.5583 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_26.5583.asc"))
+clog_asc271.356 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_271.356.asc"))
+clog_asc3.85628 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_3.85628.asc"))
+clog_asc30.8746 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_30.8746.asc"))
+clog_asc300.888 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_300.888.asc"))
+clog_asc333.863 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_333.863.asc"))
+clog_asc35.7402 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_35.7402.asc"))
+clog_asc370.689 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_370.689.asc"))
+clog_asc411.794 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_411.794.asc"))
+clog_asc457.626 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_457.626.asc"))
+clog_asc47.2119 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_47.2119.asc"))
+clog_asc5.14036 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_5.14036.asc"))
+clog_asc508.64 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_508.64.asc"))
+clog_asc53.8506 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_53.8506.asc"))
+clog_asc565.292 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_565.292.asc"))
+clog_asc6.54303 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_6.54303.asc"))
+clog_asc61.1128 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_61.1128.asc"))
+clog_asc628.026 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_628.026.asc"))
+clog_asc69.0217 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_69.0217.asc"))
+clog_asc697.259 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_697.259.asc"))
+clog_asc77.6112 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_77.6112.asc"))
+clog_asc773.368 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_773.368.asc"))
+clog_asc8.09252 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_8.09252.asc"))
+clog_asc856.679 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_856.679.asc"))
+clog_asc86.9294 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_86.9294.asc"))
+clog_asc9.82275 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_9.82275.asc"))
+clog_asc947.448 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_947.448.asc"))
+clog_asc97.0413 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_97.0413.asc"))
 
-stackavg <- stack(asc120, asc41.18, asc1.55586, asc1045.85, asc108.03, asc11.7737, asc1151.99, asc1265.86, asc13.991, asc133.076, asc1387.38, asc147.406, asc16.5253, asc163.165, asc180.55, asc19.4298, asc199.79, asc2.66768, asc22.7576, asc221.141, asc244.891, asc26.5583, asc271.356, asc3.85628, asc30.8746, asc300.888, asc333.863, asc35.7402, asc370.689, asc411.794, asc457.626, asc47.2119, asc5.14036, asc508.64, asc53.8506, asc565.292, asc6.54303, asc61.1128, asc628.026, asc69.0217, asc697.259, asc77.6112, asc773.368, asc8.09252, asc856.679, asc86.9294, asc9.82275, asc947.448, asc97.0413)
+stackavg <- stack(clog_asc120, clog_asc41.18, clog_asc1.55586, clog_asc1045.85, clog_asc108.03, clog_asc11.7737, clog_asc1151.99, clog_asc1265.86, clog_asc13.991, clog_asc133.076, clog_asc1387.38, clog_asc147.406, clog_asc16.5253, clog_asc163.165, clog_asc180.55, clog_asc19.4298, clog_asc199.79, clog_asc2.66768, clog_asc22.7576, clog_asc221.141, clog_asc244.891, clog_asc26.5583, clog_asc271.356, clog_asc3.85628, clog_asc30.8746, clog_asc300.888, clog_asc333.863, clog_asc35.7402, clog_asc370.689, clog_asc411.794, clog_asc457.626, clog_asc47.2119, clog_asc5.14036, clog_asc508.64, clog_asc53.8506, clog_asc565.292, clog_asc6.54303, clog_asc61.1128, clog_asc628.026, clog_asc69.0217, clog_asc697.259, clog_asc77.6112, clog_asc773.368, clog_asc8.09252, clog_asc856.679, clog_asc86.9294, clog_asc9.82275, clog_asc947.448, clog_asc97.0413)
 
 stsum <- sum(stackavg, na.rm = TRUE)
-  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/sum_", mthna, ".asc")), overwrite=TRUE)
-  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/sum_", mthna, ".png"))) 
+  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/clog_sum_", mthna, ".asc")), overwrite=TRUE)
+  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/clog_sum_", mthna, ".png"))) 
   dev.off()
 stavg <- calc(stackavg, fun = mean, na.rm = TRUE)
-  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".asc")), overwrite=TRUE)
-  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".png"))) 
+  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/clog_avg_", mthna, ".asc")), overwrite=TRUE)
+  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/clog_avg_", mthna, ".png"))) 
   dev.off()

### 2019-04-16 src/paper2_maxent_avec_xvalidation.Rmd
paper 2 model with xval changed
playing with aic for paper 2... doesnt work (cant loop through the folds)

--- 2019-04-12 src/maxent_avec_xvalidation.Rmd
@@
 presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+
+#remove unnecessary columns for maxent
+presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt))
 colnames(presback)
@@
 
-```{r remove unnecessary columns}
-presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt))
-colnames(presback)
-```
-
 #monthly model data selection and prep
@@
 ```{r}
-janvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_winter")
+janvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
 marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
@@
 ```{r month of interest -change}
-mthna <- "apr"
-mthno <- "4"
+mthna <- "jan"
+mthno <- "1"
 ```
@@
 normal model..
-mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE")) 
+mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "replicates=5", "replicatetype=crossvalidate"))
 mod12
 
-```{r train and test on jan inc AIC for reg val}
+```{r train and test inc AIC for reg val}
 modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
@@
 betavallist <- cbind(aicetc$regularization)
+nfolds <- 5
 
@@
   yr <- str_sub(timeslice, -4) #the year being used for testing
-  train <- subset(prback, year != yr) #training data is all years but testing
+  train <- subset(prback, year == yr) #training data is all years but testing
   trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-  test <-  subset(prback, year == yr) #creates the testing dataset
-  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
-  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
-  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
   train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
@@
     betamval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
+    mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nfolds), "replicatetype=crossvalidate", paste0("betamultiplier=", betavallist[h])), path = paste0("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxent/p2/models/", timeslice)) #here the full path is needed
 
@@
     #count number of non-zero coefficients, excluding the final four rows
-    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    lambdas_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*lambdas", full.names = TRUE)
+    samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)
+    for(fold in 1:nfolds){
+      fn <- fold
+      modn <- fn-1
+      lamf <- lambdas_list[fold]
+      sampf <- samppred_list[fold]
+      coefile <- read.csv(paste0(lamf), header = FALSE, stringsAsFactors=FALSE)
       coefile <- head(coefile, -4)
-    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+      coefile2 <- subset(coefile, V2 != 0)
       nocoeff <- nrow(coefile2)
@@
       bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
-
-    #no presences used in model (less those used for testing)
-    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    notrainingpoints <- maxresults$X.Training.samples
-    testauc <- maxresults$Test.AUC
-    trainauc <- maxresults$Training.AUC
-
       #natural log
-    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+      samplepred <- read.csv(paste0(sampf), header = TRUE)
       samplepred$nat_log <- log(samplepred$Raw.prediction)
@@
 
+      maxresults <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+
+      #no presences used in model (less those used for testing)
+
+      notrainingpoints <- maxresults$X.Training.samples[fold]
+      testauc <- maxresults$Test.AUC[fold]
+      trainauc <- maxresults$Training.AUC[fold]
+      modno <- maxresults$Species[fold]
+
    #test auc scores and other lovely things you might want to include
-    ev <- evaluate(p = testp, a= testb, model = mod)
-    evauc <- ev@auc
-    evcor <- ev@cor
-    evpcor <- ev@pcor
-    evnp <- ev@np
-    evna <- ev@na
-    evmaxtprtnr <- threshold(ev, 'spec_sens')
-    evmaxkap <- threshold(ev, 'kappa')
-    evno_omission <- threshold(ev, 'no_omission')
-    evprevalence <- threshold(ev, 'prevalence')
-    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+   # ev <- evaluate(p = testp, a= testb, model = mod)
+    #evauc <- ev@auc
+    #evcor <- ev@cor
+    #evpcor <- ev@pcor
+    #evnp <- ev@np
+    #evna <- ev@na
+    #evmaxtprtnr <- threshold(ev, 'spec_sens')
+    #evmaxkap <- threshold(ev, 'kappa')
+    #evno_omission <- threshold(ev, 'no_omission')
+    #evprevalence <- threshold(ev, 'prevalence')
+    #evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+    #evsensitivity <- threshold(ev, 'sensitivity')
 
-    for (i in 1:nrow(aicetc)){
-      if(aicetc$regularization[i] == betamval){
-      aicetc$Train.AUC[i] <- trainauc
-      aicetc$n[i] <- notrainingpoints
-      aicetc$coefficients[i] <- nocoeff
-      aicetc$sum.of.logs[i] <- sumsamplepred
-      aicetc$Test.AUC[i] <- evauc
-      aicetc$evcor[i] <- evcor
-      aicetc$evpcor[i] <- evpcor
-      aicetc$nopresence[i] <- evnp
-      aicetc$nobackground[i] <- evna
-      aicetc$maxsumsenspec[i] <- evmaxtprtnr
-      aicetc$maxsumsenspec[i] <- evmaxkap
-      aicetc$no_omission[i] <- evno_omission
-      aicetc$equal_sens_spec[i] <- evequal_sens_spec
-      aicetc$bic[i] <- bic #not sure this is correct...
-      }
-    }
-  }
+      aicetc$train_AUC <- trainauc
+      aicetc$test_AIC <- testauc
+      aicetc$no_trainng_points <- notrainingpoints
+      aicetc$test_coefficients <- nocoeff
+      aicetc$test_sum.of.logs <- sumsamplepred
+      #aicetc$test_AUC <- evauc
+      #aicetc$test_evcor <- evcor
+      # aicetc$test_evpcor <- evpcor
+      # aicetc$test_nopresence <- evnp
+      # aicetc$test_nobackground <- evna
+      # aicetc$test_maxsumsenspec <- evmaxtprtnr
+      # aicetc$test_maxkap <- evmaxkap
+      # aicetc$test_no_omission <- evno_omission
+      # aicetc$test_prevalence <- evprevalence
+      # aicetc$test_equal_sens_spec <- evequal_sens_spec
+      # aicetc$test_sensitivity <- evsensitivity
+      aicetc$test_bic <- bic #not sure this is correct...
+      aicetc$beta <- betaval
 
@@
 
-write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+    #calculate minimal predicted area using ecospat
+    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
+    }
+    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
+aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/", modn, "AICforMaxentReg.csv"), row.names = FALSE)
+
+    }
 
@@
 
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/p2/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
 
@@
     #count number of non-zero coefficients, excluding the final four rows
-    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    coefile <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
     coefile <- head(coefile, -4)
@@
     #no presences used in model (less those used for testing)
-    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    maxresults <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     notrainingpoints <- maxresults$X.Training.samples
@@
     #natural log
-    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    samplepred <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
     samplepred$nat_log <- log(samplepred$Raw.prediction)
@@
     evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+    evsensitivity <- threshold(ev, 'sensitivity')
 
-      aicetc$Train.AUC <- trainauc
-      aicetc$n <- notrainingpoints
-      aicetc$coefficients <- nocoeff
-      aicetc$sum.of.logs <- sumsamplepred
-      aicetc$Test.AUC <- evauc
-      aicetc$evcor <- evcor
-      aicetc$evpcor <- evpcor
-      aicetc$nopresence <- evnp
-      aicetc$nobackground <- evna
-      aicetc$maxsumsenspec <- evmaxtprtnr
-      aicetc$maxsumsenspec <- evmaxkap
-      aicetc$no_omission <- evno_omission
-      aicetc$equal_sens_spec <- evequal_sens_spec
-      aicetc$bic <- bic #not sure this is correct...
+      aicetc$train_AUC <- trainauc
+      aicetc$no_trainng_points <- notrainingpoints
+      aicetc$test_coefficients <- nocoeff
+      aicetc$test_sum.of.logs <- sumsamplepred
+      aicetc$test_AUC <- evauc
+      aicetc$test_evcor <- evcor
+      aicetc$test_evpcor <- evpcor
+      aicetc$test_nopresence <- evnp
+      aicetc$test_nobackground <- evna
+      aicetc$test_maxsumsenspec <- evmaxtprtnr
+      aicetc$test_maxkap <- evmaxkap
+      aicetc$test_no_omission <- evno_omission
+      aicetc$test_prevalence <- evprevalence
+      aicetc$test_equal_sens_spec <- evequal_sens_spec
+      aicetc$test_sensitivity <- evsensitivity
+      aicetc$test_bic <- bic #not sure this is correct...
       aicetc$beta <- betaval
@@
 
-    write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+    #calculate minimal predicted area using ecospat
+    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
+
+    write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
 
@@
     stackdeplst[[year]] <- stackdep
-    #write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
+    #write.csv(stackdep, paste0("../output/maxent/p2/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
     aea <- raster("../output/env/aea.tif")
@@
       #stackSave(test,unlayer_yr$stackname[td])
-      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
-      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", stackn, ".png"))) 
+      pred <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
+      predcum <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
+      prplot <- plot(pred, png(paste0("../output/maxent/p2/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
+      dev.off() # stops automatic saving of the plot to a png
+      prplotcum <- plot(predcum, png(paste0("../output/maxent/p2/predictions/", mthna, "/", "cumu_", stackn, ".png"))) 
       dev.off() # stops automatic saving of the plot to a png
@@
 stdep <- do.call(rbind, stackdeplst)
-write.csv(stdep, paste0("../output/maxent/predictions/", mthna, "/stack_depth.csv"), row.names = FALSE)
+write.csv(stdep, paste0("../output/maxent/p2/predictions/", mthna, "/stack_depth.csv"), row.names = FALSE)
 ```
@@
   if(yr == 1998){
-    aic1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic1998 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr1998 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 1999){
-    aic1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic1999 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr1999 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2000){
-    aic2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2000 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2000 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2001){
-    aic2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2001 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2001 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2002){
-    aic2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2002 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2002 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2003){
-    aic2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2003 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2003 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2004){
-    aic2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2004 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2004 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2005){
-    aic2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2005 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2005 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2006){
-    aic2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2006 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2006 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2007){
-    aic2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2007 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2007 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2008){
-    aic2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2008 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2008 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2009){
-    aic2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2009 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2009 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2010){
-    aic2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2010 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2010 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2011){
-    aic2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2011 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2011 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2012){
-    aic2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2012 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2012 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2013){
-    aic2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2013 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2013 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2014){
-    aic2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2014 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2014 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
     } else if (yr == 2015){
-    aic2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    aic2015 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+    maxr2015 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
   }
@@
 
-write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
-write.csv(maxrm, paste0("../output/maxent/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
+write.csv(avgAIC, paste0("../output/maxent/p2/models/", mthna, "avgAIC.csv"), row.names = FALSE)
+write.csv(maxrm, paste0("../output/maxent/p2/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
 ```
@@
 
-```{r avg asc layers}
-asc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "*.asc", full.names = TRUE)
+```{r avg asc layers cloglog}
+asc_list <- list.files(paste0("../output/maxent/p2/predictions/", mthna), pattern = "cloglog*.asc", full.names = TRUE)
 depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
@@
 }
-write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/asclayerbydepth.csv")), row.names = TRUE)
+write.csv(depthlayers, (paste0("../output/maxent/p2/predictions/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)
 
@@
 
-  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cloglog')
-  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".png"))) 
+  writeRaster(avg, (paste0("../output/maxent/p2/predictions/", mthna, "/avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cloglog')
+  avgplot <- plot(avg, png(paste0("../output/maxent/p2/predictions/", mthna, "/avg_", asclaydep, ".png"))) 
   dev.off()
@@
 
-ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), glob2rx(pattern = "avg_*.asc"), full.names = TRUE)
+ascdepth_list <- list.files(paste0("../output/maxent/p2/predictions/", mthna), glob2rx(pattern = "avg_*.asc"), full.names = TRUE)
 
-asc120  <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_120.asc"))
-asc41.18 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_41.18.asc"))
-asc1.55586 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1.55586.asc"))
-asc1045.85 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1045.85.asc"))
-asc108.03 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_108.03.asc"))
-asc11.7737 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_11.7737.asc"))
-asc1151.99 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1151.99.asc"))
-asc1265.86 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1265.86.asc"))
-asc13.991 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_13.991.asc"))
-asc133.076 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_133.076.asc"))
-asc1387.38 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1387.38.asc"))
-asc147.406 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_147.406.asc"))
-asc16.5253 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_16.5253.asc"))
-asc163.165 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_163.165.asc"))
-asc180.55 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_180.55.asc"))
-asc19.4298 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_19.4298.asc"))
-asc199.79 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_199.79.asc"))
-asc2.66768 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_2.66768.asc"))
-asc22.7576 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_22.7576.asc"))
-asc221.141 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_221.141.asc"))
-asc244.891 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_244.891.asc"))
-asc26.5583 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_26.5583.asc"))
-asc271.356 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_271.356.asc"))
-asc3.85628 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_3.85628.asc"))
-asc30.8746 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_30.8746.asc"))
-asc300.888 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_300.888.asc"))
-asc333.863 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_333.863.asc"))
-asc35.7402 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_35.7402.asc"))
-asc370.689 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_370.689.asc"))
-asc411.794 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_411.794.asc"))
-asc457.626 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_457.626.asc"))
-asc47.2119 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_47.2119.asc"))
-asc5.14036 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_5.14036.asc"))
-asc508.64 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_508.64.asc"))
-asc53.8506 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_53.8506.asc"))
-asc565.292 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_565.292.asc"))
-asc6.54303 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_6.54303.asc"))
-asc61.1128 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_61.1128.asc"))
-asc628.026 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_628.026.asc"))
-asc69.0217 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_69.0217.asc"))
-asc697.259 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_697.259.asc"))
-asc77.6112 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_77.6112.asc"))
-asc773.368 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_773.368.asc"))
-asc8.09252 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_8.09252.asc"))
-asc856.679 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_856.679.asc"))
-asc86.9294 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_86.9294.asc"))
-asc9.82275 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_9.82275.asc"))
-asc947.448 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_947.448.asc"))
-asc97.0413 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_97.0413.asc"))
+clog_asc120  <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_120.asc"))
+clog_asc41.18 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_41.18.asc"))
+clog_asc1.55586 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_1.55586.asc"))
+clog_asc1045.85 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_1045.85.asc"))
+clog_asc108.03 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_108.03.asc"))
+clog_asc11.7737 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_11.7737.asc"))
+clog_asc1151.99 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_1151.99.asc"))
+clog_asc1265.86 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_1265.86.asc"))
+clog_asc13.991 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_13.991.asc"))
+clog_asc133.076 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_133.076.asc"))
+clog_asc1387.38 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_1387.38.asc"))
+clog_asc147.406 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_147.406.asc"))
+clog_asc16.5253 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_16.5253.asc"))
+clog_asc163.165 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_163.165.asc"))
+clog_asc180.55 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_180.55.asc"))
+clog_asc19.4298 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_19.4298.asc"))
+clog_asc199.79 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_199.79.asc"))
+clog_asc2.66768 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_2.66768.asc"))
+clog_asc22.7576 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_22.7576.asc"))
+clog_asc221.141 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_221.141.asc"))
+clog_asc244.891 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_244.891.asc"))
+clog_asc26.5583 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_26.5583.asc"))
+clog_asc271.356 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_271.356.asc"))
+clog_asc3.85628 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_3.85628.asc"))
+clog_asc30.8746 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_30.8746.asc"))
+clog_asc300.888 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_300.888.asc"))
+clog_asc333.863 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_333.863.asc"))
+clog_asc35.7402 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_35.7402.asc"))
+clog_asc370.689 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_370.689.asc"))
+clog_asc411.794 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_411.794.asc"))
+clog_asc457.626 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_457.626.asc"))
+clog_asc47.2119 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_47.2119.asc"))
+clog_asc5.14036 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_5.14036.asc"))
+clog_asc508.64 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_508.64.asc"))
+clog_asc53.8506 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_53.8506.asc"))
+clog_asc565.292 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_565.292.asc"))
+clog_asc6.54303 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_6.54303.asc"))
+clog_asc61.1128 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_61.1128.asc"))
+clog_asc628.026 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_628.026.asc"))
+clog_asc69.0217 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_69.0217.asc"))
+clog_asc697.259 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_697.259.asc"))
+clog_asc77.6112 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_77.6112.asc"))
+clog_asc773.368 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_773.368.asc"))
+clog_asc8.09252 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_8.09252.asc"))
+clog_asc856.679 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_856.679.asc"))
+clog_asc86.9294 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_86.9294.asc"))
+clog_asc9.82275 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_9.82275.asc"))
+clog_asc947.448 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_947.448.asc"))
+clog_asc97.0413 <- raster(paste0("../output/maxent/p2/predictions/", mthna, "/avg_97.0413.asc"))
 
-stackavg <- stack(asc120, asc41.18, asc1.55586, asc1045.85, asc108.03, asc11.7737, asc1151.99, asc1265.86, asc13.991, asc133.076, asc1387.38, asc147.406, asc16.5253, asc163.165, asc180.55, asc19.4298, asc199.79, asc2.66768, asc22.7576, asc221.141, asc244.891, asc26.5583, asc271.356, asc3.85628, asc30.8746, asc300.888, asc333.863, asc35.7402, asc370.689, asc411.794, asc457.626, asc47.2119, asc5.14036, asc508.64, asc53.8506, asc565.292, asc6.54303, asc61.1128, asc628.026, asc69.0217, asc697.259, asc77.6112, asc773.368, asc8.09252, asc856.679, asc86.9294, asc9.82275, asc947.448, asc97.0413)
+stackavg <- stack(clog_asc120, clog_asc41.18, clog_asc1.55586, clog_asc1045.85, clog_asc108.03, clog_asc11.7737, clog_asc1151.99, clog_asc1265.86, clog_asc13.991, clog_asc133.076, clog_asc1387.38, clog_asc147.406, clog_asc16.5253, clog_asc163.165, clog_asc180.55, clog_asc19.4298, clog_asc199.79, clog_asc2.66768, clog_asc22.7576, clog_asc221.141, clog_asc244.891, clog_asc26.5583, clog_asc271.356, clog_asc3.85628, clog_asc30.8746, clog_asc300.888, clog_asc333.863, clog_asc35.7402, clog_asc370.689, clog_asc411.794, clog_asc457.626, clog_asc47.2119, clog_asc5.14036, clog_asc508.64, clog_asc53.8506, clog_asc565.292, clog_asc6.54303, clog_asc61.1128, clog_asc628.026, clog_asc69.0217, clog_asc697.259, clog_asc77.6112, clog_asc773.368, clog_asc8.09252, clog_asc856.679, clog_asc86.9294, clog_asc9.82275, clog_asc947.448, clog_asc97.0413)
 
-#stsum <- calc(stkasc, fun=sum)
+stsum <- sum(stackavg, na.rm = TRUE)
+  writeRaster(stsum, (paste0("../output/maxent/p2/predictions/", mthna, "/clog_sum_", mthna, ".asc")), overwrite=TRUE)
+  sumplot <- plot(stsum, png(paste0("../output/maxent/p2/predictions/", mthna, "/clog_sum_", mthna, ".png"))) 
+  dev.off()
 stavg <- calc(stackavg, fun = mean, na.rm = TRUE)
-  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".asc")), overwrite=TRUE)
-  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".png"))) 
+  writeRaster(stavg, (paste0("../output/maxent/p2/predictions/", mthna, "/clog_avg_", mthna, ".asc")), overwrite=TRUE)
+  avgplot <- plot(stavg, png(paste0("../output/maxent/p2/predictions/", mthna, "/clog_avg_", mthna, ".png"))) 
   dev.off()

### 2019-04-17 src/paper2_maxent_avec_xvalidation.Rmd
finally finished the code to go through the folds, get aic and put into a simple spreadsheet for analysis

@@
 modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
-aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
-betavallist <- cbind(aicetc$regularization)
-nfolds <- 5
+#modelmonth <- "jan2003"
+aicetc <- read.csv("../data/env/AICcCalculations_p2.csv", header = TRUE) #pre-created csv file
+betavallist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
+betavallist <- cbind(betavallist$regularization)
+nfolds <- cbind(c(1:5))
 
 for (year in 1:nrow(modelmonth)){
+
   timeslice <- modelmonth[year] #testing timeslice
+  #timeslice <- modelmonth
   yr <- str_sub(timeslice, -4) #the year being used for testing
@@
     betamval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nfolds), "replicatetype=crossvalidate", paste0("betamultiplier=", betavallist[h])), path = paste0("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxent/p2/models/", timeslice)) #here the full path is needed
+    mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)), path = paste0("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxent/p2/models/", timeslice)) #here the full path is needed
 
     #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
+
+    maxresults <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
 
@@
     samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)
-    for(fold in 1:nfolds){
-      fn <- fold
-      modn <- fn-1
-      lamf <- lambdas_list[fold]
-      sampf <- samppred_list[fold]
+
+      for(f in 1:nrow(nfolds)){
+
+      modn <- nfolds[f]-1
+      lamf <- lambdas_list[f]
+      sampf <- samppred_list[f]
       coefile <- read.csv(paste0(lamf), header = FALSE, stringsAsFactors=FALSE)
@@
 
-      maxresults <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-
       #no presences used in model (less those used for testing)
 
-      notrainingpoints <- maxresults$X.Training.samples[fold]
-      testauc <- maxresults$Test.AUC[fold]
-      trainauc <- maxresults$Training.AUC[fold]
-      modno <- maxresults$Species[fold]
-
-   #test auc scores and other lovely things you might want to include
-   # ev <- evaluate(p = testp, a= testb, model = mod)
-    #evauc <- ev@auc
-    #evcor <- ev@cor
-    #evpcor <- ev@pcor
-    #evnp <- ev@np
-    #evna <- ev@na
-    #evmaxtprtnr <- threshold(ev, 'spec_sens')
-    #evmaxkap <- threshold(ev, 'kappa')
-    #evno_omission <- threshold(ev, 'no_omission')
-    #evprevalence <- threshold(ev, 'prevalence')
-    #evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
-    #evsensitivity <- threshold(ev, 'sensitivity')
+      notrainingpoints <- maxresults$X.Training.samples[f]
+      testauc <- maxresults$Test.AUC[f]
+      trainauc <- maxresults$Training.AUC[f]
+      modno <- maxresults$Species[f]
 
       aicetc$train_AUC <- trainauc
-      aicetc$test_AIC <- testauc
-      aicetc$no_trainng_points <- notrainingpoints
-      aicetc$test_coefficients <- nocoeff
-      aicetc$test_sum.of.logs <- sumsamplepred
-      #aicetc$test_AUC <- evauc
-      #aicetc$test_evcor <- evcor
-      # aicetc$test_evpcor <- evpcor
-      # aicetc$test_nopresence <- evnp
-      # aicetc$test_nobackground <- evna
-      # aicetc$test_maxsumsenspec <- evmaxtprtnr
-      # aicetc$test_maxkap <- evmaxkap
-      # aicetc$test_no_omission <- evno_omission
-      # aicetc$test_prevalence <- evprevalence
-      # aicetc$test_equal_sens_spec <- evequal_sens_spec
-      # aicetc$test_sensitivity <- evsensitivity
-      aicetc$test_bic <- bic #not sure this is correct...
-      aicetc$beta <- betaval
+      aicetc$test_AUC <- testauc
+      aicetc$n <- notrainingpoints
+      aicetc$coefficients <- nocoeff
+      aicetc$sum.of.logs <- sumsamplepred
+
+      aicetc$bic <- bic #not sure this is correct...
+      aicetc$beta <- betamval
+      aicetc$foldno <- modn
 
@@
     aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
-    }
+
     aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
     aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
-write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/", modn, "AICforMaxentReg.csv"), row.names = FALSE)
+    write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/", betamval, "_", modn, "_", "AICforMaxentReg.csv"), row.names = FALSE)
+      }
+
+  # aa <- paste0("fold", modn, "_", betamval, "_", "AICforMaxentReg.csv")
+    aicperbetalst <- list.files(paste0("../output/maxent/p2/models/", timeslice, "/"), pattern = paste0("^", betamval, ".*AICforMaxentReg.csv"), full.names = TRUE)
+    aicperbeta <- lapply(aicperbetalst, read.csv, header = TRUE)
+    aicperbeta <- do.call(rbind, aicperbeta)
+    aicbetamean <- colMeans(aicperbeta)
+    aicperbeta <- rbind(aicperbeta, aicbetamean)
+    aicperbeta$foldno[f+1] <- f+1
+    write.csv(aicperbeta, paste0("../output/maxent/p2/models/", timeslice, "/aicperbetasum_", betamval, ".csv"), row.names = FALSE)
+
+    betaavglst <- list.files(paste0("../output/maxent/p2/models/", timeslice, "/"), pattern = "aicperbetasum_.*csv", full.names = TRUE)
+    bavg <- lapply(betaavglst, read.csv, header = TRUE)
+    bavg <- do.call(rbind, bavg)
+    bavg <- subset(bavg, foldno == f+1)
+    write.csv(bavg, paste0("../output/maxent/p2/models/", timeslice, "/avgbetalst.csv"), row.names = FALSE)
+  }
 
@@
 
-}
+#aicperfoldlst <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "fold4.*AICforMaxentReg.csv", full.names = TRUE)
+#aicperfold <- lapply(aicperfoldlst, read.csv, header = TRUE)
+#aicperfold <- do.call(rbind, aicperfold)
 

### 2019-04-17 tobedeleted.R

@@ -1,0 +1,102 @@
+modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
+aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
+betavallist <- cbind(aicetc$regularization)
+nfolds <- cbind(c(1:5))
+
+for (year in 1:nrow(modelmonth)){
+  timeslice <- modelmonth[year] #testing timeslice
+  yr <- str_sub(timeslice, -4) #the year being used for testing
+  train <- subset(prback, year == yr) #training data is all years but testing
+  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
+  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  
+  for (h in 1:nrow(betavallist)){
+    betamval <-  betavallist[h]
+    mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nfolds), "replicatetype=crossvalidate", paste0("betamultiplier=", betavallist[h])), path = paste0("../output/maxent/models/", timeslice)) #here the full path is needed
+    
+    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
+    
+    maxresults <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    
+    
+    #count number of non-zero coefficients, excluding the final four rows
+    lambdas_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*lambdas", full.names = TRUE)
+    samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)
+    for(f in 1:length(nfolds)){
+      fn <- f
+      modn <- fn-1
+      lamf <- lambdas_list[f]
+      sampf <- samppred_list[f]
[…]

### 2019-04-18 src/paper2_maxent_avec_xvalidation.Rmd
paper 2 runs work.

@@
 
-y1998 = 2.75
-y1999 = 2.25
-y2000 = 2
-y2001 = 1.75
-y2002 = 2
-y2003 = 2
-y2004 = 2
-y2005 = 2.25
-y2006 = 1.75
-y2007 = 1.75
-y2008 = 2.25
-y2009 = 2.25
-y2010 = 2.5
-y2011 =  2.5
+2003: 0.25
+2004: 1.5
+2005: 0.75
+2006: 2.25
 
@@
 modelmonth <- cbind(paste0(mthna, unique(prback$year)))
-aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
+aicetc <- read.csv("../data/env/AICcCalculations_p2.csv", header = TRUE) #pre-created csv file
 unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
+nfolds <- cbind(c(1:5))
 
@@
   yr <- str_sub(timeslice, -4) #the year being used for testing
-  train <- subset(prback, year != yr) #training data is all years but testing
+  train <- subset(prback, year == yr) #training data is all years but testing
   trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-  test <-  subset(prback, year == yr) #creates the testing dataset
-  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
-  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
-  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
   train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
 
-  if (timeslice == "y2001" | timeslice == "y2006" | timeslice == "y2007"){
-    betaval <- 1.75
-  } else if (timeslice == "y2000"| timeslice == "y2002" | timeslice == "y2003" | timeslice == "y2004"){
-    betaval <- 2
-  } else if (timeslice == "y1999" | timeslice == "y2005" | timeslice == "y2008" | timeslice == "y2009"){
+  if (yr == "2003"){
+    betaval <- 0.25
+  } else if (yr == "2004"){
+    betaval <- 1.5
+  } else if (yr == "2005"){
+    betaval <- 0.75
+  } else if (yr == "2006"){
     betaval <- 2.25
-  } else if (timeslice == "y2010" | timeslice == "y2011"){
-    betaval <- 2.5
-  } else if (timeslice == "y1998"){
-    betaval <- 2.75
   }
 
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/p2/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
+  mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betaval)), path = paste0("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxent/p2/models/", timeslice)) #here the full path is needed
 
@@
 
+    maxresults <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+
     #count number of non-zero coefficients, excluding the final four rows
-    coefile <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    lambdas_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*lambdas", full.names = TRUE)
+    samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)
+
+      for(f in 1:nrow(nfolds)){
+
+      modn <- nfolds[f]-1
+      lamf <- lambdas_list[f]
+      sampf <- samppred_list[f]
+      coefile <- read.csv(paste0(lamf), header = FALSE, stringsAsFactors=FALSE)
       coefile <- head(coefile, -4)
-    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+      coefile2 <- subset(coefile, V2 != 0)
       nocoeff <- nrow(coefile2)
-    bick <- sum(as.numeric(coefile2[, 2]))
-    bic <- (bick*log(sum(trainpa == 1)) - 2*11)
-
-    #no presences used in model (less those used for testing)
-    maxresults <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    notrainingpoints <- maxresults$X.Training.samples
-    testauc <- maxresults$Test.AUC
-    trainauc <- maxresults$Training.AUC
-
+      bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
+      bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
       #natural log
-    samplepred <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+      samplepred <- read.csv(paste0(sampf), header = TRUE)
       samplepred$nat_log <- log(samplepred$Raw.prediction)
@@
 
-    #test auc scores and other lovely things you might want to include
-    ev <- evaluate(p = testp, a= testb, model = mod)
-    evauc <- ev@auc
-    evcor <- ev@cor
-    evpcor <- ev@pcor
-    evnp <- ev@np
-    evna <- ev@na
-    evmaxtprtnr <- threshold(ev, 'spec_sens')
-    evmaxkap <- threshold(ev, 'kappa')
-    evno_omission <- threshold(ev, 'no_omission')
-    evprevalence <- threshold(ev, 'prevalence')
-    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
-    evsensitivity <- threshold(ev, 'sensitivity')
+      #no presences used in model (less those used for testing)
+
+      notrainingpoints <- maxresults$X.Training.samples[f]
+      testauc <- maxresults$Test.AUC[f]
+      trainauc <- maxresults$Training.AUC[f]
+      modno <- maxresults$Species[f]
 
       aicetc$train_AUC <- trainauc
-      aicetc$no_trainng_points <- notrainingpoints
-      aicetc$test_coefficients <- nocoeff
-      aicetc$test_sum.of.logs <- sumsamplepred
-      aicetc$test_AUC <- evauc
-      aicetc$test_evcor <- evcor
-      aicetc$test_evpcor <- evpcor
-      aicetc$test_nopresence <- evnp
-      aicetc$test_nobackground <- evna
-      aicetc$test_maxsumsenspec <- evmaxtprtnr
-      aicetc$test_maxkap <- evmaxkap
-      aicetc$test_no_omission <- evno_omission
-      aicetc$test_prevalence <- evprevalence
-      aicetc$test_equal_sens_spec <- evequal_sens_spec
-      aicetc$test_sensitivity <- evsensitivity
-      aicetc$test_bic <- bic #not sure this is correct...
+      aicetc$test_AUC <- testauc
+      aicetc$n <- notrainingpoints
+      aicetc$coefficients <- nocoeff
+      aicetc$sum.of.logs <- sumsamplepred
+
+      aicetc$bic <- bic #not sure this is correct...
       aicetc$beta <- betaval
+      aicetc$foldno <- modn
 
@@
 
-    write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
+    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+    write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/", betaval, "_", modn, "_", "AICforMaxentReg.csv"), row.names = FALSE)
+      }
+
+    aicperbetalst <- list.files(paste0("../output/maxent/p2/models/", timeslice, "/"), pattern = paste0("^", betaval, ".*AICforMaxentReg.csv"), full.names = TRUE)
+    aicperbeta <- lapply(aicperbetalst, read.csv, header = TRUE)
+    aicperbeta <- do.call(rbind, aicperbeta)
+    aicbetamean <- colMeans(aicperbeta)
+    aicperbeta <- rbind(aicperbeta, aicbetamean)
+    aicperbeta$foldno[f+1] <- f+1
+    write.csv(aicperbeta, paste0("../output/maxent/p2/models/", timeslice, "/final_model_aicperbetasum_", betaval, ".csv"), row.names = FALSE)
+
+#aicperfoldlst <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "fold4.*AICforMaxentReg.csv", full.names = TRUE)
+#aicperfold <- lapply(aicperfoldlst, read.csv, header = TRUE)
+#aicperfold <- do.call(rbind, aicperfold)
 
@@
       #stackSave(test,unlayer_yr$stackname[td])
-      pred <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
-      predcum <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
-      prplot <- plot(pred, png(paste0("../output/maxent/p2/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
+      pred <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", timeslice, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
+      predcum <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", timeslice, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
+      for (pf in 1:nrow(nfolds)){
+        plotn <- nfolds[pf]-1
+        lyr <- nfolds[pf]
+        prplot <- plot(pred[[pf]], png(paste0("../output/maxent/p2/predictions/", timeslice, "/", "cloglog_", stackn, "_", plotn, ".png")))
         dev.off() # stops automatic saving of the plot to a png
-      prplotcum <- plot(predcum, png(paste0("../output/maxent/p2/predictions/", mthna, "/", "cumu_", stackn, ".png"))) 
-      dev.off() # stops automatic saving of the plot to a png
+      }
     }
@@
 
-- the AICforMaxentReg.csv 
-- the maxentResults.csv
-- the asc files per depth layer
-
-open all maxentReg files
-
-```{r avg maxentReg}
-
-for (i in 1:length(modelmonth)){
-  timeslice <- modelmonth[i]
-  yr <- str_sub(timeslice, -4) #the year being used for testing
-  if(yr == 1998){
-    aic1998 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr1998 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 1999){
-    aic1999 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr1999 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2000){
-    aic2000 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2000 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2001){
-    aic2001 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2001 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2002){
-    aic2002 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2002 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2003){
-    aic2003 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2003 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2004){
-    aic2004 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2004 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2005){
-    aic2005 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2005 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2006){
-    aic2006 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2006 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2007){
-    aic2007 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2007 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2008){
-    aic2008 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2008 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2009){
-    aic2009 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2009 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2010){
-    aic2010 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2010 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2011){
-    aic2011 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2011 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2012){
-    aic2012 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2012 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2013){
-    aic2013 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2013 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2014){
-    aic2014 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2014 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2015){
-    aic2015 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2015 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-  }
-}
-
-aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
-maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)
-
-aicm <- colMeans(aic)
-aicm$mod <- "avg"
-avgAIC <- rbind(aic, aicm)
-
-maxr$Species <- 1
-maxrm <- colMeans(maxr)
-maxrm$Species <- "avg"
-maxrm <- rbind(maxr, maxrm)
-
-write.csv(avgAIC, paste0("../output/maxent/p2/models/", mthna, "avgAIC.csv"), row.names = FALSE)
-write.csv(maxrm, paste0("../output/maxent/p2/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
-```
+- the asc files per depth layer per year
 
@@
 then do something probably with raster package
+
+NEEDS TO BE DONE FROM HERE!!
 

### 2019-04-18 src/maxent_avec_xvalidation.Rmd
start work on binary

@@
 
-select mont
+select month
 ```{r month of interest -change}
-mthna <- "apr"
-mthno <- "4"
+mthna <- "jan"
+mthno <- "1"
 ```
@@
 
-  if (timeslice == "y2001" | timeslice == "y2006" | timeslice == "y2007"){
-    betaval <- 1.75
-  } else if (timeslice == "y2000"| timeslice == "y2002" | timeslice == "y2003" | timeslice == "y2004"){
-    betaval <- 2
-  } else if (timeslice == "y1999" | timeslice == "y2005" | timeslice == "y2008" | timeslice == "y2009"){
+  if (yr == "2003" | yr == "2004"){
+    betaval <- 1
+  } else if (yr == "2005"){
+    betaval <- 1.5
+  } else if (yr == "2006"){
     betaval <- 2.25
-  } else if (timeslice == "y2010" | timeslice == "y2011"){
-    betaval <- 2.5
-  } else if (timeslice == "y1998"){
-    betaval <- 2.75
   }
@@
       nao_sample <- projectRaster(nao_sample, aea)
-      nao_prev <- raster(unlayer_yr$nao_prev[td])
-      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
-      names(nao_prev) <- "nao_prev"
-      nao_prev <- projectRaster(nao_prev, aea)
       nao_winter <- raster(unlayer_yr$nao_winter[td])
@@
       nao_winter <- projectRaster(nao_winter, aea)
-      amo_sample <- raster(unlayer_yr$amo_sample[td])
-      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
-      names(amo_sample) <- "amo_sample"
-      amo_sample <- projectRaster(amo_sample, aea)
-      amo_winter <- raster(unlayer_yr$amo_winter[td])
-      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
-      names(amo_winter) <- "amo_winter"
-      amo_winter <- projectRaster(amo_winter, aea)
       stackn <- unlayer_yr$stackname[td]
@@
       dev.off() # stops automatic saving of the plot to a png
-      prplotcum <- plot(predcum, png(paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".png"))) 
-      dev.off() # stops automatic saving of the plot to a png
+
     }
@@
 
-aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
-maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)
+aic <- rbind(aic2003, aic2004, aic2005, aic2006)
+maxr <- rbind(maxr2003, maxr2004, maxr2005, maxr2006)
 
@@
 
-asc120
-asc41.18
-asc1.55586
-asc1045.85
-asc108.03
-asc11.7737
-asc1151.99
-asc1265.86
-asc13.991
-asc133.076
-asc1387.38
-asc147.406
-asc16.5253
-asc163.165
-asc180.55
-asc19.4298
-asc199.79
-asc2.66768
-asc22.7576
-asc221.141
-asc244.891
-asc26.5583
-asc271.356
-asc3.85628
-asc30.8746
-asc300.888
-asc333.863
-asc35.7402
-asc370.689
-asc411.794
-asc457.626
-asc47.2119
-asc5.14036
-asc508.64
-asc53.8506
-asc565.292
-asc6.54303
-asc61.1128
-asc628.026
-asc69.0217
-asc697.259
-asc77.6112
-asc773.368
-asc8.09252
-asc856.679
-asc86.9294
-asc9.82275
-asc947.448
-asc97.0413
+#binary maps
+
+there seem to be two primary ways to create binary maps.
+the most common for conservation purposes seems to be to use maxsumspensspec values, however MPA also comes recommended (but is def less common).
+probably best to use cumulative for both (merrow 2013)
+
+the thresholds are in the monthavgAIC.csv file in the model folder (not model/month)
+
+```{r binary map using maxsumsenspec}
+cumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cumu_.*.asc", full.names = TRUE)
+
+```

### 2019-04-23 src/maxent_alldata_avec_xvalidation.Rmd
trouble with running for all models.. java issue :/

--- 2019-04-18 src/maxent_avec_xvalidation.Rmd
@@
 ```{r pre-install & load packages, include=FALSE}
-options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat")
+options (java.parameters = "-Xmx2048m")
+packages <- c("rJava", "dismo", "stringr", "raster", "ecospat")
 source("../src/ipak.R")
@@
 ```{r remove unnecessary columns}
-presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt))
+presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt, month, depthlayerno, depth_layer))
 colnames(presback)
@@
 ```{r}
-janvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_winter")
-marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-
-mthnavar_lst <- list(jan = janvar, mar = marvar, apr = aprvar, may = mayvar, jun = junvar, jul = julvar, aug = augvar, sep = sepvar, oct = octvar, nov = novvar, dec = decvar)
-```
-
-select month
-```{r month of interest -change}
-mthna <- "jan"
-mthno <- "1"
-```
-
-```{r subsest data by month}
-prback <- subset(presback, month == mthno)
-prback <- subset(prback, select = mthnavar_lst[[mthna]])
+allvar <- c("occurrence", "year", "temp_surface", "temp_depth", "salinity_surface", "salinity_depth", "o2_depth", "chl_surface", "mlp_surface",  "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+presback <- subset(presback, select = c(allvar))
+colnames(presback)
 ```
@@
 ```{r remove na vals}
-prback <- na.omit(prback)
+prback <- na.omit(presback)
+```
+
+```{r}
+train <- prback
+trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
+train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.25"))
 ```
@@
 ```{r train and test inc AIC for reg val}
-modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
 aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
-betavallist <- cbind(aicetc$regularization)
-
-for (year in 1:nrow(modelmonth)){
-  timeslice <- modelmonth[year] #testing timeslice
-  yr <- str_sub(timeslice, -4) #the year being used for testing
-  train <- subset(prback, year != yr) #training data is all years but testing
+#betavallist <- cbind(aicetc$regularization)
+betavallist <- 0.25
+  timeslice <- "all_data"
+  train <- prback
   trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-  test <-  subset(prback, year == yr) #creates the testing dataset
-  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
-  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
-  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
   train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
 
-  for (h in 1:nrow(betavallist)){
-    betamval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
+  #for (h in 1:nrow(betavallist)){
+   # betamval <-  betavallist[h]
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist)))
 
@@
     #count number of non-zero coefficients, excluding the final four rows
-    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    coefile <- read.csv(paste0("../output/maxent/", timeslice, "/models/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
     coefile <- head(coefile, -4)
@@
     #no presences used in model (less those used for testing)
-    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    maxresults <- read.csv(paste0("../output/maxent/", timeslice, "/models/maxentResults.csv"), header = TRUE)
     notrainingpoints <- maxresults$X.Training.samples
@@
     #natural log
-    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    samplepred <- read.csv(paste0("../output/maxent/", timeslice, "/models/species_samplePredictions.csv"), header = TRUE)
     samplepred$nat_log <- log(samplepred$Raw.prediction)
@@
     aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
-      }
-    }
+      #}
 
@@
 
-write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
-
-}
+write.csv(aicetc, paste0("../output/maxent/", timeslice, "/models/", betaval, "_AICforMaxentReg.csv"), row.names = FALSE)
 
@@
 
-for (year in 1:nrow(modelmonth)){
-  timeslice <- modelmonth[year] #testing timeslice
-  yr <- str_sub(timeslice, -4) #the year being used for testing
+#for (year in 1:nrow(modelmonth)){
+ # timeslice <- modelmonth[year] #testing timeslice
+  #yr <- str_sub(timeslice, -4) #the year being used for testing
   train <- subset(prback, year != yr) #training data is all years but testing
@@
     aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
+    aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
 
@@
     }
-}
+#}
 
@@
 
-```{r binary map using maxsumsenspec}
+```{r binary maps}
+thresholds <- read.csv(paste0("../output/maxent/models/", mthna, "avgAIC.csv"), header = TRUE)
+thresholds <- tail(thresholds, 1)
+mpa <- thresholds$mpa_0.9_cum
+maxsesp <- thresholds$test_maxsumsenspec
 cumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cumu_.*.asc", full.names = TRUE) 
+depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
+names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
+depthlayers$depth <- as.character(depthlayers$depth)
+
+colno <- 1
+
+for (d in 1:nrow(depthlayers)){
+  dep <- depthlayers$depth[d]
+  for (a in 1:length(cumasc_list)){
+    asc <- cumasc_list[a]
+    if (grepl(dep, asc)){
+      depthlayers[d , paste0("a", colno)] <- asc
+        colno <- colno+1
+    }
+  }
+}
+write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cumu_asclayerbydepth.csv")), row.names = TRUE)
+
+for (x in 1:nrow(depthlayers)){
+  asclay <- depthlayers[x, ]
+  asclay <- asclay[!is.na(asclay)]
+  asclay <- as.data.frame(rbind(asclay))
+  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
+  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
+  asclay$V3 <- as.character(asclay$V3)
+  asclay$V4 <- as.character(asclay$V4)
+  asclay$V5 <- as.character(asclay$V5)
+
+  asclaydep <- asclay$V1
+
+  asc2003 <- raster(asclay$V2)
+  asc2004 <- raster(asclay$V3)
+  asc2005 <- raster(asclay$V4)
+  asc2006 <- raster(asclay$V5)
+
+  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
+  avg <- calc(stkasc, fun = mean)
+
+  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/cumu_avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cumulative')
+
+}
+
+cumavg_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cumu_avg_.*.asc", full.names = TRUE) 
+for(lyr in 1:length(cumavg_list)){
+  avgras <- raster(cumavg_list[lyr])
+  reclassmpa <- c(mpa, Inf, 1)
+  reclassmpa <- matrix(reclassmpa, ncol = 3, byrow = TRUE)
+  binavgras <- reclassify(avgras, reclassmpa)
+}
 

### 2019-04-24 src/maxent_alldata_avec_xvalidation.Rmd
update to all data

@@
 ```{r}
-train <- prback
-trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.25"))
+# train <- prback
+# trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
+# train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+# timeslice <- "all_data"
+# betavallist <- 0.25
+# mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.25"))
 ```
@@
 aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
-#betavallist <- cbind(aicetc$regularization)
-betavallist <- 0.25
+betavallist <- cbind(aicetc$regularization)
 timeslice <- "all_data"
@@
 train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+nfolds <- cbind(c(1:5))
 
-  #for (h in 1:nrow(betavallist)){
-   # betamval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist)))
+  for (h in 1:nrow(betavallist)){
+    betamval <-  betavallist[h]
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)))
 
@@
     #count number of non-zero coefficients, excluding the final four rows
-    coefile <- read.csv(paste0("../output/maxent/", timeslice, "/models/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    lambdas_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*lambdas", full.names = TRUE)
+    samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)
+
+for(f in 1:nrow(nfolds)){
+  modn <- nfolds[f]-1
+      lamf <- lambdas_list[f]
+      sampf <- samppred_list[f]
+      coefile <- read.csv(paste0(lamf), header = FALSE, stringsAsFactors=FALSE)
       coefile <- head(coefile, -4)
-    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+      coefile2 <- subset(coefile, V2 != 0)
       nocoeff <- nrow(coefile2)
@@
       bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
-
-    #no presences used in model (less those used for testing)
-    maxresults <- read.csv(paste0("../output/maxent/", timeslice, "/models/maxentResults.csv"), header = TRUE)
-    notrainingpoints <- maxresults$X.Training.samples
-    testauc <- maxresults$Test.AUC
-    trainauc <- maxresults$Training.AUC
-
       #natural log
-    samplepred <- read.csv(paste0("../output/maxent/", timeslice, "/models/species_samplePredictions.csv"), header = TRUE)
+      samplepred <- read.csv(paste0(sampf), header = TRUE)
       samplepred$nat_log <- log(samplepred$Raw.prediction)
@@
 
-   #test auc scores and other lovely things you might want to include
-    ev <- evaluate(p = testp, a= testb, model = mod)
-    evauc <- ev@auc
-    evcor <- ev@cor
-    evpcor <- ev@pcor
-    evnp <- ev@np
-    evna <- ev@na
-    evmaxtprtnr <- threshold(ev, 'spec_sens')
-    evmaxkap <- threshold(ev, 'kappa')
-    evno_omission <- threshold(ev, 'no_omission')
-    evprevalence <- threshold(ev, 'prevalence')
-    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
-    evsensitivity <- threshold(ev, 'sensitivity')
+      #no presences used in model (less those used for testing)
+
+      notrainingpoints <- maxresults$X.Training.samples[f]
+      testauc <- maxresults$Test.AUC[f]
+      trainauc <- maxresults$Training.AUC[f]
+      modno <- maxresults$Species[f]
 
       aicetc$train_AUC <- trainauc
-      aicetc$no_trainng_points <- notrainingpoints
-      aicetc$test_coefficients <- nocoeff
-      aicetc$test_sum.of.logs <- sumsamplepred
-      aicetc$test_AUC <- evauc
-      aicetc$test_evcor <- evcor
-      aicetc$test_evpcor <- evpcor
-      aicetc$test_nopresence <- evnp
-      aicetc$test_nobackground <- evna
-      aicetc$test_maxsumsenspec <- evmaxtprtnr
-      aicetc$test_maxkap <- evmaxkap
-      aicetc$test_no_omission <- evno_omission
-      aicetc$test_prevalence <- evprevalence
-      aicetc$test_equal_sens_spec <- evequal_sens_spec
-      aicetc$test_sensitivity <- evsensitivity
-      aicetc$test_bic <- bic #not sure this is correct...
-      aicetc$beta <- betaval
+      aicetc$test_AUC <- testauc
+      aicetc$n <- notrainingpoints
+      aicetc$coefficients <- nocoeff
+      aicetc$sum.of.logs <- sumsamplepred
+
+      aicetc$bic <- bic #not sure this is correct...
+      aicetc$beta <- betamval
+      aicetc$foldno <- modn
 
@@
     aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
-      #}
 
@@
     aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+    write.csv(aicetc, paste0("../output/maxent/", timeslice, "/models/", betamval, "_", modn, "_", "AICforMaxentReg.csv"), row.names = FALSE)
+      }
 
-write.csv(aicetc, paste0("../output/maxent/", timeslice, "/models/", betaval, "_AICforMaxentReg.csv"), row.names = FALSE)
+  # aa <- paste0("fold", modn, "_", betamval, "_", "AICforMaxentReg.csv")
+    aicperbetalst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = paste0("^", betamval, ".*AICforMaxentReg.csv"), full.names = TRUE)
+    aicperbeta <- lapply(aicperbetalst, read.csv, header = TRUE)
+    aicperbeta <- do.call(rbind, aicperbeta)
+    aicbetamean <- colMeans(aicperbeta)
+    aicperbeta <- rbind(aicperbeta, aicbetamean)
+    aicperbeta$foldno[f+1] <- f+1
+    write.csv(aicperbeta, paste0("../output/maxent/", timeslice, "/models/aicperbetasum_", betamval, ".csv"), row.names = FALSE)
+
+    betaavglst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = "aicperbetasum_.*csv", full.names = TRUE)
+    bavg <- lapply(betaavglst, read.csv, header = TRUE)
+    bavg <- do.call(rbind, bavg)
+    bavg <- subset(bavg, foldno == f+1)
+    write.csv(bavg, paste0("../output/maxent/", timeslice, "/models/avgbetalst.csv"), row.names = FALSE)
+  }
 

### 2019-04-24 src/maxent_alldata_avec_xvalidation.Rmd
yup..

@@
     #count number of non-zero coefficients, excluding the final four rows
-    lambdas_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*lambdas", full.names = TRUE)
-    samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)
+    lambdas_list <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = "*lambdas", full.names = TRUE)
+    samppred_list <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = "*samplePredictions.csv", full.names = TRUE)
 

### 2019-04-30 src/maxent_avec_xvalidation.Rmd
binary complete - just now playing with a few bits

@@
 options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat")
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps")
 source("../src/ipak.R")
@@
 
-      aicetc$train_AUC <- trainauc
-      aicetc$no_trainng_points <- notrainingpoints
-      aicetc$test_coefficients <- nocoeff
-      aicetc$test_sum.of.logs <- sumsamplepred
-      aicetc$test_AUC <- evauc
-      aicetc$test_evcor <- evcor
-      aicetc$test_evpcor <- evpcor
+      aicetc$train_AUC <- trainauc #training only AUC - not tested
+      aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
+      aicetc$test_coefficients <- nocoeff #this is to calculate the AIC
+      aicetc$test_sum.of.logs <- sumsamplepred #this is to calculate the AIC
+      aicetc$test_AUC <- evauc #testing AUC
+      aicetc$test_evcor <- evcor #testing correlation coefficient
+      aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
       aicetc$test_nopresence <- evnp 
@@
     aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
+    aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
+    aicetc$year <- yr
 
@@
     stackdeplst[[year]] <- stackdep
-    #write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
+    write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
     aea <- raster("../output/env/aea.tif")
@@
       stackn <- unlayer_yr$stackname[td]
-      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
+      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_winter)
       #stackSave(test,unlayer_yr$stackname[td])
       pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
-      predcum <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
+      predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
       prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
@@
 write.csv(stdep, paste0("../output/maxent/predictions/", mthna, "/stack_depth.csv"), row.names = FALSE)
+```
+
+#binary maps - individual layers
+
+there seem to be two primary ways to create binary maps.
+the most common for conservation purposes seems to be to use maxsumspensspec values, however MPA also comes recommended (but is def less common).
+probably best to use cumulative for both (merrow 2013)
+
+the thresholds are in the monthavgAIC.csv file in the model folder (not model/month)
+
+for some reason the cumulative output comes empty... using bossMaps
+```{r raw to cumulative}
+rawasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^raw_.*.asc", full.names = TRUE) 
+for (raw in 1:length(rawasc_list)){
+  asc <- raster(rawasc_list[raw])
+  ascname <- names(asc)
+  ascname <- str_sub(ascname, start = 5)
+  cumasc <- cumulative(asc, threshold = NULL)
+  writeRaster(cumasc, paste0("../output/maxent/predictions/", mthna, "/", "cum_", ascname, ".asc"), overwrite = TRUE)
+}
+```
+
+```{r set up threshold binary function}
+binmpa <- function(x) {
+  ifelse(x <=  mpa_thr, 0,
+  ifelse(x >  mpa_thr, 1, NA))
+  }
+```
+cum_mpathr <- calc(cumasc, fun = rc)
+plot(cum_mpathr)
+tenper_thr <- 20.338725
+
+```{r binary maps on individual layers}
+cumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cum_.*.asc", full.names = TRUE)
+
+for (modtime in 1:nrow(modelmonth)){
+  timeslice <- modelmonth[modtime]
+  print(timeslice)
+  yr <- str_sub(timeslice, -4)
+  print(yr)
+  thresholds <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+  mpa_thr <- thresholds$mpa_0.9_cum
+  print(mpa_thr)
+  for (cum in 1:length(cumasc_list)){
+    cumascyr <- str_sub(cumasc_list[cum], start = 40, end = 43) #selects the year from the file name
+    if(yr == cumascyr){
+      cumasc <- raster(cumasc_list[cum])
+      ascname <- names(cumasc)
+      print(ascname)
+      cum_mpathr <- calc(cumasc, fun = binmpa)
+      mpa_plot <- plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".png")))
+      dev.off() # stops automatic saving of the plot to a png
+      writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".asc"), overwrite = TRUE)
+    }
+  }
+}
+
 ```
@@
 ```{r avg asc layers cloglog}
-asc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "cloglog*.asc", full.names = TRUE)
+asc_list <- list.files(paste0("../output/maxent/predictions/", mthna, "/"), pattern = "^cloglog.*asc", full.names = TRUE)
 depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
@@
   asclay <- as.data.frame(rbind(asclay))
-  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
-  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
+  asclay$V1 <- as.character(asclay$V1)
+  asclay$V2 <- as.character(asclay$V2)
   asclay$V3 <- as.character(asclay$V3)
@@
   asclay$V5 <- as.character(asclay$V5)
-  asclay$V6 <- as.character(asclay$V6)
-  asclay$V7 <- as.character(asclay$V7)
-  asclay$V8 <- as.character(asclay$V8)
-  asclay$V9 <- as.character(asclay$V9)
-  asclay$V10 <- as.character(asclay$V10)
-  asclay$V11 <- as.character(asclay$V11)
-  asclay$V12 <- as.character(asclay$V12)
-  asclay$V13 <- as.character(asclay$V13)
-  asclay$V14 <- as.character(asclay$V14)
-  asclay$V15 <- as.character(asclay$V15)
-
   asclaydep <- asclay$V1
 
-  asc1998 <- raster(asclay$V2)
-  asc1999 <- raster(asclay$V3)
-  asc2000 <- raster(asclay$V4)
-  asc2001 <- raster(asclay$V5)
-  asc2002 <- raster(asclay$V6)
-  asc2003 <- raster(asclay$V7)
-  asc2004 <- raster(asclay$V8)
-  asc2005 <- raster(asclay$V9)
-  asc2006 <- raster(asclay$V10)
-  asc2007 <- raster(asclay$V11)
-  asc2008 <- raster(asclay$V12)
-  asc2009 <- raster(asclay$V13)
-  asc2010 <- raster(asclay$V14)
-  asc2011 <- raster(asclay$V15)
+  asc2003 <- raster(asclay$V2)
+  asc2004 <- raster(asclay$V3)
+  asc2005 <- raster(asclay$V4)
+  asc2006 <- raster(asclay$V5)
 
-  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)
+  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
   avg <- calc(stkasc, fun = mean)
 
-  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cloglog')
-  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".png"))) 
+  writeRaster(avg, paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".asc"), overwrite=TRUE)
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".png"))) 
+  dev.off()
+}
+```
+
+```{r avg asc layers cumulative}
+asc_list <- list.files(paste0("../output/maxent/predictions/", mthna, "/"), pattern = "^cum.*asc", full.names = TRUE)
+depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
+names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
+depthlayers$depth <- as.character(depthlayers$depth)
+
+colno <- 1
+
+for (d in 1:nrow(depthlayers)){
+  dep <- depthlayers$depth[d]
+  for (a in 1:length(asc_list)){
+    asc <- asc_list[a]
+    if (grepl(dep, asc)){
+      depthlayers[d , paste0("a", colno)] <- asc
+        colno <- colno+1
+    }
+  }
+}
+write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cum_asclayerbydepth.csv")), row.names = TRUE)
+
+for (x in 1:nrow(depthlayers)){
+  asclay <- depthlayers[x, ]
+  asclay <- asclay[!is.na(asclay)]
+  asclay <- as.data.frame(rbind(asclay))
+  asclay$V1 <- as.character(asclay$V1)
+  asclay$V2 <- as.character(asclay$V2)
+  asclay$V3 <- as.character(asclay$V3)
+  asclay$V4 <- as.character(asclay$V4)
+  asclay$V5 <- as.character(asclay$V5)
+  asclaydep <- asclay$V1
+
+  asc2003 <- raster(asclay$V2)
+  asc2004 <- raster(asclay$V3)
+  asc2005 <- raster(asclay$V4)
+  asc2006 <- raster(asclay$V5)
+
+  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
+  avg <- calc(stkasc, fun = mean)
+
+  writeRaster(avg, paste0("../output/maxent/predictions/", mthna, "/avg_cum_", asclaydep, ".asc"), overwrite=TRUE)
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_cum_", asclaydep, ".png"))) 
   dev.off()
@@
 
-ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), glob2rx(pattern = "avg_*.asc"), full.names = TRUE)
+```
 
-clog_asc120  <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_120.asc"))
-clog_asc41.18 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_41.18.asc"))
-clog_asc1.55586 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1.55586.asc"))
-clog_asc1045.85 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1045.85.asc"))
-clog_asc108.03 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_108.03.asc"))
-clog_asc11.7737 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_11.7737.asc"))
-clog_asc1151.99 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1151.99.asc"))
-clog_asc1265.86 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1265.86.asc"))
-clog_asc13.991 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_13.991.asc"))
-clog_asc133.076 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_133.076.asc"))
-clog_asc1387.38 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1387.38.asc"))
-clog_asc147.406 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_147.406.asc"))
-clog_asc16.5253 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_16.5253.asc"))
-clog_asc163.165 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_163.165.asc"))
-clog_asc180.55 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_180.55.asc"))
-clog_asc19.4298 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_19.4298.asc"))
-clog_asc199.79 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_199.79.asc"))
-clog_asc2.66768 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_2.66768.asc"))
-clog_asc22.7576 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_22.7576.asc"))
-clog_asc221.141 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_221.141.asc"))
-clog_asc244.891 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_244.891.asc"))
-clog_asc26.5583 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_26.5583.asc"))
-clog_asc271.356 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_271.356.asc"))
-clog_asc3.85628 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_3.85628.asc"))
-clog_asc30.8746 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_30.8746.asc"))
-clog_asc300.888 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_300.888.asc"))
-clog_asc333.863 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_333.863.asc"))
-clog_asc35.7402 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_35.7402.asc"))
-clog_asc370.689 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_370.689.asc"))
-clog_asc411.794 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_411.794.asc"))
-clog_asc457.626 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_457.626.asc"))
-clog_asc47.2119 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_47.2119.asc"))
-clog_asc5.14036 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_5.14036.asc"))
-clog_asc508.64 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_508.64.asc"))
-clog_asc53.8506 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_53.8506.asc"))
-clog_asc565.292 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_565.292.asc"))
-clog_asc6.54303 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_6.54303.asc"))
-clog_asc61.1128 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_61.1128.asc"))
-clog_asc628.026 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_628.026.asc"))
-clog_asc69.0217 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_69.0217.asc"))
-clog_asc697.259 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_697.259.asc"))
-clog_asc77.6112 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_77.6112.asc"))
-clog_asc773.368 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_773.368.asc"))
-clog_asc8.09252 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_8.09252.asc"))
-clog_asc856.679 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_856.679.asc"))
-clog_asc86.9294 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_86.9294.asc"))
-clog_asc9.82275 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_9.82275.asc"))
-clog_asc947.448 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_947.448.asc"))
-clog_asc97.0413 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_97.0413.asc"))
+```{r binary maps on averaged rasters}
+avgcumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cum.*.asc", full.names = TRUE)
+thresholds <- read.csv(paste0("../output/maxent/models/", mthna, "avgAIC.csv"), header = TRUE)
+thresholds <- tail(thresholds, 1) #subsset the last row
+mpa_thr <- thresholds$mpa_0.9_cum
+#print(mpa_thr)
 
-stackavg <- stack(clog_asc120, clog_asc41.18, clog_asc1.55586, clog_asc1045.85, clog_asc108.03, clog_asc11.7737, clog_asc1151.99, clog_asc1265.86, clog_asc13.991, clog_asc133.076, clog_asc1387.38, clog_asc147.406, clog_asc16.5253, clog_asc163.165, clog_asc180.55, clog_asc19.4298, clog_asc199.79, clog_asc2.66768, clog_asc22.7576, clog_asc221.141, clog_asc244.891, clog_asc26.5583, clog_asc271.356, clog_asc3.85628, clog_asc30.8746, clog_asc300.888, clog_asc333.863, clog_asc35.7402, clog_asc370.689, clog_asc411.794, clog_asc457.626, clog_asc47.2119, clog_asc5.14036, clog_asc508.64, clog_asc53.8506, clog_asc565.292, clog_asc6.54303, clog_asc61.1128, clog_asc628.026, clog_asc69.0217, clog_asc697.259, clog_asc77.6112, clog_asc773.368, clog_asc8.09252, clog_asc856.679, clog_asc86.9294, clog_asc9.82275, clog_asc947.448, clog_asc97.0413)
+for (avgcum in 1:length(avgcumasc_list)){
+    cumasc <- raster(avgcumasc_list[avgcum])
+    ascname <- names(cumasc)
+    #print(ascname)
+    cum_mpathr <- calc(cumasc, fun = binmpa)
+    mpa_plot <- plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".png")))
+    dev.off() # stops automatic saving of the plot to a png
+    writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".asc"), overwrite = TRUE, datatype="LOG1S")
+}
 
-stsum <- sum(stackavg, na.rm = TRUE)
-  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/clog_sum_", mthna, ".asc")), overwrite=TRUE)
-  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/clog_sum_", mthna, ".png"))) 
+```
+
+now to stack average binary maps at depth into one 2D representation...
+
+```{r}
+
+#this below is to get an average score for all layers per year. probably not what you want
+ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_mpa_avg_cum_.*asc", full.names = TRUE)
+
+binaryavgstack <- stack(ascdepth_list)
+
+stsum <- sum(binaryavgstack, na.rm = TRUE)
+  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".asc")), overwrite=TRUE)
+  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".png"))) 
   dev.off()
-stavg <- calc(stackavg, fun = mean, na.rm = TRUE)
-  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/clog_avg_", mthna, ".asc")), overwrite=TRUE)
-  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/clog_avg_", mthna, ".png"))) 
+stavg <- calc(binaryavgstack, fun = mean, na.rm = TRUE)
+  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".asc")), overwrite=TRUE)
+  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".png"))) 
   dev.off()
@@
 ```
-
-#binary maps
-
-there seem to be two primary ways to create binary maps.
-the most common for conservation purposes seems to be to use maxsumspensspec values, however MPA also comes recommended (but is def less common).
-probably best to use cumulative for both (merrow 2013)
-
-the thresholds are in the monthavgAIC.csv file in the model folder (not model/month)
-
-```{r binary map using maxsumsenspec}
-cumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cumu_.*.asc", full.names = TRUE)
-
-```

### 2019-04-30 src/maxent_avec_xvalidation.Rmd
plots by depth...

@@
   dev.off()
+```
 
+#plots by depth
+
+first need some statistcs... for each layer how many cells have a value of 1...
+
+```{r}
+test <- cellStats(binaryavgstack[[30]], sum)
+test
 ```

### 2019-05-01 src/maxent_avec_xvalidation.Rmd
highs and lows..

@@
 options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps")
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice")
 source("../src/ipak.R")
@@
 
-now to stack average binary maps at depth into one 2D representation...
-
-```{r}
-
-#this below is to get an average score for all layers per year. probably not what you want
-ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_mpa_avg_cum_.*asc", full.names = TRUE)
-
-binaryavgstack <- stack(ascdepth_list)
-
-stsum <- sum(binaryavgstack, na.rm = TRUE)
-  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".asc")), overwrite=TRUE)
-  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".png"))) 
-  dev.off()
-stavg <- calc(binaryavgstack, fun = mean, na.rm = TRUE)
-  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".asc")), overwrite=TRUE)
-  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".png"))) 
-  dev.off()
-```
-
 #plots by depth
@@
 
-```{r}
-test <- cellStats(binaryavgstack[[30]], sum)
-test
+```{r calc no presence cells per layer (avg)} 
+ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_mpa_avg_cum_.*asc", full.names = TRUE)
+binaryavgstack <- stack(ascdepth_list)
+stackstat <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+stackstat[ ,1] <- c(1:49)
+colnames(stackstat) <- c("layer", "cell_sum")
+
+for (lyr in 1:nlayers(binaryavgstack)){
+  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
+  stackstat$cell_sum[lyr] <- lyrsum
+}
+
+write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/cellcount_bydepth_avg.csv"), row.names = FALSE)
 ```
+
+a crude plot but ok to show slr
+
+```{r graph by depth (avg)}
+bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1), xlim = c(1, 1600))
+trellis.device(device="png",filename = paste0("../output/maxent/predictions/", mthna, "/", mthna, "_depthplot.png"))
+print(bydepth)
+dev.off()
+```
+
+what about the variation from one year to the next
+
+e.g. binary_mpa_st200401041.18.asc
+
+```{r no presence cells per layer per year}
+binaryyr_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_mpa_st.*asc", full.names = TRUE)
+yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryyr_list)))
+colnames(yrlayer) <- c("layer", "year")
+
+for (yrlyr in 1:length(binaryyr_list)){
+  yearoflyr <- str_sub(binaryyr_list[yrlyr], start = 47, end = 50) #selects the year from the file name
+  yrlayer$year[yrlyr] <- yearoflyr
+  yrlayer$layer[yrlyr] <- binaryyr_list[yrlyr]
+}
+
+y_list <- split(yrlayer, yrlayer$year)
+noyrs <- unique(yrlayer$year)
+
+stackstat_yr <- data.frame(matrix(ncol = length(noyrs) + 1, nrow = nlayers(binaryavgstack)))
+stackstat_yr[ ,1] <- c(1:49)
+colnames(stackstat_yr) <- c("layer", paste0("y_", noyrs))
+
+for (binyr in (1:length(y_list))){
+  lsta <- stack(y_list[[binyr]]$layer)
+  yr <- y_list[[binyr]]$year[1]
+  for (l in 1:nlayers(lsta)){
+    lyrsum <- cellStats(lsta[[l]], sum)
+    stcol <- paste0("y_", yr) 
+    stackstat_yr[[l, stcol]] <- lyrsum
+  }
+}
+
+#2005 coming up empty...
+
+```

### 2019-05-02 src/maxent_avec_xvalidation.Rmd
ok models good... about to add writebackgroundpredictions for cbi analysis...

--- 2019-04-30 src/maxent_avec_xvalidation.Rmd
@@
 options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps")
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr")
 source("../src/ipak.R")
@@
 
-#binary maps - individual layers
+#binary maps - individual year layers
 
@@
 the most common for conservation purposes seems to be to use maxsumspensspec values, however MPA also comes recommended (but is def less common).
-probably best to use cumulative for both (merrow 2013)
 
 the thresholds are in the monthavgAIC.csv file in the model folder (not model/month)
-
-for some reason the cumulative output comes empty... using bossMaps
-```{r raw to cumulative}
-rawasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^raw_.*.asc", full.names = TRUE) 
-for (raw in 1:length(rawasc_list)){
-  asc <- raster(rawasc_list[raw])
-  ascname <- names(asc)
-  ascname <- str_sub(ascname, start = 5)
-  cumasc <- cumulative(asc, threshold = NULL)
-  writeRaster(cumasc, paste0("../output/maxent/predictions/", mthna, "/", "cum_", ascname, ".asc"), overwrite = TRUE)
-}
-```
 
@@
 ```
-cum_mpathr <- calc(cumasc, fun = rc)
-plot(cum_mpathr)
-tenper_thr <- 20.338725
 
-```{r binary maps on individual layers}
-cumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cum_.*.asc", full.names = TRUE)
+```{r binary maps on individual layers cloglog}
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
 
@@
   thresholds <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-  mpa_thr <- thresholds$mpa_0.9_cum
+  mpa_thr <- thresholds$mpa_0.9_clog
   print(mpa_thr)
-  for (cum in 1:length(cumasc_list)){
-    cumascyr <- str_sub(cumasc_list[cum], start = 40, end = 43) #selects the year from the file name
+  for (cum in 1:length(clogasc_list)){
+    cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
     if(yr == cumascyr){
-      cumasc <- raster(cumasc_list[cum])
+      cumasc <- raster(clogasc_list[cum])
       ascname <- names(cumasc)
@@
       cum_mpathr <- calc(cumasc, fun = binmpa)
-      mpa_plot <- plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".png")))
+      writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna,  "/binary_", ascname, ".asc"), overwrite = TRUE)
+      plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/binary_", ascname, ".png")), overwrite = TRUE)
       dev.off() # stops automatic saving of the plot to a png
-      writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".asc"), overwrite = TRUE)
     }
   }
+}
+```
+
+now for each year create a 2-d representation of suitable habitat
+
+need to stack the rasters by year
+run cellStats (sum)
+save the output as they year
+
+```{r binary map per year 2D representation}
+binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_.*.asc", full.names = TRUE)
+yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryclog_list)))
+colnames(yrlayer) <- c("layer", "year")
+
+for (yrlyr in 1:length(binaryclog_list)){
+  yearoflyr <- str_sub(binaryclog_list[yrlyr], start = 51, end = 54) #selects the year from the file name
+  yrlayer$year[yrlyr] <- yearoflyr
+  yrlayer$layer[yrlyr] <- binaryclog_list[yrlyr]
+}
+
+y_list <- split(yrlayer, yrlayer$year)
+noyrs <- unique(yrlayer$year)
+
+for (binyr in (1:length(y_list))){
+  binstack <- stack(y_list[[binyr]]$layer)
+  year <- y_list[[binyr]]$year[1]
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".asc"))
+  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".png")))
+  dev.off()
 }
@@
 
-#model averageing
+#using averageing
 
@@
 ```{r avg asc layers cloglog}
-asc_list <- list.files(paste0("../output/maxent/predictions/", mthna, "/"), pattern = "^cloglog.*asc", full.names = TRUE)
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+stdep <- read.csv(paste0("../output/maxent/predictions/", mthna, "/stack_depth.csv"), header = TRUE)
 depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
@@
   dep <- depthlayers$depth[d]
-  for (a in 1:length(asc_list)){
-    asc <- asc_list[a]
+  for (a in 1:length(clogasc_list)){
+    asc <- clogasc_list[a]
     if (grepl(dep, asc)){
@@
   stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
-  avg <- calc(stkasc, fun = mean)
+  avg <- mean(stkasc, na.rm = TRUE)
 
@@
 
-```{r avg asc layers cumulative}
-asc_list <- list.files(paste0("../output/maxent/predictions/", mthna, "/"), pattern = "^cum.*asc", full.names = TRUE)
-depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
-names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
-depthlayers$depth <- as.character(depthlayers$depth)
+```{r binary maps on averaged depth layer}
+avgasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
+thresholds <- read.csv(paste0("../output/maxent/models/", mthna, "avgAIC.csv"), header = TRUE)
+thresholds <- tail(thresholds, 1) #subset the last row
+mpa_thr <- thresholds$mpa_0.9_clog
 
-colno <- 1
-
-for (d in 1:nrow(depthlayers)){
-  dep <- depthlayers$depth[d]
-  for (a in 1:length(asc_list)){
-    asc <- asc_list[a]
-    if (grepl(dep, asc)){
-      depthlayers[d , paste0("a", colno)] <- asc
-        colno <- colno+1
-    }
-  }
-}
-write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cum_asclayerbydepth.csv")), row.names = TRUE)
-
-for (x in 1:nrow(depthlayers)){
-  asclay <- depthlayers[x, ]
-  asclay <- asclay[!is.na(asclay)]
-  asclay <- as.data.frame(rbind(asclay))
-  asclay$V1 <- as.character(asclay$V1)
-  asclay$V2 <- as.character(asclay$V2)
-  asclay$V3 <- as.character(asclay$V3)
-  asclay$V4 <- as.character(asclay$V4)
-  asclay$V5 <- as.character(asclay$V5)
-  asclaydep <- asclay$V1
-
-  asc2003 <- raster(asclay$V2)
-  asc2004 <- raster(asclay$V3)
-  asc2005 <- raster(asclay$V4)
-  asc2006 <- raster(asclay$V5)
-
-  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
-  avg <- calc(stkasc, fun = mean)
-
-  writeRaster(avg, paste0("../output/maxent/predictions/", mthna, "/avg_cum_", asclaydep, ".asc"), overwrite=TRUE)
-  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_cum_", asclaydep, ".png"))) 
-  dev.off()
+for (avg in 1:length(avgasc_list)){
+    cumasc <- raster(avgasc_list[avg])
+    ascname <- names(cumasc)
+    mpathr <- calc(cumasc, fun = binmpa)
+    mpa_plot <- plot(mpathr, png(paste0("../output/maxent/predictions/", mthna, "/binary_", ascname, ".png")))
+    dev.off() # stops automatic saving of the plot to a png
+    writeRaster(mpathr, paste0("../output/maxent/predictions/", mthna, "/binary_", ascname, ".asc"), overwrite = TRUE)
 }
@@
 
-```{r binary maps on averaged rasters}
-avgcumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cum.*.asc", full.names = TRUE)
-thresholds <- read.csv(paste0("../output/maxent/models/", mthna, "avgAIC.csv"), header = TRUE)
-thresholds <- tail(thresholds, 1) #subsset the last row
-mpa_thr <- thresholds$mpa_0.9_cum
-#print(mpa_thr)
+```{r binary map averaged 2D representation}
+binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_cloglog.*.asc", full.names = TRUE)
 
-for (avgcum in 1:length(avgcumasc_list)){
-    cumasc <- raster(avgcumasc_list[avgcum])
-    ascname <- names(cumasc)
-    #print(ascname)
-    cum_mpathr <- calc(cumasc, fun = binmpa)
-    mpa_plot <- plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".png")))
-    dev.off() # stops automatic saving of the plot to a png
-    writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna, "/", "binary_mpa_", ascname, ".asc"), overwrite = TRUE, datatype="LOG1S")
-}
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.asc"))
+plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.png")))
+dev.off()
 
-```
-
-now to stack average binary maps at depth into one 2D representation...
-
-```{r}
-
-#this below is to get an average score for all layers per year. probably not what you want
-ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_mpa_avg_cum_.*asc", full.names = TRUE)
-
-binaryavgstack <- stack(ascdepth_list)
-
-stsum <- sum(binaryavgstack, na.rm = TRUE)
-  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".asc")), overwrite=TRUE)
-  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".png"))) 
-  dev.off()
-stavg <- calc(binaryavgstack, fun = mean, na.rm = TRUE)
-  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".asc")), overwrite=TRUE)
-  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".png"))) 
-  dev.off()
 ```
@@
 
-```{r}
-test <- cellStats(binaryavgstack[[30]], sum)
-test
+```{r calc no presence cells per layer (avg)} 
+ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg.*asc", full.names = TRUE)
+binaryavgstack <- stack(ascdepth_list)
+stackstat <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+stackstat[ ,1] <- c(1:49)
+colnames(stackstat) <- c("layer", "cell_sum")
+
+for (lyr in 1:nlayers(binaryavgstack)){
+  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
+  stackstat$cell_sum[lyr] <- lyrsum
+}
+
+write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/cellcount_bydepth_avg.csv"), row.names = FALSE)
 ```
+
+a crude plot but ok to show slr
+
+```{r graph by depth (avg)}
+bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1), xlim = c(0, 300))
+trellis.device(device="png", filename = paste0("../output/maxent/predictions/", mthna, "/depthplot_avg.png"))
+print(bydepth)
+dev.off()
+```

### 2019-05-02 src/maxent_avec_xvalidation.Rmd
cbi added

@@
 
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
 
@@
 ```
+
+#cbi
+use enmSdm package? see exercise 10...
+positive values are good!
+
+should move this up to evaluation....
+
+```{r}
+
+for (a in 1:length(modelmonth)){
+  timeslice <- modelmonth[a]
+  yr <- str_sub(timeslice, -4) #the year being used for testing
+  predPres <- read.csv(paste0("../output/maxent/models/", a, "/species_samplePredictions.csv"), header = TRUE)
+  predBg <- read.csv(paste0("../output/maxent/models/", a, "/species_backgroundPredictions.csv"), header = TRUE)
+  predPres <- predPres$Cloglog.prediction
+  predBg <- predBg$Cloglog
+  cbi <- contBoyce(pres = predPres, bg = predBg)
+}
+
+predPres <- read.csv("../output/maxent/models/jan2003/species_samplePredictions.csv", header = TRUE)
+predBg <- read.csv("../output/maxent/models/jan2003/species_backgroundPredictions.csv", header = TRUE)
+predPres <- predPres$Cloglog.prediction
+predBg <- predBg$Cloglog
+cbi <- contBoyce(pres = predPres, bg = predBg)
+cbi
+```

### 2019-05-04 src/data_exploration.Rmd
cbi mess, running now for other months

@@
 
+```{r}
+#jun and oct without gear and with chl_surface in 
+vif_nogear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+write.csv(vif_nogear_cor_jun, "../output/bio/vif/monthly_vifs/vif_nogear_cor_jun_chl.csv", row.names = TRUE)
+vif_nogear_cor_jun
+
+vif_nogear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
+write.csv(vif_nogear_cor_jun, "../output/bio/vif/monthly_vifs/vif_nogear_cor_oct_chl.csv", row.names = TRUE)
+vif_nogear_cor_jun
+```
+
 in the presences, which rows are missing either temp_depth, salinity_depth, or o2_depth?

### 2019-05-04 src/maxent_avec_xvalidation.Rmd
cbi mess, running now for other months

--- 2019-05-02 src/maxent_avec_xvalidation.Rmd
@@
 options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr")
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm")
 source("../src/ipak.R")
@@
 mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chll_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
 julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
@@
 sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
 novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
@@
 ```{r month of interest -change}
-mthna <- "jan"
-mthno <- "1"
+mthna <- "mar"
+mthno <- "3"
 ```
@@
 
-y1998 = 2.75
-y1999 = 2.25
-y2000 = 2
-y2001 = 1.75
-y2002 = 2
-y2003 = 2
-y2004 = 2
-y2005 = 2.25
-y2006 = 1.75
-y2007 = 1.75
-y2008 = 2.25
-y2009 = 2.25
-y2010 = 2.5
-y2011 =  2.5
-
 ```{r final models plus predict}
@@
   testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  #mess_pr <- subset(train, occurrence == 1)
+  #mess_pr  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
   train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
 
-  if (yr == "2003" | yr == "2004"){
+  if (yr == "2003"){
     betaval <- 1
-  } else if (yr == "2005"){
+  } else if (yr == "1999" | yr == "2002" | yr == "2008" | yr == "2009") {
+    betaval <- 1.25
+  } else if (yr == "2005" | yr == "2006"){
     betaval <- 1.5
-  } else if (yr == "2006"){
+  } else if (yr == "2000"){
+    betaval <- 1.75
+  } else if (yr == "2001" | yr == "2007"){
     betaval <- 2.25
+  } else if (yr == "2010"){
+    betaval <- 3.5
   }
 
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
 
@@
 
+    #continuous boyce index
+    predPres <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    predBg <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
+    predPres <- predPres$Cloglog.prediction
+    predBg <- predBg$Cloglog
+    cbi <- contBoyce(pres = predPres, bg = predBg)
+
       aicetc$train_AUC <- trainauc #training only AUC - not tested
@@
       aicetc$test_sensitivity <- evsensitivity
+      aicetc$tss <- max(ev@TPR + ev@TNR) - 1
       aicetc$test_bic <- bic #not sure this is correct...
       aicetc$beta <- betaval
+      aicetc$cbi <- cbi
 
@@
     stackdeplst[[year]] <- stackdep
-    write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
+    write.csv(stackdep, paste0("../output/maxent/models/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)
     aea <- raster("../output/env/aea.tif")
@@
       nao_winter <- projectRaster(nao_winter, aea)
+      amo_sample <- raster(unlayer_yr$amo_sample[td])
+      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(amo_sample) <- "amo_sample"
+      amo_sample <- projectRaster(amo_sample, aea)
+      amo_winter <- raster(unlayer_yr$amo_winter[td])
+      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      names(amo_winter) <- "amo_winter"
+      amo_winter <- projectRaster(amo_winter, aea)
       stackn <- unlayer_yr$stackname[td]
-      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_winter)
+      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
+
       #stackSave(test,unlayer_yr$stackname[td])
-      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
+      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=cloglog')) 
       predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
@@
 
+      #mess
+
+      #messev <- dismo::mess(x = stack, v = mess_pr, full = FALSE)
+      #messplot <- plot(messev, png(paste0("../output/maxent/predictions/", mthna, "/", "mess_", stackn, ".png")))
+      #dev.off()
+      # not working how i'd like....
     }
@@
 stdep <- do.call(rbind, stackdeplst)
-write.csv(stdep, paste0("../output/maxent/predictions/", mthna, "/stack_depth.csv"), row.names = FALSE)
+write.csv(stdep, paste0("../output/maxent/models/", timeslice, "/stack_depth.csv"), row.names = FALSE)
 ```
@@
 binmpa <- function(x) {
-  ifelse(x <=  mpa_thr, 0,
-  ifelse(x >  mpa_thr, 1, NA))
+  ifelse(x <  mpa_thr, 0,
+  ifelse(x >=  mpa_thr, 1, NA))
   }
@@
   yr <- str_sub(timeslice, -4)
-  print(yr)
   thresholds <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
   mpa_thr <- thresholds$mpa_0.9_clog
-  print(mpa_thr)
   for (cum in 1:length(clogasc_list)){
@@
       ascname <- names(cumasc)
-      print(ascname)
       cum_mpathr <- calc(cumasc, fun = binmpa)
@@
 ```{r calc no presence cells per layer (avg)} 
-ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg.*asc", full.names = TRUE)
+ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_cloglog.*asc", full.names = TRUE)
 binaryavgstack <- stack(ascdepth_list)

### 2019-05-07 src/maxent_avec_xvalidation.Rmd
missing loop for betaval back in

@@
 ```{r month of interest -change}
-mthna <- "mar"
-mthno <- "3"
+mthna <- "may"
+mthno <- "5"
 ```
@@
   for (h in 1:nrow(betavallist)){
-    betamval <-  betavallist[h]
+    betaval <-  betavallist[h]
     mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
@@
 
-      aicetc$train_AUC <- trainauc
-      aicetc$no_trainng_points <- notrainingpoints
-      aicetc$test_coefficients <- nocoeff
-      aicetc$test_sum.of.logs <- sumsamplepred
-      aicetc$test_AUC <- evauc
-      aicetc$test_evcor <- evcor
-      aicetc$test_evpcor <- evpcor
-      aicetc$test_nopresence <- evnp
-      aicetc$test_nobackground <- evna
-      aicetc$test_maxsumsenspec <- evmaxtprtnr
-      aicetc$test_maxkap <- evmaxkap
-      aicetc$test_no_omission <- evno_omission
-      aicetc$test_prevalence <- evprevalence
-      aicetc$test_equal_sens_spec <- evequal_sens_spec
-      aicetc$test_sensitivity <- evsensitivity
-      aicetc$test_bic <- bic #not sure this is correct...
-      aicetc$beta <- betaval
+    for (i in 1:nrow(aicetc)){
+    if(aicetc$regularization[i] == betaval){
+      aicetc$train_AUC[i] <- trainauc
+      aicetc$no_trainng_points[i] <- notrainingpoints
+      aicetc$test_coefficients[i] <- nocoeff
+      aicetc$test_sum.of.logs[i] <- sumsamplepred
+      aicetc$test_AUC[i] <- evauc
+      aicetc$test_evcor[i] <- evcor
+      aicetc$test_evpcor[i] <- evpcor
+      aicetc$test_nopresence[i] <- evnp
+      aicetc$test_nobackground[i] <- evna
+      aicetc$test_maxsumsenspec[i] <- evmaxtprtnr
+      aicetc$test_maxkap[i] <- evmaxkap
+      aicetc$test_no_omission[i] <- evno_omission
+      aicetc$test_prevalence[i] <- evprevalence
+      aicetc$test_equal_sens_spec[i] <- evequal_sens_spec
+      aicetc$test_sensitivity[i] <- evsensitivity
+      aicetc$test_bic[i] <- bic #not sure this is correct...
+      aicetc$beta[i] <- betaval
+      #calculate minimal predicted area using ecospat
+      aicetc$mpa_0.9_cum[i] <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
 
@@
     aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
-
-    #calculate minimal predicted area using ecospat
-    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
     }
    }
-
-aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
-aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+  }
 
@@
 
+you need to go through each aic file manually....
+
 KEEP A NOTE OF THESE IN ONENOTE
 
+#run final model
 ```{r final models plus predict}
@@
 
-  if (yr == "2003"){
-    betaval <- 1
-  } else if (yr == "1999" | yr == "2002" | yr == "2008" | yr == "2009") {
-    betaval <- 1.25
-  } else if (yr == "2005" | yr == "2006"){
-    betaval <- 1.5
-  } else if (yr == "2000"){
+  if (yr == "2001" | yr == "2006" | yr == "2007"){
     betaval <- 1.75
-  } else if (yr == "2001" | yr == "2007"){
+  } else if (yr == "2000"| yr == "2002" | yr == "2003" | yr == "2004"){
+    betaval <- 2
+  } else if (yr == "1999" | yr == "2005" | yr == "2008" | yr == "2009"){
     betaval <- 2.25
-  } else if (yr == "2010"){
-    betaval <- 3.5
+  } else if (yr == "2010" | yr == "2011"){
+    betaval <- 2.5
+  } else if (yr == "1998"){
+    betaval <- 2.75
   }
@@
 
-#binary maps - individual year layers
-
-there seem to be two primary ways to create binary maps.
-the most common for conservation purposes seems to be to use maxsumspensspec values, however MPA also comes recommended (but is def less common).
-
-the thresholds are in the monthavgAIC.csv file in the model folder (not model/month)
-
-```{r set up threshold binary function}
-binmpa <- function(x) {
-  ifelse(x <  mpa_thr, 0,
-  ifelse(x >=  mpa_thr, 1, NA))
-  }
-```
-
-```{r binary maps on individual layers cloglog}
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
-
-for (modtime in 1:nrow(modelmonth)){
-  timeslice <- modelmonth[modtime]
-  print(timeslice)
-  yr <- str_sub(timeslice, -4)
-  thresholds <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-  mpa_thr <- thresholds$mpa_0.9_clog
-  for (cum in 1:length(clogasc_list)){
-    cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
-    if(yr == cumascyr){
-      cumasc <- raster(clogasc_list[cum])
-      ascname <- names(cumasc)
-      cum_mpathr <- calc(cumasc, fun = binmpa)
-      writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna,  "/binary_", ascname, ".asc"), overwrite = TRUE)
-      plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/binary_", ascname, ".png")), overwrite = TRUE)
-      dev.off() # stops automatic saving of the plot to a png
-    }
-  }
-}
-```
-
-now for each year create a 2-d representation of suitable habitat
-
-need to stack the rasters by year
-run cellStats (sum)
-save the output as they year
-
-```{r binary map per year 2D representation}
-binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_.*.asc", full.names = TRUE)
-yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryclog_list)))
-colnames(yrlayer) <- c("layer", "year")
-
-for (yrlyr in 1:length(binaryclog_list)){
-  yearoflyr <- str_sub(binaryclog_list[yrlyr], start = 51, end = 54) #selects the year from the file name
-  yrlayer$year[yrlyr] <- yearoflyr
-  yrlayer$layer[yrlyr] <- binaryclog_list[yrlyr]
-}
-
-y_list <- split(yrlayer, yrlayer$year)
-noyrs <- unique(yrlayer$year)
-
-for (binyr in (1:length(y_list))){
-  binstack <- stack(y_list[[binyr]]$layer)
-  year <- y_list[[binyr]]$year[1]
-  stacksum <- sum(binstack, na.rm = TRUE)
-  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".asc"))
-  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".png")))
-  dev.off()
-}
-
-```
-
-#using averageing
-
-ok! now to average results:
-
-- the AICforMaxentReg.csv 
-- the maxentResults.csv
-- the asc files per depth layer
-
-open all maxentReg files
-
 ```{r avg maxentReg}
@@
 
-aic <- rbind(aic2003, aic2004, aic2005, aic2006)
-maxr <- rbind(maxr2003, maxr2004, maxr2005, maxr2006)
+aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
+maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)
 
@@
 
+#binary maps - individual year layers
+
+there seem to be two primary ways to create binary maps.
+the most common for conservation purposes seems to be to use maxsumspensspec values, however MPA also comes recommended (but is def less common).
+
+the thresholds are in the monthavgAIC.csv file in the model folder (not model/month)
+
+```{r set up threshold binary function}
+binmpa <- function(x) {
+  ifelse(x <  mpa_thr, 0,
+  ifelse(x >=  mpa_thr, 1, NA))
+  }
+```
+
+```{r binary maps on individual layers cloglog}
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+
+for (modtime in 1:nrow(modelmonth)){
+  timeslice <- modelmonth[modtime]
+  print(timeslice)
+  yr <- str_sub(timeslice, -4)
+  thresholds <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+  mpa_thr <- thresholds$mpa_0.9_clog
+  for (cum in 1:length(clogasc_list)){
+    cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
+    if(yr == cumascyr){
+      cumasc <- raster(clogasc_list[cum])
+      ascname <- names(cumasc)
+      cum_mpathr <- calc(cumasc, fun = binmpa)
+      writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna,  "/binary_", ascname, ".asc"), overwrite = TRUE)
+      plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/binary_", ascname, ".png")), overwrite = TRUE)
+      dev.off() # stops automatic saving of the plot to a png
+    }
+  }
+}
+```
+
+now for each year create a 2-d representation of suitable habitat
+
+need to stack the rasters by year
+run cellStats (sum)
+save the output as they year
+
+```{r binary map per year 2D representation}
+binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_cloglog.*.asc", full.names = TRUE)
+yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryclog_list)))
+colnames(yrlayer) <- c("layer", "year")
+
+for (yrlyr in 1:length(binaryclog_list)){
+  yearoflyr <- str_sub(binaryclog_list[yrlyr], start = 51, end = 54) #selects the year from the file name
+  yrlayer$year[yrlyr] <- yearoflyr
+  yrlayer$layer[yrlyr] <- binaryclog_list[yrlyr]
+}
+
+y_list <- split(yrlayer, yrlayer$year)
+noyrs <- unique(yrlayer$year)
+
+for (binyr in (1:length(y_list))){
+  binstack <- stack(y_list[[binyr]]$layer)
+  year <- y_list[[binyr]]$year[1]
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".png")))
+  dev.off()
+}
+
+```
+
+#using averageing
+
+ok! now to average results:
+
+- the AICforMaxentReg.csv 
+- the maxentResults.csv
+- the asc files per depth layer
+
+open all maxentReg files
+
 for this bit may need to stack rasters by depth... this will need some working with the name
@@
 clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
-stdep <- read.csv(paste0("../output/maxent/predictions/", mthna, "/stack_depth.csv"), header = TRUE)
+stdep <- read.csv(paste0("../output/maxent/models/", timeslice, "/stack_depth.csv"), header = TRUE)
 depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
@@
   asclay$V5 <- as.character(asclay$V5)
+  asclay$V6 <- as.character(asclay$V6)
+  asclay$V7 <- as.character(asclay$V7)
+  asclay$V8 <- as.character(asclay$V8)
+  asclay$V9 <- as.character(asclay$V9)
+  asclay$V10 <- as.character(asclay$V10)
+  asclay$V11 <- as.character(asclay$V11)
+  asclay$V12 <- as.character(asclay$V12)
+  asclay$V13 <- as.character(asclay$V13)
+  asclay$V14 <- as.character(asclay$V14)
+  asclay$V15 <- as.character(asclay$V15)
+
   asclaydep <- asclay$V1
 
-  asc2003 <- raster(asclay$V2)
-  asc2004 <- raster(asclay$V3)
-  asc2005 <- raster(asclay$V4)
-  asc2006 <- raster(asclay$V5)
+  asc1998 <- raster(asclay$V2)
+  asc1999 <- raster(asclay$V3)
+  asc2000 <- raster(asclay$V4)
+  asc2001 <- raster(asclay$V5)
+  asc2002 <- raster(asclay$V6)
+  asc2003 <- raster(asclay$V7)
+  asc2004 <- raster(asclay$V8)
+  asc2005 <- raster(asclay$V9)
+  asc2006 <- raster(asclay$V10)
+  asc2007 <- raster(asclay$V11)
+  asc2008 <- raster(asclay$V12)
+  asc2009 <- raster(asclay$V13)
+  asc2010 <- raster(asclay$V14)
+  asc2011 <- raster(asclay$V15)
 
-  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
+  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)
+
   avg <- mean(stkasc, na.rm = TRUE)
@@
 stacksum <- sum(binstack, na.rm = TRUE)
-writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.asc"))
+writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.asc"), overwrite=TRUE)
 plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.png")))
@@
 ```{r graph by depth (avg)}
-bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1), xlim = c(0, 300))
+bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1))
 trellis.device(device="png", filename = paste0("../output/maxent/predictions/", mthna, "/depthplot_avg.png"))

### 2019-05-07 src/maxent_avec_xvalidation.Rmd
breath... working...

@@
 ```{r pre-install & load packages, include=FALSE}
-options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
+options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
 packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm")
@@
 mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chll_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
 julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
@@
     betaval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c('plots=FALSE', "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
 
@@
       aicetc$no_trainng_points[i] <- notrainingpoints
-      aicetc$test_coefficients[i] <- nocoeff
-      aicetc$test_sum.of.logs[i] <- sumsamplepred
+      aicetc$coefficients[i] <- nocoeff
+      aicetc$sum.of.logs[i] <- sumsamplepred
       aicetc$test_AUC[i] <- evauc
@@
 
-    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
-    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+    aicetc$AIC[i] <- ((2*aicetc$coefficients[i])-(2*aicetc$sum.of.logs[i]))
+    aicetc$AICc[i] <- (-2*(aicetc$sum.of.logs[i])+((2*aicetc$coefficients[i])*(aicetc$no_trainng_points[i]/(aicetc$no_trainng_points[i] - aicetc$coefficients[i]-1))))
     }
@@
 
-  if (yr == "2001" | yr == "2006" | yr == "2007"){
+  if (yr == "2006"){
+    betaval <- 0.25
+  } else if (yr == "2005"| yr == "2008"){
+    betaval <- 1.25
+  } else if (yr == "2003" | yr == "2007"){
+    betaval <- 1.5
+  } else if (yr == "1998" | yr == "1999" | yr == "2002" | yr == "2010" | yr == "2011"){
     betaval <- 1.75
-  } else if (yr == "2000"| yr == "2002" | yr == "2003" | yr == "2004"){
+  } else if (yr == "2001" | yr == "2009"){
     betaval <- 2
-  } else if (yr == "1999" | yr == "2005" | yr == "2008" | yr == "2009"){
+  } else if (yr == "2000" | yr == "2004"){
     betaval <- 2.25
-  } else if (yr == "2010" | yr == "2011"){
-    betaval <- 2.5
-  } else if (yr == "1998"){
-    betaval <- 2.75
   }
@@
       aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
-      aicetc$test_coefficients <- nocoeff #this is to calculate the AIC
-      aicetc$test_sum.of.logs <- sumsamplepred #this is to calculate the AIC
+      aicetc$coefficients <- nocoeff #this is to calculate the AIC
+      aicetc$sum.of.logs <- sumsamplepred #this is to calculate the AIC
       aicetc$test_AUC <- evauc #testing AUC
@@
     aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
-    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$no_trainng_points/(aicetc$no_trainng_points - aicetc$coefficients-1))))
 

### 2019-05-10 src/maxent_avec_xvalidation.Rmd
minor adjustment for year variation (presence with data)

@@
 ```{r month of interest -change}
-mthna <- "may"
-mthno <- "5"
+mthna <- "dec"
+mthno <- "12"
 ```
@@
 
+note will need to reajust background points too as some years may no longer have presence data
+
 ```{r remove na vals}
 prback <- na.omit(prback)
+presyrs <-  subset(prback, occurrence == 1) #this is to get years with complete presence data 
+presyrs <- unique(presyrs$year) #make a vector of unique years
+prback <- prback[prback$year %in% presyrs, ] #keep only the data for which there is precence years
+
 ```
@@
 
-```{r train and test inc AIC for reg val}
+#```{r train and test inc AIC for reg val}
+presyrs <-  #this is to get years with complete presence data
+
 modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
@@
 
-```
+#```
 
@@
 
-  if (yr == "2006"){
+  if (yr == "1998"){
     betaval <- 0.25
-  } else if (yr == "2005"| yr == "2008"){
+  } else if (yr == "2001"){
+    betaval <- 0.75
+  } else if (yr == "2002" | yr == "2003"| yr == "2004"| yr == "2006"| yr == "2011"){
+    betaval <- 1
+  } else if (yr == "2005"){
     betaval <- 1.25
-  } else if (yr == "2003" | yr == "2007"){
+  } else if (yr == "1999" | yr == "2000"| yr == "2009"){
     betaval <- 1.5
-  } else if (yr == "1998" | yr == "1999" | yr == "2002" | yr == "2010" | yr == "2011"){
+  } else if (yr == "2007"){
     betaval <- 1.75
-  } else if (yr == "2001" | yr == "2009"){
+  } else if (yr == "2010"){
     betaval <- 2
-  } else if (yr == "2000" | yr == "2004"){
+  } else if (yr == "2008"){
     betaval <- 2.25

### 2019-05-13 src/maxent_avec_xvalidation.Rmd
probability categorised maps and plot finished

@@
 
-#```{r train and test inc AIC for reg val}
+```{r train and test inc AIC for reg val}
 presyrs <-  #this is to get years with complete presence data
@@
 
-#```
+```
 
@@
 ```
+
+#probability selected maps
+
+```{r habitat classification on probability - 3 classifications}
+
+binprob <- function(x){
+  ifelse(x >= 0.8, 3,
+  ifelse(x >= 0.7 & x <0.8, 2,
+  ifelse(x >= 0.6 & x <0.7, 1,
+  ifelse(x < 0.6, 0, NA)))) 
+}
+
+  modelmonth <- cbind(paste0(mthna, unique(prback$year)))
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+
+  for (modtime in 1:nrow(modelmonth)){
+    timeslice <- modelmonth[modtime]
+    yr <- str_sub(timeslice, -4)
+    for (cum in 1:length(clogasc_list)){
+      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
+      if(yr == cumascyr){
+        cumasc <- raster(clogasc_list[cum])
+        ascname <- names(cumasc)
+        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_prob_", ascname, ".asc"), overwrite = TRUE)
+        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_prob_", ascname, ".png")), overwrite = TRUE)
+        dev.off() # stops automatic saving of the plot to a png
+      }
+    }
+  }
+```
+
+```{r habitat classification on probability - high only}
+
+binprob <- function(x){
+  ifelse(x >= 0.8, 1,
+  ifelse(x < 0.8, 0, NA)) 
+}
+
+modelmonth <- cbind(paste0(mthna, unique(prback$year)))
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+
+  for (modtime in 1:nrow(modelmonth)){
+    timeslice <- modelmonth[modtime]
+    yr <- str_sub(timeslice, -4)
+    for (cum in 1:length(clogasc_list)){
+      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
+      if(yr == cumascyr){
+        cumasc <- raster(clogasc_list[cum])
+        ascname <- names(cumasc)
+        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_highprob_", ascname, ".asc"), overwrite = TRUE)
+        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_highprob_", ascname, ".png")), overwrite = TRUE)
+        dev.off() # stops automatic saving of the plot to a png
+      }
+    }
+  }
+
+```
+
+```{r habitat classification on probability - med only}
+binprob <- function(x){
+  ifelse(x >= 0.7 & x <0.8, 1,
+  ifelse(x < 0.7, 0, NA)) 
+}
+
+modelmonth <- cbind(paste0(mthna, unique(prback$year)))
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+
+  for (modtime in 1:nrow(modelmonth)){
+    timeslice <- modelmonth[modtime]
+    yr <- str_sub(timeslice, -4)
+    for (cum in 1:length(clogasc_list)){
+      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
+      if(yr == cumascyr){
+        cumasc <- raster(clogasc_list[cum])
+        ascname <- names(cumasc)
+        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_medprob_", ascname, ".asc"), overwrite = TRUE)
+        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_medprob_", ascname, ".png")), overwrite = TRUE)
+        dev.off() # stops automatic saving of the plot to a png
+      }
+    }
+  }
+```
+
+```{r habitat classification on probability - low only}
+binprob <- function(x){
+  ifelse(x >= 0.6 & x <0.7, 1,
+  ifelse(x < 0.6, 0, NA)) 
+}
+
+modelmonth <- cbind(paste0(mthna, unique(prback$year)))
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+
+  for (modtime in 1:nrow(modelmonth)){
+    timeslice <- modelmonth[modtime]
+    yr <- str_sub(timeslice, -4)
+    for (cum in 1:length(clogasc_list)){
+      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
+      if(yr == cumascyr){
+        cumasc <- raster(clogasc_list[cum])
+        ascname <- names(cumasc)
+        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_lowprob_", ascname, ".asc"), overwrite = TRUE)
+        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_lowprob_", ascname, ".png")), overwrite = TRUE)
+        dev.off() # stops automatic saving of the plot to a png
+      }
+    }
+  }
+```
+
+now you need to do summary maps....
+
+```{r high probability map per year 2D representation}
+binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_highprob_cloglog.*.asc", full.names = TRUE)
+yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
+colnames(yrlayer) <- c("layer", "year")
+
+for (yrlyr in 1:length(binaryprob_list)){
+  yearoflyr <- str_sub(binaryprob_list[yrlyr], start = 60, end = 63) #selects the year from the file name
+  yrlayer$year[yrlyr] <- yearoflyr
+  yrlayer$layer[yrlyr] <- binaryprob_list[yrlyr]
+}
+
+y_list <- split(yrlayer, yrlayer$year)
+noyrs <- unique(yrlayer$year)
+
+for (binyr in (1:length(y_list))){
+  binstack <- stack(y_list[[binyr]]$layer)
+  year <- y_list[[binyr]]$year[1]
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/highprob_2d_", year, ".asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/highprob_2d_", year, ".png")))
+  dev.off()
+}
+
+```
+
+```{r med probability map per year 2D representation}
+binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_medprob_cloglog.*.asc", full.names = TRUE)
+yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
+colnames(yrlayer) <- c("layer", "year")
+
+for (yrlyr in 1:length(binaryprob_list)){
+  yearoflyr <- str_sub(binaryprob_list[yrlyr], start = 59, end = 62) #selects the year from the file name
+  yrlayer$year[yrlyr] <- yearoflyr
+  yrlayer$layer[yrlyr] <- binaryprob_list[yrlyr]
+}
+
+y_list <- split(yrlayer, yrlayer$year)
+noyrs <- unique(yrlayer$year)
+
+for (binyr in (1:length(y_list))){
+  binstack <- stack(y_list[[binyr]]$layer)
+  year <- y_list[[binyr]]$year[1]
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/medprob_2d_", year, ".asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/medprob_2d_", year, ".png")))
+  dev.off()
+}
+
+```
+
+```{r low probability map per year 2D representation}
+binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_lowprob_cloglog.*.asc", full.names = TRUE)
+yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
+colnames(yrlayer) <- c("layer", "year")
+
+for (yrlyr in 1:length(binaryprob_list)){
+  yearoflyr <- str_sub(binaryprob_list[yrlyr], start = 59, end = 62) #selects the year from the file name
+  yrlayer$year[yrlyr] <- yearoflyr
+  yrlayer$layer[yrlyr] <- binaryprob_list[yrlyr]
+}
+
+y_list <- split(yrlayer, yrlayer$year)
+noyrs <- unique(yrlayer$year)
+
+for (binyr in (1:length(y_list))){
+  binstack <- stack(y_list[[binyr]]$layer)
+  year <- y_list[[binyr]]$year[1]
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/lowprob_2d_", year, ".asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/lowprob_2d_", year, ".png")))
+  dev.off()
+}
+
+```
+
+now average probability split maps - again low, med, high and all
+
+```{r habitat classification on average probability - 3 classifications}
+
+binprob <- function(x){
+  ifelse(x >= 0.8, 3,
+  ifelse(x >= 0.7 & x <0.8, 2,
+  ifelse(x >= 0.6 & x <0.7, 1,
+  ifelse(x < 0.6, 0, NA)))) 
+}
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
+
+    for (cum in 1:length(clogasc_list)){
+      cumasc <- raster(clogasc_list[cum])
+      ascname <- names(cumasc)
+      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avgprob_", ascname, ".asc"), overwrite = TRUE)
+      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avgprob_", ascname, ".png")), overwrite = TRUE)
+      dev.off() # stops automatic saving of the plot to a png
+    }
+
+```
+
+```{r habitat classification on average probability - high}
+
+binprob <- function(x){
+  ifelse(x >= 0.8, 1,
+  ifelse(x < 0.8, 0, NA))  
+}
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
+
+    for (cum in 1:length(clogasc_list)){
+      cumasc <- raster(clogasc_list[cum])
+      ascname <- names(cumasc)
+      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avg_highprob_", ascname, ".asc"), overwrite = TRUE)
+      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_highprob_", ascname, ".png")), overwrite = TRUE)
+      dev.off() # stops automatic saving of the plot to a png
+    }
+
+```
+
+```{r habitat classification on average probability - med}
+
+binprob <- function(x){
+  ifelse(x >= 0.7 & x <0.8, 1,
+  ifelse(x < 0.7, 0, NA)) 
+}
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
+
+    for (cum in 1:length(clogasc_list)){
+      cumasc <- raster(clogasc_list[cum])
+      ascname <- names(cumasc)
+      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avg_medprob_", ascname, ".asc"), overwrite = TRUE)
+      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_medprob_", ascname, ".png")), overwrite = TRUE)
+      dev.off() # stops automatic saving of the plot to a png
+    }
+
+```
+
+```{r habitat classification on average probability - low}
+
+binprob <- function(x){
+  ifelse(x >= 0.6 & x <0.7, 1,
+  ifelse(x < 0.6, 0, NA)) 
+}
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
+
+    for (cum in 1:length(clogasc_list)){
+      cumasc <- raster(clogasc_list[cum])
+      ascname <- names(cumasc)
+      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avg_lowprob_", ascname, ".asc"), overwrite = TRUE)
+      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_lowprob_", ascname, ".png")), overwrite = TRUE)
+      dev.off() # stops automatic saving of the plot to a png
+    }
+
+```
+
+```{r high probability average map per year 2D representation}
+binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_highprob.*.asc", full.names = TRUE)
+
+binstack <- stack(binaryprob_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stastaccksum, paste0("../output/maxent/predictions/", mthna, "/binary_highavg_2d.asc"), overwrite=TRUE)
+plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_highprob_avg_2d.png")))
+dev.off()
+```
+```{r med probability average map per year 2D representation}
+binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_medprob.*.asc", full.names = TRUE)
+
+binstack <- stack(binaryprob_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_medavg_2d.asc"), overwrite=TRUE)
+plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_medprob_avg_2d.png")))
+dev.off()
+```
+
+```{r low probability average map per year 2D representation}
+binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_lowprob.*.asc", full.names = TRUE)
+
+binstack <- stack(binaryprob_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_lowavg_2d.asc"), overwrite=TRUE)
+plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_lowprob_avg_2d.png")))
+dev.off()
+```
+
+first need some statistcs... for each layer how many cells have a value of 1...
+
+```{r calc no presence cells per layer (avg probability)} 
+ascdepth_list_high <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_highprob_.*asc", full.names = TRUE)
+ascdepth_list_med <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_medprob_.*asc", full.names = TRUE)
+ascdepth_list_low <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_lowprob_.*asc", full.names = TRUE)
+
+binaryavgstack <- stack(ascdepth_list_high)
+stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+stackstathigh[ ,1] <- c(1:49)
+colnames(stackstathigh) <- c("layer", "high_cell_sum")
+
+for (lyr in 1:nlayers(binaryavgstack)){
+  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
+  stackstathigh$high_cell_sum[lyr] <- lyrsum
+}
+
+binaryavgstack <- stack(ascdepth_list_med)
+stackstatmed <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+stackstatmed[ ,1] <- c(1:49)
+colnames(stackstatmed) <- c("layer", "med_cell_sum")
+
+for (lyr in 1:nlayers(binaryavgstack)){
+  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
+  stackstatmed$med_cell_sum[lyr] <- lyrsum
+}
+
+binaryavgstack <- stack(ascdepth_list_low)
+stackstatlow <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+stackstatlow[ ,1] <- c(1:49)
+colnames(stackstatlow) <- c("layer", "low_cell_sum")
+
+for (lyr in 1:nlayers(binaryavgstack)){
+  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
+  stackstatlow$low_cell_sum[lyr] <- lyrsum
+}
+
+stackstat <- cbind(stackstathigh, stackstatmed[ , 2], stackstatlow[ ,2])
+colnames(stackstat) <- c("layer", "high", "med", "low")
+stackstat <- melt(stackstat, id = "layer")
+
+write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/probability_cellcount_bydepth_avg.csv"), row.names = FALSE)
+```
+
+a crude plot but ok to show slr
+
+```{r graph by depth - prob plot (avg)}
+bydepth <- ggplot(data = stackstat, aes(x = layer, y = value, fill = variable)) + geom_bar(stat = "identity") +   scale_x_reverse() +
+coord_flip()
+png(paste0("../output/maxent/predictions/", mthna, "/depthplot_probavg.png"))
+print(bydepth)
+dev.off()
+```

### 2019-05-13 src/maxent_avec_xvalidation.Rmd
ops

@@
 ```{r month of interest -change}
-mthna <- "dec"
-mthno <- "12"
+mthna <- "jan"
+mthno <- "1"
 ```
@@
 stacksum <- sum(binstack, na.rm = TRUE)
-writeRaster(stastaccksum, paste0("../output/maxent/predictions/", mthna, "/binary_highavg_2d.asc"), overwrite=TRUE)
+writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_highavg_2d.asc"), overwrite=TRUE)
 plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_highprob_avg_2d.png")))

### 2019-05-14 src/maxent_avec_xvalidation.Rmd
permutation and jackknife plots

@@
 options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm")
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr")
 source("../src/ipak.R")
@@
 ```{r month of interest -change}
-mthna <- "jan"
-mthno <- "1"
+mthna <- "oct"
+mthno <- "10"
 ```
@@
 
+summary table of average maxent results for each month
+```{r average AIC etc for each month in one table}
+aic_lst <- list.files("../output/maxent/models/", pattern = "avgAIC.csv", full.names = TRUE)
+aicsum <- data.frame(matrix(ncol = 11, nrow = length(aic_lst)))
+colnames(aicsum) <- c("Month", "BetaVal", "AICc", "Test AUC", "Training AUC", "Correlation", "Correlation P-Val", "TSS", "CBI", "No Omissions", "Prevalence")
+
+for (a in 1:length(aic_lst)){
+  aicr <- read.csv(aic_lst[a])
+  mth <- str_sub(aic_lst[a], start = 25, end = 27) #selects the month from the filename
+  aicsum$Month[a] <- mth
+  aicres <- tail(aicr, 1)
+  aicsum$BetaVal[a] <- aicres$beta
+  aicsum$AICc[a] <- aicres$AICc
+  aicsum$`Test AUC`[a] <- aicres$test_AUC
+  aicsum$`Training AUC`[a] <- aicres$train_AUC
+  aicsum$Correlation[a] <- aicres$test_evcor
+  aicsum$`Correlation P-Val` <- aicres$test_evpcor
+  aicsum$TSS[a] <- aicres$tss
+  aicsum$CBI[a] <- aicres$cbi
+  aicsum$`No Omissions`[a] <- aicres$test_no_omission
+  aicsum$Prevalence[a] <- aicres$test_prevalence
+}
+
+write.csv(aicsum, "../output/maxent/models/aicsummary.csv", row.names = FALSE)
+```
+
+stacked permutation importance plot on average
+
+```{r permutation & jack data}
+maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
+maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
+perm <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
+colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
+jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
+colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
+
+for (m in 1:length(maxr_lst)){
+  maxr <- read.csv(maxr_lst[m])
+  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
+  perm$Month[m] <- mth
+  jack$Month[m] <- mth
+  maxt <- tail(maxr, 1)
+  perm$amo_sample[m] <- maxt$amo_sample.permutation.importance
+  perm$amo_winter[m] <- maxt$amo_winter.permutation.importance
+  perm$chl_surface[m] <- maxt$chl_surface.permutation.importance
+  perm$nao_prev[m] <- maxt$nao_prev.permutation.importance
+  perm$nao_sample[m] <- maxt$nao_sample.permutation.importance
+  perm$nao_winter[m] <- maxt$nao_winter.permutation.importance
+  perm$o2_depth[m] <- maxt$o2_depth.permutation.importance
+  perm$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
+  perm$temp_depth[m] <- maxt$temp_depth.permutation.importance
+
+  jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
+  jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
+  jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
+  jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
+  jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
+  jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
+  jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
+  jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
+  jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
+}
+
+# now for january
+
+maxr_lstj <- list.files("../output/maxent/models/", pattern = "janavgmaxentresults.csv", full.names = TRUE)
+permj <- data.frame(matrix(ncol = 7, nrow = length(maxr_lstj)))
+colnames(permj) <- c("Month", "chl_surface", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
+jackj <- data.frame(matrix(ncol = 7, nrow = length(maxr_lstj)))
+colnames(jackj) <- c("Month", "chl_surface", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
+
+for (m in 1:length(maxr_lstj)){
+  maxr <- read.csv(maxr_lstj[m])
+  mth <- str_sub(maxr_lstj[m], start = 25, end = 27) #selects the month from the filename
+  permj$Month[m] <- mth
+  jackj$Month[m] <- mth
+  maxt <- tail(maxr, 1)
+  permj$chl_surface[m] <- maxt$chl_surface.permutation.importance
+  permj$nao_sample[m] <- maxt$nao_sample.permutation.importance
+  permj$nao_winter[m] <- maxt$nao_winter.permutation.importance
+  permj$o2_depth[m] <- maxt$o2_depth.permutation.importance
+  permj$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
+  permj$temp_depth[m] <- maxt$temp_depth.permutation.importance
+
+  jackj$chl_surface[m] <- maxt$Training.gain.without.chl_surface
+  jackj$nao_sample[m] <- maxt$Training.gain.without.nao_sample
+  jackj$nao_winter[m] <- maxt$Training.gain.without.nao_winter
+  jackj$o2_depth[m] <- maxt$Training.gain.without.o2_depth
+  jackj$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
+  jackj$temp_depth[m] <- maxt$Training.gain.without.temp_depth
+}
+
+permu <- rbind.fill(perm, permj)
+jackk <- rbind.fill(jack, jackj)
+
+mthord <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
+permu <- permu[match(mthord, permu$Month), ] #reorder based on mthord
+jackk <- jackk[match(mthord, jackk$Month), ] #reorder based on mthord
+
+write.csv(permu, "../output/maxent/models/permutation_avg.csv", row.names = FALSE)
+write.csv(jackk, "../output/maxent/models/jack_avg.csv", row.names = FALSE)
+
+```
+
+```{r}
+maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
+maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
+jack_df <- list()
+maxr_lstj <- list.files("../output/maxent/models/", pattern = "janavgmaxentresults.csv", full.names = TRUE)
+jack_dfj <- list()
+
+for (m in 1:length(maxr_lst)){
+  res <- read.csv(maxr_lst[m], header = TRUE)
+  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
+  res_df <- tail(res, 1)
+  var_df <- res_df[,c(grep("Training.gain.with*",names(res_df)))]
+  var_df$id <- 1:nrow(var_df)
+  var_df <- melt(var_df, id.vars=c("id"))
+  jack_with <- var_df[grep("with.only",var_df$variable),]
+  jack_with$Type_of_Run <- "With only the variable"
+  jack_without <- var_df[grep("without",var_df$variable),]
+  jack_without$Type_of_Run <- "Without the variable"
+  gain_all <- res_df[c(grep("Regularized.",names(res_df)))]
+  gain_all$id <- 1:nrow(gain_all)
+  gain_all <- melt(gain_all, id.vars=c("id"))
+  gain_all$Type_of_Run <- "With all variables"
+  gain_all$variable <- "all_variables"
+  jack_table <- rbind(jack_with,jack_without,gain_all)
+  jack_table$state <- mth
+
+  jack_table$variable <- gsub(".*[.]","",jack_table$variable)
+  jack_df[[m]] <- jack_table
+}
+
+var_jack <- do.call(rbind, jack_df)
+
+for (m in 1:length(maxr_lstj)){
+  res <- read.csv(maxr_lstj[m], header = TRUE)
+  mth <- str_sub(maxr_lstj[m], start = 25, end = 27) #selects the month from the filename
+  res_df <- tail(res, 1)
+  var_df <- res_df[,c(grep("Training.gain.with*",names(res_df)))]
+  var_df$id <- 1:nrow(var_df)
+  var_df <- melt(var_df, id.vars=c("id"))
+  jack_with <- var_df[grep("with.only",var_df$variable),]
+  jack_with$Type_of_Run <- "With only the variable"
+  jack_without <- var_df[grep("without",var_df$variable),]
+  jack_without$Type_of_Run <- "Without the variable"
+  gain_all <- res_df[c(grep("Regularized.",names(res_df)))]
+  gain_all$id <- 1:nrow(gain_all)
+  gain_all <- melt(gain_all, id.vars=c("id"))
+  gain_all$Type_of_Run <- "With all variables"
+  gain_all$variable <- "all_variables"
+  jack_table <- rbind(jack_with,jack_without,gain_all)
+  jack_table$state <- mth
+
+  jack_table$variable <- gsub(".*[.]","",jack_table$variable)
+  jack_dfj[[m]] <- jack_table
+}
+
+var_jackj <- do.call(rbind, jack_dfj)
+
+var_jack$state <- factor(var_jack$state,levels=c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"))
+var_jack$Type_of_Run <- factor(var_jack$Type_of_Run,levels=c("With all variables", "Without the variable","With only the variable"))
+var_jackj$state <- factor(var_jackj$state,levels="jan")
+var_jackj$Type_of_Run <- factor(var_jackj$Type_of_Run,levels=c("With all variables", "Without the variable","With only the variable"))
+a <- subset(var_jack, state == "mar" | state == "apr" |state == "may" | state == "jun" | state == "jul")
+b <- subset(var_jack, state == "aug" | state == "sep" |state == "oct" | state == "nov" | state == "dec")
+c <- subset(var_jackj, state == "jan")
+
+plotlst <- list(a, b, c)
+
+color_type_1 <- c("#000000", "#009E73","#999999")
+
+for (p in 1:length(plotlst)){
+  pt <- p
+  dev.new()
+  jack <- ggplot(plotlst[[p]], aes(variable,value))+
+  geom_bar(aes(fill=Type_of_Run),position="dodge",stat="identity")+ coord_flip()+ facet_grid(state~., scales="free", space = "free")+
+  theme_bw(14)+
+  scale_fill_manual(values=color_type_1,guide=guide_legend(reverse=TRUE))+
+  theme(axis.text=element_text(size=8)) +
+  theme(legend.title=element_blank())+
+  theme(legend.position="top")+
+  theme(strip.text=element_text(size=12))+
+  theme(strip.text.y = element_text(angle = 0)) +
+  labs(x="Variable",y="Jackknife of Regularized Training Gain")
+  ggsave(filename = paste0("../output/maxent/models/jackknife_plot_", pt, ".png"), plot = jack, dpi = 900, height = 9, width = 7)
+  dev.off()
+}
+
+#png("../output/maxent/models/jackknife_plot.png")
+#print(jack)
+#dev.off()
+```
+
+```{r permulation plot}
+permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
+perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") + scale_fill_brewer(palette = "Set1")
+png("../output/maxent/models/permutation_plot.png")
+print(perpl)
+dev.off()
+```
+
+```{r jackknife plot}
+color_type_1 <- c("#000000", "#009E73","#999999")
+jackpl <- ggplot(jackk, aes(x = variable, y = value)) + geom_bar(aes(fill=Type_of_Run),position="dodge",stat="identity")+
+```
+
 #binary maps - individual year layers
@@
 ```
+
+#which cells are used in all months?
+
+start with mpa binary map first
+
+```{r}
+
+```

### 2019-05-14 src/maxent_avec_xvalidation.Rmd
stoopid merging

@@
 ```{r}
-
-```
+mthlst <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for(m in 1:length(mthlst)){
+  mthna <- mthlst[m]
+  mpabin_lst <- list.files(paste0("../output/maxent/predictions/", mthna), pattern =  "^binary_avg_cloglog.*.asc", full.names = TRUE)
+
+  stdep <- read.csv("../output/maxent/models/dec1998/stack_depth.csv", header = TRUE)
+  depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
+  names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
+  depthlayers$depth <- as.character(depthlayers$depth)
+
+  colno <- 1
+
+  for (d in 1:nrow(depthlayers)){
+    dep <- depthlayers$depth[d]
+    for (a in 1:length(mpabin_lst)){
+      asc <- mpabin_lst[a]
+      if (grepl(dep, asc)){
+        depthlayers[d , paste0("a", colno)] <- asc
+          colno <- colno+1
+      }
+    }
+  }
+  write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "_binary_avg_asclayerbydepth.csv")), row.names = TRUE)
+}
+
+a <- read.csv("../output/maxent/predictions/apr_binary_avg_asclayerbydepth.csv", header = TRUE)
+b <- read.csv("../output/maxent/predictions/may_binary_avg_asclayerbydepth.csv", header = TRUE)
+c <- merge(a, b, by=0)
+
+data <- rbindlist(lapply(list.files("../output/maxent/predictions/", pattern = ".*binary_avg_asclayerbydepth.csv", full.names = TRUE), fread))
+
+```

### 2019-05-16 src/maxent_avec_xvalidation.Rmd
cells used in all months

--- 2019-05-14 src/maxent_avec_xvalidation.Rmd
@@
 
-start with mpa binary map first
-
-```{r}
-
-```
+start with mpa binary map first - based on averages
+
+```{r get all csvs in one database}
+mthlst <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+stdep <- read.csv("../output/maxent/models/dec1998/stack_depth.csv", header = TRUE)
+depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
+names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
+depthlayers$depth <- as.character(depthlayers$depth)
+
+for(m in 1:length(mthlst)){
+  mthna <- mthlst[m]
+#mthna <- "jan"
+  mpabin_lst <- list.files(paste0("../output/maxent/predictions/", mthna), pattern =  "^binary_avg_cloglog.*.asc", full.names = TRUE)
+
+  for (d in 1:nrow(depthlayers)){
+    dep <- depthlayers$depth[d]
+    for (a in 1:length(mpabin_lst)){
+      asc <- mpabin_lst[a]
+      if (grepl(dep, asc)){
+        depthlayers[d , paste0(mthna)] <- asc
+
+      }
+    }
+  }
+  write.csv(depthlayers, ("../output/maxent/predictions/all_cells/binary_avg_asclayerbydepth.csv"), row.names = FALSE)
+}
+
+```
+
+```{r create stacks by depth}
+depthlayers <- read.csv("../output/maxent/predictions/all_cells/binary_avg_asclayerbydepth.csv", header = TRUE)
+
+for (x in 1:nrow(depthlayers)){
+  asclay <- depthlayers[x, ]
+  asclay <- asclay[!is.na(asclay)]
+  asclay <- as.data.frame(rbind(asclay))
+  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
+  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to month
+  asclay$V3 <- as.character(asclay$V3)
+  asclay$V4 <- as.character(asclay$V4)
+  asclay$V5 <- as.character(asclay$V5)
+  asclay$V6 <- as.character(asclay$V6)
+  asclay$V7 <- as.character(asclay$V7)
+  asclay$V8 <- as.character(asclay$V8)
+  asclay$V9 <- as.character(asclay$V9)
+  asclay$V10 <- as.character(asclay$V10)
+  asclay$V11 <- as.character(asclay$V11)
+  asclay$V12 <- as.character(asclay$V12)
+
+  asclaydep <- asclay$V1
+
+  ascjan <- raster(asclay$V2)
+  ascmar <- raster(asclay$V3)
+  ascapr <- raster(asclay$V4)
+  ascmay <- raster(asclay$V5)
+  ascjun <- raster(asclay$V6)
+  ascjul <- raster(asclay$V7)
+  ascaug <- raster(asclay$V8)
+  ascsep <- raster(asclay$V9)
+  ascoct <- raster(asclay$V10)
+  ascnov <- raster(asclay$V11)
+  ascdec <- raster(asclay$V12)
+
+  stkasc <- stack(ascjan, ascmar, ascapr, ascmay, ascjun, ascjul, ascaug, ascsep, ascoct, ascnov, ascdec)
+
+  avg <- sum(stkasc, na.rm = TRUE)
+
+  writeRaster(avg, paste0("../output/maxent/predictions/all_cells/count_cells_used_all_months_mpa", asclaydep, ".asc"), overwrite=TRUE)
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/all_cells/count_cells_used_all_months_mpa", asclaydep, ".png")), overwrite = TRUE) 
+  dev.off()
+}
+
+```
+
+```{r create rasters showing cells that are suitable in all 11 months}
+cellslst <- list.files("../output/maxent/predictions/all_cells", pattern =  "^count_cells_used_all_months_mpa.*.asc", full.names = TRUE)
+
+allcell <- function(x){
+  ifelse(x == 12, 1,
+  ifelse(x < 12, 0, NA)) 
+}
+
+for (r in 1:length(cellslst)){
+  ras <- raster(cellslst[r])
+  ascname <- names(ras)
+  allc <- calc(ras, fun = allcell)
+  writeRaster(allc, paste0("../output/maxent/predictions/all_cells/total_cells_used_all_months_mpa", ascname, ".asc"), overwrite = TRUE)
+  plot(allc, png(paste0("../output/maxent/predictions/all_cells/total_cells_used_all_months_mpa", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+
+}
+```
+
+there are no cells that are used in every month....

### 2019-05-16 src/maxent_avec_xvalidation.Rmd
plot generation for report

@@
 options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr")
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2")
 source("../src/ipak.R")
@@
 
+mthord <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
+aicsum <- aicsum[match(mthord, aicsum$Month), ] #reorder based on mthord
 write.csv(aicsum, "../output/maxent/models/aicsummary.csv", row.names = FALSE)
+
 ```
@@
 png("../output/maxent/models/permutation_plot.png")
+print(perpl)
+dev.off()
+
+#for jan
+
+permuj <- subset(permu, Month == "jan")
+permelt <- melt(permuj, id = "Month") #need to melt the data into this format for ggplot
+perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") + scale_fill_brewer(palette = "Set1")
+png("../output/maxent/models/permutation_plot_jan.png")
 print(perpl)
@@
 
-#plots by depth
+#plots by depth for MPA threshold
 
@@
 there are no cells that are used in every month....
+
+# quickly adding some black to some plots for mj slr
+```{r}
+mthlst <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for(m in 1:length(mthlst)){
+  mthna <- mthlst[m]
+  replst11 <- list.files(paste0("../output/maxent/predictions/", mthna), pattern =  ".*11.7737.*.asc", full.names = TRUE)
+  replst221 <- list.files(paste0("../output/maxent/predictions/", mthna), pattern =  ".*221.141.*.asc", full.names = TRUE)
+  replst508 <- list.files(paste0("../output/maxent/predictions/", mthna), pattern =  ".*508.64.*.asc", full.names = TRUE)
+  replst <- c(replst11, replst221, replst508)
+  for (r in 1:length(replst)){
+    ras <- raster(replst[r])
+    ascname <- names(ras)
+    plot(ras, colNA = 'black', png(paste0("../output/maxent/predictions/", mthna, "/", ascname, ".png")), overwrite = TRUE)
+    dev.off()
+  }
+}
+
+replst163 <- list.files("../output/maxent/predictions/jul/", pattern =  ".*163..*.asc", full.names = TRUE)
+for (r in 1:length(replst163)){
+    ras <- raster(replst163[r])
+    ascname <- names(ras)
+    plot(ras, colNA = 'black', png(paste0("../output/maxent/predictions/jul/", ascname, ".png")), overwrite = TRUE)
+    dev.off()
+}
+
+back <- raster("../output/maxent/predictions/mar/binary_avg_medprob_avg_cloglog_1.55586.asc")
+replst163 <- list.files("../output/maxent/predictions/jan/", pattern =  ".*avg_2d.*.asc", full.names = TRUE)
+for (r in 1:length(replst163)){
+    ras <- raster(replst163[r])
+    ascname <- names(ras)
+    bkras <- overlay(back, ras, fun=function(x,y){return(x+y)})
+    plot(bkras, colNA = 'black', png(paste0("../output/maxent/predictions/jan/", ascname, ".png")), overwrite = TRUE)
+    dev.off()
+}
+
+```

### 2019-05-22 src/asc_file_database.Rmd
minor adjustment

@@
 ```{r pre-install packages, message=FALSE}
-packages <- c("tidyr", "rJava", "dismo")
+packages <- c("tidyr")
 source("../src/ipak.R")
@@
 ascmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
-ascdepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38")
+ascdepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85")
 #this depth list is based on how the files are named when split into depth layers. note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers

### 2019-05-22 src/maxent_avec_xvalidation.Rmd
minor adjustment

@@
 ```{r month of interest -change}
-mthna <- "oct"
-mthno <- "10"
+mthna <- "jan"
+mthno <- "1"
 ```
@@
 ```{r binary map per year 2D representation}
+
 binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_cloglog.*.asc", full.names = TRUE)
@@
 ```{r binary map averaged 2D representation}
+
 binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_cloglog.*.asc", full.names = TRUE)
@@
 stackstat <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-stackstat[ ,1] <- c(1:49)
+stackstat[ ,1] <- c(1:46)
 colnames(stackstat) <- c("layer", "cell_sum")
@@
 write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/cellcount_bydepth_avg.csv"), row.names = FALSE)
+```
+
+attach depth in meters (layer_name in depth_bin.csv) to the stackstat based on layer (stackstat) and depthlayerno (depth_bin.csv)
+
+```{r}
+dep <- read.csv("../data/env/depth_bin.csv", header = TRUE) # the depth bin file created way back
+dep <- subset(dep, select = c("depthlayerno", "layer_name"))
+stackstat <-  read.csv(paste0("../output/maxent/predictions/", mthna, "/cellcount_bydepth_avg.csv"), header = TRUE)
+colnames(dep)[names(dep) == "depthlayerno"] <- "layer"
+stackstat <- left_join(stackstat, dep)
+stackstat[ , "layer_name"]=round(stackstat[ , "layer_name"], 2)
+head(stackstat)
+
 ```
@@
 dev.off()
+
 ```
@@
 ```{r high probability map per year 2D representation}
+
 binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_highprob_cloglog.*.asc", full.names = TRUE)
@@
 ```{r high probability average map per year 2D representation}
+
 binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_highprob.*.asc", full.names = TRUE)
@@
 dev.off()
-```
+
+```
+
 ```{r med probability average map per year 2D representation}
+
 binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_medprob.*.asc", full.names = TRUE)
@@
 ```{r low probability average map per year 2D representation}
+
 binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_lowprob.*.asc", full.names = TRUE)
@@
 stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-stackstathigh[ ,1] <- c(1:49)
+stackstathigh[ ,1] <- c(1:46)
 colnames(stackstathigh) <- c("layer", "high_cell_sum")
@@
 stackstatmed <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-stackstatmed[ ,1] <- c(1:49)
+stackstatmed[ ,1] <- c(1:46)
 colnames(stackstatmed) <- c("layer", "med_cell_sum")
@@
 stackstatlow <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-stackstatlow[ ,1] <- c(1:49)
+stackstatlow[ ,1] <- c(1:46)
 colnames(stackstatlow) <- c("layer", "low_cell_sum")
@@
 write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/probability_cellcount_bydepth_avg.csv"), row.names = FALSE)
+
 ```

### 2019-06-06 src/AO_NAO_comparison.Rmd
quick nao ao upper quartile year comparisons

# NAO

Dataset mes from [NCAR](https://climatedataguide.ucar.edu/climate-data/hurrell-north-atlantic-oscillation-nao-index-pc-based). Note I am using the monthly principal component (PC)-based indices of the North Atlantic Oscillation [(NAO)](https://climatedataguide.ucar.edu/sites/default/files/nao_pc_monthly.txt). I have already converted this file into a csv and removed unncesessary text. Note there are no headers...

```{r}
naoraw <- read.csv("./data/env/nao_pc_monthly.csv", header = FALSE) 
head(naoraw)
```

lets give some column names. col 1  year, 2:13 are the months
```{r}
colnames(naoraw) <- c("year","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
head(naoraw)
```

 NA values are represented by -999. Change these to NA
```{r}
naoraw <- replace(naoraw,naoraw==-999, NA)
```

Now subset the dataset to start in 1996 (the capelin data starts in 1998, but you want some previous time period stuff)

```{r}
nao <- naoraw[naoraw$year >= 1996, ]
head(nao) 
```

And just for fun, plot the NAO follow the tutorial here

```{r}
naocolumn <- matrix(t(nao[ ,2:13]), ncol=1) #month data is in col 2 - 13 
naocolumn1 <- as.vector(naocolumn)
naotimeseries <- ts(naocolumn1, frequency = 12, start = c(1996,1))
plot.ts(naotimeseries)
abline(h = 0)
#dev.copy(png, "../output/env/nao.png") #this prints a png of the plot
#dev.off() #this turns off the print command
```

#AO

```{r}
aoraw <- read.csv("../data/env/arctic_osc_mon.csv", header = TRUE) 
head(aoraw)
```

 NA values are represented by -999. Change these to NA
```{r}
aoraw <- replace(aoraw,aoraw==-999, NA)
```

Now subset the dataset to start in 1996 (the capelin data starts in 1998, but you want some previous time period stuff)

```{r}
ao <- aoraw[aoraw$X >= 1996, ]
head(ao) 
```

And just for fun, plot the NAO follow the tutorial here

```{r}
aocolumn <- matrix(t(ao[ ,2:13]), ncol=1) #month data is in col 2 - 13 
aocolumn1 <- as.vector(aocolumn)
aotimeseries <- ts(aocolumn1, frequency = 12, start = c(1996,1))
plot.ts(aotimeseries)
abline(h = 0)
dev.copy(png, "../output/env/ao.png") #this prints a png of the plot
dev.off() #this turns off the print command
```

```{r}
summary(nao)
```

need dplyr
```{r}
naoq1 <- summarize_all(nao, ~ quantile(.x, probs = 0.25, na.rm = TRUE))
write.csv(naoq1, "../output/env/naoq1.csv", row.names = FALSE)
naoq1 <- read.csv("../output/env/naoq1.csv", header = TRUE)

naojanq1 <- subset(nao$year, nao$Jan <= naoq1$Jan)
write.csv(naojanq1, "../output/env/nao_jan_q1.csv")        
naofebq1 <- subset(nao$year, nao$Feb <= naoq1$Feb)
write.csv(naofebq1, "../output/env/nao_feb_q1.csv") 
naomarq1 <- subset(nao$year, nao$Mar <= naoq1$Mar)
write.csv(naomarq1, "../output/env/nao_mar_q1.csv")        
naoaprq1 <- subset(nao$year, nao$Apr <= naoq1$Apr)
write.csv(naoaprq1, "../output/env/nao_apr_q1.csv")  
naomayq1 <- subset(nao$year, nao$May <= naoq1$May)
write.csv(naomayq1, "../output/env/nao_may_q1.csv")        
naojunq1 <- subset(nao$year, nao$Jun <= naoq1$Jun)
write.csv(naojunq1, "../output/env/nao_jun_q1.csv") 
naojulq1 <- subset(nao$year, nao$Jul <= naoq1$Jul)
write.csv(naojulq1, "../output/env/nao_jul_q1.csv")        
naoaugq1 <- subset(nao$year, nao$Aug <= naoq1$Aug)
write.csv(naoaugq1, "../output/env/nao_aug_q1.csv")  
naosepq1 <- subset(nao$year, nao$Sep <= naoq1$Sep)
write.csv(naosepq1, "../output/env/nao_sep_q1.csv")        
naooctq1 <- subset(nao$year, nao$Oct <= naoq1$Oct)
write.csv(naooctq1, "../output/env/nao_oct_q1.csv") 
naonovq1 <- subset(nao$year, nao$Nov <= naoq1$Nov)
write.csv(naonovq1, "../output/env/nao_nov_q1.csv")        
naodecq1 <- subset(nao$year, nao$Dec <= naoq1$Dec)
write.csv(naodecq1, "../output/env/nao_dec_q1.csv")

naoq1_all <- rbind(naojanq1, naofebq1, naomarq1, naoaprq1, naomayq1, naojunq1, naojulq1, naoaugq1, naosepq1, naooctq1, naonovq1, naodecq1)
write.csv(naoq1_all, "../output/env/nao_q1_all.csv")

naoq4 <- summarize_all(nao, ~ quantile(.x, probs = 0.75, na.rm = TRUE))
write.csv(naoq4, "../output/env/naoq4.csv", row.names = FALSE)
naoq4 <- read.csv("../output/env/naoq4.csv", header = TRUE)

naojanq4 <- subset(nao$year, nao$Jan >= naoq4$Jan)
write.csv(naojanq4, "../output/env/nao_jan_q4.csv")        
naofebq4 <- subset(nao$year, nao$Feb >= naoq4$Feb)
write.csv(naofebq4, "../output/env/nao_feb_q4.csv") 
naomarq4 <- subset(nao$year, nao$Mar >= naoq4$Mar)
write.csv(naomarq4, "../output/env/nao_mar_q4.csv")        
naoaprq4 <- subset(nao$year, nao$Apr >= naoq4$Apr)
write.csv(naoaprq4, "../output/env/nao_apr_q4.csv")  
naomayq4 <- subset(nao$year, nao$May >= naoq4$May)
write.csv(naomayq4, "../output/env/nao_may_q4.csv")        
naojunq4 <- subset(nao$year, nao$Jun >= naoq4$Jun)
write.csv(naojunq4, "../output/env/nao_jun_q4.csv") 
naojulq4 <- subset(nao$year, nao$Jul >= naoq4$Jul)
write.csv(naojulq4, "../output/env/nao_jul_q4.csv")        
naoaugq4 <- subset(nao$year, nao$Aug >= naoq4$Aug)
write.csv(naoaugq4, "../output/env/nao_aug_q4.csv")  
naosepq4 <- subset(nao$year, nao$Sep >= naoq4$Sep)
write.csv(naosepq4, "../output/env/nao_sep_q4.csv")        
naooctq4 <- subset(nao$year, nao$Oct >= naoq4$Oct)
write.csv(naooctq4, "../output/env/nao_oct_q4.csv") 
naonovq4 <- subset(nao$year, nao$Nov >= naoq4$Nov)
write.csv(naonovq4, "../output/env/nao_nov_q4.csv")        
naodecq4 <- subset(nao$year, nao$Dec >= naoq4$Dec)
write.csv(naodecq4, "../output/env/nao_dec_q4.csv")

naoq4_all <- rbind(naojanq4, naofebq4, naomarq4, naoaprq4, naomayq4, naojunq4, naojulq4, naoaugq4, naosepq4, naooctq4, naonovq4, naodecq4)
write.csv(naoq4_all, "../output/env/nao_q4_all.csv")

```

need dplyr
```{r}
aoq1 <- summarize_all(ao, ~ quantile(.x, probs = 0.25, na.rm = TRUE))
write.csv(aoq1, "../output/env/aoq1.csv", row.names = FALSE)
aoq1 <- read.csv("../output/env/aoq1.csv", header = TRUE)

aojanq1 <- subset(ao$X, ao$Jan <= aoq1$Jan)
write.csv(aojanq1, "../output/env/ao_jan_q1.csv")        
aofebq1 <- subset(ao$X, ao$Feb <= aoq1$Feb)
write.csv(aofebq1, "../output/env/ao_feb_q1.csv") 
aomarq1 <- subset(ao$X, ao$Mar <= aoq1$Mar)
write.csv(aomarq1, "../output/env/ao_mar_q1.csv")        
aoaprq1 <- subset(ao$X, ao$Apr <= aoq1$Apr)
write.csv(aoaprq1, "../output/env/ao_apr_q1.csv")  
aomayq1 <- subset(ao$X, ao$May <= aoq1$May)
write.csv(aomayq1, "../output/env/ao_may_q1.csv")        
aojunq1 <- subset(ao$X, ao$Jun <= aoq1$Jun)
write.csv(aojunq1, "../output/env/ao_jun_q1.csv") 
aojulq1 <- subset(ao$X, ao$Jul <= aoq1$Jul)
write.csv(aojulq1, "../output/env/ao_jul_q1.csv")        
aoaugq1 <- subset(ao$X, ao$Aug <= aoq1$Aug)
write.csv(aoaugq1, "../output/env/ao_aug_q1.csv")  
aosepq1 <- subset(ao$X, ao$Sep <= aoq1$Sep)
write.csv(aosepq1, "../output/env/ao_sep_q1.csv")        
aooctq1 <- subset(ao$X, ao$Oct <= aoq1$Oct)
write.csv(aooctq1, "../output/env/ao_oct_q1.csv") 
aonovq1 <- subset(ao$X, ao$Nov <= aoq1$Nov)
write.csv(aonovq1, "../output/env/ao_nov_q1.csv")        
aodecq1 <- subset(ao$X, ao$Dec <= aoq1$Dec)
write.csv(aodecq1, "../output/env/ao_dec_q1.csv")

aoq1_all <- rbind(aojanq1, aofebq1, aomarq1, aoaprq1, aomayq1, aojunq1, aojulq1, aoaugq1, aosepq1, aooctq1, aonovq1, aodecq1)
write.csv(aoq1_all, "../output/env/ao_q1_all.csv")

aoq4 <- summarize_all(ao, ~ quantile(.x, probs = 0.75, na.rm = TRUE))
write.csv(aoq4, "../output/env/aoq4.csv", row.names = FALSE)
aoq4 <- read.csv("../output/env/aoq4.csv", header = TRUE)

aojanq4 <- subset(ao$X, ao$Jan >= aoq4$Jan)
write.csv(aojanq4, "../output/env/ao_jan_q4.csv")        
aofebq4 <- subset(ao$X, ao$Feb >= aoq4$Feb)
write.csv(aofebq4, "../output/env/ao_feb_q4.csv") 
aomarq4 <- subset(ao$X, ao$Mar >= aoq4$Mar)
write.csv(aomarq4, "../output/env/ao_mar_q4.csv")        
aoaprq4 <- subset(ao$X, ao$Apr >= aoq4$Apr)
write.csv(aoaprq4, "../output/env/ao_apr_q4.csv")  
aomayq4 <- subset(ao$X, ao$May >= aoq4$May)
write.csv(aomayq4, "../output/env/ao_may_q4.csv")        
aojunq4 <- subset(ao$X, ao$Jun >= aoq4$Jun)
write.csv(aojunq4, "../output/env/ao_jun_q4.csv") 
aojulq4 <- subset(ao$X, ao$Jul >= aoq4$Jul)
write.csv(aojulq4, "../output/env/ao_jul_q4.csv")        
aoaugq4 <- subset(ao$X, ao$Aug >= aoq4$Aug)
write.csv(aoaugq4, "../output/env/ao_aug_q4.csv")  
aosepq4 <- subset(ao$X, ao$Sep >= aoq4$Sep)
write.csv(aosepq4, "../output/env/ao_sep_q4.csv")        
aooctq4 <- subset(ao$X, ao$Oct >= aoq4$Oct)
write.csv(aooctq4, "../output/env/ao_oct_q4.csv") 
aonovq4 <- subset(ao$X, ao$Nov >= aoq4$Nov)
write.csv(aonovq4, "../output/env/ao_nov_q4.csv")        
aodecq4 <- subset(ao$X, ao$Dec >= aoq4$Dec)
write.csv(aodecq4, "../output/env/ao_dec_q4.csv")

aoq4_all <- rbind(aojanq4, aofebq4, aomarq4, aoaprq4, aomayq4, aojunq4, aojulq4, aoaugq4, aosepq4, aooctq4, aonovq4, aodecq4)
write.csv(aoq4_all, "../output/env/ao_q4_all.csv")
```

so there are differences in upper quartile years...

### 2019-06-06 src/stack_test.Rmd
messing with land masses

@@
 ```
+
+------------------------
+
+# trim layer by polygon
+
+C:\Users\bunny\Documents\GitHub\Chapter_1\output\maxent\predictions\jun\avg_cloglog_147.406.asc
+C:\Users\bunny\Documents\GitHub\Chapter_1\output\bio\minimum_convex_poly\mcp_capelin_100_poly.shp
+C:\Users\bunny\Documents\GitHub\Chapter_1\output\env\aea.tif
+
+```{r}
+testras <- raster("../output/maxent/predictions/jun/avg_cloglog_147.406.asc")
+mcp <- readOGR("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp") 
+aea <- raster("../output/env/aea.tif")
+plot(testras)
+plot(mcp)
+class(mcp)
+rr <- mask(testras, mcp)
+plot(rr); plot(mcp, add = TRUE)
+plot(testras); plot(mcp, add = TRUE)
+plot(rr);plot(testras)
+```
+
+# add landmass
+
+```{r}
+us <- getData("GADM", country="USA", level=0)
+plot(us)
+class(us)
+projection(testras) <- CRS(aea)
+
+us <- spTransform(us, CRS(projection(aea)))
+
+crs(testras) <- aea
+
+rr <- mask(aea, us)
+plot(us)
+
+a <- plot(testras); plot(us, add = TRUE)
+
+```

### 2019-07-04 src/maxent_alldata_avec_xvalidation.Rmd
fold-aic-models

@@
 ```{r remove unnecessary columns}
-presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt, month, depthlayerno, depth_layer))
+prback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt, depthlayerno, depth_layer))
 colnames(presback)
@@
 ```{r}
-allvar <- c("occurrence", "year", "temp_surface", "temp_depth", "salinity_surface", "salinity_depth", "o2_depth", "chl_surface", "mlp_surface",  "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-presback <- subset(presback, select = c(allvar))
-colnames(presback)
+allvar <- c("occurrence", "year", "month", "temp_surface", "temp_depth", "salinity_surface", "salinity_depth", "o2_depth", "chl_surface", "mlp_surface",  "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+prback <- subset(prback, select = c(allvar))
+colnames(prback)
 ```
@@
 ```{r remove na vals}
-prback <- na.omit(presback)
+prback <- na.omit(prback)
+
+prback <- subset(prback, month == 8)
+
+presyrs <-  subset(prback, occurrence == 1) #this is to get years with complete presence data 
+presyrs <- unique(presyrs$year) #make a vector of unique years
+prback <- prback[prback$year %in% presyrs, ] #keep only the data for which there is precence years
 ```
@@
 aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
-betavallist <- cbind(aicetc$regularization)
+#betavallist <- cbind(aicetc$regularization)
+betavallist <- seq(0.25, 0.75, by = 0.25)
+
 timeslice <- "all_data"
@@
 train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-nfolds <- cbind(c(1:5))
+nfolds <- cbind(c(1:2))
 
-  for (h in 1:nrow(betavallist)){
+ for (h in 1:length(betavallist)){
    betamval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)))
+    #betamval <-  0.5
+    mod <- maxent(x = train, p = trainpa, plots = FALSE, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)))
 
@@
 
+          #no presences used in model (less those used for testing)
+
+    maxresults <- read.csv(paste0("../output/maxent/", timeslice, "/models/maxentResults.csv"), header = TRUE)
+
+for (i in 1:nrow(aicetc)){   
 for(f in 1:nrow(nfolds)){
@@
 
-      #no presences used in model (less those used for testing)
-
       notrainingpoints <- maxresults$X.Training.samples[f]
@@
 
-      aicetc$train_AUC <- trainauc
-      aicetc$test_AUC <- testauc
-      aicetc$n <- notrainingpoints
-      aicetc$coefficients <- nocoeff
-      aicetc$sum.of.logs <- sumsamplepred
+   if(aicetc$nfold[i] == modn){
+      aicetc$Train_AUC[i] <- trainauc
+      aicetc$Test_AUC[i] <- testauc
+      aicetc$n[i] <- notrainingpoints
+      aicetc$coefficients[i] <- nocoeff
+      aicetc$sum.of.logs[i] <- sumsamplepred
 
-      aicetc$bic <- bic #not sure this is correct...
-      aicetc$beta <- betamval
-      aicetc$foldno <- modn
-
-    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
-    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+      aicetc$bic[i] <- bic #not sure this is correct...
+      aicetc$beta[i] <- betamval
+      aicetc$foldno[i] <- modn
 
     #calculate minimal predicted area using ecospat
-    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
+    aicetc$mpa_0.9_cum[i] <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
 
-    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
-    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
-    write.csv(aicetc, paste0("../output/maxent/", timeslice, "/models/", betamval, "_", modn, "_", "AICforMaxentReg.csv"), row.names = FALSE)
+    aicetc$AIC[i] <- ((2*aicetc$coefficients[i])-(2*aicetc$sum.of.logs[i]))
+    aicetc$AICc[i] <- (-2*(aicetc$sum.of.logs[i])+((2*aicetc$coefficients[i])*(aicetc$n[i]/(aicetc$n[i] - aicetc$coefficients[i]-1))))
    }
 
-  # aa <- paste0("fold", modn, "_", betamval, "_", "AICforMaxentReg.csv")
-    aicperbetalst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = paste0("^", betamval, ".*AICforMaxentReg.csv"), full.names = TRUE)
-    aicperbeta <- lapply(aicperbetalst, read.csv, header = TRUE)
-    aicperbeta <- do.call(rbind, aicperbeta)
-    aicbetamean <- colMeans(aicperbeta)
-    aicperbeta <- rbind(aicperbeta, aicbetamean)
-    aicperbeta$foldno[f+1] <- f+1
-    write.csv(aicperbeta, paste0("../output/maxent/", timeslice, "/models/aicperbetasum_", betamval, ".csv"), row.names = FALSE)
+  }
+    write.csv(aicetc, paste0("../output/maxent/", timeslice, "/models/", "_", betamval, "_AICforMaxentReg.csv"), row.names = FALSE)
 
-    betaavglst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = "aicperbetasum_.*csv", full.names = TRUE)
-    bavg <- lapply(betaavglst, read.csv, header = TRUE)
+}
+  }
+
+    aicperbetalst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = paste0(".*AICforMaxentReg.csv"), full.names = TRUE)
+    for (beta in 1:length(aicperbetalst)){
+      aic <- read.csv(aicperbetalst[beta])
+      bval <- str_sub(aicperbetalst[beta], 35, -20) #the year being used for testing
+      aicmean <- colwise(mean)(aic)
+      write.csv(aicmean, paste0("../output/maxent/", timeslice, "/models/aic_mean_", bval, ".csv"), row.names = FALSE)
+    }
+
+    aicmeanlst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = paste0("^aic_mean"), full.names = TRUE)
+    bavg <- lapply(aicmeanlst, read.csv, header = TRUE)
     bavg <- do.call(rbind, bavg)
-    bavg <- subset(bavg, foldno == f+1)
     write.csv(bavg, paste0("../output/maxent/", timeslice, "/models/avgbetalst.csv"), row.names = FALSE)
-  }
 

### 2019-07-04 src/testing_file.Rmd
fold-aic-models

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}
presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
presences <- subset(presback, occurrence == "1") #splits the testing dataset into presences
un_mth <- unique(presences$month)
prrows <- length(presences)
presences$season <- 5

for (i in 1:nrow(presences)){
  if (presences$month[i] >= 1 & presences$month[i] <= 3){
    presences$season[i] <- 1
    } else if (presences$month[i] >= 4 & presences$month[i] <= 6){
    presences$season[i] <- 2
    } else if (presences$month[i] >= 7 & presences$month[i] <= 9){
    presences$season[i] <- 3
    } else if (presences$month[i] >= 10 & presences$month[i] <= 12){
    presences$season[i] <- 4
  }
}

write.csv(presences, "../output/bio/cluster_seasons_test.csv", row.names = FALSE)

names(presences)

pres <- subset(presences, select = c(year, month, temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter, season))
pres <- na.omit(pres)
pres2 <- na.omit(presences)
pres <- scale(pres)
```

```{r}

library("cluster")
library("factoextra")
library("magrittr")
```

```{r}
distance <- get_dist(pres)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))
```

```{r}
fviz_nbclust(pres, kmeans, method = "gap_stat")
```

```{r}
set.seed(123)
km.res <- kmeans(pres, 6, nstart = 25)
km.res
# Visualize
library("factoextra")
fviz_cluster(km.res, data = pres,
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal())
```

```{r}
library("cluster")
pam.res <- pam(pres, 6)
# Visualize
fviz_cluster(pam.res)
```

```{r}
library(rworldmap)
pres2wi <- subset(pres2, season == "1")
pres2sp <- subset(pres2, season == "2")
pres2su <- subset(pres2, season == "3")
pres2au <- subset(pres2, season == "4")
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "pres2wi", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(pres2wi$decimalLongitude, pres2wi$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
points(pres2sp$decimalLongitude, pres2sp$decimalLatitude, col = "blue") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
points(pres2su$decimalLongitude, pres2su$decimalLatitude, col = "green") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
points(pres2au$decimalLongitude, pres2au$decimalLatitude, col = "yellow") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```

```{r}
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "winter", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(pres2wi$decimalLongitude, pres2wi$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```

```{r}
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "spring", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(pres2sp$decimalLongitude, pres2sp$decimalLatitude, col = "blue") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")

```

```{r}
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "summer", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(pres2su$decimalLongitude, pres2su$decimalLatitude, col = "green") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```

```{r}
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "autumn", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(pres2au$decimalLongitude, pres2au$decimalLatitude, col = "yellow") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
```

```{r}
p98 <- subset(pres2, year == "1998")
p99 <- subset(pres2, year == "1999")
p00 <- subset(pres2, year == "2000")
p01 <- subset(pres2, year == "2001")
p02 <- subset(pres2, year == "2002")
p03 <- subset(pres2, year == "2003")
p04 <- subset(pres2, year == "2004")
p05 <- subset(pres2, year == "2005")
p06 <- subset(pres2, year == "2006")
p07 <- subset(pres2, year == "2007")
p08 <- subset(pres2, year == "2009")
p09 <- subset(pres2, year == "2009")
p10 <- subset(pres2, year == "2010")
p11 <- subset(pres2, year == "2011")
p12 <- subset(pres2, year == "2012")
p13 <- subset(pres2, year == "2013")
p14 <- subset(pres2, year == "2014")
p15 <- subset(pres2, year == "2015")

```

```{r}
map2 <- getMap(resolution = "low") #creates an object called map at low resoultion
plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "1998", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p98$decimalLongitude, p98$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_1998.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "1999", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p99$decimalLongitude, p99$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_1999.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2000", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p00$decimalLongitude, p00$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2000.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2001", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p01$decimalLongitude, p01$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2001.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2002", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p02$decimalLongitude, p02$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2002.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2003", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p03$decimalLongitude, p03$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2003.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2004", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p04$decimalLongitude, p04$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2004.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2005", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p05$decimalLongitude, p05$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2005.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2006", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p06$decimalLongitude, p06$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2006.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2007", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p07$decimalLongitude, p07$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2007.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2008", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p08$decimalLongitude, p08$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2008.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2009", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p09$decimalLongitude, p09$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2009.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2010", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p10$decimalLongitude, p10$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2010.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2011", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p11$decimalLongitude, p11$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2011.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2012", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p12$decimalLongitude, p12$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2012.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2013", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p13$decimalLongitude, p13$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2013.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2014", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p14$decimalLongitude, p14$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2014.png") #this prints a png of the plot
dev.off() #this turns off the print command

plot(map2, xlim = c(-70, -43), ylim =c(38, 70), asp = 1, main = "2015", col = "cornsilk") #the x and y lim are the long-lat bounds of the map
points(p15$decimalLongitude, p15$decimalLatitude, col = "red") #this adds points to the mapet", xlab = "Longitude", ylab = "Latitude")
dev.copy(png, "../data/bio/occurrence_map/p_2015.png") #this prints a png of the plot
dev.off() #this turns off the print command
```

```{r}
df <- iris
df <- na.omit(df)
df <- scale(df[ ,-5])

# Dissimilarity matrix
d <- dist(df, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
```

```{r}
iris2 <- pres[ ,-12]
species_labels <- pres[ ,12]
library(colorspace) # get nice colors
species_col <- rev(rainbow_hcl(4))[as.numeric(species_labels)]

# Plot a SPLOM:
pairs(iris2, col = species_col,
      lower.panel = NULL,
       cex.labels=2, pch=19, cex = 1.2)

# Add a legend
par(xpd = TRUE)
legend(x = 0.05, y = 0.4, cex = 2,
   legend = as.character(levels(species_labels)),
    fill = unique(species_col))
par(xpd = NA)
```

```{r}
par(las = 1, mar = c(4.5, 3, 3, 2) + 0.1, cex = .8)
MASS::parcoord(iris2, col = species_col, var.label = TRUE, lwd = 2)

# Add Title
title("Parallel coordinates plot of the Iris data")
# Add a legend
par(xpd = TRUE)
legend(x = 1.75, y = -.25, cex = 1,
   legend = as.character(levels(species_labels)),
    fill = unique(species_col), horiz = TRUE)
```

```{r}
d_iris <- dist(iris2) # method="man" # is a bit better
hc_iris <- hclust(d_iris, method = "complete")
iris_species <- rev(levels(pres[,12]))

library(dendextend)
dend <- as.dendrogram(hc_iris)
# order it the closest we can to the order of the observations:
dend <- rotate(dend, 1:6361)

# Color the branches based on the clusters:
dend <- color_branches(dend, k=4) #, groupLabels=iris_species)

# Manually match the labels, as much as possible, to the real classification of the flowers:
labels_colors(dend) <-
   rainbow_hcl(3)[sort_levels_values(
      as.numeric(pres[,12])[order.dendrogram(dend)]
   )]

# We shall add the flower type to the labels:
labels(dend) <- paste(as.character(iris[,5])[order.dendrogram(dend)],
                           "(",labels(dend),")", 
                           sep = "")
# We hang the dendrogram a bit:
dend <- hang.dendrogram(dend,hang_height=0.1)
# reduce the size of the labels:
# dend <- assign_values_to_leaves_nodePar(dend, 0.5, "lab.cex")
dend <- set(dend, "labels_cex", 0.5)
# And plot:
par(mar = c(3,3,3,7))
plot(dend, 
     main = "Clustered Iris data set
     (the labels give the true flower species)", 
     horiz =  TRUE,  nodePar = list(cex = .007))
legend("topleft", legend = iris_species, fill = rainbow_hcl(3))
```

```{r}
library(circlize)
par(mar = rep(0,4))
circlize_dendrogram(dend)
```

```{r}
iris3 <- pres[ ,-2]
species_labels <- pres[ ,2]
library(colorspace) # get nice colors
species_col <- rev(rainbow_hcl(12))[as.numeric(species_labels)]

# Plot a SPLOM:
pairs(iris3, col = species_col,
      lower.panel = NULL,
       cex.labels=2, pch=19, cex = 1.2)

# Add a legend
par(xpd = TRUE)
legend(x = 0.05, y = 0.4, cex = 2,
   legend = as.character(levels(species_labels)),
    fill = unique(species_col))
par(xpd = NA)
```

```{r}
d_iris2 <- dist(pres) # method="man" # is a bit better
hc_iris2 <- hclust(d_iris2, method = "complete")
iris_species2 <- rev(levels(pres[,12]))

library(dendextend)
dend2 <- as.dendrogram(hc_iris2)
# order it the closest we can to the order of the observations:
dend2 <- rotate(dend2, 1:6361)

# Color the branches based on the clusters:
dend2 <- color_branches(dend2, k=4) #, groupLabels=iris_species)

# Manually match the labels, as much as possible, to the real classification of the flowers:
labels_colors(dend2) <-
   rainbow_hcl(3)[sort_levels_values(
      as.numeric(pres[,12])[order.dendrogram(dend2)]
   )]

# We shall add the flower type to the labels:
labels(dend2) <- paste(as.character(iris[,5])[order.dendrogram(dend2)],
                           "(",labels(dend2),")", 
                           sep = "")
# We hang the dendrogram a bit:
dend2 <- hang.dendrogram(dend2,hang_height=0.1)
# reduce the size of the labels:
# dend <- assign_values_to_leaves_nodePar(dend, 0.5, "lab.cex")
dend22 <- set(dend2, "labels_cex", 0.5)
# And plot:
par(mar = c(3,3,3,7))
plot(dend2, 
     main = "Clustered Iris data set
     (the labels give the true flower species)", 
     horiz =  TRUE,  nodePar = list(cex = .007))
legend("topleft", legend = species_labels, fill = rainbow_hcl(3))
```

```{r}
library(circlize)
par(mar = rep(0,4))
circlize_dendrogram(dend2)
```

### 2019-07-15 src/maxent_alldata_avec_xvalidation.Rmd
fecky feck violin plots

@@
 ```{r pre-install & load packages, include=FALSE}
-options (java.parameters = "-Xmx2048m")
-packages <- c("rJava", "dismo", "stringr", "raster", "ecospat")
+gc() 
+options (java.parameters = "-Xmx6g")
+packages <- c("rJava", "dismo", "stringr", "raster", "ecospat", "plyr", "data.table", "RColorBrewer")
 source("../src/ipak.R")
@@
 ```{r}
-allvar <- c("occurrence", "year", "month", "temp_surface", "temp_depth", "salinity_surface", "salinity_depth", "o2_depth", "chl_surface", "mlp_surface",  "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+allvar <- c("occurrence", "year", "month", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
 prback <- subset(prback, select = c(allvar))
@@
 prback <- na.omit(prback)
-
-prback <- subset(prback, month == 8)
-
+#prback <- subset(prback, month == 8)
 presyrs <-  subset(prback, occurrence == 1) #this is to get years with complete presence data 
@@
 aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
-#betavallist <- cbind(aicetc$regularization)
-betavallist <- seq(0.25, 0.75, by = 0.25)
+betavallist <- seq(1.75, 4, by = 0.25)
 
@@
 trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-nfolds <- cbind(c(1:2))
+train  <- subset(train, select = -c(year, occurrence, month)) #don't need year and occurrence for modelling
+nfolds <- cbind(c(1:5))
 
@@
     #betamval <-  0.5
-    mod <- maxent(x = train, p = trainpa, plots = FALSE, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)))
+    mod <- maxent(x = train, p = trainpa, plots = FALSE, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=FALSE", "jackknife=FALSE", "plots=FALSE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)))
 
@@
       aic <- read.csv(aicperbetalst[beta])
-      bval <- str_sub(aicperbetalst[beta], 35, -20) #the year being used for testing
+      bval <- str_sub(aicperbetalst[beta], 35, -20) #the beta value
       aicmean <- colwise(mean)(aic)
@@
 #model selection
-next step... select the models you want for each year and run those models again (primarily based on lowest AICc but check test AUC and other things too)
 
-KEEP A NOTE OF THESE IN ONENOTE
-
-y1998 = 2.75
-y1999 = 2.25
-y2000 = 2
-y2001 = 1.75
-y2002 = 2
-y2003 = 2
-y2004 = 2
-y2005 = 2.25
-y2006 = 1.75
-y2007 = 1.75
-y2008 = 2.25
-y2009 = 2.25
-y2010 = 2.5
-y2011 =  2.5
+lowest aicc is 0.25
 
 ```{r final models plus predict}
-modelmonth <- cbind(paste0(mthna, unique(prback$year)))
-aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
-unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
 
-#for (year in 1:nrow(modelmonth)){
- # timeslice <- modelmonth[year] #testing timeslice
-  #yr <- str_sub(timeslice, -4) #the year being used for testing
-  train <- subset(prback, year != yr) #training data is all years but testing
+aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
+betaval <-  0.25
+
+timeslice <- "all_data"
+train <- prback
 trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-  test <-  subset(prback, year == yr) #creates the testing dataset
-  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
-  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
-  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+train  <- subset(train, select = -c(year, occurrence, month)) #don't need year and occurrence for modelling
+nfolds <- cbind(c(1:5))
 
-  if (yr == "2003" | yr == "2004"){
-    betaval <- 1
-  } else if (yr == "2005"){
-    betaval <- 1.5
-  } else if (yr == "2006"){
-    betaval <- 2.25
-  }
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/all_data/models"), args=c("maximumiterations=5000", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betaval)))
 
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
+    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
+
+  #count number of non-zero coefficients, excluding the final four rows
+    lambdas_list <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = "*lambdas", full.names = TRUE)
+    samppred_list <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = "*samplePredictions.csv", full.names = TRUE)
+
+          #no presences used in model (less those used for testing)
+
+    maxresults <- read.csv(paste0("../output/maxent/", timeslice, "/models/maxentResults.csv"), header = TRUE)
+
+for (i in 1:nrow(aicetc)){   
+for(f in 1:nrow(nfolds)){
+    modn <- nfolds[f]-1
+      lamf <- lambdas_list[f]
+      sampf <- samppred_list[f]
+      coefile <- read.csv(paste0(lamf), header = FALSE, stringsAsFactors=FALSE)
+      coefile <- head(coefile, -4)
+      coefile2 <- subset(coefile, V2 != 0)
+      nocoeff <- nrow(coefile2)
+      bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
+      bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
 
@@
 
-      aicetc$train_AUC <- trainauc
-      aicetc$no_trainng_points <- notrainingpoints
-      aicetc$test_coefficients <- nocoeff
-      aicetc$test_sum.of.logs <- sumsamplepred
-      aicetc$test_AUC <- evauc
-      aicetc$test_evcor <- evcor
-      aicetc$test_evpcor <- evpcor
+    #continuous boyce index
+    predPres <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    predBg <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
+    predPres <- predPres$Cloglog.prediction
+    predBg <- predBg$Cloglog
+    cbi <- contBoyce(pres = predPres, bg = predBg)
+
+      aicetc$train_AUC <- trainauc #training only AUC - not tested
+      aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
+      aicetc$coefficients <- nocoeff #this is to calculate the AIC
+      aicetc$sum.of.logs <- sumsamplepred #this is to calculate the AIC
+      aicetc$test_AUC <- evauc #testing AUC
+      aicetc$test_evcor <- evcor #testing correlation coefficient
+      aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
       aicetc$test_nopresence <- evnp 
@@
       aicetc$test_sensitivity <- evsensitivity
+      aicetc$tss <- max(ev@TPR + ev@TNR) - 1
       aicetc$test_bic <- bic #not sure this is correct...
       aicetc$beta <- betaval
+      aicetc$cbi <- cbi
 
     aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
-    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
+    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$no_trainng_points/(aicetc$no_trainng_points - aicetc$coefficients-1))))
 
@@
     aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
+    aicetc$year <- yr
 
-    write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+}
+}
 
-    unlayer_yr <- subset(unlayer, ascyears == yr)
-    unlayer_yr <- unlayer_yr[-c(1), ]
-    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
-    stackdeplst <- list()
-    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
-    stackdeplst[[year]] <- stackdep
-    #write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
+    write.csv(aicetc, paste0("../output/maxent/", timeslice, "/models/", "_", betaval, "_AICforMaxentReg.csv"), row.names = FALSE)
+
+    saveRDS(mod, file = "../output/maxent/all_data/maxmod.rds", ascii = FALSE, version = NULL, compress = TRUE, refhook = NULL)
+
+```
+
+```{r final models predict}
+
+filenames <- list.files("../data/all_data_layers/", pattern="*.csv", full.names=TRUE)
+unlayer_yrX <- rbindlist(lapply(filenames,fread))
+unlayer_yrX <- unlayer_yrX[complete.cases(unlayer_yrX), ]
 aea <- raster("../output/env/aea.tif") 
+unlayer_yr <- subset(unlayer_yrX, ascmonths == "7")
 
@@
       nao_winter <- projectRaster(nao_winter, aea)
+      amo_sample <- raster(unlayer_yr$amo_sample[td])
+      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(amo_sample) <- "amo_sample"
+      amo_sample <- projectRaster(amo_sample, aea)
+      amo_winter <- raster(unlayer_yr$amo_winter[td])
+      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      names(amo_winter) <- "amo_winter"
+      amo_winter <- projectRaster(amo_winter, aea)
       stackn <- unlayer_yr$stackname[td]
       stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
+
       #stackSave(test,unlayer_yr$stackname[td])
-      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
-      predcum <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
-      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
-      dev.off() # stops automatic saving of the plot to a png
+      pred <- predict(mod, stack, file = paste0("../output/maxent/all_data/predictions/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
+      #prplot <- plot(pred, png(paste0("../output/maxent/all_data/predictions/", "cloglog_", stackn, ".png")))
+      #dev.off() # stops automatic saving of the plot to a png
 
@@
 
-stdep <- do.call(rbind, stackdeplst)
-write.csv(stdep, paste0("../output/maxent/predictions/", mthna, "/stack_depth.csv"), row.names = FALSE)
 ```
@@
 
-```{r avg maxentReg}
-
-for (i in 1:length(modelmonth)){
-  timeslice <- modelmonth[i]
-  yr <- str_sub(timeslice, -4) #the year being used for testing
-  if(yr == 1998){
-    aic1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 1999){
-    aic1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2000){
-    aic2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2001){
-    aic2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2002){
-    aic2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2003){
-    aic2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2004){
-    aic2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2005){
-    aic2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2006){
-    aic2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2007){
-    aic2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2008){
-    aic2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2009){
-    aic2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2010){
-    aic2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2011){
-    aic2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2012){
-    aic2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2013){
-    aic2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2014){
-    aic2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2015){
-    aic2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-  }
-}
-
-aic <- rbind(aic2003, aic2004, aic2005, aic2006)
-maxr <- rbind(maxr2003, maxr2004, maxr2005, maxr2006)
-
-aicm <- colMeans(aic)
-aicm$mod <- "avg"
-avgAIC <- rbind(aic, aicm)
-
-maxr$Species <- 1
-maxrm <- colMeans(maxr)
-maxrm$Species <- "avg"
-maxrm <- rbind(maxr, maxrm)
-
-write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
-write.csv(maxrm, paste0("../output/maxent/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
-```
-
 for this bit may need to stack rasters by depth... this will need some working with the name
@@
 ```{r avg asc layers cloglog}
-asc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "cloglog*.asc", full.names = TRUE)
+stdep <- read.csv("../output/maxent/all_data/stackdepthlist.csv", header = TRUE)
+asc_list <- list.files(paste0("../output/maxent/all_data/predictions/"), pattern = "cloglog.*.asc", full.names = TRUE)
 depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
@@
 }
-write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)
+write.csv(depthlayers, ("../output/maxent/all_data/predictions/cloglog_asclayerbydepth.csv"), row.names = FALSE)
+
+#depthlayers <- read.csv("../output/maxent/all_data/predictions/cloglog_asclayerbydepth.csv", header = FALSE)
 
 for (x in 1:nrow(depthlayers)){
+  dstack <- stack()
   asclay <- depthlayers[x, ]
   asclay <- asclay[!is.na(asclay)]
+  dep <- as.character(asclay[1])
   asclay <- as.data.frame(rbind(asclay))
-  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
-  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
-  asclay$V3 <- as.character(asclay$V3)
-  asclay$V4 <- as.character(asclay$V4)
-  asclay$V5 <- as.character(asclay$V5)
-  asclay$V6 <- as.character(asclay$V6)
-  asclay$V7 <- as.character(asclay$V7)
-  asclay$V8 <- as.character(asclay$V8)
-  asclay$V9 <- as.character(asclay$V9)
-  asclay$V10 <- as.character(asclay$V10)
-  asclay$V11 <- as.character(asclay$V11)
-  asclay$V12 <- as.character(asclay$V12)
-  asclay$V13 <- as.character(asclay$V13)
-  asclay$V14 <- as.character(asclay$V14)
-  asclay$V15 <- as.character(asclay$V15)
-
-  asclaydep <- asclay$V1
-
-  asc1998 <- raster(asclay$V2)
-  asc1999 <- raster(asclay$V3)
-  asc2000 <- raster(asclay$V4)
-  asc2001 <- raster(asclay$V5)
-  asc2002 <- raster(asclay$V6)
-  asc2003 <- raster(asclay$V7)
-  asc2004 <- raster(asclay$V8)
-  asc2005 <- raster(asclay$V9)
-  asc2006 <- raster(asclay$V10)
-  asc2007 <- raster(asclay$V11)
-  asc2008 <- raster(asclay$V12)
-  asc2009 <- raster(asclay$V13)
-  asc2010 <- raster(asclay$V14)
-  asc2011 <- raster(asclay$V15)
-
-  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)
-  avg <- calc(stkasc, fun = mean)
-
-  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cloglog')
-  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".png"))) 
-  dev.off()
+  asclay[1] <- NULL
+  for (ly in 1:length(asclay)){
+    lay <- as.character(asclay[ ,40])
+    layr <- raster(lay)
+    dstack <- stack(dstack, layr)
+    if(ly == length(asclay)){
+      stavg <- calc(dstack, fun = mean, na.rm = TRUE)
+      writeRaster(stavg, (paste0("../output/maxent/all_data/predictions/avg_", dep, ".asc")), overwrite=TRUE)
+      avgplot <- plot(stavg, png(paste0("../output/maxent/all_data/predictions/avg_", dep, ".png")))
+      while (!is.null(dev.list()))  dev.off()
+    }
+  }
 }
 
-ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), glob2rx(pattern = "avg_*.asc"), full.names = TRUE)
+```
 
-clog_asc120  <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_120.asc"))
-clog_asc41.18 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_41.18.asc"))
-clog_asc1.55586 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1.55586.asc"))
-clog_asc1045.85 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1045.85.asc"))
-clog_asc108.03 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_108.03.asc"))
-clog_asc11.7737 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_11.7737.asc"))
-clog_asc1151.99 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1151.99.asc"))
-clog_asc1265.86 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1265.86.asc"))
-clog_asc13.991 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_13.991.asc"))
-clog_asc133.076 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_133.076.asc"))
-clog_asc1387.38 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1387.38.asc"))
-clog_asc147.406 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_147.406.asc"))
-clog_asc16.5253 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_16.5253.asc"))
-clog_asc163.165 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_163.165.asc"))
-clog_asc180.55 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_180.55.asc"))
-clog_asc19.4298 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_19.4298.asc"))
-clog_asc199.79 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_199.79.asc"))
-clog_asc2.66768 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_2.66768.asc"))
-clog_asc22.7576 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_22.7576.asc"))
-clog_asc221.141 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_221.141.asc"))
-clog_asc244.891 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_244.891.asc"))
-clog_asc26.5583 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_26.5583.asc"))
-clog_asc271.356 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_271.356.asc"))
-clog_asc3.85628 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_3.85628.asc"))
-clog_asc30.8746 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_30.8746.asc"))
-clog_asc300.888 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_300.888.asc"))
-clog_asc333.863 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_333.863.asc"))
-clog_asc35.7402 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_35.7402.asc"))
-clog_asc370.689 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_370.689.asc"))
-clog_asc411.794 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_411.794.asc"))
-clog_asc457.626 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_457.626.asc"))
-clog_asc47.2119 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_47.2119.asc"))
-clog_asc5.14036 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_5.14036.asc"))
-clog_asc508.64 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_508.64.asc"))
-clog_asc53.8506 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_53.8506.asc"))
-clog_asc565.292 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_565.292.asc"))
-clog_asc6.54303 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_6.54303.asc"))
-clog_asc61.1128 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_61.1128.asc"))
-clog_asc628.026 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_628.026.asc"))
-clog_asc69.0217 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_69.0217.asc"))
-clog_asc697.259 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_697.259.asc"))
-clog_asc77.6112 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_77.6112.asc"))
-clog_asc773.368 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_773.368.asc"))
-clog_asc8.09252 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_8.09252.asc"))
-clog_asc856.679 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_856.679.asc"))
-clog_asc86.9294 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_86.9294.asc"))
-clog_asc9.82275 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_9.82275.asc"))
-clog_asc947.448 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_947.448.asc"))
-clog_asc97.0413 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_97.0413.asc"))
+```{r binary maps on averaged depth layer}
+#calculate minimal predicted area using ecospat
 
-stackavg <- stack(clog_asc120, clog_asc41.18, clog_asc1.55586, clog_asc1045.85, clog_asc108.03, clog_asc11.7737, clog_asc1151.99, clog_asc1265.86, clog_asc13.991, clog_asc133.076, clog_asc1387.38, clog_asc147.406, clog_asc16.5253, clog_asc163.165, clog_asc180.55, clog_asc19.4298, clog_asc199.79, clog_asc2.66768, clog_asc22.7576, clog_asc221.141, clog_asc244.891, clog_asc26.5583, clog_asc271.356, clog_asc3.85628, clog_asc30.8746, clog_asc300.888, clog_asc333.863, clog_asc35.7402, clog_asc370.689, clog_asc411.794, clog_asc457.626, clog_asc47.2119, clog_asc5.14036, clog_asc508.64, clog_asc53.8506, clog_asc565.292, clog_asc6.54303, clog_asc61.1128, clog_asc628.026, clog_asc69.0217, clog_asc697.259, clog_asc77.6112, clog_asc773.368, clog_asc8.09252, clog_asc856.679, clog_asc86.9294, clog_asc9.82275, clog_asc947.448, clog_asc97.0413)
+samplepred1 <- read.csv(paste0("../output/maxent/all_data/models/species_1_samplePredictions.csv"), header = TRUE)
+mpa_thr1 <- ecospat.mpa(samplepred1$Cloglog.prediction, perc = 0.9) #90%
+samplepred2 <- read.csv(paste0("../output/maxent/all_data/models/species_2_samplePredictions.csv"), header = TRUE)
+mpa_thr2 <- ecospat.mpa(samplepred2$Cloglog.prediction, perc = 0.9) #90%
+samplepred3 <- read.csv(paste0("../output/maxent/all_data/models/species_3_samplePredictions.csv"), header = TRUE)
+mpa_thr3 <- ecospat.mpa(samplepred3$Cloglog.prediction, perc = 0.9) #90%
+samplepred4 <- read.csv(paste0("../output/maxent/all_data/models/species_4_samplePredictions.csv"), header = TRUE)
+mpa_thr4 <- ecospat.mpa(samplepred4$Cloglog.prediction, perc = 0.9) #90%
+samplepred5 <- read.csv(paste0("../output/maxent/all_data/models/species_0_samplePredictions.csv"), header = TRUE)
+mpa_thr5 <- ecospat.mpa(samplepred5$Cloglog.prediction, perc = 0.9) #90%
+mpa_avg <- mean(mpa_thr1, mpa_thr2, mpa_thr3, mpa_thr4, mpa_thr5)
 
-stsum <- sum(stackavg, na.rm = TRUE)
-  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/clog_sum_", mthna, ".asc")), overwrite=TRUE)
-  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/clog_sum_", mthna, ".png"))) 
-  dev.off()
-stavg <- calc(stackavg, fun = mean, na.rm = TRUE)
-  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/clog_avg_", mthna, ".asc")), overwrite=TRUE)
-  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/clog_avg_", mthna, ".png"))) 
+#binary threshold function
+binmpa <- function(x) {
+  ifelse(x <  mpa_thr, 0,
+  ifelse(x >=  mpa_thr, 1, NA))
+  }
+
+clogasc_list <- list.files(paste0("../output/maxent/all_data/predictions/"), pattern = "^avg_.*.asc", full.names = TRUE)
+
+mpa_thr <- mpa_avg
+for (cum in 1:length(clogasc_list)){
+  cumascyr <- str_sub(clogasc_list[46], start = 43) #selects the year from the file name
+  cumascyr <- str_sub(cumascyr, end = -4) #selects the year from the file name
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cum_mpathr <- calc(cumasc, fun = binmpa)
+  writeRaster(cum_mpathr, paste0("../output/maxent/all_data/predictions/binary_mpa_", ascname, ".asc"), overwrite = TRUE)
+  plot(cum_mpathr, png(paste0("../output/maxent/all_data/predictions/binary_mpa_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+  }
+
+```
+
+```{r binary map averaged 2D representation}
+
+binaryclog_list <- list.files(paste0("../output/maxent/all_data/predictions/"), pattern = "^binary_mpa_avg_.*.asc", full.names = TRUE)
+
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, paste0("../output/maxent/all_data/predictions/binary_avg_2d.asc"), overwrite=TRUE)
+plot(stacksum, png(paste0("../output/maxent/all_data/predictions/binary_avg_2d.png")))
 dev.off()
@@
 
-#binary maps
+add land to plots with standard legend (working code) THIS WORK!!!
+```{r}
+aea <- raster("../output/env/aea.tif") #the correct CRS
+maplst <- list.files("../output/maxent/all_data/predictions/", pattern = "binary_avg_2d.*.asc", full.names = TRUE)
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+#brk <- seq(0.1, 1, 0.1) #for legend
+redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
 
-there seem to be two primary ways to create binary maps.
-the most common for conservation purposes seems to be to use maxsumspensspec values, however MPA also comes recommended (but is def less common).
-probably best to use cumulative for both (merrow 2013)
+r <- raster(maplst[1])
+plot(pras)
 
-the thresholds are in the monthavgAIC.csv file in the model folder (not model/month)
-
-```{r binary maps}
-thresholds <- read.csv(paste0("../output/maxent/models/", mthna, "avgAIC.csv"), header = TRUE)
-thresholds <- tail(thresholds, 1)
-mpa <- thresholds$mpa_0.9_cum
-maxsesp <- thresholds$test_maxsumsenspec
-cumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cumu_.*.asc", full.names = TRUE) 
-depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
-names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
-depthlayers$depth <- as.character(depthlayers$depth)
-
-colno <- 1
-
-for (d in 1:nrow(depthlayers)){
-  dep <- depthlayers$depth[d]
-  for (a in 1:length(cumasc_list)){
-    asc <- cumasc_list[a]
-    if (grepl(dep, asc)){
-      depthlayers[d , paste0("a", colno)] <- asc
-        colno <- colno+1
-    }
-  }
-}
-write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cumu_asclayerbydepth.csv")), row.names = TRUE)
-
-for (x in 1:nrow(depthlayers)){
-  asclay <- depthlayers[x, ]
-  asclay <- asclay[!is.na(asclay)]
-  asclay <- as.data.frame(rbind(asclay))
-  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
-  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
-  asclay$V3 <- as.character(asclay$V3)
-  asclay$V4 <- as.character(asclay$V4)
-  asclay$V5 <- as.character(asclay$V5)
-
-  asclaydep <- asclay$V1
-
-  asc2003 <- raster(asclay$V2)
-  asc2004 <- raster(asclay$V3)
-  asc2005 <- raster(asclay$V4)
-  asc2006 <- raster(asclay$V5)
-
-  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
-  avg <- calc(stkasc, fun = mean)
-
-  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/cumu_avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cumulative')
-
-}
-
-cumavg_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cumu_avg_.*.asc", full.names = TRUE) 
-for(lyr in 1:length(cumavg_list)){
-  avgras <- raster(cumavg_list[lyr])
-  reclassmpa <- c(mpa, Inf, 1)
-  reclassmpa <- matrix(reclassmpa, ncol = 3, byrow = TRUE)
-  binavgras <- reclassify(avgras, reclassmpa)
+for (pred in 1:length(maplst)){
+  pras <- raster(maplst[pred])
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  lp <- plot(pras, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  print(lp)
+  dev.copy(png, (paste0("../output/maxent/all_data/predictions/", ascname, pred, ".png")), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
 }

### 2019-07-15 src/testing_file.Rmd
fecky feck violin plots

@@
 ```
+
+# violin plots
+
+```{r}
+#load a test raster
+
+ras <- raster("../output/maxent/predictions/apr/cloglog_st201004000120.asc")
+plot(ras)
+ras2 <- raster("../output/maxent/predictions/oct/cloglog_st201010000120.asc")
+plot(ras2)
+rasd <- as.data.frame(ras)
+rasd$month <- "4"
+rasd2 <- as.data.frame(ras2)
+rasd2$month <- "10"
+names(rasd)[names(rasd) == "cloglog_st201004000120"] <- "prob"
+names(rasd2)[names(rasd2) == "cloglog_st201010000120"] <- "prob"
+rasd$month <- as.factor(rasd$month)
+rasd2$month <- as.factor(rasd2$month)
+rd <- rbind(rasd, rasd2)
+rd
+p <- ggplot(rd, aes(x=month, y = prob)) + geom_violin(trim=FALSE)
+plot(p)
+
+mth <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+for (m in 1:length(mth)){
+  mon <- mth[m]
+  vlst <- list.files(paste0("../output/maxent/predictions/", mon), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+  for (v in 1:length(vlst)){
+    ras <- raster(vlst[v])
+    rasd <- as.data.frame(ras)
+    if (mon == "jan"){
+      rasd$month <- "1"
+      rnm <- names(ras)
+      names(rasd)[names(rasd) == rnm] <- "prob"
+      rasd$month <- as.factor(rasd$month)
+      rd1 <- rasd
+      } else if(mon == "mar"){
+        rasd$month <- "3"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd3 <- rasd
+      } else if(mon == "apr"){
+        rasd$month <- "4"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd4 <- rasd
+      } else if(mon == "may"){
+        rasd$month <- "5"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd5 <- rasd
+      } else if(mon == "jun"){
+        rasd$month <- "6"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd6 <- rasd
+      } else if(mon == "jul"){
+        rasd$month <- "7"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd7 <- rasd
+      } else if(mon == "aug"){
+        rasd$month <- "8"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd8 <- rasd
+      } else if(mon == "sep"){
+        rasd$month <- "9"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd9 <- rasd
+      } else if(mon == "oct"){
+        rasd$month <- "10"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd10 <- rasd
+      } else if(mon == "nov"){
+        rasd$month <- "11"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd11 <- rasd
+      } else if(mon == "dec"){
+        rasd$month <- "12"
+        rnm <- names(ras)
+        names(rasd)[names(rasd) == rnm] <- "prob"
+        rasd$month <- as.factor(rasd$month)
+        rd12 <- rasd
+      }
+
+      rd <- cbind(rd1, rd3, rd4, rd5, rd6, rd7, rd8, rd9, rd10, rd11, rd12)
+    #rd <- as.data.frame(rd)
+    #rd$month <- as.factor(rd$month)
+  }
+
+p <- ggplot(rd2, aes(x=month, y = prob)) + geom_violin(trim=FALSE)
+plot(p)
+
+}
+
+```

### 2019-07-15 src/testing_file.Rmd
eeekkk

@@
 
-      rd <- cbind(rd1, rd3, rd4, rd5, rd6, rd7, rd8, rd9, rd10, rd11, rd12)
+  }
+
+}
+  rd <- rbind(rd1, rd3, rd4, rd5, rd6, rd7, rd8, rd9, rd10, rd11, rd12)
     #rd <- as.data.frame(rd)
     #rd$month <- as.factor(rd$month)
-  }
-
-p <- ggplot(rd2, aes(x=month, y = prob)) + geom_violin(trim=FALSE)
+p <- ggplot(rd, aes(x=month, y = prob)) + geom_violin(trim=FALSE)
 plot(p)
 
-}
-
 ```

### 2019-07-17 src/maxent_alldata_avec_xvalidation.Rmd
qaaa

@@
 options (java.parameters = "-Xmx6g")
-packages <- c("rJava", "dismo", "stringr", "raster", "ecospat", "plyr", "data.table", "RColorBrewer")
+packages <- c("rJava", "dismo", "stringr", "raster", "ecospat", "enmSdm", "plyr", "data.table", "RColorBrewer")
 source("../src/ipak.R")
@@
 
-mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/all_data/models"), args=c("maximumiterations=5000", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betaval)))
+mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/all_data/models"), args=c("writebackgroundpredictions=TRUE", "maximumiterations=5000", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betaval)))
 
@@
 ```
+
+```{r habitat classification on probability - high only}
+
+#binary threshold function
+binprob <- function(x){
+  ifelse(x >= 0.8, 1,
+  ifelse(x < 0.8, 0, NA)) 
+}
+
+clogasc_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^avg_.*.asc", full.names = TRUE)
+
+for (cum in 1:length(clogasc_list)){
+  cumascdep <- str_sub(clogasc_list[cum], start = 47, end = -5) #selects the depth from the file name
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cum_mpathr <- calc(cumasc, fun = binprob)
+  writeRaster(cum_mpathr, paste0("../output/maxent/all_data/predictions/july/binary_p80_", cumascdep, ".asc"), overwrite = TRUE)
+  plot(cum_mpathr, png(paste0("../output/maxent/all_data/predictions/july/binary_p80_", cumascdep, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+  }
+
+```
+
+```{r binary map averaged 2D representation p80}
+
+binaryclog_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^binary_p80.*.asc", full.names = TRUE)
+
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, paste0("../output/maxent/all_data/predictions/july/bin_p80_2d.asc"), overwrite=TRUE)
+plot(stacksum, png(paste0("../output/maxent/all_data/predictions/july/bin_avg_p80_2d.png")))
+dev.off()
+
+```
+
+```{r}
+aea <- raster("../output/env/aea.tif") #the correct CRS
+maplst <- list.files("../output/maxent/all_data/predictions/", pattern = "binary_avg_2d.*.asc", full.names = TRUE)
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+#brk <- seq(0.1, 1, 0.1) #for legend
+redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+
+  pras <- raster("../output/maxent/predictions/jul/binary_highavg_2d.asc")
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  print(lp)
+  dev.copy(png, ("../output/maxent/all_data/predictions/july/oldmod_p80_2d_july.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+```
+
+```{r}
+       #continuous boyce index 
+for(f in 1:nrow(nfolds)){
+  fold <- (f - 1)
+  predPres <- read.csv(paste0("../output/maxent/all_data/models/species_", fold,  "_samplePredictions.csv"), header = TRUE)
+  predBg <- read.csv(paste0("../output/maxent/all_data/models/species_", fold, "_backgroundPredictions.csv"), header = TRUE)
+  predPres <- predPres$Cloglog.prediction
+  predBg <- predBg$Cloglog
+  cbi <- contBoyce(pres = predPres, bg = predBg)
+  write.csv(cbi, paste0("../output/maxent/all_data/cbi_", fold, ".csv"))
+
+}
+```

### 2019-07-17 src/testing_file.Rmd
qaaa

@@
 ```
+
+.
+For key predictor variables, can you plot probability of occurrence (y-axis) vs predictor (x-axis).
+I would like to see raw data which will be 1 for presence and 0 for background and then overlay the model predicted relationships on top of the raw data for the two models (full year, month).
+Functiona "abline (model name)" plots the model prediction for many models in R - not sure this will work for MaxEnt.
+This will show how the predictions vary for the different variables - perhaps it is just one driving the differences.
+Does this make sense?
+
+first need to add the same depth indicator as used in the file names
+
+```{r}
+completeFun <- function(data, desiredCols) {
+  completeVec <- complete.cases(data[, desiredCols])
+  return(data[completeVec, ])
+}
+
+data_aea <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+data_aea$depthlyr_mod <- NA
+data_aea$predicted <- NA
+data_aea <- completeFun(data_aea, "depth_layer")
+
+avgasc_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
+depths <- list()
+for (i in 1:length(avgasc_list)){
+  ascname <-  str_sub(avgasc_list[i], 48, -5)
+  depths[[paste0(i)]] <- ascname
+}
+
+for (j in 1:nrow(data_aea)){
+  d <- data_aea$depth_layer[j]
+  d3 <- str_sub(d, end = 3) #extract first 3 digits 
+   for (k in 1:length(depths)){
+    l <- depths[[k]]
+    l3 <- str_sub(l, end = 3) #extract first 3 digits
+    if (d3 == l3){
+      data_aea$depthlyr_mod[j] <- l
+    }
+  }
+}
+data_aea <- completeFun(data_aea, "depthlyr_mod")
+
+```
+
+```{r}
+mth <- c("1", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
+mthn <- c("jan", "mar", "apr", "may", "jun", "july", "aug", "sep", "oct", "nov","dec")
+aea <- raster("../output/env/aea.tif")
+
+#for (m in 1:length(mth)){
+#mont <- mth[m]
+mont <- 7
+data_aeaX <- subset(data_aea, month == mont)
+xy <- data_aeaX[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aeaX, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+avgasc_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
+for (j in 1:nrow(data_aea_sp)) { 
+for (i in 1:length(avgasc_list)) {  
+ de <- data_aea_sp$depthlyr_mod[j]  # a variable for the observation depth layer
+  mth <- (data_aea_sp$month[j])  # a variable for the observation month
+  rasdep <- str_sub(avgasc_list[i], 48, -5)
+       if (mth == mont & rasdep == de){
+          temp_brick <- raster(avgasc_list[i])
+        crs(temp_brick) <- aea@crs 
+              data_aea_sp$predicted[j] <- extract(x=temp_brick, y = data_aea_sp[j, ])
+
+              }
+          }
+}
+
+#}
+              data_aea1 <- as.data.frame(data_aea_sp)
+
+write.csv(data_aea1, "../output/maxent/all_data/testforslr_jul.csv", row.names = FALSE)
+ #}
+
+ # data_aea1 <- as.data.frame(data_aea_sp)
+
+```
+
+```{r}
+predjan <- read.csv("../output/maxent/all_data/testforslr_jan.csv", header = TRUE)
+predjanpr <- subset(predjan, occurrence == 1)
+plot(predjanpr$temp_depth, predjanpr$predicted)
+predjanbk <- subset(predjan, occurrence == 0)
+plot(predjanbk$temp_depth, predjanbk$predicted)
+```
+
+```{r}
+mod <- readRDS(file = "../output/maxent/all_data/maxmod.rds", refhook = NULL)
+px <- response(mod@models[[3]])
+```

### 2019-07-30 src/maxent_alldata_avec_xvalidation.Rmd
feckit

@@
 ```{r final models predict}
+mod <- readRDS(file = "../output/maxent/all_data/maxmod.rds", refhook = NULL)
+px <- response(mod@models[[4]])
 
@@
 aea <- raster("../output/env/aea.tif") 
-unlayer_yr <- subset(unlayer_yrX, ascmonths == "7")
+unlayer_yr <- subset(unlayer_yrX, ascmonths == "11")
 
@@
       #stackSave(test,unlayer_yr$stackname[td])
-      pred <- predict(mod, stack, file = paste0("../output/maxent/all_data/predictions/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
+      pred <- predict(mod, stack, file = paste0("../output/maxent/all_data/predictions/nov/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
       #prplot <- plot(pred, png(paste0("../output/maxent/all_data/predictions/", "cloglog_", stackn, ".png")))
@@
 stdep <- read.csv("../output/maxent/all_data/stackdepthlist.csv", header = TRUE)
-asc_list <- list.files(paste0("../output/maxent/all_data/predictions/"), pattern = "cloglog.*.asc", full.names = TRUE)
+asc_list <- list.files(paste0("../output/maxent/all_data/predictions/nov"), pattern = "cloglog.*.asc", full.names = TRUE)
 depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
@@
 }
-write.csv(depthlayers, ("../output/maxent/all_data/predictions/cloglog_asclayerbydepth.csv"), row.names = FALSE)
+write.csv(depthlayers, ("../output/maxent/all_data/predictions/nov/cloglog_asclayerbydepth.csv"), row.names = FALSE)
 
@@
       stavg <- calc(dstack, fun = mean, na.rm = TRUE)
-      writeRaster(stavg, (paste0("../output/maxent/all_data/predictions/avg_", dep, ".asc")), overwrite=TRUE)
-      avgplot <- plot(stavg, png(paste0("../output/maxent/all_data/predictions/avg_", dep, ".png")))
+      writeRaster(stavg, (paste0("../output/maxent/all_data/predictions/nov/avg_", dep, ".asc")), overwrite=TRUE)
+      avgplot <- plot(stavg, png(paste0("../output/maxent/all_data/predictions/nov/avg_", dep, ".png")))
       while (!is.null(dev.list()))  dev.off()
@@
 outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+pronly <- subset(presback, occurrence == 1)
+prjul <- subset(pronly, month == 7)
+prjan <- subset(pronly, month == 1)
 
@@
   pras[pras >0] <- 1
-  lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
-  lp <- plot(pras)
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
   lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
   lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prjul$longitude_meters, prjul$latitude_meters, add = TRUE)
   print(lp)

### 2019-07-30 src/testing_file.Rmd
feckit

@@
 
-```{r}
-df <- iris
-df <- na.omit(df)
-df <- scale(df[ ,-5])
-
-# Dissimilarity matrix
-d <- dist(df, method = "euclidean")
-
-# Hierarchical clustering using Complete Linkage
-hc1 <- hclust(d, method = "complete" )
-
-# Plot the obtained dendrogram
-plot(hc1, cex = 0.6, hang = -1)
-```
-
-```{r}
-iris2 <- pres[ ,-12]
-species_labels <- pres[ ,12]
-library(colorspace) # get nice colors
-species_col <- rev(rainbow_hcl(4))[as.numeric(species_labels)]
-
-# Plot a SPLOM:
-pairs(iris2, col = species_col,
-      lower.panel = NULL,
-       cex.labels=2, pch=19, cex = 1.2)
-
-# Add a legend
-par(xpd = TRUE)
-legend(x = 0.05, y = 0.4, cex = 2,
-   legend = as.character(levels(species_labels)),
-    fill = unique(species_col))
-par(xpd = NA)
-```
-
-```{r}
-par(las = 1, mar = c(4.5, 3, 3, 2) + 0.1, cex = .8)
-MASS::parcoord(iris2, col = species_col, var.label = TRUE, lwd = 2)
-
-# Add Title
-title("Parallel coordinates plot of the Iris data")
-# Add a legend
-par(xpd = TRUE)
-legend(x = 1.75, y = -.25, cex = 1,
-   legend = as.character(levels(species_labels)),
-    fill = unique(species_col), horiz = TRUE)
-```
-
-```{r}
-d_iris <- dist(iris2) # method="man" # is a bit better
-hc_iris <- hclust(d_iris, method = "complete")
-iris_species <- rev(levels(pres[,12]))
-
-library(dendextend)
-dend <- as.dendrogram(hc_iris)
-# order it the closest we can to the order of the observations:
-dend <- rotate(dend, 1:6361)
-
-# Color the branches based on the clusters:
-dend <- color_branches(dend, k=4) #, groupLabels=iris_species)
-
-# Manually match the labels, as much as possible, to the real classification of the flowers:
-labels_colors(dend) <-
-   rainbow_hcl(3)[sort_levels_values(
-      as.numeric(pres[,12])[order.dendrogram(dend)]
-   )]
-
-# We shall add the flower type to the labels:
-labels(dend) <- paste(as.character(iris[,5])[order.dendrogram(dend)],
-                           "(",labels(dend),")", 
-                           sep = "")
-# We hang the dendrogram a bit:
-dend <- hang.dendrogram(dend,hang_height=0.1)
-# reduce the size of the labels:
-# dend <- assign_values_to_leaves_nodePar(dend, 0.5, "lab.cex")
-dend <- set(dend, "labels_cex", 0.5)
-# And plot:
-par(mar = c(3,3,3,7))
-plot(dend, 
-     main = "Clustered Iris data set
-     (the labels give the true flower species)", 
-     horiz =  TRUE,  nodePar = list(cex = .007))
-legend("topleft", legend = iris_species, fill = rainbow_hcl(3))
-```
-
-```{r}
-library(circlize)
-par(mar = rep(0,4))
-circlize_dendrogram(dend)
-```
-
-```{r}
-iris3 <- pres[ ,-2]
-species_labels <- pres[ ,2]
-library(colorspace) # get nice colors
-species_col <- rev(rainbow_hcl(12))[as.numeric(species_labels)]
-
-# Plot a SPLOM:
-pairs(iris3, col = species_col,
-      lower.panel = NULL,
-       cex.labels=2, pch=19, cex = 1.2)
-
-# Add a legend
-par(xpd = TRUE)
-legend(x = 0.05, y = 0.4, cex = 2,
-   legend = as.character(levels(species_labels)),
-    fill = unique(species_col))
-par(xpd = NA)
-```
-
-```{r}
-d_iris2 <- dist(pres) # method="man" # is a bit better
-hc_iris2 <- hclust(d_iris2, method = "complete")
-iris_species2 <- rev(levels(pres[,12]))
-
-library(dendextend)
-dend2 <- as.dendrogram(hc_iris2)
-# order it the closest we can to the order of the observations:
-dend2 <- rotate(dend2, 1:6361)
-
-# Color the branches based on the clusters:
-dend2 <- color_branches(dend2, k=4) #, groupLabels=iris_species)
-
-# Manually match the labels, as much as possible, to the real classification of the flowers:
-labels_colors(dend2) <-
-   rainbow_hcl(3)[sort_levels_values(
-      as.numeric(pres[,12])[order.dendrogram(dend2)]
-   )]
-
-# We shall add the flower type to the labels:
-labels(dend2) <- paste(as.character(iris[,5])[order.dendrogram(dend2)],
-                           "(",labels(dend2),")", 
-                           sep = "")
-# We hang the dendrogram a bit:
-dend2 <- hang.dendrogram(dend2,hang_height=0.1)
-# reduce the size of the labels:
-# dend <- assign_values_to_leaves_nodePar(dend, 0.5, "lab.cex")
-dend22 <- set(dend2, "labels_cex", 0.5)
-# And plot:
-par(mar = c(3,3,3,7))
-plot(dend2, 
-     main = "Clustered Iris data set
-     (the labels give the true flower species)", 
-     horiz =  TRUE,  nodePar = list(cex = .007))
-legend("topleft", legend = species_labels, fill = rainbow_hcl(3))
-```
-
-```{r}
-library(circlize)
-par(mar = rep(0,4))
-circlize_dendrogram(dend2)
-```
-
 # violin plots
@@
 data_aea$depthlyr_mod <- NA
-data_aea$predicted <- NA
+data_aea$predicted_alldata <- NA
 data_aea <- completeFun(data_aea, "depth_layer")
@@
 ```{r}
-mth <- c("1", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
-mthn <- c("jan", "mar", "apr", "may", "jun", "july", "aug", "sep", "oct", "nov","dec")
+#mth <- c("1", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
+#mthn <- c("jan", "may", "apr", "may", "jun", "july", "aug", "sep", "oct", "nov","dec")
 aea <- raster("../output/env/aea.tif")
@@
 #mont <- mth[m]
-mont <- 7
+mont <- 11
 data_aeaX <- subset(data_aea, month == mont)
@@
 data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aeaX, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
-avgasc_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
+avgasc_list <- list.files(paste0("../output/maxent/all_data/predictions/nov/"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
 for (j in 1:nrow(data_aea_sp)) { 
@@
   mth <- (data_aea_sp$month[j])  # a variable for the observation month
-  rasdep <- str_sub(avgasc_list[i], 48, -5)
+  rasdep <- str_sub(avgasc_list[i], 47, -5)
        if (mth == mont & rasdep == de){
@@
 
-write.csv(data_aea1, "../output/maxent/all_data/testforslr_jul.csv", row.names = FALSE)
+write.csv(data_aea1, "../output/maxent/all_data/pred_file_nov.csv", row.names = FALSE)
  #}
@@
  # data_aea1 <- as.data.frame(data_aea_sp)
+
+```
+
+```{r}
+
+aea <- raster("../output/env/aea.tif")
+
+#for (m in 1:length(mth)){
+#mont <- mth[m]
+mont <- 11
+data_aeaX <- read.csv("../output/maxent/all_data/pred_file_nov.csv", header = TRUE)
+data_aeaX$mthpred <- NA
+xy <- data_aeaX[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aeaX, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+avgasc_list <- list.files(paste0("../output/maxent/predictions/nov/"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
+for (j in 1:nrow(data_aea_sp)) { 
+for (i in 1:length(avgasc_list)) {  
+ de <- data_aea_sp$depthlyr_mod[j]  # a variable for the observation depth layer
+  mth <- (data_aea_sp$month[j])  # a variable for the observation month
+  rasdep <- str_sub(avgasc_list[i], 46, -5)
+       if (mth == mont & rasdep == de){
+          temp_brick <- raster(avgasc_list[i])
+        crs(temp_brick) <- aea@crs 
+              data_aea_sp$mthpred[j] <- extract(x=temp_brick, y = data_aea_sp[j, ])
+
+              }
+          }
+}
+
+#}
+              data_aea1 <- as.data.frame(data_aea_sp)
+              write.csv(data_aea1, "../output/maxent/all_data/pred_file_nov.csv", row.names = FALSE)
 
@@
 mod <- readRDS(file = "../output/maxent/all_data/maxmod.rds", refhook = NULL)
-px <- response(mod@models[[3]])
+px <- response(mod@models[[4]])
 ```
+
+#get all probabiltiy values for a period of time and find the upper 20%
+
+for july with all data
+```{r}
+clogasc_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^avg_.*.asc", full.names = TRUE) # first list the files containing the data
+a <- stack(clogasc_list) #stack the files
+b <- getValues(a) # create a matrix of the values 
+c <- as.vector(b) #make into vector
+d <- na.omit(c) #remove all na values
+d <- sort(d, decreasing = TRUE) # order from high to low
+drow <- length(d) # how many values are there
+dr20 <- (drow*20)/100  # calculate which value is the upper 20%
+d20 <- d[dr20] # 0.149577....
+dmaxv <- max(d) #0.998087
+dm20 <- (dmaxv*80)/100 # calculate the 20% of 0.998097...  = 0.798477....
+
+#binary threshold function d20
+bind20 <- function(x){
+  ifelse(x >= d20, 1,
+  ifelse(x < d20, 0, NA)) 
+}
+
+#binary threshold function dm20
+bindm20 <- function(x){
+  ifelse(x >= dm20, 1,
+  ifelse(x < dm20, 0, NA)) 
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumascdep <- str_sub(clogasc_list[cum], start = 47, end = -5) #selects the depth from the file name
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cum_d20 <- calc(cumasc, fun = bind20)
+  cum_dm20 <- calc(cumasc, fun = bindm20)
+  writeRaster(cum_d20, paste0("../output/maxent/slr_analysis/july_alldata/july_alldata_d20_", ascname, ".asc"), overwrite = TRUE)
+  writeRaster(cum_dm20, paste0("../output/maxent/slr_analysis/july_alldata/july_alldata_dm20_", ascname, ".asc"), overwrite = TRUE)
+  }
+
+binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/july_alldata"), pattern = "^july_alldata_d20_.*.asc", full.names = TRUE)
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, "../output/maxent/slr_analysis/july_alldata/d20_2d.asc", overwrite=TRUE)
+plot(stacksum, png("../output/maxent/slr_analysis/july_alldata/d20_2d.png"))
+dev.off()
+binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/july_alldata"), pattern = "^july_alldata_dm20_.*.asc", full.names = TRUE)
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, "../output/maxent/slr_analysis/july_alldata/dm20_2d.asc", overwrite=TRUE)
+plot(stacksum, png("../output/maxent/slr_analysis/july_alldata/dm20_2d.png"))
+dev.off()
+
+#make map
+
+aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+#brk <- seq(0.1, 1, 0.1) #for legend
+redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+pronly <- subset(presback, occurrence == 1)
+prjul <- subset(pronly, month == 7)
+#prjan <- subset(pronly, month == 1)
+
+  pras <- raster("../output/maxent/slr_analysis/july_alldata/d20_2d.asc")
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prjul$longitude_meters, prjul$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
+  print(lp)
+  dev.copy(png, ("../output/maxent/slr_analysis/july_alldata/d20_2d.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+
+  pras <- raster("../output/maxent/slr_analysis/july_alldata/dm20_2d.asc")
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prjul$longitude_meters, prjul$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
+  print(lp)
+  dev.copy(png, ("../output/maxent/slr_analysis/july_alldata/dm20_2d.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+
+```
+
+for july with just july data
+```{r}
+clogasc_list <- list.files(paste0("../output/maxent/predictions/jul/"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) # first list the files containing the data
+a <- stack(clogasc_list) #stack the files
+b <- getValues(a) # create a matrix of the values 
+c <- as.vector(b) #make into vector
+d <- na.omit(c) #remove all na values
+d <- sort(d, decreasing = TRUE) # order from high to low
+drow <- length(d) # how many values are there
+dr20 <- (drow*20)/100  # calculate which value is the upper 20%
+d20 <- d[dr20] # 0.083319....
+dmaxv <- max(d) #0.885157....
+dm20 <- (dmaxv*80)/100 # calculate the 20% of 0.885157....  = 0.70812....
+
+#binary threshold function d20
+bind20 <- function(x){
+  ifelse(x >= d20, 1,
+  ifelse(x < d20, 0, NA)) 
+}
+
+#binary threshold function dm20
+bindm20 <- function(x){
+  ifelse(x >= dm20, 1,
+  ifelse(x < dm20, 0, NA)) 
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumascdep <- str_sub(clogasc_list[cum], start = 46, end = -5) #selects the depth from the file name
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cum_d20 <- calc(cumasc, fun = bind20)
+  cum_dm20 <- calc(cumasc, fun = bindm20)
+  writeRaster(cum_d20, paste0("../output/maxent/slr_analysis/july_alldata/july_monthdata_d20_", ascname, ".asc"), overwrite = TRUE)
+  writeRaster(cum_dm20, paste0("../output/maxent/slr_analysis/july_alldata/july_monthdata_dm20_", ascname, ".asc"), overwrite = TRUE)
+  }
+
+binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/july_alldata"), pattern = "^july_monthdata_d20_.*.asc", full.names = TRUE)
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, "../output/maxent/slr_analysis/july_alldata/d20_2d_month.asc", overwrite=TRUE)
+plot(stacksum, png("../output/maxent/slr_analysis/july_alldata/d20_2d_month.png"))
+dev.off()
+binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/july_alldata"), pattern = "^july_monthdata_dm20_.*.asc", full.names = TRUE)
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, "../output/maxent/slr_analysis/july_alldata/dm20_2d_month.asc", overwrite=TRUE)
+plot(stacksum, png("../output/maxent/slr_analysis/july_alldata/dm20_2d_month.png"))
+dev.off()
+
+#make map
+
+aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+#brk <- seq(0.1, 1, 0.1) #for legend
+redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+pronly <- subset(presback, occurrence == 1)
+prjul <- subset(pronly, month == 7)
+#prjan <- subset(pronly, month == 1)
+
+  pras <- raster("../output/maxent/slr_analysis/july_alldata/d20_2d_month.asc")
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prjul$longitude_meters, prjul$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
+  print(lp)
+  dev.copy(png, ("../output/maxent/slr_analysis/july_alldata/d20_2d_month.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+
+  pras <- raster("../output/maxent/slr_analysis/july_alldata/dm20_2d_month.asc")
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prjul$longitude_meters, prjul$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
+  print(lp)
+  dev.copy(png, ("../output/maxent/slr_analysis/july_alldata/dm20_2d_month.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+```
+
+for jan with all data
+```{r}
+clogasc_list <- list.files(paste0("../output/maxent/all_data/predictions/jan/"), pattern = "^avg_.*.asc", full.names = TRUE) # first list the files containing the data
+a <- stack(clogasc_list) #stack the files
+b <- getValues(a) # create a matrix of the values 
+c <- as.vector(b) #make into vector
+d <- na.omit(c) #remove all na values
+d <- sort(d, decreasing = TRUE) # order from high to low
+drow <- length(d) # how many values are there
+dr20 <- (drow*20)/100  # calculate which value is the upper 20%
+d20 <- d[dr20] # 0.08307....
+dmaxv <- max(d) #0.996738...
+dm20 <- (dmaxv*80)/100 # calculate the 20% of 0.996738...  = 0.7973906....
+
+#binary threshold function d20
+bind20 <- function(x){
+  ifelse(x >= d20, 1,
+  ifelse(x < d20, 0, NA)) 
+}
+
+#binary threshold function dm20
+bindm20 <- function(x){
+  ifelse(x >= dm20, 1,
+  ifelse(x < dm20, 0, NA)) 
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumascdep <- str_sub(clogasc_list[cum], start = 47, end = -5) #selects the depth from the file name
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cum_d20 <- calc(cumasc, fun = bind20)
+  cum_dm20 <- calc(cumasc, fun = bindm20)
+  writeRaster(cum_d20, paste0("../output/maxent/slr_analysis/jan_alldata/jan_alldata_d20_", ascname, ".asc"), overwrite = TRUE)
+  writeRaster(cum_dm20, paste0("../output/maxent/slr_analysis/jan_alldata/jan_alldata_dm20_", ascname, ".asc"), overwrite = TRUE)
+  }
+
+binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/jan_alldata"), pattern = "^jan_alldata_d20_.*.asc", full.names = TRUE)
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, "../output/maxent/slr_analysis/jan_alldata/d20_2d.asc", overwrite=TRUE)
+plot(stacksum, png("../output/maxent/slr_analysis/jan_alldata/d20_2d.png"))
+dev.off()
+binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/jan_alldata"), pattern = "^jan_alldata_dm20_.*.asc", full.names = TRUE)
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, "../output/maxent/slr_analysis/jan_alldata/dm20_2d.asc", overwrite=TRUE)
+plot(stacksum, png("../output/maxent/slr_analysis/jan_alldata/dm20_2d.png"))
+dev.off()
+
+#make map
+
+aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+#brk <- seq(0.1, 1, 0.1) #for legend
+redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+pronly <- subset(presback, occurrence == 1)
+#prjul <- subset(pronly, month == 7)
+prjan <- subset(pronly, month == 1)
+
+  pras <- raster("../output/maxent/slr_analysis/jan_alldata/d20_2d.asc")
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prjan$longitude_meters, prjan$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
+  print(lp)
+  dev.copy(png, ("../output/maxent/slr_analysis/jan_alldata/d20_2d.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+
+  pras <- raster("../output/maxent/slr_analysis/jan_alldata/dm20_2d.asc")
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prjan$longitude_meters, prjan$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
+  print(lp)
+  dev.copy(png, ("../output/maxent/slr_analysis/jan_alldata/dm20_2d.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+
+```
+
+for jan with just jan data
+```{r}
+clogasc_list <- list.files(paste0("../output/maxent/predictions/jan/"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) # first list the files containing the data
+a <- stack(clogasc_list) #stack the files
+b <- getValues(a) # create a matrix of the values 
+c <- as.vector(b) #make into vector
+d <- na.omit(c) #remove all na values
+d <- sort(d, decreasing = TRUE) # order from high to low
+drow <- length(d) # how many values are there
+dr20 <- (drow*20)/100  # calculate which value is the upper 20%
+d20 <- d[dr20] # 0.01165....
+dmaxv <- max(d) #0.73946....
+dm20 <- (dmaxv*80)/100 # calculate the 20% of 0.73946....  = 0.591569....
+
+#binary threshold function d20
+bind20 <- function(x){
+  ifelse(x >= d20, 1,
+  ifelse(x < d20, 0, NA)) 
+}
+
+#binary threshold function dm20
+bindm20 <- function(x){
+  ifelse(x >= dm20, 1,
+  ifelse(x < dm20, 0, NA)) 
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumascdep <- str_sub(clogasc_list[cum], start = 46, end = -5) #selects the depth from the file name
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cum_d20 <- calc(cumasc, fun = bind20)
+  cum_dm20 <- calc(cumasc, fun = bindm20)
+  writeRaster(cum_d20, paste0("../output/maxent/slr_analysis/jan_alldata/jan_monthdata_d20_", ascname, ".asc"), overwrite = TRUE)
+  writeRaster(cum_dm20, paste0("../output/maxent/slr_analysis/jan_alldata/jan_monthdata_dm20_", ascname, ".asc"), overwrite = TRUE)
+  }
+
+binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/jan_alldata"), pattern = "^jan_monthdata_d20_.*.asc", full.names = TRUE)
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, "../output/maxent/slr_analysis/jan_alldata/d20_2d_month.asc", overwrite=TRUE)
+plot(stacksum, png("../output/maxent/slr_analysis/jan_alldata/d20_2d_month.png"))
+dev.off()
+binaryclog_list <- list.files(paste0("../output/maxent/slr_analysis/jan_alldata"), pattern = "^jan_monthdata_dm20_.*.asc", full.names = TRUE)
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, "../output/maxent/slr_analysis/jan_alldata/dm20_2d_month.asc", overwrite=TRUE)
+plot(stacksum, png("../output/maxent/slr_analysis/jan_alldata/dm20_2d_month.png"))
+dev.off()
+
+#make map
+
+aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+#brk <- seq(0.1, 1, 0.1) #for legend
+redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+pronly <- subset(presback, occurrence == 1)
+prjul <- subset(pronly, month == 7)
+#prjan <- subset(pronly, month == 1)
+
+  pras <- raster("../output/maxent/slr_analysis/jan_alldata/d20_2d_month.asc")
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prjan$longitude_meters, prjan$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
+  print(lp)
+  dev.copy(png, ("../output/maxent/slr_analysis/jan_alldata/d20_2d_month.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+
+  pras <- raster("../output/maxent/slr_analysis/jan_alldata/dm20_2d_month.asc")
+  crs(pras) <- apr
+  ascname <- names(pras)
+  pras[pras >0] <- 1
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prjan$longitude_meters, prjan$latitude_meters, add = TRUE, pch=19, cex = 0.5, col = "red")
+  print(lp)
+  dev.copy(png, ("../output/maxent/slr_analysis/jan_alldata/dm20_2d_month.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+```
+
+```{r}
+
+mont <- 
+data_aeaX <- read.csv("../output/maxent/all_data/testforslr_jan.csv", header = TRUE)
+data_aeaX$predicted_mthdata <- NA
+
+xy <- data_aeaX[ ,c(19,20)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
+data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aeaX, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
+avgasc_list <- list.files(paste0("../output/maxent/predictions/jan"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
+for (j in 1:nrow(data_aea_sp)) { 
+for (i in 1:length(avgasc_list)) {  
+ de <- data_aea_sp$depthlyr_mod[j]  # a variable for the observation depth layer
+  rasdep <- str_sub(avgasc_list[i], 46, -5)
+       if (rasdep == de){
+          temp_brick <- raster(avgasc_list[i])
+        crs(temp_brick) <- aea@crs 
+              data_aea_sp$predicted_mthdata[j] <- extract(x=temp_brick, y = data_aea_sp[j, ])
+
+              }
+          }
+}
+
+           data_aea1 <- as.data.frame(data_aea_sp)
+
+write.csv(data_aea1, "../output/maxent/all_data/predictions_jan.csv", row.names = FALSE)
+ #}
+
+```
+
+#code to create benthic layers
+
+```{r}
+aea <- raster("../output/env/aea.tif") #the correct CRS
+benlst <- list.files("../data/env/ascii/benthic_layers/template", pattern = ".*.asc", full.names = TRUE)
+apr <- aea@crs
+
+for (pred in 1:length(benlst)){
+  pras <- raster(benlst[pred])
+  crs(pras) <- apr
+  ascname <- names(pras)
+  depth <-  (sapply(strsplit(ascname, "_"), "[[", 4)) # extracting the third part of the netcdf (inc.nc)
+  pras[pras >0] <- 0
+  pras[is.na(pras)] <- 1
+  writeRaster(pras, paste0("../data/env/ascii/benthic_layers/done/benthic_", depth, ".asc"), overwrite = TRUE)
+
+}
+
+#just check it works
+benlst <- list.files("../data/env/ascii/benthic_layers/done", pattern = ".*.asc", full.names = TRUE)
+
+for (pred in 1:length(benlst)){
+  pras <- raster(benlst[pred])
+  plot(pras)
+}
+
+```
+
+```{r}
+predjan <- read.csv("../output/maxent/all_data/predictions_jan.csv", header = TRUE)
+plot(predjan$occurrence, predjan$predicted_mthdata)
+plot(predjan$occurrence, predjan$predicted_alldata)
+```
+
+```{r}
+predjul <- read.csv("../output/maxent/all_data/predictions_jul.csv", header = TRUE)
+plot(predjul$occurrence, predjul$predicted_mthdata)
+plot(predjul$occurrence, predjul$predicted_alldata)
+```
+
+```{r}
+predjan <- read.csv("../output/maxent/all_data/predictions_jan.csv", header = TRUE)
+
+completeFun <- function(data, desiredCols) {
+  completeVec <- complete.cases(data[, desiredCols])
+  return(data[completeVec, ])
+}
+
+predjan <- completeFun(predjan, "predicted_mthdata")
+predjanoc <- subset(predjan, occurrence == 1)
+predjanbk <- subset(predjan, occurrence == 0)
+noc <- nrow(predjanoc)
+nbk <- nrow(predjanbk)
+janmth5plusoc <- sum(predjanoc$predicted_mthdata > 0.5)
+janmth5plusbk <- sum(predjanoc$predicted_mthdata > 0.5)
+janmthpp <- (janmth5plusoc/noc)*100
+janmthbkp <- (janmth5plusbk/nbk)*100
+
+janall5plusoc <- sum(predjanoc$predicted_alldata > 0.5)
+janall5plusbk <- sum(predjanoc$predicted_alldata > 0.5)
+janallpp <- (janall5plusoc/noc)*100
+janallbkp <- (janall5plusbk/nbk)*100
+```
+
+```{r}
+hist(predjanoc$predicted_mthdata)
+hist(predjanoc$predicted_alldata)
+hist(predjanbk$predicted_mthdata)
+hist(predjanbk$predicted_alldata)
+```
+
+```{r}
+predjul <- read.csv("../output/maxent/all_data/predictions_jul.csv", header = TRUE)
+
+completeFun <- function(data, desiredCols) {
+  completeVec <- complete.cases(data[, desiredCols])
+  return(data[completeVec, ])
+}
+
+predjul <- completeFun(predjul, "predicted_mthdata")
+predjuloc <- subset(predjul, occurrence == 1)
+predjulbk <- subset(predjul, occurrence == 0)
+noc <- nrow(predjuloc)
+nbk <- nrow(predjulbk)
+
+julmth5plusoc <- sum(predjuloc$predicted_mthdata > 0.5)
+julmth5plusbk <- sum(predjuloc$predicted_mthdata > 0.5)
+julmthpp <- (julmth5plusoc/noc)*100
+julmthbkp <- (julmth5plusbk/nbk)*100
+
+julall5plusoc <- subset(predjuloc, predicted_alldata > 0.5)
+julall5plusoc <- nrow(julall5plusoc)
+julall5plusbk <- subset(predjulbk, predicted_alldata > 0.5)
+julall5plusbk <- nrow(julall5plusbk)
+julallpp <- (julall5plusoc/noc)*100
+julallbkp <- (julall5plusbk/nbk)*100
+```
+
+```{r}
+hist(predjuloc$predicted_mthdata)
+hist(predjuloc$predicted_alldata)
+hist(predjulbk$predicted_mthdata)
+hist(predjulbk$predicted_alldata)
+```

### 2019-08-30 src/AO_NAO_comparison.Rmd
laptop update

@@
 ```{r}
-naoraw <- read.csv("./data/env/nao_pc_monthly.csv", header = FALSE) 
+naoraw <- read.csv("../data/env/nao_pc_monthlyX.csv", header = FALSE) 
 head(naoraw)
@@
 ```{r}
-colnames(naoraw) <- c("year","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
+colnames(naoraw) <- c("year","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "yr_t")
 head(naoraw)
@@
 write.csv(naodecq1, "../output/env/nao_dec_q1.csv")
+naoalltq1 <- subset(nao$year, nao$yr_t <= naoq1$yr_t)
+write.csv(naoalltq1, "../output/env/nao_yrt_q1.csv")
 
@@
 write.csv(naodecq4, "../output/env/nao_dec_q4.csv") 
+naoalltq4 <- subset(nao$year, nao$yr_t >= naoq4$yr_t)
+write.csv(naoalltq4, "../output/env/nao_yrt_q4.csv")
 
@@
 write.csv(aodecq1, "../output/env/ao_dec_q1.csv")  
+aoalltq1 <- subset(ao$X, ao$yr_t <= aoq1$yr_t)
+write.csv(aoalltq1, "../output/env/ao_yrt_q1.csv")
 
-aoq1_all <- rbind(aojanq1, aofebq1, aomarq1, aoaprq1, aomayq1, aojunq1, aojulq1, aoaugq1, aosepq1, aooctq1, aonovq1, aodecq1)
+aoq1_all <- rbind.fill(aojanq1, aofebq1, aomarq1, aoaprq1, aomayq1, aojunq1, aojulq1, aoaugq1, aosepq1, aooctq1, aonovq1, aodecq1, aoalltq1)
 write.csv(aoq1_all, "../output/env/ao_q1_all.csv")
@@
 write.csv(aodecq4, "../output/env/ao_dec_q4.csv") 
+aoalltq4 <- subset(ao$X, ao$yr_t >= aoq4$yr_t)
+write.csv(aoalltq4, "../output/env/ao_yrt_q4.csv")
 
-aoq4_all <- rbind(aojanq4, aofebq4, aomarq4, aoaprq4, aomayq4, aojunq4, aojulq4, aoaugq4, aosepq4, aooctq4, aonovq4, aodecq4)
+aoq4_all <- rbind.fill(aojanq4, aofebq4, aomarq4, aoaprq4, aomayq4, aojunq4, aojulq4, aoaugq4, aosepq4, aooctq4, aonovq4, aodecq4, aoalltq4)
 write.csv(aoq4_all, "../output/env/ao_q4_all.csv")

### 2019-08-30 src/data_exploration.Rmd
laptop update

@@
 ```
+
+## density plot with background and presence env. data by month
+
+What I want to see if if there is a marked difference between the env.
+correlates of the presence points between NAFO regions.
+This is also to try deal with sampling bias (b/c the whole region is not uniformly sampled in each month, but rather nafo regions have a strong month bias)
+
+first create NAFO-region datasets
+
+```{r load pres-background dataset}
+presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+colnames(presback)
+```
+
+this dataset contains a lot of columns you don't need for the maxent models. Remove them here
+
+```{r remove unnecessary columns}
+presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt))
+colnames(presback)
+```
+
+```{r presence by month}
+
+presence <- subset(presback, occurrence == "1")
+m1 <- subset(presence, month == "1")
+m2 <- subset(presence, month == "2")
+m3 <- subset(presence, month == "3")
+m4 <- subset(presence, month == "4")
+m5 <- subset(presence, month == "5")
+m6 <- subset(presence, month == "6")
+m7 <- subset(presence, month == "7")
+m8 <- subset(presence, month == "8")
+m9 <- subset(presence, month == "9")
+m10 <- subset(presence, month == "10")
+m11 <- subset(presence, month == "11")
+m12 <- subset(presence, month == "12")
+
+```
+
+plot by each variable, less 3m (2 Sample) and 5ze (2 samples)
+```{r presence by month by variable}
+ggplot(m1, aes(x = temp_surface, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/temp_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(m1, aes(x = temp_depth, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/temp_depth_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(m1, aes(x = chl_surface, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/chl_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(m1, aes(x = chl_depth, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/chl_depth_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(m1, aes(x = salinity_surface, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/salinity_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(m1, aes(x = salinity_depth, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/salinity_depth_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(m1, aes(x = o2_surface, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/o2_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(m1, aes(x = o2_depth, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/o2_depth_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(m1, aes(x = mlp_surface, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/mlp_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(m1, aes(x = ssh_surface, colour = month)) + geom_density(na.rm = TRUE) + geom_density(data=m2, na.rm = TRUE) + geom_density(data=m3, na.rm = TRUE) + geom_density(data=m4, na.rm = TRUE) + geom_density(data=m5, na.rm = TRUE) + geom_density(data=m6, na.rm = TRUE) + geom_density(data=m7, na.rm = TRUE) + geom_density(data=m8, na.rm = TRUE) + geom_density(data=m9, na.rm = TRUE) + geom_density(data=m10, na.rm = TRUE) + geom_density(data=m11, na.rm = TRUE) + geom_density(data=m12, na.rm = TRUE)
+dev.copy(png,"../output/env/env_by_month_plots/ssh_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```
+
+```{r}
+presence$month <- factor(presence$month, order = TRUE, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
+
+ggplot(presence, aes(temp_depth, colour = month)) + geom_density()
+dev.copy(png,"../output/env/env_by_month_plots/temp_depth_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(presence, aes(salinity_depth, colour = month)) + geom_density()
+dev.copy(png,"../output/env/env_by_month_plots/salinity_depth_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(presence, aes(o2_depth, colour = month)) + geom_density()
+dev.copy(png,"../output/env/env_by_month_plots/o2_depth_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(presence, aes(chl_surface, colour = month)) + geom_density()
+dev.copy(png,"../output/env/env_by_month_plots/chl_surface_month.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+back <- subset(presback, occurrence == "0")
+
+back$month <- factor(back$month, order = TRUE, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
+
+ggplot(back, aes(temp_depth, colour = month)) + geom_density()
+dev.copy(png,"../output/env/env_by_month_plots/temp_depth_month_background.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(back, aes(salinity_depth, colour = month)) + geom_density()
+dev.copy(png,"../output/env/env_by_month_plots/salinity_depth_month_background.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(back, aes(o2_depth, colour = month)) + geom_density()
+dev.copy(png,"../output/env/env_by_month_plots/o2_depth_month_background.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+
+ggplot(back, aes(chl_surface, colour = month)) + geom_density()
+dev.copy(png,"../output/env/env_by_month_plots/chl_surface_month_background.png") # to automatically save the plot to a png AND show it inline
+dev.off() # stops automatic saving of the plot to a png
+```

### 2019-08-30 src/maxent_alldata_avec_xvalidation.Rmd
laptop update

@@
 aea <- raster("../output/env/aea.tif") #the correct CRS
-maplst <- list.files("../output/maxent/all_data/predictions/", pattern = "binary_avg_2d.*.asc", full.names = TRUE)
+maplst <- list.files("../output/maxent/all_data/predictions/apr", pattern = "binary_avg_2d.*.asc", full.names = TRUE)
 s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
 apr <- aea@crs
-#brk <- seq(0.1, 1, 0.1) #for legend
+brk <- seq(0.1, 1, 0.1) #for legend
 redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
@@
 for (pred in 1:length(maplst)){
-  pras <- raster(maplst[pred])
+  pras <- raster("../output/maxent/all_data/predictions/apr/avg_47.2119.asc")
   crs(pras) <- apr
   ascname <- names(pras)
-  pras[pras >0] <- 1
-  lp <- plot(pras, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  #pras[pras >0] <- 1
+  lp <- plot(pras, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
   lp <- plot(pras)
@@
   print(lp)
-  dev.copy(png, (paste0("../output/maxent/all_data/predictions/", ascname, pred, ".png")), width = 1920,height = 1080)
-  dev.off() # stops automatic saving of the plot to a png))
+ # dev.copy(png, (paste0("../output/maxent/all_data/predictions/", ascname, pred, ".png")), width = 1920,height = 1080)
+  #dev.off() # stops automatic saving of the plot to a png))
 }

### 2019-08-30 src/maxent_avec_xvalidation.Rmd

@@
 ```{r avg asc layers cloglog}
+mthna <- jul
+timeslice <- jul
 clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
@@
     asc <- clogasc_list[a]
-    if (grepl(dep, asc)){
+    ascd <- str_sub(asc, 50, -5)
+    if (grepl(dep, ascd)){
       depthlayers[d , paste0("a", colno)] <- asc

### 2019-08-30 src/testing_file.Rmd

@@
 }
-  rd <- rbind(rd1, rd3, rd4, rd5, rd6, rd7, rd8, rd9, rd10, rd11, rd12)
+  #rd <- rbind(rd1, rd3, rd4, rd5, rd6, rd7, rd8, rd9, rd10, rd11, rd12)
+  rd <- rbind(rd1, rd3, rd4)
+
     #rd <- as.data.frame(rd)
@@
 ```{r}
-predjan <- read.csv("../output/maxent/all_data/testforslr_jan.csv", header = TRUE)
+predjan <- read.csv("../output/maxent/all_data/pred_file_jan.csv", header = TRUE)
 predjanpr <- subset(predjan, occurrence == 1)
-plot(predjanpr$temp_depth, predjanpr$predicted)
 predjanbk <- subset(predjan, occurrence == 0)
-plot(predjanbk$temp_depth, predjanbk$predicted)
+png("janocc_temp_vs_prob.png")
+plot(predjanpr$temp_depth, predjanpr$predicted_mthdata, xlab = "temperature", ylab = "probability")
+dev.off()
+png("janbk_temp_vs_prob.png")
+plot(predjanbk$temp_depth, predjanbk$predicted_mthdata, xlab = "temperature", ylab = "probability")
+dev.off()
 ```
@@
 mont <- 
-data_aeaX <- read.csv("../output/maxent/all_data/testforslr_jan.csv", header = TRUE)
+data_aeaX <- read.csv("../output/maxent/all_data/pred_file_nov.csv", header = TRUE)
 data_aeaX$predicted_mthdata <- NA
@@
 data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aeaX, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
-avgasc_list <- list.files(paste0("../output/maxent/predictions/jan"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
+avgasc_list <- list.files(paste0("../output/maxent/predictions/nov"), pattern = "^avg_.*.asc", full.names = TRUE) #true means the full path is included
 for (j in 1:nrow(data_aea_sp)) { 
@@
 
-write.csv(data_aea1, "../output/maxent/all_data/predictions_jan.csv", row.names = FALSE)
+write.csv(data_aea1, "../output/maxent/all_data/pred_file_nov2.csv", row.names = FALSE)
  #}
@@
 ```{r}
-predjul <- read.csv("../output/maxent/all_data/predictions_jul.csv", header = TRUE)
+predjul <- read.csv("../output/maxent/all_data/pred_file_jul.csv", header = TRUE)
 plot(predjul$occurrence, predjul$predicted_mthdata)
-plot(predjul$occurrence, predjul$predicted_alldata)
+plot(predjul$occurrence, predjul$predicted_alldata, xlab = "occurrence", ylab = "probability", main = "probability for background (o) & occurrence (1) - all data model")
 ```
@@
 ```
+
+```{r}
+predlst <- list.files(paste0("../output/maxent/all_data/"), pattern = "pred_file", full.names = TRUE)
+
+completeFun <- function(data, desiredCols) {
+  completeVec <- complete.cases(data[, desiredCols])
+  return(data[completeVec, ])
+}
+
+for (a in 1:length(predlst)){
+  pred <- read.csv(predlst[a])
+  mth <- str_sub(predlst[a], 37, -5) 
+  predc <- completeFun(pred, "predicted_mthdata")
+  predoc <- subset(predc, occurrence == 1)
+  predbk <- subset(predc, occurrence == 0)
+  png(paste0("monthly_pred_occurrence_", mth, ".png"))
+  hmthoc <- hist(predoc$predicted_mthdata, main = paste0("monthly predictions - ", mth, " occurrences"), xlab = "probability of occurrence")
+  dev.off() # stops automatic saving of the plot to a png
+  png(paste0("monthly_pred_background_", mth, ".png"))
+  hmthbk <- hist(predbk$predicted_mthdata, main = paste0("monthly predictions - ", mth, " background"), xlab = "probability of occurrence")
+  dev.off() # stops automatic saving of the plot to a png            
+  png(paste0("alldata_pred_occurrence_", mth, ".png"))
+  halloc <- hist(predoc$predicted_alldata, main = paste0("all data predictions - ", mth, " occurrences"), xlab = "probability of occurrence")
+  dev.off() # stops automatic saving of the plot to a png
+  png(paste0("alldata_pred_background_", mth, ".png"))
+  hallbk <- hist(predbk$predicted_alldata, main = paste0("all data predictions - ", mth, " background"), xlab = "probability of occurrence")
+  dev.off() # stops automatic saving of the plot to a png
+
+}
+
+```
+
+tss for all data model
+```{r}
+ev <- evaluate(p = predoc$predicted_alldata , a = predbk$predicted_alldata, model = mod)
+
+tss <- max(ev@TPR + ev@TNR) - 1
+tss
+```
+
+```{r}
+janoc <- read.csv("../output/maxent/all_data/pred_file_jan.csv")
+b <- subset(janoc, occurrence == 1)
+c <- b$predicted_alldata
+a <-max(c, na.rm = TRUE)
+a
+```
+
+```{r habitat classification on probability - high only}
+
+#binary threshold function
+binprob <- function(x){
+  ifelse(x >= 0.8, 1,
+  ifelse(x < 0.8, 0, NA)) 
+}
+
+clogasc_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^avg_.*.asc", full.names = TRUE)
+
+for (cum in 1:length(clogasc_list)){
+  cumascdep <- str_sub(clogasc_list[cum], start = 47, end = -5) #selects the depth from the file name
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cum_mpathr <- calc(cumasc, fun = binprob)
+  writeRaster(cum_mpathr, paste0("../output/maxent/all_data/predictions/july/binary_p80_", cumascdep, ".asc"), overwrite = TRUE)
+  plot(cum_mpathr, png(paste0("../output/maxent/all_data/predictions/july/binary_p80_", cumascdep, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+  }
+
+```
+
+```{r binary map averaged 2D representation p80}
+
+binaryclog_list <- list.files(paste0("../output/maxent/all_data/predictions/july/"), pattern = "^binary_p80.*.asc", full.names = TRUE)
+
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, paste0("../output/maxent/all_data/predictions/july/bin_p80_2d.asc"), overwrite=TRUE)
+plot(stacksum, png(paste0("../output/maxent/all_data/predictions/july/bin_avg_p80_2d.png")))
+dev.off()
+
+```
+
+```{r}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+prpoints <- read.csv("../output/maxent/all_data/pred_file_jul.csv", header = TRUE)
+prpoints <- subset(prpoints, occurrence == 1)
+
+completeFun <- function(data, desiredCols) {
+  completeVec <- complete.cases(data[, desiredCols])
+  return(data[completeVec, ])
+  }
+
+prpoints <- completeFun(prpoints, "predicted_mthdata")
+
+pras <- raster("../output/maxent/predictions/jul/bin_p80_2d.asc")
+  crs(pras) <- aea@crs
+  ascname <- names(pras)
+  pras <- calc(pras, fun = binprob)
+
+  pras[pras >0] <- 1
+
+  #lp <- plot(pras, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(pras, axes = FALSE, box = FALSE, legend = FALSE)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  lp <- points(prpoints$longitude_meters, prpoints$latitude_meters, add = TRUE, pch=15, cex = 0.5, col = "red")
+  print(lp)
+  dev.copy(png, ("../output/maxent/all_data/julmth_over0.8.png"), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+
+```
+
+```{r habitat classification on probability - high onlyX}
+
+#binary threshold function
+binprob <- function(x){
+  ifelse(x >= 0.8, 1,
+  ifelse(x < 0.8, 0, NA)) 
+}
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/jul/"), pattern = "^avg_.*.asc", full.names = TRUE)
+
+for (cum in 1:length(clogasc_list)){
+  cumascdep <- str_sub(clogasc_list[cum], start = 47, end = -5) #selects the depth from the file name
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cum_mpathr <- calc(cumasc, fun = binprob)
+  writeRaster(cum_mpathr, paste0("../output/maxent/predictions/jul/binary_p80_", cumascdep, ".asc"), overwrite = TRUE)
+  plot(cum_mpathr, png(paste0("../output/maxent/predictions/jul/binary_p80_", cumascdep, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+  }
+
+```
+
+```{r binary map averaged 2D representation p80x}
+
+binaryclog_list <- list.files(paste0("../output/maxent/predictions/jul/"), pattern = "^binary_p80.*.asc", full.names = TRUE)
+
+binstack <- stack(binaryclog_list)
+stacksum <- sum(binstack, na.rm = TRUE)
+writeRaster(stacksum, paste0("../output/maxent/predictions/jul/bin_p80_2d.asc"), overwrite=TRUE)
+plot(stacksum, png(paste0("../output/maxent/predictions/jul/bin_avg_p80_2d.png")))
+dev.off()
+
+```
+
+add land to plots with standard legend (working code) THIS WORK!!!
+```{r}
+aea <- raster("../output/env/aea.tif") #the correct CRS
+#maplst <- list.files("../output/maxent/all_data/predictions/", pattern = "binary_avg_2d.*.asc", full.names = TRUE)
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+brk <- seq(0.1, 1, 0.1) #for legend
+redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+
+#r <- raster(avg_77.6112.asc)
+#plot(pras)
+
+#for (pred in 1:length(maplst)){
+  pras <- raster("../output/maxent/all_data/predictions/apr/avg_77.6112.asc")
+  plot(pras)
+  crs(pras) <- apr
+  ascname <- names(pras)
+  #pras[pras >0] <- 1
+  lp <- plot(pras, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
+  lp <- plot(pras)
+  lp <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+  print(lp)
+  dev.copy(png, (paste0("../output/maxent/all_data/predictions/", ascname, pred, ".png")), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+#}
+
+```
+
+barcharts showing probability distribution at depth
+
+first bin the probabilities for cells inside the mcp
+```{r bin probabilities}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+clogasc_list <- list.files(paste0("../output/maxent/predictions/nov/"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
+
+binprob <- function(x){
+  ifelse(x > 0 & x < 0.1, 0.1,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_1_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_1_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.1 & x < 0.2, 0.2,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_2_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_2_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.2 & x < 0.3, 0.3,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_3_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_3_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.3 & x < 0.4, 0.4,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_4_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_4_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.4 & x < 0.5, 0.5,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_5_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_5_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.5 & x < 0.6, 0.6,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_6_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_6_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.6 & x < 0.7, 0.7,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_7_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_7_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.7 & x < 0.8, 0.8,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_8_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_8_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.8 & x < 0.9, 0.9,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_9_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_9_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.9 & x <= 1, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/predictions/nov/binned_10_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/predictions/nov/binned_10_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+```
+
+get data to plot probability by depth... need to do manually 1-10
+```{r}
+ascdepth_list_high <- list.files(paste0("../output/maxent/predictions/nov/"), pattern = "^binned_1_avgprob_.*asc", full.names = TRUE)
+
+binprob <- function(x){
+  ifelse(x >0,  1,
+  ifelse(x == -9999, 0, NA))
+}
+
+binaryavgstack <- stack(ascdepth_list_high)
+stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+stackstathigh[ ,1] <- c(1:49)
+colnames(stackstathigh) <- c("asclayer", "prob_bin")
+
+for (lyr in 1:nlayers(binaryavgstack)){
+  ascname <-  str_sub(ascdepth_list_high[lyr], 63, -5)
+  dlay <- binaryavgstack[[lyr]]
+  dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
+  lyrsum <- cellStats(dlay, sum)
+  stackstathigh$prob_bin[lyr] <- lyrsum
+  stackstathigh$asclayer[lyr] <- ascname
+
+}
+write.csv(stackstathigh, paste0("../output/maxent/predictions/nov/probbin_1_cellcount_bydepth_avg.csv"), row.names = FALSE)
+
+```
+
+```{r}
+
+probblst <- list.files(paste0("../output/maxent/predictions/nov"), pattern = "^probbin_.*.csv", full.names = TRUE)
+a <- read.csv(probblst[1])
+b <- read.csv(probblst[2])
+c <- read.csv(probblst[3])
+d <- read.csv(probblst[4])
+e <- read.csv(probblst[5])
+f <- read.csv(probblst[6])
+g <- read.csv(probblst[7])
+h <- read.csv(probblst[8])
+i <- read.csv(probblst[9])
+j <- read.csv(probblst[10])
+
+stackstat <- cbind(a, b[ ,2], c[ ,2], d[ ,2], e[ , 2], f[ ,2], g[ ,2], h[ ,2], i[ ,2], j[ ,2])
+colnames(stackstat) <- c("depth", "0 - 0.09", "0.1 - 0.19", "0.2 - 0.29", "0.3 - 0.39", "0.4 - 0.49", "0.5 - 0.59", "0.6 - 0.69", "0.7 - 0.79", "0.8 - 0.89", ">0.9")
+stackstat <- melt(stackstat, id = "depth")
+stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
+stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
+stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]
+
+colnames(stackstat) <- c("Depth", "Probability", "Cells")
+
+write.csv(stackstat, paste0("../output/maxent/predictions/nov/allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
+
+```
+
+count number of cells that are not NA in each raster layer that are inside the mcp
+
+```{r}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+clogasc_list <- list.files(paste0("../output/maxent/predictions/nov/"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
+cumst <- data.frame(0, 0)
+colnames(cumst) <- c("depth", "no_cells")
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumst[cum, 1] <- str_sub(ascname, 13)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  ccnt <- freq(cumasc, useNA='no')
+  ccnt <- sum(ccnt[ ,2])
+  cumst[cum, 2] <- ccnt
+}
+
+cumst$depth <- as.numeric(cumst$depth)
+cumst1 <- cumst[order(cumst$no_cells), ]
+
+write.csv(cumst1, "../output/maxent/predictions/nov/cellsperlayer.csv", row.names = FALSE)
+
+```
+
+think about how to get proportions...
+
+```{r}
+pr <- read.csv("../output/maxent/predictions/nov/allprobbin_cellcount_bydepth_avg.csv", header = TRUE)
+cumst1 <- read.csv("../output/maxent/predictions/nov/cellsperlayer.csv", header = TRUE)
+
+#order both filed from high to low depth
+pr <- pr[order(pr$Depth), ]
+cumst1 <- cumst[order(cumst$depth), ]
+
+#so there are some differences in the depth numbers... use partial matching to get the number of cells on each layer
+
+a <- pr
+for (r in 1:nrow(pr)){
+  d1 <- str_sub(pr$Depth[r], end = 4)
+  for (ro in 1:nrow(cumst1)){
+    d2 <- str_sub(cumst1$depth[ro], end = 4)
+    if (d1 == d2){
+        a$total_cells[r] <- cumst1$no_cells[ro]
+    }
+  }
+}
+
+# now work out percentage of cells for each probability
+
+for (r in 1:nrow(a)){
+  a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
+}
+
+write.csv(a, "../output/maxent/predictions/nov/dis_prob_cells.csv", row.names = FALSE)
+
+```
+
+histogram of probability by depth (straight count)
+
+```{r}
+mthna = "nov"
+bydepth <- ggplot(data = a, aes(x = Depth, y = Cells, fill = Probability)) + geom_bar(stat = "identity") +   scale_x_reverse() + coord_flip()
+png(paste0("../output/maxent/predictions/", mthna, "/depthplot_probavg.png"))
+print(bydepth)
+dev.off()
+```
+
+histogram of probability by depth - by proportion
+
+```{r}
+
+bydepth <- ggplot(a, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() #XXXXXcareful - the deeper layers aren't being displayed correctly here
+plot(bydepth)
+png(paste0("../output/maxent/predictions/", mthna, "/depthplot_probavg_prop.png"))
+print(bydepth)
+dev.off()
+```
+
+ok lets try get an probability for the whole month
+
+```{r}
+a <- read.csv("../output/maxent/predictions/jul/dis_prob_cells.csv", header = TRUE)
+
+#subset the probability ranges
+p1 <- subset(a, Probability == "0 - 0.09")
+p2 <- subset(a, Probability == "0.1 - 0.19")
+p3 <- subset(a, Probability == "0.2 - 0.29")
+p4 <- subset(a, Probability == "0.3 - 0.39")
+p5 <- subset(a, Probability == "0.4 - 0.49")
+p6 <- subset(a, Probability == "0.5 - 0.59")
+p7 <- subset(a, Probability == "0.6 - 0.69")
+p8 <- subset(a, Probability == "0.7 - 0.79")
+p9 <- subset(a, Probability == "0.8 - 0.89")
+p10 <- subset(a, Probability == ">0.9")
+
+#count the number of cells per probability.
+
+cp1 <- sum(p1$Cells)
+cp2 <- sum(p2$Cells)
+cp3 <- sum(p3$Cells)
+cp4 <- sum(p4$Cells)
+cp5 <- sum(p5$Cells)
+cp6 <- sum(p6$Cells)
+cp7 <- sum(p7$Cells)
+cp8 <- sum(p8$Cells)
+cp9 <- sum(p9$Cells)
+cp10 <- sum(p10$Cells)
+
+tc2 <- sum(cp1, cp2, cp3, cp4, cp5, cp6, cp7, cp8, cp9, cp10) #how many cells in total for novy have a value
+tcell <- sum(p1$total_cells) # how many cells are available to have a value...
+#tc2 and tcell should match
+
+#create datbase of total cells per probability bin for the month
+
+b <- unique(a$Probability) #get the unique probability bins
+c <- as.data.frame(rbind(cp1, cp2,  cp3, cp4, cp5, cp6, cp7, cp8, cp9, cp10))
+c$probability <- b
+colnames(c) <- c("cells", "prob_bin")
+rownames(c) <- c()
+c$month <- "jul"
+```
+
+make plot of probability over the month
+```{r}
+pmth <- ggplot(d, aes(month, cells, fill = prob_bin)) + geom_bar(stat='identity') 
+plot(pmth)
+png(paste0("../output/maxent/predictions/", mthna, "/prob_avg_mth.png"))
+print(pmth)
+dev.off()
+```
+
+merge two months of data together into one file then plot
+
+```{r}
+f <- rbind(c, d)
+pmth <- ggplot(f, aes(month, cells, fill = prob_bin)) + geom_bar(stat='identity') 
+plot(pmth)
+png(paste0("../output/maxent/predictions/prob_avg_mth_test.png"))
+print(pmth)
+dev.off()
+```
+
+boxplot of test auc scores test
+
+```{r}
+dec <- read.csv("../output/maxent/models/decavgAIC.csv", header = TRUE)
+mar <- read.csv("../output/maxent/models/maravgAIC.csv", header = TRUE)
+apr <- read.csv("../output/maxent/models/apravgAIC.csv", header = TRUE)
+may <- read.csv("../output/maxent/models/mayavgAIC.csv", header = TRUE)
+jun <- read.csv("../output/maxent/models/junavgAIC.csv", header = TRUE)
+jul <- read.csv("../output/maxent/models/julavgAIC.csv", header = TRUE)
+aug <- read.csv("../output/maxent/models/augavgAIC.csv", header = TRUE)
+sep <- read.csv("../output/maxent/models/sepavgAIC.csv", header = TRUE)
+oct <- read.csv("../output/maxent/models/octavgAIC.csv", header = TRUE)
+nov <- read.csv("../output/maxent/models/novavgAIC.csv", header = TRUE)
+
+mar <- mar$test_AUC
+mar <- head(mar, -1)
+apr <- apr$test_AUC
+apr <- head(apr, -1)
+may <- may$test_AUC
+may <- head(may, -1)
+jun <- jun$test_AUC
+jun <- head(jun, -1)
+jul <- jul$test_AUC
+jul <- head(jul, -1)
+aug <- aug$test_AUC
+aug <- head(aug, -1)
+sep <- sep$test_AUC
+sep <- head(sep, -1)
+oct <- oct$test_AUC
+oct <- head(oct, -1)
+nov <- nov$test_AUC
+nov <- head(nov, -1)
+dec <- dec$test_AUC
+dec <- head(dec, -1)
+
+n <- max(length(mar), length(apr), length(may), length(jun), length(jul), length(aug), length(sep), length(oct), length(nov), length(dec))
+length(mar) <- n                      
+length(apr) <- n
+length(may) <- n
+length(jun) <- n
+length(jul) <- n
+length(aug) <- n
+length(sep) <- n
+length(oct) <- n
+length(nov) <- n
+length(dec) <- n
+
+testauc <- cbind(mar, apr, may, jun, jul, aug, sep, oct, nov, dec)
+colnames(c) <-c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
+
+```
+
+boxplot of train auc scores test
+
+```{r}
+dec <- read.csv("../output/maxent/models/decavgAIC.csv", header = TRUE)
+mar <- read.csv("../output/maxent/models/maravgAIC.csv", header = TRUE)
+apr <- read.csv("../output/maxent/models/apravgAIC.csv", header = TRUE)
+may <- read.csv("../output/maxent/models/mayavgAIC.csv", header = TRUE)
+jun <- read.csv("../output/maxent/models/junavgAIC.csv", header = TRUE)
+jul <- read.csv("../output/maxent/models/julavgAIC.csv", header = TRUE)
+aug <- read.csv("../output/maxent/models/augavgAIC.csv", header = TRUE)
+sep <- read.csv("../output/maxent/models/sepavgAIC.csv", header = TRUE)
+oct <- read.csv("../output/maxent/models/octavgAIC.csv", header = TRUE)
+nov <- read.csv("../output/maxent/models/novavgAIC.csv", header = TRUE)
+
+mar <- mar$train_AUC
+mar <- head(mar, -1)
+apr <- apr$train_AUC
+apr <- head(apr, -1)
+may <- may$train_AUC
+may <- head(may, -1)
+jun <- jun$train_AUC
+jun <- head(jun, -1)
+jul <- jul$train_AUC
+jul <- head(jul, -1)
+aug <- aug$train_AUC
+aug <- head(aug, -1)
+sep <- sep$train_AUC
+sep <- head(sep, -1)
+oct <- oct$train_AUC
+oct <- head(oct, -1)
+nov <- nov$train_AUC
+nov <- head(nov, -1)
+dec <- dec$train_AUC
+dec <- head(dec, -1)
+
+n <- max(length(mar), length(apr), length(may), length(jun), length(jul), length(aug), length(sep), length(oct), length(nov), length(dec))
+length(mar) <- n                      
+length(apr) <- n
+length(may) <- n
+length(jun) <- n
+length(jul) <- n
+length(aug) <- n
+length(sep) <- n
+length(oct) <- n
+length(nov) <- n
+length(dec) <- n
+
+trainauc <- cbind(mar, apr, may, jun, jul, aug, sep, oct, nov, dec)
+colnames(c) <-c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
+```
+
+boxplot of cbi
+
+```{r}
+dec <- read.csv("../output/maxent/models/decavgAIC.csv", header = TRUE)
+mar <- read.csv("../output/maxent/models/maravgAIC.csv", header = TRUE)
+apr <- read.csv("../output/maxent/models/apravgAIC.csv", header = TRUE)
+may <- read.csv("../output/maxent/models/mayavgAIC.csv", header = TRUE)
+jun <- read.csv("../output/maxent/models/junavgAIC.csv", header = TRUE)
+jul <- read.csv("../output/maxent/models/julavgAIC.csv", header = TRUE)
+aug <- read.csv("../output/maxent/models/augavgAIC.csv", header = TRUE)
+sep <- read.csv("../output/maxent/models/sepavgAIC.csv", header = TRUE)
+oct <- read.csv("../output/maxent/models/octavgAIC.csv", header = TRUE)
+nov <- read.csv("../output/maxent/models/novavgAIC.csv", header = TRUE)
+
+mar <- mar$cbi
+mar <- head(mar, -1)
+apr <- apr$cbi
+apr <- head(apr, -1)
+may <- may$cbi
+may <- head(may, -1)
+jun <- jun$cbi
+jun <- head(jun, -1)
+jul <- jul$cbi
+jul <- head(jul, -1)
+aug <- aug$cbi
+aug <- head(aug, -1)
+sep <- sep$cbi
+sep <- head(sep, -1)
+oct <- oct$cbi
+oct <- head(oct, -1)
+nov <- nov$cbi
+nov <- head(nov, -1)
+dec <- dec$cbi
+dec <- head(dec, -1)
+
+n <- max(length(mar), length(apr), length(may), length(jun), length(jul), length(aug), length(sep), length(oct), length(nov), length(dec))
+length(mar) <- n                      
+length(apr) <- n
+length(may) <- n
+length(jun) <- n
+length(jul) <- n
+length(aug) <- n
+length(sep) <- n
+length(oct) <- n
+length(nov) <- n
+length(dec) <- n
+
+cbi <- cbind(mar, apr, may, jun, jul, aug, sep, oct, nov, dec)
+colnames(c) <-c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
+```
+
+boxplots of tss scores
+```{r}
+dec <- read.csv("../output/maxent/models/decavgAIC.csv", header = TRUE)
+mar <- read.csv("../output/maxent/models/maravgAIC.csv", header = TRUE)
+apr <- read.csv("../output/maxent/models/apravgAIC.csv", header = TRUE)
+may <- read.csv("../output/maxent/models/mayavgAIC.csv", header = TRUE)
+jun <- read.csv("../output/maxent/models/junavgAIC.csv", header = TRUE)
+jul <- read.csv("../output/maxent/models/julavgAIC.csv", header = TRUE)
+aug <- read.csv("../output/maxent/models/augavgAIC.csv", header = TRUE)
+sep <- read.csv("../output/maxent/models/sepavgAIC.csv", header = TRUE)
+oct <- read.csv("../output/maxent/models/octavgAIC.csv", header = TRUE)
+nov <- read.csv("../output/maxent/models/novavgAIC.csv", header = TRUE)
+
+mar <- mar$tss
+mar <- head(mar, -1)
+apr <- apr$tss
+apr <- head(apr, -1)
+may <- may$tss
+may <- head(may, -1)
+jun <- jun$tss
+jun <- head(jun, -1)
+jul <- jul$tss
+jul <- head(jul, -1)
+aug <- aug$tss
+aug <- head(aug, -1)
+sep <- sep$tss
+sep <- head(sep, -1)
+oct <- oct$tss
+oct <- head(oct, -1)
+nov <- nov$tss
+nov <- head(nov, -1)
+dec <- dec$tss
+dec <- head(dec, -1)
+
+n <- max(length(mar), length(apr), length(may), length(jun), length(jul), length(aug), length(sep), length(oct), length(nov), length(dec))
+length(mar) <- n                      
+length(apr) <- n
+length(may) <- n
+length(jun) <- n
+length(jul) <- n
+length(aug) <- n
+length(sep) <- n
+length(oct) <- n
+length(nov) <- n
+length(dec) <- n
+
+tss <- cbind(mar, apr, may, jun, jul, aug, sep, oct, nov, dec)
+colnames(c) <-c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
+```
+
+```{r}
+par(mfrow=c(2,2))
+
+boxplot(testauc, xlab = "Month", ylab = "Test AUC Score")
+boxplot(trainauc, xlab = "Month", ylab = "Training AUC Score")
+boxplot(cbi, xlab = "Month", ylab = "CBI Score")
+boxplot(tss, xlab = "Month", ylab = "TSS Score")
+
+```
+
+something went wrong with July aug and sep when averaging the rasters.... redo (error due to layer 120 and coconfusion with the year 2012)
+
+```{r avg asc layers cloglog}
+
+tlst <- c("sep1998", "sep1999", "sep2000", "sep2001", "sep2002", "sep2003", "sep2004", "sep2005", "sep2006", "sep2007", "sep2008", "sep2009", "sep2010", "sep2011", "sep2012", "sep2013", "sep2014", "sep2015")
+
+mthna <- "sep"
+
+for(a in length(tlst)){
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+timeslice <- tlst[1]
+stdep <- list.files(paste0("../output/maxent/models/", timeslice),  pattern = "^stackdepthlist.*.csv", full.names = TRUE)
+stdep <- read.csv(stdep)
+depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
+names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
+depthlayers$depth <- as.character(depthlayers$depth)
+
+colno <- 1
+
+for (d in 1:nrow(depthlayers)){
+  dep <- depthlayers$depth[d]
+  for (a in 1:length(clogasc_list)){
+    asc <- clogasc_list[a]
+    ascd <- str_sub(asc, 50, -5)
+    if (grepl(dep, ascd)){
+      depthlayers[d , paste0("a", colno)] <- asc
+       colno <- colno+1
+    }
+  }
+}
+write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)
+
+}
+
+for (x in 1:nrow(depthlayers)){
+  asclay <- depthlayers[x, ]
+  asclay <- asclay[!is.na(asclay)]
+  asclay <- as.data.frame(rbind(asclay))
+  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
+  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
+  asclay$V3 <- as.character(asclay$V3)
+  asclay$V4 <- as.character(asclay$V4)
+  asclay$V5 <- as.character(asclay$V5)
+  asclay$V6 <- as.character(asclay$V6)
+  asclay$V7 <- as.character(asclay$V7)
+  asclay$V8 <- as.character(asclay$V8)
+  asclay$V9 <- as.character(asclay$V9)
+  asclay$V10 <- as.character(asclay$V10)
+  asclay$V11 <- as.character(asclay$V11)
+  asclay$V12 <- as.character(asclay$V12)
+  asclay$V13 <- as.character(asclay$V13)
+  asclay$V14 <- as.character(asclay$V14)
+  asclay$V15 <- as.character(asclay$V15)  
+  asclay$V16 <- as.character(asclay$V16)  
+  asclay$V17 <- as.character(asclay$V17) 
+  asclay$V18 <- as.character(asclay$V18)  
+  asclay$V19 <- as.character(asclay$V19)
+
+  asclaydep <- asclay$V1
+
+  asc1998 <- raster(asclay$V2)
+  asc1999 <- raster(asclay$V3)
+  asc2000 <- raster(asclay$V4)
+  asc2001 <- raster(asclay$V5)
+  asc2002 <- raster(asclay$V6)
+  asc2003 <- raster(asclay$V7)
+  asc2004 <- raster(asclay$V8)
+  asc2005 <- raster(asclay$V9)
+  asc2006 <- raster(asclay$V10)
+  asc2007 <- raster(asclay$V11)
+  asc2008 <- raster(asclay$V12)
+  asc2009 <- raster(asclay$V13)
+  asc2010 <- raster(asclay$V14)
+  asc2011 <- raster(asclay$V15)
+  asc2012 <- raster(asclay$V16)
+  asc2013 <- raster(asclay$V17)
+  asc2014 <- raster(asclay$V18)
+  asc2015 <- raster(asclay$V19)
+
+  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011, asc2012, asc2013, asc2014, asc2015)
+
+  avg <- mean(stkasc, na.rm = TRUE)
+
+  writeRaster(avg, paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".asc"), overwrite=TRUE)
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".png"))) 
+  dev.off()
+
+}
+
+```
+
+Threshold tables
+```{r}
+thlst1 <- list.files(paste0("../output/maxent/models"), pattern = ".*avgAIC.csv", full.names = TRUE)
+thlst2 <- list.files(paste0("../output/maxent/models"), pattern = ".*avgmaxentresults.csv", full.names = TRUE)
+
+for (m in 1:length(thlst1)){
+    mth <- thlst1[m]
+    mth <- str_sub(mth, 25, -11) #extract first 3 digit
+    t1 <- read.csv(thlst1[m], header = TRUE)
+    t1 <- t1[-nrow(t1), -1 ]
+    t2 <- read.csv(thlst2[m], header = TRUE)
+    t2 <- t2[-nrow(t2), -1]
+    a <- subset(t1, select = c("test_maxsumsenspec", "test_equal_sens_spec", "mpa_0.9_clog"))
+    b <- subset(t2, select = c("Minimum.training.presence.Cloglog.threshold", "X10.percentile.training.presence.Cloglog.threshold"))
+    c <- cbind(a, b)
+    colnames(c) <- c("MaxSS", "ESS", "MPA", "MTP", "P10TP")
+    d <- summary(c)
+    write.csv(t(as.matrix(d)), file= paste0("../output/maxent/models/summary_test_", mth, ".csv")) 
+}
+
+thlst <- list()
+for (m in 1:length(thlst1)){
+    mth <- thlst1[m]
+    mth <- str_sub(mth, 25, -11) #extract first 3 digit
+    t1 <- read.csv(thlst1[m], header = TRUE)
+    t1 <- t1[-nrow(t1), -1 ]
+    t2 <- read.csv(thlst2[m], header = TRUE)
+    t2 <- t2[-nrow(t2), -1]
+    a <- subset(t1, select = c("test_maxsumsenspec", "test_equal_sens_spec", "mpa_0.9_clog"))
+    b <- subset(t2, select = c("Minimum.training.presence.Cloglog.threshold", "X10.percentile.training.presence.Cloglog.threshold"))
+    ca <- cbind(a, b)
+    colnames(ca) <- c("MaxSS", "ESS", "MPA", "MTP", "P10")
+    ca$month <- mth
+
+    thlst[[m]] <- ca
+  }
+
+stdep <- do.call(rbind, thlst)
+stdep <- stdep[!(stdep$month == "jan"), ] #removes jan because you're not doing jan
+
+p10p <- ggplot(data=stdep, aes(x=month, y=P10, fill=month)) + geom_boxplot() + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=3) + theme(legend.position = "none")
+mtpp <- ggplot(data=stdep, aes(x=month, y=MTP, fill=month)) + geom_boxplot() + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=3) + theme(legend.position = "none")
+plot(l)
+mpap <- ggplot(data=stdep, aes(x=month, y=MPA, fill=month)) + geom_boxplot() + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=3) + theme(legend.position = "none")
+essp <- ggplot(data=stdep, aes(x=month, y=ESS, fill=month)) + geom_boxplot() + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=3) + theme(legend.position = "none")
+maxss <- ggplot(data=stdep, aes(x=month, y=MaxSS, fill=month)) + geom_boxplot() + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=3) + theme(legend.position = "none")
+
+plot_grid(p10p, mtpp, mpap, essp, maxss, labels = "AUTO")
+
+library(reshape2)
+k <- melt(stdep, id.var = c('month'), variable.name = 'threshold')
+n <- subset(k, month == "aug")
+o <- subset(k, month == "mar")
+
+library(cowplot)
+p <-ggplot(data=n, aes(x=threshold, y=value, fill=threshold)) + geom_boxplot() + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=3) + theme(legend.position = "none")
+q <-ggplot(data=o, aes(x=threshold, y=value, fill=threshold)) + geom_boxplot() + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=3) + theme(legend.position = "none")
+plot_grid(p, q, labels = "AUTO")
+
+tsum <- stdep %>% # Start by defining the original dataframe, AND THEN...
+                group_by(month) %>% # Define the grouping variable, AND THEN...
+                summarise( # Now you define your summary variables with a name and a function...
+                          maxss.mean = mean(MaxSS), mtp.mean = mean(MTP), ess.mean = mean(ESS), mpa.mean = mean(MPA), mean.p10 = mean(P10))
+
+tsum
+```
+
+stacked maps
+https://urbandemographics.blogspot.com/2016/04/creating-tilted-and-stacked-maps-in-r.html
+
+```{r}
+# load libraries
+  library(rgeos)
+  library(UScensus2000tract)
+  library(ggplot2)
+  library(dplyr)
+  library(RColorBrewer)
+
+# load data
+  data("oregon.tract")
+
+# plot Census Tract map
+  plot(oregon.tract)
+
+oregon.tract$id=as.character(1:nrow(oregon.tract))
+
+clogasc_list <- list.files(paste0("../output/maxent/all_data/predictions/july"), pattern = "^avg_.*.asc", full.names = TRUE) # first list the files containing the data
+a <- stack(clogasc_list) #stack the files
+
+```
+ok scrap tilt just do standard
+
+need to add seabed
+
+```{r seabed files}
+slst <- list.files("../data/env/ascii/ascii2010", pattern = "temp", full.names = TRUE)
+sr <- raster(slst[1])
+plot(sr)
+```
+
+```{r}
+aea <- raster("../output/env/aea.tif") #the correct CRS
+maplst <- list.files("../output/maxent/predictions/may", pattern = "^avg_clog.*.asc", full.names = TRUE)
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+brk <- seq(0.1, 1, 0.1) #for legend
+redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+
+#r <- raster(avg_77.6112.asc)
+#plot(pras)
+
+#for (pred in 1:length(maplst)){
+  pras5 <- raster(maplst[33])
+  pras180 <- raster(maplst[14])
+  pras370 <- raster(maplst[28])
+  pras697 <- raster(maplst[41])
+  crs(pras5) <- apr
+  crs(pras180) <- apr
+  crs(pras370) <- apr
+  crs(pras697) <- apr
+  ascname5 <- names(pras5)
+  ascname180 <- names(pras180)
+  ascname370 <- names(pras370)
+  ascname697 <- names(pras697)
+
+par(mfrow=c(1,2))
+
+  #pras[pras >0] <- 1
+  lp5 <- plot(pras5, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
+  #lp <- plot(pras5)
+  lp5 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp5 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+
+  l180 <- plot(pras180, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
+  #lp <- plot(pras5)
+  lp180 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp180 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+
+  print(lp5)
+  dev.copy(png, (paste0("../output/maxent/all_data/predictions/", ascname5, pred, ".png")), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+#}
+
+  print(lp5)
+  dev.copy(png, (paste0("../output/maxent/all_data/predictions/", ascname5, pred, ".png")), width = 1920,height = 1080)
+  dev.off() # stops automatic saving of the plot to a png))
+#}
+
+```

### 2019-09-03 src/testing_file.Rmd
rastervis testing

@@
 apr <- aea@crs
-brk <- seq(0.1, 1, 0.1) #for legend
-redtrans <- rgb(218, 215, 216, 90, maxColorValue=255) 
+brk <- seq(0, 1, 0.1) #for legend
+lgc <- sequential_hcl(10, palette = "YlOrRd") #colour scheme from colorspace package
+redtrans <- rgb(218, 215, 216, 120, maxColorValue=255) #mcp
 outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
 
-#r <- raster(avg_77.6112.asc)
-#plot(pras)
-
-#for (pred in 1:length(maplst)){
-  pras5 <- raster(maplst[33])
+  pras6 <- raster(maplst[37])
   pras180 <- raster(maplst[14])
@@
   pras697 <- raster(maplst[41])
-  crs(pras5) <- apr
+  crs(pras6) <- apr
   crs(pras180) <- apr
@@
   crs(pras697) <- apr
-  ascname5 <- names(pras5)
+  ascname6 <- names(pras6)
   ascname180 <- names(pras180)
@@
 
-par(mfrow=c(1,2))
-
-  #pras[pras >0] <- 1
-  lp5 <- plot(pras5, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
-  #lp <- plot(pras5)
-  lp5 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
-  lp5 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
-
-  l180 <- plot(pras180, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
-  #lp <- plot(pras5)
+par(mfrow=c(1,4))
+
+  lp6 <- plot(pras6, breaks=brk,  col=rev(lgc), axes = FALSE, box = FALSE, legend = FALSE)
+  lp6 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp6 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+
+  lp180 <- plot(pras180, breaks=brk,  col=rev(lgc), axes = FALSE, box = FALSE, legend = FALSE)
   lp180 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
@@
 
-  print(lp5)
-  dev.copy(png, (paste0("../output/maxent/all_data/predictions/", ascname5, pred, ".png")), width = 1920,height = 1080)
+  lp370 <- plot(pras370, breaks=brk,  col=rev(lgc), axes = FALSE, box = FALSE, legend = FALSE)
+  lp370 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp370 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+
+  lp697 <- plot(pras697, breaks=brk,  col=rev(lgc), axes = FALSE, box = FALSE, legend = TRUE)
+  lp697 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp697 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+
+  print(lp6)
+  dev.copy(png, (paste0("../output/maxent/all_data/predictions/testplots/", ascname6, ".png")), width = 1920,height = 1080)
   dev.off() # stops automatic saving of the plot to a png))
@@
   print(lp5)
-  dev.copy(png, (paste0("../output/maxent/all_data/predictions/", ascname5, pred, ".png")), width = 1920,height = 1080)
+  dev.copy(png, (paste0("../output/maxent/all_data/predictions/", ascname5, ".png")), width = 1920,height = 1080)
   dev.off() # stops automatic saving of the plot to a png))
@@
 ```
+
+as above but with rcolorbrewer
+```{r}
+lp6 <- plot(pras6, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
+  lp6 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp6 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+
+  lp180 <- plot(pras180, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
+  lp180 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp180 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+
+  lp370 <- plot(pras370, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
+  lp370 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp370 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+
+  lp697 <- plot(pras697, breaks=brk, col=rev(colorRampPalette(brewer.pal(10, "Spectral"))(10)), axes = FALSE, box = FALSE, legend = TRUE)
+  lp697 <- plot(outsidemcp,  col = redtrans, border = redtrans, add = TRUE)
+  lp697 <- plot(s, col = "#474646", border = "#474646", add = TRUE)
+```
+
+with rastervis
+
+```{r}
+aea <- raster("../output/env/aea.tif") #the correct CRS
+maplst <- list.files("../output/maxent/predictions/may", pattern = "^avg_clog.*.asc", full.names = TRUE)
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+brk <- seq(0, 1, 0.1) #for legend
+lgc <- sequential_hcl(10, palette = "YlOrRd") #colour scheme from colorspace package
+redtrans <- rgb(218, 215, 216, 120, maxColorValue=255) #mcp
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+
+pras6 <- raster(maplst[37])
+pras180 <- raster(maplst[14])
+pras370 <- raster(maplst[28])
+pras697 <- raster(maplst[41])
+crs(pras6) <- apr
+crs(pras180) <- apr
+crs(pras370) <- apr
+crs(pras697) <- apr
+ascname6 <- names(pras6)
+ascname180 <- names(pras180)
+ascname370 <- names(pras370)
+ascname697 <- names(pras697)
+pstack <- stack(pras6, pras180, pras370, pras697) #stack the layers
+
+pras1 <- raster(maplst[1])
+pras2 <- raster(maplst[2])
+pras3 <- raster(maplst[3])
+pras4 <- raster(maplst[4])
+crs(pras1) <- apr
+crs(pras2) <- apr
+crs(pras3) <- apr
+crs(pras4) <- apr
+
+pstack <- stack(pras6, pras180, pras370, pras697, pras1, pras2, pras3, pras4) #stack the layers
+
+cls <- c("6m", "180m", "370m", "697m")
+rws <- c("may", "dummy")
+mapTheme <- rasterTheme(region = rev(brewer.pal(10,"RdYlGn")))
+my.at <- seq(0, 1, by = 0.1)
+
+p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(4, 2), names.attr = c("6m", "180m", "370m", "697m", "", "", "", ""))
+p + layer(sp.polygons(s, fill = "black")) + layer (sp.polygons(outsidemcp, fill = redtrans, col = NA))
+
+png("../output/maxent/predictions/testplots/rasterVis.png", height = 1024 * 0.707, width = 1024)
+p
+dev.off()
+
+```

### 2019-09-04 src/testing_file.Rmd
plot export fixed, now working on hovmoller plot (data restructuring)

@@
 ```{r}
+library(raster)
+library(rgdal)
+library(rasterVis)
+library(colorspace)
+
 aea <- raster("../output/env/aea.tif") #the correct CRS
@@
 
-p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(4, 2), names.attr = c("6m", "180m", "370m", "697m", "", "", "", ""))
+p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(4, 2), names.attr = c("6m", "180m", "370m", "697m", "", "", "", ""), margin = FALSE)
 p + layer(sp.polygons(s, fill = "black")) + layer (sp.polygons(outsidemcp, fill = redtrans, col = NA))
 
-png("../output/maxent/predictions/testplots/rasterVis.png", height = 1024 * 0.707, width = 1024)
-p
-dev.off()
-
-```
+dev.copy(png, (paste0("../output/maxent/predictions/testplots/rasterVis.png")), width = 1920,height = 1080)
+dev.off() # stops automatic saving of the plot to a png))
+
+```
+
+```{r}
+histogram(pstack)
+densityplot(pstack)
+bwplot(pstack)
+```
+
+try hovmoller plotting for depth time
+
+```{r}
+#think i need to create a 4d raster with depth and time...
+library(ncdf4)
+
+writeRaster(pstack, "../output/maxent/predictions/testplots/pstack.nc", overwrite = TRUE, format = "CDF", varname = "prob", varunit = "p", longname = "dummy netcdf for hovmollers", xname= "X", yname = "Y", zname = "date", zunit = "date")
+
+```
+
+```{r}
+pras6 <- raster::setZ(pras6, as.Date("0000-1-1"), name = "time")
+pras180 <- raster::setZ(pras180, as.Date("0000-1-1"), name = "time")
+pras370 <- raster::setZ(pras370, as.Date("0000-1-1"), name = "time")
+pras697 <- raster::setZ(pras697, as.Date("0000-1-1"), name = "time")
+pras1 <- raster::setZ(pras1, as.Date("0000-2-1"), name = "time")
+pras2 <- raster::setZ(pras2, as.Date("0000-2-1"), name = "time")
+pras3 <- raster::setZ(pras3, as.Date("0000-2-1"), name = "time")
+pras4 <- raster::setZ(pras4, as.Date("0000-2-1"), name = "time")
+```
+
+```{r}
+library(zoo)
+
+old <- setwd(tempdir())
+download.file('https://raw.github.com/oscarperpinan/spacetime-vis/master/data/SISmm2008_CMSAF.zip',
+'SISmm2008_CMSAF.zip', method='wget')
+unzip('SISmm2008_CMSAF.zip')
+listFich <- dir(pattern='\\.nc')
+stackSIS <- stack(listFich)
+stackSIS <- stackSIS*24 ##from irradiance (W/m2) to irradiation Wh/m2
+
+pbrick <- brick(pras6, pras180, pras370, pras697, pras1, pras2, pras3, pras4) #stack the layers
+
+```
+
+ok new play - try creating an individual netcdf for a month...
+```{r}
+plst <- list.files("../output/maxent/predictions/testplots/testrasters", full.names = TRUE)
+
+thlst <- list()
+anlst <- list()
+for (i in 1:length(plst)){
+  r <- raster(plst[i])
+  aname <- names(r)
+  aname <-  str_sub(aname, 13)
+  r <- setZ(r, as.numeric(aname), name = "depth")
+  thlst[[i]] <- r
+  anlst[i] <- aname
+}
+
+nstk <- stack(thlst)
+nstk <- setZ(nstk, anlst, name = "depth")
+
+writeRaster(nstk, "../output/maxent/predictions/testplots/nstk.nc", overwrite = TRUE, format = "CDF", varname = "prob", varunit = "p", longname = "dummy netcdf for hovmollers", xname= "X", yname = "Y", zname = "depth", zunit = "m")
+```

### 2019-09-09 src/paper2_maxent_avec_xvalidation.Rmd
update to p2 with auto folder creation

@@
 
-this dataset contains a lot of columns you don't need for the maxent models. Remove them here
-
 #monthly model data selection and prep
@@
 
-select mont
+select month
 ```{r month of interest -change}
-mthna <- "jan"
-mthno <- "1"
+mthna <- "mar"
+mthno <- "3"
 ```
@@
 modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
-#modelmonth <- "jan2003"
 aicetc <- read.csv("../data/env/AICcCalculations_p2.csv", header = TRUE) #pre-created csv file
@@
   timeslice <- modelmonth[year] #testing timeslice
-  #timeslice <- modelmonth
+  newf <- paste0("../output/maxent/p2/models/", timeslice)
+  dir.create(newf)
   yr <- str_sub(timeslice, -4) #the year being used for testing
@@
     betamval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)), path = paste0("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxent/p2/models/", timeslice)) #here the full path is needed
+    mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)), path = paste0("../output/maxent/p2/models/", timeslice),"overwrite=true") #here the full path is needed
 
@@
   timeslice <- modelmonth[year] #testing timeslice
+  newf <- paste0("../output/maxent/p2/models/", timeslice)
+  dir.create(newf)
   yr <- str_sub(timeslice, -4) #the year being used for testing
@@
 
-  mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betaval)), path = paste0("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxent/p2/models/", timeslice)) #here the full path is needed
+  mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betaval)), path = paste0("../output/maxent/p2/models/", timeslice)) #here the full path is needed
 

### 2019-09-09 src/testing_file.Rmd
update to p2 with auto folder creation

@@
 plst <- list.files("../output/maxent/predictions/testplots/testrasters", full.names = TRUE)
+aea <- raster("../output/env/aea.tif") #the correct CRS
+apr <- aea@crs
+
+i = 1
 
@@
   r <- setZ(r, as.numeric(aname), name = "depth")
+  crs(r) <- apr
   thlst[[i]] <- r
@@
 
-writeRaster(nstk, "../output/maxent/predictions/testplots/nstk.nc", overwrite = TRUE, format = "CDF", varname = "prob", varunit = "p", longname = "dummy netcdf for hovmollers", xname= "X", yname = "Y", zname = "depth", zunit = "m")
-```
+writeRaster(r, "../output/maxent/predictions/testplots/lstk2.nc", overwrite = TRUE, format = "CDF", varname = "prob", varunit = "p", longname = "dummy netcdf for hovmollers", xname= "X", yname = "Y", zname = "depth", zunit = "m")
+
+dirXY <- xyLayer(r, sqrt(x^2 + y^2))
+library(zoo)
+
+hovmoller(nstk,
+          at = seq(0, 1, .1),
+                    interpolate = FALSE,
+          par.settings = BuRdTheme)
+
+hovmoller(nstk, ylab ="depth", interpolate = FALSE)
+
+```
+
+try create empty netcdf
+
+```{r}
+data.time <- 3:12 #months
+
+#get depths
+dep <- read.csv("../output/maxent/models/apr1998/stackdepthlist_1998.csv", header = TRUE)
+dep <- dep$ascdepths
+
+lon <- seq( -999722.7, 1525277, by = 25000)
+lat <- seq ( 3538427, -211573.1, by = -25000)
+
+data.distance <- seq(-65,920, by=5)
+## Add dimensions and variables which accompany the dimensions (avoided by create_dimvar = FALSE)
+dimDepth <- ncdim_def(name='depth', units='m', longname='depth layer', vals=data.distance )
+dimTime <- ncdim_def('time', units='month', longname='measurement month', calendar="standard", vals=data.time)
+
+varLat <- ncvar_def(name='lat', units='meters_north', dim=list(lat), missval=NA, longname='lat_m', prec='double')
+varLon <- ncvar_def(name='lon', units='meters_east', dim=list(lon), missval=NA, longname='lon_m', prec='double')
+
+varAlt <- ncvar_def(name='depth', units='m', dim=list(dimDepth, dimTime), missval=-9999, longname='depth layer')
+
+vars <- list(varLat, varLon, varAlt)
+
+outputfile <- 'transect.nc'
+con <- nc_create(outputfile, vars)
+
+```

### 2019-09-09 src/paper2_maxent_avec_xvalidation.Rmd
updating background points for p2

@@
 ipak(packages)
+```
+
+need to create 10k background points weighted to nafo zones (as per mj)
+
+use the exising file "../output/bio/background/correct_random_files/background_point_na_removal_nona.csv" which contains centroid location for every in the mcp....
+
+```{r}
+allbk <- read.csv("../output/bio/background/correct_random_files/background_point_na_removal_nona.csv", header = TRUE)
+
+```
+
+now you need to select 10k points, weighted by nafo region...
+
+0A	2.4	240
+0B	9.6	960
+1B	0.2	20
+1C	0.8	80
+1D	1.5	150
+1E	2	200
+1F	3.8	380
+2G	6.8	680
+2H	7.2	720
+2J	8.5	850
+3K	8.8	880
+3L	7.4	740
+3M	0.5	50
+3N	2.7	270
+3O	3.9	390
+3Pn	0.3	30
+3Ps	4	400
+4R	1.9	190
+4S	3.7	370
+4T	5.1	510
+4Vn	1.1	110
+4Vs	4	400
+4W	4	400
+4X	3.8	380
+5Y	1.5	150
+5Ze	0.5	50
+HudsonStrait	4	400
+
+```{r may 1999}
+may19990a <- subset(nafo_zone == "0A")
+may19990a <- sample_n(may19990a, 14) 
+may19990b <- subset(nafo_zone == "0B")
+may19990b <- sample_n(may19990b, 58) 
+may19991B <- subset(nafo_zone == "1B")
+may19991B <- sample_n(may19991B, 1)
+may19991C <- subset(nafo_zone == "1C")
+may19991C <- sample_n(may19991C, 5) 
+may19991d <- subset(nafo_zone == "1D")
+may19991d <- sample_n(may19991d, 9)
+may19991E <- subset(nafo_zone == "1E")
+may19991E <- sample_n(may19991E, 12)
+may19991F <- subset(nafo_zone == "1F")
+may19991F <- sample_n(may19991F, 23)
+may19992G <- subset(nafo_zone == "2G")
+may19992G <- sample_n(may19992G, 40) 
+may19992H <- subset(nafo_zone == "2H")
+may19992H <- sample_n(may19992H, 43) 
+may19992J <- subset(nafo_zone == "2J")
+may19992J <- sample_n(may19992J, 50) 
+may19993k <- subset(nafo_zone == "3K")
+may19993k <- sample_n(may19993k, 52) 
+may19993L <- subset(nafo_zone == "3L")
+may19993L <- sample_n(may19993L, 44)
+may19993M <- subset(nafo_zone == "3M")
+may19993M <- sample_n(may19993M, 3)
+may19993N <- subset(nafo_zone == "3N")
+may19993N <- sample_n(may19993N, 16) 
+may19993o <- subset(nafo_zone == "3O")
+may19993o <- sample_n(may19993o, 23) 
+may19993pn <- subset(nafo_zone == "3Pn")
+may19993pn <- sample_n(may19993pn, 2) 
+may19993ps <- subset(nafo_zone == "3Ps")
+may19993ps <- sample_n(may19993ps, 24) 
+may19994r <- subset(nafo_zone == "4R")
+may19994r <- sample_n(may19994r, 11) 
+may19994s <- subset(nafo_zone == "4S")
+may19994s <- sample_n(may19994s, 22) 
+may19994t <- subset(nafo_zone == "4T")
+may19994t <- sample_n(may19994t, 30) 
+may19994vn <- subset(nafo_zone == "4Vn")
+may19994vn <- sample_n(may19994vn, 6) 
+may19994vs <- subset(nafo_zone == "4Vs")
+may19994vs <- sample_n(may19994vs, 24) 
+may19994w <- subset(nafo_zone == "4W")
+may19994w <- sample_n(may19994w, 24) 
+may19994x <- subset(nafo_zone == "4X")
+may19994x <- sample_n(may19994x, 23) 
+may19995Y <- subset(nafo_zone == "5Y")
+may19995Y <- sample_n(may19995Y, 9) 
+may19995Ze <- subset(nafo_zone == "5Ze")
+may19995Ze <- sample_n(may19995Ze, 3) 
+may1999hud <- subset(nafo_zone == "HudsonStrait")
+may1999hud <- sample_n(may1999hud, 24) 
+may1999 <- rbind(may19990a, may19990b, may19991B, may19991C, may19991d, may19991E, may19991F,  may19992G, may19992H, may19992J, may19993k, may19993L, may19993M, may19993N, may19993o, may19993pn, may19993ps, may19994r, may19994s, may19994t, may19994vn, may19994vs, may19994w, may19994x, may19995Y, may19995Ze, may1999hud)
 ```
@@
 ```{r month of interest -change}
-mthna <- "mar"
-mthno <- "3"
+mthna <- "jan"
+mthno <- "1"
 ```
@@
 prback <- na.omit(prback)
+```
+
+just checking for no background points... should be 10k per year...
+```{r delete me}
+prbackb <- subset(prback, occurrence == 0)
+byr <- unique(prbackb$year)
 ```

### 2019-09-24 src/plots for paper.Rmd
plotting for paper - error with names for depths

---
title: "plots for paper"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

Built with 'r getRversion()'..

# Package dependencies
You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). Note this function checks to see if the packages are installed first before loading.

```{r pre-install & load packages, include=FALSE}
packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster")
source("../src/ipak.R")
ipak(packages)
```

# stacked permutation importance plot and jacknife plots

```{r create permutation & jack data}
maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
perm <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")

for (m in 1:length(maxr_lst)){
  maxr <- read.csv(maxr_lst[m])
  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
  perm$Month[m] <- mth
  jack$Month[m] <- mth
  maxt <- tail(maxr, 1)
  perm$amo_sample[m] <- maxt$amo_sample.permutation.importance
  perm$amo_winter[m] <- maxt$amo_winter.permutation.importance
  perm$chl_surface[m] <- maxt$chl_surface.permutation.importance
  perm$nao_prev[m] <- maxt$nao_prev.permutation.importance
  perm$nao_sample[m] <- maxt$nao_sample.permutation.importance
  perm$nao_winter[m] <- maxt$nao_winter.permutation.importance
  perm$o2_depth[m] <- maxt$o2_depth.permutation.importance
  perm$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
  perm$temp_depth[m] <- maxt$temp_depth.permutation.importance

  jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
  jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
  jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
  jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
  jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
  jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
  jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
  jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
  jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
}

mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
permu <- perm[match(mthord, perm$Month), ] #reorder based on mthord
jackk <- jack[match(mthord, jack$Month), ] #reorder based on mthord

#rename the months so they are capitalised
permu$Month <- capitalize(permu$Month)
jackk$Month <- capitalize(jackk$Month)

#rename the col names so they are capitalised and shortened
colnames(permu)[colnames(permu)=="amo_sample"] <- "AMO_Sample"
colnames(permu)[colnames(permu)=="amo_winter"] <- "AMO_Winter"
colnames(permu)[colnames(permu)=="chl_surface"] <- "Chl_Surface"
colnames(permu)[colnames(permu)=="nao_prev"] <- "NAO_Prev"
colnames(permu)[colnames(permu)=="nao_sample"] <- "NAO_Sample"
colnames(permu)[colnames(permu)=="nao_winter"] <- "NAO_Winter"
colnames(permu)[colnames(permu)=="o2_depth"] <- "O2_Depth"
colnames(permu)[colnames(permu)=="salinity_depth"] <- "Salinity_Depth"
colnames(permu)[colnames(permu)=="temp_depth"] <- "Temp_Depth"

colnames(jackk)[colnames(jackk)=="amo_sample"] <- "AMO_Sample"
colnames(jackk)[colnames(jackk)=="amo_winter"] <- "AMO_Winter"
colnames(jackk)[colnames(jackk)=="chl_surface"] <- "Chl_Surface"
colnames(jackk)[colnames(jackk)=="nao_prev"] <- "NAO_Prev"
colnames(jackk)[colnames(jackk)=="nao_sample"] <- "NAO_Sample"
colnames(jackk)[colnames(jackk)=="nao_winter"] <- "NAO_Winter"
colnames(jackk)[colnames(jackk)=="o2_depth"] <- "O2_Depth"
colnames(jackk)[colnames(jackk)=="salinity_depth"] <- "Salinity_Depth"
colnames(jackk)[colnames(jackk)=="temp_depth"] <- "Temp_Depth"

write.csv(permu, "../output/maxent/models/permutation_avg.csv", row.names = FALSE)
write.csv(jackk, "../output/maxent/models/jack_avg.csv", row.names = FALSE)

```

```{r create jacknife plots}
maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
jack_df <- list()

for (m in 1:length(maxr_lst)){
  res <- read.csv(maxr_lst[m], header = TRUE)
  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
  mth <- capitalize(mth)
  res_df <- tail(res, 1)
  var_df <- res_df[,c(grep("Training.gain.with*",names(res_df)))]
  var_df$id <- 1:nrow(var_df)
  var_df <- melt(var_df, id.vars=c("id"))
  jack_with <- var_df[grep("with.only",var_df$variable),]
  jack_with$Type_of_Run <- "With only the variable"
  jack_without <- var_df[grep("without",var_df$variable),]
  jack_without$Type_of_Run <- "Without the variable"
  gain_all <- res_df[c(grep("Regularized.",names(res_df)))]
  gain_all$id <- 1:nrow(gain_all)
  gain_all <- melt(gain_all, id.vars=c("id"))
  gain_all$Type_of_Run <- "With all variables"
  gain_all$variable <- "all_variables"
  jack_table <- rbind(jack_with,jack_without,gain_all)
  jack_table$state <- mth

  jack_table$variable <- gsub(".*[.]","",jack_table$variable)
  jack_table$variable <- gsub("amo_sample", "AMO_Sample", jack_table$variable)
  jack_table$variable <- gsub("amo_winter", "AMO_Winter", jack_table$variable)
  jack_table$variable <- gsub("chl_surface", "Chl_Surface", jack_table$variable)
  jack_table$variable <- gsub("nao_prev", "NAO_prev", jack_table$variable)
  jack_table$variable <- gsub("nao_sample", "NAO_Sample", jack_table$variable)
  jack_table$variable <- gsub("nao_winter", "NAO_Winter", jack_table$variable)
  jack_table$variable <- gsub("o2_depth", "O2_Depth", jack_table$variable)
  jack_table$variable <- gsub("salinity_depth", "Salinity_Depth", jack_table$variable)
  jack_table$variable <- gsub("temp_depth", "Temp_Depth", jack_table$variable)

  jack_df[[m]] <- jack_table

}

var_jack <- do.call(rbind, jack_df)

var_jack$state <- factor(var_jack$state,levels=c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
var_jack$Type_of_Run <- factor(var_jack$Type_of_Run,levels=c("With all variables", "Without the variable","With only the variable"))
a <- subset(var_jack, state == "Mar" | state == "Apr" |state == "May" | state == "Jun" | state == "Jul")
b <- subset(var_jack, state == "Aug" | state == "Sep" |state == "Oct" | state == "Nov" | state == "Dec")

plotlst <- list(a, b)

color_type_1 <- c("#081d58", "#999999","#1d91c0")
varpos <- c("Temp_Depth", "Salinity_Depth", "O2_Depth", "Chl_Surface", "NAO_Sample", "NAO_Prev", "NAO_Winter", "AMO_Sample", "AMO_Winter")

for (p in 1:length(plotlst)){
  pt <- p
  dev.new()
  jack <- ggplot(plotlst[[p]], aes(variable,value))+
  geom_bar(aes(fill=Type_of_Run),position="dodge",stat="identity")+ coord_flip()+ facet_grid(state~., scales="free", space = "free")+
  theme_bw(14)+
  scale_fill_manual(values=color_type_1,guide=guide_legend(reverse=FALSE))+
  theme(axis.text=element_text(size=8)) +
  theme(legend.title=element_blank())+
  theme(legend.position="top")+
  theme(strip.text=element_text(size=12))+
  theme(strip.text.y = element_text(angle = 0)) +
  labs(x="Variable",y="Jackknife of Regularized Training Gain")
  ggsave(filename = paste0("../output/maxent/paper_plots/jackknife_plot_", pt, ".png"), plot = jack, dpi = 900, height = 9, width = 7)
  dev.off()
}

```

```{r permulation plot}
permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = -1)
png("../output/maxent/paper_plots/permutation_plot.png")
print(perpl)
dev.off()
```

# plot median probability by depth

```{r get median probability for each depth}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

meddf <- data.frame(matrix(ncol = 11, nrow = 49))
colnames(meddf) <- c("depth", mthord)

for (m in 1:length(mthord)){
  mth <- mthord[m]
  mthlst <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
  rstk <- stack(mthlst)
  rsum <- summary(rstk, na.rm = TRUE)
  rsum <- rsum["Median", ]
  meddf[mth] <- rsum
  dep <- rstk@names
  write.csv(meddf, paste0("../output/maxent/paper_plots/med_data/median.csv"), row.names = TRUE)
 # write.csv(rsum, paste0("../output/maxent/paper_plots/med_data/median_", mth, ".csv"), row.names = TRUE)
}

medlst <- list.files(paste0("../output/maxent/paper_plots/med_data"), pattern = "^median.*.csv", full.names = TRUE)
tables <- lapply(medlst, read.csv, header = TRUE)
allmed <- do.call(cbind, tables)

```

### 2019-09-24 src/testing_file.Rmd
plotting for paper - error with names for depths

@@
 
-Threshold tables
+#Threshold tables
 ```{r}

### 2019-09-25 src/plots for paper.Rmd
median plot done

@@
 
-meddf <- data.frame(matrix(ncol = 11, nrow = 49))
-colnames(meddf) <- c("depth", mthord)
+meddf <- data.frame(matrix(ncol = 3, nrow = 490))
+colnames(meddf) <- c("layer", "month", "median")
+r <- 1
 
@@
   rsum <- rsum["Median", ]
-  meddf[mth] <- rsum
-  dep <- rstk@names
-  write.csv(meddf, paste0("../output/maxent/paper_plots/med_data/median.csv"), row.names = TRUE)
- # write.csv(rsum, paste0("../output/maxent/paper_plots/med_data/median_", mth, ".csv"), row.names = TRUE)
+  for (l in 1:length(rsum)){
+    meddf$median[r] <- rsum[l]
+    meddf$month[r] <- mth
+    meddf$layer[r] <- names(rstk[[l]])
+    r <- r+1
+  }
 }
 
-medlst <- list.files(paste0("../output/maxent/paper_plots/med_data"), pattern = "^median.*.csv", full.names = TRUE)
-tables <- lapply(medlst, read.csv, header = TRUE)
-allmed <- do.call(cbind, tables)
+#rename the layers to be just the depth
 
+meddf$layer <- gsub("avg_cloglog_", "", x = meddf$layer)
+meddf <- meddf[order(meddf$layer), ]
+
+write.csv(meddf, paste0("../output/maxent/paper_plots/med_data/median.csv"), row.names = FALSE)
 ```
+
+```{r}
+a <- ggplot(data = meddf) + geom_point(aes(y = layer, x = median, colour = month))
+a
+```

### 2019-10-01 src/AO_NAO_comparison.Rmd
i hate R

@@
 ```{r}
-naoraw <- read.csv("../data/env/nao_pc_monthlyX.csv", header = FALSE) 
+naoraw <- read.csv("../data/env/nao_pc_monthly.csv", header = FALSE) 
 head(naoraw)
@@
 ```{r}
-colnames(naoraw) <- c("year","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "yr_t")
+colnames(naoraw) <- c("year","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
 head(naoraw)

### 2019-10-01 src/plots for paper.Rmd
i hate R

@@
 a
+
 ```
+
+```{r get max probability for each depth}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+meddf <- data.frame(matrix(ncol = 3, nrow = 490))
+colnames(meddf) <- c("layer", "month", "max")
+r <- 1
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  mthlst <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+  rstk <- stack(mthlst)
+  rsum <- summary(rstk, na.rm = TRUE)
+  rsum <- rsum["Max.", ]
+  for (l in 1:length(rsum)){
+    meddf$max[r] <- rsum[l]
+    meddf$month[r] <- mth
+    meddf$layer[r] <- names(rstk[[l]])
+    r <- r+1
+  }
+}
+
+#rename the layers to be just the depth
+
+meddf$layer <- gsub("avg_cloglog_", "", x = meddf$layer)
+meddf$layer <- as.numeric(meddf$layer)
+meddf <- meddf[order(meddf$layer), ]
+
+write.csv(meddf, paste0("../output/maxent/paper_plots/med_data/max.csv"), row.names = FALSE)
+```
+
+```{r}
+a <- ggplot(data = meddf) + geom_point(aes(y = layer, x = max, colour = month))
+a
+```
+
+#Histograms of probability by depth
+
+count number of cells that are not NA in each raster layer that are inside the mcp
+
+```{r}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/nov"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
+cumst <- data.frame(0, 0)
+colnames(cumst) <- c("depth", "no_cells")
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumst[cum, 1] <- str_sub(ascname, 13)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  ccnt <- freq(cumasc, useNA='no')
+  ccnt <- sum(ccnt[ ,2])
+  cumst[cum, 2] <- ccnt
+}
+
+cumst$depth <- as.numeric(cumst$depth)
+cumst1 <- cumst[order(cumst$no_cells), ]
+
+write.csv(cumst1, "../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", row.names = FALSE)
+
+```
+
+first bin the probabilities for cells inside the mcp
+```{r bin probabilities}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+
+  binprob <- function(x){
+    ifelse(x > 0 & x < 0.1, 0.1,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.1 & x < 0.2, 0.2,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.2 & x < 0.3, 0.3,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.3 & x < 0.4, 0.4,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.4 & x < 0.5, 0.5,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.5 & x < 0.6, 0.6,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.6 & x < 0.7, 0.7,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.7 & x < 0.8, 0.8,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_8_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_8_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.8 & x < 0.9, 0.9,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_9_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_9_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.9 & x <= 1, 1,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_10_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_10_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+}
+```
+
+get data to plot probability by depth... need to do manually 1-10
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") 
+binlst <- c(1:10)
+
+binprob <- function(x){
+  ifelse(x >0,  1,
+  ifelse(x == -9999, 0, NA))
+}
+
+for (m in 1:length(mthord)){
+  for (b in 1:length(binlst)){
+  mth <- mthord[m]
+  bin <- binlst[b]
+  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, "_.*asc"), full.names = TRUE)
+
+  binaryavgstack <- stack(ascdepth)
+  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+  stackstathigh[ ,1] <- c(1:49)
+  colnames(stackstathigh) <- c("asclayer", "prob_bin")
+
+  for (lyr in 1:nlayers(binaryavgstack)){
+    ascname <-  str_sub(ascdepth[lyr], 63, -5)
+    dlay <- binaryavgstack[[lyr]]
+    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
+    lyrsum <- cellStats(dlay, sum)
+    stackstathigh$prob_bin[lyr] <- lyrsum
+    stackstathigh$asclayer[lyr] <- ascname
+
+  }
+  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_probbin_", bin, "_cellcount_bydepth_avg.csv"), row.names = FALSE)
+  }
+}
+
+```
+
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, ".*probbin_.*.csv"), full.names = TRUE)
+  a <- read.csv(probblst[1])
+  b <- read.csv(probblst[2])
+  c <- read.csv(probblst[3])
+  d <- read.csv(probblst[4])
+  e <- read.csv(probblst[5])
+  f <- read.csv(probblst[6])
+  g <- read.csv(probblst[7])
+  h <- read.csv(probblst[8])
+  i <- read.csv(probblst[9])
+  j <- read.csv(probblst[10])
+
+  stackstat <- cbind(a, b[ ,2], c[ ,2], d[ ,2], e[ , 2], f[ ,2], g[ ,2], h[ ,2], i[ ,2], j[ ,2])
+  colnames(stackstat) <- c("depth", "0 - 0.09", "0.1 - 0.19", "0.2 - 0.29", "0.3 - 0.39", "0.4 - 0.49", "0.5 - 0.59", "0.6 - 0.69", "0.7 - 0.79", "0.8 - 0.89", ">0.9")
+  stackstat <- melt(stackstat, id = "depth")
+  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]
+
+  colnames(stackstat) <- c("Depth", "Probability", "Cells")
+
+  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
+}
+
+```
+
+think about how to get proportions...
+
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
+  for (r in 1:nrow(a)){
+    a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
+    d1 <- str_sub(a$Depth[r], end = 4)
+	  for (ro in 1:nrow(cumst1)){
+      d2 <- str_sub(cumst1$depth[ro], end = 4)
+      if (d1 == d2){
+          a$total_cells[r] <- cumst1$no_cells[ro]
+      }
+	  }
+    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
+  }
+    a$Depth <- as.numeric(a$Depth)
+
+  write.csv(a, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), row.names = FALSE)
+}
+
+```
+
+histogram of probability by depth - by proportion
+
+```{r}
+
+bydepth <- ggplot(a, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() #XXXXXcareful - the deeper layers aren't being displayed correctly here
+plot(bydepth)
+#png(paste0("../output/maxent/predictions/", mthna, "/depthplot_probavg_prop.png"))
+#print(bydepth)
+#dev.off()
+```
+
+ok lets try colour the probabilites to match the maps...
+
+```{r}
+probcol <- rev(brewer.pal(10,"RdYlGn"))
+bydepth <- ggplot(a, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() + scale_fill_manual(values = probcol) #XXXXXcareful - the deeper layers aren't being displayed correctly here
+plot(bydepth)
+```
+
+ok now lets lump all the 0.39 and under together...
+
+```{r}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+binprob <- function(x){
+    ifelse(x > 0 & x <0.4, 1,
+    ifelse(x <= 0, 0, NA))
+}
+
+for (m in 1:length(mthord)){
+  mth <- "dec"
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- raster::mask(x = cumasc, mask = outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_binned_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    #plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_binned_avgprob_", ascname, ".png")), overwrite = TRUE)
+    #dev.off() # stops automatic saving of the plot to a png
+  }
+}
+```
+
+get data to plot probability by depth... 
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_u0.4_binned_avgprob_", ".*asc"), full.names = TRUE)
+
+  binaryavgstack <- stack(ascdepth)
+  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+  stackstathigh[ ,1] <- c(1:49)
+  colnames(stackstathigh) <- c("asclayer", "prob_bin")
+
+  for (lyr in 1:nlayers(binaryavgstack)){
+    ascname <-  str_sub(ascdepth[lyr], 63, -5)
+    dlay <- binaryavgstack[[lyr]]
+    lyrsum <- cellStats(dlay, sum)
+    stackstathigh$prob_bin[lyr] <- lyrsum
+    stackstathigh$asclayer[lyr] <- ascname
+
+  }
+  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_probbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
+}
+
+```
+
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_u0.4_probbin_cellcount_bydepth_avg.csv"), full.names = TRUE)
+  stackstat <- read.csv(probblst[1])
+
+  colnames(stackstat) <- c("depth", "0 - 0.39")
+  stackstat <- melt(stackstat, id = "depth")
+  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]
+
+  colnames(stackstat) <- c("Depth", "Probability", "Cells")
+
+  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
+}
+
+```
+
+get proportions...
+
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
+  for (r in 1:nrow(a)){
+    a$Depth[r] <- gsub("prob_avg_cloglog_", "", x = a$Depth[r])
+    d1 <- str_sub(a$Depth[r], end = 4)
+	  for (ro in 1:nrow(cumst1)){
+      d2 <- str_sub(cumst1$depth[ro], end = 4)
+      if (d1 == d2){
+          a$total_cells[r] <- cumst1$no_cells[ro]
+      }
+	  }
+    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
+  }
+   a$Depth <- as.numeric(a$Depth)
+
+  write.csv(a, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_dis_prob_cells.csv"), row.names = FALSE)
+}
+
+```
+
+merge the u0.4 dataset with the original
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
+  b <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
+  c <- rbind(a, b)
+  c <- c[!(c$Probability == "0 - 0.09"), ]
+  c <- c[!(c$Probability == "0.1 - 0.19"), ]
+  c <- c[!(c$Probability == "0.2 - 0.29"), ]
+  c <- c[!(c$Probability == "0.3 - 0.39"), ]
+  write.csv(c, paste0("../output//maxent/paper_plots/prob_by_depth/", mth, "_u0.4plus_dis_prob_cells.csv"))
+}
+```
+
+ok no try plot
+
+```{r}
+probcol <- rev(brewer.pal(7,"RdYlGn"))
+bydepth <- ggplot(c, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() + scale_fill_manual(values = probcol) #XXXXXcareful - the deeper layers aren't being displayed correctly here
+plot(bydepth)
+```
+
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), header = TRUE)
+  head(a)
+  d <- unique(a$Depth)
+  x <- data.frame(matrix(nrow = length(d)))
+  for (s in 1:length(d)){
+    dep <- d[s]
+    sd <- subset(a, Depth == dep)
+    sd <- subset(sd, Probability == "0 - 0.09"| Probability == "0.1 - 0.19"| Probability == "0.2 - 0.29"| Probability == "0.3 - 0.39")
+    c <- sum(sd$Cells)
+    x$Depth[s] <- dep
+    x$Probability[s] <- "0 - 0.39"
+    x$cells[s] <- c
+    x$p_cells[s] <- 1
+  }
+  x <- subset(x, select = -c(matrix.nrow...length.d..))
+  x$cells[is.na(x$cells)] <- 0
+  # for (r in 1:nrow(x)){
+  #   d1 <- str_sub(x$Depth[r], end = 4)
+  #   for (ro in 1:nrow(cumst1)){
+  #     d2 <- str_sub(cumst1$depth[ro], end = 4)
+  #     if (d1 == d2){
+  #         x$total_cells[r] <- cumst1$no_cells[ro]
+  #     }
+  #   }
+  # }
+  # for (r in 1:nrow(x)){
+  #    x$p_cells[r] <- (x$Cells[r]/x$total_cells[r])*100
+
+ # }
+
+  write.csv(x, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4plus_dis_prob_cells.csv"))
+}
+```
+
+merge the u0.4 dataset with the original
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4plus_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
+  b <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
+  c <- rbind(a, b)
+  c <- c[!(c$Probability == "0 - 0.09"), ]
+  c <- c[!(c$Probability == "0.1 - 0.19"), ]
+  c <- c[!(c$Probability == "0.2 - 0.29"), ]
+  c <- c[!(c$Probability == "0.3 - 0.39"), ]
+  write.csv(c, paste0("../output//maxent/paper_plots/prob_by_depth/", mth, "_u0.4s_dis_prob_cells.csv"))
+}
+```
+
+```{r}
+probcol <- rev(brewer.pal(7,"RdYlGn"))
+bydepth <- ggplot(c, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() + scale_fill_manual(values = probcol) #XXXXXcareful - the deeper layers aren't being displayed correctly here
+plot(bydepth)
+```
+
+ok dont forget the colours will need to match the maps so you may need to manually create your own list for whatever scheme you choose...
+
+now MJ suggested to only show some depths (the most commonly predicted ones)
+
+ok lets try get an probability for the whole month
+
+```{r}
+a <- read.csv("../output/maxent/predictions/jul/dis_prob_cells.csv", header = TRUE)
+
+#subset the probability ranges
+p1 <- subset(a, Probability == "0 - 0.09")
+p2 <- subset(a, Probability == "0.1 - 0.19")
+p3 <- subset(a, Probability == "0.2 - 0.29")
+p4 <- subset(a, Probability == "0.3 - 0.39")
+p5 <- subset(a, Probability == "0.4 - 0.49")
+p6 <- subset(a, Probability == "0.5 - 0.59")
+p7 <- subset(a, Probability == "0.6 - 0.69")
+p8 <- subset(a, Probability == "0.7 - 0.79")
+p9 <- subset(a, Probability == "0.8 - 0.89")
+p10 <- subset(a, Probability == ">0.9")
+
+#count the number of cells per probability.
+
+cp1 <- sum(p1$Cells)
+cp2 <- sum(p2$Cells)
+cp3 <- sum(p3$Cells)
+cp4 <- sum(p4$Cells)
+cp5 <- sum(p5$Cells)
+cp6 <- sum(p6$Cells)
+cp7 <- sum(p7$Cells)
+cp8 <- sum(p8$Cells)
+cp9 <- sum(p9$Cells)
+cp10 <- sum(p10$Cells)
+
+tc2 <- sum(cp1, cp2, cp3, cp4, cp5, cp6, cp7, cp8, cp9, cp10) #how many cells in total for novy have a value
+tcell <- sum(p1$total_cells) # how many cells are available to have a value...
+#tc2 and tcell should match
+
+#create datbase of total cells per probability bin for the month
+
+b <- unique(a$Probability) #get the unique probability bins
+c <- as.data.frame(rbind(cp1, cp2,  cp3, cp4, cp5, cp6, cp7, cp8, cp9, cp10))
+c$probability <- b
+colnames(c) <- c("cells", "prob_bin")
+rownames(c) <- c()
+c$month <- "jul"
+```
+
+make plot of probability over the month
+```{r}
+pmth <- ggplot(d, aes(month, cells, fill = prob_bin)) + geom_bar(stat='identity') 
+plot(pmth)
+png(paste0("../output/maxent/predictions/", mthna, "/prob_avg_mth.png"))
+print(pmth)
+dev.off()
+```
+
+merge two months of data together into one file then plot
+
+```{r}
+f <- rbind(c, d)
+pmth <- ggplot(f, aes(month, cells, fill = prob_bin)) + geom_bar(stat='identity') 
+plot(pmth)
+png(paste0("../output/maxent/predictions/prob_avg_mth_test.png"))
+print(pmth)
+dev.off()
+```

### 2019-10-01 src/testing_file.Rmd
i hate R

@@
 maplst <- list.files("../output/maxent/predictions/may", pattern = "^avg_clog.*.asc", full.names = TRUE)
+maplst2 <- list.files("../output/maxent/predictions/sep", pattern = "^avg_clog.*.asc", full.names = TRUE)
 s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
@@
 pras697 <- raster(maplst[41])
+p2ras6 <- raster(maplst2[37])
+p2ras180 <- raster(maplst2[14])
+p2ras370 <- raster(maplst2[28])
+p2ras697 <- raster(maplst2[41])
 crs(pras6) <- apr
@@
 crs(pras697) <- apr
+crs(p2ras6) <- apr
+crs(p2ras180) <- apr
+crs(p2ras370) <- apr
+crs(p2ras697) <- apr
 ascname6 <- names(pras6)
@@
 ascname697 <- names(pras697)
-pstack <- stack(pras6, pras180, pras370, pras697) #stack the layers
-
-pras1 <- raster(maplst[1])
-pras2 <- raster(maplst[2])
-pras3 <- raster(maplst[3])
-pras4 <- raster(maplst[4])
-crs(pras1) <- apr
-crs(pras2) <- apr
-crs(pras3) <- apr
-crs(pras4) <- apr
-
-pstack <- stack(pras6, pras180, pras370, pras697, pras1, pras2, pras3, pras4) #stack the layers
+a2scname6 <- names(p2ras6)
+a2scname180 <- names(p2ras180)
+a2scname370 <- names(p2ras370)
+a2scname697 <- names(p2ras697)
+
+# pstack <- stack(pras6, pras180, pras370, pras697) #stack the layers
+# 
+# pras1 <- raster(maplst[1])
+# pras2 <- raster(maplst[2])
+# pras3 <- raster(maplst[3])
+# pras4 <- raster(maplst[4])
+# crs(pras1) <- apr
+# crs(pras2) <- apr
+# crs(pras3) <- apr
+# crs(pras4) <- apr
+
+pstack <- stack(pras6, pras180, pras370, pras697, p2ras6, p2ras180, p2ras370, p2ras697) #stack the layers
 
 cls <- c("6m", "180m", "370m", "697m")
-rws <- c("may", "dummy")
+rws <- c("May", "Sep")
 mapTheme <- rasterTheme(region = rev(brewer.pal(10,"RdYlGn")))
@@
 
-dev.copy(png, (paste0("../output/maxent/predictions/testplots/rasterVis.png")), width = 1920,height = 1080)
-dev.off() # stops automatic saving of the plot to a png))
+#dev.copy(png, (paste0("../output/maxent/predictions/testplots/rasterVis.png")), width = 1920,height = 1080)
+#dev.off() # stops automatic saving of the plot to a png))
 

### 2019-10-02 src/plots for paper.Rmd
still hate R...

@@
 ```{r}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+mthord <- c("mar") #, "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"
 
@@
 ```{r}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+mthord <- c("mar") #,  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"
 cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
@@
 ```{r}
-outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-
-binprob <- function(x){
-    ifelse(x > 0 & x <0.4, 1,
-    ifelse(x <= 0, 0, NA))
-}
-
-for (m in 1:length(mthord)){
-  mth <- "dec"
-  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    cumasc <- raster::mask(x = cumasc, mask = outsidemcp, inverse = TRUE)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_binned_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    #plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_binned_avgprob_", ascname, ".png")), overwrite = TRUE)
-    #dev.off() # stops automatic saving of the plot to a png
-  }
-}
-```
-
-get data to plot probability by depth... 
-```{r}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-
-for (m in 1:length(mthord)){
-  mth <- mthord[m]
-  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_u0.4_binned_avgprob_", ".*asc"), full.names = TRUE)
-
-  binaryavgstack <- stack(ascdepth)
-  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-  stackstathigh[ ,1] <- c(1:49)
-  colnames(stackstathigh) <- c("asclayer", "prob_bin")
-
-  for (lyr in 1:nlayers(binaryavgstack)){
-    ascname <-  str_sub(ascdepth[lyr], 63, -5)
-    dlay <- binaryavgstack[[lyr]]
-    lyrsum <- cellStats(dlay, sum)
-    stackstathigh$prob_bin[lyr] <- lyrsum
-    stackstathigh$asclayer[lyr] <- ascname
-
-  }
-  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_probbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
-}
-
-```
-
-```{r}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-
-for (m in 1:length(mthord)){
-  mth <- mthord[m]
-  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_u0.4_probbin_cellcount_bydepth_avg.csv"), full.names = TRUE)
-  stackstat <- read.csv(probblst[1])
-
-  colnames(stackstat) <- c("depth", "0 - 0.39")
-  stackstat <- melt(stackstat, id = "depth")
-  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
-  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
-  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]
-
-  colnames(stackstat) <- c("Depth", "Probability", "Cells")
-
-  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
-}
-
-```
-
-get proportions...
-
-```{r}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
-
-for (m in 1:length(mthord)){
-  mth <- mthord[m]
-  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
-  for (r in 1:nrow(a)){
-    a$Depth[r] <- gsub("prob_avg_cloglog_", "", x = a$Depth[r])
-    d1 <- str_sub(a$Depth[r], end = 4)
-	  for (ro in 1:nrow(cumst1)){
-      d2 <- str_sub(cumst1$depth[ro], end = 4)
-      if (d1 == d2){
-          a$total_cells[r] <- cumst1$no_cells[ro]
-      }
-	  }
-    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
-  }
-   a$Depth <- as.numeric(a$Depth)
-
-  write.csv(a, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_dis_prob_cells.csv"), row.names = FALSE)
-}
-
-```
-
-merge the u0.4 dataset with the original
-```{r}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-for (m in 1:length(mthord)){
-  mth <- mthord[m]
-  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
-  b <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
-  c <- rbind(a, b)
-  c <- c[!(c$Probability == "0 - 0.09"), ]
-  c <- c[!(c$Probability == "0.1 - 0.19"), ]
-  c <- c[!(c$Probability == "0.2 - 0.29"), ]
-  c <- c[!(c$Probability == "0.3 - 0.39"), ]
-  write.csv(c, paste0("../output//maxent/paper_plots/prob_by_depth/", mth, "_u0.4plus_dis_prob_cells.csv"))
-}
-```
-
-ok no try plot
-
-```{r}
-probcol <- rev(brewer.pal(7,"RdYlGn"))
-bydepth <- ggplot(c, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() + scale_fill_manual(values = probcol) #XXXXXcareful - the deeper layers aren't being displayed correctly here
-plot(bydepth)
-```
-
-```{r}
 mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
@@
   head(a)
-  d <- unique(a$Depth)
+  d <- unique(cumst1$depth)
   x <- data.frame(matrix(nrow = length(d)))
@@
     x$Probability[s] <- "0 - 0.39"
-    x$cells[s] <- c
+    x$Cells[s] <- c
     x$p_cells[s] <- 1
+    x$total_cells[s] <- 1
+  }
+  x$Cells[is.na(x$Cells)] <- 0
+   for (r in 1:nrow(x)){
+     d1 <- str_sub(x$Depth[r], end = 4)
+     for (ro in 1:nrow(cumst1)){
+       d2 <- str_sub(cumst1$depth[ro], end = 4)
+       if (d1 == d2){
+           x$total_cells[r] <- cumst1$no_cells[ro]
+       }
+     }
+   }
+  for (r in 1:nrow(x)){
+    x$p_cells[r] <- (x$Cells[r]/x$total_cells[r])*100
   }
 x <- subset(x, select = -c(matrix.nrow...length.d..))
-  x$cells[is.na(x$cells)] <- 0
-  # for (r in 1:nrow(x)){
-  #   d1 <- str_sub(x$Depth[r], end = 4)
-  #   for (ro in 1:nrow(cumst1)){
-  #     d2 <- str_sub(cumst1$depth[ro], end = 4)
-  #     if (d1 == d2){
-  #         x$total_cells[r] <- cumst1$no_cells[ro]
-  #     }
-  #   }
-  # }
-  # for (r in 1:nrow(x)){
-  #    x$p_cells[r] <- (x$Cells[r]/x$total_cells[r])*100
-
- # }
-
   write.csv(x, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4plus_dis_prob_cells.csv"))
@@
   a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_u0.4plus_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
+  a <- subset(a, select = -c(X))
   b <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), header = TRUE, stringsAsFactors=FALSE)
@@
 
-ok lets try get an probability for the whole month
+```{r}
+probblst <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "_u0.4s_dis_prob_cells.csv", full.names = TRUE)
+a <- read.csv(probblst[1])
+b <- read.csv(probblst[2])
+c <- read.csv(probblst[3])
+d <- read.csv(probblst[4])
+e <- read.csv(probblst[5])
+f <- read.csv(probblst[6])
+g <- read.csv(probblst[7])
+h <- read.csv(probblst[8])
+i <- read.csv(probblst[9])
+j <- read.csv(probblst[10])
+
+stackstat <- cbind(a, b[ ,4], c[ ,4], d[ ,4], e[ ,4], f[ ,4], g[ ,4], h[ ,4], i[ ,4], j[ ,4])
+
+colnames(stackstat)
+
+x <- subset(stackstat, select = -c(X, p_cells))
+x <- x[ , c(1, 2, 4, 3, 5:13)]
+for (r in 1:nrow(x)){
+x$totalm[r] <- rowSums(x[r, 4:13])
+x$totala[r] <- x$total_cells[r]*10
+x$p_cells[r] <- (x$totalm[r]/x$totala[r])*100
+}
+y <- subset(x, select = c(Depth, Probability, p_cells))
+y$Probability <- gsub(">0.9", "0.9 - 1", y$Probability)
+
+probcol <- rev(brewer.pal(7,"RdYlGn"))
+bydepth <- ggplot(y, aes(Depth, p_cells, fill = Probability)) + geom_col() + scale_x_reverse() + coord_flip() + scale_fill_manual(values = probcol) 
+plot(bydepth)
+
+z <- subset(y, Depth <510)
+
+bydepth <- ggplot(z, aes(Depth, p_cells, fill = Probability)) + geom_bar(stat = "identity") + coord_flip() + scale_x_reverse() + scale_fill_manual(values = probcol) + theme_minimal() #still missing data...
+plot(bydepth)
+
+```
 
 ```{r}
-a <- read.csv("../output/maxent/predictions/jul/dis_prob_cells.csv", header = TRUE)
+z <- subset(y, Depth <510)
 
-#subset the probability ranges
-p1 <- subset(a, Probability == "0 - 0.09")
-p2 <- subset(a, Probability == "0.1 - 0.19")
-p3 <- subset(a, Probability == "0.2 - 0.29")
-p4 <- subset(a, Probability == "0.3 - 0.39")
-p5 <- subset(a, Probability == "0.4 - 0.49")
-p6 <- subset(a, Probability == "0.5 - 0.59")
-p7 <- subset(a, Probability == "0.6 - 0.69")
-p8 <- subset(a, Probability == "0.7 - 0.79")
-p9 <- subset(a, Probability == "0.8 - 0.89")
-p10 <- subset(a, Probability == ">0.9")
-
-#count the number of cells per probability.
-
-cp1 <- sum(p1$Cells)
-cp2 <- sum(p2$Cells)
-cp3 <- sum(p3$Cells)
-cp4 <- sum(p4$Cells)
-cp5 <- sum(p5$Cells)
-cp6 <- sum(p6$Cells)
-cp7 <- sum(p7$Cells)
-cp8 <- sum(p8$Cells)
-cp9 <- sum(p9$Cells)
-cp10 <- sum(p10$Cells)
-
-tc2 <- sum(cp1, cp2, cp3, cp4, cp5, cp6, cp7, cp8, cp9, cp10) #how many cells in total for novy have a value
-tcell <- sum(p1$total_cells) # how many cells are available to have a value...
-#tc2 and tcell should match
-
-#create datbase of total cells per probability bin for the month
-
-b <- unique(a$Probability) #get the unique probability bins
-c <- as.data.frame(rbind(cp1, cp2,  cp3, cp4, cp5, cp6, cp7, cp8, cp9, cp10))
-c$probability <- b
-colnames(c) <- c("cells", "prob_bin")
-rownames(c) <- c()
-c$month <- "jul"
+bydepth <- ggplot(z, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() + scale_fill_manual(values = probcol) 
+plot(bydepth)
 ```
-
-make plot of probability over the month
-```{r}
-pmth <- ggplot(d, aes(month, cells, fill = prob_bin)) + geom_bar(stat='identity') 
-plot(pmth)
-png(paste0("../output/maxent/predictions/", mthna, "/prob_avg_mth.png"))
-print(pmth)
-dev.off()
-```
-
-merge two months of data together into one file then plot
-
-```{r}
-f <- rbind(c, d)
-pmth <- ggplot(f, aes(month, cells, fill = prob_bin)) + geom_bar(stat='identity') 
-plot(pmth)
-png(paste0("../output/maxent/predictions/prob_avg_mth_test.png"))
-print(pmth)
-dev.off()
-```

### 2019-10-03 src/plots for paper.Rmd
fuck december

@@
 ```{r pre-install & load packages, include=FALSE}
-packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster")
+packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer")
 source("../src/ipak.R")
@@
 ```{r}
-mthord <- c("mar") #, "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
 
@@
 ```{r}
-mthord <- c("mar") #,  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"
+mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
 cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
@@
 z <- subset(y, Depth <510)
+z <- z[order(z$Depth), ]
 
-bydepth <- ggplot(z, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() + scale_fill_manual(values = probcol) 
-plot(bydepth)
+y$Depth <- as.factor(y$Depth)
+flevels <- levels(y$Depth)
+flevels <- rev(flevels)
+
+ggplot(y, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevels, breaks=flevels[seq(1,length(flevels), by = 2)]) + coord_flip()
+
 ```
+
+fuck december

### 2019-10-08 src/maxent_avec_xvalidation.Rmd
redoing dec issue

@@
 ```{r avg asc layers cloglog}
-mthna <- jul
-timeslice <- jul
+#p <- read.csv("../output/bio/presab_nafo_yr.csv", header= TRUE)
+mthna <- "nov"
+mthno <- "11"
+prback <- subset(p, month == mthno)
+modelmonth <- cbind(paste0(mthna, unique(prback$year)))
+timeslice <- modelmonth[1]
+
 clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
-stdep <- read.csv(paste0("../output/maxent/models/", timeslice, "/stack_depth.csv"), header = TRUE)
+stlst <- list.files(paste0("../output/maxent/models/", timeslice), pattern = "stackdepth.*.csv", full.names = TRUE)
+stdep <- read.csv(stlst[1])
+stdep$ascdepths <- str_sub(stdep$stackname, 9)
 depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
@@
   asclay$V15 <- as.character(asclay$V15)
+  asclay$V16 <- as.character(asclay$V16)
+  asclay$V17<- as.character(asclay$V17)
+  # asclay$V18 <- as.character(asclay$V18)
+  # asclay$V19<- as.character(asclay$V19)
 
@@
   asc2011 <- raster(asclay$V15)
-
-  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)
+  asc2012 <- raster(asclay$V16)
+  asc2013 <- raster(asclay$V17)
+  # asc2014 <- raster(asclay$V18)
+  # asc2015 <- raster(asclay$V19)
+
+  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011, asc2012, asc2013) #, asc2014, asc2015)
 
@@
   dev.off()
+
+file.rename(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_000120.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.asc"))
+file.rename(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_000120.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.png"))
+file.rename(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_041.18.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.asc"))
+file.rename(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_041.18.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.png"))
 }

### 2019-10-08 src/plots for paper.Rmd
redoing dec issue

@@
 
-clogasc_list <- list.files(paste0("../output/maxent/predictions/nov"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
+clogasc_list <- list.files(paste0("../output/maxent/predictions/dec"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
 cumst <- data.frame(0, 0)
@@
 for (cum in 1:length(clogasc_list)){
-  cumasc <- raster(clogasc_list[cum])
+  cumasc <- raster(clogasc_list[6])
   ascname <- names(cumasc)
   cumst[cum, 1] <- str_sub(ascname, 13)
-  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
-  ccnt <- freq(cumasc, useNA='no')
+  cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
+  ccnt <- freq(cumasc2, useNA='no')
   ccnt <- sum(ccnt[ ,2])
@@
 
-write.csv(cumst1, "../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", row.names = FALSE)
+write.csv(cumst1, "../output/maxent/paper_plots/prob_by_depth/cellsperlayermar.csv", row.names = FALSE)
 
@@
 first bin the probabilities for cells inside the mcp
+
+XXXXFROM HERE XXXXX
 ```{r bin probabilities}
@@
 
-get data to plot probability by depth... need to do manually 1-10
+get data to plot probability by depth... 
 ```{r}
@@
   bin <- binlst[b]
-  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, "_.*asc"), full.names = TRUE)
+  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, ".*asc"), full.names = TRUE)
 
@@
 ```{r}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+mthord <- c("dec") #"mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", 
 cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
 for (m in 1:length(mthord)){
-  mth <- mthord[m]
+  mth <- "dec" #mthord[m]
   a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), header = TRUE)
@@
 
+```{r}
+mar <- read.csv("../output//maxent/paper_plots/prob_by_depth/mar_u0.4s_dis_prob_cells.csv", header = TRUE)
+mar$Probability <- gsub(">0.9", "0.9 - 1", mar$Probability)
+mar$Depth <- as.factor(mar$Depth)
+flevelsm <- levels(mar$Depth)
+flevelsm <- rev(flevelsm)
+
+jun <- read.csv("../output//maxent/paper_plots/prob_by_depth/jun_u0.4s_dis_prob_cells.csv", header = TRUE)
+jun$Probability <- gsub(">0.9", "0.9 - 1", jun$Probability)
+jun$Depth <- as.factor(jun$Depth)
+flevelsj <- levels(jun$Depth)
+flevelsj <- rev(flevelsj)
+
+sep <- read.csv("../output//maxent/paper_plots/prob_by_depth/sep_u0.4s_dis_prob_cells.csv", header = TRUE)
+sep$Probability <- gsub(">0.9", "0.9 - 1", sep$Probability)
+sep$Depth <- as.factor(sep$Depth)
+flevelss <- levels(sep$Depth)
+flevelss <- rev(flevelss)
+
+dec <- read.csv("../output//maxent/paper_plots/prob_by_depth/dec_u0.4s_dis_prob_cells.csv", header = TRUE)
+dec$Probability <- gsub(">0.9", "0.9 - 1", dec$Probability)
+dec$Depth <- as.factor(dec$Depth)
+flevelsd <- levels(dec$Depth)
+flevelsd <- rev(flevelsd)
+
+probcol <- rev(brewer.pal(7,"RdYlGn"))
+
+ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsm, breaks=flevelsm[seq(1,length(flevelsm), by = 2)]) + coord_flip()
+
+ggplot(jun, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsj, breaks=flevelsj[seq(1,length(flevelsj), by = 2)]) + coord_flip()
+
+ggplot(sep, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelss, breaks=flevelss[seq(1,length(flevelss), by = 2)]) + coord_flip()
+
+ggplot(dec, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsd, breaks=flevelsd[seq(1,length(flevelsd), by = 2)]) + coord_flip()
+```
+
 now MJ suggested to only show some depths (the most commonly predicted ones)

### 2019-10-08 src/testing_file.Rmd
redoing dec issue

@@
 ```
+
+#argh
+```{r}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/dec"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
+cumst <- data.frame(0, 0)
+colnames(cumst) <- c("depth", "no_cells")
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[6])
+  ascname <- names(cumasc)
+  cumst[cum, 1] <- str_sub(ascname, 13)
+  cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
+  ccnt <- freq(cumasc2, useNA='no')
+  ccnt <- sum(ccnt[ ,2])
+  cumst[cum, 2] <- ccnt
+}
+
+clogasc_listm <- list.files(paste0("../output/maxent/predictions/mar"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
+cumstm <- data.frame(0, 0)
+colnames(cumstm) <- c("depth", "no_cells")
+for (cum in 1:length(clogasc_listm)){
+  cumascm <- raster(clogasc_listm[6])
+  ascnamem <- names(cumascm)
+  cumstm[cum, 1] <- str_sub(ascnamem, 13)
+  cumasc2m <- raster::mask(cumascm, outsidemcp, inverse = TRUE)
+  ccntm <- freq(cumasc2, useNA='no')
+  ccntm <- sum(ccntm[ ,2])
+  cumstm[cum, 2] <- ccntm
+}
+
+cumst$depth <- as.numeric(cumst$depth)
+cumst1 <- cumst[order(cumst$no_cells), ]
+
+write.csv(cumst1, "../output/maxent/paper_plots/prob_by_depth/cellsperlayermar.csv", row.names = FALSE)
+
+```
+
+```{r avg asc layers cloglog xx}
+#p <- read.csv("../output/bio/presab_nafo_yr.csv", header= TRUE)
+mthna <- "nov"
+mthno <- "11"
+prback <- subset(p, month == mthno)
+modelmonth <- cbind(paste0(mthna, unique(prback$year)))
+timeslice <- modelmonth[1]
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+stlst <- list.files(paste0("../output/maxent/models/", timeslice), pattern = "stackdepth.*.csv", full.names = TRUE)
+stdep <- read.csv(stlst[1])
+stdep$ascdepths <- str_sub(stdep$stackname, 9)
+depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
+names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
+depthlayers$depth <- as.character(depthlayers$depth)
+
+colno <- 1
+
+for (d in 1:nrow(depthlayers)){
+  dep <- depthlayers$depth[d]
+  for (a in 1:length(clogasc_list)){
+    asc <- clogasc_list[a]
+    ascd <- str_sub(asc, 50, -5)
+    if (grepl(dep, ascd)){
+      depthlayers[d , paste0("a", colno)] <- asc
+       colno <- colno+1
+    }
+  }
+}
+write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)
+
+for (x in 1:nrow(depthlayers)){
+  asclay <- depthlayers[x, ]
+  asclay <- asclay[!is.na(asclay)]
+  asclay <- as.data.frame(rbind(asclay))
+  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
+  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
+  asclay$V3 <- as.character(asclay$V3)
+  asclay$V4 <- as.character(asclay$V4)
+  asclay$V5 <- as.character(asclay$V5)
+  asclay$V6 <- as.character(asclay$V6)
+  asclay$V7 <- as.character(asclay$V7)
+  asclay$V8 <- as.character(asclay$V8)
+  asclay$V9 <- as.character(asclay$V9)
+  asclay$V10 <- as.character(asclay$V10)
+  asclay$V11 <- as.character(asclay$V11)
+  asclay$V12 <- as.character(asclay$V12)
+  asclay$V13 <- as.character(asclay$V13)
+  asclay$V14 <- as.character(asclay$V14)
+  asclay$V15 <- as.character(asclay$V15)
+  asclay$V16 <- as.character(asclay$V16)
+  asclay$V17<- as.character(asclay$V17)
+  # asclay$V18 <- as.character(asclay$V18)
+  # asclay$V19<- as.character(asclay$V19)
+
+  asclaydep <- asclay$V1
+
+  asc1998 <- raster(asclay$V2)
+  asc1999 <- raster(asclay$V3)
+  asc2000 <- raster(asclay$V4)
+  asc2001 <- raster(asclay$V5)
+  asc2002 <- raster(asclay$V6)
+  asc2003 <- raster(asclay$V7)
+  asc2004 <- raster(asclay$V8)
+  asc2005 <- raster(asclay$V9)
+  asc2006 <- raster(asclay$V10)
+  asc2007 <- raster(asclay$V11)
+  asc2008 <- raster(asclay$V12)
+  asc2009 <- raster(asclay$V13)
+  asc2010 <- raster(asclay$V14)
+  asc2011 <- raster(asclay$V15)
+  asc2012 <- raster(asclay$V16)
+  asc2013 <- raster(asclay$V17)
+  # asc2014 <- raster(asclay$V18)
+  # asc2015 <- raster(asclay$V19)
+
+  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011, asc2012, asc2013) #, asc2014, asc2015)
+
+  avg <- mean(stkasc, na.rm = TRUE)
+
+  writeRaster(avg, paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".asc"), overwrite=TRUE)
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".png"))) 
+  dev.off()
+
+file.rename(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_000120.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.asc"))
+file.rename(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_000120.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.png"))
+file.rename(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_041.18.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.asc"))
+file.rename(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_041.18.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.png"))
+}
+
+```

### 2019-10-10 src/plots for paper.Rmd
updated histograms, now on 2d plots - need to trim polygons

--- 2019-10-03 src/plots for paper.Rmd
@@
 ```{r pre-install & load packages, include=FALSE}
-packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer")
+packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer", "cowplot")
 source("../src/ipak.R")
@@
 first bin the probabilities for cells inside the mcp
+
 ```{r bin probabilities}
 outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+mthord <- c("dec") #"mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov",
 
@@
     bin <- binlst[b]
-  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, "_.*asc"), full.names = TRUE)
+    # mth <- "dec"
+    # bin <- 1
+  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, "_avg.*asc"), full.names = TRUE)
 
@@
   for (lyr in 1:nlayers(binaryavgstack)){
-    ascname <-  str_sub(ascdepth[lyr], 63, -5)
+    ascname <-  str_sub(ascdepth[lyr], 77, -5)
     dlay <- binaryavgstack[[lyr]]
@@
   mth <- mthord[m]
-  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, ".*probbin_.*.csv"), full.names = TRUE)
+  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_probbin_.*.csv"), full.names = TRUE)
   a <- read.csv(probblst[1])
@@
   for (r in 1:nrow(a)){
-    a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
-    d1 <- str_sub(a$Depth[r], end = 4)
+    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
+    # d1 <- str_sub(a$Depth[r], end = 4)
+    d1 <- a$Depth[r]
 	  for (ro in 1:nrow(cumst1)){
-      d2 <- str_sub(cumst1$depth[ro], end = 4)
+      #d2 <- str_sub(cumst1$depth[ro], end = 4)
+	    d2 <- cumst1$depth[ro]
       if (d1 == d2){
@@
 
+"1151.99"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]
+
+```{r}
+mar <- read.csv("../output//maxent/paper_plots/prob_by_depth/mar_u0.4s_dis_prob_cells.csv", header = TRUE)
+mar <- mar[!(mar$Depth =="1151.99" | mar$Depth =="1265.86" | mar$Depth =="1387.38"), ]
+mar$Probability <- gsub(">0.9", "0.9 - 1", mar$Probability)
+mar$Depth <- as.factor(mar$Depth)
+flevelsm <- levels(mar$Depth)
+flevelsm <- rev(flevelsm)
+
+jun <- read.csv("../output//maxent/paper_plots/prob_by_depth/jun_u0.4s_dis_prob_cells.csv", header = TRUE)
+jun <- jun[!(jun$Depth =="1151.99" | jun$Depth =="1265.86" | jun$Depth =="1387.38"), ]
+jun$Probability <- gsub(">0.9", "0.9 - 1", jun$Probability)
+jun$Depth <- as.factor(jun$Depth)
+flevelsj <- levels(jun$Depth)
+flevelsj <- rev(flevelsj)
+
+sep <- read.csv("../output//maxent/paper_plots/prob_by_depth/sep_u0.4s_dis_prob_cells.csv", header = TRUE)
+sep <- sep[!(sep$Depth =="1151.99" | sep$Depth =="1265.86" | sep$Depth =="1387.38"), ]
+sep$Probability <- gsub(">0.9", "0.9 - 1", sep$Probability)
+sep$Depth <- as.factor(sep$Depth)
+flevelss <- levels(sep$Depth)
+flevelss <- rev(flevelss)
+
+dec <- read.csv("../output//maxent/paper_plots/prob_by_depth/dec_u0.4s_dis_prob_cells.csv", header = TRUE)
+dec <- dec[!(dec$Depth =="1151.99" | dec$Depth =="1265.86" | dec$Depth =="1387.38"), ]
+dec$Probability <- gsub(">0.9", "0.9 - 1", dec$Probability)
+dec$Depth <- as.factor(dec$Depth)
+flevelsd <- levels(dec$Depth)
+flevelsd <- rev(flevelsd)
+
+probcol <- rev(brewer.pal(7,"RdYlGn"))
+
+pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsm, breaks=flevelsm[seq(1,length(flevelsm), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pj <- ggplot(jun, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsj, breaks=flevelsj[seq(1,length(flevelsj), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "June") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+ps <- ggplot(sep, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelss, breaks=flevelss[seq(1,length(flevelss), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "September") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pd <- ggplot(dec, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsd, breaks=flevelsd[seq(1,length(flevelsd), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "December") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+# # arrange the four plots in a single row
+# prow <- plot_grid(
+#   pm + theme(legend.position="none"),
+#   pj + theme(legend.position="none"),
+#  # ps + theme(legend.position="none"),
+# #  pd + theme(legend.position="none"),
+#   align = 'vh',
+#   #labels = c("A", "B") #, "C", "D"),
+#   hjust = -1,
+#   nrow = 1
+# )
+# 
+# legend <- get_legend(
+#   # create some space to the left of the legend
+#   pm + theme(legend.box.margin = margin(0, 0, 0, 12))
+# )
+# 
+# # add the legend to the row we made earlier. Give it one-third of 
+# # the width of one plot (via rel_widths).
+# t <- plot_grid(prow, legend, rel_widths = c(3, .4))
+# 
+# t
+pm
+
+```
+
 now MJ suggested to only show some depths (the most commonly predicted ones)
@@
 
-fuck december
+#2d hotspot maps
+
+first create layers that are 0.5 and above
+
+```{r habitat classification on probability - >0.5}
+
+binprob <- function(x){
+  ifelse(x >= 0.5, 1,
+  ifelse(x < 0.5, 0, NA)) 
+}
+
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mthna <- mthord[m]
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    ascname <- gsub("avg_cloglog_", "", ascname)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna,  "_binary_0.5_", ascname, ".asc"), overwrite = TRUE)
+     plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_binary_0.5_", ascname, ".png")), overwrite = TRUE)
+       dev.off() # stops automatic saving of the plot to a png
+  }
+}
+
+```
+
+then create an summary map....
+```{r high probability map per year 2D representation}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mthna <- mthord[m]
+  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_binary_0.5_.*.asc$"), full.names = TRUE)
+  binstack <- stack(binaryprob_list)
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.png")))
+  dev.off()
+}
+
+```
+
+the maps need to be layered with the land etc, trimmed, and coloured to whatever is agreed
+
+with rastervis

### 2019-10-10 src/testing_file.Rmd
updated histograms, now on 2d plots - need to trim polygons

@@
 ```
+
+#2d hotspot maps
+
+first create layers that are 0.5 and above
+
+```{r habitat classification on probability - >0.5}
+
+binprob <- function(x){
+  ifelse(x >= 0.5, 1,
+  ifelse(x < 0.5, 0, NA)) 
+}
+
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mthna <- mthord[m]
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    ascname <- gsub("avg_cloglog_", "", ascname)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna,  "_binary_0.5_", ascname, ".asc"), overwrite = TRUE)
+     plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_binary_0.5_", ascname, ".png")), overwrite = TRUE)
+       dev.off() # stops automatic saving of the plot to a png
+  }
+}
+
+```
+
+then create an summary map....
+```{r high probability map per year 2D representation}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mthna <- mthord[m]
+  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_binary_0.5_.*.asc$"), full.names = TRUE)
+  binstack <- stack(binaryprob_list)
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.png")))
+  dev.off()
+}
+
+```
+
+the maps need to be layered with the land etc, trimmed, and coloured to whatever is agreed
+
+```{r}
+aea <- raster("../output/env/aea.tif") #the correct CRS
+maplst <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "_bin_0.5_2d.*.asc", full.names = TRUE)
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+apr <- aea@crs
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+
+cumasc <- raster(maplst[1])
+cumasc2 <- raster(maplst[2])
+cumasc <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
+cumasc2 <- raster::mask(cumasc2, outsidemcp, inverse = TRUE)
+ss <- raster::mask(s, outsidemcp, inverse = TRUE) #HERE NEED TO WORK OUT HOW TO CLIP LAND
+
+pstack <- stack(cumasc, cumasc2)
+
+my.at <- seq(0, 1, by = 0.1)
+mapTheme <- viridisTheme(region = viridis(10))
+
+p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(2, 1), names.attr = c("6m", "180m"), margin = FALSE)
+p + layer(sp.polygons(s, fill = "black")) 
+cumasc <- raster::mask(p, outsidemcp, inverse = TRUE)
+
+```

### 2019-10-15 src/plots for paper.Rmd
update to depth plot + land trim... now working on reversing cols

@@
 
-pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsm, breaks=flevelsm[seq(1,length(flevelsm), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsm, breaks=flevelsm[seq(1,length(flevelsm), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
 
-pj <- ggplot(jun, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsj, breaks=flevelsj[seq(1,length(flevelsj), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "June") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+pj <- ggplot(jun, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsj, breaks=flevelsj[seq(1,length(flevelsj), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "June") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
 
-ps <- ggplot(sep, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelss, breaks=flevelss[seq(1,length(flevelss), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "September") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+ps <- ggplot(sep, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelss, breaks=flevelss[seq(1,length(flevelss), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "September") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
 
-pd <- ggplot(dec, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsd, breaks=flevelsd[seq(1,length(flevelsd), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "December") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+pd <- ggplot(dec, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsd, breaks=flevelsd[seq(1,length(flevelsd), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "December") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
 

### 2019-10-15 src/testing_file.Rmd
update to depth plot + land trim... now working on reversing cols

@@
 
-the maps need to be layered with the land etc, trimmed, and coloured to whatever is agreed
-
-```{r}
+#trimming maps to mcp
+
+```{r}
+library(viridis)
+library(rasterVis)
+
 aea <- raster("../output/env/aea.tif") #the correct CRS
@@
 outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
 
@@
 cumasc2 <- raster(maplst[2])
-cumasc <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
-cumasc2 <- raster::mask(cumasc2, outsidemcp, inverse = TRUE)
-ss <- raster::mask(s, outsidemcp, inverse = TRUE) #HERE NEED TO WORK OUT HOW TO CLIP LAND
+cumasc <- raster::mask(cumasc, mcp)
+cumasc2 <- raster::mask(cumasc2, mcp)
+ss <- raster::intersect(s, mcp)
 
@@
 my.at <- seq(0, 1, by = 0.1)
-mapTheme <- viridisTheme(region = viridis(10))
-
-p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(2, 1), names.attr = c("6m", "180m"), margin = FALSE)
-p + layer(sp.polygons(s, fill = "black")) 
-cumasc <- raster::mask(p, outsidemcp, inverse = TRUE)
-
-```
+mapTheme <- viridisTheme(region = viridis(10), direction = -1)
+
+p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(2, 1), names.attr = c("6m", "180m"), margin = FALSE) + layer(sp.polygons(ss, fill = "red"))
+plot(p)
+
+pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsm, breaks=flevelsm[seq(1,length(flevelsm), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+```

### 2019-10-16 src/plots for paper.Rmd
2d plots finished

@@
 ```{r pre-install & load packages, include=FALSE}
-packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer", "cowplot")
+packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer", "cowplot", "rasterVis")
 source("../src/ipak.R")
@@
 
-the maps need to be layered with the land etc, trimmed, and coloured to whatever is agreed
+now create summary maps where values are binary - not counts of cells
+```{r}
+binprob <- function(x){
+  ifelse(x > 0, 1,
+  ifelse(x <= 0, 0, NA))
+}
 
-with rastervis
+  clogasc_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = "*_bin_0.5_2d.asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+```
+
+then plot the maps
+
+```{r}
+
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
+ss <- raster::intersect(s, mcp)  #trim to mcp
+
+mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
+apr2d <- raster("../output/maxent/paper_plots/hotspot_mapping/apr_bin_0.5_2d_single.asc")
+may2d <- raster("../output/maxent/paper_plots/hotspot_mapping/may_bin_0.5_2d_single.asc")
+jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
+jul2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jul_bin_0.5_2d_single.asc")
+aug2d <- raster("../output/maxent/paper_plots/hotspot_mapping/aug_bin_0.5_2d_single.asc")
+sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
+oct2d <- raster("../output/maxent/paper_plots/hotspot_mapping/oct_bin_0.5_2d_single.asc")
+nov2d <- raster("../output/maxent/paper_plots/hotspot_mapping/nov_bin_0.5_2d_single.asc")
+dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")
+
+#trim to mcp
+mar2d <- raster::mask(mar2d, mcp)
+apr2d <- raster::mask(apr2d, mcp)
+may2d <- raster::mask(may2d, mcp)
+jun2d <- raster::mask(jun2d, mcp)
+jul2d <- raster::mask(jul2d, mcp)
+aug2d <- raster::mask(aug2d, mcp)
+sep2d <- raster::mask(sep2d, mcp)
+oct2d <- raster::mask(oct2d, mcp)
+nov2d <- raster::mask(nov2d, mcp)
+dec2d <- raster::mask(dec2d, mcp)
+
+#add crs
+mar2d@crs <- acrs
+apr2d@crs <- acrs
+may2d@crs <- acrs
+jun2d@crs <- acrs
+jul2d@crs <- acrs
+aug2d@crs <- acrs
+sep2d@crs <- acrs
+oct2d@crs <- acrs
+nov2d@crs <- acrs
+dec2d@crs <- acrs
+
+#ratify to categorical
+mar2d <- ratify(mar2d)
+rat <- levels(mar2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(mar2d) <- rat
+apr2d <- ratify(apr2d)
+rat <- levels(apr2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(apr2d) <- rat
+may2d <- ratify(may2d)
+rat <- levels(may2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(may2d) <- rat
+jun2d <- ratify(jun2d)
+rat <- levels(jun2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(jun2d) <- rat
+jul2d <- ratify(jul2d)
+rat <- levels(jul2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(jul2d) <- rat
+aug2d <- ratify(aug2d)
+rat <- levels(aug2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(aug2d) <- rat
+sep2d <- ratify(sep2d)
+rat <- levels(sep2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(sep2d) <- rat
+oct2d <- ratify(oct2d)
+rat <- levels(oct2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(oct2d) <- rat
+nov2d <- ratify(nov2d)
+rat <- levels(nov2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(nov2d) <- rat
+dec2d <- ratify(dec2d)
+rat <- levels(dec2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(dec2d) <- rat
+
+maplst <- list(mar2d, apr2d, may2d, jun2d, jul2d, aug2d, sep2d, oct2d, nov2d, dec2d)
+
+pstack <- stack(maplst)
+
+# my.at <- seq(0, 1, by = 0.1)
+# mapTheme <- infernoTheme(region = rev(inferno(10)))
+
+p <- levelplot(pstack, col.regions=c('white', 'red'), ylab = NULL, xlab = NULL, scales=list(draw=FALSE), layout = c(5, 2), names.attr = c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+plot(p)
+```
+
+#probabilty maps (selected)

### 2019-10-16 src/testing_file.Rmd
2d plots finished

@@
 my.at <- seq(0, 1, by = 0.1)
-mapTheme <- viridisTheme(region = viridis(10), direction = -1)
-
-p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(2, 1), names.attr = c("6m", "180m"), margin = FALSE) + layer(sp.polygons(ss, fill = "red"))
+mapTheme <- infernoTheme(region = rev(inferno(10)))
+
+p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(5, 2), names.attr = c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))
 plot(p)
 
-pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsm, breaks=flevelsm[seq(1,length(flevelsm), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
-
-```
+```

### 2019-10-16 src/testing_file.Rmd
looking to add contours

@@
 
-need to add seabed
-
-```{r seabed files}
-slst <- list.files("../data/env/ascii/ascii2010", pattern = "temp", full.names = TRUE)
-sr <- raster(slst[1])
-plot(sr)
-```
-
 ```{r}
@@
 ```
+
+# add seabed?
+need to add seabed
+
+```{r seabed files}
+contwgs <- shapefile("../data/env/contours_50m.shp") #created in arcgis from glorys bottomdepth using 3d licence contour tool
+plot(contwgs)
+aea <- raster("../output/env/aea.tif") #the correct CRS
+corcrs <- aea@crs
+contaea <- raster::reproject(contwgs, aea)
+
+```

### 2019-10-22 src/plots for paper.Rmd
looking to add contours
fandangled plotting

@@
 ```{r}
-mar <- read.csv("../output//maxent/paper_plots/prob_by_depth/mar_u0.4s_dis_prob_cells.csv", header = TRUE)
+mar <- read.csv("../output/maxent/paper_plots/prob_by_depth/mar_u0.4s_dis_prob_cells.csv", header = TRUE)
 mar <- mar[!(mar$Depth =="1151.99" | mar$Depth =="1265.86" | mar$Depth =="1387.38"), ]
@@
 mar$Depth <- as.factor(mar$Depth)
-flevelsm <- levels(mar$Depth)
-flevelsm <- rev(flevelsm)
-
-jun <- read.csv("../output//maxent/paper_plots/prob_by_depth/jun_u0.4s_dis_prob_cells.csv", header = TRUE)
+flevelsmar <- levels(mar$Depth)
+flevelsmar <- rev(flevelsmar)
+
+apr <- read.csv("../output/maxent/paper_plots/prob_by_depth/apr_u0.4s_dis_prob_cells.csv", header = TRUE)
+apr <- apr[!(apr$Depth =="1151.99" | apr$Depth =="1265.86" | apr$Depth =="1387.38"), ]
+apr$Probability <- gsub(">0.9", "0.9 - 1", apr$Probability)
+apr$Depth <- as.factor(apr$Depth)
+flevelsapr <- levels(apr$Depth)
+flevelsapr <- rev(flevelsapr)
+
+may <- read.csv("../output/maxent/paper_plots/prob_by_depth/may_u0.4s_dis_prob_cells.csv", header = TRUE)
+may <- may[!(may$Depth =="1151.99" | may$Depth =="1265.86" | may$Depth =="1387.38"), ]
+may$Probability <- gsub(">0.9", "0.9 - 1", may$Probability)
+may$Depth <- as.factor(may$Depth)
+flevelsmay <- levels(may$Depth)
+flevelsmay <- rev(flevelsmay)
+
+jun <- read.csv("../output/maxent/paper_plots/prob_by_depth/jun_u0.4s_dis_prob_cells.csv", header = TRUE)
 jun <- jun[!(jun$Depth =="1151.99" | jun$Depth =="1265.86" | jun$Depth =="1387.38"), ]
@@
 jun$Depth <- as.factor(jun$Depth)
-flevelsj <- levels(jun$Depth)
-flevelsj <- rev(flevelsj)
-
-sep <- read.csv("../output//maxent/paper_plots/prob_by_depth/sep_u0.4s_dis_prob_cells.csv", header = TRUE)
+flevelsjun <- levels(jun$Depth)
+flevelsjun <- rev(flevelsjun)
+
+jul <- read.csv("../output/maxent/paper_plots/prob_by_depth/jul_u0.4s_dis_prob_cells.csv", header = TRUE)
+jul <- jul[!(jul$Depth =="1151.99" | jul$Depth =="1265.86" | jul$Depth =="1387.38"), ]
+jul$Probability <- gsub(">0.9", "0.9 - 1", jul$Probability)
+jul$Depth <- as.factor(jul$Depth)
+flevelsjul <- levels(jul$Depth)
+flevelsjul <- rev(flevelsjul)
+
+aug <- read.csv("../output/maxent/paper_plots/prob_by_depth/aug_u0.4s_dis_prob_cells.csv", header = TRUE)
+aug <- aug[!(aug$Depth =="1151.99" | aug$Depth =="1265.86" | aug$Depth =="1387.38"), ]
+aug$Probability <- gsub(">0.9", "0.9 - 1", aug$Probability)
+aug$Depth <- as.factor(aug$Depth)
+flevelsaug <- levels(aug$Depth)
+flevelsaug <- rev(flevelsaug)
+
+sep <- read.csv("../output/maxent/paper_plots/prob_by_depth/sep_u0.4s_dis_prob_cells.csv", header = TRUE)
 sep <- sep[!(sep$Depth =="1151.99" | sep$Depth =="1265.86" | sep$Depth =="1387.38"), ]
@@
 sep$Depth <- as.factor(sep$Depth)
-flevelss <- levels(sep$Depth)
-flevelss <- rev(flevelss)
-
-dec <- read.csv("../output//maxent/paper_plots/prob_by_depth/dec_u0.4s_dis_prob_cells.csv", header = TRUE)
+flevelssep <- levels(sep$Depth)
+flevelssep <- rev(flevelssep)
+
+oct <- read.csv("../output/maxent/paper_plots/prob_by_depth/oct_u0.4s_dis_prob_cells.csv", header = TRUE)
+oct <- oct[!(oct$Depth =="1151.99" | oct$Depth =="1265.86" | oct$Depth =="1387.38"), ]
+oct$Probability <- gsub(">0.9", "0.9 - 1", oct$Probability)
+oct$Depth <- as.factor(oct$Depth)
+flevelsoct <- levels(oct$Depth)
+flevelsoct <- rev(flevelsoct)
+
+nov <- read.csv("../output/maxent/paper_plots/prob_by_depth/nov_u0.4s_dis_prob_cells.csv", header = TRUE)
+nov <- nov[!(nov$Depth =="1151.99" | nov$Depth =="1265.86" | nov$Depth =="1387.38"), ]
+nov$Probability <- gsub(">0.9", "0.9 - 1", nov$Probability)
+nov$Depth <- as.factor(nov$Depth)
+flevelsnov <- levels(nov$Depth)
+flevelsnov <- rev(flevelsnov)
+
+dec <- read.csv("../output/maxent/paper_plots/prob_by_depth/dec_u0.4s_dis_prob_cells.csv", header = TRUE)
 dec <- dec[!(dec$Depth =="1151.99" | dec$Depth =="1265.86" | dec$Depth =="1387.38"), ]
@@
 dec$Depth <- as.factor(dec$Depth)
-flevelsd <- levels(dec$Depth)
-flevelsd <- rev(flevelsd)
-
-probcol <- rev(brewer.pal(7,"RdYlGn"))
-
-pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsm, breaks=flevelsm[seq(1,length(flevelsm), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
-
-pj <- ggplot(jun, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsj, breaks=flevelsj[seq(1,length(flevelsj), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "June") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
-
-ps <- ggplot(sep, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelss, breaks=flevelss[seq(1,length(flevelss), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "September") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
-
-pd <- ggplot(dec, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsd, breaks=flevelsd[seq(1,length(flevelsd), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "December") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
-
-# # arrange the four plots in a single row
-# prow <- plot_grid(
-#   pm + theme(legend.position="none"),
-#   pj + theme(legend.position="none"),
-#  # ps + theme(legend.position="none"),
-# #  pd + theme(legend.position="none"),
-#   align = 'vh',
-#   #labels = c("A", "B") #, "C", "D"),
-#   hjust = -1,
-#   nrow = 1
-# )
-# 
-# legend <- get_legend(
-#   # create some space to the left of the legend
-#   pm + theme(legend.box.margin = margin(0, 0, 0, 12))
-# )
-# 
-# # add the legend to the row we made earlier. Give it one-third of 
-# # the width of one plot (via rel_widths).
-# t <- plot_grid(prow, legend, rel_widths = c(3, .4))
-# 
-# t
+flevelsdec <- levels(dec$Depth)
+flevelsdec <- rev(flevelsdec)
+
+probcol <- viridisTheme(option = inferno, rev(10))
+
+pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsmar, breaks=flevelsmar[seq(1,length(flevelsmar), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pa <- ggplot(apr, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsapr, breaks=flevelsapr[seq(1,length(flevelsapr), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "April") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pma <- ggplot(may, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="plasma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsmay, breaks=flevelsmay[seq(1,length(flevelsmay), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "May") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pj <- ggplot(jun, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_viridis(option="inferno", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsjun, breaks=flevelsjun[seq(1,length(flevelsjun), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "June") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pju <- ggplot(jul, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsjul, breaks=flevelsjul[seq(1,length(flevelsjul), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "July") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pau <- ggplot(aug, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsaug, breaks=flevelsaug[seq(1,length(flevelsaug), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "August") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+ps <- ggplot(sep, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelssep, breaks=flevelssep[seq(1,length(flevelssep), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "September") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+po <- ggplot(oct, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="plasma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsoct, breaks=flevelsoct[seq(1,length(flevelsoct), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "October") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pn <- ggplot(nov, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsnov, breaks=flevelsnov[seq(1,length(flevelsnov), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "November") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pd <- ggplot(dec, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_viridis(option = "inferno", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsdec, breaks=flevelsdec[seq(1,length(flevelsdec), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "December") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
 pm
-
-```
-
-now MJ suggested to only show some depths (the most commonly predicted ones)
+pa
+pma
+pj
+pju
+pa
+ps
+po
+pn
+pd
+
+```
+
+and repeat without the 0.4 cut....
+
+```{r}
+mar <- read.csv("../output/maxent/paper_plots/prob_by_depth/mar_dis_prob_cells.csv", header = TRUE)
+mar <- mar[!(mar$Depth =="1151.99" | mar$Depth =="1265.86" | mar$Depth =="1387.38"), ]
+mar$Probability <- gsub(">0.9", "0.9 - 1", mar$Probability)
+mar$Depth <- as.factor(mar$Depth)
+flevelsmar <- levels(mar$Depth)
+flevelsmar <- rev(flevelsmar)
+
+apr <- read.csv("../output/maxent/paper_plots/prob_by_depth/apr_dis_prob_cells.csv", header = TRUE)
+apr <- apr[!(apr$Depth =="1151.99" | apr$Depth =="1265.86" | apr$Depth =="1387.38"), ]
+apr$Probability <- gsub(">0.9", "0.9 - 1", apr$Probability)
+apr$Depth <- as.factor(apr$Depth)
+flevelsapr <- levels(apr$Depth)
+flevelsapr <- rev(flevelsapr)
+
+may <- read.csv("../output/maxent/paper_plots/prob_by_depth/may_dis_prob_cells.csv", header = TRUE)
+may <- may[!(may$Depth =="1151.99" | may$Depth =="1265.86" | may$Depth =="1387.38"), ]
+may$Probability <- gsub(">0.9", "0.9 - 1", may$Probability)
+may$Depth <- as.factor(may$Depth)
+flevelsmay <- levels(may$Depth)
+flevelsmay <- rev(flevelsmay)
+
+jun <- read.csv("../output/maxent/paper_plots/prob_by_depth/jun_dis_prob_cells.csv", header = TRUE)
+jun <- jun[!(jun$Depth =="1151.99" | jun$Depth =="1265.86" | jun$Depth =="1387.38"), ]
+jun$Probability <- gsub(">0.9", "0.9 - 1", jun$Probability)
+jun$Depth <- as.factor(jun$Depth)
+flevelsjun <- levels(jun$Depth)
+flevelsjun <- rev(flevelsjun)
+
+jul <- read.csv("../output/maxent/paper_plots/prob_by_depth/jul_dis_prob_cells.csv", header = TRUE)
+jul <- jul[!(jul$Depth =="1151.99" | jul$Depth =="1265.86" | jul$Depth =="1387.38"), ]
+jul$Probability <- gsub(">0.9", "0.9 - 1", jul$Probability)
+jul$Depth <- as.factor(jul$Depth)
+flevelsjul <- levels(jul$Depth)
+flevelsjul <- rev(flevelsjul)
+
+aug <- read.csv("../output/maxent/paper_plots/prob_by_depth/aug_dis_prob_cells.csv", header = TRUE)
+aug <- aug[!(aug$Depth =="1151.99" | aug$Depth =="1265.86" | aug$Depth =="1387.38"), ]
+aug$Probability <- gsub(">0.9", "0.9 - 1", aug$Probability)
+aug$Depth <- as.factor(aug$Depth)
+flevelsaug <- levels(aug$Depth)
+flevelsaug <- rev(flevelsaug)
+
+sep <- read.csv("../output/maxent/paper_plots/prob_by_depth/sep_dis_prob_cells.csv", header = TRUE)
+sep <- sep[!(sep$Depth =="1151.99" | sep$Depth =="1265.86" | sep$Depth =="1387.38"), ]
+sep$Probability <- gsub(">0.9", "0.9 - 1", sep$Probability)
+sep$Depth <- as.factor(sep$Depth)
+flevelssep <- levels(sep$Depth)
+flevelssep <- rev(flevelssep)
+
+oct <- read.csv("../output/maxent/paper_plots/prob_by_depth/oct_dis_prob_cells.csv", header = TRUE)
+oct <- oct[!(oct$Depth =="1151.99" | oct$Depth =="1265.86" | oct$Depth =="1387.38"), ]
+oct$Probability <- gsub(">0.9", "0.9 - 1", oct$Probability)
+oct$Depth <- as.factor(oct$Depth)
+flevelsoct <- levels(oct$Depth)
+flevelsoct <- rev(flevelsoct)
+
+nov <- read.csv("../output/maxent/paper_plots/prob_by_depth/nov_dis_prob_cells.csv", header = TRUE)
+nov <- nov[!(nov$Depth =="1151.99" | nov$Depth =="1265.86" | nov$Depth =="1387.38"), ]
+nov$Probability <- gsub(">0.9", "0.9 - 1", nov$Probability)
+nov$Depth <- as.factor(nov$Depth)
+flevelsnov <- levels(nov$Depth)
+flevelsnov <- rev(flevelsnov)
+
+dec <- read.csv("../output/maxent/paper_plots/prob_by_depth/dec_dis_prob_cells.csv", header = TRUE)
+dec <- dec[!(dec$Depth =="1151.99" | dec$Depth =="1265.86" | dec$Depth =="1387.38"), ]
+dec$Probability <- gsub(">0.9", "0.9 - 1", dec$Probability)
+dec$Depth <- as.factor(dec$Depth)
+flevelsdec <- levels(dec$Depth)
+flevelsdec <- rev(flevelsdec)
+
+probcol <- viridisTheme(option = inferno, rev(10))
+
+pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsmar, breaks=flevelsmar[seq(1,length(flevelsmar), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pa <- ggplot(apr, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsapr, breaks=flevelsapr[seq(1,length(flevelsapr), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "April") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pma <- ggplot(may, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="plasma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsmay, breaks=flevelsmay[seq(1,length(flevelsmay), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "May") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pj <- ggplot(jun, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_viridis(option="inferno", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsjun, breaks=flevelsjun[seq(1,length(flevelsjun), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "June") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pju <- ggplot(jul, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsjul, breaks=flevelsjul[seq(1,length(flevelsjul), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "July") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pau <- ggplot(aug, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsaug, breaks=flevelsaug[seq(1,length(flevelsaug), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "August") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+ps <- ggplot(sep, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelssep, breaks=flevelssep[seq(1,length(flevelssep), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "September") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+po <- ggplot(oct, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="plasma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsoct, breaks=flevelsoct[seq(1,length(flevelsoct), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "October") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pn <- ggplot(nov, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsnov, breaks=flevelsnov[seq(1,length(flevelsnov), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "November") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pd <- ggplot(dec, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_viridis(option = "inferno", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsdec, breaks=flevelsdec[seq(1,length(flevelsdec), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "December") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))
+
+pm
+pa
+pma
+pj
+pju
+pa
+ps
+po
+pn
+pd
+
+```
+
+now MJ suggested to only show some depths (the most commonly predicted ones) NOTE DECIDED MAYBE NOT GOOD BECAUSE OF NEW GRAPHIC PLUS DEPTH CHANGES
 
@@
 then create an summary map....
-```{r high probability map per year 2D representation}
+```{r >0.5 prob monthly 2D representation}
 mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
@@
 #probabilty maps (selected)
+
+xxxx
+```{r}
+library(raster)
+library(rgdal)
+library(rasterVis)
+library(colorspace)
+
+aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
+ss <- raster::intersect(s, mcp)  #trim land file to mcp
+
+mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
+apr2d <- raster("../output/maxent/paper_plots/hotspot_mapping/apr_bin_0.5_2d_single.asc")
+may2d <- raster("../output/maxent/paper_plots/hotspot_mapping/may_bin_0.5_2d_single.asc")
+jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
+jul2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jul_bin_0.5_2d_single.asc")
+aug2d <- raster("../output/maxent/paper_plots/hotspot_mapping/aug_bin_0.5_2d_single.asc")
+sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
+oct2d <- raster("../output/maxent/paper_plots/hotspot_mapping/oct_bin_0.5_2d_single.asc")
+nov2d <- raster("../output/maxent/paper_plots/hotspot_mapping/nov_bin_0.5_2d_single.asc")
+dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")
+
+#trim to mcp
+mar2d <- raster::mask(mar2d, mcp)
+apr2d <- raster::mask(apr2d, mcp)
+may2d <- raster::mask(may2d, mcp)
+jun2d <- raster::mask(jun2d, mcp)
+jul2d <- raster::mask(jul2d, mcp)
+aug2d <- raster::mask(aug2d, mcp)
+sep2d <- raster::mask(sep2d, mcp)
+oct2d <- raster::mask(oct2d, mcp)
+nov2d <- raster::mask(nov2d, mcp)
+dec2d <- raster::mask(dec2d, mcp)
+
+#add crs
+mar2d@crs <- acrs
+apr2d@crs <- acrs
+may2d@crs <- acrs
+jun2d@crs <- acrs
+jul2d@crs <- acrs
+aug2d@crs <- acrs
+sep2d@crs <- acrs
+oct2d@crs <- acrs
+nov2d@crs <- acrs
+dec2d@crs <- acrs
+
+maplst <- list.files("../output/maxent/predictions/may", pattern = "^avg_clog.*.asc", full.names = TRUE)
+maplst2 <- list.files("../output/maxent/predictions/sep", pattern = "^avg_clog.*.asc", full.names = TRUE)
+brk <- seq(0, 1, 0.1) #for legend
+lgc <- sequential_hcl(10, palette = "YlOrRd") #colour scheme from colorspace package
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+
+pras6 <- raster(maplst[37])
+pras180 <- raster(maplst[14])
+pras370 <- raster(maplst[28])
+pras697 <- raster(maplst[41])
+p2ras6 <- raster(maplst2[37])
+p2ras180 <- raster(maplst2[14])
+p2ras370 <- raster(maplst2[28])
+p2ras697 <- raster(maplst2[41])
+crs(pras6) <- apr
+crs(pras180) <- apr
+crs(pras370) <- apr
+crs(pras697) <- apr
+crs(p2ras6) <- apr
+crs(p2ras180) <- apr
+crs(p2ras370) <- apr
+crs(p2ras697) <- apr
+ascname6 <- names(pras6)
+ascname180 <- names(pras180)
+ascname370 <- names(pras370)
+ascname697 <- names(pras697)
+a2scname6 <- names(p2ras6)
+a2scname180 <- names(p2ras180)
+a2scname370 <- names(p2ras370)
+a2scname697 <- names(p2ras697)
+
+# pstack <- stack(pras6, pras180, pras370, pras697) #stack the layers
+# 
+# pras1 <- raster(maplst[1])
+# pras2 <- raster(maplst[2])
+# pras3 <- raster(maplst[3])
+# pras4 <- raster(maplst[4])
+# crs(pras1) <- apr
+# crs(pras2) <- apr
+# crs(pras3) <- apr
+# crs(pras4) <- apr
+
+pstack <- stack(pras6, pras180, pras370, pras697, p2ras6, p2ras180, p2ras370, p2ras697) #stack the layers
+
+cls <- c("6m", "180m", "370m", "697m")
+rws <- c("May", "Sep")
+mapTheme <- rasterTheme(region = rev(brewer.pal(10,"RdYlGn")))
+my.at <- seq(0, 1, by = 0.1)
+
+p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(4, 2), names.attr = c("6m", "180m", "370m", "697m", "", "", "", ""), margin = FALSE)
+p + layer(sp.polygons(s, fill = "black")) + layer (sp.polygons(outsidemcp, fill = redtrans, col = NA))
+
+#dev.copy(png, (paste0("../output/maxent/predictions/testplots/rasterVis.png")), width = 1920,height = 1080)
+#dev.off() # stops automatic saving of the plot to a png))
+
+```
+
+#coloured histograms
+
+```{r}
+mar <- raster("../output/maxent/predictions/mar/avg_cloglog_97.0413.asc")
+#apr <- raster("../output/maxent/predictions/apr/avg_cloglog_97.0413.asc")
+
+mardf <- as.data.frame(mar) #convert raster to dataframe
+mardf <- na.omit(mardf) #remove all the land 
+for (i in 1:nrow(mardf)){
+  if (mardf$avg_cloglog_97.0413[i] <=0.1){
+    mardf$bin[i] <- 1
+  } else if (mardf$avg_cloglog_97.0413[i] >0.1 & mardf$avg_cloglog_97.0413[i] <= 0.2){
+      mardf$bin[i] <- 2
+  } else if (mardf$avg_cloglog_97.0413[i] >0.2 & mardf$avg_cloglog_97.0413[i] <= 0.3){
+      mardf$bin[i] <- 3
+  } else if (mardf$avg_cloglog_97.0413[i] >0.3 & mardf$avg_cloglog_97.0413[i] <= 0.4){
+      mardf$bin[i] <- 4
+  } else if (mardf$avg_cloglog_97.0413[i] >0.4 & mardf$avg_cloglog_97.0413[i] <= 0.5){
+      mardf$bin[i] <- 5
+  } else if (mardf$avg_cloglog_97.0413[i] >0.5 & mardf$avg_cloglog_97.0413[i] <= 0.6){
+      mardf$bin[i] <- 6
+  } else if (mardf$avg_cloglog_97.0413[i] >0.6 & mardf$avg_cloglog_97.0413[i] <= 0.7){
+      mardf$bin[i] <- 7
+  } else if (mardf$avg_cloglog_97.0413[i] >0.7 & mardf$avg_cloglog_97.0413[i] <= 0.8){
+      mardf$bin[i] <- 8
+  } else if (mardf$avg_cloglog_97.0413[i] >0.8 & mardf$avg_cloglog_97.0413[i] <= 0.9){
+      mardf$bin[i] <- 9
+  } else if (mardf$avg_cloglog_97.0413[i] >0.9){
+      mardf$bin[i] <- 10
+  }
+}
+
+mardf$bin <- as.factor(mardf$bin)
+
+brk <- seq(0.1, 1, 0.1) #bins you want
+brkl <- as.character(brk)
+
+p <- ggplot(mardf, aes(x = avg_cloglog_97.0413, fill = bin))+ geom_histogram(breaks = brk) + theme_classic() + scale_fill_viridis_d(option = "inferno", direction = -1) + theme(legend.position = "none", plot.title = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank()) + scale_x_continuous(breaks=seq(0,1,0.1))
+png("../output/maxent/paper_plots/histogram_mar_97.png", height = 1024, width = 1024)
+p
+dev.off()
+
+```
+
+# variable maps
+- temp
+- chl 
+- o2
+- sal
+- 4 different months
+- relationship of variables to each other
+- pick a depth (say the 11.7737 m)
+
+months are mar, jun, sep, dec
+pick 3 diff years say 1998, 2006, 2013
+
+```{r}
+aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
+ss <- raster::intersect(s, mcp)  #trim land file to mcp
+
+yrlst <- c("1998", "2006", "2013")
+mthlst <- c("03", "06", "09", "12")
+varlst <- c("temp", "salinity", "o2")
+
+for (y in 1:length(yrlst)){
+  yr <- yrlst[y]
+  fdr <- paste0("../data/env/ascii/ascii", yr, "/")
+  for (m in 1:length(mthlst)){
+    mt <- mthlst[m]
+    for (v in 1:length(varlst)){
+      var <- varlst[v]
+      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_11.7737.asc"))
+      lyrnm <- names(ras)
+      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      ras <- projectRaster(ras, aea)
+      ras <- raster::mask(ras, mcp)
+
+      p <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))
+
+      png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
+      print(p)
+      dev.off()
+    }
+  }
+}
+
+#chl is done seperately as it is surface
+
+yrlst <- c("1998", "2006", "2013")
+mthlst <- c("03", "06", "09", "12")
+varlst <- c("chl")
+
+for (y in 1:length(yrlst)){
+  yr <- yrlst[y]
+  fdr <- paste0("../data/env/ascii/ascii", yr, "/")
+  for (m in 1:length(mthlst)){
+    mt <- mthlst[m]
+    for (v in 1:length(varlst)){
+      var <- varlst[v]
+      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_0.50576.asc"))
+      lyrnm <- names(ras)
+      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      ras <- projectRaster(ras, aea)
+      ras <- raster::mask(ras, mcp)
+
+      p <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))
+
+      png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
+      print(p)
+      dev.off()
+    }
+  }
+}
+```

### 2019-10-22 src/testing_file.Rmd
fandangled plotting

@@
 
-boxplot of test auc scores test
+#boxplot of test auc scores test
 
@@
 
-boxplot of train auc scores test
+#boxplot of train auc scores test
 
@@
 
-boxplot of cbi
+#boxplot of cbi
 
@@
 
-boxplots of tss scores
+#boxplots of tss scores
 ```{r}
@@
     probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna,  "_binary_0.5_", ascname, ".asc"), overwrite = TRUE)
-     plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_binary_0.5_", ascname, ".png")), overwrite = TRUE)
+    writeRaster(probbin, paste0("../output/maxent/predictions/mar/", mthna,  "_binary_0.5_", ascname, ".asc"), overwrite = TRUE)
+     plot(probbin, png(paste0("../output/maxent/predictions/mar/", mthna, "_binary_0.5_", ascname, ".png")), overwrite = TRUE)
        dev.off() # stops automatic saving of the plot to a png
@@
   mthna <- mthord[m]
-  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_binary_0.5_.*.asc$"), full.names = TRUE)
+  binaryprob_list <- list.files(paste0("../output/maxent/predictions/mar/"), pattern = paste0(mthna, "*_binary_0.5_.*.asc$"), full.names = TRUE)
   binstack <- stack(binaryprob_list)
   stacksum <- sum(binstack, na.rm = TRUE)
-  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.asc"), overwrite=TRUE)
-  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.png")))
+  writeRaster(stacksum, paste0("../output/maxent/predictions/mar/", mthna, "_bin_0.5_2d.asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/predictions/mar/", mthna, "_bin_0.5_2d.png")))
   dev.off()
@@
 aea <- raster("../output/env/aea.tif") #the correct CRS
-maplst <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "_bin_0.5_2d.*.asc", full.names = TRUE)
+maplst <- list.files("../output/maxent/predictions/mar/", pattern = "_bin_0.5_2d.*.asc", full.names = TRUE)
 s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
@@
 
-# add seabed?
+# add contour?
 need to add seabed
@@
 ```
+
+#coloured histograms
+
+```{r}
+mar <- raster("../output/maxent/predictions/mar/avg_cloglog_97.0413.asc")
+#apr <- raster("../output/maxent/predictions/apr/avg_cloglog_97.0413.asc")
+
+mardf <- as.data.frame(mar) #convert raster to dataframe
+mardf <- na.omit(mardf) #remove all the land 
+for (i in 1:nrow(mardf)){
+  if (mardf$avg_cloglog_97.0413[i] <=0.1){
+    mardf$bin[i] <- 1
+  } else if (mardf$avg_cloglog_97.0413[i] >0.1 & mardf$avg_cloglog_97.0413[i] <= 0.2){
+      mardf$bin[i] <- 2
+  } else if (mardf$avg_cloglog_97.0413[i] >0.2 & mardf$avg_cloglog_97.0413[i] <= 0.3){
+      mardf$bin[i] <- 3
+  } else if (mardf$avg_cloglog_97.0413[i] >0.3 & mardf$avg_cloglog_97.0413[i] <= 0.4){
+      mardf$bin[i] <- 4
+  } else if (mardf$avg_cloglog_97.0413[i] >0.4 & mardf$avg_cloglog_97.0413[i] <= 0.5){
+      mardf$bin[i] <- 5
+  } else if (mardf$avg_cloglog_97.0413[i] >0.5 & mardf$avg_cloglog_97.0413[i] <= 0.6){
+      mardf$bin[i] <- 6
+  } else if (mardf$avg_cloglog_97.0413[i] >0.6 & mardf$avg_cloglog_97.0413[i] <= 0.7){
+      mardf$bin[i] <- 7
+  } else if (mardf$avg_cloglog_97.0413[i] >0.7 & mardf$avg_cloglog_97.0413[i] <= 0.8){
+      mardf$bin[i] <- 8
+  } else if (mardf$avg_cloglog_97.0413[i] >0.8 & mardf$avg_cloglog_97.0413[i] <= 0.9){
+      mardf$bin[i] <- 9
+  } else if (mardf$avg_cloglog_97.0413[i] >0.9){
+      mardf$bin[i] <- 10
+  }
+}
+
+mardf$bin <- as.factor(mardf$bin)
+
+brk <- seq(0.1, 1, 0.1) #bins you want
+brkl <- as.character(brk)
+
+p <- ggplot(mardf, aes(x = avg_cloglog_97.0413, fill = bin))+ geom_histogram(breaks = brk) + theme_classic() + scale_fill_viridis_d(option = "inferno", direction = -1) + theme(legend.position = "none", plot.title = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank()) + scale_x_continuous(breaks=seq(0,1,0.1))
+png(paste0("../output/maxent/paper_plots/histogram_", mthna, "_", dep, ".png"), height = 1024, width = 1024)
+p
+dev.off()
+
+```
+
+#selected maps pt2
+
+get unique depths
+```{r}
+mar <- read.csv("../output//maxent/paper_plots/prob_by_depth/mar_u0.4s_dis_prob_cells.csv", header = TRUE)
+mar <- mar[!(mar$Depth =="1151.99" | mar$Depth =="1265.86" | mar$Depth =="1387.38"), ]
+mar$Probability <- gsub(">0.9", "0.9 - 1", mar$Probability)
+mar$Depth <- as.factor(mar$Depth)
+flevelsm <- levels(mar$Depth)
+```
+
+```{r}
+library(raster)
+library(rgdal)
+library(rasterVis)
+library(colorspace)
+
+aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
+ss <- raster::intersect(s, mcp)  #trim land file to mcp
+
+#selmth <- c("mar", "jun", "sep", "dec")
+#seldep <- c("11.7737", "61.1128", "199.79", "628.026")
+
+mar11 <- raster("../output/maxent/predictions/mar/avg_cloglog_11.7737.asc")
+mar61 <- raster("../output/maxent/predictions/mar/avg_cloglog_61.1128.asc")
+mar199 <- raster("../output/maxent/predictions/mar/avg_cloglog_199.79.asc")
+mar628 <- raster("../output/maxent/predictions/mar/avg_cloglog_628.026.asc")
+mar11@crs <- acrs
+mar61@crs <- acrs
+mar199@crs <- acrs
+mar628@crs <- acrs
+mar11 <- raster::mask(mar11, mcp)
+mar61 <- raster::mask(mar61, mcp)
+mar199 <- raster::mask(mar199, mcp)
+mar628 <- raster::mask(mar628, mcp)
+
+jun11 <- raster("../output/maxent/predictions/jun/avg_cloglog_11.7737.asc")
+jun61 <- raster("../output/maxent/predictions/jun/avg_cloglog_61.1128.asc")
+jun199 <- raster("../output/maxent/predictions/jun/avg_cloglog_199.79.asc")
+jun628 <- raster("../output/maxent/predictions/jun/avg_cloglog_628.026.asc")
+jun11@crs <- acrs
+jun61@crs <- acrs
+jun199@crs <- acrs
+jun628@crs <- acrs
+jun11 <- raster::mask(jun11, mcp)
+jun61 <- raster::mask(jun61, mcp)
+jun199 <- raster::mask(jun199, mcp)
+jun628 <- raster::mask(jun628, mcp)
+
+sep11 <- raster("../output/maxent/predictions/sep/avg_cloglog_11.7737.asc")
+sep61 <- raster("../output/maxent/predictions/sep/avg_cloglog_61.1128.asc")
+sep199 <- raster("../output/maxent/predictions/sep/avg_cloglog_199.79.asc")
+sep628 <- raster("../output/maxent/predictions/sep/avg_cloglog_628.026.asc")
+sep11@crs <- acrs
+sep61@crs <- acrs
+sep199@crs <- acrs
+sep628@crs <- acrs
+sep11 <- raster::mask(sep11, mcp)
+sep61 <- raster::mask(sep61, mcp)
+sep199 <- raster::mask(sep199, mcp)
+sep628 <- raster::mask(sep628, mcp)
+
+dec11 <- raster("../output/maxent/predictions/dec/avg_cloglog_11.7737.asc")
+dec61 <- raster("../output/maxent/predictions/dec/avg_cloglog_61.1128.asc")
+dec199 <- raster("../output/maxent/predictions/dec/avg_cloglog_199.79.asc")
+dec628 <- raster("../output/maxent/predictions/dec/avg_cloglog_628.026.asc")
+dec11@crs <- acrs
+dec61@crs <- acrs
+dec199@crs <- acrs
+dec628@crs <- acrs
+dec11 <- raster::mask(dec11, mcp)
+dec61 <- raster::mask(dec61, mcp)
+dec199 <- raster::mask(dec199, mcp)
+dec628 <- raster::mask(dec628, mcp)
+
+#2d plots
+mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
+mar2d <- raster::mask(mar2d, mcp)
+mar2d@crs <- acrs
+jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
+jun2d <- raster::mask(jun2d, mcp)
+jun2d@crs <- acrs
+sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
+sep2d <- raster::mask(sep2d, mcp)
+sep2d@crs <- acrs
+dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")
+dec2d <- raster::mask(dec2d, mcp)
+dec2d@crs <- acrs
+
+#mar2d, jun2d, sep2d, dec2d,
+
+pstack <- stack(mar11, jun11, sep11, dec11, mar61, jun61, sep61, dec61, mar199, jun199, sep199, dec199, mar628, jun628, sep628, dec628) #stack the layers - printed row-wise
+
+rws <- c("11.7737 m", "61.1128 m", "199.70 m", "628.026 m", "", "", "", "", "", "", "", "", "", "", "", "")
+cls <- c("March", "June", "September", "December", "", "", "", "", "", "", "", "", "", "", "", "")
+myTheme <- rasterTheme(region = inferno(rev(10)))
+my.at <- seq(0, 1, by = 0.1)
+
+p <- levelplot(pstack, ylab = NULL, xlab = NULL, colorkey=list(space="bottom", at = my.at), par.settings = myTheme, layout = c(4, 4), margin = FALSE, scales=list(draw=FALSE), col.regions=rev(inferno(10)), at= my.at, names.attr = cls) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+png("../output/maxent/paper_plots/bydepth.png", height = 1024, width = 1024)
+p
+dev.off()
+
+```
+
+# variable maps ALSO FOR P2
+- temp
+- chl 
+- o2
+- sal
+- 4 different months
+- relationship of variables to each other
+- pick a depth (say the 11.7737 m)
+
+months are mar, jun, sep, dec
+pick 3 diff years say 1998, 2006, 2013
+
+```{r}
+aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
+ss <- raster::intersect(s, mcp)  #trim land file to mcp
+
+yrlst <- c("1998", "2006", "2013")
+mthlst <- c("03", "06", "09", "12")
+varlst <- c("temp", "salinity", "o2")
+
+for (y in 1:length(yrlst)){
+  yr <- yrlst[y]
+  fdr <- paste0("../data/env/ascii/ascii", yr, "/")
+  for (m in 1:length(mthlst)){
+    mt <- mthlst[m]
+    for (v in 1:length(varlst)){
+      var <- varlst[v]
+      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_11.7737.asc"))
+      lyrnm <- names(ras)
+      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      ras <- projectRaster(ras, aea)
+      ras <- raster::mask(ras, mcp)
+
+      p <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))
+
+      png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
+      print(p)
+      dev.off()
+    }
+  }
+}
+
+```
+
+#variable interactions... KEEP FOR P2
+
+```{r}
+yrlst <- c("1998", "2006", "2013")
+mthlst <- c("03", "06", "09", "12")
+varlst <- c("temp", "salinity", "o2")
+scbr <- stack()
+
+for (y in 1:length(yrlst)){
+  yr <- yrlst[y]
+  fdr <- paste0("../data/env/ascii/ascii", yr, "/")
+  for (m in 1:length(mthlst)){
+    mt <- mthlst[m]
+    for (v in 1:length(varlst)){
+      var <- varlst[v]
+      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_11.7737.asc"))
+      lyrnm <- names(ras)
+      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      ras <- projectRaster(ras, aea)
+      ras <- raster::mask(ras, mcp)
+      scbr <- stack(scbr, ras)
+    }
+  }
+}
+a <- xyplot(X2013_12_temp.nc_11.7737~X2013_12_salinity.nc_11.7737, data = scbr, auto.key = list(space='right'))
+b <- xyplot(X2013_06_temp.nc_11.7737~X2013_06_salinity.nc_11.7737, data = scbr, auto.key = list(space='right'))
+
+a
+b
+```
+
+```{r}
+splom(scbr)
+```
+
+#2d plots - part 2
+
+the ggplot experiment....
+
+```{r}
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
+ss <- raster::intersect(s, mcp)  #trim to mcp
+
+mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
+apr2d <- raster("../output/maxent/paper_plots/hotspot_mapping/apr_bin_0.5_2d_single.asc")
+may2d <- raster("../output/maxent/paper_plots/hotspot_mapping/may_bin_0.5_2d_single.asc")
+jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
+jul2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jul_bin_0.5_2d_single.asc")
+aug2d <- raster("../output/maxent/paper_plots/hotspot_mapping/aug_bin_0.5_2d_single.asc")
+sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
+oct2d <- raster("../output/maxent/paper_plots/hotspot_mapping/oct_bin_0.5_2d_single.asc")
+nov2d <- raster("../output/maxent/paper_plots/hotspot_mapping/nov_bin_0.5_2d_single.asc")
+dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")
+
+#trim to mcp
+mar2d <- raster::mask(mar2d, mcp)
+apr2d <- raster::mask(apr2d, mcp)
+may2d <- raster::mask(may2d, mcp)
+jun2d <- raster::mask(jun2d, mcp)
+jul2d <- raster::mask(jul2d, mcp)
+aug2d <- raster::mask(aug2d, mcp)
+sep2d <- raster::mask(sep2d, mcp)
+oct2d <- raster::mask(oct2d, mcp)
+nov2d <- raster::mask(nov2d, mcp)
+dec2d <- raster::mask(dec2d, mcp)
+
+#add crs
+mar2d@crs <- acrs
+apr2d@crs <- acrs
+may2d@crs <- acrs
+jun2d@crs <- acrs
+jul2d@crs <- acrs
+aug2d@crs <- acrs
+sep2d@crs <- acrs
+oct2d@crs <- acrs
+nov2d@crs <- acrs
+dec2d@crs <- acrs
+
+ggplot() +  ggspatial::geom_spraster_rgb(mar2d) + coord_fixed()
+```

### 2019-10-24 src/plots for paper.Rmd
plot battles continue...

@@
 ```{r permulation plot}
+permu <- read.csv("../output/maxent/models/permutation_avg.csv")
+
 permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
-perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = -1)
+perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = 1) + theme_half_open(12) + 
 png("../output/maxent/paper_plots/permutation_plot.png")
@@
 
+now just mak
+
 #probabilty maps (selected)

### 2019-10-24 src/testing_file.Rmd
plot battles continue...

@@
 
-the ggplot experiment....
-
-```{r}
+the cowplot experiment
+
+```{r}
+library(raster)
+library(rgdal)
+library(ggplot2)
+
+aea <- raster("../output/env/aea.tif") #the correct CRS
 s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
@@
 
-ggplot() +  ggspatial::geom_spraster_rgb(mar2d) + coord_fixed()
-```
+#ratify to categorical
+mar2d <- ratify(mar2d)
+rat <- levels(mar2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(mar2d) <- rat
+apr2d <- ratify(apr2d)
+rat <- levels(apr2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(apr2d) <- rat
+may2d <- ratify(may2d)
+rat <- levels(may2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(may2d) <- rat
+jun2d <- ratify(jun2d)
+rat <- levels(jun2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(jun2d) <- rat
+jul2d <- ratify(jul2d)
+rat <- levels(jul2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(jul2d) <- rat
+aug2d <- ratify(aug2d)
+rat <- levels(aug2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(aug2d) <- rat
+sep2d <- ratify(sep2d)
+rat <- levels(sep2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(sep2d) <- rat
+oct2d <- ratify(oct2d)
+rat <- levels(oct2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(oct2d) <- rat
+nov2d <- ratify(nov2d)
+rat <- levels(nov2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(nov2d) <- rat
+dec2d <- ratify(dec2d)
+rat <- levels(dec2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(dec2d) <- rat
+
+```
+
+now the histograms
+
+```{r}
+mar <- read.csv("../output/maxent/paper_plots/prob_by_depth/mar_dis_prob_cells.csv", header = TRUE)
+mar <- mar[!(mar$Depth =="1151.99" | mar$Depth =="1265.86" | mar$Depth =="1387.38"), ]
+mar$Probability <- gsub(">0.9", "0.9 - 1", mar$Probability)
+mar$Depth <- as.factor(mar$Depth)
+flevelsmar <- levels(mar$Depth)
+flevelsmar <- rev(flevelsmar)
+
+apr <- read.csv("../output/maxent/paper_plots/prob_by_depth/apr_dis_prob_cells.csv", header = TRUE)
+apr <- apr[!(apr$Depth =="1151.99" | apr$Depth =="1265.86" | apr$Depth =="1387.38"), ]
+apr$Probability <- gsub(">0.9", "0.9 - 1", apr$Probability)
+apr$Depth <- as.factor(apr$Depth)
+flevelsapr <- levels(apr$Depth)
+flevelsapr <- rev(flevelsapr)
+
+may <- read.csv("../output/maxent/paper_plots/prob_by_depth/may_dis_prob_cells.csv", header = TRUE)
+may <- may[!(may$Depth =="1151.99" | may$Depth =="1265.86" | may$Depth =="1387.38"), ]
+may$Probability <- gsub(">0.9", "0.9 - 1", may$Probability)
+may$Depth <- as.factor(may$Depth)
+flevelsmay <- levels(may$Depth)
+flevelsmay <- rev(flevelsmay)
+
+jun <- read.csv("../output/maxent/paper_plots/prob_by_depth/jun_dis_prob_cells.csv", header = TRUE)
+jun <- jun[!(jun$Depth =="1151.99" | jun$Depth =="1265.86" | jun$Depth =="1387.38"), ]
+jun$Probability <- gsub(">0.9", "0.9 - 1", jun$Probability)
+jun$Depth <- as.factor(jun$Depth)
+flevelsjun <- levels(jun$Depth)
+flevelsjun <- rev(flevelsjun)
+
+jul <- read.csv("../output/maxent/paper_plots/prob_by_depth/jul_dis_prob_cells.csv", header = TRUE)
+jul <- jul[!(jul$Depth =="1151.99" | jul$Depth =="1265.86" | jul$Depth =="1387.38"), ]
+jul$Probability <- gsub(">0.9", "0.9 - 1", jul$Probability)
+jul$Depth <- as.factor(jul$Depth)
+flevelsjul <- levels(jul$Depth)
+flevelsjul <- rev(flevelsjul)
+
+aug <- read.csv("../output/maxent/paper_plots/prob_by_depth/aug_dis_prob_cells.csv", header = TRUE)
+aug <- aug[!(aug$Depth =="1151.99" | aug$Depth =="1265.86" | aug$Depth =="1387.38"), ]
+aug$Probability <- gsub(">0.9", "0.9 - 1", aug$Probability)
+aug$Depth <- as.factor(aug$Depth)
+flevelsaug <- levels(aug$Depth)
+flevelsaug <- rev(flevelsaug)
+
+sep <- read.csv("../output/maxent/paper_plots/prob_by_depth/sep_dis_prob_cells.csv", header = TRUE)
+sep <- sep[!(sep$Depth =="1151.99" | sep$Depth =="1265.86" | sep$Depth =="1387.38"), ]
+sep$Probability <- gsub(">0.9", "0.9 - 1", sep$Probability)
+sep$Depth <- as.factor(sep$Depth)
+flevelssep <- levels(sep$Depth)
+flevelssep <- rev(flevelssep)
+
+oct <- read.csv("../output/maxent/paper_plots/prob_by_depth/oct_dis_prob_cells.csv", header = TRUE)
+oct <- oct[!(oct$Depth =="1151.99" | oct$Depth =="1265.86" | oct$Depth =="1387.38"), ]
+oct$Probability <- gsub(">0.9", "0.9 - 1", oct$Probability)
+oct$Depth <- as.factor(oct$Depth)
+flevelsoct <- levels(oct$Depth)
+flevelsoct <- rev(flevelsoct)
+
+nov <- read.csv("../output/maxent/paper_plots/prob_by_depth/nov_dis_prob_cells.csv", header = TRUE)
+nov <- nov[!(nov$Depth =="1151.99" | nov$Depth =="1265.86" | nov$Depth =="1387.38"), ]
+nov$Probability <- gsub(">0.9", "0.9 - 1", nov$Probability)
+nov$Depth <- as.factor(nov$Depth)
+flevelsnov <- levels(nov$Depth)
+flevelsnov <- rev(flevelsnov)
+
+dec <- read.csv("../output/maxent/paper_plots/prob_by_depth/dec_dis_prob_cells.csv", header = TRUE)
+dec <- dec[!(dec$Depth =="1151.99" | dec$Depth =="1265.86" | dec$Depth =="1387.38"), ]
+dec$Probability <- gsub(">0.9", "0.9 - 1", dec$Probability)
+dec$Depth <- as.factor(dec$Depth)
+flevelsdec <- levels(dec$Depth)
+flevelsdec <- rev(flevelsdec)
+
+pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsmar, breaks=flevelsmar[seq(1,length(flevelsmar), by = 2)], ) + coord_flip() + labs(title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none", axis.text.x = element_blank())
+
+pa <- ggplot(apr, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsapr, breaks=flevelsapr[seq(1,length(flevelsapr), by = 2)], ) + coord_flip() + labs(title = "April") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none", axis.text.x = element_blank(), axis.text.y = element_blank())
+
+pma <- ggplot(may, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsmay, breaks=flevelsmay[seq(1,length(flevelsmay), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "May") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none", axis.text.x = element_blank(), axis.text.y = element_blank())
+
+pj <- ggplot(jun, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsjun, breaks=flevelsjun[seq(1,length(flevelsjun), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "June") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none", axis.text.x = element_blank(), axis.text.y = element_blank())
+
+pju <- ggplot(jul, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsjul, breaks=flevelsjul[seq(1,length(flevelsjul), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "July") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none", axis.text.x = element_blank(), axis.text.y = element_blank())
+
+pau <- ggplot(aug, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsaug, breaks=flevelsaug[seq(1,length(flevelsaug), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "August") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none")
+
+ps <- ggplot(sep, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelssep, breaks=flevelssep[seq(1,length(flevelssep), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "September") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none", axis.text.y = element_blank())
+
+po <- ggplot(oct, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsoct, breaks=flevelsoct[seq(1,length(flevelsoct), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "October") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none", axis.text.y = element_blank())
+
+pn <- ggplot(nov, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() + scale_fill_viridis(option="magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsnov, breaks=flevelsnov[seq(1,length(flevelsnov), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "November") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none", axis.text.y = element_blank())
+
+pd <- ggplot(dec, aes(Depth, p_cells, fill = Probability)) + geom_col(width = 1) + coord_flip() +  scale_fill_viridis(option = "magma", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsdec, breaks=flevelsdec[seq(1,length(flevelsdec), by = 2)]) + coord_flip() +labs(title = "December") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0), axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none", axis.text.y = element_blank())
+
+pm
+pa
+pma
+pj
+pju
+pau
+ps
+po
+pn
+pd
+```
+
+merge the maps and histograms into one giant plot
+
+```{r}
+maplst <- list(mar2d, pm, apr2d, pa, may2d, pma, jun2d, pj, jul2d, pju, aug2d, pa, sep2d, ps, oct2d, po, nov2d, pn, dec2d, pd)
+
+marmap <- levelplot(mar2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+aprmap <- levelplot(apr2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+maymap <- levelplot(may2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+junmap <- levelplot(jun2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+julmap <- levelplot(jul2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+augmap <- levelplot(aug2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+sepmap <- levelplot(sep2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+octmap <- levelplot(oct2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+novmap <- levelplot(nov2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+decmap <- levelplot(dec2d, col.regions=c('white', 'red'),  colorkey = FALSE, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+
+legend <- get_legend(pd +  guides(color = guide_legend(nrow = 1)) + theme(legend.position = "bottom"))
+
+p <- plot_grid(pm, pa, pma, pj, pju, pau, ps, po, pn, pd, ncol = 5, legend)
+png("../output/maxent/paper_plots/jointtest.png", height = 1024, width = 2048) #, bg = "transparent"
+p
+dev.off()
+```
+
+#table for tss scores etc
+
+```{r}
+dec <- read.csv("../output/maxent/models/decavgAIC.csv", header = TRUE)
+colnames(dec)
+
+our_summary1 <- list("Training AUC" = list("min" = ~ min(dec$train_AUC),
+            "max" = ~ max(dec$train_AUC),
+            "mean (sd)" = ~ qwraps2::mean_sd(dec$train_AUC),
+       "Displacement" =
+       list("min" = ~ min(.data$disp),
+            "median" = ~ median(.data$disp),
+            "max" = ~ max(.data$disp),
+            "mean (sd)" = ~ qwraps2::mean_sd(.data$disp)),
+       "Weight (1000 lbs)" =
+       list("min" = ~ min(.data$wt),
+            "max" = ~ max(.data$wt),
+            "mean (sd)" = ~ qwraps2::mean_sd(.data$wt)),
+       "Forward Gears" =
+       list("Three" = ~ qwraps2::n_perc0(.data$gear == 3),
+            "Four"  = ~ qwraps2::n_perc0(.data$gear == 4),
+            "Five"  = ~ qwraps2::n_perc0(.data$gear == 5))
+       )
+```

### 2019-10-28 src/plots for paper.Rmd
minor plot changes plus some additions...

@@
 
-now just mak
+now create a table to see how many cells there are >0.5 in each month
+
+```{r count cells >0.5}
+
+binaryprob_list <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = ".*_bin_0.5_2d.asc", full.names = TRUE)
+sum2d <- setNames(data.frame(matrix(ncol = 2, nrow = binl)), c("month", "count"))
+
+for (r in 1:length(binaryprob_list)){
+  ras <- raster(binaryprob_list[r])
+  rnm <- names(ras)
+  sum2d$month[r] <- gsub("_bin_0.5_2d", "", x = rnm)
+  sum2d$count[r] <- cellStats(ras, sum, asSample = TRUE)
+}
+
+#order table from march to december
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+sum2d <- sum2d[match(mthord, sum2d$month), ]
+
+#get proportion
+tcs <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE) #file showing how many cells per depth
+tcs <- tcs[-c(1:3), ] #because the top three rows are deeper than the capelin obs...
+totc <- sum(tcs$no_cells)
+
+for (r in 1:nrow(sum2d)){
+  sum2d$proportion[r] <- (sum2d$count[r]/totc)
+}
+
+write.csv(sum2d, "../output/maxent/paper_plots/cell_count_0.5.csv", row.names = FALSE)
+
+```
 

### 2019-10-28 src/testing_file.Rmd
minor plot changes plus some additions...

@@
 
-```{r}
+forget it... quickly done in excel
+
+```{r}
+library(qwraps2)
+
 dec <- read.csv("../output/maxent/models/decavgAIC.csv", header = TRUE)
-colnames(dec)
-
-our_summary1 <- list("Training AUC" = list("min" = ~ min(dec$train_AUC),
-            "max" = ~ max(dec$train_AUC),
-            "mean (sd)" = ~ qwraps2::mean_sd(dec$train_AUC),
-       "Displacement" =
-       list("min" = ~ min(.data$disp),
-            "median" = ~ median(.data$disp),
-            "max" = ~ max(.data$disp),
-            "mean (sd)" = ~ qwraps2::mean_sd(.data$disp)),
-       "Weight (1000 lbs)" =
-       list("min" = ~ min(.data$wt),
-            "max" = ~ max(.data$wt),
-            "mean (sd)" = ~ qwraps2::mean_sd(.data$wt)),
-       "Forward Gears" =
-       list("Three" = ~ qwraps2::n_perc0(.data$gear == 3),
-            "Four"  = ~ qwraps2::n_perc0(.data$gear == 4),
-            "Five"  = ~ qwraps2::n_perc0(.data$gear == 5))
+
+our_summary1 <- list("Training AUC" = list("min" = ~ min(.data$train_AUC),
+            "max" = ~ max(.data$train_AUC),
+            "mean (sd)" = ~ qwraps2::mean_sd(.data$train_AUC)),
+       "Test AUC" =
+       list("min" = ~ min(.data$test_AUC),
+            "median" = ~ median(.data$test_AUC),
+            "max" = ~ max(.data$test_AUC),
+            "mean (sd)" = ~ qwraps2::mean_sd(.data$test_AUC))
        )
-```
+
+whole <-  summary_table(dec, our_summary1)
+whole
+
+dec <- read.csv("../output/maxent/models/decavgAIC.csv", header = TRUE)
+summary(dec)
+```
+
+#no cells >0.5 per month
+
+```{r}
+
+binaryprob_list <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = ".*_bin_0.5_2d.asc", full.names = TRUE)
+sum2d <- setNames(data.frame(matrix(ncol = 2, nrow = binl)), c("month", "count"))
+
+for (r in 1:length(binaryprob_list)){
+  ras <- raster(binaryprob_list[r])
+  rnm <- names(ras)
+  sum2d$month[r] <- gsub("_bin_0.5_2d", "", x = rnm)
+  sum2d$count[r] <- cellStats(ras, sum, asSample = TRUE)
+}
+
+#order table from march to december
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+sum2d <- sum2d[match(mthord, sum2d$month), ]
+
+#get proportion
+tcs <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE) #file showing how many cells per depth
+tcs <- tcs[-c(1:3), ] #because the top three rows are deeper than the capelin obs...
+totc <- sum(tcs$no_cells)
+
+for (r in 1:nrow(sum2d)){
+  sum2d$proportion[r] <- (sum2d$count[r]/totc)
+}
+
+write.csv(sum2d, "../output/maxent/paper_plots/cell_count_0.5.csv", row.names = FALSE)
+
+```
+
+# and get it as a proportion of habitat maybe....
+
+```{r}
+
+```

### 2019-11-10 src/plots for paper.Rmd
updates to the testing file

@@
 
+mthna <- c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
+
 permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
-perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = 1) + theme_half_open(12) + 
+perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = 1) + theme_half_open(12) + scale_x_discrete(limits = mthna) + labs(y = "Permutation Importance (%)")
 png("../output/maxent/paper_plots/permutation_plot.png")

### 2019-11-10 src/testing_file.Rmd
updates to the testing file

@@
 
-# and get it as a proportion of habitat maybe....
-
-```{r}
-
-```
+# map with nafo zons and mcp
+
+```{r}
+library(ggplot2)
+library(raster)
+
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis - aea proj
+nafo <- shapefile("../data/bio/nafo_zones/nafo_zones_wgs84.shp") #in wgs proj
+aea <- raster("../output/env/aea.tif") #the correct projections
+nafoa <- spTransform(nafo, crs(aea)) # nafo to aea
+labels <- layer(sp.text(coo))
+land <- shapefile("../data/env/landp/landpro.shp") #made in arcgis - aea proj
+
+nafoaf <- fortify(nafoa, regions = "ZONE")
+
+ggplot(nafoa) + aes() + geom_polygon()
+
+```
+
+#env plots - variable over depth by month
+
+```{r}
+library(ggplot2)
+library(reshape2)
+
+presab <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+background <- subset(presab, occurrence == "0")
+background$month <- as.factor(background$month)
+background <- subset(background, select = c(month, temp_depth, depth_layer, decimalLongitude, decimalLatitude))
+dl <- unique(background$depth_layer)
+dl <- sort(dl)
+test <- subset(background, month == "3")
+
+ggplot(test, aes(x = temp_depth, y = decimalLatitude)) + geom_point() # aes(color = month, group = month)
+plot(background)
+
+```
+
+#amo boxplots?
+
+```{r}
+amoraw <- read.csv("../data/env/amon.us.data.csv", header = TRUE) 
+colnames(amoraw) <- c("year", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
+amo <- amoraw[amoraw$year >= 1996, ]
+amo <- subset(amo, select = -c(year))
+amo2 <- melt(amo, value.name = "value")
+amo2$variable <- as.integer(amo2$variable)
+
+boxplot(amo2$value ~ amo2$variable)
+
+```

### 2019-11-10 src/testing_file.Rmd
greater than prob maps for comm meeting

@@
 ```
+
+# binned prob maps with land etc for cmeet
+
+```{r bin probabilities}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+mthord <- c("jun") #"mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov",
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+
+binprob <- function(x){
+  ifelse(x > 0.1, 0.1,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_1_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_1_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.2, 0.2,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_2_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_2_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.3, 0.3,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+ writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_3_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_3_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.4, 0.4,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_4_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_4_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.5, 0.5,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_5_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_5_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.6, 0.6,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_6_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_6_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.7, 0.7,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_7_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_7_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.8, 0.8,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_8_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_8_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+binprob <- function(x){
+  ifelse(x >= 0.9, 0.9,
+  ifelse(x <= 0, 0, NA))
+}
+
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+  probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+  writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_9_avgprob_", ascname, ".asc"), overwrite = TRUE)
+  plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mth, "_binned_9_avgprob_", ascname, ".png")), overwrite = TRUE)
+  dev.off() # stops automatic saving of the plot to a png
+}
+
+}
+
+```
+
+```{r}
+library(raster)
+library(rasterVis)
+
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+aea <- raster("../output/env/aea.tif") 
+acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
+ss <- raster::intersect(s, mcp)  #trim to mcp
+
+binlst <- list.files("../output/maxent/paper_plots/prob_by_depth", pattern = "^comm_jun_binned.*avgprob_avg_cloglog_97.0413.asc", full.names = TRUE)
+#binlst <- binlst[-2]
+
+mstack <- stack()
+
+for(m in 1:length(binlst)){
+  map <- raster(binlst[m])
+  map <- raster::mask(map, mcp)
+  map@crs <- acrs
+  mn <- names(map)
+
+  # map <- ratify(map)
+  # rat <- levels(map)[[1]]
+  # rat$suitability <- c("val")
+  # levels(map) <- rat
+  png(paste0("../output/maxent/paper_plots/prob_by_depth/comm_", mn, ".png")) #this prints a png of the plot
+  p <- levelplot(map, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE, colorkey = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+  print(p)
+  dev.off() #this turns off the print command
+  mstack <- stack(mstack, map)
+}
+
+#levelplot(mstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), layout = c(5, 2), names.attr = c("< 0.1", "0.1 - 0.2", "0.2 - 0.3", "0.3 - 0.4", "0.4 - 0.5", "0.5 - 0.6", "0.6 - 0.7", "07 - 0.8", "0.8 - 0.9"), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+#plot(p)
+
+levelplot(map, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), margin = FALSE, colorkey = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+
+```

### 2019-11-25 src/plots for paper.Rmd
plot updates

@@
 ```
+
+#boxplots of env variables - part 2
+
+```{r}
+library(ggplot2)
+library(gridExtra)
+library(dplyr)
+
+presab <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+background <- subset(presab, occurrence == "0")
+
+background <- filter(background, month !=1) #remove jan
+background <- filter(background, month !=2) #remove feb
+
+background$month <- factor(background$month) # converts month to a categorical variable
+
+```
+
+```{r}
+box_temp <- ggplot(data = background, aes(x = month, y = temp_depth)) + geom_boxplot() + xlab("Month") + ylab("Temperature (Kelvin)") + theme_classic()
+
+box_chl <- ggplot(data = background, aes(x = month, y = chl_surface)) + geom_boxplot() + xlab("Month") + ylab("Chlorophyll Concentration (mmol.m-3)") + theme_classic()
+
+box_o2 <- ggplot(data = background, aes(x = month, y = o2_depth)) + geom_boxplot() + xlab("Month") + ylab("Dissolved Oxygen (mmol.m-3)") + theme_classic()
+
+box_sal <- ggplot(data = background, aes(x = month, y = salinity_depth)) + geom_boxplot() + xlab("Month") + ylab("Salinity Concentration (PSU)") + theme_classic()
+
+box_naos <- ggplot(data = background, aes(x = month, y = nao_sample)) + geom_boxplot() + xlab("Month") + ylab("NAO (Sampling Month)") + theme_classic()
+
+box_amos <- ggplot(data = background, aes(x = month, y = amo_sample)) + geom_boxplot() + xlab("Month") + ylab("AMO (Sampling Month)") + theme_classic()
+
+box_naop <- ggplot(data = background, aes(x = month, y = nao_prev)) + geom_boxplot() + xlab("Month") + ylab("NAO (Previous Month)") + theme_classic()
+
+box_naow <- ggplot(data = background, aes(x = month, y = nao_winter)) + geom_boxplot() + xlab("Month") + ylab("NAO (Winter)") + theme_classic()
+
+box_amow <- ggplot(data = background, aes(x = month, y = amo_winter)) + geom_boxplot() + xlab("Month") + ylab("AMO (Winter)") + theme_classic()
+
+box_all <- grid.arrange(box_chl, box_o2, box_sal, box_temp, ncol=2)
+ggsave("../output/maxent/paper_plots/env_boxplots.png", plot = box_all, device = png(), dpi = 400)
+
+box_osc <- grid.arrange(box_naos, box_amos, box_naow, box_amow, box_naop, ncol = 3)
+ggsave("../output/maxent/paper_plots/env_boxplots_osc.png", plot = box_osc, device = png(), dpi = 400)
+
+box_aos <- grid.arrange(box_chl, box_o2, box_sal, box_temp,box_naos, box_amos, box_naow, box_amow, box_naop, ncol = 3)
+ggsave("../output/maxent/paper_plots/env_boxplots_allos.png", plot = box_aos, device = png(), dpi = 400)
+
+```

### 2019-11-25 src/testing_file.Rmd
plot updates

@@
 
-```{r bin probabilities}
+```{r bin probabilities greater than}
 outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
@@
 ```
+
+#calculate area encompassed by mcp
+
+```{r}
+cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE) #existing file created
+cumst1 <- cumst1[!cumst1$depth == "1387.38", ] #depth not needed
+cumst1 <- cumst1[!cumst1$depth == "1265.86", ] #depth not needed
+cumst1 <- cumst1[!cumst1$depth == "1151.99", ] #depth not needed
+
+totalcell <- sum(cumst1$no_cells) #136467
+totalcellkm <- totalcell*25 #3,461,675 km2 ?!
+```
+
+#boxplots of env variables - part 2
+
+```{r}
+library(ggplot2)
+library(gridExtra)
+presab <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+background <- subset(presab, occurrence == "0")
+
+background$month <- factor(background$month) # converts month to a categorical variable
+
+```
+
+```{r}
+box_temp <- ggplot(data = background, aes(x = month, y = temp_depth)) + geom_boxplot() + xlab("Month") + ylab("Temperature (Kelvin)") + theme_classic()
+
+box_chl <- ggplot(data = background, aes(x = month, y = chl_surface)) + geom_boxplot() + xlab("Month") + ylab("Chlorophyll Concentration (mmol.m-3)") + theme_classic()
+
+box_o2 <- ggplot(data = background, aes(x = month, y = o2_depth)) + geom_boxplot() + xlab("Month") + ylab("Dissolved Oxygen (mmol.m-3)") + theme_classic()
+
+box_sal <- ggplot(data = background, aes(x = month, y = salinity_depth)) + geom_boxplot() + xlab("Month") + ylab("Salinity Concentration (PSU)") + theme_classic()
+
+box_naos <- ggplot(data = background, aes(x = month, y = nao_sample)) + geom_boxplot() + xlab("Month") + ylab("NAO (Sampling Month)") + theme_classic()
+
+box_amos <- ggplot(data = background, aes(x = month, y = amo_sample)) + geom_boxplot() + xlab("Month") + ylab("AMO (Sampling Month)") + theme_classic()
+
+box_naop <- ggplot(data = background, aes(x = month, y = nao_prev)) + geom_boxplot() + xlab("Month") + ylab("NAO (Previous Month)") + theme_classic()
+
+box_naow <- ggplot(data = background, aes(x = month, y = nao_winter)) + geom_boxplot() + xlab("Month") + ylab("NAO (Winter)") + theme_classic()
+
+box_amow <- ggplot(data = background, aes(x = month, y = amo_winter)) + geom_boxplot() + xlab("Month") + ylab("AMO (Winter)") + theme_classic()
+
+box_all <- grid.arrange(box_chl, box_o2, box_sal, box_temp, ncol=2)
+ggsave("../output/maxent/paper_plots/env_boxplots.png", plot = box_all, device = png(), dpi = 400)
+
+box_osc <- grid.arrange(box_naos, box_amos, box_naow, box_amow, box_naop, ncol = 2)
+ggsave("../output/maxent/paper_plots/env_boxplots_osc.png", plot = box_osc, device = png(), dpi = 400)
+
+```
+
+#make map with mcp eez and all landmass
+
+```{r}
+library(ggplot2)
+library(raster)
+library(rasterVis)
+
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
+aea <- raster("../output/env/aea.tif") #the correct projections
+land <- shapefile("../data/env/landp/landpro.shp") #made in arcgis - aea proj
+eez <- shapefile("../data/env/eez/maybethere.shp") # made in arc - aea proj
+
+eezmap <- plot(mcp)
+plot(eez, add = TRUE)
+plot(land, add = TRUE)
+```
+
+#get median of env. data background points
+
+```{r}
+library(psych)
+
+#presab <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+#background <- subset(presab, occurrence == "0")
+
+temp_depth <- describeBy(background$temp_depth, background$month, mat = TRUE, digits = 2)
+temp_depth <- subset(temp_depth, select = c(group1, min, max, mean, median, range))
+names(temp_depth)[names(temp_depth)=="group1"] <- "month"
+write.csv(temp_depth, "../output/env/env_sumstats/temp_depth_background_summarystats_month.csv", row.names = FALSE)
+
+salinity_depth <- describeBy(background$salinity_depth, background$month, mat = TRUE, digits = 2)
+salinity_depth <- subset(salinity_depth, select = c(group1, min, max, mean, median, range))
+names(salinity_depth)[names(salinity_depth)=="group1"] <- "month"
+write.csv(salinity_depth, "../output/env/env_sumstats/salinity_depth_background_summarystats_month.csv", row.names = FALSE)
+
+o2_depth <- describeBy(background$o2_depth, background$month, mat = TRUE, digits = 2)
+o2_depth <- subset(o2_depth, select = c(group1, min, max, mean, median, range))
+names(o2_depth)[names(o2_depth)=="group1"] <- "month"
+write.csv(o2_depth, "../output/env/env_sumstats/o2_depth_background_summarystats_month.csv", row.names = FALSE)
+
+chl_surface <- describeBy(background$chl_surface, background$month, mat = TRUE, digits = 2)
+chl_surface <- subset(chl_surface, select = c(group1, min, max, mean, median, range))
+names(chl_surface)[names(chl_surface)=="group1"] <- "month"
+write.csv(chl_surface, "../output/env/env_sumstats/chl_surface_background_summarystats_month.csv", row.names = FALSE)
+
+amo_sample <- describeBy(background$amo_sample, background$month, mat = TRUE, digits = 2)
+amo_sample <- subset(amo_sample, select = c(group1, min, max, mean, median, range))
+names(amo_sample)[names(amo_sample)=="group1"] <- "month"
+write.csv(amo_sample, "../output/env/env_sumstats/amo_sample_background_summarystats_month.csv", row.names = FALSE)
+
+amo_winter <- describeBy(background$amo_winter, background$month, mat = TRUE, digits = 2)
+amo_winter <- subset(amo_winter, select = c(group1, min, max, mean, median, range))
+names(amo_winter)[names(amo_winter)=="group1"] <- "month"
+write.csv(amo_winter, "../output/env/env_sumstats/amo_winter_background_summarystats_month.csv", row.names = FALSE)
+
+nao_sample <- describeBy(background$nao_sample, background$month, mat = TRUE, digits = 2)
+nao_sample <- subset(nao_sample, select = c(group1, min, max, mean, median, range))
+names(nao_sample)[names(nao_sample)=="group1"] <- "month"
+write.csv(nao_sample, "../output/env/env_sumstats/nao_sample_background_summarystats_month.csv", row.names = FALSE)
+
+nao_prev <- describeBy(background$nao_prev, background$month, mat = TRUE, digits = 2)
+nao_prev <- subset(nao_prev, select = c(group1, min, max, mean, median, range))
+names(nao_prev)[names(nao_prev)=="group1"] <- "month"
+write.csv(nao_prev, "../output/env/env_sumstats/nao_prev_background_summarystats_month.csv", row.names = FALSE)
+
+nao_winter <- describeBy(background$nao_winter, background$month, mat = TRUE, digits = 2)
+nao_winter <- subset(nao_winter, select = c(group1, min, max, mean, median, range))
+names(nao_winter)[names(nao_winter)=="group1"] <- "month"
+write.csv(nao_winter, "../output/env/env_sumstats/nao_winter_background_summarystats_month.csv", row.names = FALSE)
+
+```
+
+```{r}
+backaug <- subset(background, month == "8")
+backj <- subset(background, month == "7")
+backa <- subset(background, month = "4")
+```

### 2019-11-25 tobedeleted.R

@@ -1,102 +1,0 @@
-modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
-aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
-betavallist <- cbind(aicetc$regularization)
-nfolds <- cbind(c(1:5))
-
-for (year in 1:nrow(modelmonth)){
-  timeslice <- modelmonth[year] #testing timeslice
-  yr <- str_sub(timeslice, -4) #the year being used for testing
-  train <- subset(prback, year == yr) #training data is all years but testing
-  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  
-  for (h in 1:nrow(betavallist)){
-    betamval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nfolds), "replicatetype=crossvalidate", paste0("betamultiplier=", betavallist[h])), path = paste0("../output/maxent/models/", timeslice)) #here the full path is needed
-    
-    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
-    
-    maxresults <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    
-    
-    #count number of non-zero coefficients, excluding the final four rows
-    lambdas_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*lambdas", full.names = TRUE)
-    samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)
-    for(f in 1:length(nfolds)){
-      fn <- f
-      modn <- fn-1
-      lamf <- lambdas_list[f]
-      sampf <- samppred_list[f]
[…]

### 2020-01-14 src/testing_file.Rmd
ascii combine and separate prep files (for running in bash)

@@
 ```
+
+#file renaming for the merge and split .csv files
+
+Need csv file of depth and depthlayerno (depth_bin.csv in data/env)
+
+Str sub to split filename…
+	- Year month variable
+	- Depth
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+
+Then use a loop
+
+for I in length(csv)
+Csv_dep == str_sub
+
+For j in length(list)
+	Dirdep ==
+	If csv == dirdep
+	Xx <- list[j]
+	Yy <- Str_sub Year month variable
+	Layerno <-
+
+	File.rename(xx, paste0(YY, "_", layerno, ".asc")
+
+	    filedep <- str_sub(filen, start = 1, end = 11)
+
+```{r}
+
+depthlst <- read.csv("../data/env/depth_bin.csv", header = TRUE)
+asclst <- list.files("../data/test/ASCII", full.names = FALSE)
+
+for (d in 1:length(depthlst)){
+  csvdep <- depthlst$depthlayerno[d]
+  for (f in 1:length(asclst)){
+    filen <- asclst[3]
+    filedep <- str_sub(filen, start = 16, end = -5)
+    if csvdep == filedep{
+
+    }
+
+  }
+}
+```

### 2020-01-16 src/asc_file_database_combined_files.Rmd
update to combined maxent

---
title: "asc_file_database"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Preperation database to extract and stack asc files

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there are probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.

```{r pre-install packages, message=FALSE}
packages <- c("tidyr")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

```{r base list or ascii layers}
ascyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
ascmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")

asclayerall <- crossing(ascyears, ascmonths)
asclayerall$stackname <- paste0("st", asclayerall$ascyears, asclayerall$ascmonths) #this creates the name of the stack that will be used
asclayerall$temp_depth <- NA
asclayerall$chl_depth <- NA
asclayerall$o2_depth <- NA
asclayerall$salinity_depth <- NA
# asclayerall$nao_sample <- NA
# asclayerall$nao_prev <- NA
# asclayerall$nao_winter <- NA
# asclayerall$amo_sample <- NA
# asclayerall$amo_prev <- NA
# asclayerall$amo_winter <- NA

write.csv(asclayerall, "../data/env/test-comb/asc_stacklist_combine.csv", row.names = FALSE)
```

```{r asclist for 1998}

myr <- 2014 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == myr)

ascname <- list.files(paste0("../data/env/ascii/ascii", myr, "/") , pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files(paste0("../data/env/ascii/ascii", myr, "/"), pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

tempvar <- "temp.asc"
chlvar <- "chl.asc"
o2var <- "o2.asc"
salvar <- "salinity.asc"
# naosamplevar <- "naosample"
# amosamplevar <- "amosample"
# naoprevvar <- "naoprev"
# amoprevvar <- "amoprev"
# naowintervar <- "naowinter"
# amowintervar <- "amowinter"
# surfacevar <- "0.50576"

for (i in 1:nrow(asclayer)){
    print(asclayer$stackname[i])
    year <- asclayer$ascyears[i]
    month <- asclayer$ascmonths[i]

    for (k in 1:length(ascname)){
        ascyear <- (sapply(strsplit(ascname[k], "_"), "[[", 1))
        ascmonth <- (sapply(strsplit(ascname[k], "_"), "[[", 2))
        ascvar <- (sapply(strsplit(ascname[k], "_"), "[[", 3))
        ascfile <- ascpath[k]

        if (ascyear == year & ascmonth == month & ascvar == tempvar){
          asclayer$temp_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == salvar){
          asclayer$salinity_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == chlvar){
          asclayer$chl_depth[i] <- ascfile
        } else if(ascyear == year & ascmonth == month & ascvar == o2var){
          asclayer$o2_depth[i] <- ascfile
        # } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
        #   asclayer$nao_sample[i] <- ascfile
        # } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
        #   asclayer$amo_sample[i] <- ascfile
        # } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
        #   asclayer$nao_prev[i] <- ascfile
        # } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
        #   asclayer$amo_prev[i] <- ascfile
        # } else if(ascyear == year & ascvar == naowintervar){
        #   asclayer$nao_winter[i] <- ascfile
        # } else if(ascyear == year & ascvar == amowintervar){
        #   asclayer$amo_winter[i] <- ascfile
        }
    }
  write.csv(asclayer, paste0("../data/env/test-comb/asc_layerlist", myr, ".csv"), row.names = TRUE)
}
```

it may be better to have ascii lists per month (since models are per month)

```{r}
a1998 <- read.csv("../data/env/test-comb/asc_layerlist1998.csv", header = TRUE)
a1999 <- read.csv("../data/env/test-comb/asc_layerlist1999.csv", header = TRUE)
a2000 <- read.csv("../data/env/test-comb/asc_layerlist2000.csv", header = TRUE)
a2001 <- read.csv("../data/env/test-comb/asc_layerlist2001.csv", header = TRUE)
a2002 <- read.csv("../data/env/test-comb/asc_layerlist2002.csv", header = TRUE)
a2003 <- read.csv("../data/env/test-comb/asc_layerlist2003.csv", header = TRUE)
a2004 <- read.csv("../data/env/test-comb/asc_layerlist2004.csv", header = TRUE)
a2005 <- read.csv("../data/env/test-comb/asc_layerlist2005.csv", header = TRUE)
a2006 <- read.csv("../data/env/test-comb/asc_layerlist2006.csv", header = TRUE)
a2007 <- read.csv("../data/env/test-comb/asc_layerlist2007.csv", header = TRUE)
a2008 <- read.csv("../data/env/test-comb/asc_layerlist2008.csv", header = TRUE)
a2009 <- read.csv("../data/env/test-comb/asc_layerlist2009.csv", header = TRUE)
a2010 <- read.csv("../data/env/test-comb/asc_layerlist2010.csv", header = TRUE)
a2011 <- read.csv("../data/env/test-comb/asc_layerlist2011.csv", header = TRUE)
a2012 <- read.csv("../data/env/test-comb/asc_layerlist2012.csv", header = TRUE)
a2013 <- read.csv("../data/env/test-comb/asc_layerlist2013.csv", header = TRUE)
a2014 <- read.csv("../data/env/test-comb/asc_layerlist2014.csv", header = TRUE)

as <- rbind(a1998, a1999, a2000, a2001, a2002, a2003, a2004, a2005, a2006, a2007, a2008, a2009, a2010, a2011, a2012, a2013, a2014)

asm <- split(as, as$ascmonths)
write.csv(asm$`1`, "../data/env/test-comb/asc_layerlistjan.csv", row.names = TRUE)
write.csv(asm$`2`, "../data/env/test-comb/asc_layerlistfeb.csv", row.names = TRUE)
write.csv(asm$`3`, "../data/env/test-comb/asc_layerlistmar.csv", row.names = TRUE)
write.csv(asm$`4`, "../data/env/test-comb/asc_layerlistapr.csv", row.names = TRUE)
write.csv(asm$`5`, "../data/env/test-comb/asc_layerlistmay.csv", row.names = TRUE)
write.csv(asm$`6`, "../data/env/test-comb/asc_layerlistjun.csv", row.names = TRUE)
write.csv(asm$`7`, "../data/env/test-comb/asc_layerlistjul.csv", row.names = TRUE)
write.csv(asm$`8`, "../data/env/test-comb/asc_layerlistaug.csv", row.names = TRUE)
write.csv(asm$`9`, "../data/env/test-comb/asc_layerlistsep.csv", row.names = TRUE)
write.csv(asm$`10`, "../data/env/test-comb/asc_layerlistoct.csv", row.names = TRUE)
write.csv(asm$`11`, "../data/env/test-comb/asc_layerlistnov.csv", row.names = TRUE)
write.csv(asm$`12`, "../data/env/test-comb/asc_layerlistdec.csv", row.names = TRUE)
```

### 2020-01-16 src/comb-maxent.Rmd
update to combined maxent

--- 2019-05-22 src/maxent_avec_xvalidation.Rmd
@@
----
-title: "maxent model avec xvalidation (year-based test training)"
-author: "Samantha Andrews"
-output: html_notebook
----
-
-# Overview
-A note to anyone who might happen to stumble across this...
-I am a beginner in R and have had no exposure to similar languages.
-I don't know what I'm doing.
-The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.
-
-Built with 'r getRversion()'..
-
-# Package dependencies
-You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163).
-Note this function checks to see if the packages are installed first before loading.
   
@@
 
-#load the presence-background dataset 
+load the presence-background dataset 
 ("../output/bio/presab_nafo_yr.csv")
+
+select month
+```{r month of interest -change}
+mthna <- "mar"
+mthno <- "3"
+```
 
@@
 
-#monthly model data selection and prep
+monthly model data selection and prep
 For each monthly model, subset the data to the month you want, and the variables you want (Based on spearmans and VIF)
@@
 ```{r}
-janvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_winter")
-marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
-decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
 
-mthnavar_lst <- list(jan = janvar, mar = marvar, apr = aprvar, may = mayvar, jun = junvar, jul = julvar, aug = augvar, sep = sepvar, oct = octvar, nov = novvar, dec = decvar)
-```
-
-select month
-```{r month of interest -change}
-mthna <- "jan"
-mthno <- "1"
+mthnavar_lst <- list(mar = marvar, apr = aprvar, may = mayvar, jun = junvar, jul = julvar, aug = augvar, sep = sepvar, oct = octvar, nov = novvar, dec = decvar)
 ```
@@
 
-normal model..
-mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE")) 
-mod12
-
-```{r train and test inc AIC for reg val}
-presyrs <-  #this is to get years with complete presence data
-
-modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
-aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
-betavallist <- cbind(aicetc$regularization)
-
-for (year in 1:nrow(modelmonth)){
-  timeslice <- modelmonth[year] #testing timeslice
-  yr <- str_sub(timeslice, -4) #the year being used for testing
-  train <- subset(prback, year != yr) #training data is all years but testing
-  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-  test <-  subset(prback, year == yr) #creates the testing dataset
-  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
-  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
-  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-
-  for (h in 1:nrow(betavallist)){
-    betaval <-  betavallist[h]
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c('plots=FALSE', "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
-
-    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
-
-    #count number of non-zero coefficients, excluding the final four rows
-    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
-    coefile <- head(coefile, -4)
-    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-    nocoeff <- nrow(coefile2)
-    bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
-    bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
-
-    #no presences used in model (less those used for testing)
-    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    notrainingpoints <- maxresults$X.Training.samples
-    testauc <- maxresults$Test.AUC
-    trainauc <- maxresults$Training.AUC
-
-    #natural log
-    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
-    samplepred$nat_log <- log(samplepred$Raw.prediction)
-    sumsamplepred <- sum(samplepred$nat_log)
-
-   #test auc scores and other lovely things you might want to include
-    ev <- evaluate(p = testp, a= testb, model = mod)
-    evauc <- ev@auc
-    evcor <- ev@cor
-    evpcor <- ev@pcor
-    evnp <- ev@np
-    evna <- ev@na
-    evmaxtprtnr <- threshold(ev, 'spec_sens')
-    evmaxkap <- threshold(ev, 'kappa')
-    evno_omission <- threshold(ev, 'no_omission')
-    evprevalence <- threshold(ev, 'prevalence')
-    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
-    evsensitivity <- threshold(ev, 'sensitivity')
-
-    for (i in 1:nrow(aicetc)){
-    if(aicetc$regularization[i] == betaval){
-      aicetc$train_AUC[i] <- trainauc
-      aicetc$no_trainng_points[i] <- notrainingpoints
-      aicetc$coefficients[i] <- nocoeff
-      aicetc$sum.of.logs[i] <- sumsamplepred
-      aicetc$test_AUC[i] <- evauc
-      aicetc$test_evcor[i] <- evcor
-      aicetc$test_evpcor[i] <- evpcor
-      aicetc$test_nopresence[i] <- evnp
-      aicetc$test_nobackground[i] <- evna
-      aicetc$test_maxsumsenspec[i] <- evmaxtprtnr
-      aicetc$test_maxkap[i] <- evmaxkap
-      aicetc$test_no_omission[i] <- evno_omission
-      aicetc$test_prevalence[i] <- evprevalence
-      aicetc$test_equal_sens_spec[i] <- evequal_sens_spec
-      aicetc$test_sensitivity[i] <- evsensitivity
-      aicetc$test_bic[i] <- bic #not sure this is correct...
-      aicetc$beta[i] <- betaval
-      #calculate minimal predicted area using ecospat
-      aicetc$mpa_0.9_cum[i] <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
-
-    aicetc$AIC[i] <- ((2*aicetc$coefficients[i])-(2*aicetc$sum.of.logs[i]))
-    aicetc$AICc[i] <- (-2*(aicetc$sum.of.logs[i])+((2*aicetc$coefficients[i])*(aicetc$no_trainng_points[i]/(aicetc$no_trainng_points[i] - aicetc$coefficients[i]-1))))
-    }
-   }
-  }
-
-write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
-
-}
-
-```
-
-#model selection
-next step... select the models you want for each year and run those models again (primarily based on lowest AICc but check test AUC and other things too)
-
-you need to go through each aic file manually....
-
-KEEP A NOTE OF THESE IN ONENOTE
-
-#run final model
 ```{r final models plus predict}
@@
 aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
-unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
+unlayer <- read.csv(paste0("../data/env/test-comb/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
 
@@
 
-  if (yr == "1998"){
-    betaval <- 0.25
-  } else if (yr == "2001"){
-    betaval <- 0.75
-  } else if (yr == "2002" | yr == "2003"| yr == "2004"| yr == "2006"| yr == "2011"){
+  if (yr == "2003"){
     betaval <- 1
-  } else if (yr == "2005"){
+  } else if (yr == "1999" | yr == "2002" | yr == "2008" | yr == "2009") {
     betaval <- 1.25
-  } else if (yr == "1999" | yr == "2000"| yr == "2009"){
+  } else if (yr == "2005" | yr == "2006"){
     betaval <- 1.5
-  } else if (yr == "2007"){
+  } else if (yr == "2000"){
     betaval <- 1.75
+  } else if (yr == "2001" | yr == "2007"){
+    betaval <- 2.25
   } else if (yr == "2010"){
-    betaval <- 2
-  } else if (yr == "2008"){
-    betaval <- 2.25
+    betaval <- 3.5
   }
 
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/test-comb/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
+
+    saveRDS(mod, file= paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))
 
@@
     predBg <- predBg$Cloglog
+    source("../src/contBoyce.R")
     cbi <- contBoyce(pres = predPres, bg = predBg)
@@
 
-    write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+    write.csv(aicetc, paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
 
+    mod <- readRDS(paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))
+
+    yr <- as.numeric(yr)
  unlayer_yr <- subset(unlayer, ascyears == yr)
-    unlayer_yr <- unlayer_yr[-c(1), ]
+    #unlayer_yr <- unlayer_yr[-c(1), ]
     unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
     stackdeplst <- list()
-    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
+    stackdep <- subset(unlayer_yr, select = c(stackname))
     stackdeplst[[year]] <- stackdep
-    write.csv(stackdep, paste0("../output/maxent/models/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)
+    write.csv(stackdep, paste0("../output/maxent/models/test-comb/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)
     aea <- raster("../output/env/aea.tif")
@@
       names(temp_depth) <- "temp_depth"
-      temp_depth <- projectRaster(temp_depth, aea)
-      chl_surface <- raster(unlayer_yr$chl_surface[td])
+      #temp_depth <- projectRaster(temp_depth, aea)
+      chl_surface <- raster(unlayer_yr$chl_depth[td])
       crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
       names(chl_surface) <- "chl_surface"
-      chl_surface <- projectRaster(chl_surface, aea)
+      #chl_depth <- projectRaster(chl_depth, aea)
       o2_depth <- raster(unlayer_yr$o2_depth[td])
@@
       names(o2_depth) <- "o2_depth"
-      o2_depth <- projectRaster(o2_depth, aea)
+      #o2_depth <- projectRaster(o2_depth, aea)
       salinity_depth <- raster(unlayer_yr$salinity_depth[td])
@@
       names(salinity_depth) <- "salinity_depth"
-      salinity_depth <- projectRaster(salinity_depth, aea)
-      nao_sample <- raster(unlayer_yr$nao_sample[td])
-      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
-      names(nao_sample) <- "nao_sample"
-      nao_sample <- projectRaster(nao_sample, aea)
-      nao_prev <- raster(unlayer_yr$nao_prev[td])
-      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
-      names(nao_prev) <- "nao_prev"
-      nao_prev <- projectRaster(nao_prev, aea)
-      nao_winter <- raster(unlayer_yr$nao_winter[td])
-      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
-      names(nao_winter) <- "nao_winter"
-      nao_winter <- projectRaster(nao_winter, aea)
-      amo_sample <- raster(unlayer_yr$amo_sample[td])
-      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
-      names(amo_sample) <- "amo_sample"
-      amo_sample <- projectRaster(amo_sample, aea)
-      amo_winter <- raster(unlayer_yr$amo_winter[td])
-      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
-      names(amo_winter) <- "amo_winter"
-      amo_winter <- projectRaster(amo_winter, aea)
+      #salinity_depth <- projectRaster(salinity_depth, aea)
+      # nao_sample <- raster(unlayer_yr$nao_sample[td])
+      # crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      # names(nao_sample) <- "nao_sample"
+      # nao_sample <- projectRaster(nao_sample, aea)
+      # nao_prev <- raster(unlayer_yr$nao_prev[td])
+      # crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      # names(nao_prev) <- "nao_prev"
+      # nao_prev <- projectRaster(nao_prev, aea)
+      # nao_winter <- raster(unlayer_yr$nao_winter[td])
+      # crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      # names(nao_winter) <- "nao_winter"
+      # nao_winter <- projectRaster(nao_winter, aea)
+      # amo_sample <- raster(unlayer_yr$amo_sample[td])
+      # crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      # names(amo_sample) <- "amo_sample"
+      # amo_sample <- projectRaster(amo_sample, aea)
+      # amo_winter <- raster(unlayer_yr$amo_winter[td])
+      # crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      # names(amo_winter) <- "amo_winter"
+      # amo_winter <- projectRaster(amo_winter, aea)
       stackn <- unlayer_yr$stackname[td]
-      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
+      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface)
 
       #stackSave(test,unlayer_yr$stackname[td])
-      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=cloglog')) 
-      predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
-      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
-      dev.off() # stops automatic saving of the plot to a png
-
-      #mess
-
-      #messev <- dismo::mess(x = stack, v = mess_pr, full = FALSE)
-      #messplot <- plot(messev, png(paste0("../output/maxent/predictions/", mthna, "/", "mess_", stackn, ".png")))
-      #dev.off()
-      # not working how i'd like....
-    }
-}
-
-stdep <- do.call(rbind, stackdeplst)
-write.csv(stdep, paste0("../output/maxent/models/", timeslice, "/stack_depth.csv"), row.names = FALSE)
-```
-
-```{r avg maxentReg}
-
-for (i in 1:length(modelmonth)){
-  timeslice <- modelmonth[i]
-  yr <- str_sub(timeslice, -4) #the year being used for testing
-  if(yr == 1998){
-    aic1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 1999){
-    aic1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2000){
-    aic2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2001){
-    aic2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2002){
-    aic2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2003){
-    aic2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2004){
-    aic2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2005){
-    aic2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2006){
-    aic2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2007){
-    aic2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2008){
-    aic2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2009){
-    aic2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2010){
-    aic2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2011){
-    aic2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2012){
-    aic2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2013){
-    aic2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2014){
-    aic2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-    } else if (yr == 2015){
-    aic2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-    maxr2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
-  }
-}
-
-aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
-maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)
-
-aicm <- colMeans(aic)
-aicm$mod <- "avg"
-avgAIC <- rbind(aic, aicm)
-
-maxr$Species <- 1
-maxrm <- colMeans(maxr)
-maxrm$Species <- "avg"
-maxrm <- rbind(maxr, maxrm)
-
-write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
-write.csv(maxrm, paste0("../output/maxent/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
-```
-
-summary table of average maxent results for each month
-```{r average AIC etc for each month in one table}
-aic_lst <- list.files("../output/maxent/models/", pattern = "avgAIC.csv", full.names = TRUE)
-aicsum <- data.frame(matrix(ncol = 11, nrow = length(aic_lst)))
-colnames(aicsum) <- c("Month", "BetaVal", "AICc", "Test AUC", "Training AUC", "Correlation", "Correlation P-Val", "TSS", "CBI", "No Omissions", "Prevalence")
-
-for (a in 1:length(aic_lst)){
-  aicr <- read.csv(aic_lst[a])
-  mth <- str_sub(aic_lst[a], start = 25, end = 27) #selects the month from the filename
-  aicsum$Month[a] <- mth
-  aicres <- tail(aicr, 1)
-  aicsum$BetaVal[a] <- aicres$beta
-  aicsum$AICc[a] <- aicres$AICc
-  aicsum$`Test AUC`[a] <- aicres$test_AUC
-  aicsum$`Training AUC`[a] <- aicres$train_AUC
-  aicsum$Correlation[a] <- aicres$test_evcor
-  aicsum$`Correlation P-Val` <- aicres$test_evpcor
-  aicsum$TSS[a] <- aicres$tss
-  aicsum$CBI[a] <- aicres$cbi
-  aicsum$`No Omissions`[a] <- aicres$test_no_omission
-  aicsum$Prevalence[a] <- aicres$test_prevalence
-}
-
-mthord <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
-aicsum <- aicsum[match(mthord, aicsum$Month), ] #reorder based on mthord
-write.csv(aicsum, "../output/maxent/models/aicsummary.csv", row.names = FALSE)
-
-```
-
-stacked permutation importance plot on average
-
-```{r permutation & jack data}
-maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
-maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
-perm <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
-colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
-jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
-colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
-
-for (m in 1:length(maxr_lst)){
-  maxr <- read.csv(maxr_lst[m])
-  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
-  perm$Month[m] <- mth
-  jack$Month[m] <- mth
-  maxt <- tail(maxr, 1)
-  perm$amo_sample[m] <- maxt$amo_sample.permutation.importance
-  perm$amo_winter[m] <- maxt$amo_winter.permutation.importance
-  perm$chl_surface[m] <- maxt$chl_surface.permutation.importance
-  perm$nao_prev[m] <- maxt$nao_prev.permutation.importance
-  perm$nao_sample[m] <- maxt$nao_sample.permutation.importance
-  perm$nao_winter[m] <- maxt$nao_winter.permutation.importance
-  perm$o2_depth[m] <- maxt$o2_depth.permutation.importance
-  perm$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
-  perm$temp_depth[m] <- maxt$temp_depth.permutation.importance
-
-  jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
-  jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
-  jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
-  jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
-  jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
-  jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
-  jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
-  jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
-  jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
-}
-
-# now for january
-
-maxr_lstj <- list.files("../output/maxent/models/", pattern = "janavgmaxentresults.csv", full.names = TRUE)
-permj <- data.frame(matrix(ncol = 7, nrow = length(maxr_lstj)))
-colnames(permj) <- c("Month", "chl_surface", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
-jackj <- data.frame(matrix(ncol = 7, nrow = length(maxr_lstj)))
-colnames(jackj) <- c("Month", "chl_surface", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
-
-for (m in 1:length(maxr_lstj)){
-  maxr <- read.csv(maxr_lstj[m])
-  mth <- str_sub(maxr_lstj[m], start = 25, end = 27) #selects the month from the filename
-  permj$Month[m] <- mth
-  jackj$Month[m] <- mth
-  maxt <- tail(maxr, 1)
-  permj$chl_surface[m] <- maxt$chl_surface.permutation.importance
-  permj$nao_sample[m] <- maxt$nao_sample.permutation.importance
-  permj$nao_winter[m] <- maxt$nao_winter.permutation.importance
-  permj$o2_depth[m] <- maxt$o2_depth.permutation.importance
-  permj$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
-  permj$temp_depth[m] <- maxt$temp_depth.permutation.importance
-
-  jackj$chl_surface[m] <- maxt$Training.gain.without.chl_surface
-  jackj$nao_sample[m] <- maxt$Training.gain.without.nao_sample
-  jackj$nao_winter[m] <- maxt$Training.gain.without.nao_winter
-  jackj$o2_depth[m] <- maxt$Training.gain.without.o2_depth
-  jackj$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
-  jackj$temp_depth[m] <- maxt$Training.gain.without.temp_depth
-}
-
-permu <- rbind.fill(perm, permj)
-jackk <- rbind.fill(jack, jackj)
-
-mthord <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
-permu <- permu[match(mthord, permu$Month), ] #reorder based on mthord
-jackk <- jackk[match(mthord, jackk$Month), ] #reorder based on mthord
-
-write.csv(permu, "../output/maxent/models/permutation_avg.csv", row.names = FALSE)
-write.csv(jackk, "../output/maxent/models/jack_avg.csv", row.names = FALSE)
-
-```
-
-```{r}
-maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
-maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
-jack_df <- list()
-maxr_lstj <- list.files("../output/maxent/models/", pattern = "janavgmaxentresults.csv", full.names = TRUE)
-jack_dfj <- list()
-
-for (m in 1:length(maxr_lst)){
-  res <- read.csv(maxr_lst[m], header = TRUE)
-  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
-  res_df <- tail(res, 1)
-  var_df <- res_df[,c(grep("Training.gain.with*",names(res_df)))]
-  var_df$id <- 1:nrow(var_df)
-  var_df <- melt(var_df, id.vars=c("id"))
-  jack_with <- var_df[grep("with.only",var_df$variable),]
-  jack_with$Type_of_Run <- "With only the variable"
-  jack_without <- var_df[grep("without",var_df$variable),]
-  jack_without$Type_of_Run <- "Without the variable"
-  gain_all <- res_df[c(grep("Regularized.",names(res_df)))]
-  gain_all$id <- 1:nrow(gain_all)
-  gain_all <- melt(gain_all, id.vars=c("id"))
-  gain_all$Type_of_Run <- "With all variables"
-  gain_all$variable <- "all_variables"
-  jack_table <- rbind(jack_with,jack_without,gain_all)
-  jack_table$state <- mth
-
-  jack_table$variable <- gsub(".*[.]","",jack_table$variable)
-  jack_df[[m]] <- jack_table
-}
-
-var_jack <- do.call(rbind, jack_df)
-
-for (m in 1:length(maxr_lstj)){
-  res <- read.csv(maxr_lstj[m], header = TRUE)
-  mth <- str_sub(maxr_lstj[m], start = 25, end = 27) #selects the month from the filename
-  res_df <- tail(res, 1)
-  var_df <- res_df[,c(grep("Training.gain.with*",names(res_df)))]
-  var_df$id <- 1:nrow(var_df)
-  var_df <- melt(var_df, id.vars=c("id"))
-  jack_with <- var_df[grep("with.only",var_df$variable),]
-  jack_with$Type_of_Run <- "With only the variable"
-  jack_without <- var_df[grep("without",var_df$variable),]
-  jack_without$Type_of_Run <- "Without the variable"
-  gain_all <- res_df[c(grep("Regularized.",names(res_df)))]
-  gain_all$id <- 1:nrow(gain_all)
-  gain_all <- melt(gain_all, id.vars=c("id"))
-  gain_all$Type_of_Run <- "With all variables"
-  gain_all$variable <- "all_variables"
-  jack_table <- rbind(jack_with,jack_without,gain_all)
-  jack_table$state <- mth
-
-  jack_table$variable <- gsub(".*[.]","",jack_table$variable)
-  jack_dfj[[m]] <- jack_table
-}
-
-var_jackj <- do.call(rbind, jack_dfj)
-
-var_jack$state <- factor(var_jack$state,levels=c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"))
-var_jack$Type_of_Run <- factor(var_jack$Type_of_Run,levels=c("With all variables", "Without the variable","With only the variable"))
-var_jackj$state <- factor(var_jackj$state,levels="jan")
-var_jackj$Type_of_Run <- factor(var_jackj$Type_of_Run,levels=c("With all variables", "Without the variable","With only the variable"))
-a <- subset(var_jack, state == "mar" | state == "apr" |state == "may" | state == "jun" | state == "jul")
-b <- subset(var_jack, state == "aug" | state == "sep" |state == "oct" | state == "nov" | state == "dec")
-c <- subset(var_jackj, state == "jan")
-
-plotlst <- list(a, b, c)
-
-color_type_1 <- c("#000000", "#009E73","#999999")
-
-for (p in 1:length(plotlst)){
-  pt <- p
-  dev.new()
-  jack <- ggplot(plotlst[[p]], aes(variable,value))+
-  geom_bar(aes(fill=Type_of_Run),position="dodge",stat="identity")+ coord_flip()+ facet_grid(state~., scales="free", space = "free")+
-  theme_bw(14)+
-  scale_fill_manual(values=color_type_1,guide=guide_legend(reverse=TRUE))+
-  theme(axis.text=element_text(size=8)) +
-  theme(legend.title=element_blank())+
-  theme(legend.position="top")+
-  theme(strip.text=element_text(size=12))+
-  theme(strip.text.y = element_text(angle = 0)) +
-  labs(x="Variable",y="Jackknife of Regularized Training Gain")
-  ggsave(filename = paste0("../output/maxent/models/jackknife_plot_", pt, ".png"), plot = jack, dpi = 900, height = 9, width = 7)
-  dev.off()
-}
-
-#png("../output/maxent/models/jackknife_plot.png")
-#print(jack)
-#dev.off()
-```
-
-```{r permulation plot}
-permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
-perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") + scale_fill_brewer(palette = "Set1")
-png("../output/maxent/models/permutation_plot.png")
-print(perpl)
-dev.off()
-
-#for jan
-
-permuj <- subset(permu, Month == "jan")
-permelt <- melt(permuj, id = "Month") #need to melt the data into this format for ggplot
-perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") + scale_fill_brewer(palette = "Set1")
-png("../output/maxent/models/permutation_plot_jan.png")
-print(perpl)
-dev.off()
-```
-
-```{r jackknife plot}
-color_type_1 <- c("#000000", "#009E73","#999999")
-jackpl <- ggplot(jackk, aes(x = variable, y = value)) + geom_bar(aes(fill=Type_of_Run),position="dodge",stat="identity")+
-```
-
-#binary maps - individual year layers
-
-there seem to be two primary ways to create binary maps.
-the most common for conservation purposes seems to be to use maxsumspensspec values, however MPA also comes recommended (but is def less common).
-
-the thresholds are in the monthavgAIC.csv file in the model folder (not model/month)
-
-```{r set up threshold binary function}
-binmpa <- function(x) {
-  ifelse(x <  mpa_thr, 0,
-  ifelse(x >=  mpa_thr, 1, NA))
-  }
-```
-
-```{r binary maps on individual layers cloglog}
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
-
-for (modtime in 1:nrow(modelmonth)){
-  timeslice <- modelmonth[modtime]
-  print(timeslice)
-  yr <- str_sub(timeslice, -4)
-  thresholds <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
-  mpa_thr <- thresholds$mpa_0.9_clog
-  for (cum in 1:length(clogasc_list)){
-    cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
-    if(yr == cumascyr){
-      cumasc <- raster(clogasc_list[cum])
-      ascname <- names(cumasc)
-      cum_mpathr <- calc(cumasc, fun = binmpa)
-      writeRaster(cum_mpathr, paste0("../output/maxent/predictions/", mthna,  "/binary_", ascname, ".asc"), overwrite = TRUE)
-      plot(cum_mpathr, png(paste0("../output/maxent/predictions/", mthna, "/binary_", ascname, ".png")), overwrite = TRUE)
+      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=cloglog')) 
+      # predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
+      prplot <- plot(pred, png(paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".png")))
       dev.off() # stops automatic saving of the plot to a png
@@
 }
-}
 ```
-
-now for each year create a 2-d representation of suitable habitat
-
-need to stack the rasters by year
-run cellStats (sum)
-save the output as they year
-
-```{r binary map per year 2D representation}
-
-binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_cloglog.*.asc", full.names = TRUE)
-yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryclog_list)))
-colnames(yrlayer) <- c("layer", "year")
-
-for (yrlyr in 1:length(binaryclog_list)){
-  yearoflyr <- str_sub(binaryclog_list[yrlyr], start = 51, end = 54) #selects the year from the file name
-  yrlayer$year[yrlyr] <- yearoflyr
-  yrlayer$layer[yrlyr] <- binaryclog_list[yrlyr]
-}
-
-y_list <- split(yrlayer, yrlayer$year)
-noyrs <- unique(yrlayer$year)
-
-for (binyr in (1:length(y_list))){
-  binstack <- stack(y_list[[binyr]]$layer)
-  year <- y_list[[binyr]]$year[1]
-  stacksum <- sum(binstack, na.rm = TRUE)
-  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".asc"), overwrite=TRUE)
-  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".png")))
-  dev.off()
-}
-
-```
-
-#using averageing
-
-ok! now to average results:
-
-- the AICforMaxentReg.csv 
-- the maxentResults.csv
-- the asc files per depth layer
-
-open all maxentReg files
-
-for this bit may need to stack rasters by depth... this will need some working with the name
-get unique filenames into a list
-ignore the first X characters in the filename (yr_mthna_) and stack based on the last bit of the filename
-then do something probably with raster package
-
-```{r avg asc layers cloglog}
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
-stdep <- read.csv(paste0("../output/maxent/models/", timeslice, "/stack_depth.csv"), header = TRUE)
-depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
-names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
-depthlayers$depth <- as.character(depthlayers$depth)
-
-colno <- 1
-
-for (d in 1:nrow(depthlayers)){
-  dep <- depthlayers$depth[d]
-  for (a in 1:length(clogasc_list)){
-    asc <- clogasc_list[a]
-    if (grepl(dep, asc)){
-      depthlayers[d , paste0("a", colno)] <- asc
-        colno <- colno+1
-    }
-  }
-}
-write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)
-
-for (x in 1:nrow(depthlayers)){
-  asclay <- depthlayers[x, ]
-  asclay <- asclay[!is.na(asclay)]
-  asclay <- as.data.frame(rbind(asclay))
-  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
-  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
-  asclay$V3 <- as.character(asclay$V3)
-  asclay$V4 <- as.character(asclay$V4)
-  asclay$V5 <- as.character(asclay$V5)
-  asclay$V6 <- as.character(asclay$V6)
-  asclay$V7 <- as.character(asclay$V7)
-  asclay$V8 <- as.character(asclay$V8)
-  asclay$V9 <- as.character(asclay$V9)
-  asclay$V10 <- as.character(asclay$V10)
-  asclay$V11 <- as.character(asclay$V11)
-  asclay$V12 <- as.character(asclay$V12)
-  asclay$V13 <- as.character(asclay$V13)
-  asclay$V14 <- as.character(asclay$V14)
-  asclay$V15 <- as.character(asclay$V15)
-
-  asclaydep <- asclay$V1
-
-  asc1998 <- raster(asclay$V2)
-  asc1999 <- raster(asclay$V3)
-  asc2000 <- raster(asclay$V4)
-  asc2001 <- raster(asclay$V5)
-  asc2002 <- raster(asclay$V6)
-  asc2003 <- raster(asclay$V7)
-  asc2004 <- raster(asclay$V8)
-  asc2005 <- raster(asclay$V9)
-  asc2006 <- raster(asclay$V10)
-  asc2007 <- raster(asclay$V11)
-  asc2008 <- raster(asclay$V12)
-  asc2009 <- raster(asclay$V13)
-  asc2010 <- raster(asclay$V14)
-  asc2011 <- raster(asclay$V15)
-
-  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)
-
-  avg <- mean(stkasc, na.rm = TRUE)
-
-  writeRaster(avg, paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".asc"), overwrite=TRUE)
-  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_", asclaydep, ".png"))) 
-  dev.off()
-}
-```
-
-```{r binary maps on averaged depth layer}
-avgasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
-thresholds <- read.csv(paste0("../output/maxent/models/", mthna, "avgAIC.csv"), header = TRUE)
-thresholds <- tail(thresholds, 1) #subset the last row
-mpa_thr <- thresholds$mpa_0.9_clog
-
-for (avg in 1:length(avgasc_list)){
-    cumasc <- raster(avgasc_list[avg])
-    ascname <- names(cumasc)
-    mpathr <- calc(cumasc, fun = binmpa)
-    mpa_plot <- plot(mpathr, png(paste0("../output/maxent/predictions/", mthna, "/binary_", ascname, ".png")))
-    dev.off() # stops automatic saving of the plot to a png
-    writeRaster(mpathr, paste0("../output/maxent/predictions/", mthna, "/binary_", ascname, ".asc"), overwrite = TRUE)
-}
-
-```
-
-```{r binary map averaged 2D representation}
-
-binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_cloglog.*.asc", full.names = TRUE)
-
-binstack <- stack(binaryclog_list)
-stacksum <- sum(binstack, na.rm = TRUE)
-writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.asc"), overwrite=TRUE)
-plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.png")))
-dev.off()
-
-```
-
-#plots by depth for MPA threshold
-
-first need some statistcs... for each layer how many cells have a value of 1...
-
-```{r calc no presence cells per layer (avg)} 
-ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_cloglog.*asc", full.names = TRUE)
-binaryavgstack <- stack(ascdepth_list)
-stackstat <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-stackstat[ ,1] <- c(1:46)
-colnames(stackstat) <- c("layer", "cell_sum")
-
-for (lyr in 1:nlayers(binaryavgstack)){
-  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
-  stackstat$cell_sum[lyr] <- lyrsum
-}
-
-write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/cellcount_bydepth_avg.csv"), row.names = FALSE)
-```
-
-attach depth in meters (layer_name in depth_bin.csv) to the stackstat based on layer (stackstat) and depthlayerno (depth_bin.csv)
-
-```{r}
-dep <- read.csv("../data/env/depth_bin.csv", header = TRUE) # the depth bin file created way back
-dep <- subset(dep, select = c("depthlayerno", "layer_name"))
-stackstat <-  read.csv(paste0("../output/maxent/predictions/", mthna, "/cellcount_bydepth_avg.csv"), header = TRUE)
-colnames(dep)[names(dep) == "depthlayerno"] <- "layer"
-stackstat <- left_join(stackstat, dep)
-stackstat[ , "layer_name"]=round(stackstat[ , "layer_name"], 2)
-head(stackstat)
-
-```
-
-a crude plot but ok to show slr
-
-```{r graph by depth (avg)}
-bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1))
-trellis.device(device="png", filename = paste0("../output/maxent/predictions/", mthna, "/depthplot_avg.png"))
-print(bydepth)
-dev.off()
-
-```
-
-#probability selected maps
-
-```{r habitat classification on probability - 3 classifications}
-
-binprob <- function(x){
-  ifelse(x >= 0.8, 3,
-  ifelse(x >= 0.7 & x <0.8, 2,
-  ifelse(x >= 0.6 & x <0.7, 1,
-  ifelse(x < 0.6, 0, NA)))) 
-}
-
-  modelmonth <- cbind(paste0(mthna, unique(prback$year)))
-  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
-
-  for (modtime in 1:nrow(modelmonth)){
-    timeslice <- modelmonth[modtime]
-    yr <- str_sub(timeslice, -4)
-    for (cum in 1:length(clogasc_list)){
-      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
-      if(yr == cumascyr){
-        cumasc <- raster(clogasc_list[cum])
-        ascname <- names(cumasc)
-        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_prob_", ascname, ".asc"), overwrite = TRUE)
-        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_prob_", ascname, ".png")), overwrite = TRUE)
-        dev.off() # stops automatic saving of the plot to a png
-      }
-    }
-  }
-```
-
-```{r habitat classification on probability - high only}
-
-binprob <- function(x){
-  ifelse(x >= 0.8, 1,
-  ifelse(x < 0.8, 0, NA)) 
-}
-
-modelmonth <- cbind(paste0(mthna, unique(prback$year)))
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
-
-  for (modtime in 1:nrow(modelmonth)){
-    timeslice <- modelmonth[modtime]
-    yr <- str_sub(timeslice, -4)
-    for (cum in 1:length(clogasc_list)){
-      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
-      if(yr == cumascyr){
-        cumasc <- raster(clogasc_list[cum])
-        ascname <- names(cumasc)
-        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_highprob_", ascname, ".asc"), overwrite = TRUE)
-        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_highprob_", ascname, ".png")), overwrite = TRUE)
-        dev.off() # stops automatic saving of the plot to a png
-      }
-    }
-  }
-
-```
-
-```{r habitat classification on probability - med only}
-binprob <- function(x){
-  ifelse(x >= 0.7 & x <0.8, 1,
-  ifelse(x < 0.7, 0, NA)) 
-}
-
-modelmonth <- cbind(paste0(mthna, unique(prback$year)))
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
-
-  for (modtime in 1:nrow(modelmonth)){
-    timeslice <- modelmonth[modtime]
-    yr <- str_sub(timeslice, -4)
-    for (cum in 1:length(clogasc_list)){
-      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
-      if(yr == cumascyr){
-        cumasc <- raster(clogasc_list[cum])
-        ascname <- names(cumasc)
-        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_medprob_", ascname, ".asc"), overwrite = TRUE)
-        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_medprob_", ascname, ".png")), overwrite = TRUE)
-        dev.off() # stops automatic saving of the plot to a png
-      }
-    }
-  }
-```
-
-```{r habitat classification on probability - low only}
-binprob <- function(x){
-  ifelse(x >= 0.6 & x <0.7, 1,
-  ifelse(x < 0.6, 0, NA)) 
-}
-
-modelmonth <- cbind(paste0(mthna, unique(prback$year)))
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
-
-  for (modtime in 1:nrow(modelmonth)){
-    timeslice <- modelmonth[modtime]
-    yr <- str_sub(timeslice, -4)
-    for (cum in 1:length(clogasc_list)){
-      cumascyr <- str_sub(clogasc_list[cum], start = 44, end = 47) #selects the year from the file name
-      if(yr == cumascyr){
-        cumasc <- raster(clogasc_list[cum])
-        ascname <- names(cumasc)
-        probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-        writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_lowprob_", ascname, ".asc"), overwrite = TRUE)
-        plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_lowprob_", ascname, ".png")), overwrite = TRUE)
-        dev.off() # stops automatic saving of the plot to a png
-      }
-    }
-  }
-```
-
-now you need to do summary maps....
-
-```{r high probability map per year 2D representation}
-
-binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_highprob_cloglog.*.asc", full.names = TRUE)
-yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
-colnames(yrlayer) <- c("layer", "year")
-
-for (yrlyr in 1:length(binaryprob_list)){
-  yearoflyr <- str_sub(binaryprob_list[yrlyr], start = 60, end = 63) #selects the year from the file name
-  yrlayer$year[yrlyr] <- yearoflyr
-  yrlayer$layer[yrlyr] <- binaryprob_list[yrlyr]
-}
-
-y_list <- split(yrlayer, yrlayer$year)
-noyrs <- unique(yrlayer$year)
-
-for (binyr in (1:length(y_list))){
-  binstack <- stack(y_list[[binyr]]$layer)
-  year <- y_list[[binyr]]$year[1]
-  stacksum <- sum(binstack, na.rm = TRUE)
-  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/highprob_2d_", year, ".asc"), overwrite=TRUE)
-  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/highprob_2d_", year, ".png")))
-  dev.off()
-}
-
-```
-
-```{r med probability map per year 2D representation}
-binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_medprob_cloglog.*.asc", full.names = TRUE)
-yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
-colnames(yrlayer) <- c("layer", "year")
-
-for (yrlyr in 1:length(binaryprob_list)){
-  yearoflyr <- str_sub(binaryprob_list[yrlyr], start = 59, end = 62) #selects the year from the file name
-  yrlayer$year[yrlyr] <- yearoflyr
-  yrlayer$layer[yrlyr] <- binaryprob_list[yrlyr]
-}
-
-y_list <- split(yrlayer, yrlayer$year)
-noyrs <- unique(yrlayer$year)
-
-for (binyr in (1:length(y_list))){
-  binstack <- stack(y_list[[binyr]]$layer)
-  year <- y_list[[binyr]]$year[1]
-  stacksum <- sum(binstack, na.rm = TRUE)
-  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/medprob_2d_", year, ".asc"), overwrite=TRUE)
-  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/medprob_2d_", year, ".png")))
-  dev.off()
-}
-
-```
-
-```{r low probability map per year 2D representation}
-binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_lowprob_cloglog.*.asc", full.names = TRUE)
-yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
-colnames(yrlayer) <- c("layer", "year")
-
-for (yrlyr in 1:length(binaryprob_list)){
-  yearoflyr <- str_sub(binaryprob_list[yrlyr], start = 59, end = 62) #selects the year from the file name
-  yrlayer$year[yrlyr] <- yearoflyr
-  yrlayer$layer[yrlyr] <- binaryprob_list[yrlyr]
-}
-
-y_list <- split(yrlayer, yrlayer$year)
-noyrs <- unique(yrlayer$year)
-
-for (binyr in (1:length(y_list))){
-  binstack <- stack(y_list[[binyr]]$layer)
-  year <- y_list[[binyr]]$year[1]
-  stacksum <- sum(binstack, na.rm = TRUE)
-  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/lowprob_2d_", year, ".asc"), overwrite=TRUE)
-  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/lowprob_2d_", year, ".png")))
-  dev.off()
-}
-
-```
-
-now average probability split maps - again low, med, high and all
-
-```{r habitat classification on average probability - 3 classifications}
-
-binprob <- function(x){
-  ifelse(x >= 0.8, 3,
-  ifelse(x >= 0.7 & x <0.8, 2,
-  ifelse(x >= 0.6 & x <0.7, 1,
-  ifelse(x < 0.6, 0, NA)))) 
-}
-
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
-
-    for (cum in 1:length(clogasc_list)){
-      cumasc <- raster(clogasc_list[cum])
-      ascname <- names(cumasc)
-      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avgprob_", ascname, ".asc"), overwrite = TRUE)
-      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avgprob_", ascname, ".png")), overwrite = TRUE)
-      dev.off() # stops automatic saving of the plot to a png
-    }
-
-```
-
-```{r habitat classification on average probability - high}
-
-binprob <- function(x){
-  ifelse(x >= 0.8, 1,
-  ifelse(x < 0.8, 0, NA))  
-}
-
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
-
-    for (cum in 1:length(clogasc_list)){
-      cumasc <- raster(clogasc_list[cum])
-      ascname <- names(cumasc)
-      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avg_highprob_", ascname, ".asc"), overwrite = TRUE)
-      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_highprob_", ascname, ".png")), overwrite = TRUE)
-      dev.off() # stops automatic saving of the plot to a png
-    }
-
-```
-
-```{r habitat classification on average probability - med}
-
-binprob <- function(x){
-  ifelse(x >= 0.7 & x <0.8, 1,
-  ifelse(x < 0.7, 0, NA)) 
-}
-
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
-
-    for (cum in 1:length(clogasc_list)){
-      cumasc <- raster(clogasc_list[cum])
-      ascname <- names(cumasc)
-      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avg_medprob_", ascname, ".asc"), overwrite = TRUE)
-      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_medprob_", ascname, ".png")), overwrite = TRUE)
-      dev.off() # stops automatic saving of the plot to a png
-    }
-
-```
-
-```{r habitat classification on average probability - low}
-
-binprob <- function(x){
-  ifelse(x >= 0.6 & x <0.7, 1,
-  ifelse(x < 0.6, 0, NA)) 
-}
-
-clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*.asc", full.names = TRUE)
-
-    for (cum in 1:length(clogasc_list)){
-      cumasc <- raster(clogasc_list[cum])
-      ascname <- names(cumasc)
-      probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-      writeRaster(probbin, paste0("../output/maxent/predictions/", mthna,  "/binary_avg_lowprob_", ascname, ".asc"), overwrite = TRUE)
-      plot(probbin, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_lowprob_", ascname, ".png")), overwrite = TRUE)
-      dev.off() # stops automatic saving of the plot to a png
-    }
-
-```
-
-```{r high probability average map per year 2D representation}
-
-binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_highprob.*.asc", full.names = TRUE)
-
-binstack <- stack(binaryprob_list)
-stacksum <- sum(binstack, na.rm = TRUE)
-writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_highavg_2d.asc"), overwrite=TRUE)
-plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_highprob_avg_2d.png")))
-dev.off()
-
-```
-
-```{r med probability average map per year 2D representation}
-
-binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_medprob.*.asc", full.names = TRUE)
-
-binstack <- stack(binaryprob_list)
-stacksum <- sum(binstack, na.rm = TRUE)
-writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_medavg_2d.asc"), overwrite=TRUE)
-plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_medprob_avg_2d.png")))
-dev.off()
-
-```
-
-```{r low probability average map per year 2D representation}
-
-binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_lowprob.*.asc", full.names = TRUE)
-
-binstack <- stack(binaryprob_list)
-stacksum <- sum(binstack, na.rm = TRUE)
-writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_lowavg_2d.asc"), overwrite=TRUE)
-plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_lowprob_avg_2d.png")))
-dev.off()
-
-```
-
-first need some statistcs... for each layer how many cells have a value of 1...
-
-```{r calc no presence cells per layer (avg probability)} 
-ascdepth_list_high <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_highprob_.*asc", full.names = TRUE)
-ascdepth_list_med <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_medprob_.*asc", full.names = TRUE)
-ascdepth_list_low <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_lowprob_.*asc", full.names = TRUE)
-
-binaryavgstack <- stack(ascdepth_list_high)
-stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-stackstathigh[ ,1] <- c(1:46)
-colnames(stackstathigh) <- c("layer", "high_cell_sum")
-
-for (lyr in 1:nlayers(binaryavgstack)){
-  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
-  stackstathigh$high_cell_sum[lyr] <- lyrsum
-}
-
-binaryavgstack <- stack(ascdepth_list_med)
-stackstatmed <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-stackstatmed[ ,1] <- c(1:46)
-colnames(stackstatmed) <- c("layer", "med_cell_sum")
-
-for (lyr in 1:nlayers(binaryavgstack)){
-  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
-  stackstatmed$med_cell_sum[lyr] <- lyrsum
-}
-
-binaryavgstack <- stack(ascdepth_list_low)
-stackstatlow <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-stackstatlow[ ,1] <- c(1:46)
-colnames(stackstatlow) <- c("layer", "low_cell_sum")
-
-for (lyr in 1:nlayers(binaryavgstack)){
-  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
-  stackstatlow$low_cell_sum[lyr] <- lyrsum
-}
-
-stackstat <- cbind(stackstathigh, stackstatmed[ , 2], stackstatlow[ ,2])
-colnames(stackstat) <- c("layer", "high", "med", "low")
-stackstat <- melt(stackstat, id = "layer")
-
-write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/probability_cellcount_bydepth_avg.csv"), row.names = FALSE)
-
-```
-
-a crude plot but ok to show slr
-
-```{r graph by depth - prob plot (avg)}
-bydepth <- ggplot(data = stackstat, aes(x = layer, y = value, fill = variable)) + geom_bar(stat = "identity") +   scale_x_reverse() +
-coord_flip()
-png(paste0("../output/maxent/predictions/", mthna, "/depthplot_probavg.png"))
-print(bydepth)
-dev.off()
-
-```
-
-#which cells are used in all months?
-
-start with mpa binary map first - based on averages
-
-```{r get all csvs in one database}
-mthlst <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-stdep <- read.csv("../output/maxent/models/dec1998/stack_depth.csv", header = TRUE)
-depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
-names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
-depthlayers$depth <- as.character(depthlayers$depth)
-
-for(m in 1:length(mthlst)){
-  mthna <- mthlst[m]
-#mthna <- "jan"
-  mpabin_lst <- list.files(paste0("../output/maxent/predictions/", mthna), pattern =  "^binary_avg_cloglog.*.asc", full.names = TRUE)
-
-  for (d in 1:nrow(depthlayers)){
-    dep <- depthlayers$depth[d]
-    for (a in 1:length(mpabin_lst)){
-      asc <- mpabin_lst[a]
-      if (grepl(dep, asc)){
-        depthlayers[d , paste0(mthna)] <- asc
-
-      }
-    }
-  }
-  write.csv(depthlayers, ("../output/maxent/predictions/all_cells/binary_avg_asclayerbydepth.csv"), row.names = FALSE)
-}
-
-```
-
-```{r create stacks by depth}
-depthlayers <- read.csv("../output/maxent/predictions/all_cells/binary_avg_asclayerbydepth.csv", header = TRUE)
-
-for (x in 1:nrow(depthlayers)){
-  asclay <- depthlayers[x, ]
-  asclay <- asclay[!is.na(asclay)]
-  asclay <- as.data.frame(rbind(asclay))
-  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
-  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to month
-  asclay$V3 <- as.character(asclay$V3)
-  asclay$V4 <- as.character(asclay$V4)
-  asclay$V5 <- as.character(asclay$V5)
-  asclay$V6 <- as.character(asclay$V6)
-  asclay$V7 <- as.character(asclay$V7)
-  asclay$V8 <- as.character(asclay$V8)
-  asclay$V9 <- as.character(asclay$V9)
-  asclay$V10 <- as.character(asclay$V10)
-  asclay$V11 <- as.character(asclay$V11)
-  asclay$V12 <- as.character(asclay$V12)
-
-  asclaydep <- asclay$V1
-
-  ascjan <- raster(asclay$V2)
-  ascmar <- raster(asclay$V3)
-  ascapr <- raster(asclay$V4)
-  ascmay <- raster(asclay$V5)
-  ascjun <- raster(asclay$V6)
-  ascjul <- raster(asclay$V7)
-  ascaug <- raster(asclay$V8)
-  ascsep <- raster(asclay$V9)
-  ascoct <- raster(asclay$V10)
-  ascnov <- raster(asclay$V11)
-  ascdec <- raster(asclay$V12)
-
-  stkasc <- stack(ascjan, ascmar, ascapr, ascmay, ascjun, ascjul, ascaug, ascsep, ascoct, ascnov, ascdec)
-
-  avg <- sum(stkasc, na.rm = TRUE)
-
-  writeRaster(avg, paste0("../output/maxent/predictions/all_cells/count_cells_used_all_months_mpa", asclaydep, ".asc"), overwrite=TRUE)
-  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/all_cells/count_cells_used_all_months_mpa", asclaydep, ".png")), overwrite = TRUE) 
-  dev.off()
-}
-
-```
-
-```{r create rasters showing cells that are suitable in all 11 months}
-cellslst <- list.files("../output/maxent/predictions/all_cells", pattern =  "^count_cells_used_all_months_mpa.*.asc", full.names = TRUE)
-
-allcell <- function(x){
-  ifelse(x == 12, 1,
-  ifelse(x < 12, 0, NA)) 
-}
-
-for (r in 1:length(cellslst)){
-  ras <- raster(cellslst[r])
-  ascname <- names(ras)
-  allc <- calc(ras, fun = allcell)
-  writeRaster(allc, paste0("../output/maxent/predictions/all_cells/total_cells_used_all_months_mpa", ascname, ".asc"), overwrite = TRUE)
-  plot(allc, png(paste0("../output/maxent/predictions/all_cells/total_cells_used_all_months_mpa", ascname, ".png")), overwrite = TRUE)
-  dev.off() # stops automatic saving of the plot to a png
-
-}
-```
-
-there are no cells that are used in every month....
-
-# quickly adding some black to some plots for mj slr
-```{r}
-mthlst <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-
-for(m in 1:length(mthlst)){
-  mthna <- mthlst[m]
-  replst11 <- list.files(paste0("../output/maxent/predictions/", mthna), pattern =  ".*11.7737.*.asc", full.names = TRUE)
-  replst221 <- list.files(paste0("../output/maxent/predictions/", mthna), pattern =  ".*221.141.*.asc", full.names = TRUE)
-  replst508 <- list.files(paste0("../output/maxent/predictions/", mthna), pattern =  ".*508.64.*.asc", full.names = TRUE)
-  replst <- c(replst11, replst221, replst508)
-  for (r in 1:length(replst)){
-    ras <- raster(replst[r])
-    ascname <- names(ras)
-    plot(ras, colNA = 'black', png(paste0("../output/maxent/predictions/", mthna, "/", ascname, ".png")), overwrite = TRUE)
-    dev.off()
-  }
-}
-
-replst163 <- list.files("../output/maxent/predictions/jul/", pattern =  ".*163..*.asc", full.names = TRUE)
-for (r in 1:length(replst163)){
-    ras <- raster(replst163[r])
-    ascname <- names(ras)
-    plot(ras, colNA = 'black', png(paste0("../output/maxent/predictions/jul/", ascname, ".png")), overwrite = TRUE)
-    dev.off()
-}
-
-back <- raster("../output/maxent/predictions/mar/binary_avg_medprob_avg_cloglog_1.55586.asc")
-replst163 <- list.files("../output/maxent/predictions/jan/", pattern =  ".*avg_2d.*.asc", full.names = TRUE)
-for (r in 1:length(replst163)){
-    ras <- raster(replst163[r])
-    ascname <- names(ras)
-    bkras <- overlay(back, ras, fun=function(x,y){return(x+y)})
-    plot(bkras, colNA = 'black', png(paste0("../output/maxent/predictions/jan/", ascname, ".png")), overwrite = TRUE)
-    dev.off()
-}
-
-```

### 2020-01-16 src/testing_file.Rmd
update to combined maxent

@@
 ```
+
+#editing the prediction layers for combined
+
+need to check for year and month
+
+test with march
+  if (yr == "2003"){
+    betaval <- 1
+  } else if (yr == "1999" | yr == "2002" | yr == "2008" | yr == "2009") {
+    betaval <- 1.25
+  } else if (yr == "2005" | yr == "2006"){
+    betaval <- 1.5
+  } else if (yr == "2000"){
+    betaval <- 1.75
+  } else if (yr == "2001" | yr == "2007"){
+    betaval <- 2.25
+  } else if (yr == "2010"){
+    betaval <- 3.5
+  }
+
+```{r pre-install & load packages, include=FALSE}
+options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2")
+source("../src/ipak.R")
+ipak(packages)
+```
+
+load the presence-background dataset 
+("../output/bio/presab_nafo_yr.csv")
+
+```{r load pres-background dataset}
+presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
+colnames(presback)
+```
+
+this dataset contains a lot of columns you don't need for the maxent models. Remove them here
+
+```{r remove unnecessary columns}
+presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt))
+colnames(presback)
+```
+
+monthly model data selection and prep
+For each monthly model, subset the data to the month you want, and the variables you want (Based on spearmans and VIF)
+
+```{r}
+marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+
+mthnavar_lst <- list(mar = marvar, apr = aprvar, may = mayvar, jun = junvar, jul = julvar, aug = augvar, sep = sepvar, oct = octvar, nov = novvar, dec = decvar)
+```
+
+select month
+```{r month of interest -change}
+mthna <- "mar"
+mthno <- "3"
+```
+
+```{r subsest data by month}
+prback <- subset(presback, month == mthno)
+prback <- subset(prback, select = mthnavar_lst[[mthna]])
+```
+
+now each month contains NAs - lets remove these (maxent principle doesnt allow for missing data - http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.15.5607&rep=rep1&type=pdf)
+
+note will need to reajust background points too as some years may no longer have presence data
+
+```{r remove na vals}
+prback <- na.omit(prback)
+presyrs <-  subset(prback, occurrence == 1) #this is to get years with complete presence data 
+presyrs <- unique(presyrs$year) #make a vector of unique years
+prback <- prback[prback$year %in% presyrs, ] #keep only the data for which there is precence years
+
+```
+
+```{r final models plus predict}
+modelmonth <- cbind(paste0(mthna, unique(prback$year)))
+aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
+unlayer <- read.csv(paste0("../data/env/test-comb/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
+
+for (year in 1:nrow(modelmonth)){
+  timeslice <- modelmonth[year] #testing timeslice
+  yr <- str_sub(timeslice, -4) #the year being used for testing
+  train <- subset(prback, year != yr) #training data is all years but testing
+  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
+  test <-  subset(prback, year == yr) #creates the testing dataset
+  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
+  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
+  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  #mess_pr <- subset(train, occurrence == 1)
+  #mess_pr  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+
+  if (yr == "2003"){
+    betaval <- 1
+  } else if (yr == "1999" | yr == "2002" | yr == "2008" | yr == "2009") {
+    betaval <- 1.25
+  } else if (yr == "2005" | yr == "2006"){
+    betaval <- 1.5
+  } else if (yr == "2000"){
+    betaval <- 1.75
+  } else if (yr == "2001" | yr == "2007"){
+    betaval <- 2.25
+  } else if (yr == "2010"){
+    betaval <- 3.5
+  }
+
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/test-comb/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
+
+    saveRDS(mod, file= paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))
+
+    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
+
+    #count number of non-zero coefficients, excluding the final four rows
+    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    coefile <- head(coefile, -4)
+    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+    nocoeff <- nrow(coefile2)
+    bick <- sum(as.numeric(coefile2[, 2]))
+    bic <- (bick*log(sum(trainpa == 1)) - 2*11)
+
+    #no presences used in model (less those used for testing)
+    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    notrainingpoints <- maxresults$X.Training.samples
+    testauc <- maxresults$Test.AUC
+    trainauc <- maxresults$Training.AUC
+
+    #natural log
+    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    samplepred$nat_log <- log(samplepred$Raw.prediction)
+    sumsamplepred <- sum(samplepred$nat_log)
+
+    #test auc scores and other lovely things you might want to include
+    ev <- evaluate(p = testp, a= testb, model = mod)
+    evauc <- ev@auc
+    evcor <- ev@cor
+    evpcor <- ev@pcor
+    evnp <- ev@np
+    evna <- ev@na
+    evmaxtprtnr <- threshold(ev, 'spec_sens')
+    evmaxkap <- threshold(ev, 'kappa')
+    evno_omission <- threshold(ev, 'no_omission')
+    evprevalence <- threshold(ev, 'prevalence')
+    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+    evsensitivity <- threshold(ev, 'sensitivity')
+
+    #continuous boyce index
+    predPres <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    predBg <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
+    predPres <- predPres$Cloglog.prediction
+    predBg <- predBg$Cloglog
+    source("../src/contBoyce.R")
+    cbi <- contBoyce(pres = predPres, bg = predBg)
+
+      aicetc$train_AUC <- trainauc #training only AUC - not tested
+      aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
+      aicetc$coefficients <- nocoeff #this is to calculate the AIC
+      aicetc$sum.of.logs <- sumsamplepred #this is to calculate the AIC
+      aicetc$test_AUC <- evauc #testing AUC
+      aicetc$test_evcor <- evcor #testing correlation coefficient
+      aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
+      aicetc$test_nopresence <- evnp 
+      aicetc$test_nobackground <- evna
+      aicetc$test_maxsumsenspec <- evmaxtprtnr
+      aicetc$test_maxkap <- evmaxkap
+      aicetc$test_no_omission <- evno_omission
+      aicetc$test_prevalence <- evprevalence
+      aicetc$test_equal_sens_spec <- evequal_sens_spec
+      aicetc$test_sensitivity <- evsensitivity
+      aicetc$tss <- max(ev@TPR + ev@TNR) - 1
+      aicetc$test_bic <- bic #not sure this is correct...
+      aicetc$beta <- betaval
+      aicetc$cbi <- cbi
+
+    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
+    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$no_trainng_points/(aicetc$no_trainng_points - aicetc$coefficients-1))))
+
+    #calculate minimal predicted area using ecospat
+    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
+    aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
+    aicetc$year <- yr
+
+    write.csv(aicetc, paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+
+    mod <- readRDS(paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))
+
+    yr <- as.numeric(yr)
+ unlayer_yr <- subset(unlayer, ascyears == yr)
+    #unlayer_yr <- unlayer_yr[-c(1), ]
+    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+    stackdeplst <- list()
+    stackdep <- subset(unlayer_yr, select = c(stackname))
+    stackdeplst[[year]] <- stackdep
+    write.csv(stackdep, paste0("../output/maxent/models/test-comb/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)
+    aea <- raster("../output/env/aea.tif")
+
+  for (td in 1:nrow(unlayer_yr)){
+      temp_depth <- raster(unlayer_yr$temp_depth[td])
+      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(temp_depth) <- "temp_depth"
+      #temp_depth <- projectRaster(temp_depth, aea)
+      chl_surface <- raster(unlayer_yr$chl_depth[td])
+      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(chl_surface) <- "chl_surface"
+      #chl_depth <- projectRaster(chl_depth, aea)
+      o2_depth <- raster(unlayer_yr$o2_depth[td])
+      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(o2_depth) <- "o2_depth"
+      #o2_depth <- projectRaster(o2_depth, aea)
+      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
+      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(salinity_depth) <- "salinity_depth"
+      #salinity_depth <- projectRaster(salinity_depth, aea)
+      # nao_sample <- raster(unlayer_yr$nao_sample[td])
+      # crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      # names(nao_sample) <- "nao_sample"
+      # nao_sample <- projectRaster(nao_sample, aea)
+      # nao_prev <- raster(unlayer_yr$nao_prev[td])
+      # crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      # names(nao_prev) <- "nao_prev"
+      # nao_prev <- projectRaster(nao_prev, aea)
+      # nao_winter <- raster(unlayer_yr$nao_winter[td])
+      # crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      # names(nao_winter) <- "nao_winter"
+      # nao_winter <- projectRaster(nao_winter, aea)
+      # amo_sample <- raster(unlayer_yr$amo_sample[td])
+      # crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      # names(amo_sample) <- "amo_sample"
+      # amo_sample <- projectRaster(amo_sample, aea)
+      # amo_winter <- raster(unlayer_yr$amo_winter[td])
+      # crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      # names(amo_winter) <- "amo_winter"
+      # amo_winter <- projectRaster(amo_winter, aea)
+      stackn <- unlayer_yr$stackname[td]
+      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface)
+
+      #stackSave(test,unlayer_yr$stackname[td])
+      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=cloglog')) 
+      # predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
+      prplot <- plot(pred, png(paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".png")))
+      dev.off() # stops automatic saving of the plot to a png
+}
+```
+
+need to convert ascii wgs to aea (no maybe not..)
+```{r}
+testr <- raster("../data/env/ascii-original/ascii2000/2000_01_chl.nc_3.85628.asc")
+crs(testr) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+names(testr) <- "testr"
+testr <- projectRaster(testr, aea)
+writeRaster(testr, "../data/env/test-comb/test.asc", format = 'ascii')
+
+test2 <- raster("../data/env/test-comb/test.asc")
+plot(test2)
+test2
+```
+
+set up a loop to go through all directories, take files in each directory, project, then put in a new folder
+
+```{r}
+#list directories with the ascii files that need converting in
+asciidirs <- list.dirs("../data/env/ascii-original/", full.names = TRUE, recursive = TRUE)
+asciidirs <- asciidirs[-1] #to remove the top directory
+aea <- raster("../output/env/aea.tif")
+
+for (dir in 1:length(asciidirs)){ #go through each directory
+  ascilst <- list.files(asciidirs[dir], full.names = TRUE, pattern = "\\.asc$") #make a list of files in the directory
+  dirn <- asciidirs[dir]
+  dirn <- substring(dirn, 29)
+  for (as in 1:length(ascilst)){ #for each file in the directory
+    ras <- raster(ascilst[as])
+    crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+    rasp <- projectRaster(ras, aea)
+    rnm <- names(ras)
+    rnm <- substring(rnm, 2)
+    writeRaster(rasp, paste0("../data/env/ascii-aea/", dirn, "/", rnm, ".asc"), format = 'ascii')
+    }
+}
+
+```

### 2020-01-16 src/contBoyce.r

@@ -1,0 +1,202 @@
+#' Continuous Boyce Index (CBI) with weighting
+#'
+#' This function calculates the Continuous Boyce Index (CBI), a measure of model accuracy for presence-only test data.
+#' @param pres Numeric vector. Predicted values at presence sites.
+#' @param contrast Numeric vector. Predicted values at absence/background sites.
+#' @param numBins Positive integer. Number of (overlapping) bins into which to divide predictions.
+#' @param binWidth Positive numeric. Size of a bin. Each bin will be \code{binWidth * (max - min)} where \code{max} and \code{min}. If \code{autoWindow} is \code{FALSE} (the default) then \code{min} is 0 and \code{max} is 1. If \code{autoWindow} is \code{TRUE} then \code{min} and \code{max} are the maximum and minimum value of all predictions in the background and presence sets (i.e., not necessarily 0 and 1).
+#' @param presWeight Numeric vector same length as \code{pres}. Relative weights of presence sites. The default is to assign each presence a weight of 1.
+#' @param contrastWeight Numeric vector same length as \code{contrast}. Relative weights of background sites. The default is to assign each presence a weight of 1.
+#' @param autoWindow Logical. If \code{FALSE} calculate bin boundaries starting at 0 and ending at 1 + epsilon (where epsilon is a very small number to assure inclusion of cases that equal 1 exactly). If \code{TRUE} (default) then calculate bin boundaries starting at minimum predicted value and ending at maximum predicted value.
+#' @param method Character. Type of correlation to calculate. The default is \code{'spearman'}, the Spearman rank correlation coefficient used by Boyce et al. (2002) and Hirzel et al. (2006), which is the traditional CBI. In contrast, \code{'pearson'} or \code{'kendall'} can be used instead.  See [stats::cor()] for more details.
+#' @param dropZeros Logical. If TRUE then drop all bins in which the frequency of presences is 0.
+#' @param graph Logical. If TRUE then plot P vs E and P/E versus bin.
[…]

### 2020-01-17 src/asc_file_database_combined_files.Rmd
amo and nao added, model completed

@@
 ```{r base list or ascii layers}
-ascyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
+ascyearslist <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014")) 
 #unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
@@
 asclayerall$temp_depth <- NA
-asclayerall$chl_depth <- NA
+asclayerall$chl_surface <- NA
 asclayerall$o2_depth <- NA
 asclayerall$salinity_depth <- NA
-# asclayerall$nao_sample <- NA
-# asclayerall$nao_prev <- NA
-# asclayerall$nao_winter <- NA
-# asclayerall$amo_sample <- NA
-# asclayerall$amo_prev <- NA
-# asclayerall$amo_winter <- NA
+asclayerall$nao_sample <- NA
+asclayerall$nao_prev <- NA
+asclayerall$nao_winter <- NA
+asclayerall$amo_sample <- NA
+asclayerall$amo_winter <- NA
 
@@
 
-myr <- 2014 #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
+for (ny in 1:length(ascyearslist)){
+  myr <- ascyearslist[ny] #!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME
 
@@
 salvar <- "salinity.asc"
-# naosamplevar <- "naosample"
-# amosamplevar <- "amosample"
-# naoprevvar <- "naoprev"
-# amoprevvar <- "amoprev"
-# naowintervar <- "naowinter"
-# amowintervar <- "amowinter"
-# surfacevar <- "0.50576"
+naosamplevar <- "naosample.asc"
+amosamplevar <- "amosample.asc"
+naoprevvar <- "naoprev.asc"
+naowintervar <- "naowinter.asc"
+amowintervar <- "amowinter.asc"
+surfacevar <- "0.50576"
 
@@
         } else if(ascyear == year & ascmonth == month & ascvar == chlvar){
-          asclayer$chl_depth[i] <- ascfile
+          asclayer$chl_surface[i] <- ascfile
         } else if(ascyear == year & ascmonth == month & ascvar == o2var){
           asclayer$o2_depth[i] <- ascfile
-        # } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
-        #   asclayer$nao_sample[i] <- ascfile
-        # } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
-        #   asclayer$amo_sample[i] <- ascfile
-        # } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
-        #   asclayer$nao_prev[i] <- ascfile
-        # } else if(ascyear == year & ascmonth == month & ascvar == amoprevvar){
-        #   asclayer$amo_prev[i] <- ascfile
-        # } else if(ascyear == year & ascvar == naowintervar){
-        #   asclayer$nao_winter[i] <- ascfile
-        # } else if(ascyear == year & ascvar == amowintervar){
-        #   asclayer$amo_winter[i] <- ascfile
+         } else if(ascyear == year & ascmonth == month & ascvar == naosamplevar){
+           asclayer$nao_sample[i] <- ascfile
+         } else if(ascyear == year & ascmonth == month & ascvar == amosamplevar){
+           asclayer$amo_sample[i] <- ascfile
+         } else if(ascyear == year & ascmonth == month & ascvar == naoprevvar){
+           asclayer$nao_prev[i] <- ascfile
+         } else if(ascyear == year & ascvar == naowintervar){
+           asclayer$nao_winter[i] <- ascfile
+         } else if(ascyear == year & ascvar == amowintervar){
+           asclayer$amo_winter[i] <- ascfile
         }
@@
   write.csv(asclayer, paste0("../data/env/test-comb/asc_layerlist", myr, ".csv"), row.names = TRUE)
+}
 }

### 2020-01-17 src/comb-maxent.Rmd
amo and nao added, model completed

@@
 colnames(presback)
+```
+
+remove all rows with the year 2015 (due to low number of samples)
+```{r}
+presback <- subset(presback, year!="2015")
 ```
@@
 ```{r}
-marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
-aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
-mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
-junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
-julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
-augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
-sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
-octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
-novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
-decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
+marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
+decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
 
@@
       #temp_depth <- projectRaster(temp_depth, aea)
-      chl_surface <- raster(unlayer_yr$chl_depth[td])
+      chl_surface <- raster(unlayer_yr$chl_surface[td])
       crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
@@
       #salinity_depth <- projectRaster(salinity_depth, aea)
-      # nao_sample <- raster(unlayer_yr$nao_sample[td])
-      # crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
-      # names(nao_sample) <- "nao_sample"
-      # nao_sample <- projectRaster(nao_sample, aea)
-      # nao_prev <- raster(unlayer_yr$nao_prev[td])
-      # crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
-      # names(nao_prev) <- "nao_prev"
-      # nao_prev <- projectRaster(nao_prev, aea)
-      # nao_winter <- raster(unlayer_yr$nao_winter[td])
-      # crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
-      # names(nao_winter) <- "nao_winter"
-      # nao_winter <- projectRaster(nao_winter, aea)
-      # amo_sample <- raster(unlayer_yr$amo_sample[td])
-      # crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
-      # names(amo_sample) <- "amo_sample"
-      # amo_sample <- projectRaster(amo_sample, aea)
-      # amo_winter <- raster(unlayer_yr$amo_winter[td])
-      # crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
-      # names(amo_winter) <- "amo_winter"
-      # amo_winter <- projectRaster(amo_winter, aea)
+      nao_sample <- raster(unlayer_yr$nao_sample[td])
+      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(nao_sample) <- "nao_sample"
+      #nao_sample <- projectRaster(nao_sample, aea)
+      nao_prev <- raster(unlayer_yr$nao_prev[td])
+      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(nao_prev) <- "nao_prev"
+      #nao_prev <- projectRaster(nao_prev, aea)
+      nao_winter <- raster(unlayer_yr$nao_winter[td])
+      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      names(nao_winter) <- "nao_winter"
+      #nao_winter <- projectRaster(nao_winter, aea)
+      amo_sample <- raster(unlayer_yr$amo_sample[td])
+      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+      names(amo_sample) <- "amo_sample"
+      #amo_sample <- projectRaster(amo_sample, aea)
+      amo_winter <- raster(unlayer_yr$amo_winter[td])
+      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      names(amo_winter) <- "amo_winter"
+      #amo_winter <- projectRaster(amo_winter, aea)
       stackn <- unlayer_yr$stackname[td]
-      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface)
+      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
 

### 2020-01-20 src/asc_file_database_combined_files.Rmd
splitting asc files - minor edits

@@
 
-asclayerall <- crossing(ascyears, ascmonths)
-asclayerall$stackname <- paste0("st", asclayerall$ascyears, asclayerall$ascmonths) #this creates the name of the stack that will be used
+asclayerall <- crossing(ascyearslist, ascmonths)
+asclayerall$stackname <- paste0("st", asclayerall$ascyearslist, asclayerall$ascmonths) #this creates the name of the stack that will be used
 asclayerall$temp_depth <- NA
@@
 
-asclayer <- subset(asclayerall, ascyears == myr)
+asclayer <- subset(asclayerall, ascyearslist == myr)
 
@@
     print(asclayer$stackname[i])
-    year <- asclayer$ascyears[i]
+    year <- asclayer$ascyearslist[i]
     month <- asclayer$ascmonths[i]

### 2020-01-20 src/comb-maxent.Rmd
splitting asc files - minor edits

@@
 ```{r month of interest -change}
-mthna <- "apr"
-mthno <- "4"
+mthna <- "dec"
+mthno <- "12"
 ```
@@
 
-  if (yr == "2001" | yr == "2006" | yr == "2007"){
+  if (yr == "1998"){
+    betaval <- 0.25
+  } else if (yr == "2001"){
+    betaval <- 0.75
+  } else if (yr == "2002" | yr == "2003"| yr == "2004"| yr == "2006"| yr == "2011"){
+    betaval <- 1
+  } else if (yr == "2005"){
+    betaval <- 1.25
+  } else if (yr == "1999" | yr == "2000"| yr == "2009"){
+    betaval <- 1.5
+  } else if (yr == "2007"){
     betaval <- 1.75
-  } else if (yr == "2000"| yr == "2002" | yr == "2003" | yr == "2004"){
+  } else if (yr == "2010"){
     betaval <- 2
-  } else if (yr == "1999" | yr == "2005" | yr == "2008" | yr == "2009"){
+  } else if (yr == "2008"){
     betaval <- 2.25
-  } else if (yr == "2010" | yr == "2011"){
-    betaval <- 2.5
-  } else if (yr == "1998"){
-    betaval <- 2.75
   }
@@
     yr <- as.numeric(yr)
- unlayer_yr <- subset(unlayer, ascyears == yr)
+ unlayer_yr <- subset(unlayer, ascyearslist == yr)
     #unlayer_yr <- unlayer_yr[-c(1), ]
@@
 ```
+
+all layers need to be spit in bash

### 2020-01-23 src/comb-maxent.Rmd
avg ascii layer complete. mcp reprojected

@@
     #count number of non-zero coefficients, excluding the final four rows
-    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    coefile <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
     coefile <- head(coefile, -4)
@@
     #no presences used in model (less those used for testing)
-    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    maxresults <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
     notrainingpoints <- maxresults$X.Training.samples
@@
     #natural log
-    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    samplepred <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
     samplepred$nat_log <- log(samplepred$Raw.prediction)
@@
     #continuous boyce index
-    predPres <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
-    predBg <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
+    predPres <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    predBg <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
     predPres <- predPres$Cloglog.prediction
@@
     write.csv(stackdep, paste0("../output/maxent/models/test-comb/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)
-    aea <- raster("../output/env/aea.tif")
 
@@
 
-all layers need to be spit in bash
+all layers need to be spit in bash + poweshell
+- pre-split.ps1 
+- separate.sh
+- aftersplit.ps1
+
+# average the maxent results
+
+```{r avg maxentReg}
+monthlist <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
+
+for (mth in 1:length(monthlist)){
+    mthna <- monthlist[mth]
+    modlist <- list.dirs("../output/maxent/models/test-comb", full.names = FALSE, recursive = FALSE) 
+    modlist <- modlist[ grepl(mthna, modlist) ]
+    aiclst <- list()
+    maxrlst <- list()
+    for (i in 1:length(modlist)){
+      timeslice <- modlist[i]
+      yr <- str_sub(timeslice, -4) #the year 
+      if(yr == 1998){
+        aic1998 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic1998
+        maxr1998 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr1998
+        } else if (yr == 1999){
+        aic1999 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic1999
+        maxr1999 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr1999
+        } else if (yr == 2000){
+        aic2000 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2000
+        maxr2000 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2000
+        } else if (yr == 2001){
+        aic2001 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2001
+        maxr2001 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2001
+        } else if (yr == 2002){
+        aic2002 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2002
+        maxr2002 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2002
+        } else if (yr == 2003){
+        aic2003 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2003
+        maxr2003 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2003
+        } else if (yr == 2004){
+        aic2004 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2004
+        maxr2004 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2004
+        } else if (yr == 2005){
+        aic2005 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2005
+        maxr2005 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2005
+        } else if (yr == 2006){
+        aic2006 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2006
+        maxr2006 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2006
+        } else if (yr == 2007){
+        aic2007 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2007
+        maxr2007 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2007
+        } else if (yr == 2008){
+        aic2008 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2008
+        maxr2008 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2008
+        } else if (yr == 2009){
+        aic2009 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2009
+        maxr2009 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2009
+        } else if (yr == 2010){
+        aic2010 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2010
+        maxr2010 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2010
+        } else if (yr == 2011){
+        aic2011 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2011
+        maxr2011 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2011
+        } else if (yr == 2012){
+        aic2012 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2012
+        maxr2012 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2012
+        } else if (yr == 2013){
+        aic2013 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2013
+        maxr2013 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2013
+        } else if (yr == 2014){
+        aic2014 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
+        aiclst[[i]] <- aic2014
+        maxr2014 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+        maxrlst[[i]] <- maxr2014
+        }
+    }
+  aic <- do.call(rbind, aiclst)
+  maxr <- do.call(rbind, maxrlst)
+
+  aicm <- colMeans(aic)
+  aicm$mod <- "avg"
+  avgAIC <- rbind(aic, aicm)
+
+  maxr$Species <- 1
+  maxrm <- colMeans(maxr)
+  maxrm$Species <- "avg"
+  maxrm <- rbind(maxr, maxrm)
+
+  write.csv(avgAIC, paste0("../output/maxent/models/test-comb/", mthna, "avgAIC.csv"), row.names = FALSE)
+  write.csv(maxrm, paste0("../output/maxent/models/test-comb/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
+}
+```
+
+summary table of average maxent results for each month
+```{r average AIC etc for each month in one table}
+aic_lst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
+aicsum <- data.frame(matrix(ncol = 11, nrow = length(aic_lst)))
+colnames(aicsum) <- c("Month", "BetaVal", "AICc", "Test AUC", "Training AUC", "Correlation", "Correlation P-Val", "TSS", "CBI", "No Omissions", "Prevalence")
+
+for (a in 1:length(aic_lst)){
+  aicr <- read.csv(aic_lst[a])
+  mth <- str_sub(aic_lst[a], start = 35, end = 37) #selects the month from the filename
+  aicsum$Month[a] <- mth
+  aicres <- tail(aicr, 1)
+  aicsum$BetaVal[a] <- aicres$beta
+  aicsum$AICc[a] <- aicres$AICc
+  aicsum$`Test AUC`[a] <- aicres$test_AUC
+  aicsum$`Training AUC`[a] <- aicres$train_AUC
+  aicsum$Correlation[a] <- aicres$test_evcor
+  aicsum$`Correlation P-Val` <- aicres$test_evpcor
+  aicsum$TSS[a] <- aicres$tss
+  aicsum$CBI[a] <- aicres$cbi
+  aicsum$`No Omissions`[a] <- aicres$test_no_omission
+  aicsum$Prevalence[a] <- aicres$test_prevalence
+}
+
+mthord <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
+aicsum <- aicsum[match(mthord, aicsum$Month), ] #reorder based on mthord
+write.csv(aicsum, "../output/maxent/models/test-comb/aicsummary.csv", row.names = FALSE)
+
+```
+
+# average the ascii layers 
+need a montly layer for each month-depth
+
+first need to rebuild a list of stackdepths as this does not work in the code above (due to change in ascii formatting - code above is old)
+
+```{r}
+#need an list of the month years
+p <- list.dirs("../output/maxent/models/test-comb", recursive = FALSE, full.names = FALSE)
+yrmthunique <- as.data.frame(p, stringsAsFactors=FALSE)
+
+for(time in 1:nrow(yrmthunique)){
+  slice <- yrmthunique$p[time]
+  mthna <- str_sub(slice, end = 3)
+  yr <-  str_sub(slice, start = 4)
+  mthyrclog <- list.files(paste0("../output/maxent/predictions/com-split/", mthna), pattern = yr)
+  stackd <- data.frame(mthyrclog)
+  for (st in 1:nrow(stackd)){
+    stackd$ascdepths[st] <- str_sub(stackd$mthyrclog[st], 17, -5)
+  }
+  names(stackd)[names(stackd) == "mthyrclog"] <- "stackname"
+  write.csv(stackd, paste0("../output/maxent/models/test-comb/", mthna, yr, "/stackdepthlist_", yr, ".csv"))
+}
+```
+
+```{r create list of layers}
+p <- list.dirs("../output/maxent/models/test-comb", recursive = FALSE, full.names = FALSE)
+year <- "2006"
+
+monthlist <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
+
+for (m in 1:length(monthlist)){
+  mthna <- monthlist[m]
+  timeslice <- paste(mthna, year, sep = '')
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
+  stlst <- list.files(paste0("../output/maxent/models/test-comb/", timeslice), pattern = "stackdepth.*.csv", full.names = TRUE)
+  stdep <- read.csv(stlst[1])
+  depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
+  names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
+  depthlayers$depth <- as.character(depthlayers$depth)
+
+  colno <- 1
+
+  for (d in 1:nrow(depthlayers)){
+    dep <- depthlayers$depth[d]
+    for (a in 1:length(clogasc_list)){
+      asc <- clogasc_list[a]
+      ascd <- str_sub(asc, 60, -5)
+      if (grepl(dep, ascd)){
+        depthlayers[d , paste0("a", colno)] <- asc
+         colno <- colno+1
+      }
+    }
+  }
+  write.csv(depthlayers, (paste0("../output/maxent/predictions/com-split/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)
+}
+```
+
+```{r avg asc layers cloglog}
+mthna <- "aug" #change me
+depthlayers <- read.csv(paste0("../output/maxent/predictions/com-split/", mthna, "/cloglog_asclayerbydepth.csv"), header = TRUE)
+p <- list.dirs("../output/maxent/models/test-comb", recursive = FALSE, full.names = FALSE)
+modlist <- p[ grepl(mthna, p) ]
+print(length(modlist)) #need this to modify code below
+print(modlist) #need this to modify code below
+
+for (x in 1:nrow(depthlayers)){
+  asclay <- depthlayers[x, ]
+  asclay <- asclay[!is.na(asclay)]
+  asclay <- as.data.frame(rbind(asclay))
+  asclay$V1 <- as.character(asclay$V1) #v1 just a number
+  asclay$V2 <- as.character(asclay$V2) #v2 corresponds to depth
+  asclay$V3 <- as.character(asclay$V3) #yr 1
+  asclay$V4 <- as.character(asclay$V4) #yr 2
+  asclay$V5 <- as.character(asclay$V5) #yr 3
+  asclay$V6 <- as.character(asclay$V6) #yr 4
+  asclay$V7 <- as.character(asclay$V7)
+  asclay$V8 <- as.character(asclay$V8) #yr 6
+  asclay$V9 <- as.character(asclay$V9)
+  asclay$V10 <- as.character(asclay$V10) #yr 8
+  asclay$V11 <- as.character(asclay$V11)
+  asclay$V12 <- as.character(asclay$V12) #yr 10
+  # asclay$V13 <- as.character(asclay$V13)
+  # asclay$V14 <- as.character(asclay$V14) #yr 12
+  # asclay$V15 <- as.character(asclay$V15)
+  # asclay$V16 <- as.character(asclay$V16) #yr 14
+  # asclay$V17<- as.character(asclay$V17)
+  #asclay$V18 <- as.character(asclay$V18) #yr 16
+  #asclay$V19 <- as.character(asclay$V19) #yr 17
+
+  asclaydep <- asclay$V2
+
+  # asc1998 <- raster(asclay$V3)
+  # asc1999 <- raster(asclay$V4)
+  # asc2000 <- raster(asclay$V5)
+  # asc2001 <- raster(asclay$V6)
+  # asc2002 <- raster(asclay$V7)
+  # asc2003 <- raster(asclay$V8)
+  asc2004 <- raster(asclay$V3)
+  asc2005 <- raster(asclay$V4)
+  asc2006 <- raster(asclay$V5)
+  asc2007 <- raster(asclay$V6)
+  asc2008 <- raster(asclay$V7)
+ # asc2009 <- raster(asclay$V14)
+  asc2010 <- raster(asclay$V8)
+  asc2011 <- raster(asclay$V9)
+  asc2012 <- raster(asclay$V10)
+  asc2013 <- raster(asclay$V11)
+  asc2014 <- raster(asclay$V12)
+  # asc2015 <- raster(asclay$V20)
+
+  stkasc <- stack(asc2004, asc2005, asc2006, asc2007, asc2008, asc2010, asc2011, asc2012, asc2013, asc2014) #, asc2015)   asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2009
+
+  avg <- mean(stkasc, na.rm = TRUE)
+
+  writeRaster(avg, paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_", asclaydep, ".asc"), overwrite=TRUE)
+  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_", asclaydep, ".png"))) 
+  dev.off()
+
+file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.asc"))
+file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.png"))
+file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.asc"))
+file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.png"))
+}
+```
+
+#plotting
+
+need some additional packages installed
+```{r pre-install & load packages, include=FALSE}
+packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer", "cowplot", "rasterVis")
+source("../src/ipak.R")
+ipak(packages)
+```
+
+# stacked permutation importance plot
+
+note if you change mind and want jackknife then the code is in the older plots for papers rmd
+
+first need to make the data plot-ready
+
+```{r create permutation & jack data}
+maxr_lst <- list.files("../output/maxent/models/test-comb/", pattern = "maxentresults.csv", full.names = TRUE)
+perm <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
+colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
+# jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
+# colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
+
+for (m in 1:length(maxr_lst)){
+  maxr <- read.csv(maxr_lst[m])
+  mth <- str_sub(maxr_lst[m], start = 35, end = 37) #selects the month from the filename
+  perm$Month[m] <- mth
+  # jack$Month[m] <- mth
+  maxt <- tail(maxr, 1)
+  perm$amo_sample[m] <- maxt$amo_sample.permutation.importance
+  perm$amo_winter[m] <- maxt$amo_winter.permutation.importance
+  perm$chl_surface[m] <- maxt$chl_surface.permutation.importance
+  perm$nao_prev[m] <- maxt$nao_prev.permutation.importance
+  perm$nao_sample[m] <- maxt$nao_sample.permutation.importance
+  perm$nao_winter[m] <- maxt$nao_winter.permutation.importance
+  perm$o2_depth[m] <- maxt$o2_depth.permutation.importance
+  perm$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
+  perm$temp_depth[m] <- maxt$temp_depth.permutation.importance
+
+  # jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
+  # jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
+  # jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
+  # jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
+  # jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
+  # jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
+  # jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
+  # jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
+  # jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
+}
+
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
+permu <- perm[match(mthord, perm$Month), ] #reorder based on mthord
+# jackk <- jack[match(mthord, jack$Month), ] #reorder based on mthord
+
+# rename the months so they are capitalised
+permu$Month <- capitalize(permu$Month)
+# jackk$Month <- capitalize(jackk$Month)
+
+# rename the col names so they are capitalised and shortened
+colnames(permu)[colnames(permu)=="amo_sample"] <- "AMO_Sample"
+colnames(permu)[colnames(permu)=="amo_winter"] <- "AMO_Winter"
+colnames(permu)[colnames(permu)=="chl_surface"] <- "Chl_Surface"
+colnames(permu)[colnames(permu)=="nao_prev"] <- "NAO_Prev"
+colnames(permu)[colnames(permu)=="nao_sample"] <- "NAO_Sample"
+colnames(permu)[colnames(permu)=="nao_winter"] <- "NAO_Winter"
+colnames(permu)[colnames(permu)=="o2_depth"] <- "O2_Depth"
+colnames(permu)[colnames(permu)=="salinity_depth"] <- "Salinity_Depth"
+colnames(permu)[colnames(permu)=="temp_depth"] <- "Temp_Depth"
+
+# colnames(jackk)[colnames(jackk)=="amo_sample"] <- "AMO_Sample"
+# colnames(jackk)[colnames(jackk)=="amo_winter"] <- "AMO_Winter"
+# colnames(jackk)[colnames(jackk)=="chl_surface"] <- "Chl_Surface"
+# colnames(jackk)[colnames(jackk)=="nao_prev"] <- "NAO_Prev"
+# colnames(jackk)[colnames(jackk)=="nao_sample"] <- "NAO_Sample"
+# colnames(jackk)[colnames(jackk)=="nao_winter"] <- "NAO_Winter"
+# colnames(jackk)[colnames(jackk)=="o2_depth"] <- "O2_Depth"
+# colnames(jackk)[colnames(jackk)=="salinity_depth"] <- "Salinity_Depth"
+# colnames(jackk)[colnames(jackk)=="temp_depth"] <- "Temp_Depth"
+
+write.csv(permu, "../output/maxent/models/test-comb/permutation_avg.csv", row.names = FALSE)
+# write.csv(jackk, "../output/maxent/models/test-comb/jack_avg.csv", row.names = FALSE)
+
+```
+
+now make the plot
+
+```{r permulation plot}
+permu <- read.csv("../output/maxent/models/test-comb/permutation_avg.csv")
+
+mthna <- c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
+
+permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
+permelt$variable <- factor(permelt$variable, levels = c("AMO_Sample", "AMO_Winter", "NAO_Sample", "NAO_Prev", "NAO_Winter", "Chl_Surface", "O2_Depth", "Salinity_Depth", "Temp_Depth")) #reorder the variables
+perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = 1, labels = c("AMO Sample", "AMO Winter", "NAO Sample", "NAO Prev", "NAO Winter", "Chl Surface", "O2 Depth", "Salinity Depth", "Temp Depth")) + theme_half_open(12) + scale_x_discrete(limits = mthna) + labs(y = "Permutation Importance (%)")
+png("../output/maxent/paper_plots/permutation_plot.png")
+print(perpl)
+dev.off()
+```
+
+#historgams of probability by depth
+
+count number of cells that are not NA in each raster layer that are inside the mcp
+
+```{r}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+wgsraster <- raster("../output/maxent/predictions/com-split/apr/cloglog_1998_04_0.50576.asc") #just chosen randomly to act as dummy file
+crs(wgsraster) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
+outsidemcp <- spTransform(outsidemcp, crs(wgsraster))
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/dec"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
+cumst <- data.frame(0, 0)
+colnames(cumst) <- c("depth", "no_cells")
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[6])
+  ascname <- names(cumasc)
+  cumst[cum, 1] <- str_sub(ascname, 13)
+  cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
+  ccnt <- freq(cumasc2, useNA='no')
+  ccnt <- sum(ccnt[ ,2])
+  cumst[cum, 2] <- ccnt
+}
+
+cumst$depth <- as.numeric(cumst$depth)
+cumst1 <- cumst[order(cumst$no_cells), ]
+
+write.csv(cumst1, "../output/maxent/paper_plots/prob_by_depth/cellsperlayermar.csv", row.names = FALSE)
+
+```
+
+first bin the probabilities for cells inside the mcp
+
+```{r bin probabilities}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+mthord <- c("dec") #"mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov",
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+
+  binprob <- function(x){
+    ifelse(x > 0 & x < 0.1, 0.1,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.1 & x < 0.2, 0.2,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.2 & x < 0.3, 0.3,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.3 & x < 0.4, 0.4,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.4 & x < 0.5, 0.5,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.5 & x < 0.6, 0.6,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.6 & x < 0.7, 0.7,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.7 & x < 0.8, 0.8,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_8_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_8_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.8 & x < 0.9, 0.9,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_9_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_9_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+
+  binprob <- function(x){
+    ifelse(x >= 0.9 & x <= 1, 1,
+    ifelse(x <= 0, 0, NA))
+  }
+
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_10_avgprob_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_10_avgprob_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+}
+```
+
+get data to plot probability by depth... 
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+binlst <- c(1:10)
+
+binprob <- function(x){
+  ifelse(x >0,  1,
+  ifelse(x == -9999, 0, NA))
+}
+
+for (m in 1:length(mthord)){
+  for (b in 1:length(binlst)){
+    mth <- mthord[m]
+    bin <- binlst[b]
+    # mth <- "dec"
+    # bin <- 1
+  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, "_avg.*asc"), full.names = TRUE)
+
+  binaryavgstack <- stack(ascdepth)
+  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+  stackstathigh[ ,1] <- c(1:49)
+  colnames(stackstathigh) <- c("asclayer", "prob_bin")
+
+  for (lyr in 1:nlayers(binaryavgstack)){
+    ascname <-  str_sub(ascdepth[lyr], 77, -5)
+    dlay <- binaryavgstack[[lyr]]
+    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
+    lyrsum <- cellStats(dlay, sum)
+    stackstathigh$prob_bin[lyr] <- lyrsum
+    stackstathigh$asclayer[lyr] <- ascname
+
+  }
+  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_probbin_", bin, "_cellcount_bydepth_avg.csv"), row.names = FALSE)
+  }
+}
+
+```
+
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_probbin_.*.csv"), full.names = TRUE)
+  a <- read.csv(probblst[1])
+  b <- read.csv(probblst[2])
+  c <- read.csv(probblst[3])
+  d <- read.csv(probblst[4])
+  e <- read.csv(probblst[5])
+  f <- read.csv(probblst[6])
+  g <- read.csv(probblst[7])
+  h <- read.csv(probblst[8])
+  i <- read.csv(probblst[9])
+  j <- read.csv(probblst[10])
+
+  stackstat <- cbind(a, b[ ,2], c[ ,2], d[ ,2], e[ , 2], f[ ,2], g[ ,2], h[ ,2], i[ ,2], j[ ,2])
+  colnames(stackstat) <- c("depth", "0 - 0.09", "0.1 - 0.19", "0.2 - 0.29", "0.3 - 0.39", "0.4 - 0.49", "0.5 - 0.59", "0.6 - 0.69", "0.7 - 0.79", "0.8 - 0.89", ">0.9")
+  stackstat <- melt(stackstat, id = "depth")
+  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]
+
+  colnames(stackstat) <- c("Depth", "Probability", "Cells")
+
+  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
+}
+
+```
+
+think about how to get proportions...
+
+```{r}
+mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
+  for (r in 1:nrow(a)){
+    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
+    # d1 <- str_sub(a$Depth[r], end = 4)
+    d1 <- a$Depth[r]
+	  for (ro in 1:nrow(cumst1)){
+      #d2 <- str_sub(cumst1$depth[ro], end = 4)
+	    d2 <- cumst1$depth[ro]
+      if (d1 == d2){
+          a$total_cells[r] <- cumst1$no_cells[ro]
+      }
+	  }
+    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
+  }
+    a$Depth <- as.numeric(a$Depth)
+
+  write.csv(a, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), row.names = FALSE)
+}
+
+```

### 2020-01-23 src/testing_file.Rmd
avg ascii layer complete. mcp reprojected

@@
 ```
+
+# mcp shape needs to be in wgs...
+
+```{r}
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis - in aea
+wgsraster <- raster("../output/maxent/predictions/com-split/apr/cloglog_1998_04_0.50576.asc") #just chosen randomly to act as dummy file
+crs(wgsraster) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
+outsidemcp <- spTransform(outsidemcp, crs(wgsraster))
+
+```

### 2020-01-23 src/comb-maxent.Rmd
fuck you, fuck you all fuckers

@@
 
-for (year in 1:nrow(modelmonth)){
-  timeslice <- modelmonth[year] #testing timeslice
-  yr <- str_sub(timeslice, -4) #the year being used for testing
-  train <- subset(prback, year != yr) #training data is all years but testing
-  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
-  test <-  subset(prback, year == yr) #creates the testing dataset
-  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
-  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
-  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  #mess_pr <- subset(train, occurrence == 1)
-  #mess_pr  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
-
-  if (yr == "1998"){
-    betaval <- 0.25
-  } else if (yr == "2001"){
-    betaval <- 0.75
-  } else if (yr == "2002" | yr == "2003"| yr == "2004"| yr == "2006"| yr == "2011"){
-    betaval <- 1
-  } else if (yr == "2005"){
-    betaval <- 1.25
-  } else if (yr == "1999" | yr == "2000"| yr == "2009"){
-    betaval <- 1.5
-  } else if (yr == "2007"){
-    betaval <- 1.75
-  } else if (yr == "2010"){
-    betaval <- 2
-  } else if (yr == "2008"){
-    betaval <- 2.25
-  }
-
-    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/test-comb/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
-
-    saveRDS(mod, file= paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))
-
-    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
-
-    #count number of non-zero coefficients, excluding the final four rows
-    coefile <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
-    coefile <- head(coefile, -4)
-    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
-    nocoeff <- nrow(coefile2)
-    bick <- sum(as.numeric(coefile2[, 2]))
-    bic <- (bick*log(sum(trainpa == 1)) - 2*11)
-
-    #no presences used in model (less those used for testing)
-    maxresults <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
-    notrainingpoints <- maxresults$X.Training.samples
-    testauc <- maxresults$Test.AUC
-    trainauc <- maxresults$Training.AUC
-
-    #natural log
-    samplepred <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
-    samplepred$nat_log <- log(samplepred$Raw.prediction)
-    sumsamplepred <- sum(samplepred$nat_log)
-
-    #test auc scores and other lovely things you might want to include
-    ev <- evaluate(p = testp, a= testb, model = mod)
-    evauc <- ev@auc
-    evcor <- ev@cor
-    evpcor <- ev@pcor
-    evnp <- ev@np
-    evna <- ev@na
-    evmaxtprtnr <- threshold(ev, 'spec_sens')
-    evmaxkap <- threshold(ev, 'kappa')
-    evno_omission <- threshold(ev, 'no_omission')
-    evprevalence <- threshold(ev, 'prevalence')
-    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
-    evsensitivity <- threshold(ev, 'sensitivity')
-
-    #continuous boyce index
-    predPres <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
-    predBg <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
-    predPres <- predPres$Cloglog.prediction
-    predBg <- predBg$Cloglog
-    source("../src/contBoyce.R")
-    cbi <- contBoyce(pres = predPres, bg = predBg)
-
-      aicetc$train_AUC <- trainauc #training only AUC - not tested
-      aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
-      aicetc$coefficients <- nocoeff #this is to calculate the AIC
-      aicetc$sum.of.logs <- sumsamplepred #this is to calculate the AIC
-      aicetc$test_AUC <- evauc #testing AUC
-      aicetc$test_evcor <- evcor #testing correlation coefficient
-      aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
-      aicetc$test_nopresence <- evnp 
-      aicetc$test_nobackground <- evna
-      aicetc$test_maxsumsenspec <- evmaxtprtnr
-      aicetc$test_maxkap <- evmaxkap
-      aicetc$test_no_omission <- evno_omission
-      aicetc$test_prevalence <- evprevalence
-      aicetc$test_equal_sens_spec <- evequal_sens_spec
-      aicetc$test_sensitivity <- evsensitivity
-      aicetc$tss <- max(ev@TPR + ev@TNR) - 1
-      aicetc$test_bic <- bic #not sure this is correct...
-      aicetc$beta <- betaval
-      aicetc$cbi <- cbi
-
-    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
-    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$no_trainng_points/(aicetc$no_trainng_points - aicetc$coefficients-1))))
-
-    #calculate minimal predicted area using ecospat
-    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
-    aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
-    aicetc$year <- yr
-
-    write.csv(aicetc, paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+# 
+# 
+# for (year in 1:nrow(modelmonth)){
+#   timeslice <- modelmonth[year] #testing timeslice
+#   yr <- str_sub(timeslice, -4) #the year being used for testing
+#   train <- subset(prback, year != yr) #training data is all years but testing
+#   trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
+#   test <-  subset(prback, year == yr) #creates the testing dataset
+#   testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
+#   testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+#   testb <- subset(test, occurrence == "0") #splits the testing dataset into background
+#   testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+#   #mess_pr <- subset(train, occurrence == 1)
+#   #mess_pr  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+#   train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+# 
+#   
+#   if (yr == "1998"){
+#     betaval <- 0.25
+#   } else if (yr == "2001"){
+#     betaval <- 0.75
+#   } else if (yr == "2002" | yr == "2003"| yr == "2004"| yr == "2006"| yr == "2011"){
+#     betaval <- 1
+#   } else if (yr == "2005"){
+#     betaval <- 1.25
+#   } else if (yr == "1999" | yr == "2000"| yr == "2009"){
+#     betaval <- 1.5
+#   } else if (yr == "2007"){
+#     betaval <- 1.75
+#   } else if (yr == "2010"){
+#     betaval <- 2
+#   } else if (yr == "2008"){
+#     betaval <- 2.25
+#   }
+# 
+# 
+#   
+#   
+#     mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/test-comb/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
+#     
+#     saveRDS(mod, file= paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))
+# 
+# 
+#     #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
+# 
+#     #count number of non-zero coefficients, excluding the final four rows
+#     coefile <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+#     coefile <- head(coefile, -4)
+#     coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+#     nocoeff <- nrow(coefile2)
+#     bick <- sum(as.numeric(coefile2[, 2]))
+#     bic <- (bick*log(sum(trainpa == 1)) - 2*11)
+#     
+#     #no presences used in model (less those used for testing)
+#     maxresults <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
+#     notrainingpoints <- maxresults$X.Training.samples
+#     testauc <- maxresults$Test.AUC
+#     trainauc <- maxresults$Training.AUC
+#     
+#     #natural log
+#     samplepred <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+#     samplepred$nat_log <- log(samplepred$Raw.prediction)
+#     sumsamplepred <- sum(samplepred$nat_log)
+#   
+#   
+#     #test auc scores and other lovely things you might want to include
+#     ev <- evaluate(p = testp, a= testb, model = mod)
+#     evauc <- ev@auc
+#     evcor <- ev@cor
+#     evpcor <- ev@pcor
+#     evnp <- ev@np
+#     evna <- ev@na
+#     evmaxtprtnr <- threshold(ev, 'spec_sens')
+#     evmaxkap <- threshold(ev, 'kappa')
+#     evno_omission <- threshold(ev, 'no_omission')
+#     evprevalence <- threshold(ev, 'prevalence')
+#     evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+#     evsensitivity <- threshold(ev, 'sensitivity')
+#     
+#     #continuous boyce index
+#     predPres <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+#     predBg <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
+#     predPres <- predPres$Cloglog.prediction
+#     predBg <- predBg$Cloglog
+#     source("../src/contBoyce.R")
+#     cbi <- contBoyce(pres = predPres, bg = predBg)
+#     
+#   
+#       aicetc$train_AUC <- trainauc #training only AUC - not tested
+#       aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
+#       aicetc$coefficients <- nocoeff #this is to calculate the AIC
+#       aicetc$sum.of.logs <- sumsamplepred #this is to calculate the AIC
+#       aicetc$test_AUC <- evauc #testing AUC
+#       aicetc$test_evcor <- evcor #testing correlation coefficient
+#       aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
+#       aicetc$test_nopresence <- evnp 
+#       aicetc$test_nobackground <- evna
+#       aicetc$test_maxsumsenspec <- evmaxtprtnr
+#       aicetc$test_maxkap <- evmaxkap
+#       aicetc$test_no_omission <- evno_omission
+#       aicetc$test_prevalence <- evprevalence
+#       aicetc$test_equal_sens_spec <- evequal_sens_spec
+#       aicetc$test_sensitivity <- evsensitivity
+#       aicetc$tss <- max(ev@TPR + ev@TNR) - 1
+#       aicetc$test_bic <- bic #not sure this is correct...
+#       aicetc$beta <- betaval
+#       aicetc$cbi <- cbi
+#       
+#     aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
+#     aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$no_trainng_points/(aicetc$no_trainng_points - aicetc$coefficients-1))))
+#     
+#     #calculate minimal predicted area using ecospat
+#     aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
+#     aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
+#     aicetc$year <- yr
+# 
+# 
+#     write.csv(aicetc, paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
 
@@
 
-clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/dec"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
+clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/mar"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
 cumst <- data.frame(0, 0)
@@
 for (cum in 1:length(clogasc_list)){
-  cumasc <- raster(clogasc_list[6])
+  cumasc <- raster(clogasc_list[cum])
   ascname <- names(cumasc)
@@
 
-write.csv(cumst1, "../output/maxent/paper_plots/prob_by_depth/cellsperlayermar.csv", row.names = FALSE)
+write.csv(cumst1, "../output/maxent/paper_plots/cellsperlayer.csv", row.names = FALSE)
 

### 2020-01-23 src/testing_file.Rmd
fuck you, fuck you all fuckers

@@
 ```
+
+#need to reclassify cells outside the mcp to NA for predictions
+```{r}
+outsidemcp <- shapefile("../data/env/landp/outside-mcp-wgs84.shp") #made in arcgis - in aea
+
+ascdirs <- list.dirs("../data/env/ascii-original", full.names = FALSE, recursive = FALSE)
+
+for (ad in 1:length(ascdirs)){
+  timeslice <- ascdirs[ad]
+  ascfiles <- list.files(paste0("../data/env/ascii-original/", timeslice), pattern = "\\.asc$", full.names = TRUE)
+  for (file in 1:length(ascfiles)){
+    ras<- raster(ascfiles[file])
+    crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
+    ascname <- str_sub(ascfiles[file], start = 38)
+    ras[outsidemcp] <- 500000
+    ras[ras == 500000] <- NA
+    writeRaster(ras, paste0("../data/env/asciitrim/", ascdirs[ad], "/", ascname), overwrite=TRUE)
+  }
+}
+
+```
+
+test <- raster("../output/maxent/predictions/com-split/apr/cloglog_1998_04_0.50576.asc") #just chosen randomly to act as dummy file
+crs(test) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
+
+test[outsidemcp] <- 500000
+test[test == 500000] <- NA
+plot(test)
+
+#if need to predict from the model...
+
+```{r}
+
+#monthlist <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
+
+modlist <- list.dirs("../output/maxent/models/test-comb", full.names = FALSE, recursive = FALSE) 
+#modlist <- modlist[ grepl(mthna, modlist) ]
+
+for (mth in 1:length(modlist)){
+    mthna <- str_sub(modlist[mth], end = 3)
+    yr <- str_sub(modlist[mth], start = 4)
+    timeslice <- modlist[mth]
+    mod <- readRDS(paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))
+    unlayer <- read.csv(paste0("../data/env/test-comb/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
+
+    unlayer_yr <- subset(unlayer, ascyearslist == yr)
+    #unlayer_yr <- unlayer_yr[-c(1), ]
+    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
+    stackdeplst <- list()
+    stackdep <- subset(unlayer_yr, select = c(stackname))
+    stackdeplst[[yr]] <- stackdep
+    write.csv(stackdep, paste0("../output/maxent/models/test-comb/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)
+
+    for (td in 1:nrow(unlayer_yr)){
+        temp_depth <- raster(unlayer_yr$temp_depth[td])
+        crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+        names(temp_depth) <- "temp_depth"
+        #temp_depth <- projectRaster(temp_depth, aea)
+        chl_surface <- raster(unlayer_yr$chl_surface[td])
+        crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+        names(chl_surface) <- "chl_surface"
+        #chl_depth <- projectRaster(chl_depth, aea)
+        o2_depth <- raster(unlayer_yr$o2_depth[td])
+        crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+        names(o2_depth) <- "o2_depth"
+        #o2_depth <- projectRaster(o2_depth, aea)
+        salinity_depth <- raster(unlayer_yr$salinity_depth[td])
+        crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+        names(salinity_depth) <- "salinity_depth"
+        #salinity_depth <- projectRaster(salinity_depth, aea)
+        nao_sample <- raster(unlayer_yr$nao_sample[td])
+        crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+        names(nao_sample) <- "nao_sample"
+        #nao_sample <- projectRaster(nao_sample, aea)
+        nao_prev <- raster(unlayer_yr$nao_prev[td])
+        crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+        names(nao_prev) <- "nao_prev"
+        #nao_prev <- projectRaster(nao_prev, aea)
+        nao_winter <- raster(unlayer_yr$nao_winter[td])
+        crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+        names(nao_winter) <- "nao_winter"
+        #nao_winter <- projectRaster(nao_winter, aea)
+        amo_sample <- raster(unlayer_yr$amo_sample[td])
+        crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
+        names(amo_sample) <- "amo_sample"
+        #amo_sample <- projectRaster(amo_sample, aea)
+        amo_winter <- raster(unlayer_yr$amo_winter[td])
+        crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+        names(amo_winter) <- "amo_winter"
+        #amo_winter <- projectRaster(amo_winter, aea)
+        stackn <- unlayer_yr$stackname[td]
+        stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
+
+        #stackSave(test,unlayer_yr$stackname[td])
+        pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=cloglog')) 
+        # predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
+        prplot <- plot(pred, png(paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".png")))
+        dev.off() # stops automatic saving of the plot to a png
+    }
+}
+```

### 2020-01-25 src/comb-maxent.Rmd
moving forward... back to where i was

--- 2020-01-23 src/comb-maxent.Rmd
@@
 ipak(packages)
+```
+
+#need to reclassify cells outside the mcp to NA for predictions
+```{r}
+outsidemcp <- shapefile("../data/env/landp/outside-mcp-wgs84.shp") #made in arcgis - in aea
+
+ascdirs <- list.dirs("../data/env/ascii-original", full.names = FALSE, recursive = FALSE)
+
+for (ad in 1:length(ascdirs)){
+  timeslice <- ascdirs[ad]
+  ascfiles <- list.files(paste0("../data/env/ascii-original/", timeslice), pattern = "\\.asc$", full.names = TRUE)
+  for (file in 1:length(ascfiles)){
+    ras<- raster(ascfiles[file])
+    crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
+    ascname <- str_sub(ascfiles[file], start = 38)
+    ras[outsidemcp] <- 500000
+    ras[ras == 500000] <- NA
+    writeRaster(ras, paste0("../data/env/asciitrim/", ascdirs[ad], "/", ascname), overwrite=TRUE)
+  }
+}
+
 ```
@@
 ```{r avg asc layers cloglog}
-mthna <- "aug" #change me
+mthna <- "dec" #change me
 depthlayers <- read.csv(paste0("../output/maxent/predictions/com-split/", mthna, "/cloglog_asclayerbydepth.csv"), header = TRUE)
@@
   asclay$V12 <- as.character(asclay$V12) #yr 10
-  # asclay$V13 <- as.character(asclay$V13)
-  # asclay$V14 <- as.character(asclay$V14) #yr 12
-  # asclay$V15 <- as.character(asclay$V15)
-  # asclay$V16 <- as.character(asclay$V16) #yr 14
-  # asclay$V17<- as.character(asclay$V17)
+  asclay$V13 <- as.character(asclay$V13)
+  asclay$V14 <- as.character(asclay$V14) #yr 12
+  asclay$V15 <- as.character(asclay$V15)
+  asclay$V16 <- as.character(asclay$V16) #yr 14
+  #asclay$V17<- as.character(asclay$V17)
   #asclay$V18 <- as.character(asclay$V18) #yr 16
@@
 
-  # asc1998 <- raster(asclay$V3)
-  # asc1999 <- raster(asclay$V4)
-  # asc2000 <- raster(asclay$V5)
-  # asc2001 <- raster(asclay$V6)
-  # asc2002 <- raster(asclay$V7)
-  # asc2003 <- raster(asclay$V8)
-  asc2004 <- raster(asclay$V3)
-  asc2005 <- raster(asclay$V4)
-  asc2006 <- raster(asclay$V5)
-  asc2007 <- raster(asclay$V6)
-  asc2008 <- raster(asclay$V7)
- # asc2009 <- raster(asclay$V14)
-  asc2010 <- raster(asclay$V8)
-  asc2011 <- raster(asclay$V9)
-  asc2012 <- raster(asclay$V10)
-  asc2013 <- raster(asclay$V11)
-  asc2014 <- raster(asclay$V12)
-  # asc2015 <- raster(asclay$V20)
+  asc1 <- raster(asclay$V3) #1
+  asc2 <- raster(asclay$V4)
+  asc3 <- raster(asclay$V5) #3
+  asc4 <- raster(asclay$V6)
+  asc5 <- raster(asclay$V7) #5
+  asc6 <- raster(asclay$V8)
+  asc7 <- raster(asclay$V9) #7
+  asc8 <- raster(asclay$V10)
+  asc9 <- raster(asclay$V11) #9
+  asc10 <- raster(asclay$V12)
+  asc11 <- raster(asclay$V13) #11
+  asc12 <- raster(asclay$V14)
+  asc13 <- raster(asclay$V15) #13
+  asc14 <- raster(asclay$V16)
+  #asc15 <- raster(asclay$V17) #15
+ # asc16 <- raster(asclay$V18)
+ # asc17 <- raster(asclay$V19)
 
-  stkasc <- stack(asc2004, asc2005, asc2006, asc2007, asc2008, asc2010, asc2011, asc2012, asc2013, asc2014) #, asc2015)   asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2009
+  stkasc <- stack(asc1, asc2, asc3, asc4, asc5, asc6, asc7, asc8, asc9, asc10, asc11, asc12, asc13, asc14) #, asc15) #, asc16) #, asc2015)   asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2009
 
@@
 
-file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.asc"))
-file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.png"))
-file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.asc"))
-file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.png"))
+file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.asc"), paste0("../output/maxent/predictions/com-split", mthna, "/avg_cloglog_120.asc"))
+file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.png"), paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_120.png"))
+file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.asc"), paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_41.18.asc"))
+file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.png"), paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_41.18.png"))
 }
@@
 ```{r}
-outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
-wgsraster <- raster("../output/maxent/predictions/com-split/apr/cloglog_1998_04_0.50576.asc") #just chosen randomly to act as dummy file
-crs(wgsraster) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
-outsidemcp <- spTransform(outsidemcp, crs(wgsraster))
+# outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+# wgsraster <- raster("../output/maxent/predictions/com-split/apr/cloglog_1998_04_0.50576.asc") #just chosen randomly to act as dummy file
+# crs(wgsraster) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
+# outsidemcp <- spTransform(outsidemcp, crs(wgsraster))
 
-clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/dec"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
+clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/mar"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
 cumst <- data.frame(0, 0)
@@
 for (cum in 1:length(clogasc_list)){
-  cumasc <- raster(clogasc_list[6])
+  cumasc <- raster(clogasc_list[cum])
   ascname <- names(cumasc)
   cumst[cum, 1] <- str_sub(ascname, 13)
-  cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
-  ccnt <- freq(cumasc2, useNA='no')
+  #cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
+  ccnt <- freq(cumasc, useNA='no')
   ccnt <- sum(ccnt[ ,2])
@@
 ```{r bin probabilities}
-outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
-mthord <- c("dec") #"mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov",
+#outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
 
@@
   mth <- mthord[m]
-  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
 
@@
     ascname <- names(cumasc)
-    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+    #cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
     probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
@@
     ascname <- names(cumasc)
-    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
+   # cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
     probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
@@
   stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-  stackstathigh[ ,1] <- c(1:49)
+  stackstathigh[ ,1] <- c(1:47)
   colnames(stackstathigh) <- c("asclayer", "prob_bin")
@@
 mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-cumst1 <- read.csv("../output/maxent/paper_plots/prob_by_depth/cellsperlayer.csv", header = TRUE)
+cumst1 <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE)
 
@@
 ```
+
+GO FROM HERE

### 2020-01-25 src/testing_file.Rmd
moving forward... back to where i was

@@
 
-for (mth in 1:length(modlist)){
+for (mth in 65:length(modlist)){
     mthna <- str_sub(modlist[mth], end = 3)

### 2020-01-29 src/comb-maxent.Rmd
progress on plots

@@
 ```{r}
-# outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
-# wgsraster <- raster("../output/maxent/predictions/com-split/apr/cloglog_1998_04_0.50576.asc") #just chosen randomly to act as dummy file
-# crs(wgsraster) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
-# outsidemcp <- spTransform(outsidemcp, crs(wgsraster))
 
@@
 ```{r bin probabilities}
-#outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
 mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
@@
     ascname <- names(cumasc)
-    #cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
     probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
@@
 
-GO FROM HERE
+histogram of probability by depth - by proportion
+
+based on Remi's code
+```{r}
+require(data.table)
+require(purrr)
+require(lubridate)
+require(tidyverse)
+
+filenames <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)
+
+all <- map_dfr(filenames, function(x) fread(x) %>% 
+                       mutate(Month=substr(basename(x),1,3))) %>% 
+  mutate(Probability = gsub(">0.9", "0.9 - 1", Probability),
+         Depth = factor(round(Depth,1)),
+         Depth = factor(Depth,ordered = TRUE, levels = rev(levels(Depth))),
+         Month=factor(months(parse_date_time(Month,"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))))
+
+#create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
+dlst <- unique(all$Depth)
+dlst <- as.character(dlst)
+dlst <- as.numeric(dlst)
+dlst <- sort(dlst)
+
+#make the plot
+  ggplot(all) +
+  geom_col(aes(Depth, p_cells, fill = Probability),width=1) +
+  coord_flip()+
+  facet_wrap(~Month,ncol=5)+
+  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
+  scale_fill_viridis(option="magma", discrete = TRUE, direction = -1, guide=guide_legend(nrow=2, byrow = TRUE))+
+  theme_classic()+
+  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
+  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47]))
+
+ggsave("../output/maxent/paper_plots/test.png", dpi = 600)
+```
+
+# variable maps
+- temp
+- chl 
+- o2
+- sal
+- 4 different months
+- relationship of variables to each other
+- pick a depth (say the 11.7737 m)
+
+months are mar, jun, sep, dec
+pick 3 diff years say 1998, 2006, 2013
+
+```{r individual maps}
+#aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/land-wgs84v2.shp") #made in arcgis
+crs(s) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+#acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp-wgs84.shp")
+#ss <- raster::intersect(s, mcp)  #trim land file to mcp
+
+plotlist <- list()
+
+yrlst <- c("1998", "2006", "2013")
+mthlst <- c("03", "06", "09", "12")
+varlst <- c("temp", "salinity", "o2")
+
+pn <- 1
+for (y in 1:length(yrlst)){
+  yr <- yrlst[y]
+  fdr <- paste0("../data/env/ascii-original/ascii", yr, "/")
+  for (m in 1:length(mthlst)){
+    mt <- mthlst[m]
+    for (v in 1:length(varlst)){
+      var <- varlst[v]
+      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_11.7737.asc"))
+      lyrnm <- names(ras)
+      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      #ras <- projectRaster(ras, aea)
+      ras <- raster::mask(ras, mcp)
+
+      plotlist[[pn]] <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + latticeExtra::layer(sp.polygons(s, fill = "dark grey", col = "transparent") ) #changed ss to s
+      pn <- pn+1
+
+     # png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
+     # print(p)
+     # dev.off()
+    }
+  }
+}
+
+#chl is done seperately as it is surface
+
+yrlst <- c("1998", "2006", "2013")
+mthlst <- c("03", "06", "09", "12")
+varlst <- c("chl")
+
+for (y in 1:length(yrlst)){
+  yr <- yrlst[y]
+  fdr <- paste0("../data/env/ascii-original/ascii", yr, "/")
+  for (m in 1:length(mthlst)){
+    mt <- mthlst[m]
+    for (v in 1:length(varlst)){
+      var <- varlst[v]
+      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_0.50576.asc"))
+      lyrnm <- names(ras)
+      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+      ras <- raster::mask(ras, mcp)
+      plotlist[[pn]] <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + latticeExtra::layer(sp.polygons(s, fill = "dark grey", col = "transparent"))
+      pn <- pn+1
+
+   #   png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
+    #  print(p)
+    #  dev.off()
+    }
+  }
+}
+```

### 2020-01-29 src/plots for paper.Rmd
progress on plots

@@
 binaryprob_list <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = ".*_bin_0.5_2d.asc", full.names = TRUE)
-sum2d <- setNames(data.frame(matrix(ncol = 2, nrow = binl)), c("month", "count"))
+sum2d <- setNames(data.frame(matrix(ncol = 2, nrow = binaryprob_list))), c("month", "count"))
 

### 2020-01-29 src/testing_file.Rmd
progress on plots

@@
 
-aea <- raster("../output/env/aea.tif") #the correct CRS
-s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
-acrs <- aea@crs
-mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
-ss <- raster::intersect(s, mcp)  #trim land file to mcp
+#aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/land-wgs84v2.shp") #made in arcgis
+crs(s) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+#acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp-wgs84.shp")
+#ss <- raster::intersect(s, mcp)  #trim land file to mcp
 
@@
 
-mar11 <- raster("../output/maxent/predictions/mar/avg_cloglog_11.7737.asc")
-mar61 <- raster("../output/maxent/predictions/mar/avg_cloglog_61.1128.asc")
-mar199 <- raster("../output/maxent/predictions/mar/avg_cloglog_199.79.asc")
-mar628 <- raster("../output/maxent/predictions/mar/avg_cloglog_628.026.asc")
+mar11 <- raster("../output/maxent/predictions/com-split/mar/avg_cloglog_11.7737.asc")
+mar61 <- raster("../output/maxent/predictions/com-split/mar/avg_cloglog_61.1128.asc")
+mar199 <- raster("../output/maxent/predictions/com-split/mar/avg_cloglog_199.79.asc")
+mar628 <- raster("../output/maxent/predictions/com-split/mar/avg_cloglog_628.026.asc")
 mar11@crs <- acrs
@@
 
-jun11 <- raster("../output/maxent/predictions/jun/avg_cloglog_11.7737.asc")
-jun61 <- raster("../output/maxent/predictions/jun/avg_cloglog_61.1128.asc")
-jun199 <- raster("../output/maxent/predictions/jun/avg_cloglog_199.79.asc")
-jun628 <- raster("../output/maxent/predictions/jun/avg_cloglog_628.026.asc")
+jun11 <- raster("../output/maxent/predictions/com-split/jun/avg_cloglog_11.7737.asc")
+jun61 <- raster("../output/maxent/predictions/com-split/jun/avg_cloglog_61.1128.asc")
+jun199 <- raster("../output/maxent/predictions/com-split/jun/avg_cloglog_199.79.asc")
+jun628 <- raster("../output/maxent/predictions/com-split/jun/avg_cloglog_628.026.asc")
 jun11@crs <- acrs
@@
 
-sep11 <- raster("../output/maxent/predictions/sep/avg_cloglog_11.7737.asc")
-sep61 <- raster("../output/maxent/predictions/sep/avg_cloglog_61.1128.asc")
-sep199 <- raster("../output/maxent/predictions/sep/avg_cloglog_199.79.asc")
-sep628 <- raster("../output/maxent/predictions/sep/avg_cloglog_628.026.asc")
+sep11 <- raster("../output/maxent/predictions/com-split/sep/avg_cloglog_11.7737.asc")
+sep61 <- raster("../output/maxent/predictions/com-split/sep/avg_cloglog_61.1128.asc")
+sep199 <- raster("../output/maxent/predictions/com-split/sep/avg_cloglog_199.79.asc")
+sep628 <- raster("../output/maxent/predictions/com-split//sep/avg_cloglog_628.026.asc")
 sep11@crs <- acrs
@@
 
-dec11 <- raster("../output/maxent/predictions/dec/avg_cloglog_11.7737.asc")
-dec61 <- raster("../output/maxent/predictions/dec/avg_cloglog_61.1128.asc")
-dec199 <- raster("../output/maxent/predictions/dec/avg_cloglog_199.79.asc")
-dec628 <- raster("../output/maxent/predictions/dec/avg_cloglog_628.026.asc")
+dec11 <- raster("../output/maxent/predictions/com-split/dec/avg_cloglog_11.7737.asc")
+dec61 <- raster("../output/maxent/predictions/com-split/dec/avg_cloglog_61.1128.asc")
+dec199 <- raster("../output/maxent/predictions/com-split/dec/avg_cloglog_199.79.asc")
+dec628 <- raster("../output/maxent/predictions/com-split/dec/avg_cloglog_628.026.asc")
 dec11@crs <- acrs
@@
 
-#2d plots
-mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
-mar2d <- raster::mask(mar2d, mcp)
-mar2d@crs <- acrs
-jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
-jun2d <- raster::mask(jun2d, mcp)
-jun2d@crs <- acrs
-sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
-sep2d <- raster::mask(sep2d, mcp)
-sep2d@crs <- acrs
-dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")
-dec2d <- raster::mask(dec2d, mcp)
-dec2d@crs <- acrs
+# #2d plots
+# mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
+# mar2d <- raster::mask(mar2d, mcp)
+# mar2d@crs <- acrs
+# jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
+# jun2d <- raster::mask(jun2d, mcp)
+# jun2d@crs <- acrs
+# sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
+# sep2d <- raster::mask(sep2d, mcp)
+# sep2d@crs <- acrs
+# dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")
+# dec2d <- raster::mask(dec2d, mcp)
+# dec2d@crs <- acrs
 
@@
 
-p <- levelplot(pstack, ylab = NULL, xlab = NULL, colorkey=list(space="bottom", at = my.at), par.settings = myTheme, layout = c(4, 4), margin = FALSE, scales=list(draw=FALSE), col.regions=rev(inferno(10)), at= my.at, names.attr = cls) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+p <- levelplot(pstack, ylab = NULL, xlab = NULL, colorkey=list(space="bottom", at = my.at), par.settings = myTheme, layout = c(4, 4), margin = FALSE, scales=list(draw=FALSE), col.regions=rev(inferno(10)), at= my.at, names.attr = cls) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent")) #+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
 png("../output/maxent/paper_plots/bydepth.png", height = 1024, width = 1024)
@@
 ```
+
+#probabilty maps (selected)
+
+xxxx
+```{r}
+library(raster)
+library(rgdal)
+library(rasterVis)
+library(colorspace)
+
+aea <- raster("../output/env/aea.tif") #the correct CRS
+s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
+acrs <- aea@crs
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp_capelin_100_poly.shp")
+ss <- raster::intersect(s, mcp)  #trim land file to mcp
+
+mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
+apr2d <- raster("../output/maxent/paper_plots/hotspot_mapping/apr_bin_0.5_2d_single.asc")
+may2d <- raster("../output/maxent/paper_plots/hotspot_mapping/may_bin_0.5_2d_single.asc")
+jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
+jul2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jul_bin_0.5_2d_single.asc")
+aug2d <- raster("../output/maxent/paper_plots/hotspot_mapping/aug_bin_0.5_2d_single.asc")
+sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
+oct2d <- raster("../output/maxent/paper_plots/hotspot_mapping/oct_bin_0.5_2d_single.asc")
+nov2d <- raster("../output/maxent/paper_plots/hotspot_mapping/nov_bin_0.5_2d_single.asc")
+dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")
+
+#trim to mcp
+mar2d <- raster::mask(mar2d, mcp)
+apr2d <- raster::mask(apr2d, mcp)
+may2d <- raster::mask(may2d, mcp)
+jun2d <- raster::mask(jun2d, mcp)
+jul2d <- raster::mask(jul2d, mcp)
+aug2d <- raster::mask(aug2d, mcp)
+sep2d <- raster::mask(sep2d, mcp)
+oct2d <- raster::mask(oct2d, mcp)
+nov2d <- raster::mask(nov2d, mcp)
+dec2d <- raster::mask(dec2d, mcp)
+
+#add crs
+mar2d@crs <- acrs
+apr2d@crs <- acrs
+may2d@crs <- acrs
+jun2d@crs <- acrs
+jul2d@crs <- acrs
+aug2d@crs <- acrs
+sep2d@crs <- acrs
+oct2d@crs <- acrs
+nov2d@crs <- acrs
+dec2d@crs <- acrs
+
+maplst <- list.files("../output/maxent/predictions/may", pattern = "^avg_clog.*.asc", full.names = TRUE)
+maplst2 <- list.files("../output/maxent/predictions/sep", pattern = "^avg_clog.*.asc", full.names = TRUE)
+brk <- seq(0, 1, 0.1) #for legend
+lgc <- sequential_hcl(10, palette = "YlOrRd") #colour scheme from colorspace package
+outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
+
+pras6 <- raster(maplst[37])
+pras180 <- raster(maplst[14])
+pras370 <- raster(maplst[28])
+pras697 <- raster(maplst[41])
+p2ras6 <- raster(maplst2[37])
+p2ras180 <- raster(maplst2[14])
+p2ras370 <- raster(maplst2[28])
+p2ras697 <- raster(maplst2[41])
+crs(pras6) <- apr
+crs(pras180) <- apr
+crs(pras370) <- apr
+```
+
+```{r}
+crs(pras697) <- apr
+crs(p2ras6) <- apr
+crs(p2ras180) <- apr
+crs(p2ras370) <- apr
+crs(p2ras697) <- apr
+ascname6 <- names(pras6)
+ascname180 <- names(pras180)
+ascname370 <- names(pras370)
+ascname697 <- names(pras697)
+a2scname6 <- names(p2ras6)
+a2scname180 <- names(p2ras180)
+a2scname370 <- names(p2ras370)
+a2scname697 <- names(p2ras697)
+
+# pstack <- stack(pras6, pras180, pras370, pras697) #stack the layers
+# 
+# pras1 <- raster(maplst[1])
+# pras2 <- raster(maplst[2])
+# pras3 <- raster(maplst[3])
+# pras4 <- raster(maplst[4])
+# crs(pras1) <- apr
+# crs(pras2) <- apr
+# crs(pras3) <- apr
+# crs(pras4) <- apr
+
+pstack <- stack(pras6, pras180, pras370, pras697, p2ras6, p2ras180, p2ras370, p2ras697) #stack the layers
+
+cls <- c("6m", "180m", "370m", "697m")
+rws <- c("May", "Sep")
+mapTheme <- rasterTheme(region = rev(brewer.pal(10,"RdYlGn")))
+my.at <- seq(0, 1, by = 0.1)
+
+p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(4, 2), names.attr = c("6m", "180m", "370m", "697m", "", "", "", ""), margin = FALSE)
+p + layer(sp.polygons(s, fill = "black")) + layer (sp.polygons(outsidemcp, fill = redtrans, col = NA))
+
+#dev.copy(png, (paste0("../output/maxent/predictions/testplots/rasterVis.png")), width = 1920,height = 1080)
+#dev.off() # stops automatic saving of the plot to a png))
+
+```
+
+```{r}
+streams <- raster::shapefile("../data/env/landp/land-wgs84v2.shp")
+ras <-
+
+ggplot() + layer_spatial(data = streams, fill = "dark grey", colour = "transparent") + layer_spatial() + theme_void()
+
+```
+
+#2d maps for comb-maxent
+
+#2d hotspot maps
+
+first create layers that are 0.5 and above
+
+```{r habitat classification on probability - >0.5xx}
+
+binprob <- function(x){
+  ifelse(x >= 0.5, 1,
+  ifelse(x < 0.5, 0, NA)) 
+}
+
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mthna <- mthord[m]
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mthna), pattern = "^avg_cloglog_.*asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    ascname <- gsub("avg_cloglog_", "", ascname)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna,  "_binary_0.5_", ascname, ".asc"), overwrite = TRUE)
+     plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_binary_0.5_", ascname, ".png")), overwrite = TRUE)
+       dev.off() # stops automatic saving of the plot to a png
+  }
+}
+
+```
+
+then create an summary map....
+```{r >0.5 prob monthly 2D representation}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mthna <- mthord[m]
+  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_binary_0.5_.*.asc$"), full.names = TRUE)
+  binstack <- stack(binaryprob_list)
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.png")))
+  dev.off()
+}
+
+```
+
+now create summary maps where values are binary - not counts of cells
+```{r}
+binprob <- function(x){
+  ifelse(x > 0, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+  clogasc_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = "*_bin_0.5_2d.asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+```
+
+then plot the maps
+
+```{r}
+
+ss <- shapefile("../data/env/landp/land-wgs84v2.shp") #made in arcgis
+crs(ss) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
+acrs <- crs(ss)
+mcp <- shapefile("../output/bio/minimum_convex_poly/mcp-wgs84.shp")
+#ss <- raster::intersect(s, mcp)  #trim to mcp
+
+mar2d <- raster("../output/maxent/paper_plots/hotspot_mapping/mar_bin_0.5_2d_single.asc")
+apr2d <- raster("../output/maxent/paper_plots/hotspot_mapping/apr_bin_0.5_2d_single.asc")
+may2d <- raster("../output/maxent/paper_plots/hotspot_mapping/may_bin_0.5_2d_single.asc")
+jun2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jun_bin_0.5_2d_single.asc")
+jul2d <- raster("../output/maxent/paper_plots/hotspot_mapping/jul_bin_0.5_2d_single.asc")
+aug2d <- raster("../output/maxent/paper_plots/hotspot_mapping/aug_bin_0.5_2d_single.asc")
+sep2d <- raster("../output/maxent/paper_plots/hotspot_mapping/sep_bin_0.5_2d_single.asc")
+oct2d <- raster("../output/maxent/paper_plots/hotspot_mapping/oct_bin_0.5_2d_single.asc")
+nov2d <- raster("../output/maxent/paper_plots/hotspot_mapping/nov_bin_0.5_2d_single.asc")
+dec2d <- raster("../output/maxent/paper_plots/hotspot_mapping/dec_bin_0.5_2d_single.asc")
+
+#trim to mcp
+mar2d <- raster::mask(mar2d, mcp)
+apr2d <- raster::mask(apr2d, mcp)
+may2d <- raster::mask(may2d, mcp)
+jun2d <- raster::mask(jun2d, mcp)
+jul2d <- raster::mask(jul2d, mcp)
+aug2d <- raster::mask(aug2d, mcp)
+sep2d <- raster::mask(sep2d, mcp)
+oct2d <- raster::mask(oct2d, mcp)
+nov2d <- raster::mask(nov2d, mcp)
+dec2d <- raster::mask(dec2d, mcp)
+
+#add crs
+mar2d@crs <- acrs
+apr2d@crs <- acrs
+may2d@crs <- acrs
+jun2d@crs <- acrs
+jul2d@crs <- acrs
+aug2d@crs <- acrs
+sep2d@crs <- acrs
+oct2d@crs <- acrs
+nov2d@crs <- acrs
+dec2d@crs <- acrs
+
+#ratify to categorical
+mar2d <- ratify(mar2d)
+rat <- levels(mar2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(mar2d) <- rat
+apr2d <- ratify(apr2d)
+rat <- levels(apr2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(apr2d) <- rat
+may2d <- ratify(may2d)
+rat <- levels(may2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(may2d) <- rat
+jun2d <- ratify(jun2d)
+rat <- levels(jun2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(jun2d) <- rat
+jul2d <- ratify(jul2d)
+rat <- levels(jul2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(jul2d) <- rat
+aug2d <- ratify(aug2d)
+rat <- levels(aug2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(aug2d) <- rat
+sep2d <- ratify(sep2d)
+rat <- levels(sep2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(sep2d) <- rat
+oct2d <- ratify(oct2d)
+rat <- levels(oct2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(oct2d) <- rat
+nov2d <- ratify(nov2d)
+rat <- levels(nov2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(nov2d) <- rat
+dec2d <- ratify(dec2d)
+rat <- levels(dec2d)[[1]]
+rat$suitability <- c("< 0.5", "> 0.5")
+levels(dec2d) <- rat
+
+maplst <- list(mar2d, apr2d, may2d, jun2d, jul2d, aug2d, sep2d, oct2d, nov2d, dec2d)
+
+pstack <- stack(maplst)
+
+# my.at <- seq(0, 1, by = 0.1)
+# mapTheme <- infernoTheme(region = rev(inferno(10)))
+
+p <- levelplot(pstack, col.regions=c('white', 'red'), ylab = NULL, xlab = NULL, scales=list(draw=FALSE), layout = c(5, 2), names.attr = c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), margin = FALSE) + latticeExtra::layer(sp.polygons(ss, fill = "dark grey", col = "transparent"))+ latticeExtra::layer(sp.polygons(mcp, col = "dark grey"))
+plot(p)
+```
+
+```{r}
+require(data.table)
+require(purrr)
+require(lubridate)
+require(tidyverse)
+
+filenames <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)
+
+all <- map_dfr(filenames, function(x) fread(x) %>% 
+                       mutate(Month=substr(basename(x),1,3))) %>% 
+  mutate(Probability = gsub(">0.9", "0.9 - 1", Probability),
+         Depth = factor(round(Depth,1)),
+         Depth = factor(Depth,ordered = TRUE, levels = rev(levels(Depth))),
+         Month=factor(months(parse_date_time(Month,"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))))
+
+#create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
+dlst <- unique(all$Depth)
+dlst <- as.character(dlst)
+dlst <- as.numeric(dlst)
+dlst <- sort(dlst)
+
+  ggplot(all) +
+  geom_col(aes(Depth, p_cells, fill = Probability),width=1) +
+  coord_flip()+
+  facet_wrap(~Month,ncol=5)+
+  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
+  scale_fill_viridis(option="magma", discrete = TRUE, direction = -1, guide=guide_legend(nrow=2, byrow = TRUE))+
+  theme_classic()+
+  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
+  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47]))
+
+ggsave("../output/maxent/paper_plots/test.png", dpi = 600)
+```
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+s <- st_read("../data/env/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj)
+
+mcp <- st_read("../output/bio/minimum_convex_poly/mcp-wgs84.shp")
+
+filelist <- list.files("../data/plot2-probmaps/",pattern="avg_cloglog*",recursive = TRUE,include.dirs = TRUE)
+
+x <- stack(file.path("../data/plot2-probmaps/",filelist)) %>% 
+  mask(mcp)
+
+names(x) <- filelist
+
+df <- x %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  mutate(variable=gsub(".asc","",variable),
+         month=factor(months(parse_date_time(gsub("\\..*$", "",variable),"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         depth=as.numeric(gsub("^.*_", "",variable)))
+
+#remi code stops here
+
+df$month <- str_sub(df$variable, start = 16, end = 18)
+
+CapStr <- function(y) {
+  c <- strsplit(y, " ")[[1]]
+  paste(toupper(substring(c, 1,1)), substring(c, 2),
+      sep="", collapse=" ")
+}
+df$month <- sapply(df$month, CapStr)
+
+#remi code starts here
+
+depthlab <- paste(unique(df$depth),"m")[c(1,3,2,4)]
+names(depthlab) <- unique(df$depth)[c(1,3,2,4)]
+
+ggplot(df)+
+  geom_sf(data=mcp,fill='grey')+
+  geom_raster(aes(x,y,fill=value))+
+  geom_sf(data=mcp,fill="transparent")+
+  scale_fill_viridis(option="magma",
+                       direction=-1,
+                       guide=guide_colourbar(barwidth=20))+
+  facet_grid(depth~month,
+             labeller = labeller(depth=depthlab))+
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "bottom",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank())
+```

### 2020-01-30 src/comb-maxent.Rmd
ii

@@
 options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2")
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2", "tidyverse", "patchwork", "purrr")
 source("../src/ipak.R")
@@
 
+# sample probabily maps
+
+have moved sample files into a separate folder
+march,  june, sep, dec
+11.7737, 61.1128, 199.79. 628.026
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+s <- st_read("../data/env/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj)
+
+mcp <- st_read("../output/bio/minimum_convex_poly/mcp-wgs84.shp")
+
+filelist <- list.files("../data/plot2-probmaps/",pattern="avg_cloglog*",recursive = TRUE,include.dirs = TRUE)
+
+x <- stack(file.path("../data/plot2-probmaps/",filelist)) %>% 
+  mask(mcp)
+
+names(x) <- filelist
+
+df <- x %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  mutate(variable=gsub(".asc","",variable),
+         month=factor(months(parse_date_time(gsub("\\..*$", "",variable),"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         depth=as.numeric(gsub("^.*_", "",variable)))
+
+#remi code stops here
+
+df$month <- str_sub(df$variable, start = 16, end = 18)
+
+CapStr <- function(y) {
+  c <- strsplit(y, " ")[[1]]
+  paste(toupper(substring(c, 1,1)), substring(c, 2),
+      sep="", collapse=" ")
+}
+df$month <- sapply(df$month, CapStr)
+
+#remi code starts here
+
+depthlab <- paste(unique(df$depth),"m")[c(1,3,2,4)]
+names(depthlab) <- unique(df$depth)[c(1,3,2,4)]
+
+ggplot(df)+
+  geom_sf(data=mcp,fill='grey')+
+  geom_raster(aes(x,y,fill=value))+
+  geom_sf(data=mcp,fill="transparent")+
+  scale_fill_viridis(option="magma",
+                       direction=-1, limits = c(0,1), breaks = c(0, 0.25, 0.5, 0.75, 1),
+                       guide=guide_colourbar(barwidth=0.5, barheight = 15, title = NULL, draw.ulim = FALSE, draw.llim = FALSE))+
+  facet_grid(depth~month,
+             labeller = labeller(depth=depthlab))+
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "right",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
+        strip.text.x = element_text(face = "bold", size = 12),
+        strip.text.y = element_text(face = "bold", size = 12))
+
+ggsave("../output/maxent/paper_plots/test-overview.png", width = 210, height = 297, units = "mm", dpi = 600)
+```
+
 # variable maps
@@
 pick 3 diff years say 1998, 2006, 2013
-
-```{r individual maps}
-#aea <- raster("../output/env/aea.tif") #the correct CRS
-s <- shapefile("../data/env/landp/land-wgs84v2.shp") #made in arcgis
-crs(s) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
-#acrs <- aea@crs
-mcp <- shapefile("../output/bio/minimum_convex_poly/mcp-wgs84.shp")
-#ss <- raster::intersect(s, mcp)  #trim land file to mcp
-
-plotlist <- list()
-
-yrlst <- c("1998", "2006", "2013")
-mthlst <- c("03", "06", "09", "12")
-varlst <- c("temp", "salinity", "o2")
-
-pn <- 1
-for (y in 1:length(yrlst)){
-  yr <- yrlst[y]
-  fdr <- paste0("../data/env/ascii-original/ascii", yr, "/")
-  for (m in 1:length(mthlst)){
-    mt <- mthlst[m]
-    for (v in 1:length(varlst)){
-      var <- varlst[v]
-      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_11.7737.asc"))
-      lyrnm <- names(ras)
-      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
-      #ras <- projectRaster(ras, aea)
-      ras <- raster::mask(ras, mcp)
-
-      plotlist[[pn]] <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + latticeExtra::layer(sp.polygons(s, fill = "dark grey", col = "transparent") ) #changed ss to s
-      pn <- pn+1
-
-     # png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
-     # print(p)
-     # dev.off()
-    }
-  }
-}
-
-#chl is done seperately as it is surface
-
-yrlst <- c("1998", "2006", "2013")
-mthlst <- c("03", "06", "09", "12")
-varlst <- c("chl")
-
-for (y in 1:length(yrlst)){
-  yr <- yrlst[y]
-  fdr <- paste0("../data/env/ascii-original/ascii", yr, "/")
-  for (m in 1:length(mthlst)){
-    mt <- mthlst[m]
-    for (v in 1:length(varlst)){
-      var <- varlst[v]
-      ras <- raster(paste0(fdr, yr, "_", mt, "_", var, ".nc_0.50576.asc"))
-      lyrnm <- names(ras)
-      crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
-      ras <- raster::mask(ras, mcp)
-      plotlist[[pn]] <- levelplot(ras, margin=FALSE, colorkey=list(space='bottom', axis.line=list(col='black')), par.settings=list(axis.line=list(col='transparent')), scales=list(draw=FALSE), col.regions=colorRampPalette(blue2red(20))) + latticeExtra::layer(sp.polygons(s, fill = "dark grey", col = "transparent"))
-      pn <- pn+1
-
-   #   png(paste0("../output/maxent/paper_plots/", lyrnm, ".png"), height = 1024, width = 1024)
-    #  print(p)
-    #  dev.off()
-    }
-  }
-}
-```

### 2020-01-30 src/testing_file.Rmd
ii

@@
 
+#here i capitalise the months
 df$month <- str_sub(df$variable, start = 16, end = 18)
@@
   scale_fill_viridis(option="magma",
-                       direction=-1,
-                       guide=guide_colourbar(barwidth=20))+
+                       direction=-1, limits = c(0,1), breaks = c(0, 0.25, 0.5, 0.75, 1),
+                       guide=guide_colourbar(barwidth=0.5, barheight = 15, title = NULL, draw.ulim = FALSE, draw.llim = FALSE))+
   facet_grid(depth~month,
              labeller = labeller(depth=depthlab))+
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "right",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
+        strip.text.x = element_text(face = "bold", size = 12),
+        strip.text.y = element_text(face = "bold", size = 12))
+
+ggsave("../output/maxent/paper_plots/test-overview.png", width = 210, height = 297, units = "mm", dpi = 600)
+```
+
+#variable maps
+```{r}
+
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(patchwork)
+library(tidyverse)
+
+CapStr <- function(y) {
+  c <- strsplit(y, " ")[[1]]
+  paste(toupper(substring(c, 1,1)), substring(c, 2),
+      sep="", collapse=" ")
+}
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")
+
+s <- st_read("../data/plot4-variables/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj) %>% 
+  st_intersection(mcp)
+
+filelist <- list.files("../data/plot4-variables/ascii/ascii1998/",pattern="*.asc",recursive = TRUE)
+
+x <- stack(file.path("../data/plot4-variables/ascii/ascii1998/",filelist)) %>% 
+  mask(mcp) %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  separate(variable,c('year','month','variable','depth'),'_',remove=TRUE) %>% 
+  mutate(month=factor(months(parse_date_time(month,"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         year=as.numeric(gsub("X","",year)),
+         monthyear=factor(paste0(month,'\n',year),
+                          ordered = TRUE,
+                          levels=unique(paste0(month,'\n',year))),
+         variable=gsub(".nc","",variable)
+         )
+
+vars = c('chl','o2','salinity','temp')
+
+for(v in vars){
+  if(v=='chl'){
+  assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
+           geom_sf(data=s,fill='grey')+
+           geom_raster(aes(x,y,fill=value))+
+           geom_sf(data=mcp,fill="transparent")+
+           scale_fill_viridis(option="viridis", direction = -1 , 
+           #scale_fill_distiller(palette="Spectral", direction=-1)+
+           guide_colourbar(barwidth=0.5, title = "mmol\u207B\u00B3", draw.ulim = FALSE, draw.llim = FALSE))+
+
+           facet_wrap(~monthyear,nrow=4,strip.position = "right")+
+           theme_classic()+
+           scale_x_continuous(breaks = seq(-70,-50,10))+
+           labs(title=sapply(v, CapStr))+ #i added
+
+           theme(legend.position = "bottom",
+                 axis.title.x=element_blank(),
+                 axis.title.y=element_blank(),
+                 axis.text.x=element_text(angle=45,vjust=1,hjust=1),
+                 strip.background = element_blank(),
+                 strip.text = element_blank()))
+  }else if(v=='o2'){
+      assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
+           geom_sf(data=s,fill='grey')+
+           geom_raster(aes(x,y,fill=value))+
+           geom_sf(data=mcp,fill="transparent")+
+           scale_fill_viridis(option="viridis", direction = -1 , 
+           #scale_fill_distiller(palette="Spectral", direction=-1)+
+           guide_colourbar(barwidth=0.5, title = "mmol\u207B\u00B3", draw.ulim = FALSE, draw.llim = FALSE))+
+           labs(title=sapply(v, CapStr))+ #i added
+
+           facet_wrap(~monthyear,nrow=4,strip.position = "right")+
            theme_classic()+
@@
                  axis.title.y=element_blank(),
-        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank())
-```
+                 axis.text.x=element_text(angle=45,vjust=1,hjust=1),
+                 strip.background = element_blank(),
+                 strip.text = element_blank()))
+  }else if(v=='salinity'){
+      assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
+           geom_sf(data=s,fill='grey')+
+           geom_raster(aes(x,y,fill=value))+
+           geom_sf(data=mcp,fill="transparent")+
+           scale_fill_viridis(option="viridis", direction = -1 , 
+           #scale_fill_distiller(palette="Spectral", direction=-1)+
+           guide_colourbar(barwidth=0.5, title = "PSU", draw.ulim = FALSE, draw.llim = FALSE))+
+
+           facet_wrap(~monthyear,nrow=4,strip.position = "right")+
+           labs(title=sapply(v, CapStr))+ #i added
+           theme_classic()+
+           scale_x_continuous(breaks = seq(-70,-50,10))+
+           theme(legend.position = "bottom",
+                 axis.title.x=element_blank(),
+                 axis.title.y=element_blank(),
+                 axis.text.x=element_text(angle=45,vjust=1,hjust=1),
+                 strip.background = element_blank(),
+                 strip.text = element_blank()))
+  }
+  else if(v=='temp'){
+    assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
+             geom_sf(data=s,fill='grey')+
+             geom_raster(aes(x,y,fill=value))+
+             geom_sf(data=mcp,fill="transparent")+
+             scale_fill_viridis(option="viridis", direction = -1, 
+             #scale_fill_distiller(palette="Spectral", direction=-1)+
+             guide_colourbar(barwidth=0.5, title = "Kelvin", draw.ulim = FALSE, draw.llim = FALSE))+
+             facet_wrap(~monthyear,nrow=4,strip.position = "right")+
+             labs(title=sapply(v, CapStr))+ #i added
+             theme_classic()+
+             scale_x_continuous(breaks = seq(-70,-50,10))+
+             theme(legend.position = "bottom",
+                   axis.title.x=element_blank(),
+                   axis.title.y=element_blank(),
+                   axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
+                   strip.text.y = element_text(face = "bold")))
+  }
+}
+
+chl|o2|salinity|temp
+
+ggsave("../output/maxent/paper_plots/variable_samples.png", height = 210, width = 297, units = "mm", dpi = 600)
+```
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(patchwork)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")
+
+s <- st_read("../data/plot4-variables/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj) %>% 
+  st_intersection(mcp)
+
+filelist <- list.files("../data/plot4-variables/ascii/ascii1998/",pattern="*.asc",recursive = TRUE)
+
+x <- stack(file.path("../data/plot4-variables/ascii/ascii1998/",filelist)) %>% 
+  mask(mcp) %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  separate(variable,c('year','month','variable','depth'),'_',remove=TRUE) %>% 
+  mutate(month=factor(months(parse_date_time(month,"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         year=as.numeric(gsub("X","",year)),
+         monthyear=factor(paste0(month,'\n',year),
+                          ordered = TRUE,
+                          levels=unique(paste0(month,'\n',year))),
+         variable=gsub(".nc","",variable)
+         )
+vars = c('chl','o2','salinity','temp')
+for(v in vars){
+  assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
+           geom_sf(data=s,fill='grey')+
+           geom_raster(aes(x,y,fill=value))+
+           geom_sf(data=mcp,fill="transparent")+
+scale_fill_viridis(option="viridis", direction = -1, 
+             #scale_fill_distiller(palette="Spectral", direction=-1)+
+             guide_colourbar(barwidth=0.5, title = "Kelvin", draw.ulim = FALSE, draw.llim = FALSE))+
+           facet_wrap(~monthyear,nrow=4,strip.position = "right")+
+           labs(title=v)+
+           theme_classic()+
+           scale_x_continuous(breaks = seq(-70,-50,10))+
+           theme(legend.position = "bottom",
+                 axis.title.x=element_blank(),
+                 axis.title.y=element_blank(),
+                 axis.text.x=element_text(angle=45,vjust=1,hjust=1),
+                 strip.background = element_blank(),
+                 strip.text = element_blank()))
+  if(v=='temp'){
+    assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
+             geom_sf(data=s,fill='grey')+
+             geom_raster(aes(x,y,fill=value))+
+             geom_sf(data=mcp,fill="transparent")+
+scale_fill_viridis(option="viridis", direction = -1, 
+             #scale_fill_distiller(palette="Spectral", direction=-1)+
+             guide_colourbar(barwidth=0.5, title = "Kelvin", draw.ulim = FALSE, draw.llim = FALSE))+
+             facet_wrap(~monthyear,nrow=4,strip.position = "right")+
+             labs(title=v)+
+             theme_classic()+
+             scale_x_continuous(breaks = seq(-70,-50,10))+
+             theme(legend.position = "bottom",
+                   axis.title.x=element_blank(),
+                   axis.title.y=element_blank(),
+                   axis.text.x=element_text(angle=45,vjust=1,hjust=1)))
+  }
+}
+
+chl|o2|salinity|temp
+```
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")
+
+s <- st_read("../data/plot3-2d/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj) %>% 
+  st_intersection(mcp)
+
+filelist <- list.files("../data/plot3-2d/hotspot-mapping/",include.dirs = TRUE)
+
+x <- stack(file.path("../data/plot3-2d/hotspot-mapping/",filelist)) %>% 
+  mask(mcp) %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  mutate(month=factor(months(parse_date_time(gsub("_.*$", "",variable),"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         value=ifelse(value==1,">0.5","<0.5"))
+
+df <- x %>% 
+  group_by(month) %>% 
+  summarize(total=round(sum(value==">0.5")/n(),2))
+
+ggplot(x)+
+  geom_sf(data=s,fill='grey',colour='transparent')+
+  geom_raster(aes(x,y,fill=value))+
+  scale_fill_manual(values=c('transparent','red'))+
+  geom_sf(data=mcp,fill="transparent")+
+  geom_text(data=df,x=-50,y=67,label=df$total)+
+  facet_wrap(~month,nrow=2)+#,
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "none",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1))
+```

### 2020-01-31 src/comb-maxent.Rmd
messing with Remi's code... one plot still to fiddle with

@@
 
+#2d maps
+
+this is currently set to 0.5 cut-off... not SLR and MJ would like this changed
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")
+
+s <- st_read("../data/plot3-2d/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj) %>% 
+  st_intersection(mcp)
+
+filelist <- list.files("../data/plot3-2d/hotspot-mapping/",include.dirs = TRUE)
+
+x <- stack(file.path("../data/plot3-2d/hotspot-mapping/",filelist)) %>% 
+  mask(mcp) %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  mutate(month=factor(months(parse_date_time(gsub("_.*$", "",variable),"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         value=ifelse(value==1,">0.5","<0.5"))
+
+df <- x %>% 
+  group_by(month) %>% 
+  summarize(total=round(sum(value==">0.5")/n(),2))
+
+ggplot(x)+
+  geom_sf(data=s,fill='grey',colour='transparent')+
+  geom_raster(aes(x,y,fill=value))+
+  scale_fill_manual(values=c('transparent','red'))+
+  geom_sf(data=mcp,fill="transparent")+
+  geom_text(data=df,x=-50,y=67,label=df$total)+
+  facet_wrap(~month,nrow=2)+#,
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "none",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1),
+        strip.background = element_blank(),
+        strip.text.x = element_text(face = "bold"))
+
+ggsave("../output/maxent/paper_plots/two-d.png", height = 210, width = 297, units = "mm", dpi = 600)
+```
+
 # variable maps

### 2020-01-31 src/testing_file.Rmd
messing with Remi's code... one plot still to fiddle with

@@
            geom_sf(data=mcp,fill="transparent")+
-           scale_fill_viridis(option="viridis", direction = -1 , 
+            scale_fill_viridis(option="viridis", direction = -1, 
              #scale_fill_distiller(palette="Spectral", direction=-1)+
            guide_colourbar(barwidth=0.5, title = "mmol\u207B\u00B3", draw.ulim = FALSE, draw.llim = FALSE))+
-
+           labs(title= "Chl \u03b1")+ #i added
            facet_wrap(~monthyear,nrow=4,strip.position = "right")+
@@
            scale_x_continuous(breaks = seq(-70,-50,10))+
-           labs(title=sapply(v, CapStr))+ #i added
-
            theme(legend.position = "bottom",
@@
            guide_colourbar(barwidth=0.5, title = "mmol\u207B\u00B3", draw.ulim = FALSE, draw.llim = FALSE))+
-           labs(title=sapply(v, CapStr))+ #i added
+           labs(title=paste0("Dissolved ", sapply(v, CapStr)))+ #i added
 
@@
                  strip.text = element_blank()))
-  }
-  else if(v=='temp'){
+  } else if(v=='temp'){
     assign(v,ggplot(x %>% filter(variable==v,month!='September'))+
@@
 
-ggsave("../output/maxent/paper_plots/variable_samples.png", height = 210, width = 297, units = "mm", dpi = 600)
+#ggsave("../output/maxent/paper_plots/variable_samples.png", height = 210, width = 297, units = "mm", dpi = 600)
 ```
@@
 
+#2-d map
+
 ```{r}
@@
         axis.title.y=element_blank(),
-        axis.text.x=element_text(angle=45,vjust=1,hjust=1))
-```
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1),
+        strip.background = element_blank(),
+        strip.text.x = element_text(face = "bold"))
+
+ggsave("../output/maxent/paper_plots/two-d.png", height = 210, width = 297, units = "mm", dpi = 600)
+
+```

### 2020-02-03 src/testing_file.Rmd
dong barbaric things to remi's code

@@
 
-```{r}
-streams <- raster::shapefile("../data/env/landp/land-wgs84v2.shp")
-ras <-
-
-ggplot() + layer_spatial(data = streams, fill = "dark grey", colour = "transparent") + layer_spatial() + theme_void()
-
-```
-
 #2d maps for comb-maxent
@@
 ```
+
+now create a table to see how many cells there are >0.5 in each month
+
+```{r count cells >0.5}
+
+binaryprob_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = ".*_bin_0.5_2d.asc", full.names = TRUE)
+sum2d <- data.frame(matrix(ncol = 2, nrow = 10))
+sum2d <- setNames(sum2d, c("month", "count"))
+
+for (r in 1:length(binaryprob_list)){
+  ras <- raster(binaryprob_list[r])
+  rnm <- names(ras)
+  sum2d$month[r] <- gsub("_bin_0.5_2d", "", x = rnm)
+  sum2d$count[r] <- cellStats(ras, sum, asSample = TRUE)
+}
+
+#order table from march to december
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+sum2d <- sum2d[match(mthord, sum2d$month), ]
+
+#get proportion
+tcs <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE) #file showing how many cells per depth
+#tcs <- tcs[-c(1:3), ] #because the top three rows are deeper than the capelin obs...
+totc <- sum(tcs$no_cells)
+
+for (r in 1:nrow(sum2d)){
+  sum2d$proportion[r] <- (sum2d$count[r]/totc)
+}
+
+write.csv(sum2d, "../output/maxent/paper_plots/cell_count_0.5.csv", row.names = FALSE)
+
+```
+
+#2-d map edited for value as remi cunted of the cells in the layer i sent and not with all the layers.
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")
+
+s <- st_read("../data/plot3-2d/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj) %>% 
+  st_intersection(mcp)
+
+filelist <- list.files("../data/plot3-2d/hotspot-mapping/",include.dirs = TRUE)
+
+x <- stack(file.path("../data/plot3-2d/hotspot-mapping/",filelist)) %>% 
+  mask(mcp) %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  mutate(month=factor(months(parse_date_time(gsub("_.*$", "",variable),"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         value=ifelse(value==1,">0.5","<0.5"))
+
+df <- x %>% 
+  group_by(month) %>% 
+  summarize(total=round(sum(value==">0.5")/n(),2))
+
+#quickest way might be just to inject my csv file with the correct values here....
+df2 <- read.csv("../output/maxent/paper_plots/cell_count_0.5.csv", header = TRUE)
+df$total <- df2$proportion
+df$total <- format(round(df$total, digits = 2))
+#colnames(df) <- c("month", "cellc", "total")
+
+ggplot(x)+
+  geom_sf(data=s,fill='grey',colour='transparent')+
+  geom_raster(aes(x,y,fill=value))+
+  scale_fill_manual(values=c('transparent','red'))+
+  geom_sf(data=mcp,fill="transparent")+
+  geom_text(data=df,x=-50,y=67,label=df$total)+
+  facet_wrap(~month,nrow=2)+#,
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "none",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1),
+        strip.background = element_blank(),
+        strip.text.x = element_text(face = "bold"))
+
+ggsave("../output/maxent/paper_plots/two-d.png", height = 210, width = 297, units = "mm", dpi = 600)
+
+```

### 2020-02-10 src/comb-maxent.Rmd
updating binary predictions to max kap

@@
 
+#binned probability cell count for the whole 3d space per month
+
+```{r}
+library(stringr)
+
+#calculate no of cells in the 3D space
+prob <- read.csv("../output/maxent/paper_plots/prob_by_depth/apr_dis_prob_cells.csv", header = TRUE) #just a randomly chosen dataset
+t2 <- subset(prob, Probability == "0 - 0.09") #could have chosen any probability
+tc <- sum(t2$total_cells) #173623
+
+dislst <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "dis_prob_cells.csv", full.names = TRUE)
+totlst <- list()
+for (m in 1:length(dislst)){
+  pr <- read.csv(dislst[m])
+  mth <- str_sub(dislst[m], start = 44, end = 46)
+  pr2 <- aggregate(pr$Cells, by=list(Category = pr$Probability), FUN = sum)
+  pr2$percent <- (pr2$x/tc)*100
+  pr2$month <- mth
+  write.csv(pr2, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_cellprob_totals.csv"), row.names = FALSE)
+  totlst[[m]] <- pr2
+}
+
+#then i just made a table manually in excel
+```
+
 #plotting
@@
 
-first bin the probabilities for cells inside the mcp
-
-```{r bin probabilities}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-
-for (m in 1:length(mthord)){
-  mth <- mthord[m]
+histogram of probability by depth - by proportion
+
+#max kappa thresholds
+
+```{r}
+binprob <- function(x){
+  ifelse(x > 0 & x >= mk, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+alst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
+
+for (f in 1:length(alst)){
+  mth <- str_sub(alst[f], start = 35, end = 37)
+  af <- read.csv(alst[f], header = TRUE)
+  af <- tail(af, n = 1)
+  mk <- af$test_maxkap
+
   clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
-
-  binprob <- function(x){
-    ifelse(x > 0 & x < 0.1, 0.1,
-    ifelse(x <= 0, 0, NA))
-  }
-
   for (cum in 1:length(clogasc_list)){
@@
     probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".png")), overwrite = TRUE)
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
     dev.off() # stops automatic saving of the plot to a png
   }
-
-  binprob <- function(x){
-    ifelse(x >= 0.1 & x < 0.2, 0.2,
-    ifelse(x <= 0, 0, NA))
-  }
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-
-  binprob <- function(x){
-    ifelse(x >= 0.2 & x < 0.3, 0.3,
-    ifelse(x <= 0, 0, NA))
-  }
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-
-  binprob <- function(x){
-    ifelse(x >= 0.3 & x < 0.4, 0.4,
-    ifelse(x <= 0, 0, NA))
-  }
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-
-  binprob <- function(x){
-    ifelse(x >= 0.4 & x < 0.5, 0.5,
-    ifelse(x <= 0, 0, NA))
-  }
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-
-  binprob <- function(x){
-    ifelse(x >= 0.5 & x < 0.6, 0.6,
-    ifelse(x <= 0, 0, NA))
-  }
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    #cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-
-  binprob <- function(x){
-    ifelse(x >= 0.6 & x < 0.7, 0.7,
-    ifelse(x <= 0, 0, NA))
-  }
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    #cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-
-  binprob <- function(x){
-    ifelse(x >= 0.7 & x < 0.8, 0.8,
-    ifelse(x <= 0, 0, NA))
-  }
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    #cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_8_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_8_avgprob_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-
-  binprob <- function(x){
-    ifelse(x >= 0.8 & x < 0.9, 0.9,
-    ifelse(x <= 0, 0, NA))
-  }
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    #cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_9_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_9_avgprob_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-
-  binprob <- function(x){
-    ifelse(x >= 0.9 & x <= 1, 1,
-    ifelse(x <= 0, 0, NA))
-  }
-
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-   # cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_10_avgprob_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_10_avgprob_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-}
+}
+
 ```
@@
     # bin <- 1
-  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, "_avg.*asc"), full.names = TRUE)
+  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_maxkap.*asc"), full.names = TRUE)
 
@@
   }
-  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_probbin_", bin, "_cellcount_bydepth_avg.csv"), row.names = FALSE)
+  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", bin, "_cellcount_bydepth_avg.csv"), row.names = FALSE)
   }
@@
 ```
-
-histogram of probability by depth - by proportion
-
 based on Remi's code

### 2020-02-10 src/testing_file.Rmd
updating binary predictions to max kap

@@
 library(tidyverse)
+library(viridis) #added by me
 
@@
            geom_sf(data=mcp,fill="transparent")+
-            scale_fill_viridis(option="viridis", direction = -1, 
+            scale_fill_viridis(option="viridis", direction = -1, limits = c(0.03, 4.5), 
              #scale_fill_distiller(palette="Spectral", direction=-1)+
@@
            geom_sf(data=mcp,fill="transparent")+
-           scale_fill_viridis(option="viridis", direction = -1 , 
+           scale_fill_viridis(option="viridis", direction = -1, limits = c(207, 418) , 
            #scale_fill_distiller(palette="Spectral", direction=-1)+
@@
            geom_sf(data=mcp,fill="transparent")+
-           scale_fill_viridis(option="viridis", direction = -1 , 
+           scale_fill_viridis(option="viridis", direction = -1, limits = c(22, 36), 
            #scale_fill_distiller(palette="Spectral", direction=-1)+
@@
              geom_sf(data=mcp,fill="transparent")+  
-             scale_fill_viridis(option="viridis", direction = -1, 
+             scale_fill_manual(col=blue2red(20), 
+             #scale_fill_viridis(option="viridis", direction = -1, limits = c(270, 296),
              #scale_fill_distiller(palette="Spectral", direction=-1)+
@@
 
+#2-d map edited for value as remi cunted of the cells in the layer i sent and not with all the layers.
+
 now create a table to see how many cells there are >0.5 in each month
@@
 ```
-
-#2-d map edited for value as remi cunted of the cells in the layer i sent and not with all the layers.
 
@@
 ```
+
+#binned probability cell count for the whole 3d space per month
+
+```{r}
+library(stringr)
+
+#calculate no of cells in the 3D space
+prob <- read.csv("../output/maxent/paper_plots/prob_by_depth/apr_dis_prob_cells.csv", header = TRUE) #just a randomly chosen dataset
+t2 <- subset(prob, Probability == "0 - 0.09") #could have chosen any probability
+tc <- sum(t2$total_cells) #173623
+
+dislst <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "dis_prob_cells.csv", full.names = TRUE)
+totlst <- list()
+for (m in 1:length(dislst)){
+  pr <- read.csv(dislst[m])
+  mth <- str_sub(dislst[m], start = 44, end = 46)
+  pr2 <- aggregate(pr$Cells, by=list(Category = pr$Probability), FUN = sum)
+  pr2$percent <- (pr2$x/tc)*100
+  pr2$month <- mth
+  write.csv(pr2, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_cellprob_totals.csv"), row.names = FALSE)
+  totlst[[m]] <- pr2
+}
+
+#then i just made a table manually in excel
+```
+
+#max kappa thresholds
+
+```{r}
+binprob <- function(x){
+  ifelse(x > 0 & x >= mk, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+alst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
+
+for (f in 1:length(alst)){
+  mth <- str_sub(alst[f], start = 35, end = 37)
+  af <- read.csv(alst[f], header = TRUE)
+  af <- tail(af, n = 1)
+  mk <- af$test_maxkap
+
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+}
+
+```
+
+get data to plot probability by depth... 
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+binprob <- function(x){
+  ifelse(x >0,  1,
+  ifelse(x == -9999, 0, NA))
+}
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_maxkap.*asc"), full.names = TRUE)
+  binaryavgstack <- stack(ascdepth)
+  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+  stackstathigh[ ,1] <- c(1:47)
+  colnames(stackstathigh) <- c("asclayer", "prob_bin")
+
+  for (lyr in 1:nlayers(binaryavgstack)){
+    ascname <-  str_sub(ascdepth[lyr], 67, -5)
+    dlay <- binaryavgstack[[lyr]]
+    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
+    lyrsum <- cellStats(dlay, sum)
+    stackstathigh$prob_bin[lyr] <- lyrsum
+    stackstathigh$asclayer[lyr] <- ascname
+
+  }
+  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_cellcount_bydepth_avg.csv"), row.names = FALSE)
+}
+
+```
+
+from here sam
+
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_probbin_.*.csv"), full.names = TRUE)
+  a <- read.csv(probblst[1])
+  b <- read.csv(probblst[2])
+  c <- read.csv(probblst[3])
+  d <- read.csv(probblst[4])
+  e <- read.csv(probblst[5])
+  f <- read.csv(probblst[6])
+  g <- read.csv(probblst[7])
+  h <- read.csv(probblst[8])
+  i <- read.csv(probblst[9])
+  j <- read.csv(probblst[10])
+
+  stackstat <- cbind(a, b[ ,2], c[ ,2], d[ ,2], e[ , 2], f[ ,2], g[ ,2], h[ ,2], i[ ,2], j[ ,2])
+  colnames(stackstat) <- c("depth", "0 - 0.09", "0.1 - 0.19", "0.2 - 0.29", "0.3 - 0.39", "0.4 - 0.49", "0.5 - 0.59", "0.6 - 0.69", "0.7 - 0.79", "0.8 - 0.89", ">0.9")
+  stackstat <- melt(stackstat, id = "depth")
+  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]
+
+  colnames(stackstat) <- c("Depth", "Probability", "Cells")
+
+  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
+}
+
+```

### 2020-02-11 src/comb-maxent.Rmd
adjusted kappa done. range check started

@@
 filenames <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)
+filenames <- filenames[ !grepl("maxkap", filenames) ] #remove all the maxkap files
 
@@
                       levels=months(parse_date_time(1:12,"%m"))))
-
 #create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
@@
 dlst <- sort(dlst)
-
 #make the plot
@@
   coord_flip()+
-  facet_wrap(~Month,ncol=5)+
+  facet_wrap(~Month,ncol=2)+
   labs(y = "% Ocean at Depth", x = "Depth (meters)")+
@@
 
-ggsave("../output/maxent/paper_plots/test.png", dpi = 600)
+ggsave("../output/maxent/paper_plots/prob_by_depth.png", width = 210, height = 297, units = "mm", dpi = 600)
 ```
@@
 
-#2d maps
-
-this is currently set to 0.5 cut-off... not SLR and MJ would like this changed
-
-```{r}
+#2d maps - max kappa thresholds
+
+this creates binary maps for each layer based on the maxkappa threshold
+```{r}
+binprob <- function(x){
+  ifelse(x > 0 & x >= mk, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+#get the maxkap value
+alst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
+
+for (f in 1:length(alst)){
+  mth <- str_sub(alst[f], start = 35, end = 37)
+  af <- read.csv(alst[f], header = TRUE)
+  af <- tail(af, n = 1)
+  mk <- af$test_maxkap
+
+#create binary maps per depth layer
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/hotspot_mapping/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/hotspot_mapping/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+}
+
+```
+
+then create an summary map....
+```{r >maxkap prob monthly 2D representation}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mthna <- mthord[m]
+  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_maxkap_.*.asc$"), full.names = TRUE)
+  binstack <- stack(binaryprob_list)
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.png")))
+  dev.off()
+}
+
+```
+
+now create summary maps where values are binary - not counts of cells
+```{r}
+binprob <- function(x){
+  ifelse(x > 0, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+  clogasc_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = "*_maxkap_2d.asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- str_sub(clogasc_list[cum], start = 46, end = -8)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+```
+
+now create a table to see how many cells there are >maxkap in each month (for editing plots)
+
+```{r count cells >maxkap}
+
+binaryprob_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = ".*_maxkap_2d.asc", full.names = TRUE)
+sum2d <- data.frame(matrix(ncol = 2, nrow = 10))
+sum2d <- setNames(sum2d, c("month", "count"))
+
+for (r in 1:length(binaryprob_list)){
+  ras <- raster(binaryprob_list[r])
+  rnm <- names(ras)
+  sum2d$month[r] <- gsub("_maxkap_2d", "", x = rnm)
+  sum2d$count[r] <- cellStats(ras, sum, asSample = TRUE)
+}
+
+#order table from march to december
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+sum2d <- sum2d[match(mthord, sum2d$month), ]
+
+#get proportion
+tcs <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE) #file showing how many cells per depth
+#tcs <- tcs[-c(1:3), ] #because the top three rows are deeper than the capelin obs...
+totc <- sum(tcs$no_cells)
+
+for (r in 1:nrow(sum2d)){
+  sum2d$proportion[r] <- (sum2d$count[r]/totc)*100 #this actually shows percentage
+}
+
+write.csv(sum2d, "../output/maxent/paper_plots/cell_count_maxkap.csv", row.names = FALSE)
+
+```
+
+```{r plot maxkap 2d}
 library(sf)
@@
 
-filelist <- list.files("../data/plot3-2d/hotspot-mapping/",include.dirs = TRUE)
-
-x <- stack(file.path("../data/plot3-2d/hotspot-mapping/",filelist)) %>% 
+filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", include.dirs = TRUE)
+
+x <- stack(file.path("../output/maxent/paper_plots/hotspot_mapping/",filelist)) %>% 
   mask(mcp) %>% 
@@
 
+#quickest way might be just to inject my csv file with the correct values here....
+df2 <- read.csv("../output/maxent/paper_plots/cell_count_maxkap.csv", header = TRUE)
+df$total <- df2$proportion
+df$total <- format(round(df$total, digits = 2))
+#colnames(df) <- c("month", "cellc", "total")
+
 ggplot(x)+
+  geom_raster(aes(x,y,fill=value))+
+  scale_fill_manual(values=c('light yellow','black'))+
   geom_sf(data=s,fill='grey',colour='transparent')+
-  geom_raster(aes(x,y,fill=value))+
-  scale_fill_manual(values=c('transparent','red'))+
   geom_sf(data=mcp,fill="transparent")+
@@
 
-ggsave("../output/maxent/paper_plots/two-d.png", height = 210, width = 297, units = "mm", dpi = 600)
+ggsave("../output/maxent/paper_plots/maxkap_2d.png", height = 210, width = 297, units = "mm", dpi = 600)
+
+```
+
+get data to plot maxkap binary by depth... 
+```{r maxkap by depth}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+binprob <- function(x){
+  ifelse(x >0,  1,
+  ifelse(x == -9999, 0, NA))
+}
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_maxkap.*asc"), full.names = TRUE)
+  binaryavgstack <- stack(ascdepth)
+  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+  stackstathigh[ ,1] <- c(1:47)
+  colnames(stackstathigh) <- c("asclayer", "prob_bin")
+
+  for (lyr in 1:nlayers(binaryavgstack)){
+    ascname <-  str_sub(ascdepth[lyr], 67, -5)
+    dlay <- binaryavgstack[[lyr]]
+    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
+    lyrsum <- cellStats(dlay, sum)
+    stackstathigh$prob_bin[lyr] <- lyrsum
+    stackstathigh$asclayer[lyr] <- ascname
+
+  }
+  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_cellcount_bydepth_avg.csv"), row.names = FALSE)
+}
+
+```
+
+think about how to get proportions...
+
+```{r percentage cells above and below maxkap}
+mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+cumst1 <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE)
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
+  for (r in 1:nrow(a)){
+    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
+    # d1 <- str_sub(a$Depth[r], end = 4)
+    d1 <- a$asclayer[r]
+	  for (ro in 1:nrow(cumst1)){
+      #d2 <- str_sub(cumst1$depth[ro], end = 4)
+	    d2 <- cumst1$depth[ro]
+      if (d1 == d2){
+          a$total_cells[r] <- cumst1$no_cells[ro]
+      }
+	  }
+    a$p_cells[r] <- (a$prob_bin[r]/a$total_cells[r])*100
+    a$Probability[r] <- 1
+  }
+    #a$asclayer <- as.numeric(a$asclayer)
+
+  a2 <- as.data.frame(a$asclayer)
+  a2$total_cells <- a$total_cells
+  a2$prob_bin <- a$total_cells - a$prob_bin
+  a2$p_cells <- (a2$prob_bin/a2$total_cells)*100
+  a2$Probability <- 0
+  colnames(a2)[1] <- "asclayer"
+
+  b <- rbind(a, a2)
+  colnames(b)[1] <- "Depth"
+
+  write.csv(b, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_dis_prob_cells.csv"), row.names = FALSE)
+}
+
+```
+
+based on Remi's code - maxkappa derrived habitat by depth
+```{r maxkappa habitat by depth}
+require(data.table)
+require(purrr)
+require(lubridate)
+require(tidyverse)
+require(viridis)
+
+filenames <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "_maxkap_dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)
+
+all <- map_dfr(filenames, function(x) fread(x) %>% 
+                       mutate(Month=substr(basename(x),1,3))) %>% 
+  mutate(Depth = factor(round(Depth,1)),
+         Depth = factor(Depth,ordered = TRUE, levels = rev(levels(Depth))),
+         Month=factor(months(parse_date_time(Month,"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))))
+
+#create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
+dlst <- unique(all$Depth)
+dlst <- as.character(dlst)
+dlst <- as.numeric(dlst)
+dlst <- sort(dlst)
+
+#make probability a factor - sam added this
+all$Probability <- as.factor(all$Probability)
+
+#make the plot
+  ggplot(all) +
+  geom_col(aes(Depth, p_cells, fill = Probability),width=1) +
+  coord_flip()+
+  facet_wrap(~Month,ncol=2)+
+  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
+  scale_fill_manual(aesthetics = "fill", values=c('light yellow','black'))+
+  theme_classic()+
+  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
+  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47]))
+
+ggsave("../output/maxent/paper_plots/maxkap_prob_by_depth.png", width = 210, height = 297, units = "mm", dpi = 600)
 ```

### 2020-02-11 src/testing_file.Rmd
adjusted kappa done. range check started

@@
 
+this creates binary maps based on th#max kappa thresholds
+
+this creates binary maps based on the maxkappa threshold
 ```{r}
@@
 
+#get the maxkap value
 alst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
@@
 
+#create binary maps per depth layer
   clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
@@
     probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
+    writeRaster(probbin, paste0("../output/maxent/hotspot_mapping/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/hotspot_mapping/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
     dev.off() # stops automatic saving of the plot to a png
@@
 
-get data to plot probability by depth... 
+then create an summary map....
+```{r >maxkap prob monthly 2D representation}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mthna <- mthord[m]
+  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_maxkap_.*.asc$"), full.names = TRUE)
+  binstack <- stack(binaryprob_list)
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.png")))
+  dev.off()
+}
+
+```
+
+now create summary maps where values are binary - not counts of cells
+```{r}
+binprob <- function(x){
+  ifelse(x > 0, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+  clogasc_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = "*_maxkap_2d.asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- str_sub(clogasc_list[cum], start = 46, end = -8)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+```
+
+now create a table to see how many cells there are >maxkap in each month (for editing plots)
+
+```{r count cells >maxkap}
+
+binaryprob_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = ".*_maxkap_2d.asc", full.names = TRUE)
+sum2d <- data.frame(matrix(ncol = 2, nrow = 10))
+sum2d <- setNames(sum2d, c("month", "count"))
+
+for (r in 1:length(binaryprob_list)){
+  ras <- raster(binaryprob_list[r])
+  rnm <- names(ras)
+  sum2d$month[r] <- gsub("_maxkap_2d", "", x = rnm)
+  sum2d$count[r] <- cellStats(ras, sum, asSample = TRUE)
+}
+
+#order table from march to december
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+sum2d <- sum2d[match(mthord, sum2d$month), ]
+
+#get proportion
+tcs <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE) #file showing how many cells per depth
+#tcs <- tcs[-c(1:3), ] #because the top three rows are deeper than the capelin obs...
+totc <- sum(tcs$no_cells)
+
+for (r in 1:nrow(sum2d)){
+  sum2d$proportion[r] <- (sum2d$count[r]/totc)
+}
+
+write.csv(sum2d, "../output/maxent/paper_plots/cell_count_maxkap.csv", row.names = FALSE)
+
+```
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")
+
+s <- st_read("../data/plot3-2d/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj) %>% 
+  st_intersection(mcp)
+
+filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", include.dirs = TRUE)
+
+x <- stack(file.path("../output/maxent/paper_plots/hotspot_mapping/",filelist)) %>% 
+  mask(mcp) %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  mutate(month=factor(months(parse_date_time(gsub("_.*$", "",variable),"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         value=ifelse(value==1,">0.5","<0.5"))
+
+df <- x %>% 
+  group_by(month) %>% 
+  summarize(total=round(sum(value==">0.5")/n(),2))
+
+#quickest way might be just to inject my csv file with the correct values here....
+df2 <- read.csv("../output/maxent/paper_plots/cell_count_maxkap.csv", header = TRUE)
+df$total <- df2$proportion
+df$total <- format(round(df$total, digits = 2))
+#colnames(df) <- c("month", "cellc", "total")
+
+ggplot(x)+
+  geom_sf(data=s,fill='grey',colour='transparent')+
+  geom_raster(aes(x,y,fill=value))+
+  scale_fill_manual(values=c('transparent','red'))+
+  geom_sf(data=mcp,fill="transparent")+
+  geom_text(data=df,x=-50,y=67,label=df$total)+
+  facet_wrap(~month,nrow=2)+#,
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "none",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1),
+        strip.background = element_blank(),
+        strip.text.x = element_text(face = "bold"))
+
+ggsave("../output/maxent/paper_plots/maxkap_2d.png", height = 210, width = 297, units = "mm", dpi = 600)
+
+```
+
+e maxkappa threshold
+```{r}
+binprob <- function(x){
+  ifelse(x > 0 & x >= mk, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+#get the maxkap value
+alst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
+
+for (f in 1:length(alst)){
+  mth <- str_sub(alst[f], start = 35, end = 37)
+  af <- read.csv(alst[f], header = TRUE)
+  af <- tail(af, n = 1)
+  mk <- af$test_maxkap
+
+#create binary maps per depth layer
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/hotspot_mapping/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/hotspot_mapping/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+}
+
+```
+
+then create an summary map....
+```{r >maxkap prob monthly 2D representation}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mthna <- mthord[m]
+  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_maxkap_.*.asc$"), full.names = TRUE)
+  binstack <- stack(binaryprob_list)
+  stacksum <- sum(binstack, na.rm = TRUE)
+  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.asc"), overwrite=TRUE)
+  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.png")))
+  dev.off()
+}
+
+```
+
+now create summary maps where values are binary - not counts of cells
+```{r}
+binprob <- function(x){
+  ifelse(x > 0, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+  clogasc_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = "*_maxkap_2d.asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- str_sub(clogasc_list[cum], start = 46, end = -8)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+```
+
+now create a table to see how many cells there are >maxkap in each month (for editing plots)
+
+```{r count cells >maxkap}
+
+binaryprob_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = ".*_maxkap_2d.asc", full.names = TRUE)
+sum2d <- data.frame(matrix(ncol = 2, nrow = 10))
+sum2d <- setNames(sum2d, c("month", "count"))
+
+for (r in 1:length(binaryprob_list)){
+  ras <- raster(binaryprob_list[r])
+  rnm <- names(ras)
+  sum2d$month[r] <- gsub("_maxkap_2d", "", x = rnm)
+  sum2d$count[r] <- cellStats(ras, sum, asSample = TRUE)
+}
+
+#order table from march to december
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+sum2d <- sum2d[match(mthord, sum2d$month), ]
+
+#get proportion
+tcs <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE) #file showing how many cells per depth
+#tcs <- tcs[-c(1:3), ] #because the top three rows are deeper than the capelin obs...
+totc <- sum(tcs$no_cells)
+
+for (r in 1:nrow(sum2d)){
+  sum2d$proportion[r] <- (sum2d$count[r]/totc)
+}
+
+write.csv(sum2d, "../output/maxent/paper_plots/cell_count_maxkap.csv", row.names = FALSE)
+
+```
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+mcp <- st_read("../data/plot4-variables/minimum_convex_poly/mcp-wgs84.shp")
+
+s <- st_read("../data/plot3-2d/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj) %>% 
+  st_intersection(mcp)
+
+filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", include.dirs = TRUE)
+
+x <- stack(file.path("../output/maxent/paper_plots/hotspot_mapping/",filelist)) %>% 
+  mask(mcp) %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  mutate(month=factor(months(parse_date_time(gsub("_.*$", "",variable),"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         value=ifelse(value==1,">0.5","<0.5"))
+
+df <- x %>% 
+  group_by(month) %>% 
+  summarize(total=round(sum(value==">0.5")/n(),2))
+
+#quickest way might be just to inject my csv file with the correct values here....
+df2 <- read.csv("../output/maxent/paper_plots/cell_count_maxkap.csv", header = TRUE)
+df$total <- df2$proportion
+df$total <- format(round(df$total, digits = 2))
+#colnames(df) <- c("month", "cellc", "total")
+
+ggplot(x)+
+  geom_raster(aes(x,y,fill=value))+
+  scale_fill_manual(values=c('light yellow','black'))+
+  geom_sf(data=s,fill='grey',colour='transparent')+
+  geom_sf(data=mcp,fill="transparent")+
+  geom_text(data=df,x=-50,y=67,label=df$total)+
+  facet_wrap(~month,nrow=2)+#,
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "none",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1),
+        strip.background = element_blank(),
+        strip.text.x = element_text(face = "bold"))
+
+ggsave("../output/maxent/paper_plots/maxkap_2d.png", height = 210, width = 297, units = "mm", dpi = 600)
+
+```
+
+----------------------
+
+get data to plot maxkap binary by depth... 
 ```{r}
@@
 
-from here sam
-
-```{r}
+think about how to get proportions...
+
+```{r percentage cells above and below maxkap}
 mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+cumst1 <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE)
 
@@
   mth <- mthord[m]
-  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_probbin_.*.csv"), full.names = TRUE)
-  a <- read.csv(probblst[1])
-  b <- read.csv(probblst[2])
-  c <- read.csv(probblst[3])
-  d <- read.csv(probblst[4])
-  e <- read.csv(probblst[5])
-  f <- read.csv(probblst[6])
-  g <- read.csv(probblst[7])
-  h <- read.csv(probblst[8])
-  i <- read.csv(probblst[9])
-  j <- read.csv(probblst[10])
-
-  stackstat <- cbind(a, b[ ,2], c[ ,2], d[ ,2], e[ , 2], f[ ,2], g[ ,2], h[ ,2], i[ ,2], j[ ,2])
-  colnames(stackstat) <- c("depth", "0 - 0.09", "0.1 - 0.19", "0.2 - 0.29", "0.3 - 0.39", "0.4 - 0.49", "0.5 - 0.59", "0.6 - 0.69", "0.7 - 0.79", "0.8 - 0.89", ">0.9")
-  stackstat <- melt(stackstat, id = "depth")
-  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
-  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
-  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]
-
-  colnames(stackstat) <- c("Depth", "Probability", "Cells")
-
-  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
-}
-
-```
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
+  for (r in 1:nrow(a)){
+    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
+    # d1 <- str_sub(a$Depth[r], end = 4)
+    d1 <- a$asclayer[r]
+	  for (ro in 1:nrow(cumst1)){
+      #d2 <- str_sub(cumst1$depth[ro], end = 4)
+	    d2 <- cumst1$depth[ro]
+      if (d1 == d2){
+          a$total_cells[r] <- cumst1$no_cells[ro]
+      }
+	  }
+    a$p_cells[r] <- (a$prob_bin[r]/a$total_cells[r])*100
+    a$Probability[r] <- 1
+  }
+    #a$asclayer <- as.numeric(a$asclayer)
+
+  a2 <- as.data.frame(a$asclayer)
+  a2$total_cells <- a$total_cells
+  a2$prob_bin <- a$total_cells - a$prob_bin
+  a2$p_cells <- (a2$prob_bin/a2$total_cells)*100
+  a2$Probability <- 0
+  colnames(a2)[1] <- "asclayer"
+
+  b <- rbind(a, a2)
+  colnames(b)[1] <- "Depth"
+
+  write.csv(b, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_dis_prob_cells.csv"), row.names = FALSE)
+}
+
+```
+
+based on Remi's code - maxkappa derrived habitat by depth
+
+```{r}
+require(data.table)
+require(purrr)
+require(lubridate)
+require(tidyverse)
+require(viridis)
+
+filenames <- list.files("../output/maxent/paper_plots/prob_by_depth/", pattern = "_maxkap_dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)
+
+all <- map_dfr(filenames, function(x) fread(x) %>% 
+                       mutate(Month=substr(basename(x),1,3))) %>% 
+  mutate(Depth = factor(round(Depth,1)),
+         Depth = factor(Depth,ordered = TRUE, levels = rev(levels(Depth))),
+         Month=factor(months(parse_date_time(Month,"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))))
+
+#create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
+dlst <- unique(all$Depth)
+dlst <- as.character(dlst)
+dlst <- as.numeric(dlst)
+dlst <- sort(dlst)
+
+#make probability a factor - sam added this
+all$Probability <- as.factor(all$Probability)
+
+#make the plot
+  ggplot(all) +
+  geom_col(aes(Depth, p_cells, fill = Probability),width=1) +
+  coord_flip()+
+  facet_wrap(~Month,ncol=2)+
+  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
+  scale_fill_manual(aesthetics = "fill", values=c('light yellow','black'))+
+  theme_classic()+
+  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
+  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47]))
+
+ggsave("../output/maxent/paper_plots/maxkap_prob_by_depth.png", width = 210, height = 297, units = "mm", dpi = 600)
+```
+
+#get max and min depth layers based on max kappa
+```{r}
+kapd <- list.files("../output/maxent/paper_plots/prob_by_depth/" , pattern = "_maxkap_dis_prob_cells.csv", full.names = TRUE)
+
+test <- read.csv(kapd[1])
+test <- subset(test, Probability == 1) #just keep the presence cells
+test <- test[order(test[, 1]),] #order by depth 
+#now need to find the row in which a value first appears in prob_bin column
+
+```

### 2020-02-12 src/comb-maxent.Rmd
updated plots (background point gen)

@@
   labs(y = "% Ocean at Depth", x = "Depth (meters)")+
-  scale_fill_manual(aesthetics = "fill", values=c('light yellow','black'))+
+  scale_fill_manual(aesthetics = "fill", values=c('light yellow','black'), name = "Presence")+
   theme_classic()+
@@
 ggsave("../output/maxent/paper_plots/maxkap_prob_by_depth.png", width = 210, height = 297, units = "mm", dpi = 600)
+```
+
+#get max and min depth layers based on max kappa
+```{r maxkap min max depth}
+kapd <- list.files("../output/maxent/paper_plots/prob_by_depth/" , pattern = "_maxkap_dis_prob_cells.csv", full.names = TRUE)
+kapmmd <- data.frame(mth = character(10), min = numeric(10), max = numeric(10), range = numeric(10))
+kapmmd$mth <- as.character(kapmmd$mth)
+
+for(k in 1:length(kapd)){
+  kapmmd$mth[k] <- str_sub(kapd[k], start = 44, end = 46)
+  kd <- read.csv(kapd[k])
+  kd <- subset(kd, Probability == 1) #just keep the presence cells
+  kd <- subset(kd, prob_bin > 0) # get the depths where there are presence cells
+  kd <- kd[order(kd[, 1]),] #order by depth 
+  kapmmd$min[k] <- kd$Depth[1] #minimum depth of presence cell
+  kapmmd$max[k] <- tail(kd$Depth, n = 1) #maximum depth of presence cell
+}
+
+kapmmd$range <- kapmmd$max - kapmmd$min #get depth range
+write.csv(kapmmd, "../output/maxent/paper_plots/hotspot_mapping/kappa-min-max-depth.csv")
+```
+
+#get lat long position of most northely, southerly, western, and eastern cell centroid (max kap)
+
+```{r maxkap cell centroid}
+filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", full.names = TRUE)
+lonlat <- data.frame(mth = character(10), west = numeric(10), east = numeric(10), north = numeric(10), south = numeric(10))
+lonlat$mth <- as.character(lonlat$mth)
+
+for (k in 1:length(filelist)){
+  ras <- raster(filelist[k])
+  crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" #the projection of the raster
+  lonlat$mth[k] <- str_sub(filelist[k], start = 46, end = 48)
+  cellloc <- rasterToPoints(ras, fun=function(x){x==1}) #make raster to point only for cells with the value of 1 (i.e. presence cells)
+  cellloc <- as.data.frame(cellloc) # x = lon, y = lat
+  cellloc <- cellloc[order(cellloc$x), ] #order dataframe by lon from low (west) to high (east)
+  lonlat$west[k] <- cellloc$x[1] #most western cell centroid
+  lonlat$east[k] <- tail(cellloc$x, n = 1) #most eastern cell centroid
+  cellloc <- cellloc[order(cellloc$y), ] #order dataframe by lat from low (south) to high (north)
+  lonlat$south[k] <- cellloc$y[1] #most southern cell centroid
+  lonlat$north[k] <- tail(cellloc$y, n = 1) #most northern cell centroid
+}
+
+write.csv(lonlat, "../output/maxent/paper_plots/hotspot_mapping/kappa-nswe.csv")
+
+```
+
+#get lat long position of most northely, southerly, western, and eastern cell centroid for oct, june, aug, may (to remove northern spots)
+
+```{r}
+filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", full.names = TRUE)
+lonlat <- data.frame(mth = character(10), west = numeric(10), east = numeric(10), north = numeric(10), south = numeric(10))
+lonlat$mth <- as.character(lonlat$mth)
+
+for (k in 1:length(filelist)){
+  ras <- raster(filelist[k])
+  crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" #the projection of the raster
+  lonlat$mth[k] <- str_sub(filelist[k], start = 46, end = 48)
+  cellloc <- rasterToPoints(ras, fun=function(x){x==1}) #make raster to point only for cells with the value of 1 (i.e. presence cells)
+  cellloc <- as.data.frame(cellloc) # x = lon, y = lat
+  cellloc <- subset(cellloc, y <55) #remove all the northern points
+  cellloc <- cellloc[order(cellloc$y), ] #order dataframe by lat from low (south) to high (north)
+  lonlat$south[k] <- cellloc$y[1] #most southern cell centroid
+  lonlat$north[k] <- tail(cellloc$y, n = 1) #most northern cell centroid
+  cellloc <- cellloc[order(cellloc$x), ] #order dataframe by lon from low (west) to high (east)
+  lonlat$west[k] <- cellloc$x[1] #most western cell centroid
+  lonlat$east[k] <- tail(cellloc$x, n = 1) #most eastern cell centroid
+
+}
+
+write.csv(lonlat, "../output/maxent/paper_plots/hotspot_mapping/kappa-nswe_nonorth.csv")
+
+```
+
+#background point selection plots
+
+```{r}
+test_back_df <- read.csv("../data/env/background_point_check/201406_allbackgroundpoints.csv", header = TRUE)
+```
+
+ok now create the first lot of random for 2014_06
+
+```{r}
+testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+```
+
+make long data
+```{r}
+bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
+bkpt$variable <- as.character(bkpt$variable)
+bkpt$no_points <- as.character(bkpt$no_points)
+
+bkpt$no_points[1:10000] <- "10,000"
+bkpt$no_points[10001:30000] <- "20,000"
+bkpt$no_points[30001:80000] <- "50,000"
+bkpt$no_points[80001:180000] <- "100,000"
+bkpt$no_points[180001:370000] <- "190,000"
+bkpt$variable <- "temp"
+bkpt$value[1:10000] <- testbk10000$temp_depth
+bkpt$value[10001:30000] <- testbk20000$temp_depth
+bkpt$value[30001:80000] <- testbk50000$temp_depth
+bkpt$value[80001:180000] <- testbk100000$temp_depth
+bkpt$value[180001:370000] <- testbk190000$temp_depth
+
+bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt2$variable <- as.character(bkpt2$variable)
+bkpt2$no_points[1:10000] <- "10,000"
+bkpt2$no_points[10001:30000] <- "20,000"
+bkpt2$no_points[30001:80000] <- "50,000"
+bkpt2$no_points[80001:180000] <- "100,000"
+bkpt2$no_points[180001:370000] <- "190,000"
+bkpt2$variable <- "salinity"
+bkpt2$value[1:10000] <- testbk10000$salinity_depth
+bkpt2$value[10001:30000] <- testbk20000$salinity_depth
+bkpt2$value[30001:80000] <- testbk50000$salinity_depth
+bkpt2$value[80001:180000] <- testbk100000$salinity_depth
+bkpt2$value[180001:370000] <- testbk190000$salinity_depth
+
+bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt3$variable <- as.character(bkpt3$variable)
+bkpt3$no_points[1:10000] <- "10,000"
+bkpt3$no_points[10001:30000] <- "20,000"
+bkpt3$no_points[30001:80000] <- "50,000"
+bkpt3$no_points[80001:180000] <- "100,000"
+bkpt3$no_points[180001:370000] <- "190,000"
+bkpt3$variable <- "o2"
+bkpt3$value[1:10000] <- testbk10000$o2_depth
+bkpt3$value[10001:30000] <- testbk20000$o2_depth
+bkpt3$value[30001:80000] <- testbk50000$o2_depth
+bkpt3$value[80001:180000] <- testbk100000$o2_depth
+bkpt3$value[180001:370000] <- testbk190000$o2_depth
+
+bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt4$variable <- as.character(bkpt4$variable)
+bkpt4$no_points[1:10000] <- "10,000"
+bkpt4$no_points[10001:30000] <- "20,000"
+bkpt4$no_points[30001:80000] <- "50,000"
+bkpt4$no_points[80001:180000] <- "100,000"
+bkpt4$no_points[180001:370000] <- "190,000"
+bkpt4$variable <- "chl"
+bkpt4$value[1:10000] <- testbk10000$chl_surface
+bkpt4$value[10001:30000] <- testbk20000$chl_surface
+bkpt4$value[30001:80000] <- testbk50000$chl_surface
+bkpt4$value[80001:180000] <- testbk100000$chl_surface
+bkpt4$value[180001:370000] <- testbk190000$chl_surface
+
+bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
+bk$no_points <- as.factor(bk$no_points)
+```
+
+```{r}
+ggplot(bk, aes(x = value)) +
+    stat_density(aes(group = no_points, color = no_points),position="identity",geom="line") +
+  facet_wrap(~variable,ncol=2)
+
+```
+
+problem here is that i don't want the axis shared
+
+fiddle with the dataset again
+```{r}
+t <- subset(bk, variable == "temp")
+t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+s <- subset(bk, variable == "salinity")
+s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+o <- subset(bk, variable == "o2")
+o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+c <- subset(bk, variable == "chl")
+c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+
+```
+
+plot
+```{r}
+p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+
+p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+
+ggsave("../output/maxent/paper_plots/bkpt_201406.png", width = 210, height = 297, units = "mm", dpi = 600)
+
+```
+
+200710
+
+```{r}
+test_back_df <- read.csv("../data/env/background_point_check/200710_allbackgroundpoints.csv", header = TRUE)
+```
+
+ok now create the first lot of random for 200710
+
+```{r}
+testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+```
+
+make long data
+```{r}
+bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
+bkpt$variable <- as.character(bkpt$variable)
+bkpt$no_points <- as.character(bkpt$no_points)
+
+bkpt$no_points[1:10000] <- "10,000"
+bkpt$no_points[10001:30000] <- "20,000"
+bkpt$no_points[30001:80000] <- "50,000"
+bkpt$no_points[80001:180000] <- "100,000"
+bkpt$no_points[180001:370000] <- "190,000"
+bkpt$variable <- "temp"
+bkpt$value[1:10000] <- testbk10000$temp_depth
+bkpt$value[10001:30000] <- testbk20000$temp_depth
+bkpt$value[30001:80000] <- testbk50000$temp_depth
+bkpt$value[80001:180000] <- testbk100000$temp_depth
+bkpt$value[180001:370000] <- testbk190000$temp_depth
+
+bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt2$variable <- as.character(bkpt2$variable)
+bkpt2$no_points[1:10000] <- "10,000"
+bkpt2$no_points[10001:30000] <- "20,000"
+bkpt2$no_points[30001:80000] <- "50,000"
+bkpt2$no_points[80001:180000] <- "100,000"
+bkpt2$no_points[180001:370000] <- "190,000"
+bkpt2$variable <- "salinity"
+bkpt2$value[1:10000] <- testbk10000$salinity_depth
+bkpt2$value[10001:30000] <- testbk20000$salinity_depth
+bkpt2$value[30001:80000] <- testbk50000$salinity_depth
+bkpt2$value[80001:180000] <- testbk100000$salinity_depth
+bkpt2$value[180001:370000] <- testbk190000$salinity_depth
+
+bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt3$variable <- as.character(bkpt3$variable)
+bkpt3$no_points[1:10000] <- "10,000"
+bkpt3$no_points[10001:30000] <- "20,000"
+bkpt3$no_points[30001:80000] <- "50,000"
+bkpt3$no_points[80001:180000] <- "100,000"
+bkpt3$no_points[180001:370000] <- "190,000"
+bkpt3$variable <- "o2"
+bkpt3$value[1:10000] <- testbk10000$o2_depth
+bkpt3$value[10001:30000] <- testbk20000$o2_depth
+bkpt3$value[30001:80000] <- testbk50000$o2_depth
+bkpt3$value[80001:180000] <- testbk100000$o2_depth
+bkpt3$value[180001:370000] <- testbk190000$o2_depth
+
+bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt4$variable <- as.character(bkpt4$variable)
+bkpt4$no_points[1:10000] <- "10,000"
+bkpt4$no_points[10001:30000] <- "20,000"
+bkpt4$no_points[30001:80000] <- "50,000"
+bkpt4$no_points[80001:180000] <- "100,000"
+bkpt4$no_points[180001:370000] <- "190,000"
+bkpt4$variable <- "chl"
+bkpt4$value[1:10000] <- testbk10000$chl_surface
+bkpt4$value[10001:30000] <- testbk20000$chl_surface
+bkpt4$value[30001:80000] <- testbk50000$chl_surface
+bkpt4$value[80001:180000] <- testbk100000$chl_surface
+bkpt4$value[180001:370000] <- testbk190000$chl_surface
+
+bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
+bk$no_points <- as.factor(bk$no_points)
+```
+
+fiddle with the dataset again
+```{r}
+t <- subset(bk, variable == "temp")
+t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+s <- subset(bk, variable == "salinity")
+s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+o <- subset(bk, variable == "o2")
+o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+c <- subset(bk, variable == "chl")
+c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+
+```
+
+plot
+```{r}
+p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+
+p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+
+ggsave("../output/maxent/paper_plots/bkpt_200710.png", width = 210, height = 297, units = "mm", dpi = 600)
+
+```
+199902
+
+```{r}
+test_back_df <- read.csv("../data/env/background_point_check/199902_allbackgroundpoints.csv", header = TRUE)
+```
+
+ok now create the first lot of random for 200710
+
+```{r}
+testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+```
+
+make long data
+```{r}
+bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
+bkpt$variable <- as.character(bkpt$variable)
+bkpt$no_points <- as.character(bkpt$no_points)
+
+bkpt$no_points[1:10000] <- "10,000"
+bkpt$no_points[10001:30000] <- "20,000"
+bkpt$no_points[30001:80000] <- "50,000"
+bkpt$no_points[80001:180000] <- "100,000"
+bkpt$no_points[180001:370000] <- "190,000"
+bkpt$variable <- "temp"
+bkpt$value[1:10000] <- testbk10000$temp_depth
+bkpt$value[10001:30000] <- testbk20000$temp_depth
+bkpt$value[30001:80000] <- testbk50000$temp_depth
+bkpt$value[80001:180000] <- testbk100000$temp_depth
+bkpt$value[180001:370000] <- testbk190000$temp_depth
+
+bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt2$variable <- as.character(bkpt2$variable)
+bkpt2$no_points[1:10000] <- "10,000"
+bkpt2$no_points[10001:30000] <- "20,000"
+bkpt2$no_points[30001:80000] <- "50,000"
+bkpt2$no_points[80001:180000] <- "100,000"
+bkpt2$no_points[180001:370000] <- "190,000"
+bkpt2$variable <- "salinity"
+bkpt2$value[1:10000] <- testbk10000$salinity_depth
+bkpt2$value[10001:30000] <- testbk20000$salinity_depth
+bkpt2$value[30001:80000] <- testbk50000$salinity_depth
+bkpt2$value[80001:180000] <- testbk100000$salinity_depth
+bkpt2$value[180001:370000] <- testbk190000$salinity_depth
+
+bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt3$variable <- as.character(bkpt3$variable)
+bkpt3$no_points[1:10000] <- "10,000"
+bkpt3$no_points[10001:30000] <- "20,000"
+bkpt3$no_points[30001:80000] <- "50,000"
+bkpt3$no_points[80001:180000] <- "100,000"
+bkpt3$no_points[180001:370000] <- "190,000"
+bkpt3$variable <- "o2"
+bkpt3$value[1:10000] <- testbk10000$o2_depth
+bkpt3$value[10001:30000] <- testbk20000$o2_depth
+bkpt3$value[30001:80000] <- testbk50000$o2_depth
+bkpt3$value[80001:180000] <- testbk100000$o2_depth
+bkpt3$value[180001:370000] <- testbk190000$o2_depth
+
+bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt4$variable <- as.character(bkpt4$variable)
+bkpt4$no_points[1:10000] <- "10,000"
+bkpt4$no_points[10001:30000] <- "20,000"
+bkpt4$no_points[30001:80000] <- "50,000"
+bkpt4$no_points[80001:180000] <- "100,000"
+bkpt4$no_points[180001:370000] <- "190,000"
+bkpt4$variable <- "chl"
+bkpt4$value[1:10000] <- testbk10000$chl_surface
+bkpt4$value[10001:30000] <- testbk20000$chl_surface
+bkpt4$value[30001:80000] <- testbk50000$chl_surface
+bkpt4$value[80001:180000] <- testbk100000$chl_surface
+bkpt4$value[180001:370000] <- testbk190000$chl_surface
+
+bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
+bk$no_points <- as.factor(bk$no_points)
+```
+
+fiddle with the dataset again
+```{r}
+t <- subset(bk, variable == "temp")
+t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+s <- subset(bk, variable == "salinity")
+s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+o <- subset(bk, variable == "o2")
+o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+c <- subset(bk, variable == "chl")
+c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+
+```
+
+plot
+```{r}
+p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+
+p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+
+ggsave("../output/maxent/paper_plots/bkpt_199902.png", width = 210, height = 297, units = "mm", dpi = 600)
+
 ```

### 2020-02-12 src/testing_file.Rmd
updated plots (background point gen)

@@
 
-then create an summary map....
-```{r >maxkap prob monthly 2D representation}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-
-for (m in 1:length(mthord)){
-  mthna <- mthord[m]
-  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_maxkap_.*.asc$"), full.names = TRUE)
-  binstack <- stack(binaryprob_list)
-  stacksum <- sum(binstack, na.rm = TRUE)
-  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.asc"), overwrite=TRUE)
-  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_maxkap_2d.png")))
-  dev.off()
-}
-
-```
-
-now create summary maps where values are binary - not counts of cells
-```{r}
-binprob <- function(x){
-  ifelse(x > 0, 1,
-  ifelse(x <= 0, 0, NA))
-}
-
-  clogasc_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = "*_maxkap_2d.asc", full.names = TRUE)
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- str_sub(clogasc_list[cum], start = 46, end = -8)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", ascname, "_single.png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-```
-
-now create a table to see how many cells there are >maxkap in each month (for editing plots)
-
-```{r count cells >maxkap}
-
-binaryprob_list <- list.files("../output/maxent/paper_plots/hotspot_mapping", pattern = ".*_maxkap_2d.asc", full.names = TRUE)
-sum2d <- data.frame(matrix(ncol = 2, nrow = 10))
-sum2d <- setNames(sum2d, c("month", "count"))
-
-for (r in 1:length(binaryprob_list)){
-  ras <- raster(binaryprob_list[r])
-  rnm <- names(ras)
-  sum2d$month[r] <- gsub("_maxkap_2d", "", x = rnm)
-  sum2d$count[r] <- cellStats(ras, sum, asSample = TRUE)
-}
-
-#order table from march to december
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-sum2d <- sum2d[match(mthord, sum2d$month), ]
-
-#get proportion
-tcs <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE) #file showing how many cells per depth
-#tcs <- tcs[-c(1:3), ] #because the top three rows are deeper than the capelin obs...
-totc <- sum(tcs$no_cells)
-
-for (r in 1:nrow(sum2d)){
-  sum2d$proportion[r] <- (sum2d$count[r]/totc)
-}
-
-write.csv(sum2d, "../output/maxent/paper_plots/cell_count_maxkap.csv", row.names = FALSE)
-
-```
-
 ```{r}
@@
 kapd <- list.files("../output/maxent/paper_plots/prob_by_depth/" , pattern = "_maxkap_dis_prob_cells.csv", full.names = TRUE)
-
-test <- read.csv(kapd[1])
-test <- subset(test, Probability == 1) #just keep the presence cells
-test <- test[order(test[, 1]),] #order by depth 
-#now need to find the row in which a value first appears in prob_bin column
-
-```
+kapmmd <- data.frame(mth = character(10), min = numeric(10), max = numeric(10))
+kapmmd$mth <- as.character(kapmmd$mth)
+
+for(k in 1:length(kapd)){
+  kapmmd$mth[k] <- str_sub(kapd[k], start = 44, end = 46)
+  kd <- read.csv(kapd[k])
+  kd <- subset(kd, Probability == 1) #just keep the presence cells
+  kd <- subset(kd, prob_bin > 0) # get the depths where there are presence cells
+  kd <- kd[order(kd[, 1]),] #order by depth 
+  kapmmd$min[k] <- kd$Depth[1] #minimum depth of presence cell
+  kapmmd$max[k] <- tail(kd$Depth, n = 1) #maximum depth of presence cell
+}
+
+write.csv(kapmmd, "../output/maxent/paper_plots/hotspot_mapping/kappa-min-max-depth.csv")
+```
+
+#get lat long position of most northely, southerly, western, and eastern cell centroid
+
+```{r}
+filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", full.names = TRUE)
+lonlat <- data.frame(mth = character(10), west = numeric(10), east = numeric(10), north = numeric(10), south = numeric(10))
+lonlat$mth <- as.character(lonlat$mth)
+
+for (k in 1:length(filelist)){
+  ras <- raster(filelist[9])
+  crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" #the projection of the raster
+  lonlat$mth[k] <- str_sub(filelist[k], start = 46, end = 48)
+  cellloc <- rasterToPoints(ras, fun=function(x){x==1}) #make raster to point only for cells with the value of 1 (i.e. presence cells)
+  cellloc <- as.data.frame(cellloc) # x = lon, y = lat
+  cellloc <- cellloc[order(cellloc$x), ] #order dataframe by lon from low (west) to high (east)
+  lonlat$west[k] <- cellloc$x[1] #most western cell centroid
+  lonlat$east[k] <- tail(cellloc$x, n = 1) #most eastern cell centroid
+  cellloc <- cellloc[order(cellloc$y), ] #order dataframe by lat from low (south) to high (north)
+  lonlat$south[k] <- cellloc$y[1] #most southern cell centroid
+  lonlat$north[k] <- tail(cellloc$y, n = 1) #most northern cell centroid
+}
+
+write.csv(lonlat, "../output/maxent/paper_plots/hotspot_mapping/kappa-nswe.csv")
+
+```
+
+#get lat long position of most northely, southerly, western, and eastern cell centroid for oct, june, aug, may (to remove northern spots)
+
+```{r}
+filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", full.names = TRUE)
+lonlat <- data.frame(mth = character(10), west = numeric(10), east = numeric(10), north = numeric(10), south = numeric(10))
+lonlat$mth <- as.character(lonlat$mth)
+
+for (k in 1:length(filelist)){
+  ras <- raster(filelist[k])
+  crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" #the projection of the raster
+  lonlat$mth[k] <- str_sub(filelist[k], start = 46, end = 48)
+  cellloc <- rasterToPoints(ras, fun=function(x){x==1}) #make raster to point only for cells with the value of 1 (i.e. presence cells)
+  cellloc <- as.data.frame(cellloc) # x = lon, y = lat
+  cellloc <- subset(cellloc, y <55) #remove all the northern points
+  cellloc <- cellloc[order(cellloc$y), ] #order dataframe by lat from low (south) to high (north)
+  lonlat$south[k] <- cellloc$y[1] #most southern cell centroid
+  lonlat$north[k] <- tail(cellloc$y, n = 1) #most northern cell centroid
+  cellloc <- cellloc[order(cellloc$x), ] #order dataframe by lon from low (west) to high (east)
+  lonlat$west[k] <- cellloc$x[1] #most western cell centroid
+  lonlat$east[k] <- tail(cellloc$x, n = 1) #most eastern cell centroid
+
+}
+
+write.csv(lonlat, "../output/maxent/paper_plots/hotspot_mapping/kappa-nswe_nonorth.csv")
+
+```
+
+```{r}
+t1 <- distVincentyEllipsoid(c(0,0),c(90,90))
+# on a 'Clarke 1880' ellipsoid
+test <- distVincentyEllipsoid(c(0,0),c(90,90), a=6378249.145, b=6356514.86955, f=1/293.465)
+```
+
+#background point selection plots
+
+```{r}
+test_back_df <- read.csv("../data/env/background_point_check/201406_allbackgroundpoints.csv", header = TRUE)
+```
+
+ok now create the first lot of random for 2014_06
+
+```{r}
+testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+```
+
+make long data
+```{r}
+bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
+bkpt$variable <- as.character(bkpt$variable)
+bkpt$no_points <- as.character(bkpt$no_points)
+
+bkpt$no_points[1:10000] <- "10,000"
+bkpt$no_points[10001:30000] <- "20,000"
+bkpt$no_points[30001:80000] <- "50,000"
+bkpt$no_points[80001:180000] <- "100,000"
+bkpt$no_points[180001:370000] <- "190,000"
+bkpt$variable <- "temp"
+bkpt$value[1:10000] <- testbk10000$temp_depth
+bkpt$value[10001:30000] <- testbk20000$temp_depth
+bkpt$value[30001:80000] <- testbk50000$temp_depth
+bkpt$value[80001:180000] <- testbk100000$temp_depth
+bkpt$value[180001:370000] <- testbk190000$temp_depth
+
+bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt2$variable <- as.character(bkpt2$variable)
+bkpt2$no_points[1:10000] <- "10,000"
+bkpt2$no_points[10001:30000] <- "20,000"
+bkpt2$no_points[30001:80000] <- "50,000"
+bkpt2$no_points[80001:180000] <- "100,000"
+bkpt2$no_points[180001:370000] <- "190,000"
+bkpt2$variable <- "salinity"
+bkpt2$value[1:10000] <- testbk10000$salinity_depth
+bkpt2$value[10001:30000] <- testbk20000$salinity_depth
+bkpt2$value[30001:80000] <- testbk50000$salinity_depth
+bkpt2$value[80001:180000] <- testbk100000$salinity_depth
+bkpt2$value[180001:370000] <- testbk190000$salinity_depth
+
+bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt3$variable <- as.character(bkpt3$variable)
+bkpt3$no_points[1:10000] <- "10,000"
+bkpt3$no_points[10001:30000] <- "20,000"
+bkpt3$no_points[30001:80000] <- "50,000"
+bkpt3$no_points[80001:180000] <- "100,000"
+bkpt3$no_points[180001:370000] <- "190,000"
+bkpt3$variable <- "o2"
+bkpt3$value[1:10000] <- testbk10000$o2_depth
+bkpt3$value[10001:30000] <- testbk20000$o2_depth
+bkpt3$value[30001:80000] <- testbk50000$o2_depth
+bkpt3$value[80001:180000] <- testbk100000$o2_depth
+bkpt3$value[180001:370000] <- testbk190000$o2_depth
+
+bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt4$variable <- as.character(bkpt4$variable)
+bkpt4$no_points[1:10000] <- "10,000"
+bkpt4$no_points[10001:30000] <- "20,000"
+bkpt4$no_points[30001:80000] <- "50,000"
+bkpt4$no_points[80001:180000] <- "100,000"
+bkpt4$no_points[180001:370000] <- "190,000"
+bkpt4$variable <- "chl"
+bkpt4$value[1:10000] <- testbk10000$chl_surface
+bkpt4$value[10001:30000] <- testbk20000$chl_surface
+bkpt4$value[30001:80000] <- testbk50000$chl_surface
+bkpt4$value[80001:180000] <- testbk100000$chl_surface
+bkpt4$value[180001:370000] <- testbk190000$chl_surface
+
+bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
+bk$no_points <- as.factor(bk$no_points)
+```
+
+```{r}
+ggplot(bk, aes(x = value)) +
+    stat_density(aes(group = no_points, color = no_points),position="identity",geom="line") +
+  facet_wrap(~variable,ncol=2)
+
+```
+
+problem here is that i don't want the axis shared
+
+fiddle with the dataset again
+```{r}
+t <- subset(bk, variable == "temp")
+t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+s <- subset(bk, variable == "salinity")
+s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+o <- subset(bk, variable == "o2")
+o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+c <- subset(bk, variable == "chl")
+c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+
+```
+
+plot
+```{r}
+p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+
+p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+
+ggsave("../output/maxent/paper_plots/bkpt_201406.png", width = 210, height = 297, units = "mm", dpi = 600)
+
+```
+
+200710
+
+```{r}
+test_back_df <- read.csv("../data/env/background_point_check/200710_allbackgroundpoints.csv", header = TRUE)
+```
+
+ok now create the first lot of random for 200710
+
+```{r}
+testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+```
+
+make long data
+```{r}
+bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
+bkpt$variable <- as.character(bkpt$variable)
+bkpt$no_points <- as.character(bkpt$no_points)
+
+bkpt$no_points[1:10000] <- "10,000"
+bkpt$no_points[10001:30000] <- "20,000"
+bkpt$no_points[30001:80000] <- "50,000"
+bkpt$no_points[80001:180000] <- "100,000"
+bkpt$no_points[180001:370000] <- "190,000"
+bkpt$variable <- "temp"
+bkpt$value[1:10000] <- testbk10000$temp_depth
+bkpt$value[10001:30000] <- testbk20000$temp_depth
+bkpt$value[30001:80000] <- testbk50000$temp_depth
+bkpt$value[80001:180000] <- testbk100000$temp_depth
+bkpt$value[180001:370000] <- testbk190000$temp_depth
+
+bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt2$variable <- as.character(bkpt2$variable)
+bkpt2$no_points[1:10000] <- "10,000"
+bkpt2$no_points[10001:30000] <- "20,000"
+bkpt2$no_points[30001:80000] <- "50,000"
+bkpt2$no_points[80001:180000] <- "100,000"
+bkpt2$no_points[180001:370000] <- "190,000"
+bkpt2$variable <- "salinity"
+bkpt2$value[1:10000] <- testbk10000$salinity_depth
+bkpt2$value[10001:30000] <- testbk20000$salinity_depth
+bkpt2$value[30001:80000] <- testbk50000$salinity_depth
+bkpt2$value[80001:180000] <- testbk100000$salinity_depth
+bkpt2$value[180001:370000] <- testbk190000$salinity_depth
+
+bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt3$variable <- as.character(bkpt3$variable)
+bkpt3$no_points[1:10000] <- "10,000"
+bkpt3$no_points[10001:30000] <- "20,000"
+bkpt3$no_points[30001:80000] <- "50,000"
+bkpt3$no_points[80001:180000] <- "100,000"
+bkpt3$no_points[180001:370000] <- "190,000"
+bkpt3$variable <- "o2"
+bkpt3$value[1:10000] <- testbk10000$o2_depth
+bkpt3$value[10001:30000] <- testbk20000$o2_depth
+bkpt3$value[30001:80000] <- testbk50000$o2_depth
+bkpt3$value[80001:180000] <- testbk100000$o2_depth
+bkpt3$value[180001:370000] <- testbk190000$o2_depth
+
+bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt4$variable <- as.character(bkpt4$variable)
+bkpt4$no_points[1:10000] <- "10,000"
+bkpt4$no_points[10001:30000] <- "20,000"
+bkpt4$no_points[30001:80000] <- "50,000"
+bkpt4$no_points[80001:180000] <- "100,000"
+bkpt4$no_points[180001:370000] <- "190,000"
+bkpt4$variable <- "chl"
+bkpt4$value[1:10000] <- testbk10000$chl_surface
+bkpt4$value[10001:30000] <- testbk20000$chl_surface
+bkpt4$value[30001:80000] <- testbk50000$chl_surface
+bkpt4$value[80001:180000] <- testbk100000$chl_surface
+bkpt4$value[180001:370000] <- testbk190000$chl_surface
+
+bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
+bk$no_points <- as.factor(bk$no_points)
+```
+
+fiddle with the dataset again
+```{r}
+t <- subset(bk, variable == "temp")
+t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+s <- subset(bk, variable == "salinity")
+s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+o <- subset(bk, variable == "o2")
+o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+c <- subset(bk, variable == "chl")
+c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+
+```
+
+plot
+```{r}
+p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+
+p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+
+ggsave("../output/maxent/paper_plots/bkpt_200710.png", width = 210, height = 297, units = "mm", dpi = 600)
+
+```
+199902
+
+```{r}
+test_back_df <- read.csv("../data/env/background_point_check/199902_allbackgroundpoints.csv", header = TRUE)
+```
+
+ok now create the first lot of random for 200710
+
+```{r}
+testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
+testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
+testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
+testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
+testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
+```
+
+make long data
+```{r}
+bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
+bkpt$variable <- as.character(bkpt$variable)
+bkpt$no_points <- as.character(bkpt$no_points)
+
+bkpt$no_points[1:10000] <- "10,000"
+bkpt$no_points[10001:30000] <- "20,000"
+bkpt$no_points[30001:80000] <- "50,000"
+bkpt$no_points[80001:180000] <- "100,000"
+bkpt$no_points[180001:370000] <- "190,000"
+bkpt$variable <- "temp"
+bkpt$value[1:10000] <- testbk10000$temp_depth
+bkpt$value[10001:30000] <- testbk20000$temp_depth
+bkpt$value[30001:80000] <- testbk50000$temp_depth
+bkpt$value[80001:180000] <- testbk100000$temp_depth
+bkpt$value[180001:370000] <- testbk190000$temp_depth
+
+bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt2$variable <- as.character(bkpt2$variable)
+bkpt2$no_points[1:10000] <- "10,000"
+bkpt2$no_points[10001:30000] <- "20,000"
+bkpt2$no_points[30001:80000] <- "50,000"
+bkpt2$no_points[80001:180000] <- "100,000"
+bkpt2$no_points[180001:370000] <- "190,000"
+bkpt2$variable <- "salinity"
+bkpt2$value[1:10000] <- testbk10000$salinity_depth
+bkpt2$value[10001:30000] <- testbk20000$salinity_depth
+bkpt2$value[30001:80000] <- testbk50000$salinity_depth
+bkpt2$value[80001:180000] <- testbk100000$salinity_depth
+bkpt2$value[180001:370000] <- testbk190000$salinity_depth
+
+bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt3$variable <- as.character(bkpt3$variable)
+bkpt3$no_points[1:10000] <- "10,000"
+bkpt3$no_points[10001:30000] <- "20,000"
+bkpt3$no_points[30001:80000] <- "50,000"
+bkpt3$no_points[80001:180000] <- "100,000"
+bkpt3$no_points[180001:370000] <- "190,000"
+bkpt3$variable <- "o2"
+bkpt3$value[1:10000] <- testbk10000$o2_depth
+bkpt3$value[10001:30000] <- testbk20000$o2_depth
+bkpt3$value[30001:80000] <- testbk50000$o2_depth
+bkpt3$value[80001:180000] <- testbk100000$o2_depth
+bkpt3$value[180001:370000] <- testbk190000$o2_depth
+
+bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
+bkpt4$variable <- as.character(bkpt4$variable)
+bkpt4$no_points[1:10000] <- "10,000"
+bkpt4$no_points[10001:30000] <- "20,000"
+bkpt4$no_points[30001:80000] <- "50,000"
+bkpt4$no_points[80001:180000] <- "100,000"
+bkpt4$no_points[180001:370000] <- "190,000"
+bkpt4$variable <- "chl"
+bkpt4$value[1:10000] <- testbk10000$chl_surface
+bkpt4$value[10001:30000] <- testbk20000$chl_surface
+bkpt4$value[30001:80000] <- testbk50000$chl_surface
+bkpt4$value[80001:180000] <- testbk100000$chl_surface
+bkpt4$value[180001:370000] <- testbk190000$chl_surface
+
+bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
+bk$no_points <- as.factor(bk$no_points)
+```
+
+fiddle with the dataset again
+```{r}
+t <- subset(bk, variable == "temp")
+t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+s <- subset(bk, variable == "salinity")
+s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+o <- subset(bk, variable == "o2")
+o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+c <- subset(bk, variable == "chl")
+c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
+
+```
+
+plot
+```{r}
+p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
+
+p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+
+ggsave("../output/maxent/paper_plots/bkpt_199902.png", width = 210, height = 297, units = "mm", dpi = 600)
+
+```

### 2020-02-13 src/comb-maxent.Rmd
minor change to package

@@
 options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2", "tidyverse", "patchwork", "purrr")
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2", "tidyverse", "patchwork", "purrr", "Hmisc")
 source("../src/ipak.R")
@@
 # rename the months so they are capitalised
+
 permu$Month <- capitalize(permu$Month)
@@
 ```{r permulation plot}
+library(cowplot)
 permu <- read.csv("../output/maxent/models/test-comb/permutation_avg.csv")

### 2020-02-17 src/testing_file.Rmd
building appendix example (bent)

@@
 ```
+
+#layer join images
+
+```{r}
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+s <- shapefile("../data/plot4-variables/landp/land-wgs84v2.shp")
+crs(s) <- proj
+
+temp9 <- raster("../data/env/asciitrim/ascii2004/2004_07_temp.nc_9.82275.asc")
+crs(temp9) <- proj
+temp221 <- raster("../data/env/asciitrim/ascii2004/2004_07_temp.nc_221.141.asc")
+crs(temp221) <- proj
+temp508 <- raster("../data/env/asciitrim/ascii2004/2004_07_temp.nc_508.64.asc")
+crs(temp508) <- proj
+
+ts <- stack(temp9, temp221, temp508)
+
+levelplot(ts, layout = c(1, 3), scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c("", "", ""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/temp-example.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+
+salinity9 <- raster("../data/env/asciitrim/ascii2004/2004_07_salinity.nc_9.82275.asc")
+crs(salinity9) <- proj
+salinity221 <- raster("../data/env/asciitrim/ascii2004/2004_07_salinity.nc_221.141.asc")
+crs(salinity221) <- proj
+salinity508 <- raster("../data/env/asciitrim/ascii2004/2004_07_salinity.nc_508.64.asc")
+crs(salinity508) <- proj
+
+ts <- stack(salinity9, salinity221, salinity508)
+
+levelplot(ts, layout = c(1, 3), scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c("", "", ""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/salinity-example.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+
+```
+
+```{r}
+ras <- raster("../output/maxent/paper_plots/bent_illus/tempmerge.asc")
+levelplot(ras, scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c(""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/temp-concat.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+
+ras <- raster("../output/maxent/paper_plots/bent_illus/salmerge.asc")
+levelplot(ras, scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c(""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/sal-concat.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+```

### 2020-02-18 src/comb-maxent.Rmd
figure creation

@@
 
+#plot for prediction method in appendix
+
+```{r}
+library(raster)
+library(rasterVis)
+library(viridis)
+```
+
+```{r plot sample depths}
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+crs(s) <- proj
+
+temp9 <- raster("../data/env/asciitrim/ascii2004/2004_07_temp.nc_9.82275.asc")
+crs(temp9) <- proj
+temp221 <- raster("../data/env/asciitrim/ascii2004/2004_07_temp.nc_221.141.asc")
+crs(temp221) <- proj
+temp508 <- raster("../data/env/asciitrim/ascii2004/2004_07_temp.nc_508.64.asc")
+crs(temp508) <- proj
+
+ts <- stack(temp9, temp221, temp508)
+
+levelplot(ts, layout = c(1, 3), scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c("", "", ""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/temp-example.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+
+salinity9 <- raster("../data/env/asciitrim/ascii2004/2004_07_salinity.nc_9.82275.asc")
+crs(salinity9) <- proj
+salinity221 <- raster("../data/env/asciitrim/ascii2004/2004_07_salinity.nc_221.141.asc")
+crs(salinity221) <- proj
+salinity508 <- raster("../data/env/asciitrim/ascii2004/2004_07_salinity.nc_508.64.asc")
+crs(salinity508) <- proj
+
+ts <- stack(salinity9, salinity221, salinity508)
+
+levelplot(ts, layout = c(1, 3), scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c("", "", ""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/salinity-example.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+
+```
+
+```{r plot env concat}
+ras <- raster("../output/maxent/paper_plots/bent_illus/tempmerge.asc")
+levelplot(ras, scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c(""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/temp-concat.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+
+ras <- raster("../output/maxent/paper_plots/bent_illus/salmerge.asc")
+levelplot(ras, scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c(""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/sal-concat.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+```
+
+```{r plot concat predictions}
+ras <- raster("../output/maxent/paper_plots/bent_illus/predmerge.asc")
+levelplot(ras, scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c(""), colorkey=FALSE, margin = FALSE, par.setting = magmaTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/pred-concat.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+```
+
+```{r plot split predictions}
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+pred9 <- raster("../output/maxent/paper_plots/bent_illus/pred/cloglog_2004_07_9.82275.asc")
+crs(pred9) <- proj
+pred221 <- raster("../output/maxent/paper_plots/bent_illus/pred/cloglog_2004_07_221.141.asc")
+crs(pred221) <- proj
+pred508 <- raster("../output/maxent/paper_plots/bent_illus/pred/cloglog_2004_07_508.64.asc")
+crs(pred508) <- proj
+
+ts <- stack(pred9, pred221, pred508)
+
+levelplot(ts, layout = c(1, 3), scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c("", "", ""), colorkey=FALSE, margin = FALSE, par.setting = magmaTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/pred-example.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+```
+
 # variable maps

### 2020-02-18 src/testing_file.Rmd
figure creation

--- 2020-02-12 src/testing_file.Rmd
@@
 ```
+
+#plot for prediction method in appendix
+
+```{r}
+library(raster)
+library(rasterVis)
+library(viridis)
+```
+
+```{r plot sample depths}
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+crs(s) <- proj
+
+temp9 <- raster("../data/env/asciitrim/ascii2004/2004_07_temp.nc_9.82275.asc")
+crs(temp9) <- proj
+temp221 <- raster("../data/env/asciitrim/ascii2004/2004_07_temp.nc_221.141.asc")
+crs(temp221) <- proj
+temp508 <- raster("../data/env/asciitrim/ascii2004/2004_07_temp.nc_508.64.asc")
+crs(temp508) <- proj
+
+ts <- stack(temp9, temp221, temp508)
+
+levelplot(ts, layout = c(1, 3), scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c("", "", ""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/temp-example.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+
+salinity9 <- raster("../data/env/asciitrim/ascii2004/2004_07_salinity.nc_9.82275.asc")
+crs(salinity9) <- proj
+salinity221 <- raster("../data/env/asciitrim/ascii2004/2004_07_salinity.nc_221.141.asc")
+crs(salinity221) <- proj
+salinity508 <- raster("../data/env/asciitrim/ascii2004/2004_07_salinity.nc_508.64.asc")
+crs(salinity508) <- proj
+
+ts <- stack(salinity9, salinity221, salinity508)
+
+levelplot(ts, layout = c(1, 3), scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c("", "", ""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/salinity-example.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+
+```
+
+```{r plot env concat}
+ras <- raster("../output/maxent/paper_plots/bent_illus/tempmerge.asc")
+levelplot(ras, scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c(""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/temp-concat.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+
+ras <- raster("../output/maxent/paper_plots/bent_illus/salmerge.asc")
+levelplot(ras, scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c(""), colorkey=FALSE, margin = FALSE, par.setting = viridisTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/sal-concat.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+```
+
+```{r plot concat predictions}
+ras <- raster("../output/maxent/paper_plots/bent_illus/predmerge.asc")
+levelplot(ras, scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c(""), colorkey=FALSE, margin = FALSE, par.setting = magmaTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/pred-concat.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+```
+
+```{r plot split predictions}
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+pred9 <- raster("../output/maxent/paper_plots/bent_illus/pred/cloglog_2004_07_9.82275.asc")
+crs(pred9) <- proj
+pred221 <- raster("../output/maxent/paper_plots/bent_illus/pred/cloglog_2004_07_221.141.asc")
+crs(pred221) <- proj
+pred508 <- raster("../output/maxent/paper_plots/bent_illus/pred/cloglog_2004_07_508.64.asc")
+crs(pred508) <- proj
+
+ts <- stack(pred9, pred221, pred508)
+
+levelplot(ts, layout = c(1, 3), scales=list(draw=FALSE), xlab = NULL, ylab = NULL, names.attr = c("", "", ""), colorkey=FALSE, margin = FALSE, par.setting = magmaTheme)
+dev.copy(png, (paste0("../output/maxent/paper_plots/bent_illus/pred-example.png")), width = 1080,height = 1920)
+dev.off() # stops automatic saving of the plot to a png))
+```

### 2020-05-27 src/Public-maxent.Rmd
review edit fiddling

--- 2020-02-12 src/comb-maxent.Rmd
@@
+---
+title: "maxent model plus plot creation"
+author: "Samantha Andrews"
+output:
+  html_document:
+    df_print: paged
+  html_notebook: default
+  pdf_document: default
+---
+
+# Overview
+A note to anyone who might happen to stumble across this...
+I am a beginner in R and have had no exposure to similar languages.
+I don't know what I'm doing.
+The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.
+
+Built with 'r getRversion()'..
+
+# Package dependencies
+You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163).
+Note this function checks to see if the packages are installed first before loading.
 
 ```{r pre-install & load packages, include=FALSE}
-options(java.parameters = "-Xmx2g" )#needs to run before dismo loaded - boosts memory for java to 1g
-packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "tidyr", "enmSdm", "ggplot2", "plyr", "reshape2", "tidyverse", "patchwork", "purrr")
+options(java.parameters = "-Xmx3g")#needs to run before dismo loaded - boosts memory for java to 3g
+packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice", "enmSdm", "ggplot2", "plyr", "reshape2", "tidyverse", "patchwork", "purrr", "Hmisc", "cowplot", "sf", "lubridate")
 source("../src/ipak.R")
@@
 
-#need to reclassify cells outside the mcp to NA for predictions
-```{r}
-outsidemcp <- shapefile("../data/env/landp/outside-mcp-wgs84.shp") #made in arcgis - in aea
-
-ascdirs <- list.dirs("../data/env/ascii-original", full.names = FALSE, recursive = FALSE)
-
-for (ad in 1:length(ascdirs)){
-  timeslice <- ascdirs[ad]
-  ascfiles <- list.files(paste0("../data/env/ascii-original/", timeslice), pattern = "\\.asc$", full.names = TRUE)
-  for (file in 1:length(ascfiles)){
-    ras<- raster(ascfiles[file])
-    crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
-    ascname <- str_sub(ascfiles[file], start = 38)
-    ras[outsidemcp] <- 500000
-    ras[ras == 500000] <- NA
-    writeRaster(ras, paste0("../data/env/asciitrim/", ascdirs[ad], "/", ascname), overwrite=TRUE)
-  }
-}
-
-```
-
 load the presence-background dataset
-("../output/bio/presab_nafo_yr.csv")
-
-select month
-```{r month of interest -change}
-mthna <- "dec"
-mthno <- "12"
-```
 
 ```{r load pres-background dataset}
-presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
-colnames(presback)
-```
-
-remove all rows with the year 2015 (due to low number of samples)
-```{r}
-presback <- subset(presback, year!="2015")
-```
-
-this dataset contains a lot of columns you don't need for the maxent models. Remove them here
-
-```{r remove unnecessary columns}
-presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt))
-colnames(presback)
-```
-
-monthly model data selection and prep
+presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE) # this file is available in the Figshare project
+```
+
 For each monthly model, subset the data to the month you want, and the variables you want (Based on spearmans and VIF)
+Ex: edit this list for variables used in each month (or simply replace with a single list if, as in this case, months have the same variables)
 
@@
 ```
+
+```{r train and test inc AIC for reg val}
+presyrs <-  #this is to get years with complete presence data
+
+modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
+aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
+betavallist <- cbind(aicetc$regularization)
+
+for (year in 1:nrow(modelmonth)){
+  timeslice <- modelmonth[year] #testing timeslice
+  yr <- str_sub(timeslice, -4) #the year being used for testing
+  train <- subset(prback, year != yr) #training data is all years but testing
+  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
+  test <-  subset(prback, year == yr) #creates the testing dataset
+  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
+  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
+  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
+
+  for (h in 1:nrow(betavallist)){
+    betaval <-  betavallist[h]
+    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c('plots=FALSE', "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))
+
+    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
+
+    #count number of non-zero coefficients, excluding the final four rows
+    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
+    coefile <- head(coefile, -4)
+    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
+    nocoeff <- nrow(coefile2)
+    bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
+    bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
+
+    #no presences used in model (less those used for testing)
+    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
+    notrainingpoints <- maxresults$X.Training.samples
+    testauc <- maxresults$Test.AUC
+    trainauc <- maxresults$Training.AUC
+
+    #natural log
+    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
+    samplepred$nat_log <- log(samplepred$Raw.prediction)
+    sumsamplepred <- sum(samplepred$nat_log)
+
+   #test auc scores and other lovely things you might want to include e.g.
+    ev <- evaluate(p = testp, a= testb, model = mod)
+    evauc <- ev@auc
+    evcor <- ev@cor
+    evpcor <- ev@pcor
+    evnp <- ev@np
+    evna <- ev@na
+    evmaxtprtnr <- threshold(ev, 'spec_sens')
+    evmaxkap <- threshold(ev, 'kappa')
+    evno_omission <- threshold(ev, 'no_omission')
+    evprevalence <- threshold(ev, 'prevalence')
+    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
+    evsensitivity <- threshold(ev, 'sensitivity')
+
+    for (i in 1:nrow(aicetc)){
+    if(aicetc$regularization[i] == betaval){
+      aicetc$train_AUC[i] <- trainauc
+      aicetc$no_trainng_points[i] <- notrainingpoints
+      aicetc$coefficients[i] <- nocoeff
+      aicetc$sum.of.logs[i] <- sumsamplepred
+      aicetc$test_AUC[i] <- evauc
+      aicetc$test_evcor[i] <- evcor
+      aicetc$test_evpcor[i] <- evpcor
+      aicetc$test_nopresence[i] <- evnp
+      aicetc$test_nobackground[i] <- evna
+      aicetc$test_maxsumsenspec[i] <- evmaxtprtnr
+      aicetc$test_maxkap[i] <- evmaxkap
+      aicetc$test_no_omission[i] <- evno_omission
+      aicetc$test_prevalence[i] <- evprevalence
+      aicetc$test_equal_sens_spec[i] <- evequal_sens_spec
+      aicetc$test_sensitivity[i] <- evsensitivity
+      aicetc$test_bic[i] <- bic #not sure this is correct...
+      aicetc$beta[i] <- betaval
+      aicetc$mpa_0.9_cum[i] <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
+
+    aicetc$AIC[i] <- ((2*aicetc$coefficients[i])-(2*aicetc$sum.of.logs[i]))
+    aicetc$AICc[i] <- (-2*(aicetc$sum.of.logs[i])+((2*aicetc$coefficients[i])*(aicetc$no_trainng_points[i]/(aicetc$no_trainng_points[i] - aicetc$coefficients[i]-1))))
+    }
+   }
+  }
+
+write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)
+
+}
+
+```
+
+AICc selection
+select the models you want for each year and run those models again (primarily based on lowest AICc but check test AUC and other things too)
 
@@
 
-all layers need to be spit in bash + poweshell
-- pre-split.ps1 
-- separate.sh
-- aftersplit.ps1
-
 # average the maxent results
+did this after splitting layers back into their individual depths as per Bentlage 2013 https://doi.org/10.1111/geb.12049
 
@@
 
+#the following cde was manually edited for each month
+
 for (x in 1:nrow(depthlayers)){
@@
 ```{r}
-library(stringr)
 
@@
 
-#plotting
-
-need some additional packages installed
-```{r pre-install & load packages, include=FALSE}
-packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer", "cowplot", "rasterVis")
-source("../src/ipak.R")
-ipak(packages)
-```
-
 # stacked permutation importance plot
-
-note if you change mind and want jackknife then the code is in the older plots for papers rmd
-
-first need to make the data plot-ready
+This is Figure 2
 
@@
 colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
-# jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
-# colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
 
@@
   perm$Month[m] <- mth
-  # jack$Month[m] <- mth
   maxt <- tail(maxr, 1)
@@
 
-  # jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
-  # jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
-  # jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
-  # jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
-  # jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
-  # jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
-  # jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
-  # jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
-  # jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
 }
@@
 permu <- perm[match(mthord, perm$Month), ] #reorder based on mthord
-# jackk <- jack[match(mthord, jack$Month), ] #reorder based on mthord
 
@@
 permu$Month <- capitalize(permu$Month)
-# jackk$Month <- capitalize(jackk$Month)
 
@@
 
-# colnames(jackk)[colnames(jackk)=="amo_sample"] <- "AMO_Sample"
-# colnames(jackk)[colnames(jackk)=="amo_winter"] <- "AMO_Winter"
-# colnames(jackk)[colnames(jackk)=="chl_surface"] <- "Chl_Surface"
-# colnames(jackk)[colnames(jackk)=="nao_prev"] <- "NAO_Prev"
-# colnames(jackk)[colnames(jackk)=="nao_sample"] <- "NAO_Sample"
-# colnames(jackk)[colnames(jackk)=="nao_winter"] <- "NAO_Winter"
-# colnames(jackk)[colnames(jackk)=="o2_depth"] <- "O2_Depth"
-# colnames(jackk)[colnames(jackk)=="salinity_depth"] <- "Salinity_Depth"
-# colnames(jackk)[colnames(jackk)=="temp_depth"] <- "Temp_Depth"
-
 write.csv(permu, "../output/maxent/models/test-comb/permutation_avg.csv", row.names = FALSE)
-# write.csv(jackk, "../output/maxent/models/test-comb/jack_avg.csv", row.names = FALSE)
 
@@
 now make the plot
-
 ```{r permulation plot}
@@
 
-permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
-permelt$variable <- factor(permelt$variable, levels = c("AMO_Sample", "AMO_Winter", "NAO_Sample", "NAO_Prev", "NAO_Winter", "Chl_Surface", "O2_Depth", "Salinity_Depth", "Temp_Depth")) #reorder the variables
-perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = 1, labels = c("AMO Sample", "AMO Winter", "NAO Sample", "NAO Prev", "NAO Winter", "Chl Surface", "O2 Depth", "Salinity Depth", "Temp Depth")) + theme_half_open(12) + scale_x_discrete(limits = mthna) + labs(y = "Permutation Importance (%)")
+perm2 <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
+perm2$variable <- factor(perm2$variable, levels = c("AMO_Sample", "AMO_Winter", "NAO_Sample", "NAO_Prev", "NAO_Winter", "Chl_Surface", "O2_Depth", "Salinity_Depth", "Temp_Depth")) #reorder the variables
+perpl <- ggplot(data = perm2, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = 1, labels = c("AMO Sample", "AMO Winter", "NAO Sample", "NAO Prev", "NAO Winter", "Chl Surface", "O2 Depth", "Salinity Depth", "Temp Depth")) + theme_half_open(12) + scale_x_discrete(limits = mthna) + labs(y = "Permutation Importance (%)")
 png("../output/maxent/paper_plots/permutation_plot.png")
@@
 
+# sample probabily maps
+
+This is Figure 3
+
+Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)
+
+have moved sample files into a separate folder
+march,  june, sep, dec
+11.7737, 61.1128, 199.79. 628.026
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+s <- st_read("../data/env/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj)
+
+mcp <- st_read("../output/bio/minimum_convex_poly/mcp-wgs84.shp")
+
+filelist <- list.files("../data/plot2-probmaps/",pattern="avg_cloglog*",recursive = TRUE,include.dirs = TRUE)
+
+x <- stack(file.path("../data/plot2-probmaps/",filelist)) %>% 
+  mask(mcp)
+
+names(x) <- filelist
+
+df <- x %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  mutate(variable=gsub(".asc","",variable),
+         month=factor(months(parse_date_time(gsub("\\..*$", "",variable),"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         depth=as.numeric(gsub("^.*_", "",variable)))
+
+#remi code stops here
+
+df$month <- str_sub(df$variable, start = 16, end = 18)
+
+CapStr <- function(y) {
+  c <- strsplit(y, " ")[[1]]
+  paste(toupper(substring(c, 1,1)), substring(c, 2),
+      sep="", collapse=" ")
+}
+df$month <- sapply(df$month, CapStr)
+
+#remi code starts here
+
+depthlab <- paste(unique(df$depth),"m")[c(1,3,2,4)]
+names(depthlab) <- unique(df$depth)[c(1,3,2,4)]
+
+ggplot(df)+
+  geom_sf(data=mcp,fill='grey')+
+  geom_raster(aes(x,y,fill=value))+
+  geom_sf(data=mcp,fill="transparent")+
+  scale_fill_viridis(option="magma",
+                       direction=-1, limits = c(0,1), breaks = c(0, 0.25, 0.5, 0.75, 1),
+                       guide=guide_colourbar(barwidth=0.5, barheight = 15, title = NULL, draw.ulim = FALSE, draw.llim = FALSE))+
+  facet_grid(depth~month,
+             labeller = labeller(depth=depthlab))+
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "right",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
+        strip.text.x = element_text(face = "bold", size = 12),
+        strip.text.y = element_text(face = "bold", size = 12))
+
+ggsave("../output/maxent/paper_plots/prob-overview.png", width = 210, height = 297, units = "mm", dpi = 600)
+```
+
 #historgams of probability by depth
+
+This is figure 4
 
@@
 write.csv(cumst1, "../output/maxent/paper_plots/cellsperlayer.csv", row.names = FALSE)
-
-```
-
-histogram of probability by depth - by proportion
-
-#max kappa thresholds
-
-```{r}
-binprob <- function(x){
-  ifelse(x > 0 & x >= mk, 1,
-  ifelse(x <= 0, 0, NA))
-}
-
-alst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
-
-for (f in 1:length(alst)){
-  mth <- str_sub(alst[f], start = 35, end = 37)
-  af <- read.csv(alst[f], header = TRUE)
-  af <- tail(af, n = 1)
-  mk <- af$test_maxkap
-
-  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
-  for (cum in 1:length(clogasc_list)){
-    cumasc <- raster(clogasc_list[cum])
-    ascname <- names(cumasc)
-    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
-    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
-    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
-    dev.off() # stops automatic saving of the plot to a png
-  }
-}
 
@@
 
-think about how to get proportions...
-
+Get proprotions
 ```{r}
@@
 ```
-based on Remi's code
+
+plot - Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)
+
 ```{r}
@@
 
-# sample probabily maps
-
-have moved sample files into a separate folder
-march,  june, sep, dec
-11.7737, 61.1128, 199.79. 628.026
+#2d maps - max kappa thresholds
+
+Get max kappa thresholds
 
 ```{r}
-library(sf)
-library(purrr)
-library(raster)
-library(lubridate)
-library(tidyverse)
-
-proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
-
-s <- st_read("../data/env/landp/land-wgs84v2.shp") %>% 
-  st_set_crs(proj)
-
-mcp <- st_read("../output/bio/minimum_convex_poly/mcp-wgs84.shp")
-
-filelist <- list.files("../data/plot2-probmaps/",pattern="avg_cloglog*",recursive = TRUE,include.dirs = TRUE)
-
-x <- stack(file.path("../data/plot2-probmaps/",filelist)) %>% 
-  mask(mcp)
-
-names(x) <- filelist
-
-df <- x %>% 
-  rasterToPoints() %>% 
-  as.data.frame() %>% 
-  gather(key = "variable", value = "value",-x,-y) %>% 
-  mutate(variable=gsub(".asc","",variable),
-         month=factor(months(parse_date_time(gsub("\\..*$", "",variable),"%m")),
-                      ordered = TRUE,
-                      levels=months(parse_date_time(1:12,"%m"))),
-         depth=as.numeric(gsub("^.*_", "",variable)))
-
-#remi code stops here
-
-df$month <- str_sub(df$variable, start = 16, end = 18)
-
-CapStr <- function(y) {
-  c <- strsplit(y, " ")[[1]]
-  paste(toupper(substring(c, 1,1)), substring(c, 2),
-      sep="", collapse=" ")
-}
-df$month <- sapply(df$month, CapStr)
-
-#remi code starts here
-
-depthlab <- paste(unique(df$depth),"m")[c(1,3,2,4)]
-names(depthlab) <- unique(df$depth)[c(1,3,2,4)]
-
-ggplot(df)+
-  geom_sf(data=mcp,fill='grey')+
-  geom_raster(aes(x,y,fill=value))+
-  geom_sf(data=mcp,fill="transparent")+
-  scale_fill_viridis(option="magma",
-                       direction=-1, limits = c(0,1), breaks = c(0, 0.25, 0.5, 0.75, 1),
-                       guide=guide_colourbar(barwidth=0.5, barheight = 15, title = NULL, draw.ulim = FALSE, draw.llim = FALSE))+
-  facet_grid(depth~month,
-             labeller = labeller(depth=depthlab))+
-  theme_classic()+
-  scale_x_continuous(breaks = seq(-70,-50,10))+
-  theme(legend.position = "right",
-        axis.title.x=element_blank(),
-        axis.title.y=element_blank(),
-        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
-        strip.text.x = element_text(face = "bold", size = 12),
-        strip.text.y = element_text(face = "bold", size = 12))
-
-ggsave("../output/maxent/paper_plots/test-overview.png", width = 210, height = 297, units = "mm", dpi = 600)
-```
-
-#2d maps - max kappa thresholds
+binprob <- function(x){
+  ifelse(x > 0 & x >= mk, 1,
+  ifelse(x <= 0, 0, NA))
+}
+
+alst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
+
+for (f in 1:length(alst)){
+  mth <- str_sub(alst[f], start = 35, end = 37)
+  af <- read.csv(alst[f], header = TRUE)
+  af <- tail(af, n = 1)
+  mk <- af$test_maxkap
+
+  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)
+  for (cum in 1:length(clogasc_list)){
+    cumasc <- raster(clogasc_list[cum])
+    ascname <- names(cumasc)
+    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
+    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".asc"), overwrite = TRUE)
+    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", ascname, ".png")), overwrite = TRUE)
+    dev.off() # stops automatic saving of the plot to a png
+  }
+}
+
+```
 
@@
 ```
+
+plot - Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)
+This is Figure 5
 
@@
 
-get data to plot maxkap binary by depth... 
-```{r maxkap by depth}
-mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
-
-binprob <- function(x){
-  ifelse(x >0,  1,
-  ifelse(x == -9999, 0, NA))
-}
-
-for (m in 1:length(mthord)){
-  mth <- mthord[m]
-  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_maxkap.*asc"), full.names = TRUE)
-  binaryavgstack <- stack(ascdepth)
-  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
-  stackstathigh[ ,1] <- c(1:47)
-  colnames(stackstathigh) <- c("asclayer", "prob_bin")
-
-  for (lyr in 1:nlayers(binaryavgstack)){
-    ascname <-  str_sub(ascdepth[lyr], 67, -5)
-    dlay <- binaryavgstack[[lyr]]
-    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
-    lyrsum <- cellStats(dlay, sum)
-    stackstathigh$prob_bin[lyr] <- lyrsum
-    stackstathigh$asclayer[lyr] <- ascname
-
-  }
-  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_cellcount_bydepth_avg.csv"), row.names = FALSE)
-}
-
-```
-
-think about how to get proportions...
+get data to plot maxkap binary by depth plot (Figure 6)
+
+Get proportions...
 
@@
 
-based on Remi's code - maxkappa derrived habitat by depth
+plot - Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)
+This is Figure 6
+
 ```{r maxkappa habitat by depth}
@@
 
-#get max and min depth layers based on max kappa
+#stats reported in the manuscript
+
+get max and min depth layers based on max kappa
 ```{r maxkap min max depth}
@@
 
-#get lat long position of most northely, southerly, western, and eastern cell centroid (max kap)
-
+get lat long position of most northely, southerly, western, and eastern cell centroid (max kap)
 ```{r maxkap cell centroid}
@@
 
-#get lat long position of most northely, southerly, western, and eastern cell centroid for oct, june, aug, may (to remove northern spots)
-
+get lat long position of most northely, southerly, western, and eastern cell centroid for oct, june, aug, may (to remove northern spots)
 ```{r}
@@
 ```
-
-#background point selection plots
-
-```{r}
-test_back_df <- read.csv("../data/env/background_point_check/201406_allbackgroundpoints.csv", header = TRUE)
-```
-
-ok now create the first lot of random for 2014_06
-
-```{r}
-testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
-testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
-testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
-testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
-testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
-```
-
-make long data
-```{r}
-bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
-bkpt$variable <- as.character(bkpt$variable)
-bkpt$no_points <- as.character(bkpt$no_points)
-
-bkpt$no_points[1:10000] <- "10,000"
-bkpt$no_points[10001:30000] <- "20,000"
-bkpt$no_points[30001:80000] <- "50,000"
-bkpt$no_points[80001:180000] <- "100,000"
-bkpt$no_points[180001:370000] <- "190,000"
-bkpt$variable <- "temp"
-bkpt$value[1:10000] <- testbk10000$temp_depth
-bkpt$value[10001:30000] <- testbk20000$temp_depth
-bkpt$value[30001:80000] <- testbk50000$temp_depth
-bkpt$value[80001:180000] <- testbk100000$temp_depth
-bkpt$value[180001:370000] <- testbk190000$temp_depth
-
-bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
-bkpt2$variable <- as.character(bkpt2$variable)
-bkpt2$no_points[1:10000] <- "10,000"
-bkpt2$no_points[10001:30000] <- "20,000"
-bkpt2$no_points[30001:80000] <- "50,000"
-bkpt2$no_points[80001:180000] <- "100,000"
-bkpt2$no_points[180001:370000] <- "190,000"
-bkpt2$variable <- "salinity"
-bkpt2$value[1:10000] <- testbk10000$salinity_depth
-bkpt2$value[10001:30000] <- testbk20000$salinity_depth
-bkpt2$value[30001:80000] <- testbk50000$salinity_depth
-bkpt2$value[80001:180000] <- testbk100000$salinity_depth
-bkpt2$value[180001:370000] <- testbk190000$salinity_depth
-
-bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
-bkpt3$variable <- as.character(bkpt3$variable)
-bkpt3$no_points[1:10000] <- "10,000"
-bkpt3$no_points[10001:30000] <- "20,000"
-bkpt3$no_points[30001:80000] <- "50,000"
-bkpt3$no_points[80001:180000] <- "100,000"
-bkpt3$no_points[180001:370000] <- "190,000"
-bkpt3$variable <- "o2"
-bkpt3$value[1:10000] <- testbk10000$o2_depth
-bkpt3$value[10001:30000] <- testbk20000$o2_depth
-bkpt3$value[30001:80000] <- testbk50000$o2_depth
-bkpt3$value[80001:180000] <- testbk100000$o2_depth
-bkpt3$value[180001:370000] <- testbk190000$o2_depth
-
-bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
-bkpt4$variable <- as.character(bkpt4$variable)
-bkpt4$no_points[1:10000] <- "10,000"
-bkpt4$no_points[10001:30000] <- "20,000"
-bkpt4$no_points[30001:80000] <- "50,000"
-bkpt4$no_points[80001:180000] <- "100,000"
-bkpt4$no_points[180001:370000] <- "190,000"
-bkpt4$variable <- "chl"
-bkpt4$value[1:10000] <- testbk10000$chl_surface
-bkpt4$value[10001:30000] <- testbk20000$chl_surface
-bkpt4$value[30001:80000] <- testbk50000$chl_surface
-bkpt4$value[80001:180000] <- testbk100000$chl_surface
-bkpt4$value[180001:370000] <- testbk190000$chl_surface
-
-bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
-bk$no_points <- as.factor(bk$no_points)
-```
-
-```{r}
-ggplot(bk, aes(x = value)) +
-    stat_density(aes(group = no_points, color = no_points),position="identity",geom="line") +
-  facet_wrap(~variable,ncol=2)
-
-```
-
-problem here is that i don't want the axis shared
-
-fiddle with the dataset again
-```{r}
-t <- subset(bk, variable == "temp")
-t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-s <- subset(bk, variable == "salinity")
-s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-o <- subset(bk, variable == "o2")
-o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-c <- subset(bk, variable == "chl")
-c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-
-```
-
-plot
-```{r}
-p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-
-p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
-
-ggsave("../output/maxent/paper_plots/bkpt_201406.png", width = 210, height = 297, units = "mm", dpi = 600)
-
-```
-
-200710
-
-```{r}
-test_back_df <- read.csv("../data/env/background_point_check/200710_allbackgroundpoints.csv", header = TRUE)
-```
-
-ok now create the first lot of random for 200710
-
-```{r}
-testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
-testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
-testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
-testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
-testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
-```
-
-make long data
-```{r}
-bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
-bkpt$variable <- as.character(bkpt$variable)
-bkpt$no_points <- as.character(bkpt$no_points)
-
-bkpt$no_points[1:10000] <- "10,000"
-bkpt$no_points[10001:30000] <- "20,000"
-bkpt$no_points[30001:80000] <- "50,000"
-bkpt$no_points[80001:180000] <- "100,000"
-bkpt$no_points[180001:370000] <- "190,000"
-bkpt$variable <- "temp"
-bkpt$value[1:10000] <- testbk10000$temp_depth
-bkpt$value[10001:30000] <- testbk20000$temp_depth
-bkpt$value[30001:80000] <- testbk50000$temp_depth
-bkpt$value[80001:180000] <- testbk100000$temp_depth
-bkpt$value[180001:370000] <- testbk190000$temp_depth
-
-bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
-bkpt2$variable <- as.character(bkpt2$variable)
-bkpt2$no_points[1:10000] <- "10,000"
-bkpt2$no_points[10001:30000] <- "20,000"
-bkpt2$no_points[30001:80000] <- "50,000"
-bkpt2$no_points[80001:180000] <- "100,000"
-bkpt2$no_points[180001:370000] <- "190,000"
-bkpt2$variable <- "salinity"
-bkpt2$value[1:10000] <- testbk10000$salinity_depth
-bkpt2$value[10001:30000] <- testbk20000$salinity_depth
-bkpt2$value[30001:80000] <- testbk50000$salinity_depth
-bkpt2$value[80001:180000] <- testbk100000$salinity_depth
-bkpt2$value[180001:370000] <- testbk190000$salinity_depth
-
-bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
-bkpt3$variable <- as.character(bkpt3$variable)
-bkpt3$no_points[1:10000] <- "10,000"
-bkpt3$no_points[10001:30000] <- "20,000"
-bkpt3$no_points[30001:80000] <- "50,000"
-bkpt3$no_points[80001:180000] <- "100,000"
-bkpt3$no_points[180001:370000] <- "190,000"
-bkpt3$variable <- "o2"
-bkpt3$value[1:10000] <- testbk10000$o2_depth
-bkpt3$value[10001:30000] <- testbk20000$o2_depth
-bkpt3$value[30001:80000] <- testbk50000$o2_depth
-bkpt3$value[80001:180000] <- testbk100000$o2_depth
-bkpt3$value[180001:370000] <- testbk190000$o2_depth
-
-bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
-bkpt4$variable <- as.character(bkpt4$variable)
-bkpt4$no_points[1:10000] <- "10,000"
-bkpt4$no_points[10001:30000] <- "20,000"
-bkpt4$no_points[30001:80000] <- "50,000"
-bkpt4$no_points[80001:180000] <- "100,000"
-bkpt4$no_points[180001:370000] <- "190,000"
-bkpt4$variable <- "chl"
-bkpt4$value[1:10000] <- testbk10000$chl_surface
-bkpt4$value[10001:30000] <- testbk20000$chl_surface
-bkpt4$value[30001:80000] <- testbk50000$chl_surface
-bkpt4$value[80001:180000] <- testbk100000$chl_surface
-bkpt4$value[180001:370000] <- testbk190000$chl_surface
-
-bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
-bk$no_points <- as.factor(bk$no_points)
-```
-
-fiddle with the dataset again
-```{r}
-t <- subset(bk, variable == "temp")
-t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-s <- subset(bk, variable == "salinity")
-s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-o <- subset(bk, variable == "o2")
-o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-c <- subset(bk, variable == "chl")
-c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-
-```
-
-plot
-```{r}
-p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-
-p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
-
-ggsave("../output/maxent/paper_plots/bkpt_200710.png", width = 210, height = 297, units = "mm", dpi = 600)
-
-```
-199902
-
-```{r}
-test_back_df <- read.csv("../data/env/background_point_check/199902_allbackgroundpoints.csv", header = TRUE)
-```
-
-ok now create the first lot of random for 200710
-
-```{r}
-testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
-testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
-testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
-testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
-testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
-```
-
-make long data
-```{r}
-bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
-bkpt$variable <- as.character(bkpt$variable)
-bkpt$no_points <- as.character(bkpt$no_points)
-
-bkpt$no_points[1:10000] <- "10,000"
-bkpt$no_points[10001:30000] <- "20,000"
-bkpt$no_points[30001:80000] <- "50,000"
-bkpt$no_points[80001:180000] <- "100,000"
-bkpt$no_points[180001:370000] <- "190,000"
-bkpt$variable <- "temp"
-bkpt$value[1:10000] <- testbk10000$temp_depth
-bkpt$value[10001:30000] <- testbk20000$temp_depth
-bkpt$value[30001:80000] <- testbk50000$temp_depth
-bkpt$value[80001:180000] <- testbk100000$temp_depth
-bkpt$value[180001:370000] <- testbk190000$temp_depth
-
-bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
-bkpt2$variable <- as.character(bkpt2$variable)
-bkpt2$no_points[1:10000] <- "10,000"
-bkpt2$no_points[10001:30000] <- "20,000"
-bkpt2$no_points[30001:80000] <- "50,000"
-bkpt2$no_points[80001:180000] <- "100,000"
-bkpt2$no_points[180001:370000] <- "190,000"
-bkpt2$variable <- "salinity"
-bkpt2$value[1:10000] <- testbk10000$salinity_depth
-bkpt2$value[10001:30000] <- testbk20000$salinity_depth
-bkpt2$value[30001:80000] <- testbk50000$salinity_depth
-bkpt2$value[80001:180000] <- testbk100000$salinity_depth
-bkpt2$value[180001:370000] <- testbk190000$salinity_depth
-
-bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
-bkpt3$variable <- as.character(bkpt3$variable)
-bkpt3$no_points[1:10000] <- "10,000"
-bkpt3$no_points[10001:30000] <- "20,000"
-bkpt3$no_points[30001:80000] <- "50,000"
-bkpt3$no_points[80001:180000] <- "100,000"
-bkpt3$no_points[180001:370000] <- "190,000"
-bkpt3$variable <- "o2"
-bkpt3$value[1:10000] <- testbk10000$o2_depth
-bkpt3$value[10001:30000] <- testbk20000$o2_depth
-bkpt3$value[30001:80000] <- testbk50000$o2_depth
-bkpt3$value[80001:180000] <- testbk100000$o2_depth
-bkpt3$value[180001:370000] <- testbk190000$o2_depth
-
-bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
-bkpt4$variable <- as.character(bkpt4$variable)
-bkpt4$no_points[1:10000] <- "10,000"
-bkpt4$no_points[10001:30000] <- "20,000"
-bkpt4$no_points[30001:80000] <- "50,000"
-bkpt4$no_points[80001:180000] <- "100,000"
-bkpt4$no_points[180001:370000] <- "190,000"
-bkpt4$variable <- "chl"
-bkpt4$value[1:10000] <- testbk10000$chl_surface
-bkpt4$value[10001:30000] <- testbk20000$chl_surface
-bkpt4$value[30001:80000] <- testbk50000$chl_surface
-bkpt4$value[80001:180000] <- testbk100000$chl_surface
-bkpt4$value[180001:370000] <- testbk190000$chl_surface
-
-bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
-bk$no_points <- as.factor(bk$no_points)
-```
-
-fiddle with the dataset again
-```{r}
-t <- subset(bk, variable == "temp")
-t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-s <- subset(bk, variable == "salinity")
-s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-o <- subset(bk, variable == "o2")
-o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-c <- subset(bk, variable == "chl")
-c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
-
-```
-
-plot
-```{r}
-p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
-
-p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
-
-ggsave("../output/maxent/paper_plots/bkpt_199902.png", width = 210, height = 297, units = "mm", dpi = 600)
-
-```
-
-# variable maps
-- temp
-- chl 
-- o2
-- sal
-- 4 different months
-- relationship of variables to each other
-- pick a depth (say the 11.7737 m)
-
-months are mar, jun, sep, dec
-pick 3 diff years say 1998, 2006, 2013

### 2020-05-27 src/testing_file.Rmd
review edit fiddling

@@
              ggtheme = theme_minimal())
-```
-
-```{r}
-library("cluster")
-pam.res <- pam(pres, 6)
-# Visualize
-fviz_cluster(pam.res)
 ```
@@
 ```
+
+#table 2 made as a plot
+
+gather all the percentages in a single file
+```{r}
+flst <- list.files("./output/maxent/paper_plots/prob_by_depth/", pattern = "_cellprob_totals.csv", full.names = TRUE)
+ptables <- lapply(flst, read.csv, header = TRUE)
+pall <- do.call(rbind, ptables)
+pall$month <- capitalize(pall$month)
+
+ggplot(data = pall, aes(x = month, y = percent, fill = Category)) + geom_bar(stat = "identity")
+```
+
+# find mean bottom depth of study region
+
+```{r}
+bathyrast <- raster("./data/env/bathymetric_glorys.asc") 
+plot(bathyrast)
+```
+need to reclassify cells outside the mcp to NA 
+```{r}
+outsidemcp <- shapefile("./data/env/landp/outside-mcp-wgs84.shp") 
+ascdirs <- list.dirs("./data/env/ascii", full.names = FALSE, recursive = FALSE)
+
+ras<- bathyrast
+ras[outsidemcp] <- 500000
+ras[ras == 500000] <- NA
+writeRaster(ras, "./data/env/bathy_trimmed.asc", overwrite=TRUE)
+cellStats(ras, stat = "mean", na.rm = TRUE) #1046.574
+summary(ras) #313.5291
+```
+
+find the number of folds in each month with a TSS <0.6
+```{r}
+flst <- list.files("./output/maxent/models/test-comb", pattern = "avgAIC.csv", full.names = TRUE, recursive = FALSE)
+df <- data.frame(month=character(10), nfolds=numeric(10), stringsAsFactors = FALSE)
+
+for (f in 1:length(flst)){
+  af <- read.csv(flst[f])
+  af <- head(af, -1)
+  mth <- (sapply(strsplit(flst[f], "/"), "[[", 6)) 
+  df$month[f] <- stringr::str_sub(mth, end = 3)
+  df$'x<0.6'[f] <- sum(af$tss <= 0.6)
+  df$nfolds[f] <- nrow(af)
+}
+write.csv(df, "./output/maxent/models/test-comb/tssfoldcount.csv", row.names = FALSE)
+```
+
+boxplots of depth of samples relatvie to the bottom depth
+```{r}
+presback <- read.csv("./output/bio/presab_nafo_yr.csv", header = TRUE) 
+pres <- subset(presback, occurrence == 1)
+pres <- subset(pres, select = c(month, bottom_depth_glorys, depth_layer))
+pres$depthdiff <- pres$bottom_depth_glorys - pres$depth_layer
+pres <- subset(pres, month >2)
+pres$month <- as.factor(pres$month)
+
+ggplot(pres, aes(x = month, y = depth_layer)) + geom_boxplot() + scale_y_reverse() + labs(y = "sampling depth")
+
+ggplot(pres, aes(x = month, y = depthdiff)) + geom_boxplot() + labs(y = "sampling depth relative to sea floor (0)")
+
+```
+
+#historgams of probability by depth
+
+This is figure 4
+
+count number of cells that are not NA in each raster layer that are inside the mcp
+
+```{r}
+
+clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/mar"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
+cumst <- data.frame(0, 0)
+colnames(cumst) <- c("depth", "no_cells")
+for (cum in 1:length(clogasc_list)){
+  cumasc <- raster(clogasc_list[cum])
+  ascname <- names(cumasc)
+  cumst[cum, 1] <- str_sub(ascname, 13)
+  #cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
+  ccnt <- freq(cumasc, useNA='no')
+  ccnt <- sum(ccnt[ ,2])
+  cumst[cum, 2] <- ccnt
+}
+
+cumst$depth <- as.numeric(cumst$depth)
+cumst1 <- cumst[order(cumst$no_cells), ]
+
+write.csv(cumst1, "../output/maxent/paper_plots/cellsperlayer.csv", row.names = FALSE)
+
+```
+
+get data to plot probability by depth... 
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+binlst <- c(1:10)
+
+binprob <- function(x){
+  ifelse(x >0,  1,
+  ifelse(x == -9999, 0, NA))
+}
+
+for (m in 1:length(mthord)){
+  for (b in 1:length(binlst)){
+    mth <- mthord[m]
+    bin <- binlst[b]
+    # mth <- "dec"
+    # bin <- 1
+  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_maxkap.*asc"), full.names = TRUE)
+
+  binaryavgstack <- stack(ascdepth)
+  stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
+  stackstathigh[ ,1] <- c(1:47)
+  colnames(stackstathigh) <- c("asclayer", "prob_bin")
+
+  for (lyr in 1:nlayers(binaryavgstack)){
+    ascname <-  str_sub(ascdepth[lyr], 77, -5)
+    dlay <- binaryavgstack[[lyr]]
+    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
+    lyrsum <- cellStats(dlay, sum)
+    stackstathigh$prob_bin[lyr] <- lyrsum
+    stackstathigh$asclayer[lyr] <- ascname
+
+  }
+  write.csv(stackstathigh, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_", bin, "_cellcount_bydepth_avg.csv"), row.names = FALSE)
+  }
+}
+
+```
+
+```{r}
+mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_probbin_.*.csv"), full.names = TRUE)
+  a <- read.csv(probblst[1])
+  b <- read.csv(probblst[2])
+  c <- read.csv(probblst[3])
+  d <- read.csv(probblst[4])
+  e <- read.csv(probblst[5])
+  f <- read.csv(probblst[6])
+  g <- read.csv(probblst[7])
+  h <- read.csv(probblst[8])
+  i <- read.csv(probblst[9])
+  j <- read.csv(probblst[10])
+
+  stackstat <- cbind(a, b[ ,2], c[ ,2], d[ ,2], e[ , 2], f[ ,2], g[ ,2], h[ ,2], i[ ,2], j[ ,2])
+  colnames(stackstat) <- c("depth", "0 - 0.09", "0.1 - 0.19", "0.2 - 0.29", "0.3 - 0.39", "0.4 - 0.49", "0.5 - 0.59", "0.6 - 0.69", "0.7 - 0.79", "0.8 - 0.89", ">0.9")
+  stackstat <- melt(stackstat, id = "depth")
+  stackstat <- stackstat[!(stackstat$depth == "1151.99"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
+  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]
+
+  colnames(stackstat) <- c("Depth", "Probability", "Cells")
+
+  write.csv(stackstat, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), row.names = FALSE)
+}
+
+```
+
+Get proprotions
+```{r}
+mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+cumst1 <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE)
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
+  for (r in 1:nrow(a)){
+    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
+    # d1 <- str_sub(a$Depth[r], end = 4)
+    d1 <- a$Depth[r]
+	  for (ro in 1:nrow(cumst1)){
+      #d2 <- str_sub(cumst1$depth[ro], end = 4)
+	    d2 <- cumst1$depth[ro]
+      if (d1 == d2){
+          a$total_cells[r] <- cumst1$no_cells[ro]
+      }
+	  }
+    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
+  }
+    a$Depth <- as.numeric(a$Depth)
+
+  write.csv(a, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_dis_prob_cells.csv"), row.names = FALSE)
+}
+
+```
+
+plot - Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)
+
+313.5291
+```{r}
+require(data.table)
+require(purrr)
+require(lubridate)
+require(tidyverse)
+require(viridis)
+
+filenames <- list.files("./output/maxent/paper_plots/prob_by_depth/", pattern = "dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)
+filenames <- filenames[ !grepl("maxkap", filenames) ] #remove all the maxkap files
+
+all <- map_dfr(filenames, function(x) fread(x) %>% 
+                       mutate(Month=substr(basename(x),1,3))) %>% 
+  mutate(Probability = gsub(">0.9", "0.9 - 1", Probability),
+         Depth = factor(round(Depth,1)),
+         Depth = factor(Depth,ordered = TRUE, levels = rev(levels(Depth))),
+         Month=factor(months(parse_date_time(Month,"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))))
+#create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
+dlst <- unique(all$Depth)
+dlst <- as.character(dlst)
+dlst <- as.numeric(dlst)
+dlst <- sort(dlst)
+#make the plot
+  ggplot(all) +
+  geom_col(aes(Depth, p_cells, fill = Probability),width=1) + 
+  coord_flip()+
+  facet_wrap(~Month,ncol=2)+
+  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
+  scale_fill_viridis(option="magma", discrete = TRUE, direction = -1, guide=guide_legend(nrow=2, byrow = TRUE)) +
+  theme_classic()+
+  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
+  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47])) + geom_vline(xintercept=12.5, linetype = 2, col = "light grey") # this is approximate
+
+ggsave("./output/maxent/paper_plots/prob_by_depth_line.png", width = 210, height = 297, units = "mm", dpi = 600)
+```
+
+get data to plot maxkap binary by depth plot (Figure 6)
+
+Get proportions...
+
+```{r percentage cells above and below maxkap}
+mthord <- c("mar",  "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
+cumst1 <- read.csv("../output/maxent/paper_plots/cellsperlayer.csv", header = TRUE)
+
+for (m in 1:length(mthord)){
+  mth <- mthord[m]
+  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
+  for (r in 1:nrow(a)){
+    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
+    # d1 <- str_sub(a$Depth[r], end = 4)
+    d1 <- a$asclayer[r]
+	  for (ro in 1:nrow(cumst1)){
+      #d2 <- str_sub(cumst1$depth[ro], end = 4)
+	    d2 <- cumst1$depth[ro]
+      if (d1 == d2){
+          a$total_cells[r] <- cumst1$no_cells[ro]
+      }
+	  }
+    a$p_cells[r] <- (a$prob_bin[r]/a$total_cells[r])*100
+    a$Probability[r] <- 1
+  }
+    #a$asclayer <- as.numeric(a$asclayer)
+
+  a2 <- as.data.frame(a$asclayer)
+  a2$total_cells <- a$total_cells
+  a2$prob_bin <- a$total_cells - a$prob_bin
+  a2$p_cells <- (a2$prob_bin/a2$total_cells)*100
+  a2$Probability <- 0
+  colnames(a2)[1] <- "asclayer"
+
+  b <- rbind(a, a2)
+  colnames(b)[1] <- "Depth"
+
+  write.csv(b, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_maxkap_dis_prob_cells.csv"), row.names = FALSE)
+}
+
+```
+
+plot - Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)
+This is Figure 6
+
+```{r maxkappa habitat by depth}
+
+filenames <- list.files("./output/maxent/paper_plots/prob_by_depth/", pattern = "_maxkap_dis_prob_cells.csv", recursive = TRUE,full.names = TRUE)
+
+all <- map_dfr(filenames, function(x) fread(x) %>% 
+                       mutate(Month=substr(basename(x),1,3))) %>% 
+  mutate(Depth = factor(round(Depth,1)),
+         Depth = factor(Depth,ordered = TRUE, levels = rev(levels(Depth))),
+         Month=factor(months(parse_date_time(Month,"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))))
+
+#create a list of depths so you can set breaks for the x-axis (which is showing as the y-axis...)
+dlst <- unique(all$Depth)
+dlst <- as.character(dlst)
+dlst <- as.numeric(dlst)
+dlst <- sort(dlst)
+
+#make probability a factor - sam added this
+all$Probability <- as.factor(all$Probability)
+
+#make the plot
+  ggplot(all) +
+  geom_col(aes(Depth, p_cells, fill = Probability),width=1) +
+  coord_flip()+
+  facet_wrap(~Month,ncol=2)+
+  labs(y = "% Ocean at Depth", x = "Depth (meters)")+
+  scale_fill_manual(aesthetics = "fill", values=c('light yellow','black'), name = "Presence")+
+  theme_classic()+
+  theme(legend.position = "bottom", axis.text.y = element_text(size=7), strip.background = element_blank()) +
+  scale_x_discrete(breaks = c(dlst[1], dlst[5], dlst[10], dlst[15], dlst[20], dlst[25], dlst[30], dlst[35], dlst[40], dlst[47])) +
+    geom_vline(xintercept=12.5, linetype = 2, col = "light grey") # this is approximate
+
+ggsave("./output/maxent/paper_plots/maxkap_prob_by_depth_line.png", width = 210, height = 297, units = "mm", dpi = 600)
+```
+
+This is Figure 3
+
+Code base provided by Rémi Daigle (https://remidaigle.wordpress.com/)
+
+have moved sample files into a separate folder
+march,  june, sep, dec
+11.7737, 61.1128, 199.79. 628.026
+
+```{r}
+library(sf)
+library(purrr)
+library(raster)
+library(lubridate)
+library(tidyverse)
+
+proj <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
+
+s <- st_read("./data/env/landp/land-wgs84v2.shp") %>% 
+  st_set_crs(proj)
+
+mcp <- st_read("./output/bio/minimum_convex_poly/mcp-wgs84.shp")
+
+filelist <- list.files("./data/plot2-probmaps/",pattern="avg_cloglog*",recursive = TRUE,include.dirs = TRUE)
+
+x <- stack(file.path("./data/plot2-probmaps/",filelist)) %>% 
+  mask(mcp)
+
+names(x) <- filelist
+
+df <- x %>% 
+  rasterToPoints() %>% 
+  as.data.frame() %>% 
+  gather(key = "variable", value = "value",-x,-y) %>% 
+  mutate(variable=gsub(".asc","",variable),
+         month=factor(months(parse_date_time(gsub("\\..*$", "",variable),"%m")),
+                      ordered = TRUE,
+                      levels=months(parse_date_time(1:12,"%m"))),
+         depth=as.numeric(gsub("^.*_", "",variable)))
+
+#remi code stops here
+
+df$month <- str_sub(df$variable, start = 16, end = 18)
+
+CapStr <- function(y) {
+  c <- strsplit(y, " ")[[1]]
+  paste(toupper(substring(c, 1,1)), substring(c, 2),
+      sep="", collapse=" ")
+}
+df$month <- sapply(df$month, CapStr)
+df$month_f = factor(df$month, levels=c('Mar','Jun','Sep','Dec'))
+
+#remi code starts here
+
+depthlab <- paste(unique(df$depth),"m")[c(1,3,2,4)]
+names(depthlab) <- unique(df$depth)[c(1,3,2,4)]
+
+ggplot(df)+
+  geom_sf(data=mcp,fill='grey')+
+  geom_raster(aes(x,y,fill=value))+
+  geom_sf(data=mcp,fill="transparent")+
+  scale_fill_viridis(option="magma",
+                       direction=-1, limits = c(0,1), breaks = c(0, 0.25, 0.5, 0.75, 1),
+                       guide=guide_colourbar(barwidth=0.5, barheight = 15, title = NULL, draw.ulim = FALSE, draw.llim = FALSE))+
+  facet_grid(depth~month_f,
+             labeller = labeller(depth=depthlab))+
+  theme_classic()+
+  scale_x_continuous(breaks = seq(-70,-50,10))+
+  theme(legend.position = "right",
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.x=element_text(angle=45,vjust=1,hjust=1), strip.background = element_blank(),
+        strip.text.x = element_text(face = "bold", size = 12),
+        strip.text.y = element_text(face = "bold", size = 12))
+
+ggsave("./output/maxent/paper_plots/prob-overview_swap.png", width = 210, height = 297, units = "mm", dpi = 600)
+```
